### 2011-06-17 main.tex
Bare structure, including uicthesi class (color modified) and using CC BY-NC-ND license.

\documentclass{uicthesi}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{url}
%\usepackage{euler}

\usepackage{graphicx}
\graphicspath{{./images/}}
\DeclareGraphicsExtensions{.png,.pdf,.jpg}

\title{Title}
\author{Joel Berger}
\pdegrees{of the\\University of Illinois at Chicago}
\degree{Doctor of Philosophy in Experimental Physics}

\begin{document}
\maketitle

%\copyrightpage
\newpage
\begin{figure}
  \centering
  \includegraphics{by-nc-nd}
\end{figure}
This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
To view a copy of this license, visit \url{http://creativecommons.org/licenses/by-nc-nd/3.0/us/} or send a letter to
\begin{verbatim}
Creative Commons
444 Castro Street, Suite 900
Mountain View, CA, 94041, USA.
\end{verbatim}

\dedication
The dedication.

\acknowledgment
The acknowledgment.
\initials{JAB}

\preface
The preface.

\tableofcontents
\listoftables
\listoffigures

\listofabbreviations
%\begin{list}
%{}
%{\setlength
%   {\labelwidth}{1in}
%    \setlength{\leftmargin}{1.5in}
%    \setlength{\labelsep}{.5in}
%    \setlength{\rightmargin}{\leftmargin}}
%\item[AMS\hfill] American Mathematical Society
%\item[CTAN\hfill] Comprehensive \TeX\ Archive Network
%\item[TUG\hfill] \TeX\ Users Group
%\item[UIC\hfill] University of Illinois at Chicago
%\item[UICTHESI\hfill] Thesis formatting system for use at UIC.
%\end{list}

\summary
A summary is required.

\chapter{Introduction}

%\appendices
%\appendix
%\newpage
%\input fontsym
%\bibformb
%\bibliography{uictman}
%\vita
%This is where the vita goes.  Its organization is left as an exercise.
%Hint: see the list of abbreviations.
\end{document}

### 2011-06-17 uicthesi/uictest.tex
Bare structure, including uicthesi class (color modified) and using CC BY-NC-ND license.

% UICTEST.TEX
% This is a test file for my port of UICTHESI
% to LaTeX 2e. This is based in part on UCTHESIS.
%

\documentclass{uicthesi}

\begin{document}

% Declarations for Front Matter

\title{The Elements of Theses}
\author{Perry H. Disdainful}
\pdegrees{B.A. (University of Northern South Dakota at Hoople) 1978\\
M.S. (Ed's School of Quantum Mechanics and Muffler Repair Shop) 1989}
\degree{Doctor of Philosophy in Aboriginal Basketry}

\maketitle
\copyrightpage

\dedication
{\null\vfil
{\large
\begin{center}
To myself,\\\vspace{12pt}
Perry H. Disdainful,\\\vspace{12pt}
the only person worthy of my company.
\end{center}}
\vfil\null}

\acknowledgements
{I want to ``thank'' my committee, without whose ridiculous demands, I
would have graduated so, so, very much faster.}

\preface
This preface is purely optional at UIC.

\tableofcontents
\listoftables
\listoffigures
\listofabbreviations
\begin{list}
{}
{\setlength
   {\labelwidth}{1in}
    \setlength{\leftmargin}{1.5in}
    \setlength{\labelsep}{.5in}
    \setlength{\rightmargin}{\leftmargin}}
\item[AMS\hfill] American Mathematical Society
\item[CTAN\hfill] Comprehensive \TeX\ Archive Network
\item[TUG\hfill] \TeX\ Users Group
\item[UIC\hfill] University of Illinois at Chicago
\item[UICTHESI\hfill] Thesis formatting system for use at UIC.
\end{list}

\summary
A summary is required.

\chapter{Introduction}

Every dissertation should have an introduction.  You might not realize
it, but the introduction should introduce the concepts, backgrouand,
and goals of the dissertation.

\section{A Section}

Most chapters have sections. Neat huh?

\subsection{A Subsection}

Some chapters have subsections too! Cool!

\begin{table}
\begin{center}
\begin{tabular}{|l|r|}
  \hline 
Title & Author \\
\hline
War And Peace & Leo Tolstoy \\
The Great Gatsby & F. Scott Fitzgerald \\ \hline
\end{tabular}
\caption{A normalsize table.}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{A small table.}
\begin{tabular}{|l|r|}
  \hline 
Title & Author \\
\hline
War And Peace & Leo Tolstoy \\
The Great Gatsby & F. Scott Fitzgerald \\ \hline
\end{tabular}
\end{center}
\end{table}

\chapter{Previous Work}

Some other research was once performed.

\begin{figure}
\caption{A first figure.}
\end{figure}

\begin{figure}
\caption{A second figure.}
\end{figure}

\chapter{Conclusion}

A conclusion is a good thing to have. Hopefully it is a good one!

\appendices
\newpage
\appendix

\chapter{Some Ancillary Stuff}

Ancillary material should be put in appendices.

\chapter{Some More Ancillary Stuff}

Here is yet another appendix! Wahoo!

\nocite{*}
\bibformb
\bibliography{uictest}
\newpage
\vita
This is where the vita goes.  Its organization is left as an exercise.

\end{document}

### 2011-06-17 uicthesi/uictman.tex
Bare structure, including uicthesi class (color modified) and using CC BY-NC-ND license.

%
\documentclass{uicthesi}
\usepackage{newlfont}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{euler}

%
% This is the user manual for UICTHESI CLS.
% Document date 1/10/92 (phd)
% Updated to include information on equation numbering 6/25/92 (phd)
%
% Updated to compile under LaTeX version 2e 2/21/96
%
% The material below is needed only for this document, not normally in
% a thesis prepared with this style file.
%
\def\new@fontshape#1#2#3#4#5{\expandafter
     \edef\csname#1/#2/#3\endcsname{\expandafter\noexpand
                                 \csname #4\endcsname}}
\new@fontshape{cmr}{bx}{sc}{
      <5>cmcsc8 at 5pt%
      <6>cmcsc8 at 6pt%
      <7>cmcsc8 at 7pt%
      <8>cmcsc8%
      <9>cmcsc9%
      <10>cmcsc10%
      <11>cmcsc10 at 10.95pt%
      <12>cmcsc10 at 12pt%
      <14>cmcsc10 at 14.4pt%
      <17>cmcsc10 at 17.28%
      <20>cmcsc10 at 20.736pt%
      <25>cmcsc10 at 24.8832pt%
      }{}
\mathversion{normal}
\newcommand{\ams}{{$\cal{A}\cal{M}\cal{S}$}}
\newcommand{\amslatex}{{$\cal{A}\cal{M}\cal{S}$-\LaTeX{}}}
\newcommand{\amstex}{{$\cal{A}\cal{M}\cal{S}$-\TeX{}}}
\newcommand{\BibTeX}{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\newcommand{\uicthesi}{{$\mathbb{UICTHESI}$}}

\newcommand\bs{\char '134 }   % A backslash character for \tt font
\newcommand{\lb}{\char '173 } % A left brace character for \tt font
\newcommand{\rb}{\char '175 } % A right brace character for \tt font

% one or two other commands
\def\newfont#1#2{\@ifdefinable #1{\font #1=#2\relax}}
\def\symbol#1{\char #1\relax}

\title{A Student's Guide to Thesis Formatting \\with \uicthesi{}}
\author{The Computer Center\\updated by Thomas McKibben}
\pdegrees{of the\\University of Illinois at Chicago}
\degree{Doctor of Philosophy in Thesis Formatting}

\begin{document}
\maketitle
\copyrightpage
\dedication
The dedication is optional, but if it is desired, the proper format for
it is created with the command \verb+\dedication+ followed by the
text of the dedication.
It untitled and does not appear in the Table of Contents.

\acknowledgment
The acknowledgment is optional, but if it is desired, the
proper format for it is created with the command \verb+\acknowledgment+
followed by the text of the acknowledgment.
The title {\bf ACKNOWLEDGMENT} will be centered at the top of the page.
Subsequent pages will have the heading {\bf ACKNOWLEDGMENT (continued)}.
If more than a single person is being acknowledged, the command should
be given as \verb+\acknowledgments+.  The section title and continued
page headings will then be the plural {\bf ACKNOWLEDGMENTS} and
{\bf ACKNOWLEDGMENTS (continued)}.
It will not appear in the Table of Contents.
To create the indented signature, include the line
\verb+\initials{XXX}+, where \verb+XXX+ are the initials of the author.
\initials{SSS}

\preface
The preface is optional, but if it is desired, the
proper format for it is created with the command \verb+\preface+
followed by the text of the preface.
The title {\bf PREFACE} will be centered at the top of the page.
Subsequent pages will have the heading {\bf PREFACE (continued)}.
It will not appear in the Table of Contents.

This document is intended to introduce the student and prospective
dissertation writer to the use of the \uicthesi{} system
for preparing a dissertation meeting the format requirements
of the Graduate College of the University of Illinois at Chicago.
It is based on the \LaTeX{} document preparation system, which in turn
is based on the \TeX{} typesetting system.
\TeX{} is a powerful text formatter which is especially suited for
technical works involving a lot of mathematics. It was developed
by a mathematics professor at Stanford for the publication of his
own books.
\LaTeX{} is a {\em macro} facility built on top of \TeX{} that,
while providing the typesetting power of \TeX{}, allows a user to
describe the organization of his work in logical rather than
physical terms.
To create a chapter heading, for instance, the user of \TeX{} must
provide the appropriate spacing, centering, and font size information.
A writer using \LaTeX{}
needs only to provide the text of the chapter heading.
The \LaTeX{} system already knows the physical commands needed to
format the chapter heading correctly.
This is made possible by the creation of a {\em style file} which
defines for the user just those physical instructions required for each
of the logical parts of the paper.
Those who have used Waterloo SCRIPT at UIC may be familiar with
a similar facility within that program called GML.

\uicthesi{} is a non-standard document style file, locally developed to
enable students at UIC to prepare dissertations that conform to the
requirements of the Graduate College.
For several years, we have had a thesis formatting system at UIC
based on the GML facility of Waterloo SCRIPT.
That system {\tt UICTHESI SCRIPT}
remains and in fact it has been recently upgraded.
However, some users may find that it is unsatisfactory for their
use, especially those in technical fields, where a large part of
their work consists of complex mathematical expressions. This may also be
the case for personal computer based word processors such as Word, and
WordPerfect which do not handle large documents containing graphics with
very much grace.
\uicthesi{} is a style file intended to meet the needs of those users.
The \uicthesi{} system is, however, suitable for preparing theses
in any field whether they require complex mathematical
typesetting capabilities or not.

This is not an exhaustive description of \TeX{} or \LaTeX{}.
It provides only the {\em essential} information that you
will need in order to use \uicthesi{}.
Only very basic features of \LaTeX{} are covered, and a
vast amount of detail has been omitted. In particular, many features of the new \LaTeX{} $2_\epsilon$ standard are omitted. In a document of this size
it is not possible to include everything that you might need to know.
If you intend to make extensive use \uicthesi{} you should
refer to a more complete reference.  Attempting to produce complex
documents using only the information found below will require
much more work than it should, and will probably produce a less
than satisfactory result.

This document does, however, cover the features specific to \uicthesi{}.
This document is itself a creation of the \uicthesi{} system.
The source of the document, \verb+UICTMAN TEX+, is an excellent
example of how to use the system. This has been updated for use with
\LaTeX{}$2_\epsilon$ which is the current standard.

The primary reference for \LaTeX{} is {\underl The \LaTeX{} User's Guide
and Reference Manual & }\cite{latex_guide}.
It contains just about all the information that you will ever
need to know about \LaTeX, and you will need access to a copy if
you are to use \LaTeX{} or \uicthesi{} successfully.
The book {\underl \LaTeX{} for Scientists and
Engineers & }\cite{latex_engineers}
is also a valuable general reference.
Web surfers will find CTAN, the Comprehensive \TeX\ Archive Network a
particularly rich source of information and tools. Use Netscape, the IBM
Web Explorer for OS/2 or other web browser to link to 
http://jasper.ora.com/ctan.html.

Important site-specific references you should read are available
through {\tt INFORM} on UICVM.  The most important of these documents are
{\underl Using \TeX{} at UIC & }\cite{use_tex},
{\underl Using \LaTeX{} at UIC & }\cite{use_latex},
{\underl Mainframe \TeX{} Version 3.1 and Related
Software & }\cite{tex31}, and
{\underl \TeX{}, \LaTeX{}, and \amstex{} Output on the Xerox 8790s
at UIC (TeXRoX) & }\cite{texrox}. If you are preparing your thesis on
a personal computer, then be sure to read the documentation that came
with your \TeX{} software.

The authoritative reference for thesis formatting at UIC is published
by the Graduate College\cite{thesis_dir}.
\uicthesi{} is an attempt to assist the student to conform to
those requirements,
but the Graduate College publication is the final authority in such
matters where this document and Graduate College requirements
differ.

This document incorporates information from a number of sources,
including {\underl Essential \LaTeX{} & }\cite{warbrick}
by Jon Warbrick of Plymoth Polytechnic, UK.

\tableofcontents
\listoftables
\listoffigures

\listofabbreviations
\begin{list}
{}
{\setlength
   {\labelwidth}{1in}
    \setlength{\leftmargin}{1.5in}
    \setlength{\labelsep}{.5in}
    \setlength{\rightmargin}{\leftmargin}}
\item[AMS\hfill] American Mathematical Society
\item[CTAN\hfill] Comprehensive \TeX\ Archive Network
\item[TUG\hfill] \TeX\ Users Group
\item[UIC\hfill] University of Illinois at Chicago
\item[UICTHESI\hfill] Thesis formatting system for use at UIC.
\end{list}

\summary
A summary is required.
The proper format for it is created with the command \verb+\summary+
followed by the text of the summary.
The title {\bf SUMMARY} will be centered at the top of the page.
Subsequent pages will have the heading {\bf SUMMARY (continued)}.
It will not appear in the Table of Contents.

\chapter{\LaTeX{}}\label{sec:styles}
Before there was \uicthesi{}, there was \LaTeX{}.
The \LaTeX{} system is still present on UICVM; \uicthesi{}
depends on it.
Before you can use it, you must give the command GETDISK TEX.
This will allow the user access to the newest version (3.1) of
\TeX{} and \LaTeX{}, which includes the \ams{} extensions.

A document to be prepared with \LaTeX{} should be created
with a text editor, such as {\tt XEDIT} on UICVM, {\ttfamily
 emacs} on TIGGER or ICARUS or your favorite editor if you are
using EM \TeX{} under DOS or OS/2 on a PC or Oz \TeX{} on a Mac.
If you choose to use a word processor such as DeScribe, Word, or
WordPerfect, be sure to save your files in ASCII format which is
plain text.
Any combination of 8 or less characters acceptable to UICVM as
a file name may be used, but the file type must be {\tt TEX}. Unix,
OS/2, Windows NT, and Windows 95 all allow file names with more
than 8 characters, however your implementation of \LaTeX{} may not
so it is best to stick to 8 or less until you are sure.
To create the document once the {\tt TEX} file has been created,
enter the command \verb+LATEX <fn>+, where \verb+<fn>+ is the
file name.  The \LaTeX{} processor will create several files,
including a file with the same file name but with the extension
\verb+DVI+.  The \verb+DVI+ file may be printed with the
\verb+PRINTTEX+ printer driver, i.e.{} \verb+PRINTTEX <fn>+.

The rest of this chapter is a brief discussion of standard \LaTeX{}.
For those readers familiar with the workings of \LaTeX{}, and
who wish to get right to material about \uicthesi{}, you may
skip to the next chapter.

\section{Standard Document Styles}
\LaTeX{} provides a number of standard {\em document styles\/}
that determine exactly how a document will be formatted.
Rather than occupying the student with mechanical concerns about
how your thesis should be laid out, \LaTeX{} instructions allow students
to describe its {\em logical structure\/}.
For example, you can think of a quotation embedded within your
text as an element of this logical structure: you would normally expect
a quotation to be displayed in a recognizable style to set it
off from the rest of the text.
A human typesetter would recognize the quotation and handle
it accordingly, but since \LaTeX{} is only a computer program,
it requires your help.
The \LaTeX{} system provides a command that allows the writer to
identify quotations and allow \LaTeX{} to typeset them correctly.

There are a number of good reasons for concentrating on the logical
structure rather than on the appearance of a document.  It prevents
you from making elementary typographical errors in the mistaken
idea that they improve the aesthetics of a document---you should
remember that the primary function of document design is to make
documents easier to read, not prettier.  It is more flexible, since
you only need to alter the definition of the quotation style
to change the appearance of all the quotations in a document.  Most
important of all, logical design encourages better writing.
A visual system makes it easier to create visual effects rather than
a coherent structure; logical design encourages you to concentrate on
your writing and makes it harder to use formatting as a substitute
for good writing.

There are four standard document styles available in \LaTeX:
\nobreak
\begin{description}

\item[{\em article}] intended for short documents and articles for
publication.
Articles do not have chapters, and when \verb+\maketitle+ is used to
generate a title (see Section~\ref{sec:title}) it appears at the top
of the first page rather than on a page of its own.

\item[{\em report}] intended for longer technical documents.
It is similar to {\tt article}, except that
it contains chapters and the title appears on a page of its own.

\item[{\em book}] intended as a basis for book publication.
Page layout is
adjusted assuming that the output will eventually be used to print on
both sides of the paper.

\item[{\em letter}] intended for producing personal letters.  This style
will allow you to produce all the elements of a well laid out letter:
addresses, date, signature, etc.
\end{description}

These standard styles can be modified by a number of
{\em style options\/}.
They appear in square brackets after the \verb+\documentstyle+ command.
Only one style can be used at a time, but you can have more than one
style option, in which case their names should be separated by commas.
The standard style options are:
\begin{description}

\item[{\em 11pt}]  prints the document using eleven-point type for the
running text rather that the ten-point type normally used.
Eleven-point type is about ten percent larger than ten-point.

\item[{\em 12pt}] prints the document using twelve-point type for
the running text
rather than the ten-point type normally used. Twelve-point type is about
twenty percent larger than ten-point.

\item[{\em twoside}] causes documents in the article or report styles
to be formatted for printing on both sides of the paper.
This is the default for the book style.

\item[{\em titlepage}] causes the \verb+\maketitle+ command to generate a
title on a separate page for documents in the {\tt article} style.
A separate page is always used in both the {\tt report} and {\tt book}
styles\footnote{Because file names on UICVM are limited to 8
characters, this option is named {\tt titlepag}.}.
\end{description}
\section{\amslatex{} Document Styles}
The current version of \LaTeX{} at UIC supports \amslatex{}.
Included with \amslatex{} are several new document styles:
{\tt amsart} and {\tt amsbook}.
These are similar to the standard \LaTeX{} styles {\tt article} and
{\tt book}, except that they have been specially modified to meet the
article and book requirements of AMS.  For further information on
these styles, see the document {\underl \amslatex{} Version 1.0
User's Guide & }\cite{amsltx}.

\chapter{Getting Started with \uicthesi}

\section{Overall Structure}
Some \LaTeX{} commands must appear in every document.
The actual text of the document always starts with
\verb+\begin{document}+ and ends with \verb+\end{document}+.
Everything that comes
before the \verb+\begin{document}+ command is called the
{\em preamble\/}. The preamble can only contain \LaTeX{} commands
to describe the document's style.
Anything that comes after the \verb+\end{document}+ command is ignored.

One command that must appear in the preamble is the
\verb+\documentclass+ command.
This command specifies the overall style for the document.
The standard styles are described earlier in this document.
The most important style for the student preparing a thesis at UIC
is \uicthesi{}.
This file, and the document that you wish to prepare with \uicthesi{},
are initiated by using the command \verb+\documentclass{uicthesi}+
Normally, no style options are required.
What would be style options for the standard document styles,
such as options for double spacing or for a titlepage,
are already built into \uicthesi{}.
When using \uicthesi{}, the default type size is 11~point.
Unlike the standard \LaTeX{} document styles,
there is no option to change the
default typesize to either 10~points or 12~points.
11~point size is acceptable by the graduate college.
If the special symbols or fonts included in the \ams{} font collection
are required, include the style option \verb+amssymb+, in the document
style declaration, i.e.{} \verb+\documentclass[amssymb]{uicthesi}+

\section{Running Text}
Most documents consist almost entirely of running text---words formed
into sentences, which are in turn formed into paragraphs.
Describing running text poses no problems, you just type
it in naturally. In the output that it produces, \LaTeX{} and \uicthesi{}
will fill lines and adjust the
spacing between words to give tidy left and right margins.
The spacing and distribution of the words in your input
file will have no effect at all on the eventual output.
Any number of spaces in your input file
are treated as a single space by \LaTeX{}, it also regards the
end of each line as a space between words.
A new paragraph is
indicated by a blank line in your input file, so don't leave
any blank lines unless you really wish to start a paragraph.

\LaTeX{} reserves a number of the less common keyboard characters for its
own use. The ten characters
\begin{quote}\begin{verbatim}
#  $  %  &  ~  _  ^  \  {  }
\end{verbatim}\end{quote}
should not appear as part of your text, because if they do
\LaTeX{} will get confused.

\section{\LaTeX{} and \uicthesi{} Commands}
There are a number of words in any \LaTeX{} document that
start `\verb+\+'.
These are \LaTeX{} {\em commands\/} and they describe the structure
of your document.
There are a number of things that you should realize about these
commands:
\begin{enumerate}

\item All \LaTeX{} commands consist of a `\verb+\+' followed by one
or more characters.

\item \LaTeX{} commands should be typed using the correct mixture of
upper- and lower-case letters.
\verb+\BEGIN+ is {\em not\/} the same as \verb+\begin+.

\item Some commands are placed within your text.  These are used to
switch things, like different typestyles, on and off.
The \verb+\em+ command is used like this to emphasize text, normally
by changing to an {\it italic\/} typestyle.
The command and the text are always enclosed between
`\verb+{+' and `\verb+}+'---the `\verb+{+' turns the effect on and
and the `\verb+}+' turns it off.
So when I write \verb+{\em emphasized text}+,
I get {\em emphasized text}.

\item There are other commands that look like
\begin{quote}\begin{verbatim}
\command{text}
\end{verbatim}\end{quote}
In this case the text is called the ``argument'' of the command.  The
\verb+\section+ command is like
this\footnote{The command to create footnotes is like this also.
Just give the command {\tt $\backslash$footnote$\{$...text...$\}$} where
{\tt ...text...} is the text of the footnote.}.
Sometimes you have to use curly brackets `\verb+{}+' to enclose the
argument, sometimes square brackets `\verb+[]+', and sometimes
both at
once\footnote{Note that the footnotes start with 1 for each page
and that the text of the footnote is single spaced at the bottom
of the same page.}.
There is method behind this apparent madness, but for the
time being you should be sure to copy the commands exactly as given.

\item When a command's name is made up entirely of letters, you must
make sure that the end of the command is marked by something that
isn't a letter.
This is usually either the opening bracket around the command's
argument, or it's a space.
When it's a space, that space is always ignored by \LaTeX. We
will see later that this can sometimes be a problem.

\end{enumerate}

\section{Other Things to Look At}
\LaTeX{} can print both opening and closing quote characters,
and can manage either of these either single or double.
To do this, it uses the two quote characters from your keyboard:
{\tt `} and {\tt '}.
You will probably think of
{\tt '} as the ordinary single quote character which probably looks like
{\tt\symbol{'23}} or {\tt\symbol{'15}} on your keyboard,
and {\tt `} as a ``funny''
character that probably appears as {\tt\symbol{'22}}.
You type these characters once for single quote, and twice for
double quotes. The double quote character {\tt "} itself
is almost never used.

\LaTeX{} can produce three different kinds of dashes.
A long dash, for use as a punctuation symbol, as is typed as three dash
characters in a row, like this `\verb+---+'.  A shorter dash,
used between numbers as in `10--20', is typed as two dash
characters in a row, while a single dash character is used as a hyphen.

From time to time you will need to include one or more of the \LaTeX{}
special symbols in your text.  Seven of them can be printed by
making them into commands.
To do this precede them with a backslash.
The remaining three symbols can be produced by more
advanced commands, as can symbols that do not appear on your keyboard
such as \dag, \ddag, \S, \pounds, \copyright, $\sharp$ and $\clubsuit$.

It is sometimes useful to include comments in a \LaTeX{} file, to remind
you of what you have done or why you did it.  Everything to the
right of a \verb+%+ sign is ignored by \LaTeX{}, and so it can
be used to introduce a comment.

The use of underlining is rarely seen in fine text, but the
UIC Graduate College thesis formatting requirements are derived from
manuscript form, rather than the form for a finished document.
Underlining is commonly found in such manuscripts, especially in place
of italics for emphasis.
\uicthesi{} automatically underlines section titles and those items in the
bibliography which, in finished text form, would appear in italics.
To underline an arbitrary piece of text, use the \verb+\underl+ command
as follows:
\begin{quote}\begin{verbatim}
{\underl This is text to be underlined & }
\end{verbatim}\end{quote}
The ampersand marks the place where the underlining is to end.
{\underl Note that the underlining will break for the end of lines
and wrap around & }.
The brackets at surrounding the command and the space before and after
the terminating \verb+&+ are required.

\section{Front Matter}
\subsection{Title Page}\label{sec:title}
A thesis at UIC must have a title page.
To prepare a title page for a \uicthesi{} thesis, you include
commands {\underl in the preamble & } to identify the title
\begin{quote}\verb+\title{Advances in Thesis Formatting}+\end{quote}
the author
\begin{quote}\verb+\author{Samuel S. Student}+\end{quote}
the author's prior degrees
\begin{quote}\verb+\pdegrees{B.S.University of Hither}+\end{quote}
and the degree for which the thesis is written
\begin{quote}\verb+\degree{Doctor of Philosophy in Pool}+\end{quote}
\uicthesi{} automatically produces the title and author's name
in all upper case letters
even if the writer enters them in mixed upper and lower case.
Immediately after the \verb+\begin{document}+, include the command
\verb+\maketitle+.
\ref{fig:titlepage} is an example of how to produce a title page.
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
\title{Advances in Thesis Formatting}
\author{Samuel S. Student}
\pdegrees{B.S.University of Hither,1983\\
M.S.University of Thither,1985\\
M.A.University of Yon,1986}
\degree{Doctor of Philosophy in Pool}
...
\maketitle
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
\begin{center}
\begin{singlespace}
{\normalsize ADVANCES IN THESIS FORMATTING}\\[12ex]
BY\\[1ex]
SAMUEL S. STUDENT\\B.S.University of Hither, 1983\\
M.S.University of Thither, 1985\\M.A.University of Yon, 1986\\[6ex]
THESIS\\[2ex]
Submitted in partial fulfillment of the requirements\\
for the degree of Doctor of Philosophy in Pool\\
in the Graduate College of the\\
University of Illinois at Chicago, 1996\\[4ex]
Chicago, Illinois
\end{singlespace}
\end{center}
\end{minipage}
\vspace{1em}
\caption{Creating a titlepage}\label{fig:titlepage}
\end{figure}

If you are not using \uicthesi{} but are using one of the standard \LaTeX{}
styles, the \verb+\degree+ and \verb+\pdegrees+ fields are omitted
but a \verb+\date+ field is included.
In the \verb+report+, \verb+book+ and \uicthesi{} style,
a full page title page is created, but
in the {\tt article} style it normally appears at the top
of the first page, the style option {\tt titlepage} will alter this (see
Section~\ref{sec:styles}).

\subsection{Preliminary Sections}
\subsubsection{Creating a Copyright Page}
The optional copyright page is produced by including the line
\verb+\copyrightpage+ just after \verb+\maketitle+.

\subsubsection{Creating a Dedication}
The dedication is optional, but if it is desired, the proper format for
it is created with the command \verb+\dedication+ followed by the
text of the dedication.
It untitled and does not appear in the Table of Contents.

\subsubsection{Creating an Acknowledgment}
The acknowledgment is optional, but if it is desired, the
proper format for it is created with the command \verb+\acknowledgment+
followed by the text of the acknowledgment.
The title {\bf ACKNOWLEDGMENT} will be centered at the top of the page.
Subsequent pages will have the heading {\bf ACKNOWLEDGMENT (continued)}.
If more than a single person is being acknowledged, the command should
be given as \verb+\acknowledgments+.  The section title and continued
page headings will then be the plural {\bf ACKNOWLEDGMENTS} and
{\bf ACKNOWLEDGMENTS (continued)}.
It will not appear in the Table of Contents.
To create the indented signature, include the line
\verb+\initials{XXX}+, where \verb+XXX+ are the initials of the author.

\subsubsection{Creating a Preface}
The preface is optional, but if it is desired, the
proper format for it is created with the command \verb+\preface+
followed by the text of the preface.
The title {\bf PREFACE} will be centered across the page.
Subsequent pages will have the title {\bf PREFACE (continued)}.
It will not appear in the table of contents.

\subsubsection{Creating the Table of Contents
and Lists of Figures and Tables}
\label{sec:toclflt}
Including the command \verb|\tableofcontents| in your document will cause
a contents list to be included, containing information collected from
the various sectioning commands as described in \ref{sec:section}.
You will notice that each time your document is run through \uicthesi{}
the table of contents is always made up of the headings from
the previous version of the document.
This is because \uicthesi{} collects information for the table as
it processes the document, and then includes it
the next time it is run.
This can sometimes mean that the document has to be
processed through \uicthesi{} twice to get a correct table of contents.
At the present time, only the numbered style of table of contents
is available.  The mixed letter and numbered style of section
identification is not available.
The title {\bf TABLE OF CONTENTS} will be centered at the
top of the page.
Any pages after the first page will have a
{\bf TABLE OF CONTENTS (Continued)} heading.

The commands \verb+\listoffigures+ and \verb+\listoftables+ perform a
similar function with the figures and tables defined in your text.

\subsubsection{Creating a List of Abbreviations}
The List of Abbreviations is generated by the command
\verb+\listofabbreviations+ followed by the text of the list.
Formatting of the list itself is left up to the student, although an
example of how it can be done is found in the source file for this
document.
The title {\bf LIST OF ABBREVIATIONS} will be centered at the
top of the page.
Any pages after the first page will have a
{\bf LIST OF ABBREVIATIONS (Continued)} heading.
It will not appear in the Table of Contents.

\subsubsection{Creating a Summary}
A summary is required.
The proper format for it is created with the command \verb+\summary+
followed by the text of the summary.
The title {\bf SUMMARY} will be centered at the top of the page.
Subsequent pages will have the heading {\bf SUMMARY (continued)}.
It will not appear in the Table of Contents.

\section{Sectioning Commands in the Body of the Thesis}
\label{sec:section}
Technical documents, like this one, are often divided into sections.
Each section has a heading containing a title and a number for easy
reference.
\LaTeX{} and \uicthesi{} have a series of commands that will allow you
to identify different kinds of sections.
Once you have done this \uicthesi{} takes on the
responsibility of laying out the title and of providing the numbers.

The commands that you can use are shown in \ref{fig:sections}.
\begin{figure}
\centering
\verb+\chapter+\\
\verb+\section+\\
\verb+\subsection+\\
\verb+\subsubsection+\\
\verb+\paragraph+\\
\verb+\subparagraph+
\caption{Sectioning commands}\label{fig:sections}
\end{figure}
The naming of these last two kinds of sections are
unfortunate, since they do not really have anything to do with
`paragraphs' in the normal sense of the word; they are just
lower levels of section.
Paragraphs, in the normal sense, are created by leaving a blank line
in the text.
The commands should be used in the order given, since sections are
numbered within chapters, subsections within sections, etc.

In standard document styles,
a seventh sectioning command, \verb|\part|, is also available.
Its use is always optional, and it is used to divide a large document
into series of parts.
It does not alter the numbering used for any of the other commands.
\verb|\part| is not available in \uicthesi{}.

\section{Sectioning Commands in the Appendix}
\subsection{Starting the Appendices}
When the thesis has reached that point where the main body of
the text has ended and the appendix sections are to begin, the
command \verb+\appendix+ should be used.  All \verb+\chapter+ divisions
after this point will be produce sectioning formats, headings and Table
of Contents entries for appendices rather than for regular chapters.

\subsection{Multiple Appendices}
If the thesis contains more than one appendix, the command
\verb+\appendices+ will create a page which contains only the
word {\bf APPENDICES} and the page number.
This command should immediately
precede the \verb+\appendix+ command described above.

\section{Back Matter}
\subsection{Creating a Cited Literature Section}
The Cited Literature section is created automatically by the
\verb+\bibliography{bblfile}+ command.  The \verb+bblfile+
in that command identifies the name of the external bibliography file
created by \BibTeX{}, described below.

\subsection{Using \BibTeX}
\BibTeX\ is a program for compiling a reference list for a document
from a bibliographic database.  It is run by entering
\begin{verbatim}
     BIBTEX MYFILE
\end{verbatim}
where \mbox{\tt MYFILE TEX} is the name of your \LaTeX\ input file.
This reads the file \mbox{\tt MYFILE AUX}, which was generated when you
ran \LaTeX\ on \mbox{\tt MYFILE TEX}, and produces the file \mbox{\tt
MYFILE BBL}.
The \BibTeX\ program requires a separate source file, call a \verb+BIB+
file, containing the information that will appear in the Cited Literature
section.

Instead of the \verb+\bibliographystyle+ command used with standard
\LaTeX{} styles,
the particular bibliography style is selected by the command
\verb+\bibform+{\it n}, where {\it n} is either \verb+a+, \verb+b+
or \verb+c+.
These commands correspond to the three bibliography styles described
in the Graduate School thesis manual\cite{thesis_dir}.
\verb+\bibforma+ creates text citations containing the author's name
and the year of publication and creates an unnumbered, alphabetized
Cited Literature section.
\verb+\bibformb+ creates numbered text citations and creates a
numbered Cited Literature section ordered by the order of their first
appearance in the text.
\verb+\bibformc+ creates numbered text citations like \verb+\bibformb+,
but the Cited Literature section is numbered and ordered alphabetically.

There is a simple \BibTeX{} User's Guide available through INFORM
(enter {\tt INFORM BIBTEX} in CMS)

In the Cited Literature section near the end of this manual,
there are sample entries for an article\cite{smpl_article}, a
book\cite{smpl_book}, an article in a
collection\cite{smpl_incollection}, and
a thesis\cite{smpl_masterthesis}.

\subsection{Creating a Vita}
The vita is generated by the command \verb+\vita+
followed by the text of the vita.
It is up to the user to provide the formatting commands within the vita.
It will appear in the table of contents.

\chapter{Environments}
We mentioned earlier the idea of identifying a quotation to \LaTeX{} or
\uicthesi{} so that
it could arrange to typeset it correctly. To do this you enclose the
quotation between the commands \verb|\begin{quotation}| and
\verb|\end{quotation}|.
This is an example of a \LaTeX{} construction called an {\em environment\/}.
A number of
special effects are obtained by putting text into particular environments.

\section{Quotations}
There are two environments for quotations: {\tt quote}
and {\tt quotation}.
{\tt quote} is used either for a short quotation or for a sequence of
short quotations separated by blank lines.
An illustration of how to create a quote environment is shown in
\ref{fig:quote}.
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
US presidents ... pithy remarks:
\begin{quote}
The buck stops here.

I am not a crook.
\end{quote}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
US presidents have been known for their pithy remarks:
\begin{quote}
The buck stops here.

I am not a crook.
\end{quote}
\end{minipage}
\vspace{1em}
\caption{Creating a quote}\label{fig:quote}
\end{figure}

Use the {\tt quotation} environment for quotations that consist of more
than one paragraph.  Paragraphs in the input are separated by blank
lines as usual.
An illustration of how to create a quotation environment is shown in
\ref{fig:quotation}.
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
Here is some advice to remember:
\begin{quotation}
Environments for making
...other things as well.

Many problems
...environments.
\end{quotation}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
Here is some advice to remember:
\begin{quotation}
Environments for making quotations
can be used for other things as well.

Many problems can be solved by
novel applications of existing
environments.
\end{quotation}
\end{minipage}
\vspace{1em}
\caption{Creating a quotation}\label{fig:quotation}
\end{figure}

\section{Centering and Flushing}
Text can be centered on the page by putting it within the {\tt center}
environment.
It will appear flush against the left or right margins if it
is placed within the {\tt flushleft} or {\tt flushright} environments.

Text within these environments will be formatted in the normal way, in
particular the ends of the lines that you type are just regarded as
spaces.  To
indicate a ``newline'' you need to type the \verb|\\| command.
\ref{fig:center} is an illustration of how to center text.
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
\begin{center}
one
two
three \\
four \\
five
\end{center}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
\begin{center}
one
two
three \\
four \\
five
\end{center}
\end{minipage}
\vspace{1em}
\caption{Centering text}\label{fig:center}
\end{figure}

\section{Lists}
There are three environments for constructing lists.  In each one each new
item is begun with an \verb|\item| command.  In the {\tt itemize} environment
the start of each item is given a marker, in the {\tt enumerate}
environment each item is marked by a number.  These environments can be nested
within each other in which case the amount of indentation used
is adjusted accordingly:
An illustration of how to create a an enumerated list within an itemized
list is shown in \ref{fig:list}.
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
\begin{itemize}
\item Itemized lists are handy.
\item However, don't forget
  \begin{enumerate}
  \item The `item' command.
  \item The `end' command.
  \end{enumerate}
\end{itemize}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
\begin{itemize}
\item Itemized lists are handy.
\item However, don't forget
  \begin{enumerate}
  \item The `item' command.
  \item The `end' command.
  \end{enumerate}
\end{itemize}
\end{minipage}
\vspace{1em}
\caption{Creating an enumerated list within an itemized list}
\label{fig:list}
\end{figure}

The third list making environment is {\tt description}.
In a description you specify the item labels inside square brackets
after the \verb|\item| command.
For an illustration of how to create a description list,
see \ref{fig:description}.
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
Three animals that you should
know about are:
\begin{description}
  \item[gnat] A small animal...
  \item[gnu] A large animal...
  \item[armadillo] A ...
\end{description}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
Three animals that you should
know about are:
\begin{description}
  \item[gnat] A small animal that causes no end of trouble.
  \item[gnu] A large animal that causes no end of trouble.
  \item[armadillo] A medium-sized animal.
\end{description}
\end{minipage}
\vspace{1em}
\caption{Creating a description list}
\label{fig:description}
\end{figure}

\section{Figures}
Many dissertations will require illustrative materials in the form of
figures.
The \verb+figure+ environment is used to create a figure.
You may create the content of the figure within \uicthesi{} itself
or leave empty space so that illustrative material from other sources
may be copied in the blank space.
An example of a figure environment is shown in \ref{fig:fig}.
In that example, you will notice a command
\verb+\label{example}+ following the caption.
This command allows a writer to refer to the figure by name instead of
by number in the text of the thesis at a moment when the writer may not
know what the number will be.
In this example, the writer may refer to the figure as
\verb+\ref{example}+ and it will appear in the text as ``Figure 1'',
or whatever number the figure eventually becomes.
The figure environment also generates information that will automatically
produce a List of Figures.
To specify where the List of Figures is to appear, use the command
\verb+\listoffigures+.
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
\begin{figure}
\vspace{3em}
\begin{center}
Place figure material here.
\end{center}
\vspace{3em}
\caption{This is an example.}
\label{example}
\end{figure}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
\vspace{3em}
\begin{center}
Place figure material here.
\end{center}
\vspace{3em}
\begin{center}
Figure 1. This is an example.
\end{center}
\end{minipage}
\vspace{1em}
\caption{How to create a figure.}
\label{fig:fig}
\end{figure}

\section{Tables}
Many dissertations will require the display of tabular information.
The \verb+table+ environment is used to create a table.
You may, like the figure, create the content of the figure within
\uicthesi{} itself, or leave empty space so that the tabular content from
other sources may be copied in the blank space.
An example of a table environment is shown in \ref{fig:tab}.
Like figures, a table may take a \verb+\label+ command for symbolic
reference.
It works the same way, except that instead of ``Figure 1'', the
\verb+\ref{example}+ will appear as ``Table I''.
As with the figure environment, the table environment also generates
information that will automatically produce a List of Tables.
To specify where the List of Tables is to appear, use the command
\verb+\listoftables+.
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
\begin{table}
\caption{Example Table}
\label{example}
\tablerule
\vspace{3em}
\begin{center}
Place tabular material here.
\end{center}
\vspace{3em}
\tablerule
\end{table}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
\begin{center}
TABLE I\\EXAMPLE TABLE
\end{center}
\tablerule
\vspace{3em}
\begin{center}
Place tabular material here.
\end{center}
\vspace{3em}
\tablerule
\end{minipage}
\vspace{1em}
\caption{How to create a table.}
\label{fig:tab}
\end{figure}

\section{Tabular Alignment with the Tabbing Environment}
One of the hardest parts of typesetting is the creation of aligned
material.
The \verb+table+ environment described above does nothing to
provide such alignments; it only sets off a section of material
that will be labeled as a table and entered into the list of tables.
The alignment of the material itself must be accomplished by other
means.
This section and the next will present two approaches, two environments
that may be used within a \verb+table+ environment to align information.

Because \LaTeX{} will almost always convert a sequence of spaces
into a single space,
it can be rather difficult to lay out tables.
See what happens in the example in \ref{fig:badtab}.
\nolinebreak
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
\begin{flushleft}
Income  Expenditure Result   \\
20s 0d  19s 11d     happiness \\
20s 0d  20s 1d      misery  \\
\end{flushleft}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
\begin{flushleft}
Income  Expenditure Result   \\
20s 0d  19s 11d     happiness \\
20s 0d  20s 1d      misery  \\
\end{flushleft}
\end{minipage}
\vspace{1em}
\caption{Tabular alignment: the wrong way}
\label{fig:badtab}
\end{figure}

The {\tt tabbing} environment overcomes this problem. Within it you set
tabstops and tab to them much like you do on a typewriter.  Tabstops are
set with the \verb|\=| command, and the \verb|\>| command moves to the
next stop.
The \verb|\\| command is used to separate each line.
A line that ends \verb|\kill|
produces no output, and can be used to set tabstops:
Now see what happens in the example in \ref{fig:tabbing}
\nolinebreak
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
\begin{tabbing}
Income \=Expenditure \=    \kill
Income \>Expenditure \>Result \\
20s 0d \>19s 11d \>Happiness   \\
20s 0d \>20s 1d  \>Misery    \\
\end{tabbing}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
\begin{tabbing}
Income \=Expenditure \=    \kill
Income \>Expenditure \>Result \\
20s 0d \>19s 11d \>Happiness   \\
20s 0d \>20s 1d  \>Misery    \\
\end{tabbing}
\end{minipage}
\vspace{1em}
\caption{Tabular alignment with tabbing}
\label{fig:tabbing}
\end{figure}

Unlike a typewriter's tab key, the \verb|\>| command always moves to
the next tabstop in sequence, even if this means moving to the left.
This can cause text to be overwritten if the gap between two tabstops
is too small.

\section{Tabular Alignment with the Tabular Environment}
Tabular alignments may also be created with the \verb+tabular+
environment.
See \ref{fig:tabular} for an example.
In that example, the argument in curly brackets following the
\verb+\begin{tabular}+ indicates how many columns, how the items are to
be placed in each of those columns, and whether there are vertical
separators. In this example, \verb+{|r|c|r|}+ indicates that there
is to be three columns, the first and third column are to be right
justified and the second column is to be centered.  Each of the columns
are to be separated by vertical lines, as indicated by the \verb+|+.
The \verb+\hline+ indicates the presence of horizontal lines.
Within the body of the table, \verb+&+ is used to separate the fields
from each other, and \verb+\\+ indicates the end of a line.
The \verb+\multicolumn+ command is used to create entries which span
across several fields, with its own formatting instructions.
As you can see from the example, footnotes may be created within
tables, but not with the ordinary \verb+\footnote+ command.
The \verb+\footnote+ command is used to create numbered footnotes in
the text that appear at the bottom of the page, but a manual process,
such as shown in the example, is required to create a lower-case lettered
footnote in a table that then appears at the bottom of the table.

\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{verbatim}
\begin{tabular}{|r|c|r|}
\hline
\multicolumn{3}{|c|}{AT\&T Common Stock}
   \\ \hline
Year&Price&Dividend\\ \hline
1971&41--54&\$2.60\\ \hline
2&41--54&2.70\\ \hline
3&46--55&2.87\\ \hline
4&40--53&3.24\\ \hline
5&45--52&3.40\\ \hline
6&51--59&.95\rlap{$^{a}$}\\ \hline
\multicolumn{3}{l}
 {$^{a}$\small (first quarter only)}
\end{tabular}
\end{verbatim}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
\begin{tabular}{|r|c|r|}
\hline
\multicolumn{3}{|c|}{AT\&T Common Stock}
   \\ \hline
Year&Price&Dividend\\ \hline
1971&41--54&\$2.60\\ \hline
2&41--54&2.70\\ \hline
3&46--55&2.87\\ \hline
4&40--53&3.24\\ \hline
5&45--52&3.40\\ \hline
6&51--59&.95\rlap{$^{a}$}\\ \hline
\multicolumn{3}{l}
   {$^{a}$\scriptsize (first quarter only)}
\end{tabular}
\end{minipage}
\vspace{1em}
\caption{Tabular alignment with tabular}
\label{fig:tabular}
\end{figure}

\section{Verbatim Output}
Sometimes you will want to include text exactly as it appears on a
terminal screen.
For example, you might want to include part of a computer program.
Not only do you want \LaTeX{} to stop playing around with the layout
of your text, you also want to be able to type all the characters
on your keyboard without confusing \LaTeX.
The {\tt verbatim} environment has this effect,
as shown in \ref{fig:verbatim}.
\begin{figure}
\footnotesize
\begin{minipage}[t]{0.48\textwidth}
\begin{flushleft}
\verb|The section of program in|  \\
\verb|question is:|               \\
\verb|\begin{verbatim}|           \\
\verb|{ this finds %a & %b }|     \\[2ex]

\verb|for i := 1 to 27 do|        \\
\ \ \ \verb|begin|                \\
\ \ \ \verb|table[i] := fn(i);|   \\
\ \ \ \verb|process(i)|           \\
\ \ \ \verb|end;|                 \\
\verb|\end{verbatim}|
\end{flushleft}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
The section of program in
question is:
\begin{verbatim}
{ this finds %a & %b }

for i := 1 to 27 do
   begin
   table[i] := fn(i);
   process(i)
   end;

\end{verbatim}
\end{minipage}
\vspace{1em}
\caption{Creating a verbatim environment}
\label{fig:verbatim}
\end{figure}

\section{Mathematical Expressions}
The great appeal of the \LaTeX{} typesetting is its ability to
typeset mathematical expressions of almost any complexity with
ease.  This document will not go into the details of how such
typesetting is to be done --- it would take a much longer work
than this.  However, it should be noted here that there are
three basic kinds of mathematical typesetting.
First, there is the in-text mode, produced by the \verb+math+
environment.  This produces a math expression right within a
sentence.  It can be produced by
\verb+\begin{math}+\ldots\verb+\end{math}+, or more frequently,
by one of two short forms, \verb+\(+\ldots\verb+\)+ or
\verb+$+\ldots\verb+$+.
The \verb+displaymath+ environment produces unnumbered displayed
formulas, that is, formulas that are set off by themselves,
centered on a line by themselves.  It can be produced by
\verb+\begin{displaymath}+\ldots\verb+\end{displaymath}+ or by the
short form \verb+\[+\ldots\verb+\]+.
A third variation is the \verb+equation+ environment, produced
by \verb+\begin{equation}+\ldots\verb+\end{equation}+ (there
is no short form for this environment), which are like the
displayed equations above, but are numbered.  Such equations
may be symbolically identified by the \verb+\label+ command
and refered to in the text with the \verb+\ref+ command.
By default, the equations are chapter-relative numbered, i.e.
the designation $m.n$ indicates that this is the $n$th equation
in Chapter $m$.  As an option, the user may include the
command \verb+\abseqnumberingtrue+ in the preamble to produce
absolute numbered equations, in which no chapter designation
occurs and the equation numbers increase through the entire document.

\chapter{Errors}
When you create a new input file for \LaTeX{} you will probably make
mistakes.
Everybody does, and it's nothing to be worried about.
As with most computer programs, there are two sorts of mistake that
you can make: those that \LaTeX{} notices and those that it doesn't.
To take a rather silly example, since \LaTeX{} doesn't understand
what you are saying it isn't going to be worried if
you misspell some of the words in your text.
You will just have to accurately proofread your printed output.
On the other hand, if you misspell one of the environment names in
your file then \LaTeX{} won't know what you want it to do.

When this sort of thing happens, \LaTeX{} prints an error message on your
terminal screen and then stops and waits for you to take some action.
Unfortunately, the error messages that it produces are rather
user-unfriendly.
Nevertheless, if you know where to look they
will probably tell you where the error is and what went wrong.

For example,
consider what would happen if you mistyped \verb+\begin{itemize}+ so
that it became \verb+\begin{itemie}+.
When \LaTeX{} processes this instruction, it
displays the text shown in \ref{fig:error}.
\begin{figure}
\begin{quote}\footnotesize\begin{verbatim}
LaTeX error.  See LaTeX manual for explanation.
              Type  H <return>  for immediate help.
! Environment itemie undefined.
\@latexerr ...for immediate help.}\errmessage {#1}
                                                  \endgroup
l.140 \begin{itemie}

?
\end{verbatim}\end{quote}
\vspace{1em}
\caption{Latex error report}
\label{fig:error}
\end{figure}
After typing the `?' \LaTeX{} stops and waits for you to tell it
what to do.

The first two lines of the message just tell you that the error was
detected by \LaTeX{}.
The third line, the one that starts `!' is the {\em error indicator}.
It tells you what the problem is, though until you have had some
experience of \LaTeX{} this may not mean a lot to you.
In this case it is just telling you
that it doesn't recognize an environment called {\tt itemie}.
The next two lines tell you what \LaTeX{} was doing when it found the
error, they are irrelevant at the moment and can be ignored.
The final line is called the {\em error locator}, and is
a copy of the line from your file that caused the problem.
It starts with a line number to help you to find it in your file, and
if the error was in the middle of a line it will be shown
broken at the point where \LaTeX{} realized that there was an error.
\LaTeX{} can sometimes pass the point where the real error is before
discovering that something is wrong, but it usually doesn't get very far.

At this point you could do several things.
If you knew enough about \LaTeX{} you might be able to fix the problem,
or you could type `X' and press the return key to stop \LaTeX{} running
while you go and correct the error.
The best thing to do, however, is just to press the return key.
This will allow \LaTeX{} to go on running as if nothing had happened.
If you have made one mistake, then you have probably made several
and you may as well try to find them all in one go.
It's much more efficient to do it this way than to run
\LaTeX{} over and over again fixing one error at a time.
Don't worry about remembering what the errors were---a copy of all
the error messages is being
saved in a {\em log\/} file so that you can look at them afterwards.
On CMS, the log file has the same name as the file to be processed, but
the file type is {\tt TEXLOG}.

If you look at the line that caused the error it's usually obvious
what the problem was.
If you can't work out what your problem is look at the hints
below, and if they don't help consult Chapter~6 of the manual.
It contains a list of all of the error messages that you are likely
to encounter together with some hints as to what may have caused them.

Some of the most common mistakes that cause errors are:
\begin{enumerate}
\item A misspelt command or environment name.
\item Improperly matched `\verb+{+' and `\verb+}+ ---remember
that they should always come in pairs.
\item Trying to use one of the ten special
characters \verb|# $ % & _ { } ~ ^| and \verb|\| as an ordinary
printing symbol.
\item A missing \verb|\end| command.
\item A missing command argument (that's the bit enclosed
in '\verb+{+' and `\verb+}+').
\end{enumerate}

One error can get \LaTeX{} so confused that it reports a series of
spurious errors as a result.
If you have an error that you understand, followed by a series that you
don't, try correcting the first error---the rest may vanish as if
by magic.

Sometimes \LaTeX{} may write a {\tt *} and stop without an error message.
This is normally caused by a missing \verb+\end{document}+ command,
but other errors can cause it.
If this happens type \verb+\stop+ and press the return key.

Finally, \LaTeX{} will sometimes print {\em warning\/} messages.
They report problems that were not bad enough to cause \LaTeX{} to
stop processing, but nevertheless may require investigation.
The most common problems are `overfull' and `underfull' lines of text.
A message like:
\begin{quote}\footnotesize\begin{verbatim}
Overfull \hbox (10.58649pt too wide) in paragraph at lines 172--175
[]\tenrm Mathematical for-mu-las may be dis-played. A dis-played
\end{verbatim}\end{quote}
indicates that \LaTeX{} could not find a good place to break a line
when laying out a paragraph.
As a result, it was forced to let the line stick out into the
right-hand margin, in this case by 10.6 points.
Since a point is about 1/72nd of an inch this may be rather hard to see,
but it will be there none the less.

This particular problem happens because \LaTeX{} is rather fussy about
line breaking, and it would rather generate a line that is too long
than generate a paragraph that doesn't meet its high standards.
The simplest way around the
problem is to enclose the entire offending paragraph between
\verb|\begin{sloppypar}| and \verb|\end{sloppypar}| commands.
This tells \LaTeX{} that you are happy for it to break its own rules
while it is working on that particular bit of text.

Alternatively, messages about ``Underfull \verb|\hbox'es''| may appear.
These are lines that had to have more space inserted between
words than \LaTeX{} would have liked.
In general there is not much that you can do about these.
Your output will look fine, even if the line looks a bit stretched.
About the only thing you could do is re-write the offending paragraph!

\chapter{Bugs}

There are a few known bugs in \LaTeX\ that occur very seldom and
cause the user little trouble, but would be very difficult to fix.
Moreover, given the nature of complex systems, it is likely that
the corrections would lead to even worse problems.  Therefore, these
bugs will probably not be fixed.

The bugs and ways to get around them are listed below.  Do not worry
about any of them until you are preparing the final draft, since
changes to the text are very likely to cause the problem to disappear.
\begin{enumerate}
\item In rare instances, a figure or table will be printed on the page
preceding the text where the {\tt figure} or {\tt table} environment
appears.  This can be fixed by moving the environment further
towards the end of the document.

\item A footnote can be broken across two pages when it should fit on a
single page.  This happens when there is one or more figures or tables
on the page.  The problem is corrected by moving, towards the end of the
file, the last {\tt figure} or {\tt table} environment that produces a
figure or table on the page where the footnote starts.

\item If you prepare your \verb+TEX+ file from a PC connected
to UICVM with Telnet, incorrect codes will be generated for
\verb+{+ and \verb+}+.
To correct all instances of incorrectly encoded braces, give the
command \verb+CURLYFIX+ from the Xedit command line.
This will not alter the future encoding of these symbols, only correct
the existing braces.
You must be connected to the \verb+TEXTOOLS+ disk before the
\verb+CURLYFIX+ command can be used.
Make sure to use the command \verb+GETDISK TEXTOOLS+ before you use
\verb+CURLYFIX+.
\end{enumerate}

\chapter{A Final Reminder}

You now know enough \LaTeX{} to produce a wide range of documents.
But this document has only scratched the surface of the
things that \LaTeX{} can do. In particular, now that \LaTeX $2_\epsilon$
has been released it would be a very good idea buy or borrow an up to
date reference manual.

This entire document was itself produced with
\LaTeX{} $2_\epsilon$ and \uicthesi{}
(with no sticking things in or clever use of a photocopier) and even
it hasn't used all the features that it could.
From this you may get some
feeling for the power that \LaTeX{} puts at your disposal.

Please remember what was said in the introduction: if you {\bf do} have a
complex document to produce then
{\bf go and read the manual}\cite{latex_guide}
You will be wasting your time if you rely only on what you
have read here.

One other warning: having
dabbled with \LaTeX{} your documents will never be the same again \ldots.

\appendices
\appendix
\newpage
\input fontsym
\bibformb
\bibliography{uictman}
\vita
This is where the vita goes.  Its organization is left as an exercise.
Hint: see the list of abbreviations.
\end{document}

### 2011-06-24 external_forces.tex
Began some of the mathematical sections.

\section{Extension to Include External Forces}

General form of (possibly multiple) external force ($\vec{F}_{ext}$) contributions
\begin{equation}
  \frac{\partial}{\partial t} A^{-1}_{ij} = K^{flow}_{ij} + K^{force}_{ij} + \sum K^{ext}_{ij}
\end{equation}
in the basis
\begin{equation}
  U_{i} = \lbrace x, p_x, y, p_y, z, p_z \rbrace
\end{equation}
and
\begin{equation}
  K^{ext}_{ij} = \frac{1}{N} \iint U_i U_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
\end{equation}

For example, if $\vec{F}$ is a static field given by $\frac{qV}{d}\hat{z}$ then $K^{ext} = \hat{0}$. However, if the external force is a lensing field given by $\vec{F} = -M\cdot(x\hat{x}+y\hat{y})$ then 
\begin{equation}
  K^{ext} = M \cdot 
  \begin{pmatrix}
    K^{lens}_{\smallT} & 0 & 0 \\
    0 & K^{lens}_{\smallT} & 0 \\
    0 & 0 & 0
  \end{pmatrix}
  \qquad \text{where} \qquad
  K^{ext} = 
  \begin{pmatrix}
    0 & \sigma_{\smallT} \\
    \sigma_{\smallT} & 2 \gamma_{\smallT}
  \end{pmatrix}
\end{equation}

### 2011-06-24 ms_model.tex
Began some of the mathematical sections.

\section{Model of Michalik and Sipe}

This section introduces the Analytic Gaussian model of Michalik and Sipe.\cite{michalik_analytic_2006}

For a classical set of $N$ particles (labelled $i$), distributed 6D space (3 position, 3 momentum) by 
\begin{gather}
  f_{{\scriptscriptstyle D}}(\vec{r}, \vec{p}; t) = 
  \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p}) \\
  \text{which satisfies} \quad \int d^{3}x\,d^{3}p \; f_{{\scriptscriptstyle D}}(\vec{r}, \vec{p}; t) = N
\end{gather}
of mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
\begin{gather}
  \frac{d\vec{r}_{i}(t)}{dt} = \frac{\vec{p}_{i}(t)}{m}\\
  \frac{d\vec{p}_{i}(t)}{dt} = -\sum\limits_{j} \frac{d}{d\vec{r}_{i}} V(\vec{r}_{i} - \vec{r}_{j})
\end{gather}

### 2011-07-19 ms_model.tex
More on MS model
added gaussian distribution section

@@
 
+\subsection{Formal Evolution}
+
 For a classical set of $N$ particles (labelled $i$), distributed 6D space (3 position, 3 momentum) by 
 \begin{gather}
-  f_{{\scriptscriptstyle D}}(\vec{r}, \vec{p}; t) = 
+  f_{\smallD}(\vec{r}, \vec{p}; t) = 
   \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p}) \\
-  \text{which satisfies} \quad \int d^{3}x\,d^{3}p \; f_{{\scriptscriptstyle D}}(\vec{r}, \vec{p}; t) = N
+  \text{which satisfies} \quad \int d^{3}x\,d^{3}p \; f_{\smallD}(\vec{r}, \vec{p}; t) = N
 \end{gather}
 of mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
+\begin{subequations}\label{eq:std_motion}
 \begin{gather}
@@
 \end{gather}
+\end{subequations}
+Now consider the time evolution of $f_{\smallD}$
+\begin{equation}
+  \frac{\partial f_{\smallD}}{\partial t} = 
+  \frac{d\vec{r}}{dt} \frac{\partial f_{\smallD}}{\partial \vec{r}} 
+  + \frac{\partial f_{\smallD}}{\partial \vec{p}} \frac{d\vec{p}}{dt}
+\end{equation}
+but with \ref{eq:std_motion} the final discreet evolution equation is
+%TODO find missing minus sign!
+\begin{equation}
+  \frac{d}{dt} f_{\smallD}(\vec{r}, \vec{p}; t) =
+  -\frac{\vec{p}}{m} \frac{\partial}{\partial \vec{r}} f_{\smallD}(\vec{r}, \vec{p}; t)
+  + \frac{\partial}{\partial \vec{p}} f_{\smallD} (\vec{r}, \vec{p}; t)
+  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f_{\smallD}(\vecprime{r},\vecprime{p};t)
+\end{equation}
+%TODO ensemble av and mean-field
+Final equation:
+\begin{equation}
+  \frac{d}{dt} f(\vec{r}, \vec{p}; t) =
+  -\frac{\vec{p}}{m} \frac{\partial}{\partial \vec{r}} f(\vec{r}, \vec{p}; t)
+  + \frac{\partial}{\partial \vec{p}} f (\vec{r}, \vec{p}; t)
+  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f(\vecprime{r},\vecprime{p};t)
+\end{equation}
+
+\subsection{Gaussian Distribution}
+
+Distribution function:
+\begin{equation}
+  f(\vec{r}, \vec{p}; t) = C(t)\exp \left[ - \Gamma(\vec{r}, \vec{p}; t) \right]
+\end{equation}
+where
+\begin{equation}
+  \Gamma(\vec{r}, \vec{p}; t) =
+  \frac{x^2 + y^2}{2 \sigma_{\smallT}} + \frac{z^2}{2 \sigma_{z}}
+  + \frac{
+    [p_x - (\gamma_{\smallT}/\sigma_{\smallT}) x ]^2 
+    + [p_y - (\gamma_{\smallT}/\sigma_{\smallT}) y ]^2
+  }{2 \eta_{\smallT}}
+  + \frac{ [p_z - (\gamma_{z}/\sigma_{z}) z ]^2 }{2 \eta_{z}}
+\end{equation}
+In this form, the normalization constant, given by
+\begin{equation}
+  \int f(\vec{r}, \vec{p}; t) d\vec{r} d\vec{p} = N
+\end{equation}
+is
+\begin{equation}
+  C(t) = \frac{N}{(2\pi)^3} 
+  \left( 
+    \frac{1}{\sigma_{\smallT}^2\eta_{\smallT}^2\sigma_{z}\eta_{z}}
+  \right)^{\frac{1}{2}}
+\end{equation}
+These equations may be viewed from a matrix perspective if we define the ``coordinate vection'' $u = (x, p_x, y, p_y, z, p_z)$. In this space,
+\begin{equation}
+  \Gamma(\vec{r}, \vec{p}; t) = \frac{1}{2}\sum\limits_{ij} A_{ij} u_i u_j
+\end{equation}
+where
+\begin{equation}
+  A = 
+  \begin{pmatrix}
+    a_{\smallT} & 0 & 0 \\
+    0 & a_{\smallT} & 0 \\
+    0 & 0 & a_{z}
+  \end{pmatrix}
+  \qquad \text{and} \qquad
+  a_{\alpha} = 
+  \begin{pmatrix}
+    1/\sigma_{\alpha} + \gamma_{\alpha}^2/(\eta_{\alpha} \sigma_{\alpha}) & \gamma_{\alpha}/(\eta_{\alpha} \sigma_{\alpha}) \\
+    \gamma_{\alpha}/(\eta_{\alpha} \sigma_{\alpha}) & 1/\eta_{\alpha}
+  \end{pmatrix}
+\end{equation}
+with ${\scriptstyle \alpha} = \smallT \text{ or } z$. This form allows the isolation of any of the six parameters using
+\begin{equation}
+  (A^{-1})_{ij} = \frac{1}{N} \int f(\vec{r}, \vec{p}; t) d\vec{r} d\vec{p}
+\end{equation}
+which yields
+\begin{equation}
+  A = 
+  \begin{pmatrix}
+    a_{\smallT}^{-1} & 0 & 0 \\
+    0 & a_{\smallT}^{-1} & 0 \\
+    0 & 0 & a_{z}^{-1}
+  \end{pmatrix}
+  \qquad \text{and} \qquad
+  a_{\alpha} = 
+  \begin{pmatrix}
+    \sigma_{\alpha} & \gamma_{\alpha} \\
+    \gamma_{\alpha} & \eta_{\alpha} + \gamma_{\alpha}^2/\sigma_{\alpha}
+  \end{pmatrix}
+\end{equation}

### 2011-09-24 ms_model.tex
Added license information to the top of each of the files that contain my content
Work towards mathematical model
fixed typo and added index on "coordinate vector"

@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \section{Model of Michalik and Sipe}
@@
 \end{equation}
-These equations may be viewed from a matrix perspective if we define the ``coordinate vection'' $u = (x, p_x, y, p_y, z, p_z)$. In this space,
+These equations may be viewed from a matrix perspective if we define the ``coordinate vector'' $u_i = (x, p_x, y, p_y, z, p_z)$. In this space,
 \begin{equation}
@@
   \qquad \text{and} \qquad
-  a_{\alpha} = 
+  a^{-1}_{\alpha} = 
   \begin{pmatrix}
@@
 \end{equation}
+Taking the time derivative of the components of $A^{-1}$
+\begin{equation}
+  \frac{d}{dt} a^{-1}_{\alpha} = 
+  \begin{pmatrix}
+    \dfrac{d\sigma_{\alpha}}{dt} & \dfrac{d\gamma_{\alpha}}{dt} \\
+    \dfrac{d\gamma_{\alpha}}{dt} & \dfrac{d\eta_{\alpha}}{dt} + 2\dfrac{\gamma_{\alpha}}{\sigma_{\alpha}}\dfrac{d\gamma_{\alpha}}{dt}- \dfrac{\gamma^{2}_{\alpha}}{\sigma^{2}_{\alpha}}\dfrac{d\sigma_{\alpha}}{dt}
+  \end{pmatrix}
+\end{equation}

### 2011-09-26 ms_model.tex
Added linking reference eq:dainvdt to ms_model and used in external_forces
Added linkage between formal evolution and the dainvdt
consistent def of u_i coordinate basis

@@
 Final equation:
-\begin{equation}
+\begin{equation} \label{eq:formal_evolution}
   \frac{d}{dt} f(\vec{r}, \vec{p}; t) =
@@
 \end{equation}
-These equations may be viewed from a matrix perspective if we define the ``coordinate vector'' $u_i = (x, p_x, y, p_y, z, p_z)$. In this space,
+These equations may be viewed from a matrix perspective if we define the ``coordinate vector'':
+\begin{equation} \label{eq:coordinate_vector}
+u_i = \{x, p_x, y, p_y, z, p_z\}\text{ .}
+\end{equation}
+In this space,
 \begin{equation}
@@
 Taking the time derivative of the components of $A^{-1}$
-\begin{equation}
+\begin{equation} \label{eq:dainvdt}
   \frac{d}{dt} a^{-1}_{\alpha} = 
@@
 \end{equation}
+These components, when viewed in the context of \ref{eq:formal_evolution}, form the left hand side (LHS) of the differential equations governing the evolution of $\sigma_{\alpha}$, $\gamma_{\alpha}$, and $\eta_{\alpha}$, which will be established in the upcoming sections.

### 2011-09-26 ms_model.tex
fixed cite punctuation placement
Renamed subsection title to "Evolution of a Gaussian Distribution", added abbrev. LHS and RHS
little fix

@@
 
-This section introduces the Analytic Gaussian model of Michalik and Sipe.\cite{michalik_analytic_2006}
+This section introduces the Analytic Gaussian model of Michalik and Sipe \cite{michalik_analytic_2006}.
 
@@
 
-\subsection{Gaussian Distribution}
+\subsection{Evolution of a Gaussian Distribution}
 
@@
 \end{equation}
-These equations may be viewed from a matrix perspective if we define the ``coordinate vector'':
+These equations may be viewed from a matrix perspective if we define the ``coordinate vector''
 \begin{equation} \label{eq:coordinate_vector}
@@
 \end{equation}
-In this space,
+In this space, \ref{eq:formal_evolution} becomes
 \begin{equation}

### 2011-09-26 ms_model.tex
Added text to ms_model

@@
 
-\subsection{Formal Evolution}
+\subsection{Formal Evolution} \label{sec:formal_evolution}
 
@@
 
-Distribution function:
+The formal evolution explored in \cite{sec:formal_evolution} allows for any distribution function.
+For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian, mimicing the laser pulse for which generated it.
+How correct this assumption is will be explored in \ref{sec:initial_conditions}.
+In this section we consider a Gaussian distribution function, generally of the form
 \begin{equation}
@@
 \end{equation}
-where
+where in this case we choose the argument (i.e. the function $\Gamma$) as
 \begin{equation}
@@
 \end{equation}
-In this form, the normalization constant, given by
+This parameterization defines the spatial uncertainty $\sigma$, the local momentum uncertainty $\eta$, and the momentum chirp $\gamma$ in each spatial direction.
+The ``chirp'' of a pulse quantifies the difference in local momentum from one side of the pulse to the other.
+For example, in a pulse which is normally (positively) chirped longitudinally, the front of the pulse with be moving faster than the back.
+
+The normalization constant, is found by integrating the entire distribution and knowing that this will contain all $N$ electrons
 \begin{equation}

### 2011-09-29 inc/initial_conditions/initial_conditions.tex
worked out some teething issues for import package

\subsection{Determining Appropriate Initial Conditions} \label{sec:initial_conditions}

### 2011-09-29 inc/ms_model/ms_model.tex
worked out some teething issues for import package

--- 2011-09-26 ms_model.tex
@@
 
-For a classical set of $N$ particles (labelled $i$), distributed 6D space (3 position, 3 momentum) by 
+For a classical set of $N$ particles (labelled by the index $i$), distributed 6D space (3 position, 3 momentum) by 
 \begin{gather}
@@
 
-The formal evolution explored in \cite{sec:formal_evolution} allows for any distribution function.
+The formal evolution explored in \ref{sec:formal_evolution} allows for any distribution function.
 For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian, mimicing the laser pulse for which generated it.

### 2011-09-30 inc/ms_model/ms_model.tex
Expanded on the nature of distribution functions

@@
 
-For a classical set of $N$ particles (labelled by the index $i$), distributed 6D space (3 position, 3 momentum) by 
-\begin{gather}
+Before one may describe the evolution of a pulse of electrons, one must first learn to describe the pulse.
+Rather than record the position of each electron, commonly one writes a distribution function for the electrons.
+A distribution function describes the location of a number of objects in a certain space.
+This space may be a 1D line, a 3D space, or even a 6D phase space (three spatial dimensions and three momentum dimensions).
+The functional form may be chosen to appropriately characterize the objects in a natural way, however one property is true for all distribution functions: when integrated over all relevant space, all objects should be accounted for.
+In $n$ dimensional space, with $N$ objects, this normalization condition on a distribution function $f$ can be written mathematically as
+\begin{equation}
+N = \int\limits_{\text{all space}} \!\!\! f(\vec{x}) \; d^{n}x
+\end{equation}
+which for 6D phase space is written as
+\begin{equation} \label{eq:normalization}
+n = \int f_{\smallD}(\vec{r}, \vec{p}; t) \; d^{3}x\,d^{3}p
+\end{equation}
+This notation suggests to the reader that the distribution function is implicitly a function of time as well as of the spatial and momentum dimensions.
+
+For a classical set of $N$ particles (labelled by the index $i$), distributed in 6D space (3 position, 3 momentum) by 
+\begin{equation}
   f_{\smallD}(\vec{r}, \vec{p}; t) = 
-  \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p}) \\
-  \text{which satisfies} \quad \int d^{3}x\,d^{3}p \; f_{\smallD}(\vec{r}, \vec{p}; t) = N
-\end{gather}
-of mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
+  \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p})
+\end{equation}
+which automatically satisfies \ref{eq:normalization}.
+The particles have mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
 \begin{subequations}\label{eq:std_motion}

### 2011-09-30 inc/ms_model/ms_model.tex
removed undesired index
linked normalization constant between formal and gaussian sections

@@
 \begin{equation} \label{eq:normalization}
-n = \int f_{\smallD}(\vec{r}, \vec{p}; t) \; d^{3}x\,d^{3}p
+n = \int f (\vec{r}, \vec{p}; t) \; d^{3}x\,d^{3}p
 \end{equation}
@@
 
-The normalization constant, is found by integrating the entire distribution and knowing that this will contain all $N$ electrons
-\begin{equation}
-  \int f(\vec{r}, \vec{p}; t) d\vec{r} d\vec{p} = N
-\end{equation}
-is
+The normalization constant $C(t)$, for $N$ total particles, is found by applying equation \ref{eq:formal_evolution}
 \begin{equation}
@@
 \end{equation}
-with ${\scriptstyle \alpha} = \smallT \text{ or } z$. This form allows the isolation of any of the six parameters using
+with ${\scriptstyle \alpha} = \smallT \text{ or } z$. 
+A nice property of Gaussians of this form is the convenient ``Multidimensional Gaussian Integral'' given by
 \begin{equation}
-  (A^{-1})_{ij} = \frac{1}{N} \int f(\vec{r}, \vec{p}; t) d\vec{r} d\vec{p}
+  (A^{-1})_{ij} = \frac{1}{N} \int u_i u_j f(\vec{r}, \vec{p}; t) d\vec{r} d\vec{p}
 \end{equation}
-which yields
+Application of this integral yields
 \begin{equation}

### 2011-10-04 inc/ms_model/ms_model.tex
Included comments on minus sign and A-1 eq. from discussions with Kim Frey

@@
 but with \ref{eq:std_motion} the final discreet evolution equation is
-%TODO find missing minus sign!
 \begin{equation}
@@
 \end{equation}
+where the extra minus sign comes from the change of perspective of \ref{eq:std_motion} from the frame of the individual particle $i$ to the observer frame.
 %TODO ensemble av and mean-field
@@
 \end{equation}
+which is a common form of integral analysis in Quantum Field Theory (see for example Peskin and Schroeder Equation 9.70). 
 Application of this integral yields

### 2011-11-11 inc/ms_model/ms_model.tex
cleaned up and provisionally completed the initial stages of the gaussian evolution derivation. This continues to the time derivative of the gaussian integral.

@@
 \begin{equation} \label{eq:normalization}
-n = \int f (\vec{r}, \vec{p}; t) \; d^{3}x\,d^{3}p
+N = \int f (\vec{r}, \vec{p}; t) \; d^{3}x\,d^{3}p
 \end{equation}
@@
 The formal evolution explored in \ref{sec:formal_evolution} allows for any distribution function.
-For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian, mimicing the laser pulse for which generated it.
+For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian, mimicing the laser pulse which generated it.
 How correct this assumption is will be explored in \ref{sec:initial_conditions}.
@@
 \end{equation}
-where in this case we choose the argument (i.e. the function $\Gamma$) as
-\begin{equation}
+The analysis presented here will consider a cylindrically symmetric pulse.
+For the remainder of this analysis, the subscripts $\smallT$ (transverse) and $z$ (longitudinal) will accompany many quantites, reflecting the congruence of $x$ and $y$.
+The subscript $\alpha$ will refer to either $\smallT$ or $z$.
+In this case we choose the argument (i.e. the function $\Gamma$) as
+\begin{equation} \label{eq:gaussian_exponent}
   \Gamma(\vec{r}, \vec{p}; t) =
@@
 \end{equation}
-This parameterization defines the spatial uncertainty $\sigma$, the local momentum uncertainty $\eta$, and the momentum chirp $\gamma$ in each spatial direction.
+This parameterization defines the spatial uncertainty $\sigma_{\alpha}$, the local momentum uncertainty $\eta_{\alpha}$, and the momentum chirp $\gamma_{\alpha}$ in each spatial direction.
 The ``chirp'' of a pulse quantifies the difference in local momentum from one side of the pulse to the other.
-For example, in a pulse which is normally (positively) chirped longitudinally, the front of the pulse with be moving faster than the back.
+For example, in a pulse which is normally (positively) chirped longitudinally, electrons near the front of the pulse will be moving faster (on average) than those in the back.
 
-The normalization constant $C(t)$, for $N$ total particles, is found by applying equation \ref{eq:formal_evolution}
+The normalization constant $C(t)$, for $N$ total particles, is found by applying equation \ref{eq:normalization}
 \begin{equation}
@@
 \end{equation}
-In this space, \ref{eq:formal_evolution} becomes
+In this space, \ref{eq:gaussian_exponent} becomes
 \begin{equation}
@@
 A nice property of Gaussians of this form is the convenient ``Multidimensional Gaussian Integral'' given by
-\begin{equation}
+\begin{equation} \label{eq:gaussian_integral}
   (A^{-1})_{ij} = \frac{1}{N} \int u_i u_j f(\vec{r}, \vec{p}; t) d\vec{r} d\vec{p}
@@
 \end{equation}
-These components, when viewed in the context of \ref{eq:formal_evolution}, form the left hand side (LHS) of the differential equations governing the evolution of $\sigma_{\alpha}$, $\gamma_{\alpha}$, and $\eta_{\alpha}$, which will be established in the upcoming sections.
+Since time derivative of \ref{eq:gaussian_integral} is simply
+\begin{equation} \label{eq:gaussian_integral}
+  \frac{d}{dt}(A^{-1})_{ij} = \frac{1}{N} \int u_i u_j \frac{df(\vec{r}, \vec{p}; t)}{dt} d\vec{r} d\vec{p}
+\end{equation}
+these components, when viewed in the context of \ref{eq:formal_evolution}, form the left hand side (LHS) of the system of differential equations governing the evolution of $\sigma_{\alpha}$, $\gamma_{\alpha}$, and $\eta_{\alpha}$, which will be established in the upcoming sections.

### 2011-11-11 inc/ms_model/ms_model.tex
added K equations to new section titled 'Generating the Differential Equations'

@@
 Since time derivative of \ref{eq:gaussian_integral} is simply
-\begin{equation} \label{eq:gaussian_integral}
+\begin{equation} \label{eq:gaussian_integral_dt}
   \frac{d}{dt}(A^{-1})_{ij} = \frac{1}{N} \int u_i u_j \frac{df(\vec{r}, \vec{p}; t)}{dt} d\vec{r} d\vec{p}
@@
 these components, when viewed in the context of \ref{eq:formal_evolution}, form the left hand side (LHS) of the system of differential equations governing the evolution of $\sigma_{\alpha}$, $\gamma_{\alpha}$, and $\eta_{\alpha}$, which will be established in the upcoming sections.
+
+\subsection{Generating the Differential Equations}
+
+Inserting \ref{eq:formal_evolution} into \ref{eq:gaussian_integral_dt} yields two integral terms, one related to the propagation of a non-charged Gaussian beam, the other adds the influence of the internal force due to Coulomb repulsion.
+These terms will be denoted $K^{flow}$ and $K^{int}$ respectively.
+\begin{equation} \label{eq:gaussian_integral_dt}
+  \frac{d}{dt}(A^{-1})_{ij} = K^{flow}_{ij} + K^{int}_{ij}
+\end{equation}
+Collecting and expanding terms where possible gives these forms of the $K$ terms
+\begin{gather}
+  K^{flow}_{ij} \equiv -\frac{1}{N} \int u_i u_j \frac{\vec{p}}{m} \frac{\partial f}{\partial \vec{r}} d\vec{r} d\vec{p} \\
+  K^{int}_{ij} \equiv \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial \Phi(\vec{r};t)}{\partial \vec{r}} d\vec{r} d\vec{p} \\
+  n(\vec{r};t) \equiv \int d\vec{p} f(\vec{r},\vec{p};t) 
+    = \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right) \\
+  \begin{split}
+    \Phi(\vec{r};t) \equiv & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) \\
+    = & \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \int d\vecprime{r} V(\vec{r} - \vecprime{r}) \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right)
+  \end{split}
+\end{gather}

### 2011-11-11 inc/external_forces/external_forces.tex
worked out some teething issues for import package
updated external_forces to use term 'int' rather than 'force'

--- 2011-06-24 external_forces.tex
@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \section{Extension to Include External Forces}
@@
 \begin{equation}
-  \frac{\partial}{\partial t} A^{-1}_{ij} = K^{flow}_{ij} + K^{force}_{ij} + \sum K^{ext}_{ij}
+  \frac{\partial}{\partial t} A^{-1}_{ij} = K^{flow}_{ij} + K^{int}_{ij} + \sum K^{ext}_{ij}
 \end{equation}
-in the basis
+again in the basis defined in \ref{eq:coordinate_vector} is given by
 \begin{equation}
-  U_{i} = \lbrace x, p_x, y, p_y, z, p_z \rbrace
-\end{equation}
-and
-\begin{equation}
-  K^{ext}_{ij} = \frac{1}{N} \iint U_i U_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
+  K^{ext}_{ij} = \frac{1}{N} \iint u_i u_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
 \end{equation}
@@
 \end{equation}
+which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\frac{\partial \gamma_{\alpha}}{\partial t}$ equation.

### 2011-12-12 inc/laser/laser.tex
Began porting the OE laser paper into the thesis

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

\section{Yb:KGW Laser System}
\subsection{Introduction}
% Draws from:
% High-power, femtosecond, thermal-lens-shaped Yb:KGW oscillator
% Joel A. Berger, Michael J. Greco, and W. Andreas Schroeder
% 9 June 2008 / Vol. 16, No. 12 / OPTICS EXPRESS

% page 8630
Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
As with other Yb-doped laser crystals [1], this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load. 
Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated [2-5] and are now available commercially [6,7].
Such lasers have used dichroic mirrors [2,4,8,9], polarization-coupling [5], or more complex optical geometries (e.g., thin disk lasers [6,10]) to pump the laser crystal in an efficient manner.

% page 8631
The spectroscopic properties of the biaxial Yb:KGW laser crystal [11] indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength [1].
However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) [12,13], mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis [5,11]. This optimal requirement for orthogonal linear pump and laser polarizations led Holtom [5] to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.

%TODO is this first sentance correct or should I use "I" voice?
In this section, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis [11,14,15]).
The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media [16].
It employs a simple single layer \ce{SiO2} anti-reflection coating on the gain medium to provide minimal loss for $p$-polarized intracavity 1040nm laser radiation and less than 0.5\% surface reflection for orthogonal $s$-polarized 980nm pump light.
The result is a simple and robust, soliton mode-locked and directly diode-pumped, solid-state laser oscillator delivering over 4W of average output power in a 63MHz train of pulses with a duration less than 300fs.
Frequency doubling of this laser output in a 2mm Brewster-cut Lithium triborate (LBO) crystal provides 1.65W of average power at 520nm.

The Watt-level green output power from the frequency-doubled sub-picosecond Yb:KGW oscillator corresponds to a visible peak laser pulse power in excess of 100kW, which is well-suited for further frequency conversion through harmonic generation or pumping of an ultrashort pulse optical parametric oscillator.
%TODO change voice/check ->
To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk [17,18] and cavity-dumped [19] Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators [6,7,20,21] and oscillators with rod-like Yb-fiber gain media [22] can be expected to yield comparable peak green pulse powers.
%TODO replace to "This output will then be use to ..." and link to something from previous
%In our case, the visible femtosecond pulses will be used as the driving laser radiation source for an ultrafast electron microscope [23] – a technology that seeks to combine the sub-nanometer spatial resolution of electron microscopy with the sub-picosecond temporal resolution afforded by today’s ultrashort pulse laser systems.
%Critical to this technology is likely to be the development of a femtosecond photo-electron gun with a low spatial emittance [25].
%The coherent control of photoemission through visible laser-driven plasmon excitations on large-area (a few mm2), gold or silver, nano-patterned photocathodes may provide a mechanism by which this can be achieved [25].

\subsection{Laser Head Design}

The design of the laser head follows the example of Holtom [5], which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism [16].
To ensure efficient operation of a Yb:KGW laser, the dominant absorption feature at 980nm for radiation polarized parallel to the optical $N_m$-axis ($\sim\,20^\circ$ from the crystallographic a-axis) should be pumped using high brightness diode lasers.
On the other hand, as with \ce{Yb$:$KY(WO4)2} [12,13], the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis [5]) in this biaxial material. 
%page 8632
Instead of using a polarization-coupled pumping geometry [5] to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig. 1(a), where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
At an angle of incidence of 32.3$^\circ$, this coating theoretically has zero reflection loss for $p$-polarized 1040nm laser radiation parallel to the $b$-axis propagating perpendicular to the $a$-$b$ plane of the crystal (Fig. 1(b)); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 [11,15].
The \ce{SiO2} coating also has an appropriate bandwidth for femtosecond laser operation in excess of 100nm (to the
0.1\% reflection points) for $p$-polarized radiation centered at 1040nm.
On the other hand, the $\sim$ 0.4\% reflection loss per surface for $s$-polarized 1040nm radiation is sufficient to suppress $s$-polarized oscillation in a laser resonator.
In addition, the same anti-reflection coating has less than 0.5\% reflection loss for $s$-polarized 980nm pump radiation (polarized parallel to the $a$-axis of the Yb:KGW crystal) at incidence angles between 21 and 39$^\circ$ (Fig. 1(b)).
Thus, in this Yb:KGW laser crystal geometry, the largest absorption cross-section associated with the optical $N_m$-axis is accessed with about 94\% efficiency ($\cos 20^\circ = 0.94$) by the $s$-polarized pump radiation.

% insert figure 1 here

### 2012-01-13 inc/laser/laser.tex
imported OE paper to end of page 8633

@@
 % insert figure 1 here
+
+Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH [26]).
+The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm fullwidth half-maximum (FWHM) 980nm absorption line of Yb:KGW [8,11,27].
+Moreover, the two dual-axis collimated laser diodes have full-angle $1/e^2$ intensity beam divergences of $\theta_y \approx$ 4.4mrad and $\theta_x \approx$ 6.8mrad in the horizontal ($x$-direction parallel to the emitter array) and vertical ($y$) directions respectively.
+With 100mm focal length lenses, this allows the pump radiation to be focused to a pump spot diameter 2W $\approx$ 400$\mu$m, yielding a maximum incident pump irradiance of $\sim$28kW/cm$^2$ from one laser diode array.
+The Yb dopant concentration was chosen to be 2\% atomic doping to ensure that more than 90\% of the pump radiation was absorbed over the 3mm Yb:KGW crystal thickness.
+With the employed counter-propagating pump geometry, the resultant maximum average pump irradiance over the crystal length is then roughly 3 times the $\sim$6kW/cm$^2$ saturation intensity of the dominant 980nm absorption line [28].
+The laser crystal was obtained from NovaPhase [29].
+
+To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique [16] to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal.
+For the employed non-Brewster crystal cut (Fig. 1(a)), a ray
+%page 8633
+transfer matrix analysis (with the $x$ direction parallel to the crystallographic $b$-axis and the $y$
+direction parallel to the $a$-axis) reveals that the required ellipticity for a stigmatic thermal lens
+generated by a parabolic temperature distribution, $T(x, y) = T_0 - \frac{1}{2} ( A_\smallT x^2 + B_\smallT y^2 )$, where $T_0$
+is the on-axis temperature, is given by
+%TODO make W symbol consistent with the rest of the paper
+\begin{equation} \label{eq:laser_ratio_ba}
+  \frac{B_\smallT}{A_\smallT} = 
+  \frac{ \cos^2 \theta_2 \left( \dfrac{dn}{dT} \right) + 2 \Delta \alpha_\smallT \left( n \cos\theta_2 - \cos\theta_1 \right) }
+       { \cos^2 \theta_1 \left[ \left( \dfrac{dn}{dT} \right) + 2 \Delta \alpha_\smallT \left( n \cos\theta_2 - \cos\theta_1 \right) \right] }
+  \quad\text{.}
+\end{equation}
+Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98 [11,15]), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
+expansion with coefficient $\alpha T$ [16].
+Generally, $\Delta \approx W l$ , where $W$ is the transverse size (radius) of the pumped region [30].
+In our case, $l \approx$ 3mm and $W \approx$ 0.2mm, so that $\Delta < 0.1$, implying that the thermal duct due to a non-zero $\frac{dn}{dT}$ primarily determines the required $B_T$/$A_T$ ratio.
+In other words, for the potassium gadolinium tungstate (\ce{KGd(WO4)2}) crystal geometry of Fig.
+1(a), we expect from \ref{eq:laser_ratio_ba} to require $\dfrac{B_\smallT}{A_\smallT} \approx \dfrac{ \cos^2 \theta_2 }{ \cos^2 \theta_1 } = 1.3$.
+
+The spatial pump distribution needed to generate the required $B_T/A_T$ ratio can be found using a cosine Fourier series solution to the thermal diffusion equation for our two-dimensional problem,
+\begin{equation} \label{eq:heat_eqn}
+  \kappa \nabla^2 T(x,y) = Q(x,y)
+  \quad\text{,}
+\end{equation}
+where $\kappa$ is the thermal diffusion coefficient.
+Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry [16].
+The specifications of our dual-axis collimated diodes [26] indicate that heat source can be written as
+\begin{equation}
+  Q(x,y) = Q_0 \exp \left[ - \frac{x^2}{W_x^2} - \left( \frac{y^2}{W_y^2} \right)^3 \right]
+  \quad\text{;}
+\end{equation}
+that is, the focused pump radiation is described accurately by a Gaussian irradiance distribution in the horizontal ($x$) direction with half-width 1/e maximum (HW1/eM) spot size $W_x$ and a super-Gaussian of order 3 in the vertical ($y$) direction with spot size $W_y$.
+\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique [16]; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
+For a fixed vertical pump spot size of $W_y = 0.2$mm, our analysis indicates that in order to achieve the required ratio $B_T / A_T = 1.3$ in the parabolic approximation for $T(x,y)$ about $(x,y) = (0,0)$ a horizontal pump spot size $W_x = 0.3$mm is needed; in other words, the required $Q(x,y)$ heat source spot size ratio (or ellipticity) $W_x / W_y = 1.5$.
+This analytical result for the central region around the laser beam axis is shown in Fig. 2; displayed are the
+%page 8634

### 2012-01-13 inc/laser/laser.tex
fixed atomic doping phrasing

@@
 With 100mm focal length lenses, this allows the pump radiation to be focused to a pump spot diameter 2W $\approx$ 400$\mu$m, yielding a maximum incident pump irradiance of $\sim$28kW/cm$^2$ from one laser diode array.
-The Yb dopant concentration was chosen to be 2\% atomic doping to ensure that more than 90\% of the pump radiation was absorbed over the 3mm Yb:KGW crystal thickness.
+The Yb concentration was chosen to be 2\% atomic doping to ensure that more than 90\% of the pump radiation was absorbed over the 3mm Yb:KGW crystal thickness.
 With the employed counter-propagating pump geometry, the resultant maximum average pump irradiance over the crystal length is then roughly 3 times the $\sim$6kW/cm$^2$ saturation intensity of the dominant 980nm absorption line [28].

### 2012-01-13 inc/laser/laser.tex
import OE paper up to (not inc) subsection 3

@@
 is the on-axis temperature, is given by
-%TODO make W symbol consistent with the rest of the paper
+%TODO check W symbol consistent with the rest of the paper
 \begin{equation} \label{eq:laser_ratio_ba}
@@
 %page 8634
+horizontal ($y = 0$, Fig. 2(a)) and vertical ($x = 0$, Fig. 2(b)) sections, respectively, through the $Q(x,y)$ and $T(x,y)$ distributions.
+Also shown are the parabolic approximations to the pump-induced temperature distribution (dashed lines) and representative TEM$_{00}$ Gaussian laser modes (shaded area) in the Yb:KGW crystal with a HW1/eM field mode sizes of $w_y = 160\mu$m and $w_x = 182 \mu$m (14\% larger due to refraction (Fig.
+1(a)).
+This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium [1]; in particular, in the vertical direction (Fig.
+2(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
+
+% insert figure 2
+
+The analysis clearly shows that the TEM$_{00}$ laser mode will experience a near perfect pump-induced thermal lens upon propagation through the gain medium --- the deviation of $T(x,y)$ from the parabolic approximation, and hence a perfect dominant thermo-optic (GRIN) duct, being insignificant over the size of the laser mode in both the horizontal and vertical directions.
+Moreover, the fact that $W_x \approx 1.6 w_x$ implies that some degree of horizontal spatial walk-off between the pump and laser beams can be tolerated in this TLS geometry.
+For example, when the pump and laser beams are exactly overlapped at the center of the Yb:KGW crystal, an external angle of 0.1rad (= 5.7$^\circ$) would generate an effective spatial walk-off of only about 75$\mu$m over our 3mm gain crystal length.
+Clearly, this can be readily accommodated by the $W_x = 0.3$mm horizontal pump spot size when $w_x \approx 180\mu$m.
+However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium [1].
+
+We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) [11,14] and stress-induced refractive index changes [31].
+For the employed crystal cut in our TLS geometry (Fig.
+1(a)), the anisotropy in $\kappa$ [14,27] is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
+The refractive index
+%page 8635
+changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index [21].
+Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping [31,34].
+Consequently, for our laser polarized along the optical $N_p$-axis [5], there is insufficient information to include accurately the effects of stress-induced refractive index changes.
+
+\subsection{The Yb:KGW laser cavity}

### 2012-01-16 inc/laser/laser.tex
Added to section 4 of oe paper

@@
 \subsection{The Yb:KGW laser cavity}
+
+The diode-pumped TLS Yb:KGW laser head is placed in the 2.35m-long, asymmetric z-fold cavity shown in Fig. 3(a).
+The two focusing gain section mirrors each have a group velocity dispersion (GVD) of -1300($\pm$150) fs$^2$ (Layertec GmbH [32]), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
+The 5$^\circ$ angle of incidence on the gain section mirrors results in minimal astigmatism, although any net round-trip cavity astigmatism can be compensated for by the TLS technique.
+The longer 75cm arm of the resonator is terminated by a concave ($R$ = 1m and angle of incidence $< 2^\circ$) high reflector
+% insert figure 3
+%page 8636
+focusing the intracavity radiation on a saturable Bragg reflector (SBR) with a 1\% reflectivity modulation depth at 1040nm (BATOP GmbH [33]) positioned a distance $z$ from the focusing mirror.
+The shorter 63cm arm is terminated by a plane output coupler with a 7\% transmission at 1040nm.
+As a free-running oscillator (SBR focusing section replaced by flat high reflector at $d_2$ = 75cm), the laser produced over 6W of TEM$_{00}$ ($M^2 < 1.2$) output power.
+
+The stability analysis of our diode-pumped Yb:KGW laser (Fig.
+3(b)), clearly indicates that the resonator configuration is unstable for thermal focal lengths less than about $-6m^{-1}$ (diopters) and is unlikely to oscillate for negative thermal focal lengths as this requires a vertical TEM$_{00}$-mode size $w_y$ greater than 180$\mu$m in the Yb:KGW gain medium --- a value that would result in strong absorptive losses due to the restrictive $W_y = 200\mu$m super-Gaussian vertical pump beam size in the Yb-doped crystal.
+The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
+This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW [11], but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis [5,34].
+
+\subsection{Mode-locked laser operation}

### 2012-01-16 inc/laser/laser.tex
added up to figure 4 from oe paper

@@
 The diode-pumped TLS Yb:KGW laser head is placed in the 2.35m-long, asymmetric z-fold cavity shown in Fig. 3(a).
-The two focusing gain section mirrors each have a group velocity dispersion (GVD) of -1300($\pm$150) fs$^2$ (Layertec GmbH [32]), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
+The two focusing gain section mirrors each have a group velocity dispersion (GVD) of $-1300 ( \pm 150 ) \text{fs}^2$ (Layertec GmbH [32]), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
 The 5$^\circ$ angle of incidence on the gain section mirrors results in minimal astigmatism, although any net round-trip cavity astigmatism can be compensated for by the TLS technique.
@@
 \subsection{Mode-locked laser operation}
+
+In the cavity configuration depicted in Fig. 3(a), self-starting mode-locked (ML) operation was observed for output powers greater than $\sim$2W and $z \approx 42$cm.
+The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (Fig. 4(a)).
+The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH [32]) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
+For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (Fig.
+4(b)) at an optimum mode-locked output power of $5.0$W.
+This trend is consistent with the expected ‘soliton’ modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity [2].
+
+The laser is capable of generating shorter ML pulse durations.
+Replacement of one of the $-1300 ( \pm 150 ) \text{fs}^2$ z-fold mirrors with a conventional $R = 50$cm high reflector and the insertion of $-880\text{fs}^2$ of negative GVD per round trip (a pair of $-220\text{fs}^2$ dispersion compensating mirrors) generated $\sim220$fs pulses at 24A of diode current, 2W of output power, and a SBR position $z = 43.2$cm.
+However, ML operation was significantly less stable under these conditions, indicating that either the cavity is operating close to the net zero GVD point or that the finite gain bandwidth is influencing the ML operation.
+We also note that in this case the optimum SBR position $z$ is larger than that for higher power 3.5W ML operation with a negative cavity GVD of $-5,200\text{fs}^2$ per round-trip (Fig.
+4(a)), but is always less than the 50cm focal length of the concave high reflector.
+This is due to the need to match the absorption-induced thermal bowing of the SBR (i.e., the radius of curvature of its front surface mirror) to the radius of curvature of the intra-cavity laser mode [35].
+In general, good ML operation is achieved with the SBR positioned within $\pm$1cm from the optimum $z$ distance.
+
+Figure 5 shows the laser's output power as a function of diode current for the cavity configuration depicted in Fig.
+3(a); a total negative GVD of $-5,200\text{fs}^2$ per round-trip and a 7\% transmission output coupler.
+Soliton mode-locked (ML) operation is observed for diode currents greater than 23A, corresponding to a continuous-wave (cw) output power of 1.8W.
+Between diode currents (output powers) of 23 and 25A (1.8 and 2.3W), the resonator operates mainly in the cw mode, but will transition to ML operation upon being subjected to a perturbation.
+The switching to ML operation is accompanied by a 250mW increase in output
+%page 8637
+power indicating a higher cavity Q for ML operation.
+This is consistent both with the laser accessing more of the available inhomogeneous gain bandwidth and with our resonator design favoring operation with a strong Kerr lens in the Yb:KGW gain medium.
+Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal’s optical $N_p$-axis [36-38] and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
+From Fig.
+3(b), we see that the resulting increase in the strength of the lensing in the Yb:KGW crystal will result in a reduction of the laser mode size in the gain medium, which produces a better spatial match to gain duct produced by the laser diode pumps (less absorptive loss), thus generating an increase in output power.
+
+% figure 4

### 2012-01-17 inc/laser/laser.tex
import oe paper to frequency double section

@@
 % figure 4
+
+Above a diode current of 25A (i.e., an output power of 2.3W), the laser exhibits spontaneous self-starting ML operation.
+At a diode current around 29A, a maximum (or optimum) average mode-locked output power of $\sim$3.5W is obtained.
+The 250fs pulses generated at this output power (Fig.
+4(a)) correspond to a Kerr lens strength in the Yb:KGW crystal of $\sim4 \text{m}^{-1}$, which is readily accommodated by the resonator design (Fig.
+3(b)).
+For this $\sim$3.5W ML laser with a negative cavity GVD of $-5,200\text{fs}^2$ per round-trip, the full-width at
+%page 8638
+half maximum of the mode-locked spectrum centered at 1046nm is 4.7nm (Fig. 4(c)).
+The laser is therefore generates pulses with a time-bandwidth product very close to the Fourier limit of 0.32 for a sech$^2$ pulse shape.
+At output powers greater than 3.5W, a DC spectral component emerges at the short wavelength end of the ML spectrum (Fig. 4(d)) and there is a slow roll-over in the laser output power (Fig. 5).
+This trend is expected for a Yb:KGW soliton ML laser; in this case, the soliton power limit being reached at about 29A, and the asymmetric spectral gain of Yb:KGW forcing excess CW laser power oscillation at wavelengths shorter than the peak of the ML spectrum.
+Similar laser performance is observed with the insertion of an additional $-5,600\text{fs}^2$ of cavity GVD per round trip, which generates 347fs sech$^2$ pulses (Fig.
+4(b)); that is, a DC spectral component appears on the short wavelength side of a narrower $\sim$3.5nm ML spectrum at output powers above 5W.
+
+\subsection{Frequency doubling}

### 2012-01-17 inc/laser/laser.tex
All main text now ported from oe paper. Still need references, figures (and captions).

@@
 
+% figure 5
+
 \subsection{Frequency doubling}
+
+The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
+The crystal is uncoated and cut for frequency doubling with the fundamental $p$-polarized radiation at 1046nm incident at Brewster’s angle --- a geometry that ensures both minimum loss for the fundamental radiation and, more importantly, a high damage threshold.
+The doubling crystal is positioned at the focus of a $z$-fold comprising a 1040nm high reflector with a radius of curvature of 75mm followed by a collimating mirror with a radius of curvature of 100mm reflective at both the fundamental and second harmonic wavelengths.
+At the 3.5W ML output power, with a pulse duration of 250fs, the estimated focal irradiance for the fundamental is $\sim50\text{GW}/\text{cm}^2$ in the LBO doubling crystal.
+Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH [32]), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
+After the collimating
+%page 8639
+mirror, a standard flat dichroic beamsplitter then separates the reflected $s$-polarized 523nm radiation from the transmitted $p$-polarized fundamental.
+
+This frequency doubling geometry for LBO is chosen to be close to the optimum for efficient short pulse doubling [39,40] defined by limits imposed by group-velocity walk-off between the fundamental and second harmonic pulses, spatial walk-off in the critical phase-matching, and the ratio of the confocal parameter of the focused fundamental wave to the LBO crystal length.
+In addition, the $z$-fold configuration allows for astigmatism compensation (e.g., due to the Brewster LBO crystal) of the second harmonic output by adjustment of the angle of incidence on the second collimating mirror.
+
+Figure 5 shows the resultant frequency-doubled green power for the ML Yb:KGW laser.
+The measured average power at 523nm increases rapidly from 0.5W at the 23A mode-locking threshold to over 1.3W at a diode drive current of 27A.
+At and above 30A, the second harmonic power saturates at 1.65W --- an effect that is clearly consistent with the observed emergence of a DC spectral component at diode currents above 29A since only the ML pulse will be efficiently frequency doubled.
+Also plotted in Fig.
+5 is the frequency doubling efficiency, which is calculated taking into account the 17\% loss suffered by the $s$-polarized 523nm radiation upon exiting the Brewster-cut LBO crystal and the 4\% loss (to diagnostics) of the fundamental radiation prior to the $z$-fold doubling geometry.
+The maximum doubling efficiency of over 60\% occurs at 27-28A of diode drive current when the ML laser output power is $\sim$3W.
+At a ML output power of 3.5W (29-30A), the frequency doubling efficiency is still $\sim$60\% and slowly decreases above 30A as the DC spectral component absorbs the residual power above the soliton power limit.
+
+\subsection{Summary}
+
+A robust, high-power ($\sim$5W), diode-pumped and thermal-lens-shaped, femtosecond (220-350fs) Yb:KGW oscillator is presented.
+The ML operation of the laser is self-starting above a critical output power of $\sim$2W, being initiated by a 1\% modulation depth SBR and subsequently supported in the soliton regime through a balance of self-phase modulation (Kerr lensing) in the gain medium and negative GVD (introduced using dispersion compensating mirrors).
+The use of the TLS technique in the diode-pumped laser head provides for astigmatism-compensated TEM$_{00}$-mode operation (a spatial beam quality $M^2 <1.2$), while proper cavity design accommodates the strong Kerr lens (up to $\sim4\text{m}^{-1}$) in the Yb:KGW crystal under ML operation.
+A full analysis of the thermal conduction in our 3x3x10mm Yb:KGW crystal is used to determine the correct TLS pump conditions, which lead to both excellent pump-laser beam overlap for the near collinear pump geometry and a near perfect pump-induced thermal lens (i.e., parabolic temperature distribution) experienced by the intracavity laser beam in the gain medium.
+
+The laser head design also employs a simple single layer \ce{SiO2} anti-reflection coating on the gain medium to provide simultaneously (i) minimal loss for the intracavity 1040nm laser radiation $p$-polarized along the optical $N_p$-axis (crystallographic $b$-axis) associated with the largest gain bandwidth and (ii) less than 0.5\% surface reflection for the orthogonally $s$-polarized 980nm laser diode radiation (polarized parallel to the crystallographic $a$-axis) to enable efficient pumping of the strongest absorption line along the optical $N_m$-axis (rotated by $\sim20^\circ$ from the crystal $a$-axis).
+In this near optimal pump-emission Yb:KGW crystal geometry for ML operation, the pump-induced thermal lens in the gain medium is observed to be positive, indicating that the sum of the thermo-optical coefficient and the stress-generated refractive index change with temperature is positive for 1$\mu$m radiation polarized parallel to the optical $N_p$-axis (crystallographic $b$-axis).
+
+The 250fs ML pulses generated at 3.5W of average output power and a 63.5MHz repetition rate are frequency doubled in a 2mm-thick, Brewster-cut and critically phase-matched Lithium triborate (LBO) crystal.
+The resultant efficient frequency doubling, which exceeds 60\% at a ML fundamental power of $\sim$3W, yields 1.65W of average power at 523nm ---
+%page 8640
+a peak green pulse power of $\sim$130kW.
+This femtosecond visible output will be used as the driving pulsed laser radiation source for an ultrafast electron microscope [23].
+
+%% Acknowledgements
+% This research was supported in part by the National Science Foundation under award DMR-
+% 0619573 and the Department of Energy under award DE-FG52-06NA26213. J.A.B.
+% acknowledges support from the Department of Education under the Graduate Assistantships
+% in Areas of National Need (GAANN) program.

### 2012-01-20 inc/laser/laser.tex
Began porting OE references into laser.tex. Currently through first 5

@@
 Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
-As with other Yb-doped laser crystals [1], this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
-Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated [2-5] and are now available commercially [6,7].
-Such lasers have used dichroic mirrors [2,4,8,9], polarization-coupling [5], or more complex optical geometries (e.g., thin disk lasers [6,10]) to pump the laser crystal in an efficient manner.
+As with other Yb-doped laser crystals\cite{Brenier_new_criteria}, this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
+Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially [6,7].
+Such lasers have used dichroic mirrors\cite{Brunner_diode_pumped,Major_femtosecond_2006} [8,9], polarization-coupling\cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers [6,10]) to pump the laser crystal in an efficient manner.
 
 % page 8631
-The spectroscopic properties of the biaxial Yb:KGW laser crystal [11] indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength [1].
-However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) [12,13], mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis [5,11].
-This optimal requirement for orthogonal linear pump and laser polarizations led Holtom [5] to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
+The spectroscopic properties of the biaxial Yb:KGW laser crystal [11] indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength\cite{Brenier_new_criteria}.
+However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) [12,13], mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis\cite{Holtom_mode_locked_2006} [11].
+This optimal requirement for orthogonal linear pump and laser polarizations led Holtom\cite{Holtom_mode_locked_2006} to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
 
@@
 
-The design of the laser head follows the example of Holtom [5], which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism [16].
+The design of the laser head follows the example of Holtom\cite{Holtom_mode_locked_2006}, which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism [16].
 To ensure efficient operation of a Yb:KGW laser, the dominant absorption feature at 980nm for radiation polarized parallel to the optical $N_m$-axis ($\sim\,20^\circ$ from the crystallographic a-axis) should be pumped using high brightness diode lasers.
-On the other hand, as with \ce{Yb$:$KY(WO4)2} [12,13], the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis [5]) in this biaxial material.
+On the other hand, as with \ce{Yb$:$KY(WO4)2} [12,13], the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis\cite{Holtom_mode_locked_2006}) in this biaxial material.
 %page 8632
-Instead of using a polarization-coupled pumping geometry [5] to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig.
+Instead of using a polarization-coupled pumping geometry\cite{Holtom_mode_locked_2006} to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig.
 1(a), where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
@@
 1(a)).
-This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium [1]; in particular, in the vertical direction (Fig.
+This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium\cite{Brenier_new_criteria}; in particular, in the vertical direction (Fig.
 2(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
@@
 Clearly, this can be readily accommodated by the $W_x = 0.3$mm horizontal pump spot size when $w_x \approx 180\mu$m.
-However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium [1].
+However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium\cite{Brenier_new_criteria}.
 
@@
 Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping [31,34].
-Consequently, for our laser polarized along the optical $N_p$-axis [5], there is insufficient information to include accurately the effects of stress-induced refractive index changes.
+Consequently, for our laser polarized along the optical $N_p$-axis\cite{Holtom_mode_locked_2006}, there is insufficient information to include accurately the effects of stress-induced refractive index changes.
 
@@
 The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
-This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW [11], but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis [5,34].
+This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW [11], but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis\cite{Holtom_mode_locked_2006} [34].
 
@@
 4(b)) at an optimum mode-locked output power of $5.0$W.
-This trend is consistent with the expected ‘soliton’ modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity [2].
+This trend is consistent with the expected ‘soliton’ modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity\cite{Brunner_diode_pumped}.
 

### 2012-01-20 inc/laser/laser.tex
Ported OE paper refs up to #15 inclusive

@@
 As with other Yb-doped laser crystals\cite{Brenier_new_criteria}, this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
-Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially [6,7].
-Such lasers have used dichroic mirrors\cite{Brunner_diode_pumped,Major_femtosecond_2006} [8,9], polarization-coupling\cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers [6,10]) to pump the laser crystal in an efficient manner.
+Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially\cite{website_amplitude,website_solar}.
+Such lasers have used dichroic mirrors\cite{Brunner_diode_pumped,Major_femtosecond_2006,Paunescu_diode_2004,Major_extended_2006}, polarization-coupling\cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers\cite{website_amplitude,Brunner_pulses_2002}) to pump the laser crystal in an efficient manner.
 
 % page 8631
-The spectroscopic properties of the biaxial Yb:KGW laser crystal [11] indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength\cite{Brenier_new_criteria}.
-However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) [12,13], mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis\cite{Holtom_mode_locked_2006} [11].
+The spectroscopic properties of the biaxial Yb:KGW laser crystal\cite{Biswal_thermo_optical_05} indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength\cite{Brenier_new_criteria}.
+However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis\cite{Holtom_mode_locked_2006,Biswal_thermo_optical_05}.
 This optimal requirement for orthogonal linear pump and laser polarizations led Holtom\cite{Holtom_mode_locked_2006} to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
@@
 %TODO is this first sentance correct or should I use "I" voice?
-In this section, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis [11,14,15]).
+In this section, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis\cite{Biswal_thermo_optical_05,mochalov_laser_1997,pujol_crystalline_1999}).
 The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media [16].
@@
 %TODO change voice/check ->
-To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk [17,18] and cavity-dumped [19] Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators [6,7,20,21] and oscillators with rod-like Yb-fiber gain media [22] can be expected to yield comparable peak green pulse powers.
+To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk [17,18] and cavity-dumped [19] Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators\cite{website_amplitude,website_solar} [20,21] and oscillators with rod-like Yb-fiber gain media [22] can be expected to yield comparable peak green pulse powers.
 %TODO replace to "This output will then be use to ..." and link to something from previous
@@
 To ensure efficient operation of a Yb:KGW laser, the dominant absorption feature at 980nm for radiation polarized parallel to the optical $N_m$-axis ($\sim\,20^\circ$ from the crystallographic a-axis) should be pumped using high brightness diode lasers.
-On the other hand, as with \ce{Yb$:$KY(WO4)2} [12,13], the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis\cite{Holtom_mode_locked_2006}) in this biaxial material.
+On the other hand, as with \ce{Yb$:$KY(WO4)2} \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis\cite{Holtom_mode_locked_2006}) in this biaxial material.
 %page 8632
@@
 At an angle of incidence of 32.3$^\circ$, this coating theoretically has zero reflection loss for $p$-polarized 1040nm laser radiation parallel to the $b$-axis propagating perpendicular to the $a$-$b$ plane of the crystal (Fig.
-1(b)); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 [11,15].
+1(b)); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 \cite{Biswal_thermo_optical_05,pujol_crystalline_1999}.
 The \ce{SiO2} coating also has an appropriate bandwidth for femtosecond laser operation in excess of 100nm (to the
@@
 Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH [26]).
-The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm fullwidth half-maximum (FWHM) 980nm absorption line of Yb:KGW [8,11,27].
+The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm fullwidth half-maximum (FWHM) 980nm absorption line of Yb:KGW \cite{Paunescu_diode_2004,Biswal_thermo_optical_05} [27].
 Moreover, the two dual-axis collimated laser diodes have full-angle $1/e^2$ intensity beam divergences of $\theta_y \approx$ 4.4mrad and $\theta_x \approx$ 6.8mrad in the horizontal ($x$-direction parallel to the emitter array) and vertical ($y$) directions respectively.
@@
 \end{equation}
-Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98 [11,15]), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
+Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98 \cite{Biswal_thermo_optical_05,pujol_crystalline_1999}), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
 expansion with coefficient $\alpha T$ [16].
@@
 
-We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) [11,14] and stress-induced refractive index changes [31].
+We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) \cite{Biswal_thermo_optical_05,mochalov_laser_1997} and stress-induced refractive index changes [31].
 For the employed crystal cut in our TLS geometry (Fig.
-1(a)), the anisotropy in $\kappa$ [14,27] is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
+1(a)), the anisotropy in $\kappa$ \cite{mochalov_laser_1997} [27] is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
 The refractive index
@@
 The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
-This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW [11], but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis\cite{Holtom_mode_locked_2006} [34].
+This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW \cite{Biswal_thermo_optical_05}, but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis\cite{Holtom_mode_locked_2006} [34].
 

### 2012-01-20 inc/laser/laser.tex
ported oe refs to #21 inclusive

@@
 In this section, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis\cite{Biswal_thermo_optical_05,mochalov_laser_1997,pujol_crystalline_1999}).
-The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media [16].
+The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media\cite{Rimington_thermal_lens_2004}.
 It employs a simple single layer \ce{SiO2} anti-reflection coating on the gain medium to provide minimal loss for $p$-polarized intracavity 1040nm laser radiation and less than 0.5\% surface reflection for orthogonal $s$-polarized 980nm pump light.
@@
 %TODO change voice/check ->
-To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk [17,18] and cavity-dumped [19] Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators\cite{website_amplitude,website_solar} [20,21] and oscillators with rod-like Yb-fiber gain media [22] can be expected to yield comparable peak green pulse powers.
+To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media [22] can be expected to yield comparable peak green pulse powers.
 %TODO replace to "This output will then be use to ..." and link to something from previous
@@
 
-The design of the laser head follows the example of Holtom\cite{Holtom_mode_locked_2006}, which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism [16].
+The design of the laser head follows the example of Holtom\cite{Holtom_mode_locked_2006}, which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism\cite{Rimington_thermal_lens_2004}.
 To ensure efficient operation of a Yb:KGW laser, the dominant absorption feature at 980nm for radiation polarized parallel to the optical $N_m$-axis ($\sim\,20^\circ$ from the crystallographic a-axis) should be pumped using high brightness diode lasers.
@@
 
-To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique [16] to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal.
+To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique\cite{Rimington_thermal_lens_2004} to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal.
 For the employed non-Brewster crystal cut (Fig. 1(a)), a ray
@@
 Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98 \cite{Biswal_thermo_optical_05,pujol_crystalline_1999}), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
-expansion with coefficient $\alpha T$ [16].
+expansion with coefficient $\alpha T$ \cite{Rimington_thermal_lens_2004}.
 Generally, $\Delta \approx W l$ , where $W$ is the transverse size (radius) of the pumped region [30].
@@
 where $\kappa$ is the thermal diffusion coefficient.
-Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry [16].
+Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry\cite{Rimington_thermal_lens_2004}.
 The specifications of our dual-axis collimated diodes [26] indicate that heat source can be written as
@@
 that is, the focused pump radiation is described accurately by a Gaussian irradiance distribution in the horizontal ($x$) direction with half-width 1/e maximum (HW1/eM) spot size $W_x$ and a super-Gaussian of order 3 in the vertical ($y$) direction with spot size $W_y$.
-\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique [16]; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
+\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique\cite{Rimington_thermal_lens_2004}; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
 For a fixed vertical pump spot size of $W_y = 0.2$mm, our analysis indicates that in order to achieve the required ratio $B_T / A_T = 1.3$ in the parabolic approximation for $T(x,y)$ about $(x,y) = (0,0)$ a horizontal pump spot size $W_x = 0.3$mm is needed; in other words, the required $Q(x,y)$ heat source spot size ratio (or ellipticity) $W_x / W_y = 1.5$.
@@
 %page 8635
-changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index [21].
+changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index\cite{website_time_bandwidth}.
 Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping [31,34].

### 2012-01-21 inc/laser/laser.tex
Finished porting oe references. N.B numbers diverged, resorted to using old word doc for cross-check

@@
 %TODO change voice/check ->
-To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media [22] can be expected to yield comparable peak green pulse powers.
+To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media\cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.
 %TODO replace to "This output will then be use to ..." and link to something from previous
-%In our case, the visible femtosecond pulses will be used as the driving laser radiation source for an ultrafast electron microscope [23] – a technology that seeks to combine the sub-nanometer spatial resolution of electron microscopy with the sub-picosecond temporal resolution afforded by today’s ultrashort pulse laser systems.
-%Critical to this technology is likely to be the development of a femtosecond photo-electron gun with a low spatial emittance [25].
-%The coherent control of photoemission through visible laser-driven plasmon excitations on large-area (a few mm2), gold or silver, nano-patterned photocathodes may provide a mechanism by which this can be achieved [25].
+%In our case, the visible femtosecond pulses will be used as the driving laser radiation source for an ultrafast electron microscope \cite{king_ultrafast_2005} – a technology that seeks to combine the sub-nanometer spatial resolution of electron microscopy with the sub-picosecond temporal resolution afforded by today’s ultrashort pulse laser systems.
+%Critical to this technology is likely to be the development of a femtosecond photo-electron gun with a low spatial emittance\cite{zawadzka_evanescent_2001}.
+%The coherent control of photoemission through visible laser-driven plasmon excitations on large-area (a few mm2), gold or silver, nano-patterned photocathodes may provide a mechanism by which this can be achieved\cite{zawadzka_evanescent_2001}.
 
@@
 
-Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH [26]).
-The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm fullwidth half-maximum (FWHM) 980nm absorption line of Yb:KGW \cite{Paunescu_diode_2004,Biswal_thermo_optical_05} [27].
+Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH\cite{website_limo}).
+The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm fullwidth half-maximum (FWHM) 980nm absorption line of Yb:KGW \cite{Paunescu_diode_2004,Biswal_thermo_optical_05,website_EKSPLA}.
 Moreover, the two dual-axis collimated laser diodes have full-angle $1/e^2$ intensity beam divergences of $\theta_y \approx$ 4.4mrad and $\theta_x \approx$ 6.8mrad in the horizontal ($x$-direction parallel to the emitter array) and vertical ($y$) directions respectively.
@@
 The Yb concentration was chosen to be 2\% atomic doping to ensure that more than 90\% of the pump radiation was absorbed over the 3mm Yb:KGW crystal thickness.
-With the employed counter-propagating pump geometry, the resultant maximum average pump irradiance over the crystal length is then roughly 3 times the $\sim$6kW/cm$^2$ saturation intensity of the dominant 980nm absorption line [28].
-The laser crystal was obtained from NovaPhase [29].
+With the employed counter-propagating pump geometry, the resultant maximum average pump irradiance over the crystal length is then roughly 3 times the $\sim$6kW/cm$^2$ saturation intensity of the dominant 980nm absorption line\cite{Brunner_powerful_2004}.
+The laser crystal was obtained from NovaPhase\cite{website_nova}.
 
@@
 expansion with coefficient $\alpha T$ \cite{Rimington_thermal_lens_2004}.
-Generally, $\Delta \approx W l$ , where $W$ is the transverse size (radius) of the pumped region [30].
+Generally, $\Delta \approx W l$ , where $W$ is the transverse size (radius) of the pumped region\cite{Koechner_thermal_1970}.
 In our case, $l \approx$ 3mm and $W \approx$ 0.2mm, so that $\Delta < 0.1$, implying that the thermal duct due to a non-zero $\frac{dn}{dT}$ primarily determines the required $B_T$/$A_T$ ratio.
@@
 Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry\cite{Rimington_thermal_lens_2004}.
-The specifications of our dual-axis collimated diodes [26] indicate that heat source can be written as
+The specifications of our dual-axis collimated diodes\cite{website_limo} indicate that heat source can be written as
 \begin{equation}
@@
 
-We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) \cite{Biswal_thermo_optical_05,mochalov_laser_1997} and stress-induced refractive index changes [31].
+We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) \cite{Biswal_thermo_optical_05,mochalov_laser_1997} and stress-induced refractive index changes\cite{Yumashev_laser_2007}.
 For the employed crystal cut in our TLS geometry (Fig.
-1(a)), the anisotropy in $\kappa$ \cite{mochalov_laser_1997} [27] is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
+1(a)), the anisotropy in $\kappa$ \cite{mochalov_laser_1997,website_EKSPLA} is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
 The refractive index
@@
 changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index\cite{website_time_bandwidth}.
-Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping [31,34].
+Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping\cite{Yumashev_laser_2007,Hellstrom_efficient_2006}.
 Consequently, for our laser polarized along the optical $N_p$-axis\cite{Holtom_mode_locked_2006}, there is insufficient information to include accurately the effects of stress-induced refractive index changes.
@@
 The diode-pumped TLS Yb:KGW laser head is placed in the 2.35m-long, asymmetric z-fold cavity shown in Fig. 3(a).
-The two focusing gain section mirrors each have a group velocity dispersion (GVD) of $-1300 ( \pm 150 ) \text{fs}^2$ (Layertec GmbH [32]), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
+The two focusing gain section mirrors each have a group velocity dispersion (GVD) of $-1300 ( \pm 150 ) \text{fs}^2$ (Layertec GmbH\cite{website_layertec}), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
 The 5$^\circ$ angle of incidence on the gain section mirrors results in minimal astigmatism, although any net round-trip cavity astigmatism can be compensated for by the TLS technique.
@@
 %page 8636
-focusing the intracavity radiation on a saturable Bragg reflector (SBR) with a 1\% reflectivity modulation depth at 1040nm (BATOP GmbH [33]) positioned a distance $z$ from the focusing mirror.
+focusing the intracavity radiation on a saturable Bragg reflector (SBR) with a 1\% reflectivity modulation depth at 1040nm (BATOP GmbH\cite{website_BATOP}) positioned a distance $z$ from the focusing mirror.
 The shorter 63cm arm is terminated by a plane output coupler with a 7\% transmission at 1040nm.
@@
 The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
-This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW \cite{Biswal_thermo_optical_05}, but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis\cite{Holtom_mode_locked_2006} [34].
+This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW \cite{Biswal_thermo_optical_05}, but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis\cite{Holtom_mode_locked_2006,Hellstrom_efficient_2006}.
 
@@
 The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (Fig. 4(a)).
-The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH [32]) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
+The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH\cite{website_layertec}) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
 For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (Fig.
@@
 4(a)), but is always less than the 50cm focal length of the concave high reflector.
-This is due to the need to match the absorption-induced thermal bowing of the SBR (i.e., the radius of curvature of its front surface mirror) to the radius of curvature of the intra-cavity laser mode [35].
+This is due to the need to match the absorption-induced thermal bowing of the SBR (i.e., the radius of curvature of its front surface mirror) to the radius of curvature of the intra-cavity laser mode\cite{Schieffer_dual_passive_2006}.
 In general, good ML operation is achieved with the SBR positioned within $\pm$1cm from the optimum $z$ distance.
@@
 This is consistent both with the laser accessing more of the available inhomogeneous gain bandwidth and with our resonator design favoring operation with a strong Kerr lens in the Yb:KGW gain medium.
-Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal’s optical $N_p$-axis [36-38] and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
+Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal’s optical $N_p$-axis\cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
 From Fig.
@@
 At the 3.5W ML output power, with a pulse duration of 250fs, the estimated focal irradiance for the fundamental is $\sim50\text{GW}/\text{cm}^2$ in the LBO doubling crystal.
-Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH [32]), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
+Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH\cite{website_layertec}), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
 After the collimating
@@
 
-This frequency doubling geometry for LBO is chosen to be close to the optimum for efficient short pulse doubling [39,40] defined by limits imposed by group-velocity walk-off between the fundamental and second harmonic pulses, spatial walk-off in the critical phase-matching, and the ratio of the confocal parameter of the focused fundamental wave to the LBO crystal length.
+This frequency doubling geometry for LBO is chosen to be close to the optimum for efficient short pulse doubling \cite{wang_efficiency_2003,Saltiel_second_harmonic_2004} defined by limits imposed by group-velocity walk-off between the fundamental and second harmonic pulses, spatial walk-off in the critical phase-matching, and the ratio of the confocal parameter of the focused fundamental wave to the LBO crystal length.
 In addition, the $z$-fold configuration allows for astigmatism compensation (e.g., due to the Brewster LBO crystal) of the second harmonic output by adjustment of the angle of incidence on the second collimating mirror.
@@
 a peak green pulse power of $\sim$130kW.
-This femtosecond visible output will be used as the driving pulsed laser radiation source for an ultrafast electron microscope [23].
+This femtosecond visible output will be used as the driving pulsed laser radiation source for an ultrafast electron microscope\cite{king_ultrafast_2005}.
 

### 2012-01-23 inc/laser/laser.tex
moved oe figure 1a (tikz form) into laser.tex (input) added figure 1 caption

--- 2012-01-20 inc/laser/laser.tex
@@
 Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
-As with other Yb-doped laser crystals\cite{Brenier_new_criteria}, this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
-Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially\cite{website_amplitude,website_solar}.
-Such lasers have used dichroic mirrors\cite{Brunner_diode_pumped,Major_femtosecond_2006,Paunescu_diode_2004,Major_extended_2006}, polarization-coupling\cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers\cite{website_amplitude,Brunner_pulses_2002}) to pump the laser crystal in an efficient manner.
+As with other Yb-doped laser crystals [1], this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
+Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated [2-5] and are now available commercially [6,7].
+Such lasers have used dichroic mirrors [2,4,8,9], polarization-coupling [5], or more complex optical geometries (e.g., thin disk lasers [6,10]) to pump the laser crystal in an efficient manner.
 
 % page 8631
-The spectroscopic properties of the biaxial Yb:KGW laser crystal\cite{Biswal_thermo_optical_05} indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength\cite{Brenier_new_criteria}.
-However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis\cite{Holtom_mode_locked_2006,Biswal_thermo_optical_05}.
-This optimal requirement for orthogonal linear pump and laser polarizations led Holtom\cite{Holtom_mode_locked_2006} to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
+The spectroscopic properties of the biaxial Yb:KGW laser crystal [11] indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength [1].
+However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) [12,13], mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis [5,11].
+This optimal requirement for orthogonal linear pump and laser polarizations led Holtom [5] to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
 
 %TODO is this first sentance correct or should I use "I" voice?
-In this section, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis\cite{Biswal_thermo_optical_05,mochalov_laser_1997,pujol_crystalline_1999}).
+In this section, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis [11,14,15]).
 The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media [16].
@@
 %TODO change voice/check ->
-To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk [17,18] and cavity-dumped [19] Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators\cite{website_amplitude,website_solar} [20,21] and oscillators with rod-like Yb-fiber gain media [22] can be expected to yield comparable peak green pulse powers.
+To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk [17,18] and cavity-dumped [19] Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators [6,7,20,21] and oscillators with rod-like Yb-fiber gain media [22] can be expected to yield comparable peak green pulse powers.
 %TODO replace to "This output will then be use to ..." and link to something from previous
@@
 
-The design of the laser head follows the example of Holtom\cite{Holtom_mode_locked_2006}, which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism [16].
+The design of the laser head follows the example of Holtom [5], which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism [16].
 To ensure efficient operation of a Yb:KGW laser, the dominant absorption feature at 980nm for radiation polarized parallel to the optical $N_m$-axis ($\sim\,20^\circ$ from the crystallographic a-axis) should be pumped using high brightness diode lasers.
-On the other hand, as with \ce{Yb$:$KY(WO4)2} \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis\cite{Holtom_mode_locked_2006}) in this biaxial material.
+On the other hand, as with \ce{Yb$:$KY(WO4)2} [12,13], the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis [5]) in this biaxial material.
 %page 8632
-Instead of using a polarization-coupled pumping geometry\cite{Holtom_mode_locked_2006} to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig.
+Instead of using a polarization-coupled pumping geometry [5] to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig.
 1(a), where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
 At an angle of incidence of 32.3$^\circ$, this coating theoretically has zero reflection loss for $p$-polarized 1040nm laser radiation parallel to the $b$-axis propagating perpendicular to the $a$-$b$ plane of the crystal (Fig.
-1(b)); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 \cite{Biswal_thermo_optical_05,pujol_crystalline_1999}.
+1(b)); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 [11,15].
 The \ce{SiO2} coating also has an appropriate bandwidth for femtosecond laser operation in excess of 100nm (to the
@@
 
-% insert figure 1 here
+\begin{figure}
+  \subfloat[][]{
+    \label{fig:crystal_geo}
+    \input{crystal_geo}
+  }
+  \caption{
+\subref{fig:crystal_geo} Top view of the 3x3x10mm AR coated Yb:KGW laser crystal showing the crystal cut, the direction of propagation of p-polarized laser radiation at 1040nm, and the $x$ (tangential) and $y$ (saggital) directions (in blue) employed in the thermal analysis;
+(b) Calculated reflectivity of the 193nm \ce{SiO2} AR coating as a function of incidence angle for $p$-polarized 1040nm laser radiation (solid line) and $s$-polarized 980nm pump radiation (dashed line).
+}
+\end{figure}
 
 Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH [26]).
-The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm fullwidth half-maximum (FWHM) 980nm absorption line of Yb:KGW \cite{Paunescu_diode_2004,Biswal_thermo_optical_05} [27].
+The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm fullwidth half-maximum (FWHM) 980nm absorption line of Yb:KGW [8,11,27].
 Moreover, the two dual-axis collimated laser diodes have full-angle $1/e^2$ intensity beam divergences of $\theta_y \approx$ 4.4mrad and $\theta_x \approx$ 6.8mrad in the horizontal ($x$-direction parallel to the emitter array) and vertical ($y$) directions respectively.
@@
 \end{equation}
-Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98 \cite{Biswal_thermo_optical_05,pujol_crystalline_1999}), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
+Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98 [11,15]), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
 expansion with coefficient $\alpha T$ [16].
@@
 1(a)).
-This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium\cite{Brenier_new_criteria}; in particular, in the vertical direction (Fig.
+This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium [1]; in particular, in the vertical direction (Fig.
 2(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
@@
 Clearly, this can be readily accommodated by the $W_x = 0.3$mm horizontal pump spot size when $w_x \approx 180\mu$m.
-However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium\cite{Brenier_new_criteria}.
+However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium [1].
 
-We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) \cite{Biswal_thermo_optical_05,mochalov_laser_1997} and stress-induced refractive index changes [31].
+We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) [11,14] and stress-induced refractive index changes [31].
 For the employed crystal cut in our TLS geometry (Fig.
-1(a)), the anisotropy in $\kappa$ \cite{mochalov_laser_1997} [27] is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
+1(a)), the anisotropy in $\kappa$ [14,27] is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
 The refractive index
@@
 Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping [31,34].
-Consequently, for our laser polarized along the optical $N_p$-axis\cite{Holtom_mode_locked_2006}, there is insufficient information to include accurately the effects of stress-induced refractive index changes.
+Consequently, for our laser polarized along the optical $N_p$-axis [5], there is insufficient information to include accurately the effects of stress-induced refractive index changes.
 
@@
 The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
-This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW \cite{Biswal_thermo_optical_05}, but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis\cite{Holtom_mode_locked_2006} [34].
+This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW [11], but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis [5,34].
 
@@
 4(b)) at an optimum mode-locked output power of $5.0$W.
-This trend is consistent with the expected ‘soliton’ modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity\cite{Brunner_diode_pumped}.
+This trend is consistent with the expected ‘soliton’ modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity [2].
 

### 2012-01-23 inc/laser/laser.tex

--- 2012-01-21 inc/laser/laser.tex
@@
 
-% insert figure 1 here
+\begin{figure}
+  \subfloat[][]{
+    \label{fig:crystal_geo}
+    \input{crystal_geo}
+  }
+  \caption{
+\subref{fig:crystal_geo} Top view of the 3x3x10mm AR coated Yb:KGW laser crystal showing the crystal cut, the direction of propagation of p-polarized laser radiation at 1040nm, and the $x$ (tangential) and $y$ (saggital) directions (in blue) employed in the thermal analysis;
+(b) Calculated reflectivity of the 193nm \ce{SiO2} AR coating as a function of incidence angle for $p$-polarized 1040nm laser radiation (solid line) and $s$-polarized 980nm pump radiation (dashed line).
+}
+\end{figure}
 

### 2012-02-03 inc/laser/laser.tex
stitched schematic into laser section

--- 2012-01-21 inc/laser/laser.tex
@@
 
-% insert figure 1 here
+\begin{figure}
+  \subfloat[][]{
+    \label{fig:crystal_geo}
+    \input{crystal_geo}
+  }
+  \caption{
+\protect\subref{fig:crystal_geo} Top view of the 3x3x10mm AR coated Yb:KGW laser crystal showing the crystal cut, the direction of propagation of p-polarized laser radiation at 1040nm, and the $x$ (tangential) and $y$ (saggital) directions (in blue) employed in the thermal analysis;
+(b) Calculated reflectivity of the 193nm \ce{SiO2} AR coating as a function of incidence angle for $p$-polarized 1040nm laser radiation (solid line) and $s$-polarized 980nm pump radiation (dashed line).
+}
+\end{figure}
 
@@
 The longer 75cm arm of the resonator is terminated by a concave ($R$ = 1m and angle of incidence $< 2^\circ$) high reflector
-% insert figure 3
+
+\begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:laser_schematic}
+    \input{schematic}
+  }
+  \caption{
+\protect\subref{fig:laser_schematic} Schematic of the femtosecond TLS Yb:KGW oscillator (LD = laser diode, OC = output coupler, SBR = saturable Bragg reflector).
+(b) Laser mode size in the center of the Yb:KGW gain medium ($w$) as a function of the strength of the lensing ($1/f$) in the gain medium (thermal and Kerr) obtained from a resonator stability analysis with $z \approx 42$cm for both the tangential ($x$, red line) and saggital ($y$, dashed black line) directions.
+The 0.2mm (HW1/eM) vertical super-Gaussian pump spot size is indicated by the horizontal line.
+}
+\end{figure}
+
 %page 8636

### 2012-02-24 inc/magnetic_lenses/maglens.tex
Added magnetic lenses section. So far, includes some motivation and introduction, now moving to designing via el-kareh book

\subsection{Motivation}
% montgomery_some_1961
% el-kareh_electron_1970

In many ways, an electron beam can be thought of in terms of an optical equivalent system. 
Of course there are notable exceptions; photons don't repel each other; photon energies are hard to change.
Electron optical systems do have an analogous system to the lens, in this case using magnetic fields rather than glass. 
The magnetic lens is a common element in electron microscopes and other electron-optical systems.
%TODO refs needed

Magnetic lenses present some challenges for Ultrafast Electron Microscopy.
First, the beams are likely to be much larger and thus a UEM will need large-aperture lenses.
Second, and more importantly, near a lens crossover (i.e. focus), the charge density can increase rapidly.
This charge density can lead to an unacceptable amount of space-charge interaction (see section \ref{sec:model_results}).

\subsection{Introduction to Magnetic Lenses}

A simple magnetic lens is just a solenoid.
The axis the lens is collinear with the axis of the microscope column.
Initially one would think that electrons moving parallel to this axis would not be affected by the magnetic field lines in the solenoid, however the contribution of the fringing field must be taken into consideration.
As electrons propagate through the solenoid the fringing magnetic field imparts an azimuthal velocity.
This ``tumbling'' around the axis, along with the axial magnetic field, is what causes the inward-directed force.

A full description of the motion of the electrons is easily derived \cite{el-kareh_electron_1970},
\begin{subequations} \label{eq:lens_eq_of_motion}
\begin{gather}
  m \ddot{r} = -e B_z r \dot{\theta} + m r \dot{\theta}^2                 \\
  \frac{d}{dt} ( m r^2 \dot{\theta} ) = e r \dot{r} B_z - e r B_r \dot{z} \\
  m \ddot{z} = e B_r r \dot{\theta}
\end{gather}
\end{subequations}
however, attempting to further seperate these equations quickly highlights the tightly coupled nature of the motion of the electron in this system.
The full description of the off-axis field of a real (non-infinite) air-core solenoid is also surprisingly complicated.
Montgomery and Terrell \cite{montgomery_some_1961} provide a full treatment for many lens configurations, including a tractable treatment for a single current loop of radius $a$ in cylindrical coordinates, 
\begin{subequations} \label{eq:field_of_loop}
\begin{gather}
  H_z = \frac{2I}{\sqrt{Q}} \left(   F(k) + \frac{ a^2 - r^2 - z^2 }{ (1-k^2) Q } E(k) \right) \\
  H_r = \frac{2I}{\sqrt{Q}} \left( - F(k) + \frac{ a^2 + r^2 + z^2 }{ (1-k^2) Q } E(k) \right) \\
  Q \equiv (a+r)^2 + z^2 \\
  k \equiv \sqrt{ 4 a r / Q }
\end{gather}
\end{subequations}
%%TODO I can't decide whether the Q and k definitions should be numbered or not
%where
%\begin{gather*}
%  Q \equiv (a+r)^2 + z^2 \\
%  k \equiv \sqrt{ 4 a r / Q }
%\end{gather*}
and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
An even more accurate solenoid can be modeled by appropriately summing the fields of many such loops.

\subsection{Design}

While in principle one could design a magnetic lens using equations \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
Further --- as suggested by the use of the $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility; this component is called a ``pole piece''.

### 2012-02-24 inc/magnetic_lenses/maglens.tex
Continue magnetic lenses. added schematic figure

@@
 
-While in principle one could design a magnetic lens using equations \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
-Further --- as suggested by the use of the $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility; this component is called a ``pole piece''.
+While in principle one could design a magnetic lens using \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
+Further --- as suggested by the use of the $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility.
+These materials, called ``magnetic shielding'' since it captures the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.
+The field lines exit the shielding in the gap and it is this field that acts as the lens.
+This gap can be much smaller than the total width of the coil turns, and yet contain in a field commensurate with the full coil.
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+  \foreach \n in {0,0.1,...,0.6}{
+    \draw [red] (0.75,-0.5) ellipse [x radius=\n,y radius=0.6];
+  }
+  \foreach \n in {0,0.1,...,0.6}{
+    \draw [red] (-0.75,-0.5) ellipse [x radius=\n,y radius=0.6];
+  }
+  \draw [fill=blue!50] 
+    (-0.5,0)
+    -- ++(0,2)
+    -- ++(-2,0)
+    -- ++(0,-5)
+      node [pos=0.5,left,align=right] {Magnetic\\Shielding}
+    -- ++(2,0)
+    -- ++(0,2)
+      coordinate [pos=0] (D left)
+    -- ++(-0.5,0)
+      coordinate [pos=0.5] (S top)
+    -- ++(0,-1.5)
+    -- ++(-1,0)
+    -- ++(0,4)
+    -- ++(1,0)
+    -- ++(0,-1.5)
+    -- ++(0.5,0)
+      coordinate [pos=0.5] (S bottom)
+    -- cycle
+  ;
+  \draw [fill=blue!50] 
+    (0.5,0)
+    -- ++(0,2)
+    -- ++(2,0)
+    -- ++(0,-5)
+    -- ++(-2,0)
+    -- ++(0,2)
+      coordinate [pos=0] (D right)
+    -- ++(0.5,0)
+    -- ++(0,-1.5)
+    -- ++(1,0)
+    -- ++(0,4)
+    -- ++(-1,0)
+    -- ++(0,-1.5)
+    -- cycle
+  ;
+  \draw [thick, dashed]
+    (0,3)
+    -- (0,-3)
+  ;
+  \draw [very thick,<->] 
+    (S top) 
+    -- (S bottom)
+      node [pos=0.5,fill=white] {$S$}
+  ;
+  \draw [very thick,<->] 
+    (D left) 
+    -- (D right)
+      node [below,pos=0.5,align=center] {$D$\\$(=2R)$}
+  ;
+  \node [rotate=90] at (-1.6,-0.5) {Windings};
+  \node [red,rotate=90] at (1.6,-0.5) {Field Lines};
+\end{tikzpicture}
+\end{figure}
+
+El-Kareh and El-Kareh \cite{el-kareh_electron_1970} provides an analysis of symmetric lenses with magnetic shielding and a pole-piece in chapter 8, with the most useful information being in section 8.7.

### 2012-02-24 inc/magnetic_lenses/maglens.tex
added specifics about the our lenses (parameters).

@@
     -- (0,-3)
+      node [pos=0,above] {Axis of Electron Microscope Column}
   ;
@@
 \end{tikzpicture}
+\caption{Side-view schematic of a shielded magnetic lens with a pole-piece.
+The windings (current loops) are contained within the shielding.
+The magnetic field created by these windings is contained inside the shielding except in the gap.}
 \end{figure}
 
-El-Kareh and El-Kareh \cite{el-kareh_electron_1970} provides an analysis of symmetric lenses with magnetic shielding and a pole-piece in chapter 8, with the most useful information being in section 8.7.
+El-Kareh and El-Kareh \cite{el-kareh_electron_1970} provide an analysis of symmetric lenses with magnetic shielding and a pole-piece in chapter 8, with the most useful information being in section 8.7.
+They define
+\begin{subequations}
+\begin{gather}
+  k^2 \equiv \frac{e}{8 m V_r} B_0^2 R^2 \\
+  \beta \equiv k^2 V_r / (NI)^2 \label{eq:elkareh_k2}
+\end{gather}
+\end{subequations}
+where $N$ is the number of loops, $I$ is the current in a single loop, $B_0$ is the maximum axial magnetic field strength, $R$ is the bore radius and $V_r$ is the relativity corrected acceleration potential, given by
+\begin{equation}
+  V_r = V ( 1 + \frac{0.978 \cdot 10^{-6}}{\text{Volt}} V )
+\end{equation}
+They then derive a number of formulae and tabulate the results for a variety of parameters.
+Using these tabulated results, and for a choice of parameters one can determine the number of ampere-turns ($NI$) to create a certain focal length.
+
+%TODO check S and D as I seem to remember we have S=0.1in
+%TODO need a description of the steel inner tube causing an effective diameter larger than the actual bore
+For the UEM, I chose a bore diameter $D$=0.3in and a pole-piece gap $S$=0.5in, for an ratio $S/D=0.6$.
+By their table 8.2a this results in $\beta=0.0104$. Then, to create a focal length of 4in ($f/R=16$) table 8.17 gives a value for $k^2 \approx 0.05$.
+Finally substituting this value into \ref{eq:elkareh_k2}, with $V=30$kV ($V_r\approx30.88$kV), one can deduce that $NI\approx385$A.
+
+To further ease the burden on the lenses, I actually created two, one placed to either side of the RF cavity. As each has 100 turns, a current of only 2A is needed to drive the two lenses in series.

### 2012-03-16 inc/magnetic_lenses/maglens.tex
update magnetic lens numbers because my notes were incorrect

@@
 %TODO need a description of the steel inner tube causing an effective diameter larger than the actual bore
-For the UEM, I chose a bore diameter $D$=0.3in and a pole-piece gap $S$=0.5in, for an ratio $S/D=0.6$.
-By their table 8.2a this results in $\beta=0.0104$. Then, to create a focal length of 4in ($f/R=16$) table 8.17 gives a value for $k^2 \approx 0.05$.
-Finally substituting this value into \ref{eq:elkareh_k2}, with $V=30$kV ($V_r\approx30.88$kV), one can deduce that $NI\approx385$A.
+For the UEM, I chose a bore diameter $D=2R=$0.1in and a pole-piece gap $S$=0.5in, for an ratio $S/D=0.2$.
+By their table 8.2a this results in $\beta=0.0146$. Then, to create a focal length of 4in ($f/R=16$) table 8.13 gives a value for $k^2 \approx 0.06$.
+Finally substituting this value into \ref{eq:elkareh_k2}, with $V=30$kV ($V_r\approx30.88$kV), one can deduce that $NI\approx356$A.
 
-To further ease the burden on the lenses, I actually created two, one placed to either side of the RF cavity. As each has 100 turns, a current of only 2A is needed to drive the two lenses in series.
+To further ease the burden on the lenses, I actually created two, one placed to either side of the RF cavity.
+As each has 100 turns, a current of only 2A is needed to drive the two lenses in series.
+The oxygen-free copper wire coils remain cool to the touch under these operating conditions.

### 2012-04-07 inc/semicond/semicond.tex
Included recent progress on upper conduction band absorption

\subsection{Theoretical Basis}

\subsubsection{Band Structure}

\subsubsection{Population of Upper Conduction Band}

\subsubsection{Absorption into Upper Conduction Band}

In order to quantify the behavior of emission from this upper conduction band it is necessary to calculate the absorption $\alpha_0$ into this band.
The text `Optoelectronics' [cite] has a formula for absorption in terms of the energy gap $E_g$, the effective masses and a few optical properties.
\begin{equation} \label{eq:absorption_optoelectronics}
  \alpha_0(\omega) = \frac{ q^2 x_{vc}^2 (2 m_r)^{3/2} }{ \lambda \varepsilon_0 \hbar^3 n_{op} } \sqrt{\hbar\omega - E_g}
\end{equation}
Here $\omega$ and $\lambda$ are the frequency and wavelength of the incoming photons respectively.
$m_r$ is the reduced mass of the system given by
\begin{equation}
  \frac{1}{m_r} = \frac{1}{m_c} + \frac{1}{m_v}
\end{equation}
in terms of the effective mass of the valence holes ($m_v$) and conduction electrons ($m_c$).
The quantity $x_{vc}^2$ is
\begin{equation}
  x_{vc}^2 = \frac{1}{3} \frac{\hbar^2}{E_g^2} \frac{E_p}{m_0}
\end{equation}
$m_0$ is the standard electron mass. $E_p$ is the Kane energy which is normally calculated from the matrix elements of the band structure.
It can be estimated using [Optoelectronics] (5.C.19)
\begin{equation}
  \frac{m_c}{m_0} = \left( 1 + \frac{E_p}{E_g} \right)^{-1}
\end{equation}
A further approximation is made owing to the difficulty in finding accurate data on these upper conduction bands.
Visually one can recognize that the curvature of the valence band this upper conduction band are comparable, meaning that approximating $ m_c \approx m_v $ is reasonable for order of magnitude calculations. With this approximation, \ref{eq:absorption_optoelectronics} becomes 
\begin{equation}
    \alpha_0(\omega) = \frac{ q^2 x_{vc}^2 m_v^{3/2} }{ \lambda \varepsilon_0 \hbar^3 n_{op} } \sqrt{\hbar\omega - E_g}
\end{equation}

\begin{table} 
  \begin{tabular}{|l|l|c|c|c|}
      \hline
    Material  & UC Band & $m_v$ $(m_0)$ & $E_g$ (eV) & $n_{op}$ \\
      \hline\hline
    \ce{GaSb} & $\Gamma(8c)$ & 0.28 & 3.77 & 1.45 \\
    \ce{InSb} & $\Gamma(8c)$ & 0.45 & 3.45 & 1.40 \\
      \hline
  \end{tabular}
  \caption{Data and results from analysis of the absorption into upper conduction bands of useful III-V semiconductors.}
  \label{table:absorption}
\end{table}
Using data collected from [Springer Ref], \ref{table:absorption} shows the results of these calculations for useful III-V semiconductors.

### 2012-04-18 inc/magnetic_lenses/maglens.tex
separate polepiece schematic into input'ed file

@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \subsection{Motivation}
@@
   \centering
-\begin{tikzpicture}
-  \foreach \n in {0,0.1,...,0.6}{
-    \draw [red] (0.75,-0.5) ellipse [x radius=\n,y radius=0.6];
+  \input{polepiece-schematic}
+  \caption{
+    Side-view schematic of a shielded magnetic lens with a pole-piece.
+    The windings (current loops) are contained within the shielding.
+    The magnetic field created by these windings is contained inside the shielding except in the gap.
   }
-  \foreach \n in {0,0.1,...,0.6}{
-    \draw [red] (-0.75,-0.5) ellipse [x radius=\n,y radius=0.6];
-  }
-  \draw [fill=blue!50] 
-    (-0.5,0)
-    -- ++(0,2)
-    -- ++(-2,0)
-    -- ++(0,-5)
-      node [pos=0.5,left,align=right] {Magnetic\\Shielding}
-    -- ++(2,0)
-    -- ++(0,2)
-      coordinate [pos=0] (D left)
-    -- ++(-0.5,0)
-      coordinate [pos=0.5] (S top)
-    -- ++(0,-1.5)
-    -- ++(-1,0)
-    -- ++(0,4)
-    -- ++(1,0)
-    -- ++(0,-1.5)
-    -- ++(0.5,0)
-      coordinate [pos=0.5] (S bottom)
-    -- cycle
-  ;
-  \draw [fill=blue!50] 
-    (0.5,0)
-    -- ++(0,2)
-    -- ++(2,0)
-    -- ++(0,-5)
-    -- ++(-2,0)
-    -- ++(0,2)
-      coordinate [pos=0] (D right)
-    -- ++(0.5,0)
-    -- ++(0,-1.5)
-    -- ++(1,0)
-    -- ++(0,4)
-    -- ++(-1,0)
-    -- ++(0,-1.5)
-    -- cycle
-  ;
-  \draw [thick, dashed]
-    (0,3)
-    -- (0,-3)
-      node [pos=0,above] {Axis of Electron Microscope Column}
-  ;
-  \draw [very thick,<->] 
-    (S top) 
-    -- (S bottom)
-      node [pos=0.5,fill=white] {$S$}
-  ;
-  \draw [very thick,<->] 
-    (D left) 
-    -- (D right)
-      node [below,pos=0.5,align=center] {$D$\\$(=2R)$}
-  ;
-  \node [rotate=90] at (-1.6,-0.5) {Windings};
-  \node [red,rotate=90] at (1.6,-0.5) {Field Lines};
-\end{tikzpicture}
-\caption{Side-view schematic of a shielded magnetic lens with a pole-piece.
-The windings (current loops) are contained within the shielding.
-The magnetic field created by these windings is contained inside the shielding except in the gap.}
 \end{figure}

### 2012-05-03 inc/ms_model/ms_model.tex
added K^{flow} matrix result and began the final M&S AG equation set
Added new Liouville result (big gamma)

@@
 \end{gather}
+Obviously the $K^{flow}$ term is far easier to evaluate
+\begin{equation}
+  K^{flow} = 
+  \begin{pmatrix}
+    k^{flow}_{\smallT} & 0 & 0 \\
+    0 & k^{flow}_{\smallT} & 0 \\
+    0 & 0 & k^{flow}_{z}
+  \end{pmatrix}
+  \quad \text{and} \quad
+  k^{flow}_{\alpha} = 
+  \begin{pmatrix}
+    \dfrac{2\sigma_{\alpha}}{m} & \dfrac{1}{m} \left(\eta_{\alpha} + \dfrac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) \\
+    \dfrac{1}{m} \left(\eta_{\alpha} + \dfrac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) & 0
+  \end{pmatrix}
+\end{equation}
+
+\subsection{The Differential Equation System}
+
+The results of the previous analyses culminate in the differential equation set
+\begin{subequations} \label{eq:ag_original}
+\begin{gather}
+  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
+  \frac{d\gamma_{\alpha}}{dt} = \frac{1}{m} \left(\eta_{\alpha} + \frac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) 
+    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi)\\
+  \frac{d\eta_{\alpha}}{dt} = - \frac{2 \gamma_{\alpha} \eta_{\alpha}}{m \sigma_{\alpha}}
+\end{gather}
+\end{subequations}
+which depends on the ellipticity parameter
+\begin{equation}
+  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}}
+\end{equation}
+
+\subsection{Implications of Liouville's Theorem}
+
+One of the most well known theorems in physics (and mathematics) is Liouville's Theorem.
+This theorem can be summarized succinctly as a `the total phase space of a system is conserved.'
+%TODO find a good reference for this
+
+To evaluate the effect of this theorem on these equations, consider the time derivative of the product $\sigma_{\alpha} \eta_{\alpha}$.
+Using \ref{eq:ag_original} it is easy to show that
+\begin{gather}
+  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0\\
+  \label{eq:define_big_gamma}
+  \therefore \sigma_{\alpha} \eta_{\alpha} \equiv \Gamma_{\alpha}^2 = \text{Constant}
+\end{gather}
+This analysis shows that the Analytic Gaussian model explicitly obeys the Liouville Theorem.
+
+Not only is this an interesting result, but it can be both instructive and computationally useful.
+We can rewrite \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$ and substitute into \ref{eq:ag_original}.
+\begin{subequations} \label{eq:ag_big_gamma}
+\begin{gather}
+  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
+  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
+    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi)
+\end{gather}
+\end{subequations}
+These four equations now contain all of the information previously contained in six, making this much less computationally intensive.
+Further, now one can see that these $\Gamma$ constants can be thought of as some minimum/intrinsic action of the pulse. 
+This will become more apparent when in future sections
+%TODO include internal reference
+it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.

### 2012-05-03 inc/ms_model/ms_model.tex
expanded on the original model

@@
 \end{equation}
+These analyses will be referred to while deriving the influence of external forces.
+%TODO internal reference
+
+The $K^{force}$ integrals, related to the internal force of the pulse's Coulombic repulsion, are far more complicated to evaluate.
+For more information about the derivation, see \cite{michalik_analytic_2006}.
+The final matrices are
+\begin{equation}
+  K^{force} = 
+  \begin{pmatrix}
+    k^{force}_{\smallT} & 0 & 0 \\
+    0 & k^{force}_{\smallT} & 0 \\
+    0 & 0 & k^{force}_{z}
+  \end{pmatrix}
+\end{equation}
+where
+\begin{equation}
+  k^{force}_{\alpha} = 
+  \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{2\pi^{1/2}\sigma_{\alpha}^{1/2}} L_{\alpha}(\xi)
+  \begin{pmatrix}
+    0 & 1 \\
+    1 & \dfrac{2 \gamma_{\alpha}}{\sigma_{\alpha}}
+  \end{pmatrix}
+\end{equation}
+The parameter $\xi$ is the ellipticity parameter, defined as
+\begin{equation}
+  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}}
+\end{equation}
+Note that the original paper of Michalik and Sipe contained an error which incorrectly defined $\xi$; this was corrected in an erratum\cite{michalik_erratum:_2008}.
+The family of $L$ functions are
+\begin{gather}
+  L_{\smallT}(\xi) = \frac{3}{2} \left( L(\xi) + \frac{\xi^2 L(\xi) - \xi}{ 1 - \xi^2 } \right) \\
+  L_{z}(\xi) = \frac{ 3 \xi^2 }{ \xi^2 - 1} \left( \xi L(\xi) - 1 \right) \\
+  L(\xi) = \frac{1}{2} \int_0^{\pi} \frac{1}{1+\xi\sin(\theta)} d\theta
+\end{gather}
 
@@
 Not only is this an interesting result, but it can be both instructive and computationally useful.
-We can rewrite \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$ and substitute into \ref{eq:ag_original}.
+By rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$ and substituting into \ref{eq:ag_original}.
 \begin{subequations} \label{eq:ag_big_gamma}
@@
 \end{subequations}
-These four equations now contain all of the information previously contained in six, making this much less computationally intensive.
+these four equations now contain all of the information previously contained in six.
+Having fewer equations, and fewer mathematical operations, makes this set of equations much less computationally intensive.
 Further, now one can see that these $\Gamma$ constants can be thought of as some minimum/intrinsic action of the pulse. 

### 2012-05-03 inc/ms_model/ms_model.tex
nearly completed the inclusion of M&S AG derivation

@@
 \end{equation}
-Note that the original paper of Michalik and Sipe contained an error which incorrectly defined $\xi$; this was corrected in an erratum\cite{michalik_erratum:_2008}.
+Note that the original paper of Michalik and Sipe contained an error which incorrectly defined $\xi$; this was corrected in an erratum \cite{michalik_erratum:_2008}.
 The family of $L$ functions are
@@
 \end{gather}
+The integral function $L(\xi)$ can be evaluated to the piecewise function
+\begin{equation}
+  L(\xi) = 
+  \begin{cases}
+    \frac{ \arcsin \sqrt{1-\xi^2} }{ \sqrt{1-\xi^2} } & 0 \le \xi \le 1 \\
+    \frac{ \ln\left( \xi + \sqrt{\xi^2 - 1} \right) }{ \sqrt{\xi^2 - 1} } & \xi \ge 1
+  \end{cases}
+\end{equation}
+which, although piecewise, is a smooth and well behaved function.
 
@@
 \end{gather}
-\end{subequations}
-which depends on the ellipticity parameter
-\begin{equation}
-  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}}
-\end{equation}
 

### 2012-05-03 inc/ms_model/ms_model.tex
fixed typo added internal linkage between Kflow and ExtForces
Most of the important bits of the original AG model is now included.

@@
 \end{equation}
-Taking the time derivative of the components of $A^{-1}$
+
+Since time derivative of \ref{eq:gaussian_integral} is simply
+\begin{equation} \label{eq:gaussian_integral_dt}
+  \frac{d}{dt}(A^{-1})_{ij} = \frac{1}{N} \int u_i u_j \frac{df(\vec{r}, \vec{p}; t)}{dt} d\vec{r} d\vec{p}
+\end{equation}
+and the time derivative of the components of $A^{-1}$ are
 \begin{equation} \label{eq:dainvdt}
@@
 \end{equation}
-Since time derivative of \ref{eq:gaussian_integral} is simply
-\begin{equation} \label{eq:gaussian_integral_dt}
-  \frac{d}{dt}(A^{-1})_{ij} = \frac{1}{N} \int u_i u_j \frac{df(\vec{r}, \vec{p}; t)}{dt} d\vec{r} d\vec{p}
-\end{equation}
 these components, when viewed in the context of \ref{eq:formal_evolution}, form the left hand side (LHS) of the system of differential equations governing the evolution of $\sigma_{\alpha}$, $\gamma_{\alpha}$, and $\eta_{\alpha}$, which will be established in the upcoming sections.
@@
 These terms will be denoted $K^{flow}$ and $K^{int}$ respectively.
-\begin{equation} \label{eq:gaussian_integral_dt}
+\begin{equation} \label{eq:gaussian_integral_dt_k}
   \frac{d}{dt}(A^{-1})_{ij} = K^{flow}_{ij} + K^{int}_{ij}
@@
 \end{equation}
-These analyses will be referred to while deriving the influence of external forces.
-%TODO internal reference
+These analyses will be referred to while deriving the influence of external forces (Section \ref{sec:external_forces}).
 
@@
 
-The results of the previous analyses culminate in the differential equation set
+When these $K$ matrices are summed (\ref{eq:gaussian_integral_dt_k}) to form the right hand side (RHS) of the differential equations and compared to LHS derived above (\ref{eq:dainvdt}), the non-zero elements finally yield a system of differential equations which govern the dynamics of these Gaussian beam parameters.
+
 \begin{subequations} \label{eq:ag_original}
@@
 \end{gather}
+\end{subequations}
 

### 2012-05-03 inc/external_forces/external_forces.tex
fixed typo added internal linkage between Kflow and ExtForces
Improved generalized extension for external forces

@@
 
-\section{Extension to Include External Forces}
+\subsection{Extension to Include External Forces} \label{sec:external_forces}
 
-General form of (possibly multiple) external force ($\vec{F}_{ext}$) contributions
+\ref{eq:potential_integral} is the entry point for the internal Coulombing repulsion force; since both potentials and integrals are linear, one may add additional potentials so that it becomes
+\begin{equation}
+  \begin{split}
+    \Phi(\vec{r};t) \rightarrow & \int d\vecprime{r} \left( V(\vec{r} - \vecprime{r}) + \sum V_{ext}(\vec{r}) \right) n(\vecprime{r}; t) \\
+    = & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) + \sum \int d\vecprime{r} V_{ext}(\vec{r}) n(\vecprime{r}; t) \\
+    = & \Phi(\vec{r};t) + \Phi_{ext}(\vec{r};t)
+  \end{split}
+\end{equation}
+This form of $\Phi(\vec{r};t)$ can be substituted into the definition of $K^{force}$ (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector} 
+\begin{equation}
+  \begin{split}
+    K^{force}_{ij} \rightarrow & \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial}{\partial \vec{r}} \left( \Phi(\vec{r};t) + \Phi_{ext}(\vec{r};t) \right) d\vec{r} d\vec{p} \\
+    \equiv & K^{int}_{ij} + \sum K^{ext}_{ij}
+  \end{split}
+\end{equation}
+where $K^{int}$ is simply the old $K^{force}$. Noting that the spatial derivative of a potential is just the force, we can now define the matrix contributed by each external force is given by
+\begin{equation}
+  K^{ext}_{ij} = \frac{1}{N} \iint u_i u_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
+\end{equation}
+Finally, the general form of \ref{eq:gaussian_integral_dt_k} for (possibly multiple) external force ($\vec{F}_{ext}$) contributions is now
 \begin{equation}
@@
 \end{equation}
-again in the basis defined in \ref{eq:coordinate_vector} is given by
-\begin{equation}
-  K^{ext}_{ij} = \frac{1}{N} \iint u_i u_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
-\end{equation}
 
-For example, if $\vec{F}$ is a static field given by $\frac{qV}{d}\hat{z}$ then $K^{ext} = \hat{0}$.
-However, if the external force is a lensing field given by $\vec{F} = -M\cdot(x\hat{x}+y\hat{y})$ then
+\subsection{Specific Forms of External Forces}
+If $\vec{F}$ is a static field given by $\frac{qV}{d}\hat{z}$ then $K^{ext} = \hat{0}$. However, if the external force is a lensing field given by $\vec{F} = -M\cdot(x\hat{x}+y\hat{y})$ then 
 \begin{equation}

### 2012-05-21 inc/external_forces/external_forces.tex
fixed typo

@@
 
-\ref{eq:potential_integral} is the entry point for the internal Coulombing repulsion force; since both potentials and integrals are linear, one may add additional potentials so that it becomes
+\ref{eq:potential_integral} is the entry point for the internal Coulombic repulsion force; since both potentials and integrals are linear, one may add additional potentials so that it becomes
 \begin{equation}

### 2012-05-21 inc/initial_conditions/initial_conditions.tex
added CC notice to initial_conditions
added the simplisitc intial conditions from AG model paper

@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \subsection{Determining Appropriate Initial Conditions} \label{sec:initial_conditions}
+For a DTEM or UED apparatus, where the electron pulse is produced in a laser-driven photo-electron gun, the implementation of the AG model of Michalik and Sipe \cite{michalik_analytic_2006} requires knowledge of the initial ($ t = 0 $) bunch parameters immediately after generation at the photocathode surface inside the acceleration region of the gun.
+For the purposes of this paper, we will make a simple approximation to the behavior of the gun in order to provide generic initial conditions for the electron pulse.
+
+For single-photon photoemission process with a spatially cylindrically symmetric incident Gaussian laser beam with Gaussian temporal irradiance profile 
+\begin{equation}\label{eq:laser_form}
+I ( x , y , t ) = I_{0} \exp \left ( - \frac{ ( x^{2} +y^{2} ) }{ w^{2} } - \frac{ t^{ 2 } }{ \tau^{ 2 } } \right ) \text{ ,}
+\end{equation}
+where $w$ is the half-width 1/e maximum (HW1/eM) of the spatial irradiance distribution, $\tau$ is the HW1/e pulse duration and $I_{0}$ is the peak pulse irradiance, the electron pulse's initial spatial variance is
+\begin{equation}
+\sigma_{\smallT } ( 0 ) = ( \Delta x )^{2} = ( \Delta y)^{2} = \frac{ w^{2} }{ 2 } \text{ .}
+\end{equation}
+For the longitudinal direction, and in the limit where dispersion effects in the acceleration region of the electron gun can be neglected, we may write
+\begin{equation} \label{eq:initial_sigma_z}
+\sigma_{ z } ( 0 ) = ( \Delta z)^{2} = \frac{ ( v_{{ \scriptscriptstyle 0}} \tau )^{2} }{ 2 } \text{ ,}
+\end{equation}
+with the velocity of the pulse in the lab frame $v_{{ \scriptscriptstyle 0}} = \sqrt{ 2 e V / m_{e} d } $, where $ V $ and $ d $ are the potential and cathode-anode separation of the gun's DC acceleration region respectively.
+For spatially-uniform photoemission from a planar photocathode, we have $ \gamma_{\smallT} ( 0 ) = 0 $; i.e., there is no initial spatial momentum chirp in the electron bunch.
+Immediately after photo-generation, however, it is strictly inaccurate to say that $ \gamma_{ z } ( 0 ) = 0 $.
+This is because the excess photoemission energy $ \Delta E = \hbar \omega - \Phi $, associated with the electron emission from a material with work function $\Phi$ using photons of energy $ \hbar \omega $ (which gives the maximum initial velocity of an electron, $ v_{max} = \sqrt{ 2 \Delta E/m_{e} } $), can produce an initial momentum chirp.
+Nonetheless, the initial non-zero nature of $ \gamma_{z} $ is quickly dominated by the evolution of the pulse, so this approximation will suffice in the region far from the gun.
+
+The single-photon photoemission analysis of Jensen \textit{et al.}\cite{jensen_theoretical_2006} indicates that for a metallic photocathode the transverse momentum variance at $ t = 0 $ is given by
+\begin{equation} \label{eq:jensen_eta}
+( \Delta p_{x} )^{2} = ( \Delta p_{y} )^{2} = \eta_{ \smallT } ( 0 ) \approx \frac{ m_{e} \epsilon_{\smallF} ( \hbar \omega - \Phi ) } { \hbar \omega + \epsilon_{\smallF} } \; \text{,}
+\end{equation}
+in the limit where the Fermi energy $ \epsilon_{\smallF} $ of the metal photocathode and excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ are significantly greater that the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
+From a simplistic viewpoint, only the electrons generated with positive longitudinal momentum will contribute to the pulse since those electrons generated with negative momentum do not leave the photocathode.
+This implies that the uncertainty in the longitudinal direction is half that of the transverse directions; that is to say
+\begin{equation}
+( \Delta p_{ z } )^{2} = \left ( \frac{\Delta p_{ x , y }}{ 2 } \right )^{2} \text{,}
+\end{equation}
+so that
+\begin{equation}
+\eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 )}{ 4 } \; \text{.}
+\end{equation}
+
+These initial conditions at $t=0$ for an incident Gaussian laser pulse of the form given in \ref{eq:laser_form}, will be used for the remainder of this paper.
+\begin{subequations} \label{eq:summary}
+\begin{gather}
+\sigma_{ \smallT } ( 0 ) = \frac{ w^{ 2 } }{ 2 } \\
+\sigma_{ z } ( 0 ) = \frac{ ( v_{{ \scriptscriptstyle 0}} \tau )^{ 2 } }{ 2 } \\
+%alternative formulation + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
+\gamma_{ \smallT } ( 0 ) = 0 \\
+\gamma_{ z } ( 0 ) = 0 \\
+\eta_{ \smallT } ( 0 ) = \frac{ m_{e} \epsilon_{ F } ( \hbar \omega - \Phi ) }{ 3 ( \hbar \omega + \epsilon_{ \smallF } ) } \\
+\eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 ) }{ 4 }
+\end{gather}
+\end{subequations}
+
+We recognize that these initial conditions for an electron pulse generated by a laser-driven DC gun are idealized in that they neglect several important effects; in particular, the interplay of dispersive ($ \eta_{i}(0) \neq 0 $) and space-charge effects in the acceleration region of the photoelectron gun and transverse and longitudinal effects due to the divergence of the DC field in the vicinity of the gun anode.\cite{berger_dc_2009,togawa_ceb6_2007}
+Such effects and the limits of the validity of \ref{eq:initial_sigma_z}, within the context of the AG model, will be presented elsewhere as they require \textit{a priori} this presented extension of the AG model.
+However, \ref{eq:summary} also represent a defined set of initial conditions that allows the effect of electron optics (i.e., magnetic lenses and RF cavities) on electron pulse propagation to be more readily understood.
+Moreover, they reflect the fact that the majority of DC photoelectron guns are driven by laser pulses with spatial and temporal profiles that are close to Gaussian.\cite{williamson_clocking_1997,sciaini_electronic_2009}
+Only for specialized cases, such as the generation of uniform ellipsoid electron bunch shapes, is spatial and temporal manipulation of the incident laser pulse considered.\cite{luiten_how_2004,li_generating_2008}

### 2012-05-21 inc/liouvilles_theorem/liouvilles_theorem.tex
Move Liouville's theorem implications to extension (ie my contributions) section

--- 2012-05-03 inc/ms_model/ms_model.tex
@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
-
-\section{Model of Michalik and Sipe}
-
-This section introduces the Analytic Gaussian model of Michalik and Sipe \cite{michalik_analytic_2006}.
-
-\subsection{Formal Evolution} \label{sec:formal_evolution}
-
-Before one may describe the evolution of a pulse of electrons, one must first learn to describe the pulse.
-Rather than record the position of each electron, commonly one writes a distribution function for the electrons.
-A distribution function describes the location of a number of objects in a certain space.
-This space may be a 1D line, a 3D space, or even a 6D phase space (three spatial dimensions and three momentum dimensions).
-The functional form may be chosen to appropriately characterize the objects in a natural way, however one property is true for all distribution functions: when integrated over all relevant space, all objects should be accounted for.
-In $n$ dimensional space, with $N$ objects, this normalization condition on a distribution function $f$ can be written mathematically as
-\begin{equation}
-N = \int\limits_{\text{all space}} \!\!\! f(\vec{x}) \; d^{n}x
-\end{equation}
-which for 6D phase space is written as
-\begin{equation} \label{eq:normalization}
-N = \int f (\vec{r}, \vec{p}; t) \; d^{3}x\,d^{3}p
-\end{equation}
-This notation suggests to the reader that the distribution function is implicitly a function of time as well as of the spatial and momentum dimensions.
-
-For a classical set of $N$ particles (labelled by the index $i$), distributed in 6D space (3 position, 3 momentum) by 
-\begin{equation}
-  f_{\smallD}(\vec{r}, \vec{p}; t) = 
-  \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p})
-\end{equation}
-which automatically satisfies \ref{eq:normalization}.
-The particles have mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
-\begin{subequations}\label{eq:std_motion}
-\begin{gather}
-  \frac{d\vec{r}_{i}(t)}{dt} = \frac{\vec{p}_{i}(t)}{m}\\
-  \frac{d\vec{p}_{i}(t)}{dt} = -\sum\limits_{j} \frac{d}{d\vec{r}_{i}} V(\vec{r}_{i} - \vec{r}_{j})
-\end{gather}
-\end{subequations}
-Now consider the time evolution of $f_{\smallD}$
-\begin{equation}
-  \frac{\partial f_{\smallD}}{\partial t} = 
-  \frac{d\vec{r}}{dt} \frac{\partial f_{\smallD}}{\partial \vec{r}} 
-  + \frac{\partial f_{\smallD}}{\partial \vec{p}} \frac{d\vec{p}}{dt}
-\end{equation}
-but with \ref{eq:std_motion} the final discreet evolution equation is
-\begin{equation}
-  \frac{d}{dt} f_{\smallD}(\vec{r}, \vec{p}; t) =
-  -\frac{\vec{p}}{m} \frac{\partial}{\partial \vec{r}} f_{\smallD}(\vec{r}, \vec{p}; t)
-  + \frac{\partial}{\partial \vec{p}} f_{\smallD} (\vec{r}, \vec{p}; t)
-  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f_{\smallD}(\vecprime{r},\vecprime{p};t)
-\end{equation}
-where the extra minus sign comes from the change of perspective of \ref{eq:std_motion} from the frame of the individual particle $i$ to the observer frame.
-%TODO ensemble av and mean-field
-Final equation:
-\begin{equation} \label{eq:formal_evolution}
-  \frac{d}{dt} f(\vec{r}, \vec{p}; t) =
-  -\frac{\vec{p}}{m} \frac{\partial}{\partial \vec{r}} f(\vec{r}, \vec{p}; t)
-  + \frac{\partial}{\partial \vec{p}} f (\vec{r}, \vec{p}; t)
-  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f(\vecprime{r},\vecprime{p};t)
-\end{equation}
-
-\subsection{Evolution of a Gaussian Distribution}
-
-The formal evolution explored in \ref{sec:formal_evolution} allows for any distribution function.
-For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian, mimicing the laser pulse which generated it.
-How correct this assumption is will be explored in \ref{sec:initial_conditions}.
-In this section we consider a Gaussian distribution function, generally of the form
-\begin{equation}
-  f(\vec{r}, \vec{p}; t) = C(t)\exp \left[ - \Gamma(\vec{r}, \vec{p}; t) \right]
-\end{equation}
-The analysis presented here will consider a cylindrically symmetric pulse.
-For the remainder of this analysis, the subscripts $\smallT$ (transverse) and $z$ (longitudinal) will accompany many quantites, reflecting the congruence of $x$ and $y$.
-The subscript $\alpha$ will refer to either $\smallT$ or $z$.
-In this case we choose the argument (i.e. the function $\Gamma$) as
-\begin{equation} \label{eq:gaussian_exponent}
-  \Gamma(\vec{r}, \vec{p}; t) =
-  \frac{x^2 + y^2}{2 \sigma_{\smallT}} + \frac{z^2}{2 \sigma_{z}}
-  + \frac{
-    [p_x - (\gamma_{\smallT}/\sigma_{\smallT}) x ]^2 
-    + [p_y - (\gamma_{\smallT}/\sigma_{\smallT}) y ]^2
-  }{2 \eta_{\smallT}}
-  + \frac{ [p_z - (\gamma_{z}/\sigma_{z}) z ]^2 }{2 \eta_{z}}
-\end{equation}
-This parameterization defines the spatial uncertainty $\sigma_{\alpha}$, the local momentum uncertainty $\eta_{\alpha}$, and the momentum chirp $\gamma_{\alpha}$ in each spatial direction.
-The ``chirp'' of a pulse quantifies the difference in local momentum from one side of the pulse to the other.
-For example, in a pulse which is normally (positively) chirped longitudinally, electrons near the front of the pulse will be moving faster (on average) than those in the back.
-
-The normalization constant $C(t)$, for $N$ total particles, is found by applying equation \ref{eq:normalization}
-\begin{equation}
-  C(t) = \frac{N}{(2\pi)^3} 
-  \left( 
-    \frac{1}{\sigma_{\smallT}^2\eta_{\smallT}^2\sigma_{z}\eta_{z}}
-  \right)^{\frac{1}{2}}
-\end{equation}
-These equations may be viewed from a matrix perspective if we define the ``coordinate vector''
-\begin{equation} \label{eq:coordinate_vector}
-u_i = \{x, p_x, y, p_y, z, p_z\}\text{ .}
-\end{equation}
-In this space, \ref{eq:gaussian_exponent} becomes
-\begin{equation}
-  \Gamma(\vec{r}, \vec{p}; t) = \frac{1}{2}\sum\limits_{ij} A_{ij} u_i u_j
-\end{equation}
-where
-\begin{equation}
-  A = 
-  \begin{pmatrix}
-    a_{\smallT} & 0 & 0 \\
-    0 & a_{\smallT} & 0 \\
-    0 & 0 & a_{z}
-  \end{pmatrix}
-  \qquad \text{and} \qquad
-  a_{\alpha} = 
-  \begin{pmatrix}
-    1/\sigma_{\alpha} + \gamma_{\alpha}^2/(\eta_{\alpha} \sigma_{\alpha}) & \gamma_{\alpha}/(\eta_{\alpha} \sigma_{\alpha}) \\
-    \gamma_{\alpha}/(\eta_{\alpha} \sigma_{\alpha}) & 1/\eta_{\alpha}
-  \end{pmatrix}
-\end{equation}
-with ${\scriptstyle \alpha} = \smallT \text{ or } z$. 
-A nice property of Gaussians of this form is the convenient ``Multidimensional Gaussian Integral'' given by
-\begin{equation} \label{eq:gaussian_integral}
-  (A^{-1})_{ij} = \frac{1}{N} \int u_i u_j f(\vec{r}, \vec{p}; t) d\vec{r} d\vec{p}
-\end{equation}
-which is a common form of integral analysis in Quantum Field Theory (see for example Peskin and Schroeder Equation 9.70). 
-Application of this integral yields
-\begin{equation}
-  A = 
-  \begin{pmatrix}
-    a_{\smallT}^{-1} & 0 & 0 \\
-    0 & a_{\smallT}^{-1} & 0 \\
-    0 & 0 & a_{z}^{-1}
-  \end{pmatrix}
-  \qquad \text{and} \qquad
-  a^{-1}_{\alpha} = 
-  \begin{pmatrix}
-    \sigma_{\alpha} & \gamma_{\alpha} \\
-    \gamma_{\alpha} & \eta_{\alpha} + \gamma_{\alpha}^2/\sigma_{\alpha}
-  \end{pmatrix}
-\end{equation}
-Taking the time derivative of the components of $A^{-1}$
-\begin{equation} \label{eq:dainvdt}
-  \frac{d}{dt} a^{-1}_{\alpha} = 
-  \begin{pmatrix}
-    \dfrac{d\sigma_{\alpha}}{dt} & \dfrac{d\gamma_{\alpha}}{dt} \\
-    \dfrac{d\gamma_{\alpha}}{dt} & \dfrac{d\eta_{\alpha}}{dt} + 2\dfrac{\gamma_{\alpha}}{\sigma_{\alpha}}\dfrac{d\gamma_{\alpha}}{dt}- \dfrac{\gamma^{2}_{\alpha}}{\sigma^{2}_{\alpha}}\dfrac{d\sigma_{\alpha}}{dt}
-  \end{pmatrix}
-\end{equation}
-Since time derivative of \ref{eq:gaussian_integral} is simply
-\begin{equation} \label{eq:gaussian_integral_dt}
-  \frac{d}{dt}(A^{-1})_{ij} = \frac{1}{N} \int u_i u_j \frac{df(\vec{r}, \vec{p}; t)}{dt} d\vec{r} d\vec{p}
-\end{equation}
-these components, when viewed in the context of \ref{eq:formal_evolution}, form the left hand side (LHS) of the system of differential equations governing the evolution of $\sigma_{\alpha}$, $\gamma_{\alpha}$, and $\eta_{\alpha}$, which will be established in the upcoming sections.
-
-\subsection{Generating the Differential Equations}
-
-Inserting \ref{eq:formal_evolution} into \ref{eq:gaussian_integral_dt} yields two integral terms, one related to the propagation of a non-charged Gaussian beam, the other adds the influence of the internal force due to Coulomb repulsion.
-These terms will be denoted $K^{flow}$ and $K^{int}$ respectively.
-\begin{equation} \label{eq:gaussian_integral_dt}
-  \frac{d}{dt}(A^{-1})_{ij} = K^{flow}_{ij} + K^{int}_{ij}
-\end{equation}
-Collecting and expanding terms where possible gives these forms of the $K$ terms
-\begin{gather}
-  K^{flow}_{ij} \equiv -\frac{1}{N} \int u_i u_j \frac{\vec{p}}{m} \frac{\partial f}{\partial \vec{r}} d\vec{r} d\vec{p} \\
-  K^{int}_{ij} \equiv \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial \Phi(\vec{r};t)}{\partial \vec{r}} d\vec{r} d\vec{p} \\
-  n(\vec{r};t) \equiv \int d\vec{p} f(\vec{r},\vec{p};t) 
-    = \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right) \\
-  \begin{split}
-    \Phi(\vec{r};t) \equiv & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) \\
-    = & \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \int d\vecprime{r} V(\vec{r} - \vecprime{r}) \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right)
-  \end{split}
-\end{gather}
-Obviously the $K^{flow}$ term is far easier to evaluate
-\begin{equation}
-  K^{flow} = 
-  \begin{pmatrix}
-    k^{flow}_{\smallT} & 0 & 0 \\
-    0 & k^{flow}_{\smallT} & 0 \\
-    0 & 0 & k^{flow}_{z}
-  \end{pmatrix}
-  \quad \text{and} \quad
-  k^{flow}_{\alpha} = 
-  \begin{pmatrix}
-    \dfrac{2\sigma_{\alpha}}{m} & \dfrac{1}{m} \left(\eta_{\alpha} + \dfrac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) \\
-    \dfrac{1}{m} \left(\eta_{\alpha} + \dfrac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) & 0
-  \end{pmatrix}
-\end{equation}
-These analyses will be referred to while deriving the influence of external forces.
-%TODO internal reference
-
-The $K^{force}$ integrals, related to the internal force of the pulse's Coulombic repulsion, are far more complicated to evaluate.
-For more information about the derivation, see \cite{michalik_analytic_2006}.
-The final matrices are
-\begin{equation}
-  K^{force} = 
-  \begin{pmatrix}
-    k^{force}_{\smallT} & 0 & 0 \\
-    0 & k^{force}_{\smallT} & 0 \\
-    0 & 0 & k^{force}_{z}
-  \end{pmatrix}
-\end{equation}
-where
-\begin{equation}
-  k^{force}_{\alpha} = 
-  \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{2\pi^{1/2}\sigma_{\alpha}^{1/2}} L_{\alpha}(\xi)
-  \begin{pmatrix}
-    0 & 1 \\
-    1 & \dfrac{2 \gamma_{\alpha}}{\sigma_{\alpha}}
-  \end{pmatrix}
-\end{equation}
-The parameter $\xi$ is the ellipticity parameter, defined as
-\begin{equation}
-  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}}
-\end{equation}
-Note that the original paper of Michalik and Sipe contained an error which incorrectly defined $\xi$; this was corrected in an erratum\cite{michalik_erratum:_2008}.
-The family of $L$ functions are
-\begin{gather}
-  L_{\smallT}(\xi) = \frac{3}{2} \left( L(\xi) + \frac{\xi^2 L(\xi) - \xi}{ 1 - \xi^2 } \right) \\
-  L_{z}(\xi) = \frac{ 3 \xi^2 }{ \xi^2 - 1} \left( \xi L(\xi) - 1 \right) \\
-  L(\xi) = \frac{1}{2} \int_0^{\pi} \frac{1}{1+\xi\sin(\theta)} d\theta
-\end{gather}
-
-\subsection{The Differential Equation System}
-
-The results of the previous analyses culminate in the differential equation set
-\begin{subequations} \label{eq:ag_original}
-\begin{gather}
-  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
-  \frac{d\gamma_{\alpha}}{dt} = \frac{1}{m} \left(\eta_{\alpha} + \frac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) 
-    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi)\\
-  \frac{d\eta_{\alpha}}{dt} = - \frac{2 \gamma_{\alpha} \eta_{\alpha}}{m \sigma_{\alpha}}
-\end{gather}
-\end{subequations}
-which depends on the ellipticity parameter
-\begin{equation}
-  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}}
-\end{equation}
 

### 2012-05-21 inc/ms_model/ms_model.tex
Improved generalized extension for external forces
Move Liouville's theorem implications to extension (ie my contributions) section

@@
 Inserting \ref{eq:formal_evolution} into \ref{eq:gaussian_integral_dt} yields two integral terms, one related to the propagation of a non-charged Gaussian beam, the other adds the influence of the internal force due to Coulomb repulsion.
-These terms will be denoted $K^{flow}$ and $K^{int}$ respectively.
+These terms will be denoted $K^{flow}$ and $K^{force}$ respectively.
 \begin{equation} \label{eq:gaussian_integral_dt_k}
@@
   K^{flow}_{ij} \equiv -\frac{1}{N} \int u_i u_j \frac{\vec{p}}{m} \frac{\partial f}{\partial \vec{r}} d\vec{r} d\vec{p} \\
-  K^{int}_{ij} \equiv \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial \Phi(\vec{r};t)}{\partial \vec{r}} d\vec{r} d\vec{p} \\
+  \label{eq:Kforce}
+  K^{force}_{ij} \equiv \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial \Phi(\vec{r};t)}{\partial \vec{r}} d\vec{r} d\vec{p} \\
   n(\vec{r};t) \equiv \int d\vec{p} f(\vec{r},\vec{p};t) 
     = \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right) \\
-  \begin{split}
-    \Phi(\vec{r};t) \equiv & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) \\
-    = & \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \int d\vecprime{r} V(\vec{r} - \vecprime{r}) \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right)
-  \end{split}
+    \label{eq:potential_integral}
+    \Phi(\vec{r};t) \equiv \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t)
 \end{gather}
@@
 \end{equation}
-These analyses will be referred to while deriving the influence of external forces (Section \ref{sec:external_forces}).
 
@@
 \end{subequations}
-
-\subsection{Implications of Liouville's Theorem}
-
-One of the most well known theorems in physics (and mathematics) is Liouville's Theorem.
-This theorem can be summarized succinctly as a `the total phase space of a system is conserved.'
-%TODO find a good reference for this
-
-To evaluate the effect of this theorem on these equations, consider the time derivative of the product $\sigma_{\alpha} \eta_{\alpha}$.
-Using \ref{eq:ag_original} it is easy to show that
-\begin{gather}
-  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0\\
-  \label{eq:define_big_gamma}
-  \therefore \sigma_{\alpha} \eta_{\alpha} \equiv \Gamma_{\alpha}^2 = \text{Constant}
-\end{gather}
-This analysis shows that the Analytic Gaussian model explicitly obeys the Liouville Theorem.
-
-Not only is this an interesting result, but it can be both instructive and computationally useful.
-By rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$ and substituting into \ref{eq:ag_original}.
-\begin{subequations} \label{eq:ag_big_gamma}
-\begin{gather}
-  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
-  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
-    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi)
-\end{gather}
-\end{subequations}
-these four equations now contain all of the information previously contained in six.
-Having fewer equations, and fewer mathematical operations, makes this set of equations much less computationally intensive.
-Further, now one can see that these $\Gamma$ constants can be thought of as some minimum/intrinsic action of the pulse. 
-This will become more apparent when in future sections
-%TODO include internal reference
-it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.

### 2012-05-21 inc/external_forces/external_forces.tex
insert commentary from JAP paper to relevant places

@@
 
+\subsubsection{Validity and Limitations}
+
+The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations.\cite{michalik_analytic_2006}
+As a result, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
+Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid.\cite{luiten_how_2004}
+This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
+Consequently, the AG approach is applicable to both the single electron per pulse limit,\cite{lobastov_four-dimensional_2005} where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate.\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}
+It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
+Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
+
+\subsubsection{Mathematical Formulation of External Force Contribution}
 \ref{eq:potential_integral} is the entry point for the internal Coulombic repulsion force; since both potentials and integrals are linear, one may add additional potentials so that it becomes
@@
 
-\subsection{Specific Forms of External Forces}
+\subsubsection{Specific Forms of External Forces}
 If $\vec{F}$ is a static field given by $\frac{qV}{d}\hat{z}$ then $K^{ext} = \hat{0}$. However, if the external force is a lensing field given by $\vec{F} = -M\cdot(x\hat{x}+y\hat{y})$ then 

### 2012-05-21 inc/ms_model/ms_model.tex
insert commentary from JAP paper to relevant places

@@
 \end{subequations}
+
+\subsection{Validity and Limitations}
+
+This AG formalism of Michalik and Sipe\cite{michalik_analytic_2006} employs a mean-field approximation in the evolution of the dynamical effect of the electron pulse's internal space-charge field.
+This is of course not strictly accurate for a Gaussian charge distribution and the error leads to a distortion of the pulse shape.
+Even so, the self-similar AG model of charge bunch dynamics has been benchmarked against particle tracking simulations for a number of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid which explicitly features a linear internal space-charge field.\cite{luiten_how_2004}

### 2012-05-21 inc/external_forces/external_forces.tex
Comment on relativistic contributions

@@
 Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
+
+Further, these analyses are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
+Extension to include relativistic effects is conceptually straightforward using standard transformations between the laboratory and electron pulse reference frames.
 

### 2012-05-21 inc/external_forces/external_forces.tex
Constant and linear external forces

@@
 where $K^{int}$ is simply the old $K^{force}$. Noting that the spatial derivative of a potential is just the force, we can now define the matrix contributed by each external force is given by
-\begin{equation}
+\begin{equation} \label{eq:Kint}
   K^{ext}_{ij} = \frac{1}{N} \iint u_i u_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
@@
 
-\subsubsection{Specific Forms of External Forces}
-If $\vec{F}$ is a static field given by $\frac{qV}{d}\hat{z}$ then $K^{ext} = \hat{0}$. However, if the external force is a lensing field given by $\vec{F} = -M\cdot(x\hat{x}+y\hat{y})$ then 
+\subsubsection{Constant External Forces}
+
+For constant static forces having no dependence in $ x $, $ y $, $ z $, or $ t $, one can readily show that there is no effect on the pulse propagation parameters $ \sigma_{ \alpha } $, $ \eta_{ \alpha } $, and $ \gamma_{ \alpha } $.
+For example, consider a field consider a field $ \vec{E} = -E \hat{z} $ such that electrons are accelerated in the positive $ z $ direction by $ \vec{F}_{ext} = qE \hat{z} = q ( V / d ) \hat{z} $, so that from \ref{eq:Kint} we have
 \begin{equation}
-  K^{ext} = M \cdot 
-  \begin{pmatrix}
-    K^{lens}_{\smallT} & 0 & 0 \\
-    0 & K^{lens}_{\smallT} & 0 \\
-    0 & 0 & 0
-  \end{pmatrix}
-  \qquad \text{where} \qquad
+ \begin{split} 
+  K^{ext}_{ij} = \frac{1} {N} \dfrac{qV} {d} \int{ u_{i} u_{j} \left( \frac{p_{z}} {\eta_{z}} - \frac{\gamma_{z} z} {\sigma_{z} \eta_{z}} \right) f d^{3}x   d^{3}p } ;
+ \end{split}
+\end{equation}
+but it is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ (or equivalently $ K^{ext} = \hat{0} $) meaning that this static field has no effect on the parameters of the pulse.
+Of course, this is only true in the pulse frame, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m_{e} $ is unchanged.
+However, in the lab frame, where  $ \Delta E = p \Delta p / m_{e} $, there is an increase in the momentum spread, as the pulse now has the large momentum $p$ imparted by the accelerating field.
+
+\subsubsection{Linear External Forces}
+
+For the case of a perfect lens, or in the longitudinal direction a perfect RF compression element, with a parabolic potential, the external force is linear with distance from the center of the electron pulse:
+\begin{equation} \label{eq:linear_force}
+\vec{F}_{ext} = - M_{ \smallT } r \hat{\operatorname{r}} - M_{z} z \hat{\operatorname{z}} \text{ .}
+\end{equation}
+Here, $M_{ \smallT }$ and $M_{z}$ characterize the strength of a lens or a compressor, respectively, and are chosen to be positive for pulse focusing.
+Substitution into the integral equation for the $ K^{ext} $ matrix elements,
+\begin{equation}
+ \begin{split} 
+  K^{ext}_{ij} & \equiv \dfrac{1} {N} \int u_{i} u_{j} ( \vec{\nabla}_{p} f ) \cdot \vec{F}_{ext} d^{3} x d^{3} p \\ & = \dfrac{1} {N} \int u_{i} u_{j} \biggl[ x \left( \dfrac{ p_{x} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } x } { \sigma_{ \smallT } \eta_{ \smallT } } \right) + y \left( \dfrac{ p_{y} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } y } { \sigma_{ \smallT } \eta_{ \smallT } } \right) \\ & \quad + z \left( \dfrac{ p_{z} } { \eta_{z} } - \dfrac{ \gamma_{z} z } { \sigma_{z} \eta_{z} } \right) \biggr] f d^{3} x d^{3} p ,
+ \end{split}
+\end{equation}
+leads to the $6\times6$ matrix
+\begin{equation}
 K^{ext} = 
 \begin{pmatrix}
-    0 & \sigma_{\smallT} \\
-    \sigma_{\smallT} & 2 \gamma_{\smallT}
-  \end{pmatrix}
+\hat{k}^{lin}_{\smallT} & \hat{0} & \hat{0} \\
+\hat{0} & \hat{k}^{lin}_{\smallT} & \hat{0} \\
+\hat{0} & \hat{0} & \hat{k}^{lin}_{z}
+\end{pmatrix} \text{ ,}
 \end{equation}
-which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\frac{\partial \gamma_{\alpha}}{\partial t}$ equation.
+where
+\begin{equation}
+\hat{k}^{lin}_{\alpha} =
+\begin{pmatrix}
+0 & M_{\alpha} \sigma_{\alpha} \\
+M_{\alpha} \sigma_{\alpha} & 2 M_{\alpha} \gamma_{\alpha} \text{.}
+\end{pmatrix} \text{ .}
+\end{equation}
+which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.

### 2012-06-22 inc/external_forces/external_forces.tex
a little reorg
Split external forces into formal and practical sections

@@
 
-\subsection{Extension to Include External Forces} \label{sec:external_forces}
-
-\subsubsection{Validity and Limitations}
-
-The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations.\cite{michalik_analytic_2006}
-As a result, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
-Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid.\cite{luiten_how_2004}
-This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
-Consequently, the AG approach is applicable to both the single electron per pulse limit,\cite{lobastov_four-dimensional_2005} where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate.\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}
-It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
-Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
-
-Further, these analyses are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
-Extension to include relativistic effects is conceptually straightforward using standard transformations between the laboratory and electron pulse reference frames.
+\subsection{Formal Extension to Include Generic External Forces} \label{sec:external_forces}
 
@@
 
+%External forces that vary with position in the pulse reference frame (i.e., are not constant) act to change the electron pulse propagation dynamics, most often by affecting $\gamma_{\alpha}$ --- the spatial dependence of the local relative mean electron momenta.
+%Below, we consider two examples of electron optical elements that are being used, or are under consideration, for controlling and manipulating the space-time dynamics of electron pulses in DTEM and UED instrumentation; namely, magnetic lenses for spatial focusing and RF pulse compression cavities.
+
 For the case of a perfect lens, or in the longitudinal direction a perfect RF compression element, with a parabolic potential, the external force is linear with distance from the center of the electron pulse:
@@
 which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.
+
+\subsubsection{Validity and Limitations}
+
+The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations.\cite{michalik_analytic_2006}
+As a result, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
+Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid.\cite{luiten_how_2004}
+This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
+Consequently, the AG approach is applicable to both the single electron per pulse limit,\cite{lobastov_four-dimensional_2005} where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate.\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}
+It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
+Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
+
+Further, these analyses are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
+Extension to include relativistic effects is conceptually straightforward using standard transformations between the laboratory and electron pulse reference frames.
+
+\subsection{Practical Extension to Include Magnetic Lenses and RF Cavities}
+
+\subsubsection{Region of Influence}
+
+To specify the physical region of influence of the lenses, it is convenient to redefine $M_{i}$ to include a super-Gaussian envelope in $z^{\prime}$, the position of the peak of the pulse in the lab frame;
+\begin{equation} \label{eq:reg_of_influence}
+  M_{i}	\to M_{i} \exp \left [ - \left (  \frac{ z^{\prime} - z_{lens}^{\prime} }{ L_{lens} / 2 } \right )^{ 2 n } \right ] \text{ ,}
+\end{equation}
+where $z_{lens}^{\prime}$ is the position of the center of the lens, $L_{lens}$ is the length of the lens and $n$ is the super-Gaussian order parameter.
+The order parameter defines the ``sharpness'' of the edge of the region of influence with integer values ranging from 1 (a simple Gaussian) through $\infty$ (a top-hat).
+We connect $z^{\prime}$ to $t$ through $ v_{{ \scriptscriptstyle 0}} $; the kinetic equations are simply defined for the position of the peak of the pulse.
+
+\subsubsection{Magnetic Lenses}
+
+\subsubsection{RF Cavities}
+
+In the longitudinal ($ z $) direction, a resonant radio-frequency (RF) cavity can be used in a UEM column or UED experiment \cite{oudheusden_electron_2007,fill_sub-fs_2006} to act as a pulse compression or `temporal focusing' element.
+Although there are many possible resonant cavity structures,\cite{oudheusden_electron_2007,humphries_principles_1986} for simplicity, this analysis will consider only the evacuated cylindrical (or `pillbox') $\text{TM}_{010}$ cavity for which the electrons propagate down the axis.\cite{fill_sub-fs_2006,humphries_principles_1986}
+The known mathematical forms of the electric and magnetic components of this resonant mode together with a super-Gaussian envelope function (\Eq \eqref{eq:reg_of_influence}) allow the three main effects of a pill-box RF cavity element on the electron pulse to be described within the approximation of the linear force extension to the AG model: the primary action of the pulse compression (or acceleration and expansion dependent upon the relative RF phase), and the transverse lens effects due to the magnetic component of the $\text{TM}_{010}$ resonance and the electric RF cavity fringe fields at its entrance and exit apertures.\cite{kim_rf_1989}
+
+The spatial dependence of the axial RF electric field experienced by the electron pulse in a TM$_{010}$ cavity of radius $a$ may be written as
+\begin{equation} \label{eq:RF_Efield}
+  E ( r , z , z^{\prime} ) = E_{0} \sin \left ( \frac{ \Omega ( z^{\prime} - z ) }{ v_{{ \scriptscriptstyle 0}} } + \phi \right )\operatorname{J_{0}} \left ( \frac{ k r }{ a } \right ) \text{,}
+\end{equation}
+where $ E_{ 0 } $ is the oscillating field amplitude, $\Omega$ is the RF frequency and $\phi$ is a phase factor determining whether the RF cavity acts as a pulse compressor $ ( \phi \approx 0 ) $ or an electron accelerator $ ( \phi \approx \pm \pi/2 ) $.
+The coordinate $z^{\prime} $ describes the position of the center of the pulse in the RF cavity, which extends from $ z^{\prime} = - d / 2 $ to $ z^{\prime} = d / 2 $, and $z$ is the position of any electron in the pulse with respect to the center of the pulse.
+For the $\text{TM}_{010}$ oscillation mode, the RF frequency is given by
+\begin{equation}
+  \Omega = \frac{ c k }{a} \text{,}
+\end{equation}
+where $c$ is the speed of light in a vacuum and $ k = 2.405 $ is the first zero of the $\operatorname{ J_{0} }$ Bessel function.
+
+For the usual case of a spatially compact electron pulse with respect to the dimension of the RF cavity (i.e., $ \Omega z \ll v_{{ \scriptscriptstyle 0}} $ and $ \sqrt{\sigma_{\smallT}} \ll a $), \Eq \eqref{eq:RF_Efield} can be rewritten as
+\begin{equation} \label{eq:RF_Efield_approx}
+  E ( r , z , z^{\prime} ) = E_{0} \biggl [ \sin \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right ) - \frac{ \Omega z }{ v_{{ \scriptscriptstyle 0}} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right ) \biggr ] \text{,}
+\end{equation}
+which demonstrates that the RF cavity acts both as a pulse acceleration (first term on RHS of \Eq \eqref{eq:RF_Efield_approx}) and a pulse compression or expansion (second term on RHS of \eqref{eq:RF_Efield_approx}) device.
+The second term then gives the desired form (see \Eq \eqref{eq:linear_force}) for the force manipulating the temporal dynamics and characteristics during electron pulse compression;
+\begin{equation} \label{eq:RF_Force_z}
+  \boldsymbol{F}_{ z , RF } = - \left [ \frac{ e \Omega E_{0} }{ v_{{ \scriptscriptstyle 0}} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right ) \right ] z \hat{\operatorname{ \boldsymbol{z} }} \text{.}
+\end{equation}
+
+By considering the impulse exerted by the RF cavity field on the electrons in the pulse, it is straightforward to show that pulse compression (when $\phi = 0$) is maximized for $ \Omega d / (2 v_{{ \scriptscriptstyle 0}} ) = \pi /2 $; that is, the time of flight of the electron pulse through the RF cavity of length $ d $ is equal to half the RF period.
+Such a calculation assumes that the pulse velocity $v_{{ \scriptscriptstyle 0}}$ is constant; so that, the condition that $ \phi = 0 $ when the pulse is at the center of the RF cavity (i.e., when the pulse is at $ z^{\prime} = 0 $) is readily met.
+As suggested by the acceleration term in \Eq \eqref{eq:RF_Efield_approx}, this is only a good approximation if the RF field amplitude $E_{0}$ is sufficiently small to not affect the velocity of the peak of the pulse as it propagates through the cavity.
+As a result, in practice, the phase $\phi$ of the RF field is adjusted to optimize the performance of the cavity.
+
+The two main effects contributing to the transverse lens of a $\text{TM}_{010}$-mode RF cavity, its magnetic field component and electric fringe field aperture effects, may also be included within the linear force approximation.
+The azimuthal magnetic component of the resonant mode oscillates 90$^{\circ}$ out of phase with the electric component and with a $\operatorname{ J_{1} }$ Bessel radial form; i.e., $ \operatorname{ J_{1} }( k r / a ) \sim k r / a $ for small $r$.
+Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $,\cite{kim_rf_1989} where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\Eq \eqref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\Eq \eqref{eq:reg_of_influence}).
+In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures, as illustrated by $1 \frac{1}{2}$-cell RF photo-gun designs.\cite{mcdonald_design_1988}
+For the case of an evacuated cavity, these two contributions result in a net force,
+\begin{equation} \label{eq:RF_Force_r}
+  \boldsymbol{F}_{ r , RF } = e E_{0} \left [ 
+    \frac{ v_{{ \scriptscriptstyle 0}} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right )
+   \right ] r \hat{\operatorname{ \boldsymbol{r} }} \text{.}
+\end{equation}
+In obtaining \Eq \eqref{eq:RF_Force_r}, which is also of the form for use in \Eq \eqref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \Eq \eqref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
+For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{{\scriptscriptstyle 0}} / \Omega $,\cite{kim_rf_1989} even though its magnitude is maximized for $\phi = 0$.
+Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.
+
+% excluded example section from JAP paper as it will need new graphics

### 2012-06-22 inc/magnetic_lenses/maglens.tex

@@
 
-To further ease the burden on the lenses, I actually created two, one placed to either side of the RF cavity.
-As each has 100 turns, a current of only 2A is needed to drive the two lenses in series.
-The oxygen-free copper wire coils remain cool to the touch under these operating conditions.
+To manage the divergence created by an RF cavity, two lenses are employed (one before and one after the cavity), forming a lens system of a similar effective focal length.
+For the purposes of design, one can make a reasonable approximation that the two lenses together may together provide the total number of needed amp-turns.
+As each has 100 turns, a current of only about 2A is needed to drive the two lenses in series.
+
+\subsection{Thermal Properties}
+
+%TODO need reference for equation and \alpha_{Cu}
+Since the lenses have no cooling devices, it was prudent to characterize the thermal response to a moderate current load.
+Resistance of metals changes as a function of temperature by $ R = R_0 [ 1 + \alpha ( T - T_0 ) ] $, where $\alpha$ is a material contant; for copper, $\alpha = 3.9 * 10^{-3} \text{K}^{-1}$. 
+One can calculate the temperature increase of the system by monitoring the change in resistance.
+Maintaining a constant current of 4A through the lens, the voltage drop increased from 1.22V ($R_0 = 0.305\Omega$) to 1.25V ($R=0.3125\Omega$) as the lens warmed.
+The resultant temperature increase is only $T - T_0 = 2\text{K}$.
+This result is easily confirmed manually; the oxygen-free copper wire coils remain cool to the touch under these operating conditions.

### 2012-06-27 inc/external_forces/external_forces.tex
fix some errors from undefined macros

@@
 Although there are many possible resonant cavity structures,\cite{oudheusden_electron_2007,humphries_principles_1986} for simplicity, this analysis will consider only the evacuated cylindrical (or `pillbox') $\text{TM}_{010}$ cavity for which the electrons propagate down the axis.\cite{fill_sub-fs_2006,humphries_principles_1986}
-The known mathematical forms of the electric and magnetic components of this resonant mode together with a super-Gaussian envelope function (\Eq \eqref{eq:reg_of_influence}) allow the three main effects of a pill-box RF cavity element on the electron pulse to be described within the approximation of the linear force extension to the AG model: the primary action of the pulse compression (or acceleration and expansion dependent upon the relative RF phase), and the transverse lens effects due to the magnetic component of the $\text{TM}_{010}$ resonance and the electric RF cavity fringe fields at its entrance and exit apertures.\cite{kim_rf_1989}
+The known mathematical forms of the electric and magnetic components of this resonant mode together with a super-Gaussian envelope function (\ref{eq:reg_of_influence}) allow the three main effects of a pill-box RF cavity element on the electron pulse to be described within the approximation of the linear force extension to the AG model: the primary action of the pulse compression (or acceleration and expansion dependent upon the relative RF phase), and the transverse lens effects due to the magnetic component of the $\text{TM}_{010}$ resonance and the electric RF cavity fringe fields at its entrance and exit apertures.\cite{kim_rf_1989}
 
@@
 
-For the usual case of a spatially compact electron pulse with respect to the dimension of the RF cavity (i.e., $ \Omega z \ll v_{{ \scriptscriptstyle 0}} $ and $ \sqrt{\sigma_{\smallT}} \ll a $), \Eq \eqref{eq:RF_Efield} can be rewritten as
+For the usual case of a spatially compact electron pulse with respect to the dimension of the RF cavity (i.e., $ \Omega z \ll v_{{ \scriptscriptstyle 0}} $ and $ \sqrt{\sigma_{\smallT}} \ll a $), \ref{eq:RF_Efield} can be rewritten as
 \begin{equation} \label{eq:RF_Efield_approx}
@@
 \end{equation}
-which demonstrates that the RF cavity acts both as a pulse acceleration (first term on RHS of \Eq \eqref{eq:RF_Efield_approx}) and a pulse compression or expansion (second term on RHS of \eqref{eq:RF_Efield_approx}) device.
-The second term then gives the desired form (see \Eq \eqref{eq:linear_force}) for the force manipulating the temporal dynamics and characteristics during electron pulse compression;
+which demonstrates that the RF cavity acts both as a pulse acceleration (first term on RHS of \ref{eq:RF_Efield_approx}) and a pulse compression or expansion (second term on RHS of \eqref{eq:RF_Efield_approx}) device.
+The second term then gives the desired form (see \ref{eq:linear_force}) for the force manipulating the temporal dynamics and characteristics during electron pulse compression;
 \begin{equation} \label{eq:RF_Force_z}
@@
 Such a calculation assumes that the pulse velocity $v_{{ \scriptscriptstyle 0}}$ is constant; so that, the condition that $ \phi = 0 $ when the pulse is at the center of the RF cavity (i.e., when the pulse is at $ z^{\prime} = 0 $) is readily met.
-As suggested by the acceleration term in \Eq \eqref{eq:RF_Efield_approx}, this is only a good approximation if the RF field amplitude $E_{0}$ is sufficiently small to not affect the velocity of the peak of the pulse as it propagates through the cavity.
+As suggested by the acceleration term in \ref{eq:RF_Efield_approx}, this is only a good approximation if the RF field amplitude $E_{0}$ is sufficiently small to not affect the velocity of the peak of the pulse as it propagates through the cavity.
 As a result, in practice, the phase $\phi$ of the RF field is adjusted to optimize the performance of the cavity.
@@
 The azimuthal magnetic component of the resonant mode oscillates 90$^{\circ}$ out of phase with the electric component and with a $\operatorname{ J_{1} }$ Bessel radial form; i.e., $ \operatorname{ J_{1} }( k r / a ) \sim k r / a $ for small $r$.
-Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $,\cite{kim_rf_1989} where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\Eq \eqref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\Eq \eqref{eq:reg_of_influence}).
+Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $,\cite{kim_rf_1989} where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\ref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\ref{eq:reg_of_influence}).
 In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures, as illustrated by $1 \frac{1}{2}$-cell RF photo-gun designs.\cite{mcdonald_design_1988}
@@
 \end{equation}
-In obtaining \Eq \eqref{eq:RF_Force_r}, which is also of the form for use in \Eq \eqref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \Eq \eqref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
+In obtaining \ref{eq:RF_Force_r}, which is also of the form for use in \ref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \ref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
 For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{{\scriptscriptstyle 0}} / \Omega $,\cite{kim_rf_1989} even though its magnitude is maximized for $\phi = 0$.

### 2012-07-10 inc/external_forces/external_forces.tex
Split AG model chapter into two different chapters, one for previous efforts and one for my contributions.

@@
 
+A magnetic lens may be modeled using one linear force contribution.
+Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is very complicated to describe\cite{montgomery_some_1961} and the ``lensing'' action is actually a compound force (first rotating the beam, then lensing).
+We have found that in practice, characterizing a specific magnetic lens ($M_{\smallT}(I)$ where $I$ is the current in the magnetic lens) is much easier than attempting to find a generic functional form.
+
 \subsubsection{RF Cavities}
@@
 In the longitudinal ($ z $) direction, a resonant radio-frequency (RF) cavity can be used in a UEM column or UED experiment \cite{oudheusden_electron_2007,fill_sub-fs_2006} to act as a pulse compression or `temporal focusing' element.
+The model of the RF cavity is different from the magnetic lens in two distinct ways.
+First, the RF cavity contains several different fields and thus contributes several different linear force terms to the overall equation set.
+Second, though there are more terms, their magnitudes are easily connected to the physical characteristics of the cavity.
+
 Although there are many possible resonant cavity structures,\cite{oudheusden_electron_2007,humphries_principles_1986} for simplicity, this analysis will consider only the evacuated cylindrical (or `pillbox') $\text{TM}_{010}$ cavity for which the electrons propagate down the axis.\cite{fill_sub-fs_2006,humphries_principles_1986}

### 2012-07-10 inc/external_forces/external_forces.tex
A little more reoganization related to splitting the AG model chapter in two. Futher I have started an introduction to the extension chapter

@@
 
-\subsection{Formal Extension to Include Generic External Forces} \label{sec:external_forces}
+\section{Formal Extension to Include Generic External Forces} \label{sec:external_forces}
 
-\subsubsection{Mathematical Formulation of External Force Contribution}
+\subsection{Mathematical Formulation of External Force Contribution}
 \ref{eq:potential_integral} is the entry point for the internal Coulombic repulsion force; since both potentials and integrals are linear, one may add additional potentials so that it becomes
@@
 
-\subsubsection{Constant External Forces}
+\subsection{Constant External Forces}
 
@@
 
-\subsubsection{Linear External Forces}
+\subsection{Linear External Forces}
 
@@
 
-\subsubsection{Validity and Limitations}
+\subsection{Validity and Limitations}
 
@@
 
-\subsection{Practical Extension to Include Magnetic Lenses and RF Cavities}
+\section{Practical Extension to Include Magnetic Lenses and RF Cavities}
 
-\subsubsection{Region of Influence}
+\subsection{Region of Influence}
 
@@
 
-\subsubsection{Magnetic Lenses}
+\subsection{Magnetic Lenses}
 
@@
 
-\subsubsection{RF Cavities}
+\subsection{RF Cavities}
 

### 2012-07-10 inc/initial_conditions/initial_conditions.tex
insert commentary from JAP paper to relevant places
A little more reoganization related to splitting the AG model chapter in two. Futher I have started an introduction to the extension chapter

@@
 
-\subsection{Determining Appropriate Initial Conditions} \label{sec:initial_conditions}
+\section{Determining Appropriate Initial Conditions} \label{sec:initial_conditions}
 For a DTEM or UED apparatus, where the electron pulse is produced in a laser-driven photo-electron gun, the implementation of the AG model of Michalik and Sipe \cite{michalik_analytic_2006} requires knowledge of the initial ($ t = 0 $) bunch parameters immediately after generation at the photocathode surface inside the acceleration region of the gun.

### 2012-07-10 main.tex
minor
Added basic chapter/section structure
Added bibliography from model_lenses paper as a starting point. Added abstraction code for a choice between bibtex and biblatex.
removed fontsym.tex as it really was just part of the manual.
Began some of the mathematical sections.
More on MS model
added gaussian distribution section
Renamed subsection title to "Evolution of a Gaussian Distribution", added abbrev. LHS and RHS
Added text to ms_model
switched to import for directory control
worked out some teething issues for import package
Began porting the OE laser paper into the thesis
Began porting

@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \documentclass{uicthesi}
-\usepackage{amsfonts}
-\usepackage{amssymb}
+\usepackage{amsmath,amsfonts,amssymb}
+\usepackage{esint}
+\usepackage{hyperref}
 \usepackage{url}
+\usepackage{ifthen}
+\usepackage{pgffor}
+
 %\usepackage{euler}
+%\usepackage{breqn}
+
+\usepackage[version=3]{mhchem}
+
+\usepackage{import}
+\usepackage{subfig}
 
@@
 \DeclareGraphicsExtensions{.png,.pdf,.jpg}
+\input{mytikz}
+
+%% My convenience fuctions
+\providecommand{\vecprime}[1]{\vec{#1}\,^{\prime}}
+
+\providecommand{\smallT}{\ensuremath{{\scriptscriptstyle T}}}
+\providecommand{\smallD}{\ensuremath{{\scriptscriptstyle D}}}
+\providecommand{\smallF}[0]{ { \scriptscriptstyle F } }
+%\providecommand{\smallZ}{\ensuremath{{\scriptscriptstyle Z}}}
+
+\providecommand{\abbr}[2]{\item[#1\hfill] #2}
+
+%% Special handling for using biblatex vs bibtex ==
+\newboolean{usebiblatex}
+\setboolean{usebiblatex}{false}
+
+\ifthenelse{\boolean{usebiblatex}}{
+  \usepackage{biblatex}
+  \addbibresource{bibliography.bib}
+}{}
+%% ==
 
@@
 \pdegrees{of the\\University of Illinois at Chicago}
-\degree{Doctor of Philosophy in Experimental Physics}
+\degree{Doctor of Philosophy in Physics}
 
@@
 \listofabbreviations
-%\begin{list}
-%{}
-%{\setlength
-%   {\labelwidth}{1in}
-%    \setlength{\leftmargin}{1.5in}
-%    \setlength{\labelsep}{.5in}
-%    \setlength{\rightmargin}{\leftmargin}}
-%\item[AMS\hfill] American Mathematical Society
-%\item[CTAN\hfill] Comprehensive \TeX\ Archive Network
-%\item[TUG\hfill] \TeX\ Users Group
-%\item[UIC\hfill] University of Illinois at Chicago
-%\item[UICTHESI\hfill] Thesis formatting system for use at UIC.
-%\end{list}
+\begin{list}
+  {}
+  {\setlength{\labelwidth}{1in}
+   \setlength{\leftmargin}{1.5in}
+   \setlength{\labelsep}{.5in}
+   \setlength{\rightmargin}{\leftmargin}}
+
+  \abbr{AG}{Analytic Gaussian (Model)}
+  \abbr{AR}{Anti-Reflection}
+  \abbr{DC}{Direct Current (i.e. Static Electric Field)}
+  \abbr{DTEM}{Dynamic Transmission Electron Microscopy}
+  \abbr{FWHM}{Full-Width Half Maximum}
+  \abbr{GRIN}{Gradient Index (of Refraction)}
+  \abbr{GVD}{Group Velocity Dispersion}
+  \abbr{HW1/eM}{Half-Width 1/e Maximum}
+  \abbr{LHS}{Left Hand Side}
+  \abbr{ML}{Mode-Locked}
+  \abbr{Q}{Quality Factor}
+  \abbr{RHS}{Right Hand Side}
+  \abbr{RF}{Radio Frequency}
+  \abbr{SBR}{Saturable Bragg Reflector}
+  \abbr{TLS}{Thermal Lens Shaping}
+  \abbr{TEM$_{00}$}{Tranvserse ElectroMagnetic 00 Mode}
+  \abbr{TM$_{010}$}{Tranvserse Magnetic 010 Mode}
+  \abbr{UED}{Ultrafast Electron Diffraction}
+  \abbr{UEM}{Ultrafast Electron Microscope}
+  \abbr{UIC}{University of Illinois at Chicago}
+\end{list}
 
@@
 
+\chapter{Review of Foundational Modeling Efforts} \label{chap:previous_models}
+
+  \subimport{inc/ms_model/}{ms_model}
+
+\chapter{Extension of the Analytic Gaussian Model\\for UEM Column Modeling}
+
+Modeling a system as complex as an Ultrafast Electron Microscope is a task too large for one person.
+In the previous chapter\ref{chap:previous_models} I have highlighted some previous efforts at modeling electron pulse evolution.
+While these are excellent works in their own right, none provide the complete picture of the evolution of an electron pulse through the entire column.
+In the following chapter I will present a model which builds on and extends those individual pieces to form a single model; one which is generic enough to model many different columns and which is computationally fast enough to be used to design and optimize a column.
+
+  \subimport{inc/liouvilles_theorem/}{liouvilles_theorem}
+
+  \subimport{inc/initial_conditions/}{initial_conditions}
+
+  \subimport{inc/external_forces/}{external_forces}
+
+\section{Model Results, Predictions and Implications} \label{sec:model_results}
+
+\chapter{Prototype Instrument}
+
+  \subimport{inc/laser/}{laser}
+
+\section{Chamber Design}
+
+\section{Accelerator Design}
+
+\section{Magnetic Lenses}
+
+  \subimport{inc/magnetic_lenses/}{maglens}
+
+\section{Deflection Plates}
+
+\chapter{Photocathode Engineering}
+
+\section{Flat Metal Photocathodes}
+
+\section{Nanopatterned (Plasmonic) Photocathodes}
+
+\section{Semiconductor Photocathodes}
+
+  \subimport{inc/semicond/}{semicond}
+
+\chapter{Experimental Results}
+
 %\appendices
@@
 %\newpage
-%\input fontsym
-%\bibformb
-%\bibliography{uictman}
+
+%% Special handling for using biblatex vs bibtex ==
+\ifthenelse{\boolean{usebiblatex}}{
+  \printbibliography
+}{
+  \bibformc
+  \bibliography{bibliography,biblaser}
+}
+%% ==
+
 %\vita

### 2012-07-10 inc/external_forces/external_forces.tex
continuing to move \section directives into subimported files

@@
 A magnetic lens may be modeled using one linear force contribution.
-Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is very complicated to describe\cite{montgomery_some_1961} and the ``lensing'' action is actually a compound force (first rotating the beam, then lensing).
+Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is very complicated to describe\cite{montgomery_some_1961} and the ``lensing'' action is actually a compound force (first rotating the beam, then lensing, see Section \ref{sec:mag_lens}).
 We have found that in practice, characterizing a specific magnetic lens ($M_{\smallT}(I)$ where $I$ is the current in the magnetic lens) is much easier than attempting to find a generic functional form.

### 2012-07-10 inc/liouvilles_theorem/liouvilles_theorem.tex
insert commentary from JAP paper to relevant places
A little more reoganization related to splitting the AG model chapter in two. Futher I have started an introduction to the extension chapter
highlight that \Gamma^2 constants are indeed constants, evaluated at time t=0

@@
 
-\subsection{Implications of Liouville's Theorem}
+\section{Implications of Liouville's Theorem}
 
@@
 Not only is this an interesting result, but it can be both instructive and computationally useful.
-By rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$ and substituting into \ref{eq:ag_original}.
+By rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$, evaulating at some time (presumably $t=0$) and substituting into \ref{eq:ag_original}.
 \begin{subequations} \label{eq:ag_big_gamma}
@@
   \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
-  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
+  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2(0) + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
     + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi)
@@
 it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.
+
+In the transverse dimension, this is consistent with the definition of `coherent fluence' employed by Reed et al.\cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the spatial emittance of an electron beam\cite{jensen_theoretical_2006,siwick_ultrafast_2002}.

### 2012-07-10 inc/external_forces/external_forces.tex
(hopefully) clarified some of the notation and explaination of the purely mathematical parts of the extension

@@
 
+Besides a set of initial conditions, the Analytic Gaussian model presented by Michalik and Sipe (see Section \ref{sec:ms_model}) is only useful for modeling free-space propagation.
+This is because there is no allowance for external forces, such as those imparted by the ``optical'' elements of an electron microscope column.
+
 \subsection{Mathematical Formulation of External Force Contribution}
-\ref{eq:potential_integral} is the entry point for the internal Coulombic repulsion force; since both potentials and integrals are linear, one may add additional potentials so that it becomes
+
+While the free-space AG model does not contain any external forces, it does include an internal force.
+The Coulomb force is included to model the repulsion of the pulse's electrons from each other.
+Because this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identically ``Gaussian'' optics.\cite{michalik_analytic_2006}
+
+The internal Coulombic repulsion force enters the model in \ref{eq:potential_integral}.
+Since both potentials and integrals are linear, one may add additional potentials so that it becomes
+\begin{equation}
+  V(\vec{r} - \vecprime{r}) \rightarrow V(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r})
+\end{equation}
+Following this new term through the derivation, we see that
 \begin{equation}
   \begin{split}
-    \Phi(\vec{r};t) \rightarrow & \int d\vecprime{r} \left( V(\vec{r} - \vecprime{r}) + \sum V_{ext}(\vec{r}) \right) n(\vecprime{r}; t) \\
-    = & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) + \sum \int d\vecprime{r} V_{ext}(\vec{r}) n(\vecprime{r}; t) \\
-    = & \Phi(\vec{r};t) + \Phi_{ext}(\vec{r};t)
+    \Phi(\vec{r};t) = & \int d\vecprime{r} \left( V(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r}) \right) n(\vecprime{r}; t) \\
+    = & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) + \sum_{k} \int d\vecprime{r} V_{k}^{ext}(\vec{r}) n(\vecprime{r}; t) \\
+    \equiv & \Phi^{int}(\vec{r};t) + \sum_{k} \Phi_{k}^{ext}(\vec{r};t)
   \end{split}
 \end{equation}
-This form of $\Phi(\vec{r};t)$ can be substituted into the definition of $K^{force}$ (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector} 
+Therefore each external force contributes an additional term defined as
+\begin{equation}
+  \Phi^{ext}(\vec{r};t) \equiv \int d\vecprime{r} V^{ext}(\vec{r}) n(\vecprime{r}; t)
+\end{equation}
+These external terms are not self-influencing, and thus are not subject to the complexities of the internal Coulombic repulsion force.
+For clarity we now define $\Phi^{int}(\vec{r};t)$ as the old $\Phi(\vec{r};t)$, which results from the Coulomb force, while our new $K^{force}$ now includes these exteral terms.
+
+This form of $\Phi(\vec{r};t)$ can now be substituted into the definition of $K^{force}$ matrix (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector} 
 \begin{equation}
   \begin{split}
-    K^{force}_{ij} \rightarrow & \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial}{\partial \vec{r}} \left( \Phi(\vec{r};t) + \Phi_{ext}(\vec{r};t) \right) d\vec{r} d\vec{p} \\
-    \equiv & K^{int}_{ij} + \sum K^{ext}_{ij}
+    K^{force}_{ij} = & \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial}{\partial \vec{r}} \left( \Phi(\vec{r};t) + \sum_{k} \Phi_{k}^{ext}(\vec{r};t) \right) d\vec{r} d\vec{p} \\
+    \equiv & K^{int}_{ij} + \sum_{k} (K_{k}^{ext})_{ij}
   \end{split}
 \end{equation}
-where $K^{int}$ is simply the old $K^{force}$. Noting that the spatial derivative of a potential is just the force, we can now define the matrix contributed by each external force is given by
+Like the potential integrals, we are left with an additional matrix for each external force.
+Again, for clarity, we now define $K^{int}$ as the $K$ matrix resulting from the internal Coulombic repulsion.
+
+Noting that the spatial derivative of a potential is just the force, we can now define the matrix contributed by each external force is given by
 \begin{equation} \label{eq:Kint}

### 2012-08-22 inc/plasmon/plasmon.tex
added plasmon file, include cc warning on pole piece figure

--- 2012-05-21 inc/liouvilles_theorem/liouvilles_theorem.tex
@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
-
-\subsection{Implications of Liouville's Theorem}
-
-One of the most well known theorems in physics (and mathematics) is Liouville's Theorem.
-This theorem can be summarized succinctly as a `the total phase space of a system is conserved.'
-%TODO find a good reference for this
-
-To evaluate the effect of this theorem on these equations, consider the time derivative of the product $\sigma_{\alpha} \eta_{\alpha}$.
-Using \ref{eq:ag_original} it is easy to show that
-\begin{gather}
-  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0\\
-  \label{eq:define_big_gamma}
-  \therefore \sigma_{\alpha} \eta_{\alpha} \equiv \Gamma_{\alpha}^2 = \text{Constant}
-\end{gather}
-This analysis shows that the Analytic Gaussian model explicitly obeys the Liouville Theorem.
-
-Not only is this an interesting result, but it can be both instructive and computationally useful.
-By rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$ and substituting into \ref{eq:ag_original}.
-\begin{subequations} \label{eq:ag_big_gamma}
-\begin{gather}
-  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
-  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
-    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi)
-\end{gather}
-\end{subequations}
-these four equations now contain all of the information previously contained in six.
-Having fewer equations, and fewer mathematical operations, makes this set of equations much less computationally intensive.
-Further, now one can see that these $\Gamma$ constants can be thought of as some minimum/intrinsic action of the pulse. 
-This will become more apparent when in future sections
-%TODO include internal reference
-it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.

### 2012-08-22 inc/plasmon/plasmon.tex
expand on plasmon

@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
+\section{Introduction}
+
+In the search for alternative mechanisms to reduce the intrinsic divergence of the electron beam upon photoemission, a possible area of interest was Plasmon-Assisted Photoemission (PAPE). 
+This avenue of investigation was motivated by \fixme{cite} in which authors \fixme{authors} indicate that when driving a surface plasmon on gold they witnessed a reduction in the emission angle of photoemitted electrons.
+In their geometry (called the Kretschmann geometry) the plasmon is excited by a laser through a glass substrate, however due to fears about heat transportation, it was decided to use the ``grating coupling'' geometry.
+
+\subsection{Driving the Plasmon}
+
+Every material has an intrinsic plasma frequency given by 
+\begin{equation}
+  \omega_{P} = \sqrt{\frac{n e^2}{\varepsilon_{0} m^*}}
+\end{equation}
+In order to drive a plasmon on the surface of a material, the portion of the laser wave-vector parallel to the plasmon ($\vec{k}_{laser,x}$) must be matched to the wave-vector of the surface plasmon.

### 2012-08-22 inc/plasmon/plasmon.tex
added grating damage image

--- 2012-08-22 inc/plasmon/plasmon.tex
@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
+\section{Introduction}
+
+In the search for alternative mechanisms to reduce the intrinsic divergence of the electron beam upon photoemission, a possible area of interest was Plasmon-Assisted Photoemission (PAPE). 
+This avenue of investigation was motivated by Zawadzka et al \cite{zawadzka_evanescent_2001} who indicate that when driving a surface plasmon on gold they witnessed a reduction in the emission angle of photoemitted electrons.
+In their paper, the authors mount a gold foil on a prism; the laser is incident on the gold through the prism.
+This geometry is called the Kretschmann geometry.
+
+While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, that UEM applications should explore the ``grating coupling'' geometry.
+In this geometry, the plasmon is driven on a periodic surface rather than a flat surface attached to glass.
+Futher, in this geometry the back surface is now free to be attached to a heat sink.
+
+\subsection{Driving the Plasmon}
+
+Every material has an intrinsic plasma frequency given by 
+\begin{equation}
+  \omega_{P} = \sqrt{\frac{n e^2}{\varepsilon_{0} m^*}}
+\end{equation}
+where $n$ is the electron density and $m*$ is the effective mass of the material.
+In order to drive a plasmon on the surface of a material, the portion of the laser wave-vector parallel to the plasmon ($\vec{k}_{laser,x}$) must be matched to the wave-vector of the surface plasmon.
+
+\section{Experimental Attempts}
+
+\subsection{Commerical Grating on Glass}
+
+For proof-of-concept purposes the first experiment was run on a commercially available gold foil on glass holographic grating.
+Like the Kretschmann geometry, a foil on glass grating is vulnerable to heating problems; a plasmon creates a large amount of heat but the glass is a poor heat conductor.
+The grating was operational as the angles were tuned attempting to drive the plasmon.
+At the instant that the plasmon started, a bright flash was visible on the scintillator screen then disappeared.
+
+After this, the grating showed signs of damage.
+\ref{fig:grating-damage} shows the results of Scanning Electron Microscopy (SEM) on the grating.
+The damage pattern is round and of the size of the laser spot used to drive the plasmon.
+The extent of the damage indicates a large temperature increase, which given the other evidence of the test, can be assumed to be from the sudden appearance of a surface plasmon.
+
+The most important result was that a plasmon was driven successfully; this created the heat that destroyed the grating.
+Further this experiment validated the decision not to persue a Kretschmann geometry for fear of heat damage at these high powers.
+
+\begin{figure}
+  \centering
+  \includegraphics{damage.png}
+  \caption{Damage to commerical gold-on-glass holographic grating.}
+  \label{fig:grating-damage}
+\end{figure}
+
+\subsection{Sinusoidal Grating}
+
+\subsection{Sinusoidal Grating, Rotated}
+
+\subsection{Trenched Grating}

### 2012-08-29 inc/laser/laser.tex
removed some errant unicode

@@
 Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
-As with other Yb-doped laser crystals\cite{Brenier_new_criteria}, this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
+As with other Yb-doped laser crystals\cite{Brenier_new_criteria}, this is primarily due to the gain medium's advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
 Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially\cite{website_amplitude,website_solar}.
@@
 %TODO replace to "This output will then be use to ..." and link to something from previous
-%In our case, the visible femtosecond pulses will be used as the driving laser radiation source for an ultrafast electron microscope \cite{king_ultrafast_2005} – a technology that seeks to combine the sub-nanometer spatial resolution of electron microscopy with the sub-picosecond temporal resolution afforded by today’s ultrashort pulse laser systems.
+%In our case, the visible femtosecond pulses will be used as the driving laser radiation source for an ultrafast electron microscope \cite{king_ultrafast_2005} --- a technology that seeks to combine the sub-nanometer spatial resolution of electron microscopy with the sub-picosecond temporal resolution afforded by today's ultrashort pulse laser systems.
 %Critical to this technology is likely to be the development of a femtosecond photo-electron gun with a low spatial emittance\cite{zawadzka_evanescent_2001}.
@@
 4(b)) at an optimum mode-locked output power of $5.0$W.
-This trend is consistent with the expected ‘soliton’ modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity\cite{Brunner_diode_pumped}.
+This trend is consistent with the expected `soliton' modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity\cite{Brunner_diode_pumped}.
 
@@
 This is consistent both with the laser accessing more of the available inhomogeneous gain bandwidth and with our resonator design favoring operation with a strong Kerr lens in the Yb:KGW gain medium.
-Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal’s optical $N_p$-axis\cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
+Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal's optical $N_p$-axis\cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
 From Fig.
@@
 The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
-The crystal is uncoated and cut for frequency doubling with the fundamental $p$-polarized radiation at 1046nm incident at Brewster’s angle --- a geometry that ensures both minimum loss for the fundamental radiation and, more importantly, a high damage threshold.
+The crystal is uncoated and cut for frequency doubling with the fundamental $p$-polarized radiation at 1046nm incident at Brewster's angle --- a geometry that ensures both minimum loss for the fundamental radiation and, more importantly, a high damage threshold.
 The doubling crystal is positioned at the focus of a $z$-fold comprising a 1040nm high reflector with a radius of curvature of 75mm followed by a collimating mirror with a radius of curvature of 100mm reflective at both the fundamental and second harmonic wavelengths.

### 2012-09-05 inc/plasmon/plasmon.tex
remove some unicode

@@
 where $n$ is the electron density and $m*$ is the effective mass of the material.
-In order to drive a plasmon on the surface of a material, the portion of the laser wave-vector parallel to the plasmon ($\vec{k}_{laser,x}$) must be matched to the wave-vector of the surface plasmon.
+In order to drive a plasmon on the surface of a material, the portion of the laser wave-vector parallel to the plasmon ($k_{Lx}$) must be matched to the wave-vector of the surface plasmon.
+\begin{equation}
+  k_{P} = k_{Lx} + i k_{G}
+\end{equation}
 

### 2012-09-05 inc/plasmon/plasmon.tex
improving the introduction to SPP

@@
 
-\subsection{Driving the Plasmon}
+\subsection{Driving the Surface Plasmon}
+
+As the name suggests, a plasmon is the quasiparticle of plasma oscillation.
+Electrons in a material can couple with electromagnetic fields, in this case a laser field, and oscillate accordingly.
+When this occurs at the surface of a material, the resultant quasiparticles are called ``surface plasmon-polaritons.''
+Since these will be the only plasmons discussed in this paper, the word ``plasmon'' will be used interchangeably for the name ``surface plasmon-polaritons.''
+%TODO should this be formalized, either one way (SPP) or the other
+
+The mathematical formalism of surface plasmon-polaritons is a lengthy subject and well described elsewhere \cite{cottam_introduction_2004,concepts_2002}.
 
@@
 \begin{equation}
-  \omega_{P} = \sqrt{\frac{n e^2}{\varepsilon_{0} m^*}}
+  \omega_{P} = \sqrt{\frac{\rho_{e} e^2}{\varepsilon_{0} m^*}}
 \end{equation}
-where $n$ is the electron density and $m*$ is the effective mass of the material.
+where $\rho_{e}$ is the electron density and $m*$ is the effective mass of the material.
+And in the most simple (undamped) approximations, the dielectric function of the material may be written in terms of this constant.
+\begin{equation} \label{eq:dielectric_as_omega}
+  \varepsilon(\omega) = 1 - \frac{ \omega_{P}^2 }{ \omega^2 }
+\end{equation}
+
+At the interface of a vacuum and such a simple material material, contstraints on the material parameters which may allow surface plasmons result in the condition that the dielectric constant at the oscillation frequency of the surface plasmon $ \varepsilon(\omega_{SP}) = -1 $.
+Substituting this result into \ref{eq:dielectric_as_omega} show that this frequency of oscillation is simply related to intrinsic plasma frequency by
+\begin{equation}
+  \omega_{SP} = \omega_{P} / \sqrt{2}
+\end{equation}
+
+Further, plasmons carry momentum whose magnitude is given by
+\begin{equation}
+  k_{P}^2 = \frac{\omega^2}{c^2} \left( \frac{\varepsilon_{1} \varepsilon_{2}}{\varepsilon_{1} + \varepsilon_{2}} \right)
+\end{equation}
+but of course for a vacuum interface this is simplified to 
+\begin{equation}
+  k_{SP}^2 = \frac{\omega^2}{c^2} \left( \frac{ \varepsilon }{ 1 + \varepsilon } \right)
+\end{equation}
+
 In order to drive a plasmon on the surface of a material, the portion of the laser wave-vector parallel to the plasmon ($k_{Lx}$) must be matched to the wave-vector of the surface plasmon.
 \begin{equation}
-  k_{P} = k_{Lx} + i k_{G}
+  k_{SP} = k_{Lx} + \imath k_{G}
 \end{equation}

### 2012-09-12 inc/plasmon/plasmon.tex
further describe the mathematics of plasmons

@@
 
-In order to drive a plasmon on the surface of a material, the portion of the laser wave-vector parallel to the plasmon ($k_{Lx}$) must be matched to the wave-vector of the surface plasmon.
+In the grating coupling scheme (shown in \ref{fig:plasmon_schematic}) a laser of wavelength $\lambda$ is incident on a material with a grating of period $d$, making an angle $\theta$ from the surface normal.
+A surface plasmon is driven when the portion of the laser wave-vector parallel to the plasmon ($k_{Lx}$) and a fourier component ($k_{G} = 2 \pi / d$) of the grating matches the wave-vector of the surface plasmon
 \begin{equation}
-  k_{SP} = k_{Lx} + \imath k_{G}
+  k_{SP} = k_{Lx} + n k_{G}
 \end{equation}
+where $n$ here is an integer.
+Which reduces to 
+\begin{equation} \label{eq:plasmon_condition}
+  \frac{ \varepsilon }{ 1 + \varepsilon } = \sin \theta + \frac{ n \lambda }{ d }
+\end{equation}
+For a given grating period, \ref{eq:plasmon_condition} determines the angle of incidence that the laser must make with the surface in order to drive the plasmon.
+Further, it is used to design the grating, allowing for angles that the microscope column will accept.
+
+\begin{figure}
+  \centering
+  \input{grating_schematic}
+  \caption{Schematic diagram of a grating coupling geometry for driving a surface plasmon}
+  \label{fig:plasmon_schematic}
+\end{figure}
 

### 2012-09-24 inc/external_forces/external_forces.tex
use space between word and \cite

@@
 A magnetic lens may be modeled using one linear force contribution.
-Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is very complicated to describe\cite{montgomery_some_1961} and the ``lensing'' action is actually a compound force (first rotating the beam, then lensing, see Section \ref{sec:mag_lens}).
+Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is very complicated to describe \cite{montgomery_some_1961} and the ``lensing'' action is actually a compound force (first rotating the beam, then lensing, see Section \ref{sec:mag_lens}).
 We have found that in practice, characterizing a specific magnetic lens ($M_{\smallT}(I)$ where $I$ is the current in the magnetic lens) is much easier than attempting to find a generic functional form.

### 2012-09-24 inc/laser/laser.tex
use space between word and \cite

@@
 Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
-As with other Yb-doped laser crystals\cite{Brenier_new_criteria}, this is primarily due to the gain medium's advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
-Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially\cite{website_amplitude,website_solar}.
-Such lasers have used dichroic mirrors\cite{Brunner_diode_pumped,Major_femtosecond_2006,Paunescu_diode_2004,Major_extended_2006}, polarization-coupling\cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers\cite{website_amplitude,Brunner_pulses_2002}) to pump the laser crystal in an efficient manner.
+As with other Yb-doped laser crystals \cite{Brenier_new_criteria}, this is primarily due to the gain medium's advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
+Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated \cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially \cite{website_amplitude,website_solar}.
+Such lasers have used dichroic mirrors \cite{Brunner_diode_pumped,Major_femtosecond_2006,Paunescu_diode_2004,Major_extended_2006}, polarization-coupling \cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers \cite{website_amplitude,Brunner_pulses_2002}) to pump the laser crystal in an efficient manner.
 
 % page 8631
-The spectroscopic properties of the biaxial Yb:KGW laser crystal\cite{Biswal_thermo_optical_05} indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength\cite{Brenier_new_criteria}.
-However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis\cite{Holtom_mode_locked_2006,Biswal_thermo_optical_05}.
-This optimal requirement for orthogonal linear pump and laser polarizations led Holtom\cite{Holtom_mode_locked_2006} to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
+The spectroscopic properties of the biaxial Yb:KGW laser crystal \cite{Biswal_thermo_optical_05} indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength \cite{Brenier_new_criteria}.
+However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis \cite{Holtom_mode_locked_2006,Biswal_thermo_optical_05}.
+This optimal requirement for orthogonal linear pump and laser polarizations led Holtom \cite{Holtom_mode_locked_2006} to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
 
 %TODO is this first sentance correct or should I use "I" voice?
-In this section, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis\cite{Biswal_thermo_optical_05,mochalov_laser_1997,pujol_crystalline_1999}).
-The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media\cite{Rimington_thermal_lens_2004}.
+In this section, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis \cite{Biswal_thermo_optical_05,mochalov_laser_1997,pujol_crystalline_1999}).
+The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media \cite{Rimington_thermal_lens_2004}.
 It employs a simple single layer \ce{SiO2} anti-reflection coating on the gain medium to provide minimal loss for $p$-polarized intracavity 1040nm laser radiation and less than 0.5\% surface reflection for orthogonal $s$-polarized 980nm pump light.
@@
 %TODO change voice/check ->
-To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media\cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.
+To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk \cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped \cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators \cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media \cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.
 %TODO replace to "This output will then be use to ..." and link to something from previous
 %In our case, the visible femtosecond pulses will be used as the driving laser radiation source for an ultrafast electron microscope \cite{king_ultrafast_2005} --- a technology that seeks to combine the sub-nanometer spatial resolution of electron microscopy with the sub-picosecond temporal resolution afforded by today's ultrashort pulse laser systems.
-%Critical to this technology is likely to be the development of a femtosecond photo-electron gun with a low spatial emittance\cite{zawadzka_evanescent_2001}.
-%The coherent control of photoemission through visible laser-driven plasmon excitations on large-area (a few mm2), gold or silver, nano-patterned photocathodes may provide a mechanism by which this can be achieved\cite{zawadzka_evanescent_2001}.
+%Critical to this technology is likely to be the development of a femtosecond photo-electron gun with a low spatial emittance \cite{zawadzka_evanescent_2001}.
+%The coherent control of photoemission through visible laser-driven plasmon excitations on large-area (a few mm2), gold or silver, nano-patterned photocathodes may provide a mechanism by which this can be achieved \cite{zawadzka_evanescent_2001}.
 
@@
 
-The design of the laser head follows the example of Holtom\cite{Holtom_mode_locked_2006}, which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism\cite{Rimington_thermal_lens_2004}.
+The design of the laser head follows the example of Holtom \cite{Holtom_mode_locked_2006}, which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism \cite{Rimington_thermal_lens_2004}.
 To ensure efficient operation of a Yb:KGW laser, the dominant absorption feature at 980nm for radiation polarized parallel to the optical $N_m$-axis ($\sim\,20^\circ$ from the crystallographic a-axis) should be pumped using high brightness diode lasers.
-On the other hand, as with \ce{Yb$:$KY(WO4)2} \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis\cite{Holtom_mode_locked_2006}) in this biaxial material.
+On the other hand, as with \ce{Yb$:$KY(WO4)2} \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis \cite{Holtom_mode_locked_2006}) in this biaxial material.
 %page 8632
-Instead of using a polarization-coupled pumping geometry\cite{Holtom_mode_locked_2006} to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig.
+Instead of using a polarization-coupled pumping geometry \cite{Holtom_mode_locked_2006} to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig.
 1(a), where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
@@
 
-Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH\cite{website_limo}).
+Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH \cite{website_limo}).
 The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm fullwidth half-maximum (FWHM) 980nm absorption line of Yb:KGW \cite{Paunescu_diode_2004,Biswal_thermo_optical_05,website_EKSPLA}.
@@
 The Yb concentration was chosen to be 2\% atomic doping to ensure that more than 90\% of the pump radiation was absorbed over the 3mm Yb:KGW crystal thickness.
-With the employed counter-propagating pump geometry, the resultant maximum average pump irradiance over the crystal length is then roughly 3 times the $\sim$6kW/cm$^2$ saturation intensity of the dominant 980nm absorption line\cite{Brunner_powerful_2004}.
-The laser crystal was obtained from NovaPhase\cite{website_nova}.
+With the employed counter-propagating pump geometry, the resultant maximum average pump irradiance over the crystal length is then roughly 3 times the $\sim$6kW/cm$^2$ saturation intensity of the dominant 980nm absorption line \cite{Brunner_powerful_2004}.
+The laser crystal was obtained from NovaPhase \cite{website_nova}.
 
-To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique\cite{Rimington_thermal_lens_2004} to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal.
+To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique \cite{Rimington_thermal_lens_2004} to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal.
 For the employed non-Brewster crystal cut (Fig. 1(a)), a ray
@@
 expansion with coefficient $\alpha T$ \cite{Rimington_thermal_lens_2004}.
-Generally, $\Delta \approx W l$ , where $W$ is the transverse size (radius) of the pumped region\cite{Koechner_thermal_1970}.
+Generally, $\Delta \approx W l$ , where $W$ is the transverse size (radius) of the pumped region \cite{Koechner_thermal_1970}.
 In our case, $l \approx$ 3mm and $W \approx$ 0.2mm, so that $\Delta < 0.1$, implying that the thermal duct due to a non-zero $\frac{dn}{dT}$ primarily determines the required $B_T$/$A_T$ ratio.
@@
 where $\kappa$ is the thermal diffusion coefficient.
-Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry\cite{Rimington_thermal_lens_2004}.
-The specifications of our dual-axis collimated diodes\cite{website_limo} indicate that heat source can be written as
+Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry \cite{Rimington_thermal_lens_2004}.
+The specifications of our dual-axis collimated diodes \cite{website_limo} indicate that heat source can be written as
 \begin{equation}
@@
 that is, the focused pump radiation is described accurately by a Gaussian irradiance distribution in the horizontal ($x$) direction with half-width 1/e maximum (HW1/eM) spot size $W_x$ and a super-Gaussian of order 3 in the vertical ($y$) direction with spot size $W_y$.
-\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique\cite{Rimington_thermal_lens_2004}; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
+\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique \cite{Rimington_thermal_lens_2004}; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
 For a fixed vertical pump spot size of $W_y = 0.2$mm, our analysis indicates that in order to achieve the required ratio $B_T / A_T = 1.3$ in the parabolic approximation for $T(x,y)$ about $(x,y) = (0,0)$ a horizontal pump spot size $W_x = 0.3$mm is needed; in other words, the required $Q(x,y)$ heat source spot size ratio (or ellipticity) $W_x / W_y = 1.5$.
@@
 1(a)).
-This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium\cite{Brenier_new_criteria}; in particular, in the vertical direction (Fig.
+This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium \cite{Brenier_new_criteria}; in particular, in the vertical direction (Fig.
 2(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
@@
 Clearly, this can be readily accommodated by the $W_x = 0.3$mm horizontal pump spot size when $w_x \approx 180\mu$m.
-However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium\cite{Brenier_new_criteria}.
+However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium \cite{Brenier_new_criteria}.
 
-We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) \cite{Biswal_thermo_optical_05,mochalov_laser_1997} and stress-induced refractive index changes\cite{Yumashev_laser_2007}.
+We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) \cite{Biswal_thermo_optical_05,mochalov_laser_1997} and stress-induced refractive index changes \cite{Yumashev_laser_2007}.
 For the employed crystal cut in our TLS geometry (Fig.
@@
 %page 8635
-changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index\cite{website_time_bandwidth}.
-Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping\cite{Yumashev_laser_2007,Hellstrom_efficient_2006}.
-Consequently, for our laser polarized along the optical $N_p$-axis\cite{Holtom_mode_locked_2006}, there is insufficient information to include accurately the effects of stress-induced refractive index changes.
+changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index \cite{website_time_bandwidth}.
+Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping \cite{Yumashev_laser_2007,Hellstrom_efficient_2006}.
+Consequently, for our laser polarized along the optical $N_p$-axis \cite{Holtom_mode_locked_2006}, there is insufficient information to include accurately the effects of stress-induced refractive index changes.
 
@@
 The diode-pumped TLS Yb:KGW laser head is placed in the 2.35m-long, asymmetric z-fold cavity shown in Fig. 3(a).
-The two focusing gain section mirrors each have a group velocity dispersion (GVD) of $-1300 ( \pm 150 ) \text{fs}^2$ (Layertec GmbH\cite{website_layertec}), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
+The two focusing gain section mirrors each have a group velocity dispersion (GVD) of $-1300 ( \pm 150 ) \text{fs}^2$ (Layertec GmbH \cite{website_layertec}), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
 The 5$^\circ$ angle of incidence on the gain section mirrors results in minimal astigmatism, although any net round-trip cavity astigmatism can be compensated for by the TLS technique.
@@
 %page 8636
-focusing the intracavity radiation on a saturable Bragg reflector (SBR) with a 1\% reflectivity modulation depth at 1040nm (BATOP GmbH\cite{website_BATOP}) positioned a distance $z$ from the focusing mirror.
+focusing the intracavity radiation on a saturable Bragg reflector (SBR) with a 1\% reflectivity modulation depth at 1040nm (BATOP GmbH \cite{website_BATOP}) positioned a distance $z$ from the focusing mirror.
 The shorter 63cm arm is terminated by a plane output coupler with a 7\% transmission at 1040nm.
@@
 The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
-This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW \cite{Biswal_thermo_optical_05}, but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis\cite{Holtom_mode_locked_2006,Hellstrom_efficient_2006}.
+This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW \cite{Biswal_thermo_optical_05}, but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis \cite{Holtom_mode_locked_2006,Hellstrom_efficient_2006}.
 
@@
 The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (Fig. 4(a)).
-The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH\cite{website_layertec}) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
+The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH \cite{website_layertec}) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
 For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (Fig.
 4(b)) at an optimum mode-locked output power of $5.0$W.
-This trend is consistent with the expected `soliton' modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity\cite{Brunner_diode_pumped}.
+This trend is consistent with the expected `soliton' modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity \cite{Brunner_diode_pumped}.
 
@@
 4(a)), but is always less than the 50cm focal length of the concave high reflector.
-This is due to the need to match the absorption-induced thermal bowing of the SBR (i.e., the radius of curvature of its front surface mirror) to the radius of curvature of the intra-cavity laser mode\cite{Schieffer_dual_passive_2006}.
+This is due to the need to match the absorption-induced thermal bowing of the SBR (i.e., the radius of curvature of its front surface mirror) to the radius of curvature of the intra-cavity laser mode \cite{Schieffer_dual_passive_2006}.
 In general, good ML operation is achieved with the SBR positioned within $\pm$1cm from the optimum $z$ distance.
@@
 This is consistent both with the laser accessing more of the available inhomogeneous gain bandwidth and with our resonator design favoring operation with a strong Kerr lens in the Yb:KGW gain medium.
-Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal's optical $N_p$-axis\cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
+Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal's optical $N_p$-axis \cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
 From Fig.
@@
 At the 3.5W ML output power, with a pulse duration of 250fs, the estimated focal irradiance for the fundamental is $\sim50\text{GW}/\text{cm}^2$ in the LBO doubling crystal.
-Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH\cite{website_layertec}), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
+Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH \cite{website_layertec}), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
 After the collimating
@@
 a peak green pulse power of $\sim$130kW.
-This femtosecond visible output will be used as the driving pulsed laser radiation source for an ultrafast electron microscope\cite{king_ultrafast_2005}.
+This femtosecond visible output will be used as the driving pulsed laser radiation source for an ultrafast electron microscope \cite{king_ultrafast_2005}.
 

### 2012-09-24 inc/liouvilles_theorem/liouvilles_theorem.tex
use space between word and \cite

@@
 
-In the transverse dimension, this is consistent with the definition of `coherent fluence' employed by Reed et al.\cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the spatial emittance of an electron beam\cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
+In the transverse dimension, this is consistent with the definition of `coherent fluence' employed by Reed et al.\cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the spatial emittance of an electron beam \cite{jensen_theoretical_2006,siwick_ultrafast_2002}.

### 2012-09-24 inc/ms_model/ms_model.tex
Constant and linear external forces
Split AG model chapter into two different chapters, one for previous efforts and one for my contributions.
(hopefully) clarified some of the notation and explaination of the purely mathematical parts of the extension
use space between word and \cite

@@
 
-\section{Model of Michalik and Sipe}
+\section{The Analytic Gaussian Model of Michalik and Sipe} \label{sec:ms_model}
 
@@
   \begin{pmatrix}
-    k^{force}_{\smallT} & 0 & 0 \\
-    0 & k^{force}_{\smallT} & 0 \\
-    0 & 0 & k^{force}_{z}
+    \hat{k}^{force}_{\smallT} & 0 & 0 \\
+    0 & \hat{k}^{force}_{\smallT} & 0 \\
+    0 & 0 & \hat{k}^{force}_{z}
   \end{pmatrix}
@@
 \begin{equation}
-  k^{force}_{\alpha} = 
+  \hat{k}^{force}_{\alpha} = 
   \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{2\pi^{1/2}\sigma_{\alpha}^{1/2}} L_{\alpha}(\xi)
@@
 
-This AG formalism of Michalik and Sipe\cite{michalik_analytic_2006} employs a mean-field approximation in the evolution of the dynamical effect of the electron pulse's internal space-charge field.
+This AG formalism of Michalik and Sipe \cite{michalik_analytic_2006} employs a mean-field approximation in the evolution of the dynamical effect of the electron pulse's internal space-charge field.
 This is of course not strictly accurate for a Gaussian charge distribution and the error leads to a distortion of the pulse shape.

### 2012-10-31 inc/este/este.tex
removed old semiconductor section and add apl paper as este

--- 2012-09-05 inc/plasmon/plasmon.tex
@@
 
-\section{Introduction}
-
-In the search for alternative mechanisms to reduce the intrinsic divergence of the electron beam upon photoemission, a possible area of interest was Plasmon-Assisted Photoemission (PAPE). 
-This avenue of investigation was motivated by Zawadzka et al \cite{zawadzka_evanescent_2001} who indicate that when driving a surface plasmon on gold they witnessed a reduction in the emission angle of photoemitted electrons.
-In their paper, the authors mount a gold foil on a prism; the laser is incident on the gold through the prism.
-This geometry is called the Kretschmann geometry.
-
-While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, that UEM applications should explore the ``grating coupling'' geometry.
-In this geometry, the plasmon is driven on a periodic surface rather than a flat surface attached to glass.
-Futher, in this geometry the back surface is now free to be attached to a heat sink.
-
-\subsection{Driving the Plasmon}
-
-Every material has an intrinsic plasma frequency given by 
-\begin{equation}
-  \omega_{P} = \sqrt{\frac{n e^2}{\varepsilon_{0} m^*}}
-\end{equation}
-where $n$ is the electron density and $m*$ is the effective mass of the material.
-In order to drive a plasmon on the surface of a material, the portion of the laser wave-vector parallel to the plasmon ($k_{Lx}$) must be matched to the wave-vector of the surface plasmon.
-\begin{equation}
-  k_{P} = k_{Lx} + i k_{G}
-\end{equation}
-
-\section{Experimental Attempts}
-
-\subsection{Commerical Grating on Glass}
-
-For proof-of-concept purposes the first experiment was run on a commercially available gold foil on glass holographic grating.
-Like the Kretschmann geometry, a foil on glass grating is vulnerable to heating problems; a plasmon creates a large amount of heat but the glass is a poor heat conductor.
-The grating was operational as the angles were tuned attempting to drive the plasmon.
-At the instant that the plasmon started, a bright flash was visible on the scintillator screen then disappeared.
-
-After this, the grating showed signs of damage.
-\ref{fig:grating-damage} shows the results of Scanning Electron Microscopy (SEM) on the grating.
-The damage pattern is round and of the size of the laser spot used to drive the plasmon.
-The extent of the damage indicates a large temperature increase, which given the other evidence of the test, can be assumed to be from the sudden appearance of a surface plasmon.
-
-The most important result was that a plasmon was driven successfully; this created the heat that destroyed the grating.
-Further this experiment validated the decision not to persue a Kretschmann geometry for fear of heat damage at these high powers.
-
-\begin{figure}
-  \centering
-  \includegraphics{damage.png}
-  \caption{Damage to commerical gold-on-glass holographic grating.}
-  \label{fig:grating-damage}
-\end{figure}
-
-\subsection{Sinusoidal Grating}
-
-\subsection{Sinusoidal Grating, Rotated}
-
-\subsection{Trenched Grating}
+\section{Excited State Themionic Emission Photocathodes}

### 2012-10-31 inc/este/este.tex
move more of apl paper to este

@@
 \section{Excited State Themionic Emission Photocathodes}
+
+%TODO consider moving this paragraph to chapter intro 
+Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED) [1], ultrafast transmission electron microscopy (UEM) [2], and pulsed X-ray free electron lasers [3].
+The spatial quality of these sources is often characterized in terms of their normalized transverse rms emittance [4,5]; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
+As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) [6], significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness [7]) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source – an intrinsic property of the emission source [4,5].
+
+In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_x$ developed in Refs.
+4 and 5 for thermionic electron emission should be rewritten to read $\sqrt{m^* k_B T_e}$, where $T_e$ is the temperature of the emitting electron distribution and $k_B$ is Boltzmann's constant.
+Direct observation of the transverse momentum distribution for excited-state thermionic emission (ESTE) from GaSb and InSb, two similar III-V zincblende semiconductors, demonstrates the $m^*$ dependence.
+A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m_0$ [8], confirms the temperature dependence of $\Delta p_x$.
+Simulations of the experiment using our extended analytical Gaussian (AG) model of electron pulse propagation (see Section \ref{sec:external_forces}) are in close agreement with the data.
+
+Figure 1 depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
+The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator [10] delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate (see Section \ref{sec:laser}).
+This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV).
+Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
+Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface [7].
+The electron gun design is based on the that of Togawa et al. [7,11] and has an acceleration gap of 25($\pm$2)mm.
+After acceleration in the DC photo-gun, the electrons pass through a set of electrostatic deflection plates to ensure that the electron beam passes, as close as possible, axially through the center of two, 6.35mm-bore, thin and round magnetic lenses with counter-propagating currents to minimize image rotation effects.
+A YAG scintillator placed 27.5cm after the second magnetic lens detects the spatial electron beam profile, and its visible fluorescence is 1:1 imaged onto a CCD detector with 5.4$\mu$m pixels.
+
+To ensure that the transverse momentum distribution is observed, the YAG scintillator must be positioned at the back focal plane (the Fourier plane) of the lens system.
+This is readily achieved by appropriate adjustment of the current $I$ in the magnetic lens coils; specifically, after determining the current required to focus the electron beam onto the scintillator, optical imaging relations together with the $I^2$ dependence of the magnetic lens strength are used to find the current that ensures that the average lens to scintillator distance equals the focal length of the lens system.
+Simulation of the experimental apparatus (employing the extended AG model, see Section \ref{sec:external_forces} ), also shown in Figure 1, confirms that this procedure places the YAG scintillator at the back focal plane of the lens system to within a $\pm$0.5cm error.
+This simulation also confirms that for our experimental conditions (less than 5,000 electrons/pulse and $\Delta x > $ 30$\mu$m) space-charge effects do not influence the observations.
+Additionally, and as expected from optical considerations, the modeling shows that the spatial beam size in the Fourier plane is independent of the incident laser spot size on the photocathode (i.e., the electron source size, $\Delta x_0$).

### 2012-10-31 inc/este/este.tex
more of apl paper into este

@@
 The spatial quality of these sources is often characterized in terms of their normalized transverse rms emittance [4,5]; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
-As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) [6], significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness [7]) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source – an intrinsic property of the emission source [4,5].
+As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) [6], significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness [7]) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source [4,5].
 
@@
 Additionally, and as expected from optical considerations, the modeling shows that the spatial beam size in the Fourier plane is independent of the incident laser spot size on the photocathode (i.e., the electron source size, $\Delta x_0$).
+
+To verify the efficacy of the experimental technique, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
+Due to the 60$^{\circ}$ angle of incidence, the circular laser beam is focused by the 30cm focal length lens to an elliptical 50x100$\mu$m (half-width 1/e maximum (HWe$^{-1}$M) of the field) spot size on the photocathode surface.
+The results, the measured Fourier plane beam size and electron yield, are displayed in Figure 2 as a function of the incident laser pulse energy, corrected to account for 10\% optical loss (mainly from the uncoated vacuum system windows).
+
+The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
+This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water [12] or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface [13].
+Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function [12] ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
+Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds [14].
+The solid line in Figure 2 is the predicted variation derived using a simple zero free parameter model of this effect.
+Using the known optical properties of gold [8], which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm [15].
+This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (Figure 1) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
+The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
+The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m_0$ [8].
+
+Figure 3 depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
+Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
+Prior studies [16] indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included [4].
+Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
+
+The known optical properties of the two semiconductors [17] indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb [18] and InSb [19], is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
+Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m_0$ (0.5$m_0$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV).
+For the $\sim$0.1nJ incident UV pulse energies, $\sim40\mu$m laser spot size, and $p$-polarized reflectance of $\sim$40\% for both semiconductors at our 60$^{\circ}$ incidence angle, the photo-injected electron density is $\sim10^{18}cm^{-3}$.
+At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization [20] and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
+
+Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above $\phi$, thereby curtailing the observed ESTE.
+By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band [21], we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
+This means that even without population decay mechanisms the temperature of the exited electron distribution $T_e$ drops at a rate of $\sim$1,600K/ps in both semiconductors.
+We note that the close proximity of the lower and unconfined (in momentum space) $\Gamma_7$ upper conduction band to the $\Gamma_8$ band will likely result in a fast population decay, so that the observed ESTE process should have an intrinsic latency of much less than the $\sim$4ps UV laser pulse duration.
+
+Figure 3 also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
+More significantly, using $\Delta p_x = \sqrt{m_0 k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m_0$ in the semiconductor), simulation of the experiment with the extended AG model [9] indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\phi/(k_B T_e)] \sim 10^{-15}$).
+On the other hand, employing a mass of 0.3$m_0$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
+In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (Figure 3).
+We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy [4,5]) irradiated at the same 261nm wavelength.

### 2012-10-31 inc/este/este.tex
finish importing body text from APL paper. still need figures+captions and references

@@
 
+% the content of this section is a modification of our APL paper on ESTE
 \section{Excited State Themionic Emission Photocathodes}
@@
 We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy [4,5]) irradiated at the same 261nm wavelength.
+
+Very similar results are observed with the $p$-type InSb photocathode except that the Fourier plane spot size is 30-40\%  larger and the electron yield is a factor of 1.7 higher.
+The increased rms transverse momentum for the same ESTE process is consistent with the larger $\sim$0.5$m_0$ effective mass of the upper $\Gamma_8$ conduction band in InSb, the expected higher initial temperature of the electron distribution, and the marginally slower cooling process associated with the lower 24meV LO phonon energy.
+The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation [17].
+
+The dependence of $\Delta p_x$ on $m^*$ is readily explained through a consideration of energy and momentum conservation in transmission across a boundary with a potential step associated with the effective photoelectric work function $\phi$.
+Inside the material, an electron with an energy $E$ greater than $\phi$ has a maximum momentum of $\sqrt{2 m^* (E-\phi) }$ parallel to the boundary if it is to be emitted.
+As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_x$ of the electron source clearly must scale with $\sqrt{m_*}$ as observed in the experiment.
+This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies [22] where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
+We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes [23], where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m_0$)  to allow for direct emission of electrons excited into the conduction band.
+For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_x = \sqrt{m^* ( \hbar \omega - \phi ) / 3 }$.
+Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$ps, 261nm laser radiation source.

### 2012-10-31 inc/este/este.tex
add apl bibliography up to entry 13

@@
 %TODO consider moving this paragraph to chapter intro 
-Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED) [1], ultrafast transmission electron microscopy (UEM) [2], and pulsed X-ray free electron lasers [3].
-The spatial quality of these sources is often characterized in terms of their normalized transverse rms emittance [4,5]; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
-As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) [6], significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness [7]) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source [4,5].
+Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED)\cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers\cite{dowell_cathode_2010}.
+The spatial quality of these sources is often characterized in terms of their normalized transverse rms emittance\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
+As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field)\cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness\cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source\cite{dowell_quantum_2009,jensen_emittance_2010}.
 
@@
 Direct observation of the transverse momentum distribution for excited-state thermionic emission (ESTE) from GaSb and InSb, two similar III-V zincblende semiconductors, demonstrates the $m^*$ dependence.
-A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m_0$ [8], confirms the temperature dependence of $\Delta p_x$.
+A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m_0$\cite{johnson_optical_1972}, confirms the temperature dependence of $\Delta p_x$.
 Simulations of the experiment using our extended analytical Gaussian (AG) model of electron pulse propagation (see Section \ref{sec:external_forces}) are in close agreement with the data.
@@
 Figure 1 depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
-The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator [10] delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate (see Section \ref{sec:laser}).
+The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate (see Section \ref{sec:laser}).
 This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV).
 Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
-Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface [7].
-The electron gun design is based on the that of Togawa et al. [7,11] and has an acceleration gap of 25($\pm$2)mm.
+Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface\cite{berger_dc_2009}.
+The electron gun design is based on the that of Togawa et al.\cite{berger_dc_2009,togawa_ceb6_2007} and has an acceleration gap of 25($\pm$2)mm.
 After acceleration in the DC photo-gun, the electrons pass through a set of electrostatic deflection plates to ensure that the electron beam passes, as close as possible, axially through the center of two, 6.35mm-bore, thin and round magnetic lenses with counter-propagating currents to minimize image rotation effects.
@@
 The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
-This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water [12] or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface [13].
-Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function [12] ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
+This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water\cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface\cite{eastman_photoelectric_1970}.
+Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function\cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
 Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds [14].
@@
 Figure 3 also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
-More significantly, using $\Delta p_x = \sqrt{m_0 k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m_0$ in the semiconductor), simulation of the experiment with the extended AG model [9] indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\phi/(k_B T_e)] \sim 10^{-15}$).
+More significantly, using $\Delta p_x = \sqrt{m_0 k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m_0$ in the semiconductor), simulation of the experiment with the extended AG model\cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\phi/(k_B T_e)] \sim 10^{-15}$).
 On the other hand, employing a mass of 0.3$m_0$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.

### 2012-10-31 inc/este/este.tex
add references through 15

@@
 Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function\cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
-Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds [14].
+Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds\cite{chen_semiclassical_2006}.
 The solid line in Figure 2 is the predicted variation derived using a simple zero free parameter model of this effect.
-Using the known optical properties of gold [8], which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm [15].
+Using the known optical properties of gold\cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm\cite{seah_quantitative_1979}.
 This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (Figure 1) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
 The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
-The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m_0$ [8].
+The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m_0$\cite{johnson_optical_1972}.
 
@@
 Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
-Prior studies [16] indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included [4].
+Prior studies [16] indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included\cite{dowell_quantum_2009}.
 Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
@@
 In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (Figure 3).
-We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy [4,5]) irradiated at the same 261nm wavelength.
+We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy\cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.
 

### 2012-10-31 inc/este/este.tex
finished importing apl references into bibeste

@@
 Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
-Prior studies [16] indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included\cite{dowell_quantum_2009}.
+Prior studies\cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included\cite{dowell_quantum_2009}.
 Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
 
-The known optical properties of the two semiconductors [17] indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb [18] and InSb [19], is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
+The known optical properties of the two semiconductors\cite{aspnes_dielectric_1983} indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb\cite{chelikowsky_nonlocal_1976} and InSb\cite{chelikowsky_erratum_1984}, is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
 Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m_0$ (0.5$m_0$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV).
 For the $\sim$0.1nJ incident UV pulse energies, $\sim40\mu$m laser spot size, and $p$-polarized reflectance of $\sim$40\% for both semiconductors at our 60$^{\circ}$ incidence angle, the photo-injected electron density is $\sim10^{18}cm^{-3}$.
-At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization [20] and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
+At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization\cite{portella_k-space_1992} and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
 
 Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above $\phi$, thereby curtailing the observed ESTE.
-By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band [21], we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
+By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band\cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
 This means that even without population decay mechanisms the temperature of the exited electron distribution $T_e$ drops at a rate of $\sim$1,600K/ps in both semiconductors.
@@
 The increased rms transverse momentum for the same ESTE process is consistent with the larger $\sim$0.5$m_0$ effective mass of the upper $\Gamma_8$ conduction band in InSb, the expected higher initial temperature of the electron distribution, and the marginally slower cooling process associated with the lower 24meV LO phonon energy.
-The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation [17].
+The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation\cite{aspnes_dielectric_1983}.
 
@@
 As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_x$ of the electron source clearly must scale with $\sqrt{m_*}$ as observed in the experiment.
-This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies [22] where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
-We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes [23], where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m_0$)  to allow for direct emission of electrons excited into the conduction band.
+This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies\cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
+We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes\cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m_0$)  to allow for direct emission of electrons excited into the conduction band.
 For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_x = \sqrt{m^* ( \hbar \omega - \phi ) / 3 }$.

### 2012-11-13 inc/este/este.tex
investigations towards using gnuplot/tikz figures
include figure 1 from APL paper, including data and gnuplot file to generate a TikZ plot
added caption from APL figure 1

@@
 
-Figure 1 depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
+\begin{figure}
+  \centering
+  \input{fig1}
+  \caption{
+    Simulations of the electron pulse propagation through the apparatus using the extended AG model.  
+    Solid line: The spatial beam width (normalized to that at the photocathode $\Delta x_0$) with the magnetic lens strengths adjusted to ensure that the detector is at their back focal plane (Fourier plane).
+    Dashed line: The case when the strength of the magnetic lenses is increased by a factor of two to focus the electron beam on the YAG scintillator detector.
+  }
+  \label{fig:este-sim}
+\end{figure}
+
+\ref{fig:este-sim} depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
 The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate (see Section \ref{sec:laser}).
@@
 This is readily achieved by appropriate adjustment of the current $I$ in the magnetic lens coils; specifically, after determining the current required to focus the electron beam onto the scintillator, optical imaging relations together with the $I^2$ dependence of the magnetic lens strength are used to find the current that ensures that the average lens to scintillator distance equals the focal length of the lens system.
-Simulation of the experimental apparatus (employing the extended AG model, see Section \ref{sec:external_forces} ), also shown in Figure 1, confirms that this procedure places the YAG scintillator at the back focal plane of the lens system to within a $\pm$0.5cm error.
+Simulation of the experimental apparatus (employing the extended AG model, see Section \ref{sec:external_forces} ), also shown in \ref{fig:este-sim}, confirms that this procedure places the YAG scintillator at the back focal plane of the lens system to within a $\pm$0.5cm error.
 This simulation also confirms that for our experimental conditions (less than 5,000 electrons/pulse and $\Delta x > $ 30$\mu$m) space-charge effects do not influence the observations.
@@
 Using the known optical properties of gold\cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm\cite{seah_quantitative_1979}.
-This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (Figure 1) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
+This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:este-sim}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
 The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.

### 2012-11-14 inc/este/este.tex
import APL figure 2

@@
 
+\begin{figure}
+  \centering
+  \input{fig2}
+  \caption{
+    Measured Fourier plane electron beam size (HWe$^{-1}$M; squares) and electron yield (circles) as a function of the incident $\sim$200fs, 523nm laser pulse energy for the 300nm-thick gold photocathode.
+    The electron yield is proportional to the square of laser pulse energy (solid line) and the dependence of the beam size with the laser pulse energy fits that predicted by a zero free parameter electron heating model (solid line; dashed lines represent $\pm$10\% error in the laser spot size).
+    A representative raw Fourier plane beam image is also shown.
+  }
+  \label{fig:gold-emission}
+\end{figure}
+
 To verify the efficacy of the experimental technique, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
 Due to the 60$^{\circ}$ angle of incidence, the circular laser beam is focused by the 30cm focal length lens to an elliptical 50x100$\mu$m (half-width 1/e maximum (HWe$^{-1}$M) of the field) spot size on the photocathode surface.
-The results, the measured Fourier plane beam size and electron yield, are displayed in Figure 2 as a function of the incident laser pulse energy, corrected to account for 10\% optical loss (mainly from the uncoated vacuum system windows).
+The results, the measured Fourier plane beam size and electron yield, are displayed in \ref{fig:gold-emission} as a function of the incident laser pulse energy, corrected to account for 10\% optical loss (mainly from the uncoated vacuum system windows).
 
@@
 Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds\cite{chen_semiclassical_2006}.
-The solid line in Figure 2 is the predicted variation derived using a simple zero free parameter model of this effect.
+The solid line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
 Using the known optical properties of gold\cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm\cite{seah_quantitative_1979}.

### 2012-11-19 inc/este/este.tex
include APL figure 3

@@
 
-Figure 3 depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
+\begin{figure}
+  \centering
+  \input{fig3}
+  \caption{
+    Number of electrons emitted per pulse as a function of the incident 261nm, UV laser pulse energy for GaSb (open squares) and InSb (filled circles): Linear efficiency dependences are shown by the solid lines.
+    The laser pulse energy invariance of the Fourier plane electron beam size (HWe$^{-1}$M right axis; open squares) for GaSb and a representative raw Fourier plane beam image is also shown.
+  }
+  \label{fig:este-semicond}
+\end{figure}
+
+\ref{fig:este-semicond} depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
 Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
@@
 
-Figure 3 also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
+\ref{fig:este-semicond} also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
 More significantly, using $\Delta p_x = \sqrt{m_0 k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m_0$ in the semiconductor), simulation of the experiment with the extended AG model\cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\phi/(k_B T_e)] \sim 10^{-15}$).
 On the other hand, employing a mass of 0.3$m_0$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
-In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (Figure 3).
+In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (\ref{fig:este-semicond}).
 We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy\cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.

### 2012-11-20 inc/laser/laser.tex
move more of apl paper to este
added several of the opt express graphics

@@
 
-\section{Yb:KGW Laser System}
+\section{Yb:KGW Laser System} \label{sec:laser}
 \subsection{Introduction}
@@
 %page 8632
-Instead of using a polarization-coupled pumping geometry \cite{Holtom_mode_locked_2006} to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in Fig.
-1(a), where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
-At an angle of incidence of 32.3$^\circ$, this coating theoretically has zero reflection loss for $p$-polarized 1040nm laser radiation parallel to the $b$-axis propagating perpendicular to the $a$-$b$ plane of the crystal (Fig.
-1(b)); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 \cite{Biswal_thermo_optical_05,pujol_crystalline_1999}.
+Instead of using a polarization-coupled pumping geometry \cite{Holtom_mode_locked_2006} to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in \ref{fig:crystal}\subref{fig:crystal_geo}, where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
+At an angle of incidence of 32.3$^\circ$, this coating theoretically has zero reflection loss for $p$-polarized 1040nm laser radiation parallel to the $b$-axis propagating perpendicular to the $a$-$b$ plane of the crystal (\ref{fig:crystal}\subref{fig:crystal_coating}); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 \cite{Biswal_thermo_optical_05,pujol_crystalline_1999}.
 The \ce{SiO2} coating also has an appropriate bandwidth for femtosecond laser operation in excess of 100nm (to the
@@
 On the other hand, the $\sim$ 0.4\% reflection loss per surface for $s$-polarized 1040nm radiation is sufficient to suppress $s$-polarized oscillation in a laser resonator.
-In addition, the same anti-reflection coating has less than 0.5\% reflection loss for $s$-polarized 980nm pump radiation (polarized parallel to the $a$-axis of the Yb:KGW crystal) at incidence angles between 21 and 39$^\circ$ (Fig.
-1(b)).
+In addition, the same anti-reflection coating has less than 0.5\% reflection loss for $s$-polarized 980nm pump radiation (polarized parallel to the $a$-axis of the Yb:KGW crystal) at incidence angles between 21 and 39$^\circ$ (\ref{fig:crystal}\subref{fig:crystal_coating}).
 Thus, in this Yb:KGW laser crystal geometry, the largest absorption cross-section associated with the optical $N_m$-axis is accessed with about 94\% efficiency ($\cos 20^\circ = 0.94$) by the $s$-polarized pump radiation.
@@
   }
+  \subfloat[][]{
+    \label{fig:crystal_coating}
+    \includegraphics{coating.jpg}
+  }
   \caption{
     \protect\subref{fig:crystal_geo} Top view of the 3x3x10mm AR coated Yb:KGW laser crystal showing the crystal cut, the direction of propagation of p-polarized laser radiation at 1040nm, and the $x$ (tangential) and $y$ (saggital) directions (in blue) employed in the thermal analysis;
-(b) Calculated reflectivity of the 193nm \ce{SiO2} AR coating as a function of incidence angle for $p$-polarized 1040nm laser radiation (solid line) and $s$-polarized 980nm pump radiation (dashed line).
+    \protect\subref{fig:crystal_coating} Calculated reflectivity of the 193nm \ce{SiO2} AR coating as a function of incidence angle for $p$-polarized 1040nm laser radiation (solid line) and $s$-polarized 980nm pump radiation (dashed line).
   }
+  \label{fig:crystal}
 \end{figure}
@@
 To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique \cite{Rimington_thermal_lens_2004} to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal.
-For the employed non-Brewster crystal cut (Fig. 1(a)), a ray
+For the employed non-Brewster crystal cut (\ref{fig:crystal}\subref{fig:crystal_geo}), a ray
 %page 8633
@@
 In our case, $l \approx$ 3mm and $W \approx$ 0.2mm, so that $\Delta < 0.1$, implying that the thermal duct due to a non-zero $\frac{dn}{dT}$ primarily determines the required $B_T$/$A_T$ ratio.
-In other words, for the potassium gadolinium tungstate (\ce{KGd(WO4)2}) crystal geometry of Fig.
-1(a), we expect from \ref{eq:laser_ratio_ba} to require $\dfrac{B_\smallT}{A_\smallT} \approx \dfrac{ \cos^2 \theta_2 }{ \cos^2 \theta_1 } = 1.3$.
+In other words, for the potassium gadolinium tungstate (\ce{KGd(WO4)2}) crystal geometry of \ref{fig:crystal}\subref{fig:crystal_geo}, we expect from \ref{eq:laser_ratio_ba} to require $\dfrac{B_\smallT}{A_\smallT} \approx \dfrac{ \cos^2 \theta_2 }{ \cos^2 \theta_1 } = 1.3$.
 
@@
 For a fixed vertical pump spot size of $W_y = 0.2$mm, our analysis indicates that in order to achieve the required ratio $B_T / A_T = 1.3$ in the parabolic approximation for $T(x,y)$ about $(x,y) = (0,0)$ a horizontal pump spot size $W_x = 0.3$mm is needed; in other words, the required $Q(x,y)$ heat source spot size ratio (or ellipticity) $W_x / W_y = 1.5$.
-This analytical result for the central region around the laser beam axis is shown in Fig. 2; displayed are the
+This analytical result for the central region around the laser beam axis is shown in \ref{fig:laser-thermal}; displayed are the
 %page 8634
-horizontal ($y = 0$, Fig. 2(a)) and vertical ($x = 0$, Fig. 2(b)) sections, respectively, through the $Q(x,y)$ and $T(x,y)$ distributions.
-Also shown are the parabolic approximations to the pump-induced temperature distribution (dashed lines) and representative TEM$_{00}$ Gaussian laser modes (shaded area) in the Yb:KGW crystal with a HW1/eM field mode sizes of $w_y = 160\mu$m and $w_x = 182 \mu$m (14\% larger due to refraction (Fig.
-1(a)).
-This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium \cite{Brenier_new_criteria}; in particular, in the vertical direction (Fig.
-2(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
+horizontal ($y = 0$, \ref{fig:laser-thermal}(a)) and vertical ($x = 0$, \ref{fig:laser-thermal}(b)) sections, respectively, through the $Q(x,y)$ and $T(x,y)$ distributions.
+Also shown are the parabolic approximations to the pump-induced temperature distribution (dashed lines) and representative TEM$_{00}$ Gaussian laser modes (shaded area) in the Yb:KGW crystal with a HW1/eM field mode sizes of $w_y = 160\mu$m and $w_x = 182 \mu$m (14\% larger due to refraction (\ref{fig:crystal}\subref{fig:crystal_geo}).
+This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium \cite{Brenier_new_criteria}; in particular, in the vertical direction (\ref{fig:laser-thermal}(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
 
-% insert figure 2
+\begin{figure}
+  \centering
+  \includegraphics{thermal.jpg}
+  \caption{
+    Analytical evaluation of the temperature distribution $T(x,y)$ due to the Gaussian/super-Gaussian heat source $Q(x,y)$ in the (a) horizontal ($y=0$) and (b) vertical ($x=0$) directions: $Q(x,y)$ - red-brown line; $T(x,y)$ - black line; parabolic approximation to $T(x,y)$ - dashed black line; representative TEM$_{00}$ laser field modes - solid green.
+  }
+  \label{fig:laser-thermal}
+\end{figure}
 
@@
 We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) \cite{Biswal_thermo_optical_05,mochalov_laser_1997} and stress-induced refractive index changes \cite{Yumashev_laser_2007}.
-For the employed crystal cut in our TLS geometry (Fig.
-1(a)), the anisotropy in $\kappa$ \cite{mochalov_laser_1997,website_EKSPLA} is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
+For the employed crystal cut in our TLS geometry (\ref{fig:crystal}\subref{fig:crystal_geo}), the anisotropy in $\kappa$ \cite{mochalov_laser_1997,website_EKSPLA} is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
 The refractive index
@@
 
-The diode-pumped TLS Yb:KGW laser head is placed in the 2.35m-long, asymmetric z-fold cavity shown in Fig. 3(a).
+The diode-pumped TLS Yb:KGW laser head is placed in the 2.35m-long, asymmetric z-fold cavity shown in \ref{fig:laser-cavity}\subref{fig:laser_schematic}.
 The two focusing gain section mirrors each have a group velocity dispersion (GVD) of $-1300 ( \pm 150 ) \text{fs}^2$ (Layertec GmbH \cite{website_layertec}), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
@@
   }
+  \\
+  \subfloat[][]{
+    \label{fig:laser_stability}
+    \includegraphics{stability.jpg}
+  }
   \caption{
     \protect\subref{fig:laser_schematic} Schematic of the femtosecond TLS Yb:KGW oscillator (LD = laser diode, OC = output coupler, SBR = saturable Bragg reflector).
-(b) Laser mode size in the center of the Yb:KGW gain medium ($w$) as a function of the strength of the lensing ($1/f$) in the gain medium (thermal and Kerr) obtained from a resonator stability analysis with $z \approx 42$cm for both the tangential ($x$, red line) and saggital ($y$, dashed black line) directions.
+    \protect\subref{fig:laser_stability} Laser mode size in the center of the Yb:KGW gain medium ($w$) as a function of the strength of the lensing ($1/f$) in the gain medium (thermal and Kerr) obtained from a resonator stability analysis with $z \approx 42$cm for both the tangential ($x$, red line) and saggital ($y$, dashed black line) directions.
     The 0.2mm (HW1/eM) vertical super-Gaussian pump spot size is indicated by the horizontal line.
   }
+  \label{fig:laser-cavity}
 \end{figure}
@@
 
-The stability analysis of our diode-pumped Yb:KGW laser (Fig.
-3(b)), clearly indicates that the resonator configuration is unstable for thermal focal lengths less than about $-6m^{-1}$ (diopters) and is unlikely to oscillate for negative thermal focal lengths as this requires a vertical TEM$_{00}$-mode size $w_y$ greater than 180$\mu$m in the Yb:KGW gain medium --- a value that would result in strong absorptive losses due to the restrictive $W_y = 200\mu$m super-Gaussian vertical pump beam size in the Yb-doped crystal.
+The stability analysis of our diode-pumped Yb:KGW laser (\ref{fig:laser-cavity}\subref{fig:laser_schematic}), clearly indicates that the resonator configuration is unstable for thermal focal lengths less than about $-6m^{-1}$ (diopters) and is unlikely to oscillate for negative thermal focal lengths as this requires a vertical TEM$_{00}$-mode size $w_y$ greater than 180$\mu$m in the Yb:KGW gain medium --- a value that would result in strong absorptive losses due to the restrictive $W_y = 200\mu$m super-Gaussian vertical pump beam size in the Yb-doped crystal.
 The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
@@
 
-In the cavity configuration depicted in Fig. 3(a), self-starting mode-locked (ML) operation was observed for output powers greater than $\sim$2W and $z \approx 42$cm.
-The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (Fig. 4(a)).
+In the cavity configuration depicted in \ref{fig:laser-cavity}\subref{fig:laser_schematic}, self-starting mode-locked (ML) operation was observed for output powers greater than $\sim$2W and $z \approx 42$cm.
+The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (\ref{fig:laser-profile}(a)).
 The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH \cite{website_layertec}) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
-For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (Fig.
-4(b)) at an optimum mode-locked output power of $5.0$W.
+For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (\ref{fig:laser-profile}(b)) at an optimum mode-locked output power of $5.0$W.
 This trend is consistent with the expected `soliton' modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity \cite{Brunner_diode_pumped}.
@@
 However, ML operation was significantly less stable under these conditions, indicating that either the cavity is operating close to the net zero GVD point or that the finite gain bandwidth is influencing the ML operation.
-We also note that in this case the optimum SBR position $z$ is larger than that for higher power 3.5W ML operation with a negative cavity GVD of $-5,200\text{fs}^2$ per round-trip (Fig.
-4(a)), but is always less than the 50cm focal length of the concave high reflector.
+We also note that in this case the optimum SBR position $z$ is larger than that for higher power 3.5W ML operation with a negative cavity GVD of $-5,200\text{fs}^2$ per round-trip \ref{fig:laser-profile}(a)), but is always less than the 50cm focal length of the concave high reflector.
 This is due to the need to match the absorption-induced thermal bowing of the SBR (i.e., the radius of curvature of its front surface mirror) to the radius of curvature of the intra-cavity laser mode \cite{Schieffer_dual_passive_2006}.
@@
 
-Figure 5 shows the laser's output power as a function of diode current for the cavity configuration depicted in Fig.
-3(a); a total negative GVD of $-5,200\text{fs}^2$ per round-trip and a 7\% transmission output coupler.
+\ref{fig:laser-performance} shows the laser's output power as a function of diode current for the cavity configuration depicted in \ref{fig:laser-cavity}\subref{fig:laser_schematic}; a total negative GVD of $-5,200\text{fs}^2$ per round-trip and a 7\% transmission output coupler.
 Soliton mode-locked (ML) operation is observed for diode currents greater than 23A, corresponding to a continuous-wave (cw) output power of 1.8W.
@@
 Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal's optical $N_p$-axis \cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
-From Fig.
-3(b), we see that the resulting increase in the strength of the lensing in the Yb:KGW crystal will result in a reduction of the laser mode size in the gain medium, which produces a better spatial match to gain duct produced by the laser diode pumps (less absorptive loss), thus generating an increase in output power.
+From \ref{fig:laser-cavity}\subref{fig:laser_stability}, we see that the resulting increase in the strength of the lensing in the Yb:KGW crystal will result in a reduction of the laser mode size in the gain medium, which produces a better spatial match to gain duct produced by the laser diode pumps (less absorptive loss), thus generating an increase in output power.
 
-% figure 4
+\begin{figure}
+  \centering
+  \includegraphics{profile.jpg}
+  \caption{
+Second harmonic autocorrelation measurements and output spectra of the ML Yb:KGW laser; (a) the 250fs laser pulse duration (and fit assuming a sech$^2$ pulse shape) for the cavity depicted in \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} at an output power of 3.5W, (b) the 347fs pulse duration (and fit assuming a sech$^2$ pulse shape) obtained with an additional -5600fs$^2$ of round-trip GVD at an output power of 5W, (c) the 4.7nm full-width half maximum spectrum associated with the 250fs ML pulse of (a), and (d) spectral output for the ML resonator of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} at a laser power of 3.7W.
+  }
+  \label{fig:laser-profile}
+\end{figure}
 
@@
 At a diode current around 29A, a maximum (or optimum) average mode-locked output power of $\sim$3.5W is obtained.
-The 250fs pulses generated at this output power (Fig.
-4(a)) correspond to a Kerr lens strength in the Yb:KGW crystal of $\sim4 \text{m}^{-1}$, which is readily accommodated by the resonator design (Fig.
-3(b)).
+The 250fs pulses generated at this output power (\ref{fig:laser-profile}(a)) correspond to a Kerr lens strength in the Yb:KGW crystal of $\sim4 \text{m}^{-1}$, which is readily accommodated by the resonator design (\ref{fig:laser-cavity}\subref{fig:laser_stability}).
 For this $\sim$3.5W ML laser with a negative cavity GVD of $-5,200\text{fs}^2$ per round-trip, the full-width at
 %page 8638
-half maximum of the mode-locked spectrum centered at 1046nm is 4.7nm (Fig. 4(c)).
+half maximum of the mode-locked spectrum centered at 1046nm is 4.7nm (\ref{fig:laser-profile}(c)).
 The laser is therefore generates pulses with a time-bandwidth product very close to the Fourier limit of 0.32 for a sech$^2$ pulse shape.
-At output powers greater than 3.5W, a DC spectral component emerges at the short wavelength end of the ML spectrum (Fig. 4(d)) and there is a slow roll-over in the laser output power (Fig. 5).
+At output powers greater than 3.5W, a DC spectral component emerges at the short wavelength end of the ML spectrum (\ref{fig:laser-profile}(d)) and there is a slow roll-over in the laser output power (\ref{fig:laser-performance}).
 This trend is expected for a Yb:KGW soliton ML laser; in this case, the soliton power limit being reached at about 29A, and the asymmetric spectral gain of Yb:KGW forcing excess CW laser power oscillation at wavelengths shorter than the peak of the ML spectrum.
-Similar laser performance is observed with the insertion of an additional $-5,600\text{fs}^2$ of cavity GVD per round trip, which generates 347fs sech$^2$ pulses (Fig.
-4(b)); that is, a DC spectral component appears on the short wavelength side of a narrower $\sim$3.5nm ML spectrum at output powers above 5W.
+Similar laser performance is observed with the insertion of an additional $-5,600\text{fs}^2$ of cavity GVD per round trip, which generates 347fs sech$^2$ pulses (\ref{fig:laser-profile}(b)); that is, a DC spectral component appears on the short wavelength side of a narrower $\sim$3.5nm ML spectrum at output powers above 5W.
 
-% figure 5
+\begin{figure}
+  \centering
+  \includegraphics{performance.jpg}
+  \caption{
+Performance of the diode-pumped, TLS, Yb:KGW laser (in the cavity configuration of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} ) as a function of applied diode current: continuous-wave (CW) output power (black squares); mode locked (ML) output power (red circles); frequency doubled power (green triangles); and second harmonic generation (SHG) efficiency (line).
+  }
+  \label{fig:laser-performance}
+\end{figure}
 
@@
 
-Figure 5 shows the resultant frequency-doubled green power for the ML Yb:KGW laser.
+\ref{fig:laser-performance} shows the resultant frequency-doubled green power for the ML Yb:KGW laser.
 The measured average power at 523nm increases rapidly from 0.5W at the 23A mode-locking threshold to over 1.3W at a diode drive current of 27A.
 At and above 30A, the second harmonic power saturates at 1.65W --- an effect that is clearly consistent with the observed emergence of a DC spectral component at diode currents above 29A since only the ML pulse will be efficiently frequency doubled.
-Also plotted in Fig.
-5 is the frequency doubling efficiency, which is calculated taking into account the 17\% loss suffered by the $s$-polarized 523nm radiation upon exiting the Brewster-cut LBO crystal and the 4\% loss (to diagnostics) of the fundamental radiation prior to the $z$-fold doubling geometry.
+Also plotted in \ref{fig:laser-performance} is the frequency doubling efficiency, which is calculated taking into account the 17\% loss suffered by the $s$-polarized 523nm radiation upon exiting the Brewster-cut LBO crystal and the 4\% loss (to diagnostics) of the fundamental radiation prior to the $z$-fold doubling geometry.
 The maximum doubling efficiency of over 60\% occurs at 27-28A of diode drive current when the ML laser output power is $\sim$3W.

### 2012-12-04 inc/photocathode/plasmon/plasmon.tex
move photocathode topics into subdirectory

--- 2012-09-12 inc/plasmon/plasmon.tex
@@
 
-\section{Introduction}
+\section{Nanopatterned (Plasmonic) Photocathodes}
+
+\subsection{Introduction}
 
@@
 
-\subsection{Driving the Surface Plasmon}
+%\subsection{Driving the Surface Plasmon}
 
@@
 \begin{equation} \label{eq:plasmon_condition}
-  \frac{ \varepsilon }{ 1 + \varepsilon } = \sin \theta + \frac{ n \lambda }{ d }
+  \sqrt{ \frac{ \varepsilon }{ 1 + \varepsilon } } = \sin \theta + \frac{ n \lambda }{ d }
 \end{equation}
@@
 
-\section{Experimental Attempts}
+\subsection{Experimental Attempts}
 
-\subsection{Commerical Grating on Glass}
+\subsubsection{Commerical Grating on Glass}
 
-For proof-of-concept purposes the first experiment was run on a commercially available gold foil on glass holographic grating.
+For proof-of-concept purposes the first experiment was run on a commercially available gold foil on glass holographic grating (750 lines/mm).
 Like the Kretschmann geometry, a foil on glass grating is vulnerable to heating problems; a plasmon creates a large amount of heat but the glass is a poor heat conductor.
-The grating was operational as the angles were tuned attempting to drive the plasmon.
+The aforementioned laser (emitting $\sim$200fs pulses of green light at 30MW/cm$^2$) was used to drive the plasmon.
+
+The input laser angle was slowly changed until the plasmon oscillation is visible.
 At the instant that the plasmon started, a bright flash was visible on the scintillator screen then disappeared.
 
-After this, the grating showed signs of damage.
+After this, the grating showed visual signs of damage, both on the surface and in the reflected light pattern.
 \ref{fig:grating-damage} shows the results of Scanning Electron Microscopy (SEM) on the grating.
@@
 
-The most important result was that a plasmon was driven successfully; this created the heat that destroyed the grating.
-Further this experiment validated the decision not to persue a Kretschmann geometry for fear of heat damage at these high powers.
+Though this conclusively shows that a plasmon was driven successfully, the heat generated by the plasmon oscillation destroyed the grating.
+Further this experiment validated the decision not to pursue a Kretschmann geometry for fear of heat damage at these high powers.
 
@@
 
-\subsection{Sinusoidal Grating}
+\subsubsection{Sinusoidal Grating}
 
-\subsection{Sinusoidal Grating, Rotated}
+After the destruction of the gold foil on glass grating, new gratings were designed to have a higher thermal conductivity.
+The second generation grating was constructed from a silicon substract then coated with a gold film.
+The thermal conductivity of silicon is approximately 100 times that of glass. %TODO reference needed?
 
-\subsection{Trenched Grating}
+Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division used a defocused Zeiss 1540XB focused ion beam (FIB) to iteratively mill an approximately sinusoidal shape into the surface of a silicon wafer.
+Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
+An SEM image of one such region is shown in \ref{fig:fib-si-sem}.
+
+Using the Varian e-beam deposition chamber at the UIC Nano Core Facility (NCF), I deposited a thin layer of chromium as a binding agent, then a 300nm film of gold.
+This thickness of gold is several times thicker than the skin depth of green light on gold, therefore the laser radiation cannot interact with the silicon.
+
+\begin{figure}
+  \centering
+  \includegraphics{HighMagSEM.jpg}
+  \caption{
+    SEM image of sinusoidal grating on silicon.
+    Pattern milled by Focused Ion Beam (FIB) milling at Argonne National Lab (ANL).
+    Sample was later coated with 300 nm gold film.
+  }
+  \label{fig:fib-si-sem}
+\end{figure}
+
+\subsubsection{Sinusoidal Grating, Rotated}
+
+\subsubsection{Trenched Grating}

### 2012-12-04 inc/photocathode/este/este.tex
moved experimental setup information from este paper to beginning of photocathode engineering section

--- 2012-11-19 inc/este/este.tex
@@
 
-%TODO consider moving this paragraph to chapter intro 
-Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED)\cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers\cite{dowell_cathode_2010}.
-The spatial quality of these sources is often characterized in terms of their normalized transverse rms emittance\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
-As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field)\cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness\cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source\cite{dowell_quantum_2009,jensen_emittance_2010}.
-
 In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_x$ developed in Refs.
@@
 Simulations of the experiment using our extended analytical Gaussian (AG) model of electron pulse propagation (see Section \ref{sec:external_forces}) are in close agreement with the data.
-
-\begin{figure}
-  \centering
-  \input{fig1}
-  \caption{
-    Simulations of the electron pulse propagation through the apparatus using the extended AG model.  
-    Solid line: The spatial beam width (normalized to that at the photocathode $\Delta x_0$) with the magnetic lens strengths adjusted to ensure that the detector is at their back focal plane (Fourier plane).
-    Dashed line: The case when the strength of the magnetic lenses is increased by a factor of two to focus the electron beam on the YAG scintillator detector.
-  }
-  \label{fig:este-sim}
-\end{figure}
-
-\ref{fig:este-sim} depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
-The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate (see Section \ref{sec:laser}).
-This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV).
-Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
-Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface\cite{berger_dc_2009}.
-The electron gun design is based on the that of Togawa et al.\cite{berger_dc_2009,togawa_ceb6_2007} and has an acceleration gap of 25($\pm$2)mm.
-After acceleration in the DC photo-gun, the electrons pass through a set of electrostatic deflection plates to ensure that the electron beam passes, as close as possible, axially through the center of two, 6.35mm-bore, thin and round magnetic lenses with counter-propagating currents to minimize image rotation effects.
-A YAG scintillator placed 27.5cm after the second magnetic lens detects the spatial electron beam profile, and its visible fluorescence is 1:1 imaged onto a CCD detector with 5.4$\mu$m pixels.
-
-To ensure that the transverse momentum distribution is observed, the YAG scintillator must be positioned at the back focal plane (the Fourier plane) of the lens system.
-This is readily achieved by appropriate adjustment of the current $I$ in the magnetic lens coils; specifically, after determining the current required to focus the electron beam onto the scintillator, optical imaging relations together with the $I^2$ dependence of the magnetic lens strength are used to find the current that ensures that the average lens to scintillator distance equals the focal length of the lens system.
-Simulation of the experimental apparatus (employing the extended AG model, see Section \ref{sec:external_forces} ), also shown in \ref{fig:este-sim}, confirms that this procedure places the YAG scintillator at the back focal plane of the lens system to within a $\pm$0.5cm error.
-This simulation also confirms that for our experimental conditions (less than 5,000 electrons/pulse and $\Delta x > $ 30$\mu$m) space-charge effects do not influence the observations.
-Additionally, and as expected from optical considerations, the modeling shows that the spatial beam size in the Fourier plane is independent of the incident laser spot size on the photocathode (i.e., the electron source size, $\Delta x_0$).
 
@@
 
-To verify the efficacy of the experimental technique, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
+To verify the efficacy of the experimental technique (see \ref{sec:photocathode-method}) for these purposes, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
 Due to the 60$^{\circ}$ angle of incidence, the circular laser beam is focused by the 30cm focal length lens to an elliptical 50x100$\mu$m (half-width 1/e maximum (HWe$^{-1}$M) of the field) spot size on the photocathode surface.
@@
 Using the known optical properties of gold\cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm\cite{seah_quantitative_1979}.
-This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:este-sim}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
+This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
 The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.

### 2012-12-04 inc/photocathode/setup/setup.tex
moved experimental setup information from este paper to beginning of photocathode engineering section

--- 2012-11-13 inc/este/este.tex
@@
 
-% the content of this section is a modification of our APL paper on ESTE
-\section{Excited State Themionic Emission Photocathodes}
-
-%TODO consider moving this paragraph to chapter intro 
-Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED)\cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers\cite{dowell_cathode_2010}.
-The spatial quality of these sources is often characterized in terms of their normalized transverse rms emittance\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
+%Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED)\cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers\cite{dowell_cathode_2010}.
+The spatial quality of these sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
 As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field)\cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness\cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source\cite{dowell_quantum_2009,jensen_emittance_2010}.
 
-In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_x$ developed in Refs.
-4 and 5 for thermionic electron emission should be rewritten to read $\sqrt{m^* k_B T_e}$, where $T_e$ is the temperature of the emitting electron distribution and $k_B$ is Boltzmann's constant.
-Direct observation of the transverse momentum distribution for excited-state thermionic emission (ESTE) from GaSb and InSb, two similar III-V zincblende semiconductors, demonstrates the $m^*$ dependence.
-A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m_0$\cite{johnson_optical_1972}, confirms the temperature dependence of $\Delta p_x$.
-Simulations of the experiment using our extended analytical Gaussian (AG) model of electron pulse propagation (see Section \ref{sec:external_forces}) are in close agreement with the data.
+\section{Experimental Methodology} \label{sec:photocathode-method}
 
@@
   }
-  \label{fig:este-sim}
+  \label{fig:transverse-measurement}
 \end{figure}
 
-\ref{fig:este-sim} depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
+\ref{fig:transverse-measurement} depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
 The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate (see Section \ref{sec:laser}).
@@
 This is readily achieved by appropriate adjustment of the current $I$ in the magnetic lens coils; specifically, after determining the current required to focus the electron beam onto the scintillator, optical imaging relations together with the $I^2$ dependence of the magnetic lens strength are used to find the current that ensures that the average lens to scintillator distance equals the focal length of the lens system.
-Simulation of the experimental apparatus (employing the extended AG model, see Section \ref{sec:external_forces} ), also shown in \ref{fig:este-sim}, confirms that this procedure places the YAG scintillator at the back focal plane of the lens system to within a $\pm$0.5cm error.
+Simulation of the experimental apparatus (employing the extended AG model, see Section \ref{sec:external_forces} ), also shown in \ref{fig:transverse-measurement}, confirms that this procedure places the YAG scintillator at the back focal plane of the lens system to within a $\pm$0.5cm error.
 This simulation also confirms that for our experimental conditions (less than 5,000 electrons/pulse and $\Delta x > $ 30$\mu$m) space-charge effects do not influence the observations.
 Additionally, and as expected from optical considerations, the modeling shows that the spatial beam size in the Fourier plane is independent of the incident laser spot size on the photocathode (i.e., the electron source size, $\Delta x_0$).
-
-To verify the efficacy of the experimental technique, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
-Due to the 60$^{\circ}$ angle of incidence, the circular laser beam is focused by the 30cm focal length lens to an elliptical 50x100$\mu$m (half-width 1/e maximum (HWe$^{-1}$M) of the field) spot size on the photocathode surface.
-The results, the measured Fourier plane beam size and electron yield, are displayed in Figure 2 as a function of the incident laser pulse energy, corrected to account for 10\% optical loss (mainly from the uncoated vacuum system windows).
-
-The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
-This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water\cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface\cite{eastman_photoelectric_1970}.
-Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function\cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
-Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds\cite{chen_semiclassical_2006}.
-The solid line in Figure 2 is the predicted variation derived using a simple zero free parameter model of this effect.
-Using the known optical properties of gold\cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm\cite{seah_quantitative_1979}.
-This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:este-sim}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
-The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
-The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m_0$\cite{johnson_optical_1972}.
-
-Figure 3 depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
-Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
-Prior studies\cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included\cite{dowell_quantum_2009}.
-Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
-
-The known optical properties of the two semiconductors\cite{aspnes_dielectric_1983} indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb\cite{chelikowsky_nonlocal_1976} and InSb\cite{chelikowsky_erratum_1984}, is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
-Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m_0$ (0.5$m_0$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV).
-For the $\sim$0.1nJ incident UV pulse energies, $\sim40\mu$m laser spot size, and $p$-polarized reflectance of $\sim$40\% for both semiconductors at our 60$^{\circ}$ incidence angle, the photo-injected electron density is $\sim10^{18}cm^{-3}$.
-At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization\cite{portella_k-space_1992} and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
-
-Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above $\phi$, thereby curtailing the observed ESTE.
-By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band\cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
-This means that even without population decay mechanisms the temperature of the exited electron distribution $T_e$ drops at a rate of $\sim$1,600K/ps in both semiconductors.
-We note that the close proximity of the lower and unconfined (in momentum space) $\Gamma_7$ upper conduction band to the $\Gamma_8$ band will likely result in a fast population decay, so that the observed ESTE process should have an intrinsic latency of much less than the $\sim$4ps UV laser pulse duration.
-
-Figure 3 also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
-More significantly, using $\Delta p_x = \sqrt{m_0 k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m_0$ in the semiconductor), simulation of the experiment with the extended AG model\cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\phi/(k_B T_e)] \sim 10^{-15}$).
-On the other hand, employing a mass of 0.3$m_0$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
-In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (Figure 3).
-We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy\cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.
-
-Very similar results are observed with the $p$-type InSb photocathode except that the Fourier plane spot size is 30-40\%  larger and the electron yield is a factor of 1.7 higher.
-The increased rms transverse momentum for the same ESTE process is consistent with the larger $\sim$0.5$m_0$ effective mass of the upper $\Gamma_8$ conduction band in InSb, the expected higher initial temperature of the electron distribution, and the marginally slower cooling process associated with the lower 24meV LO phonon energy.
-The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation\cite{aspnes_dielectric_1983}.
-
-The dependence of $\Delta p_x$ on $m^*$ is readily explained through a consideration of energy and momentum conservation in transmission across a boundary with a potential step associated with the effective photoelectric work function $\phi$.
-Inside the material, an electron with an energy $E$ greater than $\phi$ has a maximum momentum of $\sqrt{2 m^* (E-\phi) }$ parallel to the boundary if it is to be emitted.
-As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_x$ of the electron source clearly must scale with $\sqrt{m_*}$ as observed in the experiment.
-This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies\cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
-We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes\cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m_0$)  to allow for direct emission of electrons excited into the conduction band.
-For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_x = \sqrt{m^* ( \hbar \omega - \phi ) / 3 }$.
-Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$ps, 261nm laser radiation source.

### 2012-12-04 inc/photocathode/este/este.tex
remove hardcoding of certain references from este section

@@
 In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_x$ developed in Refs.
-4 and 5 for thermionic electron emission should be rewritten to read $\sqrt{m^* k_B T_e}$, where $T_e$ is the temperature of the emitting electron distribution and $k_B$ is Boltzmann's constant.
+\cite{dowell_quantum_2009} and \cite{jensen_emittance_2010} for thermionic electron emission should be rewritten to read $\sqrt{m^* k_B T_e}$, where $T_e$ is the temperature of the emitting electron distribution and $k_B$ is Boltzmann's constant.
 Direct observation of the transverse momentum distribution for excited-state thermionic emission (ESTE) from GaSb and InSb, two similar III-V zincblende semiconductors, demonstrates the $m^*$ dependence.

### 2012-12-04 inc/photocathode/este/este.tex
fix some nits

@@
 
-To verify the efficacy of the experimental technique (see \ref{sec:photocathode-method}) for these purposes, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
+To verify the efficacy of the experimental technique (see Section \ref{sec:photocathode-method}) for these purposes, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
 Due to the 60$^{\circ}$ angle of incidence, the circular laser beam is focused by the 30cm focal length lens to an elliptical 50x100$\mu$m (half-width 1/e maximum (HWe$^{-1}$M) of the field) spot size on the photocathode surface.

### 2012-12-04 inc/photocathode/setup/setup.tex
fix some nits

@@
 
+% the content of this section is a modification of our APL paper on ESTE
+
 %Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED)\cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers\cite{dowell_cathode_2010}.
-The spatial quality of these sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
+The spatial quality of the sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
 As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field)\cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness\cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source\cite{dowell_quantum_2009,jensen_emittance_2010}.
@@
 The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate (see Section \ref{sec:laser}).
-This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV).
+This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV) (see Section \ref{sec:laser}).
 Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
@@
 A YAG scintillator placed 27.5cm after the second magnetic lens detects the spatial electron beam profile, and its visible fluorescence is 1:1 imaged onto a CCD detector with 5.4$\mu$m pixels. 
+%TODO internal cross-references to appropriate sections above
 

### 2012-12-04 inc/photocathode/plasmon/plasmon.tex
include PAPE data and PAPE explaination (that from 2010 ANL report)

@@
 \section{Nanopatterned (Plasmonic) Photocathodes}
-
-\subsection{Introduction}
 
@@
 
-\subsection{Experimental Attempts}
-
-\subsubsection{Commerical Grating on Glass}
+\subsection{Commerical Grating on Glass}
 
@@
 
-\subsubsection{Sinusoidal Grating}
+\subsection{Sinusoidal Grating}
 
@@
 
-Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division used a defocused Zeiss 1540XB focused ion beam (FIB) to iteratively mill an approximately sinusoidal shape into the surface of a silicon wafer.
+Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division used a defocused Gallium focused ion beam (FIB) (Zeiss 1540XB) to iteratively mill an approximately sinusoidal shape into the surface of a silicon wafer.
 Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
+Given a laser wavelength of 523nm ($\varepsilon_{Au} \approx –3.95$ \cite{johnson_optical_1972}), this choice of periodicity corresponds by \ref{eq:plasmon_condition} to an incidence angle of $\sim39^{\circ}$, which is within the acceptance range of our experimental setup (35-75$^{\circ}$).
+%TODO point to Togawa section
 An SEM image of one such region is shown in \ref{fig:fib-si-sem}.
+
+\begin{figure}
+  \centering
+  \includegraphics{HighMagSEM.jpg}
+  \caption{
+    SEM image of sinusoidal grating on silicon.
+    Pattern milled by focused ion beam milling at Argonne National Lab.
+    Sample was later coated with 300 nm gold film.
+  }
+  \label{fig:fib-si-sem}
+\end{figure}
 
@@
   \centering
-  \includegraphics{HighMagSEM.jpg}
-  \caption{
-    SEM image of sinusoidal grating on silicon.
-    Pattern milled by Focused Ion Beam (FIB) milling at Argonne National Lab (ANL).
-    Sample was later coated with 300 nm gold film.
-  }
-  \label{fig:fib-si-sem}
+  \input{power_seq}
+  \caption{}
+  \label{fig:pape}
 \end{figure}
 
-\subsubsection{Sinusoidal Grating, Rotated}
+% Content taken from 2010 ANL report
+Strong enhanced phototemission was observed for only a narrow $\pm$5mrad range of incidence angles $\theta$ around the expected plasmon resonance angle of 39$^{\circ}$.
+Moreover, this enhanced laser-driven electron emission was only observed for $p$-polarized incident 523nm femtosecond laser radiation.
+These two properties are known characteristics of plasmon-assisted photoemission.
 
-\subsubsection{Trenched Grating}
+As seen in \ref{fig:pape}, at low incident laser intensities, below about 25MW/cm$^2$, the emission from the nano-patterned photocathode at the plasmon resonance condition is very similar to that observed from a flat gold surface neighboring the FIB-milled region.
+Specifically, the strength of both emissions have a quadratic dependence on the incident laser intensity.
+This is consistent with a two-photon field emission process --- it requires the absorption of two 2.37eV (green) photons for an electron to acquire sufficient energy to allow any significant field emission as gold has a work function of 5.1eV.
+Moreover, both emissions have similar momentum distributions \fixme{(images A and E)}, again indicating a similar emission process.
+The only difference is that the emission from the nano-patterned region is 3-4 times more efficient.
+This is readily explained by the fact that the periodic surface modulation results in a local enhancement of the $\sim$1MV/m DC gun field at the peaks of the modulation which, through the Schottky effect, then presents a reduced barrier for field emission.
+
+Above $\sim$32MW/cm$^2$, there is a rapid and dramatic increase in the emission efficiency for the nano-patterned photocathode.
+This can be explained by a suppression of the barrier to photoemission below the two-photon excitation energy, thus allowing direct two-photon photoemission.
+This is likely caused by the laser-excited surface plasmon since a similar increase from the neighboring flat gold surface is not present.
+The observed distinct changes in the momentum distributions \fixme{(images B and C)} in this laser intensity range support this argument.
+In \fixme{image B}, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves).
+This broadening becomes the dominant feature of the electron emission above $\sim$32MW/cm$^2$; indeed, it becomes so severe that at laser intensities above $\sim$37MW/cm$^2$ the 6mm-diameter bore of the magnetic lenses becomes a strong limiting aperture for the resulting beam --- as is clearly seen in \fixme{image D} and by the cut-off in the detected electron signal (YAG fluorescence).
+
+\subsection{Sinusoidal Grating, Rotated}
+
+\subsection{Trenched Grating}

### 2012-12-05 inc/photocathode/plasmon/plasmon.tex
added diagram describing the rotated geometry
more work on the plasmonic emission discussion

@@
   \input{power_seq}
-  \caption{}
+  \caption{
+    Photoemission characteristics of laser-driven plasmons on a one-dimensional periodic gold surface.
+    Top: Electron emission strength as a function of incident 523nm femtosecond laser pulse intensity for coherently-driven plasmons (top curve)
+      % with parabolic fit to initial dependence (dashed curve)
+      and comparative emission from a plane gold surface under the same irradiantion conditions.
+    Bottom: Two-dimensional electron momentum distributions for plasmon-assisted emission (A-D) and two-photon field emission from gold (E) at the laser intensities indicated on the top graph.
+    The circular aperture visible in B-D is the 6mm-diameter bore of the magnetic focusing lens.
+    The orientation of the grating lines of the periodic gold surface with respect to the emission in A-D is parallel to the broadening seen in B-D.
+  }
   \label{fig:pape}
@@
 % Content taken from 2010 ANL report
+Note that for this measurement, relative emission efficiencies are calculated from the integration of the luminosity detected on the YAG scintillator screen.
+The scintillator and CCD camera detection scheme was also shown to be linear (i.e., integrated signal proportional to number of electrons per pulse) for at least $10^4$ electrons per pulse, using the known linearity of single-photon photoemission from Ta irradiated by our 261nm ultraviolet laser source.
+As a result, the detection scheme will accurately determine any laser power law dependence to the electron emission from the photocathode.
+
 Strong enhanced phototemission was observed for only a narrow $\pm$5mrad range of incidence angles $\theta$ around the expected plasmon resonance angle of 39$^{\circ}$.
@@
 This is consistent with a two-photon field emission process --- it requires the absorption of two 2.37eV (green) photons for an electron to acquire sufficient energy to allow any significant field emission as gold has a work function of 5.1eV.
-Moreover, both emissions have similar momentum distributions \fixme{(images A and E)}, again indicating a similar emission process.
+Moreover, both emissions have similar momentum distributions (\ref{fig:pape} A and E), again indicating a similar emission process.
 The only difference is that the emission from the nano-patterned region is 3-4 times more efficient.
@@
 This is likely caused by the laser-excited surface plasmon since a similar increase from the neighboring flat gold surface is not present.
-The observed distinct changes in the momentum distributions \fixme{(images B and C)} in this laser intensity range support this argument.
-In \fixme{image B}, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves).
-This broadening becomes the dominant feature of the electron emission above $\sim$32MW/cm$^2$; indeed, it becomes so severe that at laser intensities above $\sim$37MW/cm$^2$ the 6mm-diameter bore of the magnetic lenses becomes a strong limiting aperture for the resulting beam --- as is clearly seen in \fixme{image D} and by the cut-off in the detected electron signal (YAG fluorescence).
+The observed distinct changes in the momentum distributions (\ref{fig:pape} B and C) in this laser intensity range support this argument.
+In \ref{fig:pape} B, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves).
+This broadening becomes the dominant feature of the electron emission above $\sim$32MW/cm$^2$; indeed, it becomes so severe that at laser intensities above $\sim$37MW/cm$^2$ the 6mm-diameter bore of the magnetic lenses becomes a strong limiting aperture for the resulting beam --- as is clearly seen in \ref{fig:pape} D and by the cut-off in the detected electron signal (YAG fluorescence).
 
@@
 
+In principle this unexpected increase in transverse momentum could be caused by either of two effects.
+The first is the divergence of the plasmon's electric field ($\partial E_{plasmon} / \partial x$) parallel to the surface plasmon vector $k_{SP}$ (assumed here to be in the $x$ direction).
+The second is the one dimensional local field enhancement of the accelerating electric field due to the peaks in the sinusoidal shape of the surface (parallel to the grating vector $k_G$); this is commonly known as the ``lightning rod'' effect.
+
+Given the geometry of the previous measurement, notably the colinear arrangement of plasmon vector and the grating vector, the true cause is difficult to determine.
+However, if instead the grating is rotated --- and the laser incidence angle is corrected to account for the longer effective periodicity of the surface --- the directionality of these effects may be separated.
+
+\begin{figure}
+  \centering
+  \input{rotated}
+  \caption{
+    Comparison of the colinear (above) and rotated (below) plasmonic excitation geometries.
+    The vector diagrams (left) are related to the corresponding momentum distribution image (right).
+    That the distribution is shown to be rotated by the same $45^{\circ}$ as the grating (the original colinear line is shown dashed) is evidence that the grating is the cause of the increased momentum distribution.
+  }
+  \label{fig:rotated}
+\end{figure}
+
+\ref{fig:rotated} shows the comparison of the grating vector geometries for the original scheme (top) and one in which the grating is rotated by $45^{\circ}$ (bottom).
+For this rotation, the laser angle must be increased from $39^{\circ}$ to $47^{\circ}$.
+
+The results are shown to the right in \ref{fig:rotated}.
+As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
+
+This result is commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
+As that geometry does not employ a patterned surface, it would is not subject to such local field enhancements of the accelerating electric field.
+
 \subsection{Trenched Grating}

### 2012-12-05 inc/photocathode/plasmon/plasmon.tex
fix a couple nits

@@
 
-Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division used a defocused Gallium focused ion beam (FIB) (Zeiss 1540XB) to iteratively mill an approximately sinusoidal shape into the surface of a silicon wafer.
+Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division employed a defocused Gallium focused ion beam (FIB) (Zeiss 1540XB) to iteratively mill an approximately sinusoidal shape into the surface of a silicon wafer.
 Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
@@
 This thickness of gold is several times thicker than the skin depth of green light on gold, therefore the laser radiation cannot interact with the silicon.
+%TODO should I comment on the coating acting to smooth the surface to a proper sinusoid?
 

### 2012-12-11 inc/photocathode/plasmon/plasmon.tex
include discussion of trenched gratings and conclusion of nano-patterned sources

@@
 
-In principle this unexpected increase in transverse momentum could be caused by either of two effects.
-The first is the divergence of the plasmon's electric field ($\partial E_{plasmon} / \partial x$) parallel to the surface plasmon vector $k_{SP}$ (assumed here to be in the $x$ direction).
-The second is the one dimensional local field enhancement of the accelerating electric field due to the peaks in the sinusoidal shape of the surface (parallel to the grating vector $k_G$); this is commonly known as the ``lightning rod'' effect.
+In principle this unexpected increase in transverse momentum could be caused by any of three effects.
+The first, and most easily ruled out, is that the electric field of the laser itself could affect the electrons, causing the divergence.
+The possibility is easily dismissed as the effect is not observed when off the grating.
+The second possible cause is the divergence of the plasmon's own electric field ($\partial E_{plasmon} / \partial x$).
+This field is parallel to the surface plasmon vector $k_{SP}$ (assumed here to be in the $x$ direction).
+The third possible cause is local field enhancement of the accelerating electric field due to the peaks in the sinusoidal shape of the surface, commonly known as the ``lightning rod'' effect.
+For the employed one dimensional grating, this effect is necessarily parallel to the grating vector $k_G$.
 
@@
     Comparison of the colinear (above) and rotated (below) plasmonic excitation geometries.
-    The vector diagrams (left) are related to the corresponding momentum distribution image (right).
-    That the distribution is shown to be rotated by the same $45^{\circ}$ as the grating (the original colinear line is shown dashed) is evidence that the grating is the cause of the increased momentum distribution.
+    This rotation separates the directionality of effects caused by the laser, plasmon field and grating.
+    Each vector diagrams (left) is shown next to a representative observed momentum distribution image (right).
+    The increased divergence is shown to be rotated by the same $45^{\circ}$ as the grating (the original colinear line is shown dashed); evidence that the grating is the cause of the deleterious effect.
   }
@@
 \subsection{Trenched Grating}
+
+In an attempt to mitigate the effect of the surface peaks on the accelerating field, a new pattern was commissioned and prepared in a similar manner to the original.
+This new photocathode featured long trenches of an approximately square profile, arranged periodically across the surface.
+This periodicity should have been sufficient to facilitate the generation of the plasmon.
+Although the structure should be less disruptive to the accelerating electric field, this photocathode performed markedly worse than its sinusoidal predecessor.
+Indeed, no evidence of sucessful plasmon oscillation was demonstrated after repeated attempts.
+
+Though the cause of the failure may never be known for certain, several causes are suspected.
+Now that the surface is not purely sinusoidal, the surface has multiple Fourier components.
+This may mean that the laser may not couple as efficiently to the plasmon, requiring increased laser power for oscillation. 
+Additionally, the non-smooth surface may cause increaded difficulty for propagating the plasmon for large distances.
+Finally, upon further investigation of the surface by SEM, evidence seen of incomplete filling of the gold into the trenches, which may have interrupted creation or propagation of the plasmon.
+
+\subsection{Discussion}
+
+Although plasmonic photocathodes are still an interesting field of research in photocathode physics, we have concluded that they are not a promising candidate for an electron source for UEM at this time.
+They have been difficult to manufacture and thus would be expensive for owners.
+They are very intolerant to misalignment and would therefore be problematic for users.
+Finally, the cannot be shown to produce the performance needed to justify the previous concerns.
+
+Other nano-engineered surfaces, including quantum-wells, are still being considered for future work.

### 2012-12-12 inc/extension/implementation/Cooke.tex
added comment on Physics::UEMColumn and example script

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{}!/usr/bin/env perl}

\PY{k}{use} \PY{n}{strict}\PY{p}{;}
\PY{k}{use} \PY{n}{warnings}\PY{p}{;}

\PY{k}{use} \PY{n+nn}{Physics::UEMColumn} \PY{n}{alias} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{':standard'}\PY{p}{;}
\PY{k}{use} \PY{n+nn}{Physics::UEMColumn::Auxiliary} \PY{l+s}{':materials'}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{laser} \PY{o}{=} \PY{n}{Laser}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{width}    \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'1 mm'}\PY{p}{,}
  \PY{n}{duration} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'1 ps'}\PY{p}{,}
  \PY{n}{energy}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'4.75 eV'}\PY{p}{,}
\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{acc} \PY{o}{=} \PY{n}{DCAccelerator}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{length}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'20 mm'}\PY{p}{,}
  \PY{n}{voltage} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'20 kilovolts'}\PY{p}{,}
\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{column} \PY{o}{=} \PY{n}{Column}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{length}       \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'400 mm'}\PY{p}{,} 
  \PY{n}{laser}        \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{laser}\PY{p}{,}
  \PY{n}{accelerator}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{acc}\PY{p}{,}
  \PY{n}{photocathode} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{Photocathode}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}\PY{n}{Ta}\PY{p}{)}\PY{p}{,}
\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim} \PY{o}{=} \PY{n+nn}{Physics::}\PY{n}{UEMColumn}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{column}      \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{column}\PY{p}{,}
  \PY{n}{number}      \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mi}{1}\PY{p}{,}
  \PY{n}{debug}       \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mi}{1}\PY{p}{,}
\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{z\PYZus{}rf}       \PY{o}{=} \PY{l+m+mi}{20}\PY{p}{;} \PY{c+c1}{\PYZsh{}cm}
\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{l\PYZus{}mag\PYZus{}lens} \PY{o}{=} \PY{l+s}{'1 in'}\PY{p}{;}
\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{cooke\PYZus{}sep}  \PY{o}{=} \PY{l+m+mi}{5}\PY{p}{;} \PY{c+c1}{\PYZsh{}cm}
\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{str\PYZus{}mag}    \PY{o}{=} \PY{l+m+mf}{33e-13}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{lens1} \PY{o}{=} \PY{n}{MagneticLens}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{location} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{z\PYZus{}rf} \PY{o}{-} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{cooke\PYZus{}sep}\PY{p}{)} \PY{o}{.} \PY{l+s}{'cm'}\PY{p}{,}
  \PY{n}{length}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{l\PYZus{}mag\PYZus{}lens}\PY{p}{,}
  \PY{n}{strength} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{str\PYZus{}mag}\PY{p}{,}
\PY{p}{)}\PY{p}{;}
\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{lens2} \PY{o}{=} \PY{n}{MagneticLens}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{location} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{z\PYZus{}rf} \PY{o}{+} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{cooke\PYZus{}sep}\PY{p}{)} \PY{o}{.} \PY{l+s}{'cm'}\PY{p}{,}
  \PY{n}{length}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{l\PYZus{}mag\PYZus{}lens}\PY{p}{,}
  \PY{n}{strength} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{str\PYZus{}mag}\PY{p}{,}
\PY{p}{)}\PY{p}{;}
\PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{add\PYZus{}element}\PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{lens1}\PY{p}{)}\PY{p}{;}
\PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{add\PYZus{}element}\PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{lens2}\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{rf\PYZus{}cav} \PY{o}{=} \PY{n}{RFCavity}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
  \PY{n}{location}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{z\PYZus{}rf} \PY{o}{.} \PY{l+s}{'cm'}\PY{p}{,}
  \PY{n}{length}    \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'2 cm'}\PY{p}{,}
  \PY{n}{strength}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'230 kilovolts / m'}\PY{p}{,}
  \PY{n}{frequency} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'3 gigahertz'}\PY{p}{,}
\PY{p}{)}\PY{p}{;}
\PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{add\PYZus{}element}\PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{rf\PYZus{}cav}\PY{p}{)}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result} \PY{o}{=} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{propagate}\PY{p}{;}
\end{Verbatim}

### 2012-12-12 inc/extension/implementation/implementation.tex
added comment on Physics::UEMColumn and example script

\section{Implementation}

%The model presented in this chapter is deceptively hard to implement.
%As noted by Michalik and Sipe in their original paper \cite{michalik_analytic_2006}, in the absence of space-charge, the solution of the system of equations in \ref{eq:ag_original} may be solved analytically and result in the standard equations of Gaussian optics.
%With the inclusion of space-charge (that is the inclusion of electron-electron repulsion) the system must be solved by numerical methods.
%The scaling of the terms requires special consideration of error-handling.

%This problem is exacerbated when including the external forces.
%The necessity of limiting the region of influence of electron-optical elements (\ref{eq:reg_of_influence}) adds terms with Gaussian envelopes

The model presented in this chapter was initially implemented in Wolfram's Mathematica.
While this implementation provided the results needed, it was inflexible and overly confusing.
The limitations of the Mathematica language precluded abstracting the model sufficiently to cover all common use cases.

Recently the model was rewritten in Perl, and is called \verb!Physics::UEMColumn!.
Perl is a dynamic scripting language known for its versitility and flexibility.
In this implementation, the model is implemented as a decoupled set of high-level object-oriented classes.

\verb!Physics::UEMColumn! has been released as open source software under the terms of the Artistic License (a derivative of the GNU General Public License), which allows for free use of the software.
The current state of development of the simulation toolkit is publically available at \url{https://github.com/jberger/Physics-UEMColumn}.
Releases of the toolkit are available via the Comprehensive Perl Archive Network (CPAN), which is a central repository of software developed for Perl, a standard tool for the language.
The CPAN has several tracker sites, see for example \url{https://metacpan.org/module/Physics::UEMColumn}.

A simulation script is composed of creating instances of the classes that represent the column, the pulse, electron-optical elements like magnetic lenses and RF cavities.
These objects act as containers for the information about the physical object that they represent.
Further they manage progammatic tasks related to those objects.
For example, electron-optical element objects manage the Gaussian region-of-influence (\ref{eq:reg_of_influence}) and calculate the force that their fields apply to the pulse.
This allows for clean, managable and extensible code.

These objects are managed by a central ``simulation'' class (called \verb!Physics::UEMColumn!) which handles the time evolution of the system.
The source package contains an example script analogous to the optical Cooke triplet, containing a magnetic lens, RF cavity and another magnetic lens.
This script is included as Appendix \ref{sec:script}.

### 2012-12-12 pygments.tex
added comment on Physics::UEMColumn and example script

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.13,0.55,0.13}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.41,0.55}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.71,0.32,0.80}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.71,0.32,0.80}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.55,0.00,0.55}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.41,0.55}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.71,0.32,0.80}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.12,0.53,0.61}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.67,0.00}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.33,0.33}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@ul=\underline\def\PY@tc##1{\textcolor[rgb]{0.00,0.55,0.27}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.41,0.55}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.55,0.00}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.55,0.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.55,0.27}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.44,0.48,0.49}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.55,0.27}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.55,0.27}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.33,0.33}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.11,0.49,0.44}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.41,0.55}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.55,0.00,0.55}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.41,0.55}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.33,0.33}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.33,0.33,0.33}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.55,0.00,0.55}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.42,0.13}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.55,0.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.13,0.55,0.13}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.55,0.00,0.55}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.13,0.55,0.13}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.71,0.32,0.80}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.65,0.09,0.09}{##1}}\def\PY@bc##1{\setlength{\fboxsep}{0pt}\colorbox[rgb]{0.89,0.82,0.82}{\strut ##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.55,0.00,0.55}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.33,0.33}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.11,0.49,0.44}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.71,0.32,0.80}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.71,0.32,0.80}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.55,0.00,0.55}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.55,0.00,0.55}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.33,0.33}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.55,0.00,0.55}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.65,0.65,0.65}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.33,0.33}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.33,0.33}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.55,0.00,0.55}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.33,0.33}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.80,0.33,0.33}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

### 2012-12-12 inc/extension/implementation/Cooke.pl

@@ -1,0 +1,60 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+use Physics::UEMColumn alias => ':standard';
+use Physics::UEMColumn::Auxiliary ':materials';
+
+my $laser = Laser->new(
+  width    => '1 mm',
+  duration => '1 ps',
+  energy   => '4.75 eV',
+);
+
+my $acc = DCAccelerator->new(
+  length  => '20 mm',
+  voltage => '20 kilovolts',
+);
+
+my $column = Column->new(
+  length       => '400 mm', 
+  laser        => $laser,
+  accelerator  => $acc,
+  photocathode => Photocathode->new(Ta),
+);
+
+my $sim = Physics::UEMColumn->new(
+  column      => $column,
+  number      => 1,
+  debug       => 1,
+);
+
+my $z_rf       = 20; #cm
+my $l_mag_lens = '1 in';
+my $cooke_sep  = 5; #cm
+my $str_mag    = 33e-13;
+
+my $lens1 = MagneticLens->new(
+  location => ($z_rf - $cooke_sep) . 'cm',
+  length   => $l_mag_lens,
+  strength => $str_mag,
+);
+my $lens2 = MagneticLens->new(
+  location => ($z_rf + $cooke_sep) . 'cm',
+  length   => $l_mag_lens,
+  strength => $str_mag,
+);
+$sim->add_element($lens1);
+$sim->add_element($lens2);
+
+my $rf_cav = RFCavity->new(
+  location  => $z_rf . 'cm',
+  length    => '2 cm',
+  strength  => '230 kilovolts / m',
+  frequency => '3 gigahertz',
+);
+$sim->add_element($rf_cav);
+
+my $result = $sim->propagate;
+

### 2012-12-19 inc/gun/sharpness.pl
created figure comparing the tanh "sharpness" in the gun model

@@ -1,0 +1,50 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+use Physics::UEMColumn alias => ':standard';
+use Physics::UEMColumn::Auxiliary ':materials';
+use PDL;
+
+sub for_sharpness {
+  my $laser = Laser->new(
+    width    => '1 mm',
+    duration => '1 ps',
+    energy   => '4.75 eV',
+  );
+
+  my $acc = DCAccelerator->new(
+    length  => '20 mm',
+    voltage => '20 kilovolts',
+    sharpness => shift || 10,
+  );
+
+  my $column = Column->new(
+    length       => '25 mm', 
+    laser        => $laser,
+    accelerator  => $acc,
+    photocathode => Photocathode->new(Ta),
+  );
+
+  my $sim = Physics::UEMColumn->new(
+    column      => $column,
+    number      => 1,
+    debug       => 1,
+  );
+
+  my $result = $sim->propagate;
+
+  return pdl $result;
+}
+
+for ( qw/ 5 10 20 / ) {
+  my $pdl = for_sharpness($_)->dice([1,3,4])->sever;
+  my $slice = $pdl->slice('1:2');
+  $slice *= 2;
+  $slice **= 0.5;
+  $pdl->slice('0:1,') *= 1e3;
+  $pdl->slice('2,') *= 1e6;
+  $pdl->xchg(0,1)->wcols("sharpness$_.dat");
+}
+

### 2012-12-20 inc/gun/sharpness.pl
massaged onaxis data to compare nicely with sharpness data

@@ -15,13 +15,13 @@ sub for_sharpness {
   );
 
   my $acc = DCAccelerator->new(
-    length  => '20 mm',
-    voltage => '20 kilovolts',
+    length  => '0.25 in',
+    voltage => '10 kilovolts',
     sharpness => shift || 10,
   );
 
   my $column = Column->new(
-    length       => '25 mm', 
+    length       => '10 mm', 
     laser        => $laser,
     accelerator  => $acc,
     photocathode => Photocathode->new(Ta),

### 2012-12-21 inc/gun/gun_design.tex
began work on modeling accelerator, its in the gun folder now, depending on overlap might move to external_forces

\section{Accelerator Design} \label{sec:gun_design}

\begin{figure}
  \centering
  \input{gunfield}
\end{figure}

\begin{figure}
  \centerline{
    \subfloat[][]{
      \label{fig:field_on_axis}
      \input{onaxis}
    }
    \subfloat[][]{
      \label{fig:tanh_sharpness}
      \input{sharpness}
    }
  }
  %\caption{
  %  \protect\subref{fig:laser_schematic}
  %}
  \label{fig:laser-cavity}
\end{figure}

### 2012-12-21 inc/gun/gun_model.tex
began work on modeling accelerator, its in the gun folder now, depending on overlap might move to external_forces

\subsection{Acceleration Region}

Obviously an overall acceleration occurs due to the applied electic field in the acceleration gap.
The anode must contain an aperture to allow the electrons to continue down the column.
The electric field that nears this aperture turns outward as expected, causing the anode itself to act as a negative lens. %TODO reference
What may not be so obvious is that the falling field longitudinally acts as a pulse compression element.

To model an acceleration region, one must assume a form for the accelerating field.
From that form one may derive the terms governing the pulse dynamics. Initially, assume that the on-axis field is purely in the longitudinal direction, which may be thought of as an implication of radial symmetry.
Mathematically, that is
\begin{equation}
  \vec{F}(r^{\prime}=0,z^{\prime}) = F_z (0,z^{\prime}) \hat{z}
\end{equation}
where as before, the primed coordinates are the location of the center of the pulse.

To first model the longitudinal (compression) action at the anode %TODO ref
on electrons that are a distance $z$ ahead or behind of the center of the pulse feel a slightly different field, 
\begin{equation}
  F_z(0,z^{\prime} - z) \approx F_z(0,z^{\prime} - z) + \frac{\partial F_z}{\partial z^{\prime}} z
\end{equation}
The leading term does not contribute to the intrapulse dynamic (see Section \ref{sec:const_force}), however the second term is of the proper form to say that
\begin{equation}
  M_z = \frac{\partial F_z}{\partial z^{\prime}}
\end{equation}

To model the radial action (negative lensing), we can expand for electrons that are a distance $r$ radially outward from the center of the pulse,
\begin{equation}
  F_r(r,z^{\prime}) \approx F_r(r,z^{\prime}) + \frac{\partial F_r}{\partial r^{\prime}} r
\end{equation}
without an analytic form for $F$, we can appeal to Gauss' law (in absense of external charge)
\begin{equation}
  \frac{\partial F_x}{\partial x^{\prime}} + \frac{\partial F_y}{\partial y^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 2 \frac{\partial F_{\smallT}}{\partial r^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 0
\end{equation}
to see that
\begin{equation}
  M_{\smallT} = -\frac{1}{2}\frac{\partial F_z}{\partial z^{\prime}}
\end{equation}

We may choose to assume that the on-axis field is defined as a smoothed step down function
\begin{equation}
  F_z(0,z^{\prime}) = \frac{qV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right)
\end{equation}
where $V$ is the accelerating potential, $z^{\prime}_A$ is the position of the anode (the cathode assumed to be at $z^{\prime} = 0$) and $s$ is a constant quantifying the ``sharpness'' of the fall-off of the electric field.
Indeed, this form is analgous to the super-Gaussian envalope (\ref{eq:reg_of_influence}), however it is open ended on the cathode side, which is necessary for the electrons to be accelerated away from the photocathode.
Given this form (justified in Section \ref{sec:gun_design}) the lensing constants are then
\begin{gather}
  M_{\smallT} = -\frac{qVs}{4z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \\
  M_z = \frac{qVs}{2z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right)
\end{gather}

### 2012-12-28 inc/gun/gun_model.tex
improved gun model description

@@
 
-Obviously an overall acceleration occurs due to the applied electic field in the acceleration gap.
-The anode must contain an aperture to allow the electrons to continue down the column.
-The electric field that nears this aperture turns outward as expected, causing the anode itself to act as a negative lens. %TODO reference
-What may not be so obvious is that the falling field longitudinally acts as a pulse compression element.
+In the acceleration region of a UEM column, a constant electric field is applied to the newly created electron pulse, increasing its velocity until it achieves its ultimate velocity just upon leaving the region at the anode aperture.
+As seen in Section \ref{sec:const_force}, this constant electric field does not affect the intrapulse dynamics, however the electric field that nears this aperture turns outward, causing the anode itself to act as a negative lens.
+%TODO reference
+Additionally, the longitudinal electric field stregth sharply decreases near the anode.
+Because the front of the pulse sees this change before the back, this changing field acts as a pulse compression element.
+
+Since the aforementioned effects are localized in the area very near the anode, the following analysis is only asserted for pulses that are of approximately the same length as the acceleration region or shorter.
+This assumption is very reasonable for UEM and for some of the shorter-pulse DTEM systems.
+Though these effects are expected for longer pulses, they are unlikely to be predominant factors in the dynamics.
 
@@
 \end{equation}
-The leading term does not contribute to the intrapulse dynamic (see Section \ref{sec:const_force}), however the second term is of the proper form to say that
+As noted before, the constant leading term does not contribute to the intrapulse dynamic, however the second term is of the proper form to say that in this region
 \begin{equation}
@@
 Indeed, this form is analgous to the super-Gaussian envalope (\ref{eq:reg_of_influence}), however it is open ended on the cathode side, which is necessary for the electrons to be accelerated away from the photocathode.
-Given this form (justified in Section \ref{sec:gun_design}) the lensing constants are then
+Given this form (demonstrated in Section \ref{sec:gun_design}) the lensing constants are then
 \begin{gather}
@@
 \end{gather}
+
+\begin{figure}
+  \centerline{
+    \subfloat[][]{
+      \label{fig:field_on_axis}
+      \input{onaxis}
+    }
+    \subfloat[][]{
+      \label{fig:tanh_sharpness}
+      \input{sharpness}
+    }
+  }
+  \caption{
+    An example of a tahn form acceleration region (overall length 6.35mm) for three different ``sharpness'' values.
+    The square data points in \protect\subref{fig:field_on_axis} represent the on-axis electric field data from FEM simulations, see Section \ref{sec:gun_design}.
+    The colored electric fields in \protect\subref{fig:field_on_axis} correspond to the pulse width (solid) and length (dashed) in \protect\subref{fig:tanh_sharpness}.
+  }
+  \label{fig:laser-cavity}
+\end{figure}
+
+\ref{fig:laser-cavity} shows an example acceleration region for three values of the ``sharpness'' $s$.
+What is clear from this figure is that the dynamics are not highly coupled to the value of this hard-to-quantify parameter.
+For all values of $s$ the divergence of the beam is equal (thought the size may vary slightly).
+Interestingly the length of the pulse is stabilized on exiting the anode, coming very nearly to a constant length (not expanding nor contracting).
+This effect is due to the slower electrons remaining in the cavity longer and thus receiving an additional kick, as described in \cite{oudheusden_electron_2007}.

### 2012-12-28 inc/extension/implications/spacecharge.pl
recreate fig 5 from M&M paper (DC gun)

@@ -1,0 +1,52 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+use Physics::UEMColumn alias => ':standard';
+use Physics::UEMColumn::Auxiliary ':materials';
+
+sub for_number {
+  my $num = shift;
+
+  my $laser = Laser->new(
+    width    => '1 mm',
+    duration => '1 ps',
+    energy   => '4.75 eV',
+  );
+
+  my $acc = DCAccelerator->new(
+    length  => '20 mm',
+    voltage => '20 kilovolts',
+  );
+
+  my $column = Column->new(
+    length       => '15 cm', 
+    laser        => $laser,
+    accelerator  => $acc,
+    photocathode => Photocathode->new(Ta),
+  );
+
+  my $sim = Physics::UEMColumn->new(
+    column      => $column,
+    number      => $num,
+  );
+
+  my $result = $sim->propagate;
+  return $result;
+}
+
+open my $fh, '>', 'spacecharge.dat';
+
+my @nums = map { ("1e$_", "2.2e$_", "4.6e$_") } (0..9);
+pop @nums for 1..2;
+
+for my $num ( @nums ) {
+  my $result = for_number($num);
+  for (@$result) {
+    next unless $_->[1] >= 0.015;
+    print {$fh} $num . ' ' . sqrt( 2 * $_->[3] ) * 1e3 . ' ' . sqrt( 2 * $_->[4] ) * 1e3 . "\n";
+    last;
+  }
+}
+

### 2013-01-09 inc/extension/initial_conditions/initial_conditions.tex
use Dowell calcuated initial transverse emission formulation

--- 2012-07-10 inc/initial_conditions/initial_conditions.tex
@@
 
-The single-photon photoemission analysis of Jensen \textit{et al.}\cite{jensen_theoretical_2006} indicates that for a metallic photocathode the transverse momentum variance at $ t = 0 $ is given by
+The careful analysis of Dowell and Schmerge \cite{dowell_quantum_2009} indicates that for a metallic photocathode the transverse momentum variance at $ t = 0 $ is given by
 \begin{equation} \label{eq:jensen_eta}
-( \Delta p_{x} )^{2} = ( \Delta p_{y} )^{2} = \eta_{ \smallT } ( 0 ) \approx \frac{ m_{e} \epsilon_{\smallF} ( \hbar \omega - \Phi ) } { \hbar \omega + \epsilon_{\smallF} } \; \text{,}
+  \eta_{ \smallT } ( 0 ) \approx \frac{ m_{e} }{ 3 } ( \hbar \omega - \Phi ) \; \text{,}
 \end{equation}
-in the limit where the Fermi energy $ \epsilon_{\smallF} $ of the metal photocathode and excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ are significantly greater that the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
+%in the limit where the Fermi energy $ \epsilon_{\smallF} $ of the metal photocathode and excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ are significantly greater that the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
+in the limit where the temperature is low enough to consider the system well described by Fermi-Dirac statistics where the chemical potential is nearly equal to the Fermi energy.
 From a simplistic viewpoint, only the electrons generated with positive longitudinal momentum will contribute to the pulse since those electrons generated with negative momentum do not leave the photocathode.
@@
 \gamma_{ z } ( 0 ) = 0 \\
-\eta_{ \smallT } ( 0 ) = \frac{ m_{e} \epsilon_{ F } ( \hbar \omega - \Phi ) }{ 3 ( \hbar \omega + \epsilon_{ \smallF } ) } \\
+\eta_{ \smallT } ( 0 ) = \frac{ m_{e} } { 3 } ( \hbar \omega - \Phi ) \\
 \eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 ) }{ 4 }

### 2013-01-10 inc/extension/initial_conditions/bin_model_figure.tex
Implemented bin model result plots in the paper

\begin{figure}
  \centerline {
    \subfloat[][]{
      \begin{tikzpicture}
        \input{Acc}
      \end{tikzpicture}
    }
    \subfloat[][]{
      \begin{tikzpicture}
        \input{Vmax}
      \end{tikzpicture}
    }
    \\
  }
  \centerline{
    \subfloat[][]{
      \begin{tikzpicture}
        \input{Equal}
      \end{tikzpicture}
    }
  }
\end{figure}

### 2013-01-12 inc/extension/initial_conditions/initial_conditions.tex
Implemented bin model result plots in the paper
work on DC dominated longitudinal initial conditions

@@
 
-For single-photon photoemission process with a spatially cylindrically symmetric incident Gaussian laser beam with Gaussian temporal irradiance profile 
+% transverse
+Consider a single-photon photoemission process with a spatially cylindrically symmetric incident Gaussian laser beam with Gaussian temporal irradiance profile 
 \begin{equation}\label{eq:laser_form}
@@
 \end{equation}
-where $w$ is the half-width 1/e maximum (HW1/eM) of the spatial irradiance distribution, $\tau$ is the HW1/e pulse duration and $I_{0}$ is the peak pulse irradiance, the electron pulse's initial spatial variance is
+where $w$ is the half-width 1/e maximum (HW1/eM) of the spatial irradiance distribution, $\tau$ is the HW1/e pulse duration and $I_{0}$ is the peak pulse irradiance incident on a flat metallic photocathode.
+The initial tranverse ($\smallT$) properties of the beam are easily understood.
+The electron pulse's initial spatial variance is simply
 \begin{equation}
@@
 \end{equation}
-For the longitudinal direction, and in the limit where dispersion effects in the acceleration region of the electron gun can be neglected, we may write
-\begin{equation} \label{eq:initial_sigma_z}
-\sigma_{ z } ( 0 ) = ( \Delta z)^{2} = \frac{ ( v_{{ \scriptscriptstyle 0}} \tau )^{2} }{ 2 } \text{ ,}
-\end{equation}
-with the velocity of the pulse in the lab frame $v_{{ \scriptscriptstyle 0}} = \sqrt{ 2 e V / m_{e} d } $, where $ V $ and $ d $ are the potential and cathode-anode separation of the gun's DC acceleration region respectively.
 For spatially-uniform photoemission from a planar photocathode, we have $ \gamma_{\smallT} ( 0 ) = 0 $; i.e., there is no initial spatial momentum chirp in the electron bunch.
-Immediately after photo-generation, however, it is strictly inaccurate to say that $ \gamma_{ z } ( 0 ) = 0 $.
-This is because the excess photoemission energy $ \Delta E = \hbar \omega - \Phi $, associated with the electron emission from a material with work function $\Phi$ using photons of energy $ \hbar \omega $ (which gives the maximum initial velocity of an electron, $ v_{max} = \sqrt{ 2 \Delta E/m_{e} } $), can produce an initial momentum chirp.
-Nonetheless, the initial non-zero nature of $ \gamma_{z} $ is quickly dominated by the evolution of the pulse, so this approximation will suffice in the region far from the gun.
-
-The careful analysis of Dowell and Schmerge \cite{dowell_quantum_2009} indicates that for a metallic photocathode the transverse momentum variance at $ t = 0 $ is given by
+Finally, thanks to the careful analysis of Dowell and Schmerge \cite{dowell_quantum_2009}, we know that metallic photocathode's initial transverse momentum variance at $ t = 0 $ is given by
 \begin{equation} \label{eq:jensen_eta}
@@
 \end{equation}
-%in the limit where the Fermi energy $ \epsilon_{\smallF} $ of the metal photocathode and excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ are significantly greater that the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
-in the limit where the temperature is low enough to consider the system well described by Fermi-Dirac statistics where the chemical potential is nearly equal to the Fermi energy.
+in the limit where the excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ are significantly greater that the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
+%in the limit where the temperature is low enough to consider the system well described by Fermi-Dirac statistics where the chemical potential is nearly equal to the Fermi energy.
+
+% longitudinal
+In the longitudinal direction ($z$), the interplay between acceleration, $ a = e E_{DC}/m $, in the DC gun and the physics of the photoemission process determine the properties of the electron bunch.
+Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) by the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than that traveled by electrons with the maximum photoemission velocity $ v_{max} $ alone over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
+
+% dc dominated
+In the long-pulse regime, the DC field dominated case, we assume that the intrinsic velocity of the emitted electrons is negligable.
+Clearly the first emitted electrons travel a distance $ \Delta z = \frac{ 1 }{ 2 } a ( \Delta t_{L} )^{ 2 } = \sqrt{ \sigma_{ z } ( 0 ) } $, during the  $ \Delta t_{ L } $ rms laser pulse duration: for a laser pulse described above, we have $ ( \Delta t_{L} )^{ 2 } = \tau^{ 2 } / 2 $.
+This result may be restated as,
+\begin{equation}
+  \sigma_{ z } ( 0 ) = \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 }
+\end{equation}
+By a similar logic, we can say that the `front' of the pulse then has a velocity equal to $ a \Delta t_{ L } $, while the `back' of the pulse (a distance $ \frac{ 1 }{ 2 } \sqrt{ \sigma_{ z } ( 0 ) } $ from the electron bunch peak) has zero velocity since it has just been generated.
+From this one sees that the initial linear chirp across the pulse is given by
+\begin{equation}
+  \gamma_{ z } ( 0 ) = m \sigma_{ z } ( 0 ) \frac{ \Delta v }{ \Delta z } = m a \Delta t_{ L } \sqrt{ \sigma_{ z } ( 0 ) } = \frac{ e E_{ DC } \tau }{ \sqrt{ 2 } }\, .
+\end{equation} 
 From a simplistic viewpoint, only the electrons generated with positive longitudinal momentum will contribute to the pulse since those electrons generated with negative momentum do not leave the photocathode.
-This implies that the uncertainty in the longitudinal direction is half that of the transverse directions; that is to say
+This implies that the local momentum uncertainty in the longitudinal direction is half that of the transverse directions; that is to say
 \begin{equation}
@@
 
+% v_0 dominated
+with the velocity of the pulse in the lab frame $v_{{ \scriptscriptstyle 0}} = \sqrt{ 2 e V / m_{e} d } $, where $ V $ and $ d $ are the potential and cathode-anode separation of the gun's DC acceleration region respectively.
+Immediately after photo-generation, however, it is strictly inaccurate to say that $ \gamma_{ z } ( 0 ) = 0 $.
+This is because the excess photoemission energy $ \Delta E = \hbar \omega - \Phi $, associated with the electron emission from a material with work function $\Phi$ using photons of energy $ \hbar \omega $ (which gives the maximum initial velocity of an electron, $ v_{max} = \sqrt{ 2 \Delta E/m_{e} } $), can produce an initial momentum chirp.
+Nonetheless, the initial non-zero nature of $ \gamma_{z} $ is quickly dominated by the evolution of the pulse, so this approximation will suffice in the region far from the gun.
+
 These initial conditions at $t=0$ for an incident Gaussian laser pulse of the form given in \ref{eq:laser_form}, will be used for the remainder of this paper.
@@
     \sigma_{ \smallT } ( 0 ) = \frac{ w^{ 2 } }{ 2 } \\
-\sigma_{ z } ( 0 ) = \frac{ ( v_{{ \scriptscriptstyle 0}} \tau )^{ 2 } }{ 2 } \\
-%alternative formulation + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
+    \sigma_{ z } ( 0 ) = \frac{ ( v_{ max } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
     \gamma_{ \smallT } ( 0 ) = 0 \\
-\gamma_{ z } ( 0 ) = 0 \\
-\eta_{ \smallT } ( 0 ) = \frac{ m_{e} } { 3 } ( \hbar \omega - \Phi ) \\
-\eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 ) }{ 4 }
+    \gamma_{ z } ( 0 ) = \sqrt{ \sigma_{ z } ( 0 ) } \left ( 0.06 m v_{ max } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
+    \eta_{ \smallT } ( 0 ) = \frac{ m \epsilon_{ F } ( \hbar \omega - \Phi ) }{ 3 ( \hbar \omega + \epsilon_{ F } ) } \\
+    \eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 ) }{ 3.85 }
   \end{gather}
@@
 Only for specialized cases, such as the generation of uniform ellipsoid electron bunch shapes, is spatial and temporal manipulation of the incident laser pulse considered.\cite{luiten_how_2004,li_generating_2008}
+
+\input{bin_model_figure}

### 2013-01-14 inc/extension/initial_conditions/initial_conditions.tex
more work on initial conditions

@@
 In the longitudinal direction ($z$), the interplay between acceleration, $ a = e E_{DC}/m $, in the DC gun and the physics of the photoemission process determine the properties of the electron bunch.
-Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) by the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than that traveled by electrons with the maximum photoemission velocity $ v_{max} $ alone over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
+Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) by the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than that traveled by electrons with the maximum photoemission velocity $ v_{0} $ alone over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
 
@@
 
-% v_0 dominated
-with the velocity of the pulse in the lab frame $v_{{ \scriptscriptstyle 0}} = \sqrt{ 2 e V / m_{e} d } $, where $ V $ and $ d $ are the potential and cathode-anode separation of the gun's DC acceleration region respectively.
-Immediately after photo-generation, however, it is strictly inaccurate to say that $ \gamma_{ z } ( 0 ) = 0 $.
-This is because the excess photoemission energy $ \Delta E = \hbar \omega - \Phi $, associated with the electron emission from a material with work function $\Phi$ using photons of energy $ \hbar \omega $ (which gives the maximum initial velocity of an electron, $ v_{max} = \sqrt{ 2 \Delta E/m_{e} } $), can produce an initial momentum chirp.
-Nonetheless, the initial non-zero nature of $ \gamma_{z} $ is quickly dominated by the evolution of the pulse, so this approximation will suffice in the region far from the gun.
+%v0 dominated
+In the short pulse limit, where the acceleration in the DC gun field over the incident laser pulse duration can be neglected, the initial electron bunch properties are determined primarily by the photoemission physics.
+%TODO describe new model
+Careful analysis of the electron generation using the longitudinal $ v^{3} $ photoemission velocity distribution (Appendix \ref{sec:step_model}) and iterative time steps through the incident laser pulse reveals that, although the spatial electron bunch distribution is slightly skewed with respect to the symmetric incident laser pulse, its initial variance $ \sqrt{ \sigma_{ z } ( 0 ) } $ (or width) is nearly equal to that of the laser pulse of the normalization by the velocity ratio $ v_{ 0 }/c $; in other words,
+\begin{equation} \label{eq:initial_sigma_z}
+\sqrt{ \sigma_{ z } ( 0 ) } = v_{ 0 } \Delta t_{L} = \frac{ v_{ 0 } \tau }{ \sqrt{ 2 } } \, \text{.}
+\end{equation}
+Not unexpectedly, the analysis also indicates that the initial electron bunch has a velocity chirp.
+%TODO description again
+Its value is only roughly 6\% of $ v_{ 0 }/\sqrt{ \sigma_{z} ( 0 ) } $ since the photoemission generates the $ v^{ 3 } $ velocity distribution throughout the laser pulse; thus 
+\begin{equation}
+\gamma_{ z } ( 0 ) = m \sigma_{ z } ( 0 ) \frac{ d v }{ d z } \approx 0.06 m v_{ 0 } \sqrt{ \sigma_{ z } ( 0 ) } \, ,
+\end{equation}
+in this regime.
+Additionally the iterative time-step analysis clearly shows that the local longitudinal momentum variance $ \eta_{ z } $ is not constant over the inital photo-generated electron bunch.
+In fact, $ \sqrt{ \eta_{ z } ( 0 ) } $ varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch with the narrowest momentum distribution at the front of the bunch where only the faster electrons with velocity near $ v_{ 0 } $ initially reside.
+Unfortunately, the mean-field theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch.
+Consequently, we will assume that $ \eta_{ z } $ is a constant over the bunch with a value given by $ \eta_{\smallT } ( 0 ) \cong 3.85 \eta{ z } ( 0 ) $ (see Appendix \ref{sec:step_model}): indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.
 
@@
     \sigma_{ \smallT } ( 0 ) = \frac{ w^{ 2 } }{ 2 } \\
-    \sigma_{ z } ( 0 ) = \frac{ ( v_{ max } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
+    \sigma_{ z } ( 0 ) = \frac{ ( v_{ 0 } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
     \gamma_{ \smallT } ( 0 ) = 0 \\
-    \gamma_{ z } ( 0 ) = \sqrt{ \sigma_{ z } ( 0 ) } \left ( 0.06 m v_{ max } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
+    \gamma_{ z } ( 0 ) = \sqrt{ \sigma_{ z } ( 0 ) } \left ( 0.06 m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
     \eta_{ \smallT } ( 0 ) = \frac{ m \epsilon_{ F } ( \hbar \omega - \Phi ) }{ 3 ( \hbar \omega + \epsilon_{ F } ) } \\

### 2013-01-25 inc/extension/external_forces/external_forces.tex
began work on modeling accelerator, its in the gun folder now, depending on overlap might move to external_forces
small fixes

--- 2012-07-10 inc/external_forces/external_forces.tex
@@
 
-\subsection{Constant External Forces}
+\subsection{Constant External Forces} \label{sec:const_force}
 
@@
 A magnetic lens may be modeled using one linear force contribution.
-Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is very complicated to describe\cite{montgomery_some_1961} and the ``lensing'' action is actually a compound force (first rotating the beam, then lensing, see Section \ref{sec:mag_lens}).
+Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is very complicated to describe \cite{montgomery_some_1961} and the ``lensing'' action is actually a compound force (first rotating the beam, then lensing).
 We have found that in practice, characterizing a specific magnetic lens ($M_{\smallT}(I)$ where $I$ is the current in the magnetic lens) is much easier than attempting to find a generic functional form.
+Further discussion of magnetic lenses can be found in Section \ref{sec:mag_lens}.
 

### 2013-01-25 inc/extension/initial_conditions/initial_conditions.tex
fix equation, add longitudinal model datas and plot/fitting
small fixes

@@
 %TODO describe new model
-Careful analysis of the electron generation using the longitudinal $ v^{3} $ photoemission velocity distribution (Appendix \ref{sec:step_model}) and iterative time steps through the incident laser pulse reveals that, although the spatial electron bunch distribution is slightly skewed with respect to the symmetric incident laser pulse, its initial variance $ \sqrt{ \sigma_{ z } ( 0 ) } $ (or width) is nearly equal to that of the laser pulse of the normalization by the velocity ratio $ v_{ 0 }/c $; in other words,
+Careful analysis of the electron generation using the longitudinal $ v^{3} $ photoemission velocity distribution %TODO reference bin model
+and iterative time steps through the incident laser pulse reveals that, although the spatial electron bunch distribution is slightly skewed with respect to the symmetric incident laser pulse, its initial variance $ \sqrt{ \sigma_{ z } ( 0 ) } $ (or width) is nearly equal to that of the laser pulse of the normalization by the velocity ratio $ v_{ 0 }/c $; in other words,
 \begin{equation} \label{eq:initial_sigma_z}
@@
 Unfortunately, the mean-field theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch.
-Consequently, we will assume that $ \eta_{ z } $ is a constant over the bunch with a value given by $ \eta_{\smallT } ( 0 ) \cong 3.85 \eta{ z } ( 0 ) $ (see Appendix \ref{sec:step_model}): indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.
+Consequently, we will assume that $ \eta_{ z } $ is a constant over the bunch with a value given by $ \eta_{\smallT } ( 0 ) \cong 3.85 \eta{ z } ( 0 ) $: %TODO reference bin model
+indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.
 
@@
     \gamma_{ z } ( 0 ) = \sqrt{ \sigma_{ z } ( 0 ) } \left ( 0.06 m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
-    \eta_{ \smallT } ( 0 ) = \frac{ m \epsilon_{ F } ( \hbar \omega - \Phi ) }{ 3 ( \hbar \omega + \epsilon_{ F } ) } \\
+    \eta_{ \smallT } ( 0 ) = \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \\
     \eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 ) }{ 3.85 }

### 2013-01-25 inc/gun/gun_design.tex
improved gun model description
began describing the Togawa gun

@@
 \section{Accelerator Design} \label{sec:gun_design}
+
+The acceleration region of our prototype UEM system has several requirement which are unique in comparison to standard electron guns.
+Primarily, given the limitations on the beam discussed earlier, %TODO link to discussion of Childs law beam size limitation
+the employ laser pulse must be a relatively large area.
+This concern leads to several additional criteria.
+To accomodate this large beam, the anode aperture must be similarly large; in fact, this will be larger than the emission source due to beam expansion during acceleration.
+Also, in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the accelerating electic field given that it will span a wider proportion of the accelerator than thin beams used in standard TEM.
 
@@
   \input{gunfield}
+  \caption{
+    Schematic of the employed Togawa inspired cathode ($-V$) and anode (0V) pair.
+    The emission source is inserted behind the cathode Wehnelt.
+    Also shown are simulated equipotential lines generated by FEM modeling.
+    The separation of the cathode and anode is adjustable to allow a greater range of input laser angles if needed. 
+  }
+  \label{fig:gun-field}
 \end{figure}
 
-\begin{figure}
-  \centerline{
-    \subfloat[][]{
-      \label{fig:field_on_axis}
-      \input{onaxis}
-    }
-    \subfloat[][]{
-      \label{fig:tanh_sharpness}
-      \input{sharpness}
-    }
-  }
-  %\caption{
-  %  \protect\subref{fig:laser_schematic}
-  %}
-  \label{fig:laser-cavity}
-\end{figure}
+These requirements lead to the fabrication of cathode/Wehnelt and anode pair based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007}.
+A schematic of this design, overlayed by a Finite Element Model (FEM) simulation of the electic field, is shown in \ref{fig:gun-field}.
+The schematic shows that the aperture concerns are easily satisfied.
+From the field equipotentials, it is clear that this design has a consistently flat electric field for large displacements off the central axis.
+
+%TODO comment on alignment
+%TODO comment on high voltage
+%TODO comment on EV parts etc

### 2013-01-25 inc/gun/gun_design.tex
more gun design

@@
 
+Of course, as derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electic field.
+This field will cause a divergence of the beam.
+Future work will include investigating the designs of Butler %TODO reference
+which may be able to build in an intrinsic lensing to compensate for this effect.
+
 %TODO comment on alignment

### 2013-01-25 inc/gun/gun_design.tex
comment on alignment

@@
 %TODO comment on EV parts etc
+
+Re-alignment of cathode and anode is sometimes necessary to direct the emitted beam down the center axis of the column.
+To facilitate this, the prototype column has the photocathode and Wehnelt mounted on the end of the high voltage feedthrough, which is in turn mounted on a two axis alignment mechanism.
+Although this mechanism, called a ``port aligner,'' is typically used to couple two non-colinear vacuum ports, for use in this instrument it has been modified to be able to manually adjust the aligner during use.
+In contrast to this flexibility, the anode is mounted to column using Kimball Physics's ``eV Parts'' mounting systems.
+These strictly manufactured mounting hardwares ensure that the anode is properly aligned with the remainder of the microscope column.
+Thus, by iteratively repositioning the input laser and then the gun alignment, the beam may be walked to pass as near as possible down the axis of the column.

### 2013-01-28 inc/hardware/hardware.tex
begin hardware section

\section{Column Design and Construction}

Apart from the laser system, the majority of an Ultrafast Electron Microscope column is either vacuum hardware or contained within the vacuum system.
It is common in the UEM field to modify an existing Transmission Electron Microscope column, adapting it for laser-stimulated electron generation.
The system presented in the following section has been designed and built either from scatch or from commercially available parts.
The reasons for this are two-fold, the ease of prototyping inherent in using replacable parts and the physical limitations our unique choices have imposed on our beam profile.

The vacuum chamber itself is almost entirely 4.5 in standard ConFlat hardware.
Much has been purchased from off-the-shelf hardware suppliers.
More complicated chamber components have been purchased from Kimball Physics Inc.

Kimball's line of Multi-CF hardware has been especially useful in building the prototype column.
The main chamber is a 4.5 in ``spherical cube,'' having six 4.5 in ConFlat ports on a cube, with eight 1.33 in ports on the corners.
All ports on Multi-CF hardware also contain a set of grooves, designed as mounting points for a system of erector-set-like equipment called ``eV Parts.''

### 2013-01-29 inc/hardware/column.tex
generalized hardware folder

--- 2013-01-28 inc/hardware/hardware.tex
@@
 All ports on Multi-CF hardware also contain a set of grooves, designed as mounting points for a system of erector-set-like equipment called ``eV Parts.''
+This system is used to mount the accelerator anode and the the deflector plates which are both to be described later.

### 2013-01-30 inc/hardware/deflector.tex
added deflector content and schematic

\section{Deflection Plates}

In order to redirect the beam down the column, in conjunction with the accelerator alignment mechanisms, custom large area deflector plates were employed.
Four plates are arranged on a hollow tube with a square profile, one on each side.
One of each pair is held at ground while the other has a variable potential.
By varying each potential the beam may be turned in flight.

\begin{figure}
  \centering
  \include{deflector_schematic}
  \caption{
    Schematic diagram of one pair of deflection plates.
    The blue ray represents the path of the pulse as a whole (not to be confused with the plot of a lens).
  }
  \label{fig:deflector_schematic}
\end{figure}

A schematic description of such a system is shown in \ref{fig:deflector_schematic}.
To estimate the required voltage on the plates $V$, consider the applied impulse provided to the beam by the plate's electric field
\begin{gather}
  F \Delta t = m \Delta v \\
  \frac{ e V }{ d } \frac{ L }{ v_z } = m_e \Delta v_x
\end{gather}
where (as seen in the figure) $L$ is the length of the plate (in the direction of travel $\hat{z}$) and $d$ is the separation of the plates.
Then in the small angle approximation, the angle that the plate may correct for is approximately
\begin{equation}
  \theta = \frac{ \Delta v_x }{ v_z } = \frac{ e }{ m_e } \frac{ L }{ d } \frac{ V }{ v_z^2 }
\end{equation}
Using this result, when $L \sim d$ and $v_z = c/3$ a voltage of 1kV will deflect the beam approximately one degree.
These plates therefore are used as a fine adjustment, as opposed to the coarse adjustment provided by aligning the acceleration gap.

### 2013-01-30 inc/photocathode/este/este.tex
typo

@@
 Inside the material, an electron with an energy $E$ greater than $\phi$ has a maximum momentum of $\sqrt{2 m^* (E-\phi) }$ parallel to the boundary if it is to be emitted.
-As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_x$ of the electron source clearly must scale with $\sqrt{m_*}$ as observed in the experiment.
+As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_x$ of the electron source clearly must scale with $\sqrt{m^*}$ as observed in the experiment.
 This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies\cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.

### 2013-01-31 inc/magnetic_lenses/maglens.tex
continuing to move \section directives into subimported files
added magnetic lens picture

@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
+\section{Magnetic Lenses} \label{sec:mag_lens}
 
@@
 
+A picture of this system is shown in \ref{fig:maglens-pic}.
+In that picture the two lenses are shown, however a placeholder is inserted between them rather than the RF cavity.
+
+\begin{figure}
+  \centering
+  \includegraphics{maglens.jpg}
+  \caption{
+    Side view of the two custom large-bore magnetic lenses designed and built at UIC, seen installed in the prototype UEM.
+    The coils are visible as is high permeability shielding used for the pole-piece.
+    The outer (larger) plates and struts are for structural integrity only.
+  }
+  \label{fig:maglens-pic}
+\end{figure}
+
 \subsection{Thermal Properties}

### 2013-01-31 main.tex
continuing to move \section directives into subimported files
remove stray unicode characters from bibliography files
tracking down some more unicode
added plasmon file, include cc warning on pole piece figure
expand on plasmon
remove some unicode
use space between word and \cite
added some more section stubs
removed old semiconductor section and add apl paper as este
add apl bibliography up to entry 13
added \inputdata function and inputdata boolean to include larger datasets or not (for speed)
redo some headings in plasmon section
created separate abbreviations file to keep main cleaner. add

@@
 \input{mytikz}
+\setboolean{inputdata}{true}
+\usepackage{fancyvrb}
+\input{pygments}
 
@@
 
+\usepackage{color}
+\providecommand{\fixme}[1]{\textcolor{red}{#1}}
+
 %% Special handling for using biblatex vs bibtex ==
@@
   \usepackage{biblatex}
+  %\usepackage[backend=biber]{biblatex-chicago}
   \addbibresource{bibliography.bib}
+  \addbibresource{biblaser.bib}
+  \addbibresource{bibeste.bib}
 }{}
@@
 
-\listofabbreviations
-\begin{list}
-  {}
-  {\setlength{\labelwidth}{1in}
-   \setlength{\leftmargin}{1.5in}
-   \setlength{\labelsep}{.5in}
-   \setlength{\rightmargin}{\leftmargin}}
-
-  \abbr{AG}{Analytic Gaussian (Model)}
-  \abbr{AR}{Anti-Reflection}
-  \abbr{DC}{Direct Current (i.e. Static Electric Field)}
-  \abbr{DTEM}{Dynamic Transmission Electron Microscopy}
-  \abbr{FWHM}{Full-Width Half Maximum}
-  \abbr{GRIN}{Gradient Index (of Refraction)}
-  \abbr{GVD}{Group Velocity Dispersion}
-  \abbr{HW1/eM}{Half-Width 1/e Maximum}
-  \abbr{LHS}{Left Hand Side}
-  \abbr{ML}{Mode-Locked}
-  \abbr{Q}{Quality Factor}
-  \abbr{RHS}{Right Hand Side}
-  \abbr{RF}{Radio Frequency}
-  \abbr{SBR}{Saturable Bragg Reflector}
-  \abbr{TLS}{Thermal Lens Shaping}
-  \abbr{TEM$_{00}$}{Tranvserse ElectroMagnetic 00 Mode}
-  \abbr{TM$_{010}$}{Tranvserse Magnetic 010 Mode}
-  \abbr{UED}{Ultrafast Electron Diffraction}
-  \abbr{UEM}{Ultrafast Electron Microscope}
-  \abbr{UIC}{University of Illinois at Chicago}
-\end{list}
+\input{abbreviations}
 
@@
 
-  \subimport{inc/liouvilles_theorem/}{liouvilles_theorem}
+  \subimport{inc/extension/initial_conditions/}{initial_conditions}
 
-  \subimport{inc/initial_conditions/}{initial_conditions}
+  \subimport{inc/extension/liouvilles_theorem/}{liouvilles_theorem}
 
-  \subimport{inc/external_forces/}{external_forces}
+\section{Free Expansion Results}
+
+\subsection{Comparison to Other Articles}
+
+\subsection{Trends for Initial Shapes}
+
+  \subimport{inc/extension/external_forces/}{external_forces}
+
+  \subimport{inc/gun/}{gun_model}
+
+  \subimport{inc/extension/implementation/}{implementation}
 
@@
 
+  \begin{figure}
+    \centering
+    \includegraphics{inc/hardware/column.jpg}
+    \caption{
+      Image of the majority of the prototype UEM column at UIC.
+      The main vacuum chamber is laid sideways, with the high voltage (gun) at the left and the imaging system on the right.
+      The magnetic levitated turbo pump hangs below the column, a scroll pump used for backing is not pictured.
+      The laser system is in the background.
+      All of the above is mounted on a custom shaped active vibration-canceling optical table.
+      The electronic control systems are not pictured.
+    }
+    \label{fig:column-pic}
+  \end{figure}
+
   \subimport{inc/laser/}{laser}
 
-\section{Chamber Design}
+  \subimport{inc/hardware/}{column}
 
-\section{Accelerator Design}
-
-\section{Magnetic Lenses}
+  \subimport{inc/gun/}{gun_design}
 
@@
 
-\section{Deflection Plates}
+  \subimport{inc/hardware/}{deflector}
 
@@
 
+  \subimport{inc/photocathode/setup/}{setup}
+
 \section{Flat Metal Photocathodes}
 
-\section{Nanopatterned (Plasmonic) Photocathodes}
+  \subimport{inc/photocathode/plasmon/}{plasmon}
 
-\section{Semiconductor Photocathodes}
-
-  \subimport{inc/semicond/}{semicond}
+  \subimport{inc/photocathode/este/}{este}
 
@@
 
+%uncomment when multiple appedicies are used 
 %\appendices
-%\appendix
-%\newpage
+\appendix
+  \subimport{inc/extension/implementation/}{appendix-script}
+
+\newpage
 
@@
   \bibformc
-  \bibliography{bibliography,biblaser}
+  \bibliography{bibliography,biblaser,bibeste}
 }

### 2013-01-31 inc/gun/gun_design.tex
added anode and cathode pictures

@@
 Also, in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the accelerating electic field given that it will span a wider proportion of the accelerator than thin beams used in standard TEM.
+
+\begin{figure}
+  \centering
+  \centerline{\includegraphics{cathode.jpg} \includegraphics{anode.jpg}}
+  \caption{
+    Pictured are the cathode (left) and anode (right) of the accelerator built at UIC from a design by Togawa \protect\textit{et al.} \protect\cite{togawa_ceb6_2007}.
+    The photocathode is visible in the cathode Wehnelt aperture.
+    The anode is viewed as installed in the column, through the opened port normally occupied by the cathode and high voltage feedthrough.
+    The white material seen in both images helps to prevent electrical arcing under high voltage.
+  }
+  \label{fig:togawa-pic}
+\end{figure}
+
+These requirements lead to the fabrication of cathode/Wehnelt and anode pair based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007} (pictured in \ref{fig:togawa-pic}).
+A schematic of this design, overlayed by a Finite Element Model (FEM) simulation of the electic field, is shown in \ref{fig:gun-field}.
+The schematic shows that the aperture concerns are easily satisfied.
+From the field equipotentials, it is clear that this design has a consistently flat electric field for large displacements off the central axis.
 
@@
 \end{figure}
-
-These requirements lead to the fabrication of cathode/Wehnelt and anode pair based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007}.
-A schematic of this design, overlayed by a Finite Element Model (FEM) simulation of the electic field, is shown in \ref{fig:gun-field}.
-The schematic shows that the aperture concerns are easily satisfied.
-From the field equipotentials, it is clear that this design has a consistently flat electric field for large displacements off the central axis.
 

### 2013-01-31 inc/model_results/spacecharge/spacecharge.tex
a little reorg to start free-space space-charge section

\subsection{Effect of Space Charge} \label{sec:free_spacecharge}

### 2013-01-31 inc/model_results/spacecharge/spacecharge.pl
found problem in spacecharge plot simulation

@@ -38,13 +38,13 @@ sub for_number {
 
 open my $fh, '>', 'spacecharge.dat';
 
-my @nums = map { ("1e$_", "2.2e$_", "4.6e$_") } (0..9);
+my @nums = map { ("1e$_", "2.2e$_", "4.6e$_") } (0..8);
 pop @nums for 1..2;
 
 for my $num ( @nums ) {
   my $result = for_number($num);
   for (@$result) {
-    next unless $_->[1] >= 0.015;
+    next unless $_->[1] >= 0.15;
     print {$fh} $num . ' ' . sqrt( 2 * $_->[3] ) * 1e3 . ' ' . sqrt( 2 * $_->[4] ) * 1e3 . "\n";
     last;
   }

### 2013-01-31 inc/model_results/shapes/shapes.pl
began work on generating data for initial pulse shape plot

@@ -1,0 +1,56 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+use Physics::UEMColumn alias => ':standard';
+use Physics::UEMColumn::Auxiliary ':constants';
+
+sub gen_xi_series {
+  my ($n, $vol, $eta_t, $length) = @_; 
+
+  return sub {
+    my $xi = shift;
+
+    my $wt = ( $vol / $xi ) ** (1/3);
+    my $wz = ( $xi ** 2 * $vol ) ** (1/3);
+
+    my $pulse = Pulse->new(
+      velocity => '1e8 m/s',
+      number   => $n,
+      gamma_t  => 0,
+      gamma_z  => 0,
+      eta_t    => $eta_t,
+      eta_z    => $eta_t / 4,
+      sigma_t  => $wt**2 / 2,
+      sigma_z  => $wz**2 / 2,
+    );
+
+    my $column = Column->new(
+      length => $length, 
+    );
+
+    my $sim = Physics::UEMColumn->new(
+      column => $column,
+      pulse  => $pulse,
+    );
+
+    my $result = $sim->propagate;
+    return $result;
+  }
+}
+
+my $length = 0.15; # 15 cm
+my $eta_t = me * qe * (0.5) / 3; # 0.5 eV excess energy
+my $vol = (1e-3)**3;
+
+my $for_xi = gen_xi_series( 1, $vol, $eta_t, $length );
+
+for my $xi ( 1/3, 1, 3 ) {
+  my $result = $for_xi($xi);
+  for (@$result) {
+    next unless $_->[1] >= $length;
+    print {$fh} $num . ' ' . sqrt( 2 * $_->[3] ) * 1e3 . ' ' . sqrt( 2 * $_->[4] ) * 1e3 . "\n";
+    last;
+  }
+}

### 2013-02-01 inc/model_results/shapes/shapes.pl
generate data for shapes plot

@@ -3,7 +3,7 @@
 use strict;
 use warnings;
 
-use Physics::UEMColumn alias => ':standard';
+use Physics::UEMColumn alias => [qw':standard Pulse'];
 use Physics::UEMColumn::Auxiliary ':constants';
 
 sub gen_xi_series {
@@ -40,17 +40,23 @@ sub gen_xi_series {
   }
 }
 
+my $prefix = 'shape_';
+
 my $length = 0.15; # 15 cm
 my $eta_t = me * qe * (0.5) / 3; # 0.5 eV excess energy
 my $vol = (1e-3)**3;
 
 my $for_xi = gen_xi_series( 1, $vol, $eta_t, $length );
 
-for my $xi ( 1/3, 1, 3 ) {
-  my $result = $for_xi($xi);
+for my $spec ( [1/3, 'prolate' ], [1.0001, 'sphere'], [3, 'oblate'] ) {
+  my ( $xi, $shape ) = @$spec;
+  my $file = "${prefix}${shape}.dat";
+  open my $fh, '>', $file or die "Cannot open $file: $!\n";
+
+  my $result = $for_xi->($xi);
+  my ($st0, $sz0) = @{$result->[0]}[3,4];
   for (@$result) {
-    next unless $_->[1] >= $length;
-    print {$fh} $num . ' ' . sqrt( 2 * $_->[3] ) * 1e3 . ' ' . sqrt( 2 * $_->[4] ) * 1e3 . "\n";
-    last;
+    print {$fh} $_->[1] . ' ' . sqrt( $_->[3] / $st0 ) . ' ' . sqrt( $_->[4] / $sz0 ) . "\n";
+    last if $_->[1] >= $length;
   }
 }

### 2013-02-04 inc/model_results/shapes/shapes.tex
added shapes plot

--- 2011-06-17 main.tex
@@
-\documentclass{uicthesi}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{url}
-%\usepackage{euler}
+\subsection{Trends for Initial Shapes}
 
-\usepackage{graphicx}
-\graphicspath{{./images/}}
-\DeclareGraphicsExtensions{.png,.pdf,.jpg}
+\begin{figure}
 
-\title{Title}
-\author{Joel Berger}
-\pdegrees{of the\\University of Illinois at Chicago}
-\degree{Doctor of Philosophy in Experimental Physics}
+\include{figure_shapes}
 
-\begin{document}
-\maketitle
-
-%\copyrightpage
-\newpage
-\begin{figure}
-  \centering
-  \includegraphics{by-nc-nd}
 \end{figure}
-This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
-To view a copy of this license, visit \url{http://creativecommons.org/licenses/by-nc-nd/3.0/us/} or send a letter to
-\begin{verbatim}
-Creative Commons
-444 Castro Street, Suite 900
-Mountain View, CA, 94041, USA.
-\end{verbatim}
-
-\dedication
-The dedication.
-
-\acknowledgment
-The acknowledgment.
-\initials{JAB}
-
-\preface
-The preface.
-
-\tableofcontents
-\listoftables
-\listoffigures
-
-\listofabbreviations
-%\begin{list}
-%{}
-%{\setlength
-%   {\labelwidth}{1in}
-%    \setlength{\leftmargin}{1.5in}
-%    \setlength{\labelsep}{.5in}
-%    \setlength{\rightmargin}{\leftmargin}}
-%\item[AMS\hfill] American Mathematical Society
-%\item[CTAN\hfill] Comprehensive \TeX\ Archive Network
-%\item[TUG\hfill] \TeX\ Users Group
-%\item[UIC\hfill] University of Illinois at Chicago
-%\item[UICTHESI\hfill] Thesis formatting system for use at UIC.
-%\end{list}
-
-\summary
-A summary is required.
-
-\chapter{Introduction}
-
-%\appendices
-%\appendix
-%\newpage
-%\input fontsym
-%\bibformb
-%\bibliography{uictman}
-%\vita
-%This is where the vita goes.  Its organization is left as an exercise.
-%Hint: see the list of abbreviations.
-\end{document}

### 2013-02-04 inc/model_results/lens_charge/lens_charge.pl
started a lens length vs charge plot

@@ -1,0 +1,68 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+use v5.10;
+
+use Physics::UEMColumn alias => ':standard';
+
+use PDL;
+use PDL::Graphics::Prima::Simple;
+
+sub gen_sim {
+  my ($lens_position, $strength) = @_;
+  my $column_length = 3 * $lens_position;
+
+  return sub {
+    my $num = shift;
+    my $laser = Laser->new(
+      width    => '1 mm',
+      duration => '1 ps',
+      energy   => '4.75 eV',
+    );
+
+    my $acc = DCAccelerator->new(
+      length  => '20 mm',
+      voltage => '20 kilovolts',
+    );
+
+    my $column = Column->new(
+      length       => $column_length, 
+      laser        => $laser,
+      accelerator  => $acc,
+      photocathode => Photocathode->new( work_function => '4.25 eV' ), # Ta
+    );
+
+    my $sim = Physics::UEMColumn->new(
+      column => $column,
+      number => $num,
+    );
+
+    my $lens1 = MagneticLens->new(
+      location => $lens_position,
+      length   => '1 in',
+      strength => $strength,
+    );
+
+    $sim->add_element($lens1);
+
+    return pdl $sim->propagate;
+  }
+}
+
+my $out;
+my $sim = gen_sim( 0.20, 26.5e-13 );
+
+for ( qw/ 1e0 1e4 1e5 1e6 1e7 / ) {
+  my $result = $sim->($_);
+  my $z  = $result->slice('(1),');
+  my $st = $result->slice('(3),');
+
+  wcols $z, $st, "lens_20cm_n$_.dat";
+
+  my $min_ind = $st->minimum_ind;
+  say $z->at($min_ind) / 0.40;
+}
+
+#line_plot $z, $st;
+

### 2013-02-05 inc/model_results/shapes/shapes.tex
updated shapes plot
work on shape comparison text

@@
 
+The initial `space-time' eccentricity \ref{eq:define_xi}, $\xi ( 0 ) = v_{{ \scriptscriptstyle 0}} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
+\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and `cigar-like' ($ \sigma_{ z } ( 0 ) > \sigma_{ \smallT } ( 0 ) $) pulses.
+%TODO rewrite using w?
+Unlike most simulations in this paper, the pulses in this simulation are not ``generated'' from laser parameters, but are defined directly.
+The initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ while in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
+Since the initial pulse volumn is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_N106} ($ N = 10^{6} $ electrons in the bunch) than in \ref{fig:compare_shape}\subref{fig:compare_N1} ($ N = 10^{0} $).
+The pulse velocity is $ v_{{ \scriptscriptstyle 0}} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.
+
 \begin{figure}
+  \centering
+  \include{figure_shapes}
+  \caption{
+    Transverse (solid) and longitudinal (dashed) dynamics of three example pulses for three length/width ratios ($\xi$). 
+    The red lines represent a prolate pulse (a ``pancake'') with $\xi=1/3$, green lines for a spherical pulse with $\xi=1$, and blue for oblate (a ``cigar'') having $\xi=3$.
+  }
+  \label{fig:compare_shape}
+\end{figure}
 
-\include{figure_shapes}
+%\begin{figure}[t]
+%\centering
+%\subfloat[][]
+%{
+%
+%\label{fig:compare_N1}
+%\includegraphics{compare_N1}
+%}
+%\\
+%\subfloat[][]
+%{
+%
+%\label{fig:compare_N106}
+%\includegraphics{compare_N106}
+%}
+%\caption
+%{
+%Comparison of the propagation dynamics of disk-like ($ \xi (0) = 0.1 $), spherical ($ \xi (0) = 1 $) and cigar-like ($ \xi (0) = 10 $) pulses, generated from a 261nm laser driven Ta photocathode in a 20kV DC electron gun, at \subref{fig:compare_N1} negligible charge densities ($ N = 10^{0} $) and \subref{fig:compare_N106} charge densities suitable for single shot UED studies ($ N = 10^{6} $): the normalized transverse beam width ($ \sqrt{ \sigma_{\smallT} / \sigma_{\smallT} ( 0 ) } $) (solid line) and normalized longitudinal pulse width ($ \sqrt{ \sigma_{z} / \sigma_{z} ( 0 ) } $) (dashed line).
+%}
+%\label{fig:compare_shape}
+%\end{figure}
 
-\end{figure}
+%The bunch propagation dynamics for negligible pulse charge density ($ N = 10^{0} $) shown in \Fig \ref{fig:compare_shape}\subref{fig:compare_N1} display the trends expected from the initial photoemission conditions.
+In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{i} $ and the zero initial pulse chirp ($ \gamma_{i} = 0 $).
+Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{i} / \sigma_{i} (0) } $ for comparable $ \eta_{i} $ is greatest for the smallest initial $ \sigma_{i} (0) $; that is both disk- and cigar-like pulses are initially driven towards the spherical regime.
+In all cases, this intrinsic electron pulse broadening due to non-zero $ \eta_{i} $ generates a momentum chirp across the electron bunch; in the longitudinal dimension the faster electrons photoemitted with larger excess energy leading the slower electrons.
+
+%For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects can dominate the propagation dynamics.
+In \Fig \ref{fig:compare_shape}\subref{fig:compare_N106}, we display the predicted temporal dynamics of electron pulses with the same initial size and photoemission conditions as \Fig \ref{fig:compare_shape}\subref{fig:compare_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
+%Above is probably true for UED but never done
+%The significant influence of space-charge effects for the initial peak pulse charge density of $0.0287\text{C}/\text{m}^{3}$ is clearly evident.
+In the spherical case, space-charge induced bunch broadening overwhelms the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing an almost uniform three-dimensional `Coulomb explosion' of the electron pulse with a one-dimensional bunch broadening rate roughly one order of magnitude greater than that for $ N = 10^{0} $ (\Fig \ref{fig:compare_shape}\subref{fig:compare_N1}).
+For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation.
+This is to be expected, because the initial $\eta_{i}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
+In fact, both the $ N = 10^{6} $ disk- and cigar-like pulse examples shown in \Fig \ref{fig:compare_shape}\subref{fig:compare_N106} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
+The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{i} (0) } $.
+
+%The data presented in \Fig \ref{fig:compare_shape} using the AG model of Michalik and Sipe \cite{michalik_analytic_2006} are entirely consistent with prior works.\cite{reed_femtosecond_2006,siwick_ultrafast_2002} In femtosecond UED experiments, significant efforts are made to offset space-charge induced temporal (longitudinal, $z$) pulse broadening by reducing the time of flight from the gun photocathode to the sample.\cite{siwick_ultrafast_2002,reed_evolution_2009} Moreover, efforts are now underway to compensate for temporal electron pulse broadening using RF cavity pulse compression techniques.\cite{oudheusden_electron_2007} In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers,\cite{lobastov_four-dimensional_2005} deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
+Even for nanosecond DTEM (the extreme cigar-like pulse ($\xi \gg 1 $)), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008} In this paper, however, we will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the disk-like regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\Fig \ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}

### 2013-02-05 inc/photocathode/plasmon/plasmon.tex
some bugfixes

@@
 Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
-Given a laser wavelength of 523nm ($\varepsilon_{Au} \approx –3.95$ \cite{johnson_optical_1972}), this choice of periodicity corresponds by \ref{eq:plasmon_condition} to an incidence angle of $\sim39^{\circ}$, which is within the acceptance range of our experimental setup (35-75$^{\circ}$).
+Given a laser wavelength of 523nm ($\varepsilon_{Au} \approx -3.95$ \cite{johnson_optical_1972}), this choice of periodicity corresponds by \ref{eq:plasmon_condition} to an incidence angle of $\sim39^{\circ}$, which is within the acceptance range of our experimental setup (35-75$^{\circ}$).
 %TODO point to Togawa section

### 2013-02-05 inc/model_results/shapes/shapes.tex
added high charge shape plot and continue with text

--- 2013-02-04 inc/model_results/shapes/shapes.tex
@@
 
+The initial `space-time' eccentricity \ref{eq:define_xi}, $\xi ( 0 ) = v_{{ \scriptscriptstyle 0}} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
+\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and `cigar-like' ($ \sigma_{ z } ( 0 ) > \sigma_{ \smallT } ( 0 ) $) pulses.
+%TODO rewrite using w?
+Unlike most simulations in this paper, the pulses in this simulation are not ``generated'' from laser parameters, but are defined directly.
+The initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ while in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
+Since the initial pulse volumn is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} ($ N = 10^{6} $ electrons in the bunch) than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} ($ N = 10^{0} $).
+The pulse velocity is $ v_{{ \scriptscriptstyle 0}} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.
+
 \begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:compare_shape_N1}
+    \input{shapes_N1}
+  }
+  \\
+  \subfloat[][]{
+    \label{fig:compare_shape_N1e6}
+    \input{shapes_N1e6}
+  }
+  \caption{
+    Transverse (solid) and longitudinal (dashed) dynamics of three example pulses for three length/width ratios ($\xi$). 
+    The red lines represent a prolate pulse (a ``pancake'') with $\xi=1/3$, green lines for a spherical pulse with $\xi=1$, and blue for oblate (a ``cigar'') having $\xi=3$.
+  }
+  \label{fig:compare_shape}
+\end{figure}
 
-\include{figure_shapes}
+%\begin{figure}[t]
+%\centering
+%\subfloat[][]
+%{
+%
+%\label{fig:compare_shape_N1}
+%\includegraphics{compare_N1}
+%}
+%\\
+%\subfloat[][]
+%{
+%
+%\label{fig:compare_shape_N1e6}
+%\includegraphics{compare_N106}
+%}
+%\caption
+%{
+%Comparison of the propagation dynamics of disk-like ($ \xi (0) = 0.1 $), spherical ($ \xi (0) = 1 $) and cigar-like ($ \xi (0) = 10 $) pulses, generated from a 261nm laser driven Ta photocathode in a 20kV DC electron gun, at \subref{fig:compare_shape_N1} negligible charge densities ($ N = 10^{0} $) and \subref{fig:compare_shape_N1e6} charge densities suitable for single shot UED studies ($ N = 10^{6} $): the normalized transverse beam width ($ \sqrt{ \sigma_{\smallT} / \sigma_{\smallT} ( 0 ) } $) (solid line) and normalized longitudinal pulse width ($ \sqrt{ \sigma_{z} / \sigma_{z} ( 0 ) } $) (dashed line).
+%}
+%\label{fig:compare_shape}
+%\end{figure}
 
-\end{figure}
+%The bunch propagation dynamics for negligible pulse charge density ($ N = 10^{0} $) shown in \Fig \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
+In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{i} $ and the zero initial pulse chirp ($ \gamma_{i} = 0 $).
+Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{i} / \sigma_{i} (0) } $ for comparable $ \eta_{i} $ is greatest for the smallest initial $ \sigma_{i} (0) $; that is both disk- and cigar-like pulses are initially driven towards the spherical regime.
+In all cases, this intrinsic electron pulse broadening due to non-zero $ \eta_{i} $ generates a momentum chirp across the electron bunch; in the longitudinal dimension the faster electrons photoemitted with larger excess energy leading the slower electrons.
+
+%For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects can dominate the propagation dynamics.
+In \Fig \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, we display the predicted temporal dynamics of electron pulses with the same initial size and photoemission conditions as \Fig \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
+%Above is probably true for UED but never done
+%The significant influence of space-charge effects for the initial peak pulse charge density of $0.0287\text{C}/\text{m}^{3}$ is clearly evident.
+In the spherical case, space-charge induced bunch broadening overwhelms the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing an almost uniform three-dimensional `Coulomb explosion' of the electron pulse with a one-dimensional bunch broadening rate roughly one order of magnitude greater than that for $ N = 10^{0} $ (\Fig \ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
+For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation.
+This is to be expected, because the initial $\eta_{i}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
+In fact, both the $ N = 10^{6} $ disk- and cigar-like pulse examples shown in \Fig \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
+The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{i} (0) } $.
+
+%The data presented in \Fig \ref{fig:compare_shape} using the AG model of Michalik and Sipe \cite{michalik_analytic_2006} are entirely consistent with prior works.\cite{reed_femtosecond_2006,siwick_ultrafast_2002} In femtosecond UED experiments, significant efforts are made to offset space-charge induced temporal (longitudinal, $z$) pulse broadening by reducing the time of flight from the gun photocathode to the sample.\cite{siwick_ultrafast_2002,reed_evolution_2009} Moreover, efforts are now underway to compensate for temporal electron pulse broadening using RF cavity pulse compression techniques.\cite{oudheusden_electron_2007} In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers,\cite{lobastov_four-dimensional_2005} deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
+Even for nanosecond DTEM (the extreme cigar-like pulse ($\xi \gg 1 $)), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008} In this paper, however, we will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the disk-like regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\Fig \ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}

### 2013-02-05 inc/model_results/shapes/shapes.pl

@@ -6,6 +6,27 @@ use warnings;
 use Physics::UEMColumn alias => [qw':standard Pulse'];
 use Physics::UEMColumn::Auxiliary ':constants';
 
+my $n = shift or die "Please specify a number of electrons";
+
+my $length = 0.15; # 15 cm
+my $eta_t = me * qe * (0.5) / 3; # 0.5 eV excess energy
+my $vol = (1e-3)**3;
+
+my $for_xi = gen_xi_series( $n, $vol, $eta_t, $length );
+
+for my $spec ( [1/3, 'prolate' ], [1.0001, 'sphere'], [3, 'oblate'] ) {
+  my ( $xi, $shape ) = @$spec;
+  my $file = "shape_${shape}_N$n.dat";
+  open my $fh, '>', $file or die "Cannot open $file: $!\n";
+
+  my $result = $for_xi->($xi);
+  my ($st0, $sz0) = @{$result->[0]}[3,4];
+  for (@$result) {
+    print {$fh} $_->[1] . ' ' . sqrt( $_->[3] / $st0 ) . ' ' . sqrt( $_->[4] / $sz0 ) . "\n";
+    last if $_->[1] >= $length;
+  }
+}
+
 sub gen_xi_series {
   my ($n, $vol, $eta_t, $length) = @_; 
 
@@ -40,23 +61,3 @@ sub gen_xi_series {
   }
 }
 
-my $prefix = 'shape_';
-
-my $length = 0.15; # 15 cm
-my $eta_t = me * qe * (0.5) / 3; # 0.5 eV excess energy
-my $vol = (1e-3)**3;
-
-my $for_xi = gen_xi_series( 1, $vol, $eta_t, $length );
-
-for my $spec ( [1/3, 'prolate' ], [1.0001, 'sphere'], [3, 'oblate'] ) {
-  my ( $xi, $shape ) = @$spec;
-  my $file = "${prefix}${shape}.dat";
-  open my $fh, '>', $file or die "Cannot open $file: $!\n";
-
-  my $result = $for_xi->($xi);
-  my ($st0, $sz0) = @{$result->[0]}[3,4];
-  for (@$result) {
-    print {$fh} $_->[1] . ' ' . sqrt( $_->[3] / $st0 ) . ' ' . sqrt( $_->[4] / $sz0 ) . "\n";
-    last if $_->[1] >= $length;
-  }
-}

### 2013-02-05 inc/model_results/shapes/shapes.tex
finished importing JAP paper shape section, it still needs some polish

--- 2013-02-04 inc/model_results/shapes/shapes.tex
@@
 
+The initial `space-time' eccentricity \ref{eq:define_xi}, $\xi ( 0 ) = v_{{ \scriptscriptstyle 0}} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
+\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and `cigar-like' ($ \sigma_{ z } ( 0 ) > \sigma_{ \smallT } ( 0 ) $) pulses.
+%TODO rewrite using w?
+Unlike most simulations in this paper, the pulses in this simulation are not calculated from laser and photocathode parameters, but are rather defined directly.
+The initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ while in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
+Since the initial pulse volumn is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} ($ N = 10^{6} $ electrons in the bunch) than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} ($ N = 10^{0} $).
+The pulse velocity is $ v_{{ \scriptscriptstyle 0}} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.
+
 \begin{figure}
+  \centering
+  \subfloat[][]{
+    \label{fig:compare_shape_N1}
+    \input{shapes_N1}
+  }
+  \\
+  \subfloat[][]{
+    \label{fig:compare_shape_N1e6}
+    \input{shapes_N1e6}
+  }
+  \caption{
+    Transverse (solid) and longitudinal (dashed) dynamics of three example pulses for three length/width ratios ($\xi$). 
+    The red lines represent a prolate pulse (a ``pancake'') with $\xi=1/3$, green lines for a spherical pulse with $\xi=1$, and blue for oblate (a ``cigar'') having $\xi=3$.
+  }
+  \label{fig:compare_shape}
+%\caption
+%{
+%Comparison of the propagation dynamics of disk-like ($ \xi (0) = 0.1 $), spherical ($ \xi (0) = 1 $) and cigar-like ($ \xi (0) = 10 $) pulses, generated from a 261nm laser driven Ta photocathode in a 20kV DC electron gun, at \subref{fig:compare_shape_N1} negligible charge densities ($ N = 10^{0} $) and \subref{fig:compare_shape_N1e6} charge densities suitable for single shot UED studies ($ N = 10^{6} $): the normalized transverse beam width ($ \sqrt{ \sigma_{\smallT} / \sigma_{\smallT} ( 0 ) } $) (solid line) and normalized longitudinal pulse width ($ \sqrt{ \sigma_{z} / \sigma_{z} ( 0 ) } $) (dashed line).
+%}
+\end{figure}
 
-\include{figure_shapes}
+The bunch propagation dynamics for negligible pulse charge density ($ N = 10^{0} $) shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
+%In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the zero initial pulse chirp ($ \gamma_{\alpha} = 0 $).
+In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands more rapidly than the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ ( with zero initial pulse chirp, i.e.
+$ \gamma_{\alpha} = 0 $).
+Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both disk- and cigar-like pulses are initially driven towards the spherical regime.
+In all cases, this intrinsic electron pulse broadening due to non-zero $ \eta_{\alpha} $ generates a momentum chirp across the electron bunch; in the longitudinal dimension the faster electrons photoemitted with larger excess energy leading the slower electrons.
 
-\end{figure}
+For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects can dominate the propagation dynamics.
+In \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, we display the predicted temporal dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
+The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
+%TODO revisit this next sentance, the data doesn't seem to bear it out
+In the spherical case, space-charge induced bunch broadening overwhelms the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing an almost uniform three-dimensional `Coulomb explosion' of the electron pulse with a one-dimensional bunch broadening rate roughly one order of magnitude greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
+For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation.
+This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
+In fact, both the $ N = 10^{6} $ disk- and cigar-like pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
+The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.
+
+The data presented in \ref{fig:compare_shape} using the AG model of Michalik and Sipe \cite{michalik_analytic_2006} are entirely consistent with prior works.\cite{reed_femtosecond_2006,siwick_ultrafast_2002}
+In femtosecond UED experiments, significant efforts are made to offset space-charge induced temporal (longitudinal, $z$) pulse broadening by reducing the time of flight from the gun photocathode to the sample.\cite{siwick_ultrafast_2002,reed_evolution_2009}
+Moreover, efforts are now underway to compensate for temporal electron pulse broadening using RF cavity pulse compression techniques.\cite{oudheusden_electron_2007}
+In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers,\cite{lobastov_four-dimensional_2005} deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
+Even for nanosecond DTEM (the extreme cigar-like pulse ($\xi \gg 1 $)), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
+In this paper, however, we will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the disk-like regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}

### 2013-02-07 inc/model_results/shapes/shapes.tex
rephrase about spherical pulse increase

@@
 %TODO rewrite using w?
-Unlike most simulations in this paper, the pulses in this simulation are not calculated from laser and photocathode parameters, but are rather defined directly.
+For simplicity, and unlike most simulations in this paper, the pulses in this simulation are not calculated from laser and photocathode parameters, but are rather defined directly.
 The initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ while in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
@@
   \label{fig:compare_shape}
-%\caption
-%{
-%Comparison of the propagation dynamics of disk-like ($ \xi (0) = 0.1 $), spherical ($ \xi (0) = 1 $) and cigar-like ($ \xi (0) = 10 $) pulses, generated from a 261nm laser driven Ta photocathode in a 20kV DC electron gun, at \subref{fig:compare_shape_N1} negligible charge densities ($ N = 10^{0} $) and \subref{fig:compare_shape_N1e6} charge densities suitable for single shot UED studies ($ N = 10^{6} $): the normalized transverse beam width ($ \sqrt{ \sigma_{\smallT} / \sigma_{\smallT} ( 0 ) } $) (solid line) and normalized longitudinal pulse width ($ \sqrt{ \sigma_{z} / \sigma_{z} ( 0 ) } $) (dashed line).
-%}
 \end{figure}
@@
 The bunch propagation dynamics for negligible pulse charge density ($ N = 10^{0} $) shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
-%In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the zero initial pulse chirp ($ \gamma_{\alpha} = 0 $).
-In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands more rapidly than the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ ( with zero initial pulse chirp, i.e.
-$ \gamma_{\alpha} = 0 $).
+In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands roughly twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp ($ \gamma_{\alpha} = 0 $).
 Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both disk- and cigar-like pulses are initially driven towards the spherical regime.

### 2013-02-07 inc/magnetic_lenses/maglens.tex
a little reorg to start free-space space-charge section
work on short titles for figures (continuing)

@@
 Second, and more importantly, near a lens crossover (i.e. focus), the charge density can increase rapidly.
-This charge density can lead to an unacceptable amount of space-charge interaction (see section \ref{sec:model_results}).
+This charge density can lead to an unacceptable amount of space-charge interaction (see section \ref{sec:free_spacecharge}).
 
@@
   \input{polepiece-schematic}
-  \caption{
+  \caption[Side-view schematic of a shielded magnetic lens with a pole-piece]{
     Side-view schematic of a shielded magnetic lens with a pole-piece.
@@
   \includegraphics{maglens.jpg}
-  \caption{
-    Side view of the two custom large-bore magnetic lenses designed and built at UIC, seen installed in the prototype UEM.
+  \caption[Side-view picture of the employed two custom large-bore magnetic lenses]{
+    Side-view of the two custom large-bore magnetic lenses designed and built at UIC, seen installed in the prototype UEM.
     The coils are visible as is high permeability shielding used for the pole-piece.

### 2013-02-07 inc/photocathode/este/este.tex
work on short titles for figures (continuing)

@@
   \input{fig2}
-  \caption{
+  \caption[Measured Fourier plane electron beam size and yield vs incident laser pulse energy on gold photocathode]{
     Measured Fourier plane electron beam size (HWe$^{-1}$M; squares) and electron yield (circles) as a function of the incident $\sim$200fs, 523nm laser pulse energy for the 300nm-thick gold photocathode.
@@
   \input{fig3}
-  \caption{
+  \caption[Electron emission vs UV laser pulse energy for GaSb and InSb]{
     Number of electrons emitted per pulse as a function of the incident 261nm, UV laser pulse energy for GaSb (open squares) and InSb (filled circles): Linear efficiency dependences are shown by the solid lines.

### 2013-02-07 inc/photocathode/plasmon/plasmon.tex
work on short titles for figures (continuing)

@@
   \input{grating_schematic}
-  \caption{Schematic diagram of a grating coupling geometry for driving a surface plasmon}
+  \caption[Grating coupling geometry for driving a surface plasmon]{Schematic diagram of a grating coupling geometry for driving a surface plasmon}
   \label{fig:plasmon_schematic}
@@
   \includegraphics{damage.png}
-  \caption{Damage to commerical gold-on-glass holographic grating.}
+  \caption{Damage to commerical gold-on-glass holographic grating}
   \label{fig:grating-damage}
@@
   \includegraphics{HighMagSEM.jpg}
-  \caption{
+  \caption[SEM image of sinusoidal grating on silicon]{
     SEM image of sinusoidal grating on silicon.
@@
   \input{power_seq}
-  \caption{
+  \caption[Photoemission characteristics of laser-driven plasmons on a one-dimensional periodic gold surface]{
     Photoemission characteristics of laser-driven plasmons on a one-dimensional periodic gold surface.
@@
   \input{rotated}
-  \caption{
+  \caption[Comparison of the colinear and rotated plasmonic excitation geometries]{
     Comparison of the colinear (above) and rotated (below) plasmonic excitation geometries.

### 2013-02-08 inc/gun/gun_design.tex
added aligner picture
added short titles for the remaining figures

@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \section{Accelerator Design} \label{sec:gun_design}
@@
   \centerline{\includegraphics{cathode.jpg} \includegraphics{anode.jpg}}
-  \caption{
+  \caption[Pictures of the employed Togawa-inspired custom cathode-anode pair]{
     Pictured are the cathode (left) and anode (right) of the accelerator built at UIC from a design by Togawa \protect\textit{et al.} \protect\cite{togawa_ceb6_2007}.
@@
   \input{gunfield}
-  \caption{
-    Schematic of the employed Togawa inspired cathode ($-V$) and anode (0V) pair.
+  \caption[Schematic of the employed Togawa-inspired custom cathode-anode pair]{
+    Schematic of the employed Togawa-inspired cathode ($-V$) and anode (0V) pair.
     The emission source is inserted behind the cathode Wehnelt.
@@
 Thus, by iteratively repositioning the input laser and then the gun alignment, the beam may be walked to pass as near as possible down the axis of the column.
+
+\begin{figure}
+  \centering
+  \includegraphics{aligner.jpg}
+  \caption[Picture of the high-voltage side of the prototype UEM]{
+    The high voltage passthrough (left) is mounted on the main chamber (right) by a ``port aligner'' which has been modified to allow coarse alignment of the cathode (mounted to the high voltage) and anode (mounted in the chamber).
+  }
+  \label{fig:aligner-pic}
+\end{figure}

### 2013-02-08 inc/gun/gun_model.tex
began describing the Togawa gun
more gun design
added short titles for the remaining figures

@@
-\subsection{Acceleration Region}
+\subsection{Acceleration Region} \label{sec:gun_model}
 
@@
   }
-  \caption{
+  \caption[An example of a tahn form acceleration region for three different ``sharpness'' values]{
     An example of a tahn form acceleration region (overall length 6.35mm) for three different ``sharpness'' values.
@@
   }
-  \label{fig:laser-cavity}
+  \label{fig:gun-sharpness}
 \end{figure}
 
-\ref{fig:laser-cavity} shows an example acceleration region for three values of the ``sharpness'' $s$.
+\ref{fig:gun-sharpness} shows an example acceleration region for three values of the ``sharpness'' $s$.
 What is clear from this figure is that the dynamics are not highly coupled to the value of this hard-to-quantify parameter.

### 2013-02-08 inc/laser/laser.tex
added short titles for the remaining figures

@@
   }
-  \caption{
+  \caption[Laser crystal and anti-reflection coating]{
     \protect\subref{fig:crystal_geo} Top view of the 3x3x10mm AR coated Yb:KGW laser crystal showing the crystal cut, the direction of propagation of p-polarized laser radiation at 1040nm, and the $x$ (tangential) and $y$ (saggital) directions (in blue) employed in the thermal analysis;
@@
   \includegraphics{thermal.jpg}
-  \caption{
+  \caption[Thermal profile and thermal lens shaping of KGW crystal pumping]{
     Analytical evaluation of the temperature distribution $T(x,y)$ due to the Gaussian/super-Gaussian heat source $Q(x,y)$ in the (a) horizontal ($y=0$) and (b) vertical ($x=0$) directions: $Q(x,y)$ - red-brown line; $T(x,y)$ - black line; parabolic approximation to $T(x,y)$ - dashed black line; representative TEM$_{00}$ laser field modes - solid green.
@@
   }
-  \caption{
+  \caption[Schematic and stability plot of employed Yb:KGW ultrafast laser]{
     \protect\subref{fig:laser_schematic} Schematic of the femtosecond TLS Yb:KGW oscillator (LD = laser diode, OC = output coupler, SBR = saturable Bragg reflector).
@@
   \includegraphics{profile.jpg}
-  \caption{
+  \caption[Autocorrelation measurements of employed ultrafast laser's pulse duration]{
 Second harmonic autocorrelation measurements and output spectra of the ML Yb:KGW laser; (a) the 250fs laser pulse duration (and fit assuming a sech$^2$ pulse shape) for the cavity depicted in \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} at an output power of 3.5W, (b) the 347fs pulse duration (and fit assuming a sech$^2$ pulse shape) obtained with an additional -5600fs$^2$ of round-trip GVD at an output power of 5W, (c) the 4.7nm full-width half maximum spectrum associated with the 250fs ML pulse of (a), and (d) spectral output for the ML resonator of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} at a laser power of 3.7W.
@@
   \includegraphics{performance.jpg}
-  \caption{
+  \caption[Power measuments of employed ultrast laser]{
 Performance of the diode-pumped, TLS, Yb:KGW laser (in the cavity configuration of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} ) as a function of applied diode current: continuous-wave (CW) output power (black squares); mode locked (ML) output power (red circles); frequency doubled power (green triangles); and second harmonic generation (SHG) efficiency (line).

### 2013-02-08 inc/model_results/spacecharge/spacecharge.tex
added spacecharge plot (beam size vs number of electrons)
added short titles for the remaining figures

@@
 \subsection{Effect of Space Charge} \label{sec:free_spacecharge}
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \input{spacecharge_data}
+  \end{tikzpicture}
+  \caption[Free-space pulse evolution vs initial shape]{
+    The final HW1/eM width (solid) and length (dashed) of similar initial size pulses for different number of electrons in the pulse.
+    The pulse is generated by 1 mm (HW1/eM) width 1 ps laser pulses, then traverse a 20mm acceleration region and a 15cm drift region.
+  }
+  \label{fig:spacecharge}
+\end{figure}

### 2013-02-09 inc/model_results/shapes/shapes.tex
added short titles for the remaining figures
continue discussion on shape/density

@@
-\subsection{Trends for Initial Shapes}
+\subsection{Trends for Initial Shapes} \label{sec:initial_shapes}
 
@@
   }
-  \caption{
+  \caption[AG simulation of free-space pulse expansion]{
     Transverse (solid) and longitudinal (dashed) dynamics of three example pulses for three length/width ratios ($\xi$). 
@@
 In this paper, however, we will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the disk-like regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}
+
+I have shown that a prolate spheroid pulse shape is preferred for UEM, both for generation in SECTION %TODO link to Child law discussion
+and in this section for propagation.
+Further I have shown in %TODO link to section where I show that initial conditions create shorter pulses than laser
+that electron pulses shorter than the laser pulse duration are likely occurances.
+Henceforth I shall only consider prolate pulses when a choice of initial pulse shape is possible.

### 2013-02-09 inc/model_results/spacecharge/spacecharge.tex
continue discussion on shape/density

--- 2013-01-31 inc/model_results/spacecharge/spacecharge.tex
@@
 \subsection{Effect of Space Charge} \label{sec:free_spacecharge}
+
+A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
+Much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements.
+
+Purely for illustrative purposes I am including \ref{fig:spacecharge} to provide an intuition about the order of magnitude of these effects.
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \input{spacecharge_data}
+  \end{tikzpicture}
+  \caption[Free-space pulse evolution vs charge density]{
+    The final HW1/eM width (solid) and length (dashed) of similar initial size pulses for different number of electrons in the pulse.
+    The pulse is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy. The pulse then traverses a 20mm acceleration region and a 15cm drift region.
+  }
+  \label{fig:spacecharge}
+\end{figure}

### 2013-02-12 inc/extension/initial_conditions/initial_conditions.tex
lots of surgery on the initial conditions

@@
 For the purposes of this paper, we will make a simple approximation to the behavior of the gun in order to provide generic initial conditions for the electron pulse.
+
+\subsection{A Simplistic Analytic Form}
 
@@
 \end{equation}
+A laser pulse of this form has a temporal uncertainty given simply $ ( \Delta t_{L} )^{ 2 } = \tau^{ 2 } / 2 $.
+
 For spatially-uniform photoemission from a planar photocathode, we have $ \gamma_{\smallT} ( 0 ) = 0 $; i.e., there is no initial spatial momentum chirp in the electron bunch.
@@
 % longitudinal
-In the longitudinal direction ($z$), the interplay between acceleration, $ a = e E_{DC}/m $, in the DC gun and the physics of the photoemission process determine the properties of the electron bunch.
-Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) by the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than that traveled by electrons with the maximum photoemission velocity $ v_{0} $ alone over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
+In the longitudinal direction ($z$), the interplay between acceleration, $ a = e E_{DC}/m $, in the DC gun and the physics of the photoemission process determine the properties $\sigma_z$ and $\gamma_z$ of the electron bunch.
+Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) by the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than that traveled by electrons with the maximum photoemission velocity $ v_{0} = \sqrt{ 2 \Delta E / m } $ alone over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
 
@@
 In the long-pulse regime, the DC field dominated case, we assume that the intrinsic velocity of the emitted electrons is negligable.
-Clearly the first emitted electrons travel a distance $ \Delta z = \frac{ 1 }{ 2 } a ( \Delta t_{L} )^{ 2 } = \sqrt{ \sigma_{ z } ( 0 ) } $, during the  $ \Delta t_{ L } $ rms laser pulse duration: for a laser pulse described above, we have $ ( \Delta t_{L} )^{ 2 } = \tau^{ 2 } / 2 $.
-This result may be restated as,
+The emitted electrons travel an average distance $ \Delta z $, during the $ \Delta t_{ L } $ rms laser pulse duration given by,
 \begin{equation}
-  \sigma_{ z } ( 0 ) = \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 }
+  \sqrt{ \sigma_{ z } ( 0 ) } = \Delta z = \frac{ 1 }{ 2 } a ( \Delta t_{L} )^{ 2 }  = \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m }
 \end{equation}
@@
 \begin{equation}
-  \gamma_{ z } ( 0 ) = m \sigma_{ z } ( 0 ) \frac{ \Delta v }{ \Delta z } = m a \Delta t_{ L } \sqrt{ \sigma_{ z } ( 0 ) } = \frac{ e E_{ DC } \tau }{ \sqrt{ 2 } }\, .
+  \gamma_{ z } ( 0 ) = m \Delta v \Delta z = m a \Delta t_{ L } \sqrt{ \sigma_{ z } ( 0 ) } = \frac{ e E_{ DC } \tau }{ \sqrt{ 2 } }\sqrt{ \sigma_{ z } ( 0 ) }\, .
 \end{equation}
-From a simplistic viewpoint, only the electrons generated with positive longitudinal momentum will contribute to the pulse since those electrons generated with negative momentum do not leave the photocathode.
-This implies that the local momentum uncertainty in the longitudinal direction is half that of the transverse directions; that is to say
+
+%v0 dominated
+In the short pulse limit, where the acceleration in the DC gun field over the incident laser pulse duration can be neglected, the initial electron bunch properties are determined primarily by the photoemission physics.
+Since the electric field has no time to influence the front of the pulse before the back is also generated we can say that the front of the pulse has a velocity equal to $v_0$ (the velocity associated with the excess photoemission energy) and the back is essentially stationary just after generation.
+The pulse's initial spatial variance $\Delta z$ is therefore equal to the distance that the front of the pulse travels during the laser pulse duration $\Delta t_L$,
+\begin{equation} \label{eq:initial_sigma_z}
+  \sqrt{ \sigma_{ z } ( 0 ) } = \Delta z = v_{ 0 } \Delta t_{L} = \frac{ v_{ 0 } \tau }{ \sqrt{ 2 } } \, \text{.}
+\end{equation}
+Once again we can use this information to determine a velocity chirp across the pulse in this regime,
+\begin{equation}
+  \gamma_{ z } ( 0 ) = m \Delta v \Delta z = m v_{ 0 } \sqrt{ \sigma_{ z } ( 0 ) } \, ,
+\end{equation}
+
+Finally, from a simplistic viewpoint, only the electrons generated with positive longitudinal momentum will contribute to the pulse since those electrons generated with negative momentum do not leave the photocathode.
+This implies that the final initial condition, the local momentum uncertainty in the longitudinal direction $\eta_z$, is half that of the transverse directions; that is to say
 \begin{equation}
@@
 
-%v0 dominated
-In the short pulse limit, where the acceleration in the DC gun field over the incident laser pulse duration can be neglected, the initial electron bunch properties are determined primarily by the photoemission physics.
-%TODO describe new model
-Careful analysis of the electron generation using the longitudinal $ v^{3} $ photoemission velocity distribution %TODO reference bin model
-and iterative time steps through the incident laser pulse reveals that, although the spatial electron bunch distribution is slightly skewed with respect to the symmetric incident laser pulse, its initial variance $ \sqrt{ \sigma_{ z } ( 0 ) } $ (or width) is nearly equal to that of the laser pulse of the normalization by the velocity ratio $ v_{ 0 }/c $; in other words,
-\begin{equation} \label{eq:initial_sigma_z}
-\sqrt{ \sigma_{ z } ( 0 ) } = v_{ 0 } \Delta t_{L} = \frac{ v_{ 0 } \tau }{ \sqrt{ 2 } } \, \text{.}
-\end{equation}
-Not unexpectedly, the analysis also indicates that the initial electron bunch has a velocity chirp.
-%TODO description again
-Its value is only roughly 6\% of $ v_{ 0 }/\sqrt{ \sigma_{z} ( 0 ) } $ since the photoemission generates the $ v^{ 3 } $ velocity distribution throughout the laser pulse; thus 
-\begin{equation}
-\gamma_{ z } ( 0 ) = m \sigma_{ z } ( 0 ) \frac{ d v }{ d z } \approx 0.06 m v_{ 0 } \sqrt{ \sigma_{ z } ( 0 ) } \, ,
-\end{equation}
-in this regime.
-Additionally the iterative time-step analysis clearly shows that the local longitudinal momentum variance $ \eta_{ z } $ is not constant over the inital photo-generated electron bunch.
-In fact, $ \sqrt{ \eta_{ z } ( 0 ) } $ varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch with the narrowest momentum distribution at the front of the bunch where only the faster electrons with velocity near $ v_{ 0 } $ initially reside.
-Unfortunately, the mean-field theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch.
-Consequently, we will assume that $ \eta_{ z } $ is a constant over the bunch with a value given by $ \eta_{\smallT } ( 0 ) \cong 3.85 \eta{ z } ( 0 ) $: %TODO reference bin model
-indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.
-
 These initial conditions at $t=0$ for an incident Gaussian laser pulse of the form given in \ref{eq:laser_form}, will be used for the remainder of this paper.
@@
     \gamma_{ \smallT } ( 0 ) = 0 \\
-    \gamma_{ z } ( 0 ) = \sqrt{ \sigma_{ z } ( 0 ) } \left ( 0.06 m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
+    \gamma_{ z } ( 0 ) = \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
     \eta_{ \smallT } ( 0 ) = \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \\
-    \eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 ) }{ 3.85 }
+    \eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 ) }{ 4 }
   \end{gather}
@@
 
+\subsection{Iterative Spatial/Temporal Bin Model}
+
+Additionally the iterative time-step analysis clearly shows that the local longitudinal momentum variance $ \eta_{ z } $ is not constant over the inital photo-generated electron bunch.
+In fact, $ \sqrt{ \eta_{ z } ( 0 ) } $ varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch with the narrowest momentum distribution at the front of the bunch where only the faster electrons with velocity near $ v_{ 0 } $ initially reside.
+Unfortunately, the mean-field theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch.
+Consequently, we will assume that $ \eta_{ z } $ is a constant over the bunch with a value given by $ \eta_{\smallT } ( 0 ) \cong 3.85 \eta{ z } ( 0 ) $: %TODO reference bin model
+indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.
+
 \input{bin_model_figure}

### 2013-02-13 inc/extension/initial_conditions/initial_conditions.tex
typo

@@
 \end{equation}
-where $w$ is the half-width 1/e maximum (HW1/eM) of the spatial irradiance distribution, $\tau$ is the HW1/e pulse duration and $I_{0}$ is the peak pulse irradiance incident on a flat metallic photocathode.
+where $w$ is the half-width 1/e maximum (HW1/eM) of the spatial irradiance distribution, $\tau$ is the HW1/eM pulse duration and $I_{0}$ is the peak pulse irradiance incident on a flat metallic photocathode.
 The initial tranverse ($\smallT$) properties of the beam are easily understood.

### 2013-02-14 inc/extension/initial_conditions/initial_conditions.tex
more work on initial conditions

@@
 \end{equation}
-A laser pulse of this form has a temporal uncertainty given simply $ ( \Delta t_{L} )^{ 2 } = \tau^{ 2 } / 2 $.
+A laser pulse of this form has a temporal uncertainty $ \Delta t_{L} = \tau / \sqrt{2} $.
 
-For spatially-uniform photoemission from a planar photocathode, we have $ \gamma_{\smallT} ( 0 ) = 0 $; i.e., there is no initial spatial momentum chirp in the electron bunch.
+For spatially-uniform photoemission from a planar photocathode, from symmetry we have $ \gamma_{\smallT} ( 0 ) = 0 $; i.e., there is no initial transverse spatial momentum chirp across the electron bunch.
 Finally, thanks to the careful analysis of Dowell and Schmerge \cite{dowell_quantum_2009}, we know that metallic photocathode's initial transverse momentum variance at $ t = 0 $ is given by
@@
 \begin{equation}
-  \sqrt{ \sigma_{ z } ( 0 ) } = \Delta z = \frac{ 1 }{ 2 } a ( \Delta t_{L} )^{ 2 }  = \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m }
+  \sqrt{ \sigma_{ z } ( 0 ) } \approx \Delta z = \frac{ 1 }{ 2 } a ( \Delta t_{L} )^{ 2 }  = \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m }
 \end{equation}
@@
 \begin{equation}
-  \gamma_{ z } ( 0 ) = m \Delta v \Delta z = m a \Delta t_{ L } \sqrt{ \sigma_{ z } ( 0 ) } = \frac{ e E_{ DC } \tau }{ \sqrt{ 2 } }\sqrt{ \sigma_{ z } ( 0 ) }\, .
+  \gamma_{ z } ( 0 ) \approx \Delta v \Delta z = m a \Delta t_{ L } \sqrt{ \sigma_{ z } ( 0 ) } = \frac{ e E_{ DC } \tau }{ \sqrt{ 2 } }\sqrt{ \sigma_{ z } ( 0 ) }\, .
 \end{equation}
@@
 Since the electric field has no time to influence the front of the pulse before the back is also generated we can say that the front of the pulse has a velocity equal to $v_0$ (the velocity associated with the excess photoemission energy) and the back is essentially stationary just after generation.
-The pulse's initial spatial variance $\Delta z$ is therefore equal to the distance that the front of the pulse travels during the laser pulse duration $\Delta t_L$,
-\begin{equation} \label{eq:initial_sigma_z}
-  \sqrt{ \sigma_{ z } ( 0 ) } = \Delta z = v_{ 0 } \Delta t_{L} = \frac{ v_{ 0 } \tau }{ \sqrt{ 2 } } \, \text{.}
+Further, since the photoemission efficiency goes like $\Delta E^2$ \cite{shalaev_electron_1994} the distribution of emitted electrons vs velocity goes like $v^4$; therefore for a first approximation it is reasonable to say that in this case $\Delta v \approx v_0$.
+The pulse's initial spatial variance $\Delta z$ is therefore approximately to the spread of distances traveled during the laser pulse duration $\Delta t_L$,
+\begin{equation} 
+  \sqrt{ \sigma_{ z } ( 0 ) } \approx \Delta z = \Delta v \Delta t_{L} = \frac{ v_{ 0 } \tau }{ \sqrt{ 2 } } \, \text{.}
 \end{equation}
@@
 \begin{equation}
-  \gamma_{ z } ( 0 ) = m \Delta v \Delta z = m v_{ 0 } \sqrt{ \sigma_{ z } ( 0 ) } \, ,
+  \gamma_{ z } ( 0 ) \approx m \Delta v \Delta z = m v_{ 0 } \sqrt{ \sigma_{ z } ( 0 ) } \, ,
 \end{equation}
+
+For simplicity of use, one may combine the initial conditions resulting from these two cases.
+Because they arise from independent processes, the initial spatial uncertainties $\sigma_{ z } ( 0 )$ may be added naively.
+\begin{equation} \label{eq:initial_sigma_z}
+  \sigma_{ z } ( 0 ) \approx \frac{ ( v_{ 0 } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 }
+\end{equation}
+While it is perhaps more rigorously correct to naively add the initial momentum chirp terms $\gamma_{ z } ( 0 )$ with the spatial variance evaluated in each respective limit, we propose that the evaluation be done in terms of the resultant form presented in \ref{eq:initial_sigma_z}.
+\begin{equation}
+  \gamma_{ z } ( 0 ) \approx \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right )
+\end{equation}
+In this way the the model may, in an intermediate case, more closely reflect the pulse as modeled rather than the pulse as idealized in each case.
 
@@
 \begin{equation}
-  \eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 )}{ 4 } \; \text{.}
+  \eta_{ z } ( 0 ) \approx \frac{ \eta_{ \smallT } ( 0 )}{ 4 } \; \text{.}
 \end{equation}
@@
   \begin{gather}
-    \sigma_{ \smallT } ( 0 ) = \frac{ w^{ 2 } }{ 2 } \\
-    \sigma_{ z } ( 0 ) = \frac{ ( v_{ 0 } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
-    \gamma_{ \smallT } ( 0 ) = 0 \\
-    \gamma_{ z } ( 0 ) = \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
-    \eta_{ \smallT } ( 0 ) = \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \\
-    \eta_{ z } ( 0 ) = \frac{ \eta_{ \smallT } ( 0 ) }{ 4 }
+    \sigma_{ \smallT } ( 0 ) \approx \frac{ w^{ 2 } }{ 2 } \\
+    \sigma_{ z } ( 0 ) \approx \frac{ ( v_{ 0 } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
+    \gamma_{ \smallT } ( 0 ) \approx 0 \\
+    \gamma_{ z } ( 0 ) \approx \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
+    \eta_{ \smallT } ( 0 ) \approx \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \\
+    \eta_{ z } ( 0 ) \approx \frac{ \eta_{ \smallT } ( 0 ) }{ 4 }
   \end{gather}

### 2013-02-15 inc/extension/initial_conditions/initial_conditions.tex
implement a \punct macro for end-of-math punctuation, this may be a bad idea, but at least it would now be searchable

@@
 \begin{equation}\label{eq:laser_form}
-I ( x , y , t ) = I_{0} \exp \left ( - \frac{ ( x^{2} +y^{2} ) }{ w^{2} } - \frac{ t^{ 2 } }{ \tau^{ 2 } } \right ) \text{ ,}
+I ( x , y , t ) = I_{0} \exp \left ( - \frac{ ( x^{2} +y^{2} ) }{ w^{2} } - \frac{ t^{ 2 } }{ \tau^{ 2 } } \right ) \punct{,}
 \end{equation}
@@
 \begin{equation}
-\sigma_{\smallT } ( 0 ) = ( \Delta x )^{2} = ( \Delta y)^{2} = \frac{ w^{2} }{ 2 } \text{ .}
+\sigma_{\smallT } ( 0 ) = ( \Delta x )^{2} = ( \Delta y)^{2} = \frac{ w^{2} }{ 2 } \punct{.}
 \end{equation}
@@
 \begin{equation} \label{eq:jensen_eta}
-  \eta_{ \smallT } ( 0 ) \approx \frac{ m_{e} }{ 3 } ( \hbar \omega - \Phi ) \; \text{,}
+  \eta_{ \smallT } ( 0 ) \approx \frac{ m_{e} }{ 3 } ( \hbar \omega - \Phi ) \punct{,}
 \end{equation}
@@
 \begin{equation}
-  \sqrt{ \sigma_{ z } ( 0 ) } \approx \Delta z = \frac{ 1 }{ 2 } a ( \Delta t_{L} )^{ 2 }  = \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m }
+  \sqrt{ \sigma_{ z } ( 0 ) } \approx \Delta z = \frac{ 1 }{ 2 } a ( \Delta t_{L} )^{ 2 }  = \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \punct{.}
 \end{equation}
@@
 \begin{equation}
-  \gamma_{ z } ( 0 ) \approx \Delta v \Delta z = m a \Delta t_{ L } \sqrt{ \sigma_{ z } ( 0 ) } = \frac{ e E_{ DC } \tau }{ \sqrt{ 2 } }\sqrt{ \sigma_{ z } ( 0 ) }\, .
+  \gamma_{ z } ( 0 ) \approx \Delta v \Delta z = m a \Delta t_{ L } \sqrt{ \sigma_{ z } ( 0 ) } = \frac{ e E_{ DC } \tau }{ \sqrt{ 2 } }\sqrt{ \sigma_{ z } ( 0 ) } \punct{.}
 \end{equation}
@@
 \begin{equation} 
-  \sqrt{ \sigma_{ z } ( 0 ) } \approx \Delta z = \Delta v \Delta t_{L} = \frac{ v_{ 0 } \tau }{ \sqrt{ 2 } } \, \text{.}
+  \sqrt{ \sigma_{ z } ( 0 ) } \approx \Delta z = \Delta v \Delta t_{L} = \frac{ v_{ 0 } \tau }{ \sqrt{ 2 } } \punct{.}
 \end{equation}
@@
 \begin{equation}
-  \gamma_{ z } ( 0 ) \approx m \Delta v \Delta z = m v_{ 0 } \sqrt{ \sigma_{ z } ( 0 ) } \, ,
+  \gamma_{ z } ( 0 ) \approx m \Delta v \Delta z = m v_{ 0 } \sqrt{ \sigma_{ z } ( 0 ) } \punct{,}
 \end{equation}
@@
 For simplicity of use, one may combine the initial conditions resulting from these two cases.
-Because they arise from independent processes, the initial spatial uncertainties $\sigma_{ z } ( 0 )$ may be added naively.
+Because they arise from independent processes, the initial spatial uncertainties $\sigma_{ z } ( 0 )$ may be added naively 
 \begin{equation} \label{eq:initial_sigma_z}
-  \sigma_{ z } ( 0 ) \approx \frac{ ( v_{ 0 } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 }
+  \sigma_{ z } ( 0 ) \approx \frac{ ( v_{ 0 } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \punct{.}
 \end{equation}
-While it is perhaps more rigorously correct to naively add the initial momentum chirp terms $\gamma_{ z } ( 0 )$ with the spatial variance evaluated in each respective limit, we propose that the evaluation be done in terms of the resultant form presented in \ref{eq:initial_sigma_z}.
+While it is perhaps more rigorously correct to naively add the initial momentum chirp terms $\gamma_{ z } ( 0 )$ with the spatial variance evaluated in each respective limit, we propose that the evaluation be done in terms of the resultant form presented in \ref{eq:initial_sigma_z}
 \begin{equation}
-  \gamma_{ z } ( 0 ) \approx \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right )
+  \gamma_{ z } ( 0 ) \approx \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \punct{.}
 \end{equation}
@@
 \begin{equation}
-  ( \Delta p_{ z } )^{2} = \left ( \frac{\Delta p_{ x , y }}{ 2 } \right )^{2} \text{,}
+  ( \Delta p_{ z } )^{2} = \left ( \frac{\Delta p_{ x , y }}{ 2 } \right )^{2} \punct{,}
 \end{equation}
@@
 \begin{equation}
-  \eta_{ z } ( 0 ) \approx \frac{ \eta_{ \smallT } ( 0 )}{ 4 } \; \text{.}
+  \eta_{ z } ( 0 ) \approx \frac{ \eta_{ \smallT } ( 0 )}{ 4 } \punct{.}
 \end{equation}
 
-These initial conditions at $t=0$ for an incident Gaussian laser pulse of the form given in \ref{eq:laser_form}, will be used for the remainder of this paper.
+In conclusion, the resulting set of analytic initial conditions for an incident Gaussian laser pulse of the form given in \ref{eq:laser_form} are
 \begin{subequations} \label{eq:summary}
@@
     \eta_{ \smallT } ( 0 ) \approx \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \\
-    \eta_{ z } ( 0 ) \approx \frac{ \eta_{ \smallT } ( 0 ) }{ 4 }
+    \eta_{ z } ( 0 ) \approx \frac{ \eta_{ \smallT } ( 0 ) }{ 4 } \punct{.}
   \end{gather}
 \end{subequations}
+Unless otherwise noted, these initial conditions will be used in all simulations for the remainder of this paper.
 

### 2013-02-15 inc/extension/initial_conditions/initial_conditions.tex
comment on bin model

@@
 
+To test the validity of the simplistic model presented above I also designed a bin model that depends on nothing other than basic physics and a ``transmission'' function.
+This function is the likelihood of emission by emission energy, the maximum energy being the the excess photoemission energy.
+
+In the model the space in front of the photocathode is binned into segments in ranges of distance from the photocathode.
+The laser pulse duration is then binned into slices of time for a total duration of $N \tau$ where the peak of the pulse occurs at $N \tau / 2$.
+For each slice of time, one can determine the relative number of electrons generated in that slice (normalized to the total number that will be created by the end of the simulation), and the velocity distribution of those electrons.
+Knowing the remaining time in the simulation and the acceleration due to the electric field, the velocity distribution can be used to place these electrons into final spatial bins.
+Additionally, as the electrons are placed into these bins, their average velocity is stored as well.
+
+This process repeats for each time slice until the finish time, at which point the laser has essentially stopped creating electrons.
+At this time, the earliest fastest electrons will be a distance 
+\begin{equation}
+  d_{max} = v_0 N \tau + \frac{e E_{ DC }}{2 m} (N \tau)^2 
+\end{equation}
+away from the the photocathode, and the the electrons created in the last slice will be at $d=0$.
+By counting the electrons in each spatial bin and fitting to a gaussian, we can determine $\sigma_z(0)$, the inital spatial uncertainty in the pulse.
+Taking a weighted average of the velocity of the electrons in each spatial bin and fitting to a linear function across the pulse, we can determine the velocity (momentum) chirp $\gamma_z(0)$.
+The shapes of the actual output compare well with the fitted functional form for both of these parameters for all reasonable ranges of field and laser pulse duration.
+
+\input{bin_model_figure}
+
+Finally the velocity uncertainty in each pulse can be computed, however unlike the previous parameters, the AG model only allows for a constant value of local momentum uncertainty across the pulse $\eta_z(0)$.
+As seen in \ref{fig:bin_model} this limitation on functional form severely limits the accuracy with which the determined value of $\eta_z(0)$ represents the actual distribution of momentum uncertainties computed by the binning model.
+
 Additionally the iterative time-step analysis clearly shows that the local longitudinal momentum variance $ \eta_{ z } $ is not constant over the inital photo-generated electron bunch.
@@
 indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.
-
-\input{bin_model_figure}

### 2013-02-22 inc/model_results/lens_charge/lens_charge.tex
began commentary on lens charge interaction

\subsection{Space-charge Effects in Magnetic Lenses}

In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses.\cite{oudheusden_electron_2007,lagrange_nanosecond_2008}
In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry.\cite{berger_dc_2009}
The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity).

In the preceding sections it has already been established that prolate eccentricities are preferred for UEM in general.
When focusing the pulse it can be shown that prolate pulses will also have markedly better performance.
Therefore, in the sake of simplicity, only the prolate case will be demonstrated here.
For more information see reference \cite{berger_semi-analytic_2010}.

\begin{figure}
  \centering
  \subfloat[][] {
    \label{fig:focus_lens_charge_long}
    \input{lens_60mm}
  }
  \\
  \subfloat[][] { 
    \label{fig:focus_lens_charge_short}
    \input{lens_6mm}
  }
  \caption{ PLACEHOLDER
  %Focal behavior of perfect magnetic lenses predicted by the AG model (with DC acceleration voltage of 20kV) as a function of pulse charge $N$: \subref{fig:LongPancake} $f = $ 63.5mm and disk-like ($ w = 500 \mu \text{m}$ and $ \xi ( 0 ) = 0.1 $); \subref{fig:LongCigar} $ f = $ 63.5mm and cigar-like ($ w = 107.7 \mu \text{m}$ and $ \xi ( 0 ) = 10 $); \subref{fig:ShortPancake} $ f = $ 6.35mm and disk-like ($ w = 500 \mu \text{m}$ and $ \xi ( 0 ) = 0.1 $); \subref{fig:ShortCigar} $ f = $ 6.35mm and cigar-like ($ w = 107.7 \mu \text{m}$ and $ \xi ( 0 ) = 10 $). A logarithmic scale is used for comparative clarity.
  }
  \label{fig:focus_lens_charge}
\end{figure}

\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for prolate pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_long}) and a `short' focal length of 6.0mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_short}).
%In all cases, the pulse is assumed to be accelerated to 20kV and generated from a Ta photocathode with a negligible excess photoemission energy (i.e., $ \eta_{i} \rightarrow 0 $) in order to highlight the effect of space-charge. For both lens strengths, the deleterious effects of space-charge are evident as the number of electrons/pulse $ N $ is increased. The single-electron ($ N = 1 $) focal spot sizes of $\sim$6 microns for the $ f = 63.5 \text{mm} $ lens and $\sim$0.6 microns for the shorter $ f = 6.35 \text{mm} $ lens increase markedly for $ N \gtrsim 10^{ 5 } $ and $ N \gtrsim 10^{ 6 }$, respectively. This order of magnitude difference is fundamentally related to the greater time of flight before the focus of the $ f = 63.5\text{mm} $ lens; it increases the impulse that the internal Coulomb force can exert on the pulse as its charge density is increased under focusing. A consequent further effect is a shift in the focal position to $ z > f $.  A more subtle effect is that disk-like pulses are more readily focused than cigar-like pulses --- the rate of transverse pulse broadening due to space-charge effects already being intrinsically greater in the latter (\Fig \ref{fig:compare}). Thus, the electron pulse propagation model predicts that higher fidelity focusing will be achieved with shorter focal length magnetic lenses, higher acceleration energy electrons (time of flight reduction), and pulses with $ \sigma_{ \smallT } \gg \sigma_{ z } $ at the entrance aperture of the lens.

### 2013-02-22 main.tex
a little reorg to start free-space space-charge section
added shapes plot
a little reorg
work on short titles for figures (continuing)
implement a \punct macro for end-of-math punctuation, this may be a bad idea, but at least it would now be searchable
began commentary on lens charge interaction

@@
 %\providecommand{\smallZ}{\ensuremath{{\scriptscriptstyle Z}}}
+
+\providecommand{\punct}[1]{\ensuremath{\, \text{#1}}}
 
@@
 
-\section{Free Expansion Results}
-
-\subsection{Comparison to Other Articles}
-
-\subsection{Trends for Initial Shapes}
+  %TODO refer from here to Free expansion results
 
@@
 
-\section{Model Results, Predictions and Implications} \label{sec:model_results}
+\chapter{Model Results, Predictions and Implications}
+
+\section{Free Expansion Propagation}
+
+  \subimport{inc/model_results/shapes/}{shapes}
+
+  \subimport{inc/model_results/spacecharge/}{spacecharge}
+
+\section{Propagtion under External Forces}
+
+  \subimport{inc/model_results/lens_charge/}{lens_charge}
 
@@
     \includegraphics{inc/hardware/column.jpg}
-    \caption{
-      Image of the majority of the prototype UEM column at UIC.
+    \caption[Picture of the prototype UEM column at UIC]{
+      Picture of the majority of the prototype UEM column at UIC.
       The main vacuum chamber is laid sideways, with the high voltage (gun) at the left and the imaging system on the right.

### 2013-02-23 inc/model_results/lens_charge/lens_charge.tex
improve lens_charge section

@@
   }
-  \caption{ PLACEHOLDER
-  %Focal behavior of perfect magnetic lenses predicted by the AG model (with DC acceleration voltage of 20kV) as a function of pulse charge $N$: \subref{fig:LongPancake} $f = $ 63.5mm and disk-like ($ w = 500 \mu \text{m}$ and $ \xi ( 0 ) = 0.1 $); \subref{fig:LongCigar} $ f = $ 63.5mm and cigar-like ($ w = 107.7 \mu \text{m}$ and $ \xi ( 0 ) = 10 $); \subref{fig:ShortPancake} $ f = $ 6.35mm and disk-like ($ w = 500 \mu \text{m}$ and $ \xi ( 0 ) = 0.1 $); \subref{fig:ShortCigar} $ f = $ 6.35mm and cigar-like ($ w = 107.7 \mu \text{m}$ and $ \xi ( 0 ) = 10 $).
+  \caption[AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$]{
+    Focal behavior of perfect magnetic lenses predicted by the AG model (with DC acceleration voltage of 20kV) as a function of pulse charge $N$.
+    Pulses begin at the lens' front focal plane with HW1/eM width $ 500 \mu \text{m}$ and duration 500 fs at 20 kV velocity ($c/3$) or ellipticiy $ \xi ( 0 ) = 0.1 $ (prolate).
+    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth, \subref{fig:focus_lens_charge_long} $f = $ 60mm and \subref{fig:focus_lens_charge_short} $ f = $ 6.35mm.
     A logarithmic scale is used for comparative clarity.
@@
 \ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for prolate pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_long}) and a `short' focal length of 6.0mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_short}).
-%In all cases, the pulse is assumed to be accelerated to 20kV and generated from a Ta photocathode with a negligible excess photoemission energy (i.e., $ \eta_{i} \rightarrow 0 $) in order to highlight the effect of space-charge.
 For both lens strengths, the deleterious effects of space-charge are evident as the number of electrons/pulse $ N $ is increased.
-The single-electron ($ N = 1 $) focal spot sizes of $\sim$6 microns for the $ f = 63.5 \text{mm} $ lens and $\sim$0.6 microns for the shorter $ f = 6.35 \text{mm} $ lens increase markedly for $ N \gtrsim 10^{ 5 } $ and $ N \gtrsim 10^{ 6 }$, respectively.
-This order of magnitude difference is fundamentally related to the greater time of flight before the focus of the $ f = 63.5\text{mm} $ lens; it increases the impulse that the internal Coulomb force can exert on the pulse as its charge density is increased under focusing.
+To make an effective comparison, highlighting the effect of space-charge alone, the pulses used are at 20kV velocity with $\Delta E = $ 0.5eV but are not simulated using an accelerator.
+The length of the accelerator and the anode lensing would make useful comparison difficult; as such the simulation presented approximates a beam which has been perfectly collimated at the front-focal plane of the lens.
+
+The single-electron ($ N = 1 $) focal spot sizes of $\sim$150 microns for the $ f = 60 \text{mm} $ lens and $\sim$15 microns for the shorter $ f = 6.0 \text{mm} $ lens show marked increase starting around $ N \approx 10^{ 5 } $ and $ N \approx 10^{ 6 }$, respectively, and worsening dramatically above these thresholds.
+This order of magnitude difference is fundamentally related to the greater time of flight before the focus of the $ f = 60\text{mm} $ lens; it increases the impulse that the internal Coulomb force can exert on the pulse as its charge density is increased under focusing.
 A consequent further effect is a shift in the focal position to $ z > f $.
-A more subtle effect is that disk-like pulses are more readily focused than cigar-like pulses --- the rate of transverse pulse broadening due to space-charge effects already being intrinsically greater in the latter (\Fig \ref{fig:compare}).
+%A more subtle effect is that disk-like pulses are more readily focused than cigar-like pulses --- the rate of transverse pulse broadening due to space-charge effects already being intrinsically greater in the latter (\Fig \ref{fig:compare}).
 Thus, the electron pulse propagation model predicts that higher fidelity focusing will be achieved with shorter focal length magnetic lenses, higher acceleration energy electrons (time of flight reduction), and pulses with $ \sigma_{ \smallT } \gg \sigma_{ z } $ at the entrance aperture of the lens.

### 2013-02-23 inc/model_results/lens_charge/lens_charge.pl

@@ -31,7 +31,7 @@ sub gen_sim {
       number => $num,
       velocity => vc / 3,
       sigma_t => (500**2/2) . 'um^2',
-      sigma_z => ((0.6*1e8)**2/2) . ' (ps m / s)^2',
+      sigma_z => ((0.5*1e8)**2/2) . ' (ps m / s)^2',
       eta_t => (me * 0.5 / 3) . 'kg eV',
     );
 
@@ -82,7 +82,7 @@ foreach my $set ( @sets ) {
     wcols $z, $w, "lens_${lens}mm_n$_.dat";
 
     my $min_ind = $w->minimum_ind;
-    say "\t" . $z->at($min_ind);
+    say "\t" . $w->at($min_ind) . 'mm at z = ' . $z->at($min_ind) . 'f';
   }
 
 }

### 2013-02-23 inc/model_results/lens_charge/lens_charge.tex
more improvements to lens_charge plot

@@
   \centering
+  \centerline{
     \subfloat[][] {
@@
     }
-  \\
     \subfloat[][] { 
@@
     }
+  }
   \caption[AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$]{
-    Focal behavior of perfect magnetic lenses predicted by the AG model (with DC acceleration voltage of 20kV) as a function of pulse charge $N$.
-    Pulses begin at the lens' front focal plane with HW1/eM width $ 500 \mu \text{m}$ and duration 500 fs at 20 kV velocity ($c/3$) or ellipticiy $ \xi ( 0 ) = 0.1 $ (prolate).
-    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth, \subref{fig:focus_lens_charge_long} $f = $ 60mm and \subref{fig:focus_lens_charge_short} $ f = $ 6.35mm.
+    AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$.
+    Pulses begin at the lens' front focal plane with HW1/eM width $ 500 \mu \text{m}$ and duration 500 fs at 20 kV velocity ($c/3$) for an ellipticiy $ \xi ( 0 ) = 0.1 $ (prolate).
+    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth, \subref{fig:focus_lens_charge_long} $f = $ 60mm and \subref{fig:focus_lens_charge_short} $ f = $ 6.0mm.
     A logarithmic scale is used for comparative clarity.

### 2013-02-23 inc/model_results/lens_charge/lens_charge.tex
improve plots for lens_charge

@@
 %A more subtle effect is that disk-like pulses are more readily focused than cigar-like pulses --- the rate of transverse pulse broadening due to space-charge effects already being intrinsically greater in the latter (\Fig \ref{fig:compare}).
-Thus, the electron pulse propagation model predicts that higher fidelity focusing will be achieved with shorter focal length magnetic lenses, higher acceleration energy electrons (time of flight reduction), and pulses with $ \sigma_{ \smallT } \gg \sigma_{ z } $ at the entrance aperture of the lens.
+
+%TODO? include simulations for different voltages to establish TOF argument?
+
+Thus, the extended AG electron pulse propagation model predicts that higher fidelity focusing will be achieved with shorter focal length magnetic lenses, higher acceleration energy electrons (time of flight reduction), and pulses with $ \sigma_{ \smallT } \gg \sigma_{ z } $ at the entrance aperture of the lens.

### 2013-02-25 inc/ms_model/ms_model.tex
some bugfixes
added hand-made changes to MS model section

@@
 \end{equation}
-This notation suggests to the reader that the distribution function is implicitly a function of time as well as of the spatial and momentum dimensions.
+This notation suggests to the reader that the distribution function is implicitly a function of time as well as of the spatial and momentum coordinates.
 
-For a classical set of $N$ particles (labelled by the index $i$), distributed in 6D space (3 position, 3 momentum) by 
+For a classical set of $N$ discreet particles (labelled by the index $i$), distributed in 6D space (3 position, 3 momentum) by 
 \begin{equation}
@@
 but with \ref{eq:std_motion} the final discreet evolution equation is
-\begin{equation}
+\begin{equation} \label{eq:evolution_discreet}
   \frac{d}{dt} f_{\smallD}(\vec{r}, \vec{p}; t) =
@@
 where the extra minus sign comes from the change of perspective of \ref{eq:std_motion} from the frame of the individual particle $i$ to the observer frame.
-%TODO ensemble av and mean-field
-Final equation:
+
+This analysis can be extended to continuous distributions by taking the ensemble average over \ref{eq:evolution_discreet}.
+In order to simplify the resulting equation to a useful result, one applies the usual mean-field approximation
+\begin{equation}
+  \left < f_{\smallD} (\vec{r}, \vec{p}; t) f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right > = \left < f_{\smallD} (\vec{r}, \vec{p}; t) \right > \left < f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right >
+\end{equation}
+for a final formal evolution equation
 \begin{equation} \label{eq:formal_evolution}
@@
 
-The formal evolution explored in \ref{sec:formal_evolution} allows for any distribution function.
+The formal evolution explored in Section \ref{sec:formal_evolution} allows for any distribution function.
 For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian, mimicing the laser pulse which generated it.
@@
 \begin{equation}
-  f(\vec{r}, \vec{p}; t) = C(t)\exp \left[ - \Gamma(\vec{r}, \vec{p}; t) \right]
+  f(\vec{r}, \vec{p}; t) = C(t)\exp \left[ - G(\vec{r}, \vec{p}; t) \right]
 \end{equation}
@@
 The subscript $\alpha$ will refer to either $\smallT$ or $z$.
-In this case we choose the argument (i.e. the function $\Gamma$) as
+In order to model several important physical characteristics of the pulse, we choose the argument (i.e. the function $G$) as
 \begin{equation} \label{eq:gaussian_exponent}
-  \Gamma(\vec{r}, \vec{p}; t) =
+  G(\vec{r}, \vec{p}; t) =
   \frac{x^2 + y^2}{2 \sigma_{\smallT}} + \frac{z^2}{2 \sigma_{z}}
@@
 \end{equation}
-This parameterization defines the spatial uncertainty $\sigma_{\alpha}$, the local momentum uncertainty $\eta_{\alpha}$, and the momentum chirp $\gamma_{\alpha}$ in each spatial direction.
-The ``chirp'' of a pulse quantifies the difference in local momentum from one side of the pulse to the other.
+This parameterization defines the spatial uncertainty $\sigma_{\alpha}$, the local momentum uncertainty $\eta_{\alpha}$, and the momentum chirp $\gamma_{\alpha}$ in each spatial direction $\alpha$.
+The ``chirp'' of a pulse quantifies the difference in average local momentum from one side of the pulse to the other.
 For example, in a pulse which is normally (positively) chirped longitudinally, electrons near the front of the pulse will be moving faster (on average) than those in the back.
-
-The normalization constant $C(t)$, for $N$ total particles, is found by applying equation \ref{eq:normalization}
+The normalization constant $C(t)$, for this distributions of $N$ total particles, is found by applying equation \ref{eq:normalization}
 \begin{equation}
@@
 \end{equation}
-These equations may be viewed from a matrix perspective if we define the ``coordinate vector''
+
+To help simplify what is already becoming a complex problem, these equations may be viewed from a matrix perspective if we define the ``coordinate vector''
 \begin{equation} \label{eq:coordinate_vector}
@@
 \end{equation}
-In this space, \ref{eq:gaussian_exponent} becomes
+In this space, \ref{eq:gaussian_exponent} becomes a 9x9 matrix
 \begin{equation}
-  \Gamma(\vec{r}, \vec{p}; t) = \frac{1}{2}\sum\limits_{ij} A_{ij} u_i u_j
+  G(\vec{r}, \vec{p}; t) = \frac{1}{2}\sum\limits_{ij} A_{ij} u_i u_j
 \end{equation}
@@
 \end{equation}
-with ${\scriptstyle \alpha} = \smallT \text{ or } z$. 
+where again ${\scriptstyle \alpha} = \smallT \text{ or } z$. 
 A nice property of Gaussians of this form is the convenient ``Multidimensional Gaussian Integral'' given by
@@
 which is a common form of integral analysis in Quantum Field Theory (see for example Peskin and Schroeder Equation 9.70). 
-Application of this integral yields
+Application of this integral yields another 9x9 matrix
 \begin{equation}
@@
 The parameter $\xi$ is the ellipticity parameter, defined as
-\begin{equation}
+\begin{equation} \label{eq:define_xi}
   \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}}
 \end{equation}
-Note that the original paper of Michalik and Sipe contained an error which incorrectly defined $\xi$; this was corrected in an erratum \cite{michalik_erratum:_2008}.
+Note that the original paper of Michalik and Sipe (Reference \cite{michalik_analytic_2006}) contained an error which incorrectly defined $\xi$; this was corrected in an erratum (Reference \cite{michalik_erratum:_2008}).
 The family of $L$ functions are
@@
 \end{equation}
-which, although piecewise, is a smooth and well behaved function.
+which, although piecewise, is a smooth and well behaved function for all physically meaningful $\xi$.
 
@@
 When these $K$ matrices are summed (\ref{eq:gaussian_integral_dt_k}) to form the right hand side (RHS) of the differential equations and compared to LHS derived above (\ref{eq:dainvdt}), the non-zero elements finally yield a system of differential equations which govern the dynamics of these Gaussian beam parameters.
-
 \begin{subequations} \label{eq:ag_original}
@@
 \end{subequations}
+These equations are the crux of the model presented by Michalik and Sipe for the time evolution of a Gaussian electron pulse propagating through free space (no external forces).
+They form the basis upon which much of my work in the following chapters is founded.
 
@@
 This is of course not strictly accurate for a Gaussian charge distribution and the error leads to a distortion of the pulse shape.
-Even so, the self-similar AG model of charge bunch dynamics has been benchmarked against particle tracking simulations for a number of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid which explicitly features a linear internal space-charge field.\cite{luiten_how_2004}
+The ``self-similar'' nature of the model, the fact that the overall shape of the pulse cannot change, prevents inclusion of these distortions.
+Even so, the AG model of charge bunch dynamics has been benchmarked against particle tracking simulations for a number of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid which explicitly features a linear internal space-charge field.\cite{luiten_how_2004}

### 2013-02-25 inc/ms_model/ms_model.tex
more hand-written changes

@@
 \end{equation}
-In this space, \ref{eq:gaussian_exponent} becomes a 9x9 matrix
+In this space, \ref{eq:gaussian_exponent} becomes
 \begin{equation}
@@
 \end{equation}
-where
+where the 9 x 9 matrix $A$ is defined as
 \begin{equation}
@@
 \end{equation}
-where again ${\scriptstyle \alpha} = \smallT \text{ or } z$. 
+and again ${\scriptstyle \alpha} = \smallT \text{ or } z$. 
 A nice property of Gaussians of this form is the convenient ``Multidimensional Gaussian Integral'' given by
@@
 which is a common form of integral analysis in Quantum Field Theory (see for example Peskin and Schroeder Equation 9.70). 
-Application of this integral yields another 9x9 matrix
+Application of this integral yields another 9 x 9 matrix
 \begin{equation}

### 2013-02-25 main.tex
more hand-written changes

@@
 
-\chapter{Extension of the Analytic Gaussian Model\\for UEM Column Modeling}
+\chapter{Extension of the Analytic Gaussian Model\\for UEM Column Modeling} \label{chap:extension}
 
 Modeling a system as complex as an Ultrafast Electron Microscope is a task too large for one person.
-In the previous chapter\ref{chap:previous_models} I have highlighted some previous efforts at modeling electron pulse evolution.
+In Chapter \ref{chap:previous_models} I have highlighted some previous efforts at modeling electron pulse evolution.
 While these are excellent works in their own right, none provide the complete picture of the evolution of an electron pulse through the entire column.

### 2013-02-25 inc/extension/initial_conditions/initial_conditions.tex
more brush-up

@@
 where $w$ is the half-width 1/e maximum (HW1/eM) of the spatial irradiance distribution, $\tau$ is the HW1/eM pulse duration and $I_{0}$ is the peak pulse irradiance incident on a flat metallic photocathode.
-The initial tranverse ($\smallT$) properties of the beam are easily understood.
+The initial tranverse properties of the beam are easily understood.
 The electron pulse's initial spatial variance is simply
@@
 in the limit where the excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ are significantly greater that the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
-%in the limit where the temperature is low enough to consider the system well described by Fermi-Dirac statistics where the chemical potential is nearly equal to the Fermi energy.
+In other words, this is the limit where the temperature is low enough to consider the system well described by Fermi-Dirac statistics where the chemical potential is nearly equal to the Fermi energy.
 
@@
 In the longitudinal direction ($z$), the interplay between acceleration, $ a = e E_{DC}/m $, in the DC gun and the physics of the photoemission process determine the properties $\sigma_z$ and $\gamma_z$ of the electron bunch.
-Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) by the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than that traveled by electrons with the maximum photoemission velocity $ v_{0} = \sqrt{ 2 \Delta E / m } $ alone over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
+Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) when the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than the distance traveled by electrons moving with a constant velocity $ v_{0} = \sqrt{ 2 \Delta E / m } $, the maximum photoemission velocity, over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
 
@@
 \end{equation}
-While it is perhaps more rigorously correct to naively add the initial momentum chirp terms $\gamma_{ z } ( 0 )$ with the spatial variance evaluated in each respective limit, we propose that the evaluation be done in terms of the resultant form presented in \ref{eq:initial_sigma_z}
+In the case of the initial momentum chirp terms $\gamma_{ z } ( 0 )$, while it is perhaps more rigorously correct to naively add  the terms with the spatial variance evaluated in each respective limit, we propose that the evaluation be done in terms of the resultant form presented in \ref{eq:initial_sigma_z}, keeping $\sigma_{z}$ unevaluated,
 \begin{equation}
@@
 \end{equation}
-In this way the the model may, in an intermediate case, more closely reflect the pulse as modeled rather than the pulse as idealized in each case.
+In this way the model may, in an intermediate case, more closely reflect the pulse length as modeled rather than the pulse length as idealized in each limiting case.
 
-Finally, from a simplistic viewpoint, only the electrons generated with positive longitudinal momentum will contribute to the pulse since those electrons generated with negative momentum do not leave the photocathode.
+Finally, and again from a simplistic viewpoint, only the electrons generated with positive longitudinal momentum will contribute to the pulse since those electrons generated with negative momentum do not leave the photocathode.
 This implies that the final initial condition, the local momentum uncertainty in the longitudinal direction $\eta_z$, is half that of the transverse directions; that is to say
@@
 To test the validity of the simplistic model presented above I also designed a bin model that depends on nothing other than basic physics and a ``transmission'' function.
-This function is the likelihood of emission by emission energy, the maximum energy being the the excess photoemission energy.
+This function represents the likelihood of emission versus emission energy; the maximum energy being the full excess photoemission energy $\Delta E$.
 
 In the model the space in front of the photocathode is binned into segments in ranges of distance from the photocathode.
-The laser pulse duration is then binned into slices of time for a total duration of $N \tau$ where the peak of the pulse occurs at $N \tau / 2$.
+The laser pulse duration is then binned into slices of time for a total duration of $N \tau$ where the peak of the pulse occurs at $N \tau / 2$; $N$ is a number selected to accumulate enough of the total laser pulse for an accurate simulation.
 For each slice of time, one can determine the relative number of electrons generated in that slice (normalized to the total number that will be created by the end of the simulation), and the velocity distribution of those electrons.
@@
 By counting the electrons in each spatial bin and fitting to a gaussian, we can determine $\sigma_z(0)$, the inital spatial uncertainty in the pulse.
-Taking a weighted average of the velocity of the electrons in each spatial bin and fitting to a linear function across the pulse, we can determine the velocity (momentum) chirp $\gamma_z(0)$.
-The shapes of the actual output compare well with the fitted functional form for both of these parameters for all reasonable ranges of field and laser pulse duration.
+Taking a weighted average of the velocity of the electrons in each spatial bin and fitting to a linear function across the pulse (as allowed by the AG model), we can determine the velocity (momentum) chirp $\gamma_z(0)$.
+As seen in \ref{fig:bin_model}, the shapes of the actual output compare well with the fitted functional form for both of these parameters for all reasonable ranges of field and laser pulse duration.
 
@@
 Finally the velocity uncertainty in each pulse can be computed, however unlike the previous parameters, the AG model only allows for a constant value of local momentum uncertainty across the pulse $\eta_z(0)$.
-As seen in \ref{fig:bin_model} this limitation on functional form severely limits the accuracy with which the determined value of $\eta_z(0)$ represents the actual distribution of momentum uncertainties computed by the binning model.
+\ref{fig:bin_model} shows that this limitation on functional form severely limits the accuracy with which the determined value of $\eta_z(0)$ represents the actual distribution of momentum uncertainties computed by the binning model.
 
-Additionally the iterative time-step analysis clearly shows that the local longitudinal momentum variance $ \eta_{ z } $ is not constant over the inital photo-generated electron bunch.
+Additionally the iterative time-step analysis clearly demonstrates that the local longitudinal momentum variance $ \eta_{ z } $ is not constant over the inital photo-generated electron bunch.
 In fact, $ \sqrt{ \eta_{ z } ( 0 ) } $ varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch with the narrowest momentum distribution at the front of the bunch where only the faster electrons with velocity near $ v_{ 0 } $ initially reside.

### 2013-02-25 inc/extension/initial_conditions/bin_model_figure.tex
comment on bin model
more improvements to initial_conditions section

@@
     \subfloat[][]{
+      \label{fig:bin_model_acc}
       \begin{tikzpicture}
@@
     \subfloat[][]{
+      \label{fig:bin_model_vmax}
       \begin{tikzpicture}
@@
     \subfloat[][]{
+      \label{fig:bin_model_equal}
       \begin{tikzpicture}
@@
   }
+  \caption[Results of bin model for computing initial conditions]{
+    Comparison of the results of the bin model and AG model allowed functional forms for three different input parameters cases.
+    Compared are: \subref{fig:bin_model_acc} the DC acceleration dominated case, \subref{fig:bin_model_vmax} the photoemission energy dominated case, and \subref{fig:bin_model_equal} an intermediate case.
+    For $\sigma_z$ (red) and $\gamma_z$ (green) the simulated data (solid) and the fit to the AG allowed form (dashed) are decently well matched in all cases.
+    Unfortunately, the model only allows for a constant $\eta_z$ (green) and in all but the extreme acceleration dominated case \subref{fig:bin_model_acc} the model cannot well represent the simulated data.
+  }
+  \label{fig:bin_model}
 \end{figure}

### 2013-02-25 inc/extension/initial_conditions/initial_conditions.tex
more improvements to initial_conditions section

@@
 
-\subsection{Iterative Spatial/Temporal Bin Model}
+\subsection{Iterative Spatial/Temporal Bin Model} \label{sec:bin_model}
 
@@
 
-\input{bin_model_figure}
+\input{bin_model_figure} %TODO do another run of this data, so I can have the run parameters
 
@@
 In fact, $ \sqrt{ \eta_{ z } ( 0 ) } $ varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch with the narrowest momentum distribution at the front of the bunch where only the faster electrons with velocity near $ v_{ 0 } $ initially reside.
-Unfortunately, the mean-field theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch.
-Consequently, we will assume that $ \eta_{ z } $ is a constant over the bunch with a value given by $ \eta_{\smallT } ( 0 ) \cong 3.85 \eta{ z } ( 0 ) $: %TODO reference bin model
-indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.
+Unfortunately, the self-similar theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch; to allow a more realistic fit to the simulated initial data, the pulse would eventually evolve to be non-Gaussian.
+%Consequently, we will assume that $ \eta_{ z } $ is a constant over the bunch with a value given by $ \eta_{\smallT } ( 0 ) \cong 3.85 \eta{ z } ( 0 ) $: indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.

### 2013-02-25 inc/extension/implementation/implementation.tex
comment on Physics::Photoemission in implementation section

@@
 This script is included as Appendix \ref{sec:script}.
+
+Many scripts which use \verb!Physics::UEMColumn! were need to generate data used in the writing of this paper.
+I have included most of them in the source repository which is available at \url{https://github.com/jberger/Thesis}.
+
+Additionally, a development version of the bin model presented in Section \ref{sec:bin_model} is available at \url{https://github.com/jberger/Physics-Photoemission}.
+Once I am sufficently confident in its implementation, and more importantly its voracity, I intend to release it to the CPAN as well, most likely under the name \verb!Physics::Photoemission!.
+This module depends on the much larger Perl project PDL (\url{http://pdl.perl.org}), a matrix-oriented numerical library, making it slightly more difficult to install, however the speed and implementational simplicy benefit added by that dependency is marked.
+Usefully, it also comes with a subclass of \verb!Physics::UEMColumn::Photoemission! which uses this model to generate the pulse's $z$ direction initial conditions.

### 2013-02-25 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
tweak liouvilles_theorem section

--- 2012-09-24 inc/liouvilles_theorem/liouvilles_theorem.tex
@@
 
-To evaluate the effect of this theorem on these equations, consider the time derivative of the product $\sigma_{\alpha} \eta_{\alpha}$.
+In the AG model this phase space product is simply $\sigma_{\alpha} \eta_{\alpha}$
+To evaluate the effect of this theorem on these equations, consider the time derivative of this product
 Using \ref{eq:ag_original} it is easy to show that
@@
 \end{gather}
-This analysis shows that the Analytic Gaussian model explicitly obeys the Liouville Theorem.
+showing that the Analytic Gaussian model explicitly obeys the Liouville Theorem.
 

### 2013-02-25 inc/extension/external_forces/external_forces.tex
improve formal external_forces section

@@
 
-Besides a set of initial conditions, the Analytic Gaussian model presented by Michalik and Sipe (see Section \ref{sec:ms_model}) is only useful for modeling free-space propagation.
+Simulation of electron pulse propagation in UEM and modern UED apparatus\cite{siwick_ultrafast_2002,oudheusden_electron_2007,reed_femtosecond_2006} requires the ability to model electron-optical elements designed to control the three-dimensional space-charge and dispersion induced broadening of the pulse, and hence deliver electron pulses of the desired quality to the specimen.
+The Analytic Gaussian model as presented by Michalik and Sipe (see Section \ref{sec:ms_model}) is only useful for modeling free-space propagation.
 This is because there is no allowance for external forces, such as those imparted by the ``optical'' elements of an electron microscope column.
+In this section, I show that the AG model can be extended to include external forces on the pulse, such as the influence of magnetic electron lenses and RF pulse compression cavities.\cite{oudheusden_electron_2007,veisz_hybrid_2007}
 
@@
 \begin{equation}
-  V(\vec{r} - \vecprime{r}) \rightarrow V(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r})
+  V(\vec{r} - \vecprime{r}) \rightarrow V^{int}(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r})
 \end{equation}
-Following this new term through the derivation, we see that
+In doing so, we rename the old potential, the internal repulsive Coulomb force, $V^{int}$.
+Following these new terms through the derivation, we see that
 \begin{equation}
   \begin{split}
-    \Phi(\vec{r};t) = & \int d\vecprime{r} \left( V(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r}) \right) n(\vecprime{r}; t) \\
-    = & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) + \sum_{k} \int d\vecprime{r} V_{k}^{ext}(\vec{r}) n(\vecprime{r}; t) \\
+    \Phi(\vec{r};t) = & \int d\vecprime{r} \left( V^{int}(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r}) \right) n(\vecprime{r}; t) \\
+    = & \int d\vecprime{r} V^{int}(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) + \sum_{k} \int d\vecprime{r} V_{k}^{ext}(\vec{r}) n(\vecprime{r}; t) \\
     \equiv & \Phi^{int}(\vec{r};t) + \sum_{k} \Phi_{k}^{ext}(\vec{r};t)
@@
 These external terms are not self-influencing, and thus are not subject to the complexities of the internal Coulombic repulsion force.
-For clarity we now define $\Phi^{int}(\vec{r};t)$ as the old $\Phi(\vec{r};t)$, which results from the Coulomb force, while our new $K^{force}$ now includes these exteral terms.
+For consistency we have now defined $\Phi^{int}(\vec{r};t)$ as the old $\Phi(\vec{r};t)$, which results from the Coulomb force.
 
-This form of $\Phi(\vec{r};t)$ can now be substituted into the definition of $K^{force}$ matrix (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector} 
+This new form of $\Phi(\vec{r};t)$ can be substituted into the definition of $K^{force}$ matrix (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector} 
 \begin{equation}
@@
 Like the potential integrals, we are left with an additional matrix for each external force.
-Again, for clarity, we now define $K^{int}$ as the $K$ matrix resulting from the internal Coulombic repulsion.
+Again, for consistency, we have defined $K^{int}$ as the $K$ matrix resulting from the internal Coulombic repulsion.
 
@@
 but it is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ (or equivalently $ K^{ext} = \hat{0} $) meaning that this static field has no effect on the parameters of the pulse.
-Of course, this is only true in the pulse frame, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m_{e} $ is unchanged.
-However, in the lab frame, where  $ \Delta E = p \Delta p / m_{e} $, there is an increase in the momentum spread, as the pulse now has the large momentum $p$ imparted by the accelerating field.
+Of course, this is only true in the pulse frame, the frame of the simulation, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m_{e} $ is unchanged.
+In the lab frame however, where  $ \Delta E = p \Delta p / m_{e} $, there is an increase in the momentum spread, as the pulse now has the large momentum $p$ imparted by the accelerating field.
 
@@
 \end{equation}
-Here, $M_{ \smallT }$ and $M_{z}$ characterize the strength of a lens or a compressor, respectively, and are chosen to be positive for pulse focusing.
+Here, $M_{ \smallT }$ and $M_{z}$ characterize the strength of a lens or a compressor, respectively, and are chosen to be positive for focusing/compression elements.
 Substitution into the integral equation for the $ K^{ext} $ matrix elements,
@@
 which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.
+This derivation has been for a single force, in practice however there may be multiple forces, therefore \ref{eq:ag_big_gamma_gamma} may be restated as
+\begin{equation}
+  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2(0) + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
+    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sum M_{\alpha} \sigma_{\alpha}
+\end{equation}
+
+Since these terms are only added to $\dfrac{d \gamma_{\alpha}}{d t}$, the analysis of Liouville's Theorem, shown in Section \ref{sec:liouville}, is unaffected.
+The extended AG model still explicitly satisfies Liouvilles's Theorem.
 

### 2013-02-26 inc/extension/external_forces/external_forces.tex
more updates to formal extension section

@@
 The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations.\cite{michalik_analytic_2006}
-As a result, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
+As a result, we emphasize that the presented AG model extension is valid only for the action of linear forces (i.e., forces that have a linear dependence on distance in the pulse frame) on the electron pulse due to the model's self-similar Gaussian approximation.
+The inclusion of the action of nonlinear forces (e.g., due to space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations, which should be employed to gauge the accuracy (and hence limits) of our extended AG approach at some point.
+
+It should also be noted that these analyses are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
+Extension to include relativistic effects is conceptually straightforward using standard transformations between the laboratory and electron pulse reference frames.
+
 Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid.\cite{luiten_how_2004}
@@
 Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
-
-Further, these analyses are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
-Extension to include relativistic effects is conceptually straightforward using standard transformations between the laboratory and electron pulse reference frames.
 

### 2013-02-26 inc/extension/external_forces/external_forces.tex
improve magnetic lenses section of external forces

@@
   \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2(0) + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
-    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sum M_{\alpha} \sigma_{\alpha}
+    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum M_{\alpha} 
 \end{equation}
@@
 
-\section{Practical Extension to Include Magnetic Lenses and RF Cavities}
+\section{Practical Extension to Include Magnetic Lenses, RF Cavities and DC Accelerators}
 
@@
 
-A magnetic lens may be modeled using one linear force contribution.
+In the transverse ($ \smallT $) direction, a magnetic lens may be modeled using one linear force contribution.
 Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is very complicated to describe \cite{montgomery_some_1961} and the ``lensing'' action is actually a compound force (first rotating the beam, then lensing).
-We have found that in practice, characterizing a specific magnetic lens ($M_{\smallT}(I)$ where $I$ is the current in the magnetic lens) is much easier than attempting to find a generic functional form.
-Further discussion of magnetic lenses can be found in Section \ref{sec:mag_lens}.
+Further discussion of the mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens}.
+We have found that in practice, characterizing a specific magnetic lens ($M_{\smallT}^{mag}(I)$ where $I$ is the current in the magnetic lens) is much easier than attempting to find a generic functional form.
+
+For localizing the region of influence of the lens, I have commonly modelled a thin magnetic lens as having a super-Gaussian order value $n=1$, a simple Gaussian; of course this may depend on the specifics of the lens.
+To simulate the effect of magnetic lenses in a UEM by the extended AG model, for each magnetic lens, add a term like
+\begin{equation} \label{eq:reg_of_influence}
+  M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \text{ ,}
+\end{equation}
+to the set of $M_{\smallT}$ parameters, where $z_{mag}^{\prime}$ and $L_{mag}$ are the position and length of the lens in question respectively.
 

### 2013-02-26 inc/extension/external_forces/external_forces.tex
lens section tweaks

@@
 In the transverse ($ \smallT $) direction, a magnetic lens may be modeled using one linear force contribution.
-Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is very complicated to describe \cite{montgomery_some_1961} and the ``lensing'' action is actually a compound force (first rotating the beam, then lensing).
+Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is surprisingly complicated to describe \cite{montgomery_some_1961} and the ``lensing'' action is in fact a compound force (first rotating the beam, then lensing).
 Further discussion of the mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens}.
@@
 \begin{equation} \label{eq:reg_of_influence}
-  M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \text{ ,}
+  M_{\smallT}^{mag} = M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \text{ ,}
 \end{equation}

### 2013-03-01 inc/ms_model/ms_model.tex
added andreas' comments on ms_model

@@
 
-\section{The Analytic Gaussian Model of Michalik and Sipe} \label{sec:ms_model}
+\section{The Analytic Gaussian Model} \label{sec:ms_model}
 
 This section introduces the Analytic Gaussian model of Michalik and Sipe \cite{michalik_analytic_2006}.
+%TODO more here (obviously), comparison with GPT, extended in ch. 3, etc
 
@@
 \end{equation}
-This notation suggests to the reader that the distribution function is implicitly a function of time as well as of the spatial and momentum coordinates.
+This notation suggests to the reader that the distribution function is implicitly a function of time as well as of the spatial and momentum coordinates; for most circumstances however, the total number of particles $N$ will not change with time.
 
-For a classical set of $N$ discreet particles (labelled by the index $i$), distributed in 6D space (3 position, 3 momentum) by 
+For a classical set of $N$ discreet particles (labelled by the index $i$), distributed in 6D space (3 position, 3 momentum) the distribution function is 
 \begin{equation}
@@
 which automatically satisfies \ref{eq:normalization}.
-The particles have mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
+If the particles have mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
 \begin{subequations}\label{eq:std_motion}
@@
   \frac{d\vec{r}_{i}(t)}{dt} = \frac{\vec{p}_{i}(t)}{m}\\
-  \frac{d\vec{p}_{i}(t)}{dt} = -\sum\limits_{j} \frac{d}{d\vec{r}_{i}} V(\vec{r}_{i} - \vec{r}_{j})
+  \frac{d\vec{p}_{i}(t)}{dt} = -\sum\limits_{j} \frac{d}{d\vec{r}_{i}} V(\vec{r}_{i} - \vec{r}_{j}) \label{eq:std_motion_v}
 \end{gather}
 \end{subequations}
-Now consider the time evolution of $f_{\smallD}$
+Now let us consider the time evolution of $f_{\smallD}$;
 \begin{equation}
@@
 \end{equation}
-where the extra minus sign comes from the change of perspective of \ref{eq:std_motion} from the frame of the individual particle $i$ to the observer frame.
+where the extra minus sign comes from the change of perspective of \ref{eq:std_motion_v} from the frame of the individual particle $i$ to the observer frame.
 
@@
 \end{equation}
-for a final formal evolution equation
+so that the formal evolution of a continuous distribution is given by
 \begin{equation} \label{eq:formal_evolution}
@@
 \end{equation}
+Although this equation looks very similar to \ref{eq:evolution_discreet}, notice that it contains approximations, mean-field that has been seen and in practice the self-similarity of the distribution function, that the discreet discription did not.
 
@@
 The formal evolution explored in Section \ref{sec:formal_evolution} allows for any distribution function.
-For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian, mimicing the laser pulse which generated it.
+For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian in space and time, mimicing the laser pulse which generated it.
 How correct this assumption is will be explored in \ref{sec:initial_conditions}.
@@
 \end{equation}
-This parameterization defines the spatial uncertainty $\sigma_{\alpha}$, the local momentum uncertainty $\eta_{\alpha}$, and the momentum chirp $\gamma_{\alpha}$ in each spatial direction $\alpha$.
+This parameterization defines the spatial variance $\sigma_{\alpha}$, the local momentum variance $\eta_{\alpha}$, and the momentum chirp $\gamma_{\alpha}$ in each spatial direction $\alpha$.
 The ``chirp'' of a pulse quantifies the difference in average local momentum from one side of the pulse to the other.
 For example, in a pulse which is normally (positively) chirped longitudinally, electrons near the front of the pulse will be moving faster (on average) than those in the back.
-The normalization constant $C(t)$, for this distributions of $N$ total particles, is found by applying equation \ref{eq:normalization}
+The normalization constant $C(t)$, for this distributions of $N$ total particles, is found by applying equation \ref{eq:normalization}:
 \begin{equation}
@@
 
-Inserting \ref{eq:formal_evolution} into \ref{eq:gaussian_integral_dt} yields two integral terms, one related to the propagation of a non-charged Gaussian beam, the other adds the influence of the internal force due to Coulomb repulsion.
+Inserting \ref{eq:formal_evolution} into \ref{eq:gaussian_integral_dt} yields two integral terms, one related to the propagation of a non-charged Gaussian beam, the other adds the influence of the internal force due to Coulomb repulsion for an electron pulse.
 These terms will be denoted $K^{flow}$ and $K^{force}$ respectively.
@@
 \end{gather}
-Obviously the $K^{flow}$ term is far easier to evaluate
+Obviously the $K^{flow}$ term is far easier to evaluate;
 \begin{equation}
@@
 The $K^{force}$ integrals, related to the internal force of the pulse's Coulombic repulsion, are far more complicated to evaluate.
-For more information about the derivation, see \cite{michalik_analytic_2006}.
+For more information about the derivation, see \cite{michalik_analytic_2006}. %TODO link to an appendix with Bessel etc for Kforce
 The final matrices are
@@
 \end{equation}
-The parameter $\xi$ is the ellipticity parameter, defined as
+The parameter $\xi$ quantifies the spatial ellipticity of the electron pulse, defined as
 \begin{equation} \label{eq:define_xi}
@@
 \end{equation}
-Note that the original paper of Michalik and Sipe (Reference \cite{michalik_analytic_2006}) contained an error which incorrectly defined $\xi$; this was corrected in an erratum (Reference \cite{michalik_erratum:_2008}).
+[Note that the original paper of Michalik and Sipe (Reference \cite{michalik_analytic_2006}) contained an error which incorrectly defined $\xi$; this was corrected in an erratum (Reference \cite{michalik_erratum:_2008}).]
 The family of $L$ functions are
+\begin{subequations}
   \begin{gather}
@@
     L_{z}(\xi) = \frac{ 3 \xi^2 }{ \xi^2 - 1} \left( \xi L(\xi) - 1 \right) \\
-  L(\xi) = \frac{1}{2} \int_0^{\pi} \frac{1}{1+\xi\sin(\theta)} d\theta
+    L(\xi) = \frac{1}{2} \int_0^{\pi} \frac{1}{1+\xi\sin(\theta)} d\theta \label{eq:ag_l}
   \end{gather}
-The integral function $L(\xi)$ can be evaluated to the piecewise function
+\end{subequations}
+The integral function $L(\xi)$ (\ref{eq:ag_l}) can be further evaluated to the piecewise function
 \begin{equation}
@@
 \end{equation}
-which, although piecewise, is a smooth and well behaved function for all physically meaningful $\xi$.
+which, although defined piecewise, is a smooth and well behaved function for all physically meaningful $\xi$.
 
@@
 These equations are the crux of the model presented by Michalik and Sipe for the time evolution of a Gaussian electron pulse propagating through free space (no external forces).
-They form the basis upon which much of my work in the following chapters is founded.
+They form the basis upon which much of the work in the following chapters is founded.
 
@@
 Even so, the AG model of charge bunch dynamics has been benchmarked against particle tracking simulations for a number of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid which explicitly features a linear internal space-charge field.\cite{luiten_how_2004}
+%TODO add the error found in these benchmarks

### 2013-03-01 inc/extension/initial_conditions/initial_conditions.tex
start merging andreas' comments on initial conditions section

@@
 For a DTEM or UED apparatus, where the electron pulse is produced in a laser-driven photo-electron gun, the implementation of the AG model of Michalik and Sipe \cite{michalik_analytic_2006} requires knowledge of the initial ($ t = 0 $) bunch parameters immediately after generation at the photocathode surface inside the acceleration region of the gun.
-For the purposes of this paper, we will make a simple approximation to the behavior of the gun in order to provide generic initial conditions for the electron pulse.
+For the purposes of this thesis, we will make a simple approximation to the behavior of the gun in order to provide generic initial conditions for the electron pulse.
 
@@
 \end{equation}
-A laser pulse of this form has a temporal uncertainty $ \Delta t_{L} = \tau / \sqrt{2} $.
+Similarly, a laser pulse of this form has a temporal uncertainty $ \Delta t_{L} = \tau / \sqrt{2} $.
 

### 2013-03-01 main.tex
start merging andreas' comments on initial conditions section

@@
 
-Modeling a system as complex as an Ultrafast Electron Microscope is a task too large for one person.
-In Chapter \ref{chap:previous_models} I have highlighted some previous efforts at modeling electron pulse evolution.
+Modeling a system as complex as an Ultrafast Electron Microscope is an enourmous task.
+In Chapter \ref{chap:previous_models}, I have highlighted some previous efforts at modeling electron pulse evolution.
 While these are excellent works in their own right, none provide the complete picture of the evolution of an electron pulse through the entire column.
-In the following chapter I will present a model which builds on and extends those individual pieces to form a single model; one which is generic enough to model many different columns and which is computationally fast enough to be used to design and optimize a column.
+In the following chapter, I will present a model which builds on and extends those individual pieces to form a single model; one which is generic enough to model many different columns and is computationally fast enough to be used to design and optimize a column.
 

### 2013-03-01 inc/extension/initial_conditions/initial_conditions.tex
added more of andreas' comments on initial conditions section

@@
 where $w$ is the half-width 1/e maximum (HW1/eM) of the spatial irradiance distribution, $\tau$ is the HW1/eM pulse duration and $I_{0}$ is the peak pulse irradiance incident on a flat metallic photocathode.
-The initial tranverse properties of the beam are easily understood.
+A laser pulse of this form has a temporal uncertainty $ \Delta t_{L} = \tau / \sqrt{2} $.
+
+The initial tranverse properties of the emitted electron beam are easily understood.
 The electron pulse's initial spatial variance is simply
@@
 \end{equation}
-Similarly, a laser pulse of this form has a temporal uncertainty $ \Delta t_{L} = \tau / \sqrt{2} $.
-
 For spatially-uniform photoemission from a planar photocathode, from symmetry we have $ \gamma_{\smallT} ( 0 ) = 0 $; i.e., there is no initial transverse spatial momentum chirp across the electron bunch.
-Finally, thanks to the careful analysis of Dowell and Schmerge \cite{dowell_quantum_2009}, we know that metallic photocathode's initial transverse momentum variance at $ t = 0 $ is given by
+Finally, thanks to the careful analysis of Dowell and Schmerge \cite{dowell_quantum_2009} (confirmed by Jensen \textit{et.
+al.} \cite{jensen_emittance_2010}), we know that metallic photocathode's initial transverse momentum variance at $ t = 0 $ is given by
 \begin{equation} \label{eq:jensen_eta}
-  \eta_{ \smallT } ( 0 ) \approx \frac{ m_{e} }{ 3 } ( \hbar \omega - \Phi ) \punct{,}
+  \eta_{ \smallT } ( 0 ) \approx \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \punct{,}
 \end{equation}
-in the limit where the excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ are significantly greater that the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
-In other words, this is the limit where the temperature is low enough to consider the system well described by Fermi-Dirac statistics where the chemical potential is nearly equal to the Fermi energy.
+in the limit where the maximum excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ is significantly greater than the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
+%TODO include TE value for $\Delta E < 0$  is $\eta_{\smallT} = m k_B T$.
 
 % longitudinal
-In the longitudinal direction ($z$), the interplay between acceleration, $ a = e E_{DC}/m $, in the DC gun and the physics of the photoemission process determine the properties $\sigma_z$ and $\gamma_z$ of the electron bunch.
+In the longitudinal direction ($z$), the interplay between acceleration, $ a = e E_{DC}/m $, in the DC gun and the physics of the photoemission process determine the initial values of $\sigma_z$, the longitudinal spatial variance, and $\gamma_z$, the longitudinal momentum chirp, of an electron bunch.
 Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) when the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than the distance traveled by electrons moving with a constant velocity $ v_{0} = \sqrt{ 2 \Delta E / m } $, the maximum photoemission velocity, over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
@@
 In the short pulse limit, where the acceleration in the DC gun field over the incident laser pulse duration can be neglected, the initial electron bunch properties are determined primarily by the photoemission physics.
-Since the electric field has no time to influence the front of the pulse before the back is also generated we can say that the front of the pulse has a velocity equal to $v_0$ (the velocity associated with the excess photoemission energy) and the back is essentially stationary just after generation.
-Further, since the photoemission efficiency goes like $\Delta E^2$ \cite{shalaev_electron_1994} the distribution of emitted electrons vs velocity goes like $v^4$; therefore for a first approximation it is reasonable to say that in this case $\Delta v \approx v_0$.
+Since the electric field has no time to influence the front of the pulse before the back is also generated we can say that the front of the pulse has a velocity equal to $v_0$ and the back is essentially stationary just after generation.
+Further, since the photoemission efficiency is proportional to $\Delta E^2$ \cite{shalaev_electron_1994} the distribution of emitted electrons vs velocity is proportional to $v^4$; therefore to a first approximation it is reasonable to say that in this case $\Delta v \approx v_0$.
 The pulse's initial spatial variance $\Delta z$ is therefore approximately to the spread of distances traveled during the laser pulse duration $\Delta t_L$,
@@
 \end{equation}
-Once again we can use this information to determine a velocity chirp across the pulse in this regime,
+Once again, we can use this information to determine a velocity chirp across the pulse in this regime,
 \begin{equation}
@@
 For simplicity of use, one may combine the initial conditions resulting from these two cases.
-Because they arise from independent processes, the initial spatial uncertainties $\sigma_{ z } ( 0 )$ may be added naively 
+Since they arise from independent processes, the initial spatial uncertainties $\sigma_{ z } ( 0 )$ may be added naively 
 \begin{equation} \label{eq:initial_sigma_z}

### 2013-03-05 inc/extension/initial_conditions/initial_conditions.tex
more andreas changes

@@
 We recognize that these initial conditions for an electron pulse generated by a laser-driven DC gun are idealized in that they neglect several important effects; in particular, the interplay of dispersive ($ \eta_{i}(0) \neq 0 $) and space-charge effects in the acceleration region of the photoelectron gun and transverse and longitudinal effects due to the divergence of the DC field in the vicinity of the gun anode.\cite{berger_dc_2009,togawa_ceb6_2007}
-Such effects and the limits of the validity of \ref{eq:initial_sigma_z}, within the context of the AG model, will be presented elsewhere as they require \textit{a priori} this presented extension of the AG model.
+Such effects and the limits of the validity of \ref{eq:initial_sigma_z}, within the context of the AG model, will be presented in Section \ref{sec:bin_model}.
 However, \ref{eq:summary} also represent a defined set of initial conditions that allows the effect of electron optics (i.e., magnetic lenses and RF cavities) on electron pulse propagation to be more readily understood.
@@
 
-To test the validity of the simplistic model presented above I also designed a bin model that depends on nothing other than basic physics and a ``transmission'' function.
+To test the validity of the simplistic model presented above, I also developed a bin model that depends on nothing other than basic Newtonian kinematics and a simple quantum-mechanical ``transmission'' function over a barrier.
 This function represents the likelihood of emission versus emission energy; the maximum energy being the full excess photoemission energy $\Delta E$.
 
-In the model the space in front of the photocathode is binned into segments in ranges of distance from the photocathode.
-The laser pulse duration is then binned into slices of time for a total duration of $N \tau$ where the peak of the pulse occurs at $N \tau / 2$; $N$ is a number selected to accumulate enough of the total laser pulse for an accurate simulation.
+In the model, the space in front of the photocathode is binned into segments in ranges of distance from the photocathode.
+The total laser pulse duration is then binned into slices of time for a total duration of $N \tau$, a few times the HW1/eM duration $\tau$; $N$ is a number selected to accumulate enough of the total laser pulse for an accurate simulation.
+In this notation the peak of the pulse occurs at time $N \tau / 2$.
 For each slice of time, one can determine the relative number of electrons generated in that slice (normalized to the total number that will be created by the end of the simulation), and the velocity distribution of those electrons.
-Knowing the remaining time in the simulation and the acceleration due to the electric field, the velocity distribution can be used to place these electrons into final spatial bins.
+Knowing the remaining time in the simulation and the acceleration due to the electric field, the velocity distribution can be used to place these electrons into their final spatial bins (ignoring Coulomb repulsion on these short time scales).
 Additionally, as the electrons are placed into these bins, their average velocity is stored as well.
 
-This process repeats for each time slice until the finish time, at which point the laser has essentially stopped creating electrons.
+This simulation repeats for each time slice until the finish time, at which point the laser has essentially stopped creating electrons.
 At this time, the earliest fastest electrons will be a distance 
 \begin{equation}
-  d_{max} = v_0 N \tau + \frac{e E_{ DC }}{2 m} (N \tau)^2 
+  z_{max} = v_0 N \tau + \frac{e E_{ DC }}{2 m} (N \tau)^2 
 \end{equation}
-away from the the photocathode, and the the electrons created in the last slice will be at $d=0$.
+away from the the photocathode, and the the electrons created in the last slice will be at $z=0$.
 By counting the electrons in each spatial bin and fitting to a gaussian, we can determine $\sigma_z(0)$, the inital spatial uncertainty in the pulse.

### 2013-03-05 inc/extension/initial_conditions/bin_model_figure.tex
all notes from andreas included for initial conditions section

@@
     Compared are: \subref{fig:bin_model_acc} the DC acceleration dominated case, \subref{fig:bin_model_vmax} the photoemission energy dominated case, and \subref{fig:bin_model_equal} an intermediate case.
-    For $\sigma_z$ (red) and $\gamma_z$ (green) the simulated data (solid) and the fit to the AG allowed form (dashed) are decently well matched in all cases.
-    Unfortunately, the model only allows for a constant $\eta_z$ (green) and in all but the extreme acceleration dominated case \subref{fig:bin_model_acc} the model cannot well represent the simulated data.
+    For the Gaussian spatial distribution whose variance is $\sigma_z$ (red) and the momemtum chirp $\gamma_z$ (green) the simulated data (solid) and the fit to the AG allowed form (dashed) are decently well matched in all cases.
+    Unfortunately, the model only allows for a constant local momentum variance $\eta_z$ (green) and in all but the extreme acceleration dominated case \subref{fig:bin_model_acc} the model cannot well represent the simulated data.
   }

### 2013-03-05 inc/extension/initial_conditions/initial_conditions.tex
all notes from andreas included for initial conditions section

@@
 \begin{equation}
-  \eta_{ z } ( 0 ) \approx \frac{ \eta_{ \smallT } ( 0 )}{ 4 } \punct{.}
+  \eta_{ z } ( 0 ) \approx \frac{ \eta_{ \smallT } ( 0 )}{ 4 } \punct{,}
 \end{equation}
+which is certainly expected to be reasonably accurate in the short-pulse limit, where the acceleration has had no effect (see Section \ref{sec:bin_model} for further investigation).
 
@@
 We recognize that these initial conditions for an electron pulse generated by a laser-driven DC gun are idealized in that they neglect several important effects; in particular, the interplay of dispersive ($ \eta_{i}(0) \neq 0 $) and space-charge effects in the acceleration region of the photoelectron gun and transverse and longitudinal effects due to the divergence of the DC field in the vicinity of the gun anode.\cite{berger_dc_2009,togawa_ceb6_2007}
-Such effects and the limits of the validity of \ref{eq:initial_sigma_z}, within the context of the AG model, will be presented in Section \ref{sec:bin_model}.
+Such effects and the limits of the validity of the longitudinal terms of \ref{eq:summary}, within the context of the AG model, will be presented in Section \ref{sec:bin_model}.
 However, \ref{eq:summary} also represent a defined set of initial conditions that allows the effect of electron optics (i.e., magnetic lenses and RF cavities) on electron pulse propagation to be more readily understood.
@@
 away from the the photocathode, and the the electrons created in the last slice will be at $z=0$.
-By counting the electrons in each spatial bin and fitting to a gaussian, we can determine $\sigma_z(0)$, the inital spatial uncertainty in the pulse.
-Taking a weighted average of the velocity of the electrons in each spatial bin and fitting to a linear function across the pulse (as allowed by the AG model), we can determine the velocity (momentum) chirp $\gamma_z(0)$.
+By counting the electrons in each spatial bin and fitting to a gaussian, we can determine $\sigma_z(0)$, the inital longitudinal spatial variance of the pulse.
+Taking a weighted average of the velocity of the electrons in each spatial bin and fitting to a linear function across the pulse (as defined by the AG model), we can determine the initial longitudinal velocity (momentum) chirp $\gamma_z(0)$.
 As seen in \ref{fig:bin_model}, the shapes of the actual output compare well with the fitted functional form for both of these parameters for all reasonable ranges of field and laser pulse duration.
@@
 
-Finally the velocity uncertainty in each pulse can be computed, however unlike the previous parameters, the AG model only allows for a constant value of local momentum uncertainty across the pulse $\eta_z(0)$.
-\ref{fig:bin_model} shows that this limitation on functional form severely limits the accuracy with which the determined value of $\eta_z(0)$ represents the actual distribution of momentum uncertainties computed by the binning model.
+Finally, the velocity uncertainty in each bin can be computed, however unlike the previous parameters, the AG model only allows for a constant value of $\eta_z(0)$, the local momentum variance across the pulse.
+Unfortunately, \ref{fig:bin_model} clearly demonstrates that the local momentum variance is not constant over the inital photo-generated electron bunch; in fact it varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch.
+The narrowest momentum distribution are clearly at the front of the bunch where only the earliest generated electrons with initial velocity near $ v_{ 0 } $ initially reside; the largest distribution is at the back where the last electrons generated still haven't had time to seperate from one another.
 
-Additionally the iterative time-step analysis clearly demonstrates that the local longitudinal momentum variance $ \eta_{ z } $ is not constant over the inital photo-generated electron bunch.
-In fact, $ \sqrt{ \eta_{ z } ( 0 ) } $ varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch with the narrowest momentum distribution at the front of the bunch where only the faster electrons with velocity near $ v_{ 0 } $ initially reside.
-Unfortunately, the self-similar theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch; to allow a more realistic fit to the simulated initial data, the pulse would eventually evolve to be non-Gaussian.
+This limitation on functional form severely limits the accuracy with which the determined value of $\eta_z(0)$ represents the actual distribution of momentum uncertainties computed by the binning model.
+The self-similar theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch; with a more realistic fit to the simulated initial data, the pulse would eventually evolve to be non-Gaussian.
 %Consequently, we will assume that $ \eta_{ z } $ is a constant over the bunch with a value given by $ \eta_{\smallT } ( 0 ) \cong 3.85 \eta{ z } ( 0 ) $: indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.

### 2013-03-05 inc/extension/initial_conditions/initial_conditions.tex
a few more tweaks

@@
 
-This limitation on functional form severely limits the accuracy with which the determined value of $\eta_z(0)$ represents the actual distribution of momentum uncertainties computed by the binning model.
+This limitation on functional form severely limits the accuracy with which a determined spatially constant value of $\eta_z(0)$ represents the actual distribution of momentum uncertainties computed by the binning model.
 The self-similar theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch; with a more realistic fit to the simulated initial data, the pulse would eventually evolve to be non-Gaussian.

### 2013-03-05 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
improve formal external_forces section
merging comments on Liouville's theorem

--- 2012-09-24 inc/liouvilles_theorem/liouvilles_theorem.tex
@@
 
-\section{Implications of Liouville's Theorem}
+\section{Implications of Liouville's Theorem} \label{sec:liouville}
 
@@
 %TODO find a good reference for this
+In the AG model the component phase-space products are simply $\sigma_{\alpha} \eta_{\alpha}$ where as before $ \alpha = \smallT, z$.
 
-To evaluate the effect of this theorem on these equations, consider the time derivative of the product $\sigma_{\alpha} \eta_{\alpha}$.
+In fact it is easy to show that the AG model explicitly obeys the Liouville Theorem.
+To do so, consider the time derivative of these component phase-space products.
 Using \ref{eq:ag_original} it is easy to show that
-\begin{gather}
-  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0\\
-  \label{eq:define_big_gamma}
-  \therefore \sigma_{\alpha} \eta_{\alpha} \equiv \Gamma_{\alpha}^2 = \text{Constant}
-\end{gather}
-This analysis shows that the Analytic Gaussian model explicitly obeys the Liouville Theorem.
+\begin{equation}
+  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0
+\end{equation}
+and therefore we can define the per-simulation constants $\Gamma_{\alpha}$ as
+\begin{equation} \label{eq:define_big_gamma}
+  \Gamma_{\alpha}^2 \equiv \sigma_{\alpha} \eta_{\alpha} = \text{Constant .}
+\end{equation}
 
 Not only is this an interesting result, but it can be both instructive and computationally useful.
-By rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$, evaulating at some time (presumably $t=0$) and substituting into \ref{eq:ag_original}.
+By evaluating $\Gamma_{\alpha}$ at some time (presumably $t=0$) and rewriting \ref{eq:define_big_gamma} as $\eta_{\alpha} = \Gamma_{\alpha}^2 / \sigma_{\alpha}$, we can rewrite \ref{eq:ag_original} as
 \begin{subequations} \label{eq:ag_big_gamma}
@@
   \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
-  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2(0) + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
-    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi)
+  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
+    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) \, \text{.} \label{eq:ag_big_gamma_gamma}
 \end{gather}
 \end{subequations}
-these four equations now contain all of the information previously contained in six.
+These four equations now contain all of the information previously contained in six (\ref{eq:ag_original}).
 Having fewer equations, and fewer mathematical operations, makes this set of equations much less computationally intensive.
-Further, now one can see that these $\Gamma$ constants can be thought of as some minimum/intrinsic action of the pulse. 
+Further, when viewed in comparison to the chirps $\gamma_{\alpha}$, one can now see that these $\Gamma_{\alpha}$ constants can be thought of as some minimum/intrinsic action of the pulse.
 This will become more apparent when in future sections

### 2013-03-05 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
merge the rest of the comments from andreas in sections 2-3.2

@@
 
-In the transverse dimension, this is consistent with the definition of `coherent fluence' employed by Reed et al.\cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the spatial emittance of an electron beam \cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
+We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed et al.\cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the trasverse spatial emittance of an electron beam \cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
+%TODO include more commentary on $\varepsilon_{\smallT}$?

### 2013-03-05 inc/extension/external_forces/external_forces.tex
consistent use of v_max vs v_0

@@
 The order parameter defines the ``sharpness'' of the edge of the region of influence with integer values ranging from 1 (a simple Gaussian) through $\infty$ (a top-hat).
-We connect $z^{\prime}$ to $t$ through $ v_{{ \scriptscriptstyle 0}} $; the kinetic equations are simply defined for the position of the peak of the pulse.
+We connect $z^{\prime}$ to $t$ through $ v_{\smallzero} $; the kinetic equations are simply defined for the position of the peak of the pulse.
 
@@
 \begin{equation} \label{eq:RF_Efield}
-  E ( r , z , z^{\prime} ) = E_{0} \sin \left ( \frac{ \Omega ( z^{\prime} - z ) }{ v_{{ \scriptscriptstyle 0}} } + \phi \right )\operatorname{J_{0}} \left ( \frac{ k r }{ a } \right ) \text{,}
+  E ( r , z , z^{\prime} ) = E_{0} \sin \left ( \frac{ \Omega ( z^{\prime} - z ) }{ v_{\smallzero} } + \phi \right )\operatorname{J_{0}} \left ( \frac{ k r }{ a } \right ) \text{,}
 \end{equation}
@@
 
-For the usual case of a spatially compact electron pulse with respect to the dimension of the RF cavity (i.e., $ \Omega z \ll v_{{ \scriptscriptstyle 0}} $ and $ \sqrt{\sigma_{\smallT}} \ll a $), \ref{eq:RF_Efield} can be rewritten as
+For the usual case of a spatially compact electron pulse with respect to the dimension of the RF cavity (i.e., $ \Omega z \ll v_{\smallzero} $ and $ \sqrt{\sigma_{\smallT}} \ll a $), \ref{eq:RF_Efield} can be rewritten as
 \begin{equation} \label{eq:RF_Efield_approx}
-  E ( r , z , z^{\prime} ) = E_{0} \biggl [ \sin \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right ) - \frac{ \Omega z }{ v_{{ \scriptscriptstyle 0}} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right ) \biggr ] \text{,}
+  E ( r , z , z^{\prime} ) = E_{0} \biggl [ \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) - \frac{ \Omega z }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \biggr ] \text{,}
 \end{equation}
@@
 \begin{equation} \label{eq:RF_Force_z}
-  \boldsymbol{F}_{ z , RF } = - \left [ \frac{ e \Omega E_{0} }{ v_{{ \scriptscriptstyle 0}} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right ) \right ] z \hat{\operatorname{ \boldsymbol{z} }} \text{.}
+  \boldsymbol{F}_{ z , RF } = - \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] z \hat{\operatorname{ \boldsymbol{z} }} \text{.}
 \end{equation}
 
-By considering the impulse exerted by the RF cavity field on the electrons in the pulse, it is straightforward to show that pulse compression (when $\phi = 0$) is maximized for $ \Omega d / (2 v_{{ \scriptscriptstyle 0}} ) = \pi /2 $; that is, the time of flight of the electron pulse through the RF cavity of length $ d $ is equal to half the RF period.
-Such a calculation assumes that the pulse velocity $v_{{ \scriptscriptstyle 0}}$ is constant; so that, the condition that $ \phi = 0 $ when the pulse is at the center of the RF cavity (i.e., when the pulse is at $ z^{\prime} = 0 $) is readily met.
+By considering the impulse exerted by the RF cavity field on the electrons in the pulse, it is straightforward to show that pulse compression (when $\phi = 0$) is maximized for $ \Omega d / (2 v_{\smallzero} ) = \pi /2 $; that is, the time of flight of the electron pulse through the RF cavity of length $ d $ is equal to half the RF period.
+Such a calculation assumes that the pulse velocity $v_{\smallzero}$ is constant; so that, the condition that $ \phi = 0 $ when the pulse is at the center of the RF cavity (i.e., when the pulse is at $ z^{\prime} = 0 $) is readily met.
 As suggested by the acceleration term in \ref{eq:RF_Efield_approx}, this is only a good approximation if the RF field amplitude $E_{0}$ is sufficiently small to not affect the velocity of the peak of the pulse as it propagates through the cavity.
@@
   \boldsymbol{F}_{ r , RF } = e E_{0} \left [ 
-    \frac{ v_{{ \scriptscriptstyle 0}} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{{ \scriptscriptstyle 0}} } + \phi \right )
+    \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
    \right ] r \hat{\operatorname{ \boldsymbol{r} }} \text{.}
@@
 In obtaining \ref{eq:RF_Force_r}, which is also of the form for use in \ref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \ref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
-For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{{\scriptscriptstyle 0}} / \Omega $,\cite{kim_rf_1989} even though its magnitude is maximized for $\phi = 0$.
+For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $,\cite{kim_rf_1989} even though its magnitude is maximized for $\phi = 0$.
 Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.

### 2013-03-05 inc/extension/initial_conditions/initial_conditions.tex
consistent use of v_max vs v_0

@@
 In the longitudinal direction ($z$), the interplay between acceleration, $ a = e E_{DC}/m $, in the DC gun and the physics of the photoemission process determine the initial values of $\sigma_z$, the longitudinal spatial variance, and $\gamma_z$, the longitudinal momentum chirp, of an electron bunch.
-Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) when the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than the distance traveled by electrons moving with a constant velocity $ v_{0} = \sqrt{ 2 \Delta E / m } $, the maximum photoemission velocity, over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
+Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) when the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than the distance traveled by electrons moving with a constant velocity $ v_{max} = \sqrt{ 2 \Delta E / m } $, the maximum photoemission velocity, over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
 
@@
 In the short pulse limit, where the acceleration in the DC gun field over the incident laser pulse duration can be neglected, the initial electron bunch properties are determined primarily by the photoemission physics.
-Since the electric field has no time to influence the front of the pulse before the back is also generated we can say that the front of the pulse has a velocity equal to $v_0$ and the back is essentially stationary just after generation.
-Further, since the photoemission efficiency is proportional to $\Delta E^2$ \cite{shalaev_electron_1994} the distribution of emitted electrons vs velocity is proportional to $v^4$; therefore to a first approximation it is reasonable to say that in this case $\Delta v \approx v_0$.
+Since the electric field has no time to influence the front of the pulse before the back is also generated we can say that the front of the pulse has a velocity equal to $v_{max}$ and the back is essentially stationary just after generation.
+Further, since the photoemission efficiency is proportional to $\Delta E^2$ \cite{shalaev_electron_1994} the distribution of emitted electrons vs velocity is proportional to $v^4$; therefore to a first approximation it is reasonable to say that in this case $\Delta v \approx v_{max}$.
 The pulse's initial spatial variance $\Delta z$ is therefore approximately to the spread of distances traveled during the laser pulse duration $\Delta t_L$,
 \begin{equation} 
-  \sqrt{ \sigma_{ z } ( 0 ) } \approx \Delta z = \Delta v \Delta t_{L} = \frac{ v_{ 0 } \tau }{ \sqrt{ 2 } } \punct{.}
+  \sqrt{ \sigma_{ z } ( 0 ) } \approx \Delta z = \Delta v \Delta t_{L} = \frac{ v_{max} \tau }{ \sqrt{ 2 } } \punct{.}
 \end{equation}
@@
 \begin{equation}
-  \gamma_{ z } ( 0 ) \approx m \Delta v \Delta z = m v_{ 0 } \sqrt{ \sigma_{ z } ( 0 ) } \punct{,}
+  \gamma_{ z } ( 0 ) \approx m \Delta v \Delta z = m v_{max} \sqrt{ \sigma_{ z } ( 0 ) } \punct{,}
 \end{equation}
@@
 \begin{equation} \label{eq:initial_sigma_z}
-  \sigma_{ z } ( 0 ) \approx \frac{ ( v_{ 0 } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \punct{.}
+  \sigma_{ z } ( 0 ) \approx \frac{ ( v_{max} \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \punct{.}
 \end{equation}
@@
 \begin{equation}
-  \gamma_{ z } ( 0 ) \approx \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \punct{.}
+  \gamma_{ z } ( 0 ) \approx \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{max} + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \punct{.}
 \end{equation}
@@
     \sigma_{ \smallT } ( 0 ) \approx \frac{ w^{ 2 } }{ 2 } \\
-    \sigma_{ z } ( 0 ) \approx \frac{ ( v_{ 0 } \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
+    \sigma_{ z } ( 0 ) \approx \frac{ ( v_{max} \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
     \gamma_{ \smallT } ( 0 ) \approx 0 \\
-    \gamma_{ z } ( 0 ) \approx \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{ 0 } + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
+    \gamma_{ z } ( 0 ) \approx \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{max} + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
     \eta_{ \smallT } ( 0 ) \approx \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \\
@@
 \begin{equation}
-  z_{max} = v_0 N \tau + \frac{e E_{ DC }}{2 m} (N \tau)^2 
+  z_{max} = v_{max}N \tau + \frac{e E_{ DC }}{2 m} (N \tau)^2 
 \end{equation}
@@
 Unfortunately, \ref{fig:bin_model} clearly demonstrates that the local momentum variance is not constant over the inital photo-generated electron bunch; in fact it varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch.
-The narrowest momentum distribution are clearly at the front of the bunch where only the earliest generated electrons with initial velocity near $ v_{ 0 } $ initially reside; the largest distribution is at the back where the last electrons generated still haven't had time to seperate from one another.
+The narrowest momentum distribution are clearly at the front of the bunch where only the earliest generated electrons with initial velocity near $ v_{max} $ initially reside; the largest distribution is at the back where the last electrons generated still haven't had time to seperate from one another.
 

### 2013-03-05 inc/model_results/shapes/shapes.tex
consistent use of v_max vs v_0

@@
 
-The initial `space-time' eccentricity \ref{eq:define_xi}, $\xi ( 0 ) = v_{{ \scriptscriptstyle 0}} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
+The initial `space-time' eccentricity \ref{eq:define_xi}, $\xi ( 0 ) = v_{\smallzero} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
 \ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and `cigar-like' ($ \sigma_{ z } ( 0 ) > \sigma_{ \smallT } ( 0 ) $) pulses.
@@
 Since the initial pulse volumn is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} ($ N = 10^{6} $ electrons in the bunch) than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} ($ N = 10^{0} $).
-The pulse velocity is $ v_{{ \scriptscriptstyle 0}} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.
+The pulse velocity is $ v_{\smallzero} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.
 

### 2013-03-07 inc/extension/external_forces/formal.tex
added magnetic lens loops comparison figure and moved some of the related content from the hardware section to the extension

--- 2013-03-05 inc/extension/external_forces/external_forces.tex
@@
 Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
-
-\section{Practical Extension to Include Magnetic Lenses, RF Cavities and DC Accelerators}
-
-\subsection{Region of Influence}
-
-To specify the physical region of influence of the lenses, it is convenient to redefine $M_{i}$ to include a super-Gaussian envelope in $z^{\prime}$, the position of the peak of the pulse in the lab frame;
-\begin{equation} \label{eq:reg_of_influence}
-  M_{i}	\to M_{i} \exp \left [ - \left (  \frac{ z^{\prime} - z_{lens}^{\prime} }{ L_{lens} / 2 } \right )^{ 2 n } \right ] \text{ ,}
-\end{equation}
-where $z_{lens}^{\prime}$ is the position of the center of the lens, $L_{lens}$ is the length of the lens and $n$ is the super-Gaussian order parameter.
-The order parameter defines the ``sharpness'' of the edge of the region of influence with integer values ranging from 1 (a simple Gaussian) through $\infty$ (a top-hat).
-We connect $z^{\prime}$ to $t$ through $ v_{\smallzero} $; the kinetic equations are simply defined for the position of the peak of the pulse.
-
-\subsection{Magnetic Lenses}
-
-In the transverse ($ \smallT $) direction, a magnetic lens may be modeled using one linear force contribution.
-Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is surprisingly complicated to describe \cite{montgomery_some_1961} and the ``lensing'' action is in fact a compound force (first rotating the beam, then lensing).
-Further discussion of the mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens}.
-We have found that in practice, characterizing a specific magnetic lens ($M_{\smallT}^{mag}(I)$ where $I$ is the current in the magnetic lens) is much easier than attempting to find a generic functional form.
-
-For localizing the region of influence of the lens, I have commonly modelled a thin magnetic lens as having a super-Gaussian order value $n=1$, a simple Gaussian; of course this may depend on the specifics of the lens.
-To simulate the effect of magnetic lenses in a UEM by the extended AG model, for each magnetic lens, add a term like
-\begin{equation} \label{eq:reg_of_influence}
-  M_{\smallT}^{mag} = M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \text{ ,}
-\end{equation}
-to the set of $M_{\smallT}$ parameters, where $z_{mag}^{\prime}$ and $L_{mag}$ are the position and length of the lens in question respectively.
-
-\subsection{RF Cavities}
-
-In the longitudinal ($ z $) direction, a resonant radio-frequency (RF) cavity can be used in a UEM column or UED experiment \cite{oudheusden_electron_2007,fill_sub-fs_2006} to act as a pulse compression or `temporal focusing' element.
-The model of the RF cavity is different from the magnetic lens in two distinct ways.
-First, the RF cavity contains several different fields and thus contributes several different linear force terms to the overall equation set.
-Second, though there are more terms, their magnitudes are easily connected to the physical characteristics of the cavity.
-
-Although there are many possible resonant cavity structures,\cite{oudheusden_electron_2007,humphries_principles_1986} for simplicity, this analysis will consider only the evacuated cylindrical (or `pillbox') $\text{TM}_{010}$ cavity for which the electrons propagate down the axis.\cite{fill_sub-fs_2006,humphries_principles_1986}
-The known mathematical forms of the electric and magnetic components of this resonant mode together with a super-Gaussian envelope function (\ref{eq:reg_of_influence}) allow the three main effects of a pill-box RF cavity element on the electron pulse to be described within the approximation of the linear force extension to the AG model: the primary action of the pulse compression (or acceleration and expansion dependent upon the relative RF phase), and the transverse lens effects due to the magnetic component of the $\text{TM}_{010}$ resonance and the electric RF cavity fringe fields at its entrance and exit apertures.\cite{kim_rf_1989}
-
-The spatial dependence of the axial RF electric field experienced by the electron pulse in a TM$_{010}$ cavity of radius $a$ may be written as
-\begin{equation} \label{eq:RF_Efield}
-  E ( r , z , z^{\prime} ) = E_{0} \sin \left ( \frac{ \Omega ( z^{\prime} - z ) }{ v_{\smallzero} } + \phi \right )\operatorname{J_{0}} \left ( \frac{ k r }{ a } \right ) \text{,}
-\end{equation}
-where $ E_{ 0 } $ is the oscillating field amplitude, $\Omega$ is the RF frequency and $\phi$ is a phase factor determining whether the RF cavity acts as a pulse compressor $ ( \phi \approx 0 ) $ or an electron accelerator $ ( \phi \approx \pm \pi/2 ) $.
-The coordinate $z^{\prime} $ describes the position of the center of the pulse in the RF cavity, which extends from $ z^{\prime} = - d / 2 $ to $ z^{\prime} = d / 2 $, and $z$ is the position of any electron in the pulse with respect to the center of the pulse.
-For the $\text{TM}_{010}$ oscillation mode, the RF frequency is given by
-\begin{equation}
-  \Omega = \frac{ c k }{a} \text{,}
-\end{equation}
-where $c$ is the speed of light in a vacuum and $ k = 2.405 $ is the first zero of the $\operatorname{ J_{0} }$ Bessel function.
-
-For the usual case of a spatially compact electron pulse with respect to the dimension of the RF cavity (i.e., $ \Omega z \ll v_{\smallzero} $ and $ \sqrt{\sigma_{\smallT}} \ll a $), \ref{eq:RF_Efield} can be rewritten as
-\begin{equation} \label{eq:RF_Efield_approx}
-  E ( r , z , z^{\prime} ) = E_{0} \biggl [ \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) - \frac{ \Omega z }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \biggr ] \text{,}
-\end{equation}
-which demonstrates that the RF cavity acts both as a pulse acceleration (first term on RHS of \ref{eq:RF_Efield_approx}) and a pulse compression or expansion (second term on RHS of \eqref{eq:RF_Efield_approx}) device.
-The second term then gives the desired form (see \ref{eq:linear_force}) for the force manipulating the temporal dynamics and characteristics during electron pulse compression;
-\begin{equation} \label{eq:RF_Force_z}
-  \boldsymbol{F}_{ z , RF } = - \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] z \hat{\operatorname{ \boldsymbol{z} }} \text{.}
-\end{equation}
-
-By considering the impulse exerted by the RF cavity field on the electrons in the pulse, it is straightforward to show that pulse compression (when $\phi = 0$) is maximized for $ \Omega d / (2 v_{\smallzero} ) = \pi /2 $; that is, the time of flight of the electron pulse through the RF cavity of length $ d $ is equal to half the RF period.
-Such a calculation assumes that the pulse velocity $v_{\smallzero}$ is constant; so that, the condition that $ \phi = 0 $ when the pulse is at the center of the RF cavity (i.e., when the pulse is at $ z^{\prime} = 0 $) is readily met.
-As suggested by the acceleration term in \ref{eq:RF_Efield_approx}, this is only a good approximation if the RF field amplitude $E_{0}$ is sufficiently small to not affect the velocity of the peak of the pulse as it propagates through the cavity.
-As a result, in practice, the phase $\phi$ of the RF field is adjusted to optimize the performance of the cavity.
-
-The two main effects contributing to the transverse lens of a $\text{TM}_{010}$-mode RF cavity, its magnetic field component and electric fringe field aperture effects, may also be included within the linear force approximation.
-The azimuthal magnetic component of the resonant mode oscillates 90$^{\circ}$ out of phase with the electric component and with a $\operatorname{ J_{1} }$ Bessel radial form; i.e., $ \operatorname{ J_{1} }( k r / a ) \sim k r / a $ for small $r$.
-Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $,\cite{kim_rf_1989} where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\ref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\ref{eq:reg_of_influence}).
-In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures, as illustrated by $1 \frac{1}{2}$-cell RF photo-gun designs.\cite{mcdonald_design_1988}
-For the case of an evacuated cavity, these two contributions result in a net force,
-\begin{equation} \label{eq:RF_Force_r}
-  \boldsymbol{F}_{ r , RF } = e E_{0} \left [ 
-    \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
-   \right ] r \hat{\operatorname{ \boldsymbol{r} }} \text{.}
-\end{equation}
-In obtaining \ref{eq:RF_Force_r}, which is also of the form for use in \ref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \ref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
-For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $,\cite{kim_rf_1989} even though its magnitude is maximized for $\phi = 0$.
-Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.
-
-% excluded example section from JAP paper as it will need new graphics

### 2013-03-07 inc/extension/external_forces/practical.tex
added magnetic lens loops comparison figure and moved some of the related content from the hardware section to the extension

--- 2013-03-05 inc/extension/external_forces/external_forces.tex
@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
-
-\section{Formal Extension to Include Generic External Forces} \label{sec:external_forces}
-
-Simulation of electron pulse propagation in UEM and modern UED apparatus\cite{siwick_ultrafast_2002,oudheusden_electron_2007,reed_femtosecond_2006} requires the ability to model electron-optical elements designed to control the three-dimensional space-charge and dispersion induced broadening of the pulse, and hence deliver electron pulses of the desired quality to the specimen.
-The Analytic Gaussian model as presented by Michalik and Sipe (see Section \ref{sec:ms_model}) is only useful for modeling free-space propagation.
-This is because there is no allowance for external forces, such as those imparted by the ``optical'' elements of an electron microscope column.
-In this section, I show that the AG model can be extended to include external forces on the pulse, such as the influence of magnetic electron lenses and RF pulse compression cavities.\cite{oudheusden_electron_2007,veisz_hybrid_2007}
-
-\subsection{Mathematical Formulation of External Force Contribution}
-
-While the free-space AG model does not contain any external forces, it does include an internal force.
-The Coulomb force is included to model the repulsion of the pulse's electrons from each other.
-Because this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identically ``Gaussian'' optics.\cite{michalik_analytic_2006}
-
-The internal Coulombic repulsion force enters the model in \ref{eq:potential_integral}.
-Since both potentials and integrals are linear, one may add additional potentials so that it becomes
-\begin{equation}
-  V(\vec{r} - \vecprime{r}) \rightarrow V^{int}(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r})
-\end{equation}
-In doing so, we rename the old potential, the internal repulsive Coulomb force, $V^{int}$.
-Following these new terms through the derivation, we see that
-\begin{equation}
-  \begin{split}
-    \Phi(\vec{r};t) = & \int d\vecprime{r} \left( V^{int}(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r}) \right) n(\vecprime{r}; t) \\
-    = & \int d\vecprime{r} V^{int}(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) + \sum_{k} \int d\vecprime{r} V_{k}^{ext}(\vec{r}) n(\vecprime{r}; t) \\
-    \equiv & \Phi^{int}(\vec{r};t) + \sum_{k} \Phi_{k}^{ext}(\vec{r};t)
-  \end{split}
-\end{equation}
-Therefore each external force contributes an additional term defined as
-\begin{equation}
-  \Phi^{ext}(\vec{r};t) \equiv \int d\vecprime{r} V^{ext}(\vec{r}) n(\vecprime{r}; t)
-\end{equation}
-These external terms are not self-influencing, and thus are not subject to the complexities of the internal Coulombic repulsion force.
-For consistency we have now defined $\Phi^{int}(\vec{r};t)$ as the old $\Phi(\vec{r};t)$, which results from the Coulomb force.
-
-This new form of $\Phi(\vec{r};t)$ can be substituted into the definition of $K^{force}$ matrix (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector} 
-\begin{equation}
-  \begin{split}
-    K^{force}_{ij} = & \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial}{\partial \vec{r}} \left( \Phi(\vec{r};t) + \sum_{k} \Phi_{k}^{ext}(\vec{r};t) \right) d\vec{r} d\vec{p} \\
-    \equiv & K^{int}_{ij} + \sum_{k} (K_{k}^{ext})_{ij}
-  \end{split}
-\end{equation}
-Like the potential integrals, we are left with an additional matrix for each external force.
-Again, for consistency, we have defined $K^{int}$ as the $K$ matrix resulting from the internal Coulombic repulsion.
-
-Noting that the spatial derivative of a potential is just the force, we can now define the matrix contributed by each external force is given by
-\begin{equation} \label{eq:Kint}
-  K^{ext}_{ij} = \frac{1}{N} \iint u_i u_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
-\end{equation}
-Finally, the general form of \ref{eq:gaussian_integral_dt_k} for (possibly multiple) external force ($\vec{F}_{ext}$) contributions is now
-\begin{equation}
-  \frac{\partial}{\partial t} A^{-1}_{ij} = K^{flow}_{ij} + K^{int}_{ij} + \sum K^{ext}_{ij}
-\end{equation}
-
-\subsection{Constant External Forces} \label{sec:const_force}
-
-For constant static forces having no dependence in $ x $, $ y $, $ z $, or $ t $, one can readily show that there is no effect on the pulse propagation parameters $ \sigma_{ \alpha } $, $ \eta_{ \alpha } $, and $ \gamma_{ \alpha } $.
-For example, consider a field consider a field $ \vec{E} = -E \hat{z} $ such that electrons are accelerated in the positive $ z $ direction by $ \vec{F}_{ext} = qE \hat{z} = q ( V / d ) \hat{z} $, so that from \ref{eq:Kint} we have
-\begin{equation}
- \begin{split} 
-  K^{ext}_{ij} = \frac{1} {N} \dfrac{qV} {d} \int{ u_{i} u_{j} \left( \frac{p_{z}} {\eta_{z}} - \frac{\gamma_{z} z} {\sigma_{z} \eta_{z}} \right) f d^{3}x   d^{3}p } ;
- \end{split}
-\end{equation}
-but it is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ (or equivalently $ K^{ext} = \hat{0} $) meaning that this static field has no effect on the parameters of the pulse.
-Of course, this is only true in the pulse frame, the frame of the simulation, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m_{e} $ is unchanged.
-In the lab frame however, where  $ \Delta E = p \Delta p / m_{e} $, there is an increase in the momentum spread, as the pulse now has the large momentum $p$ imparted by the accelerating field.
-
-\subsection{Linear External Forces}
-
-%External forces that vary with position in the pulse reference frame (i.e., are not constant) act to change the electron pulse propagation dynamics, most often by affecting $\gamma_{\alpha}$ --- the spatial dependence of the local relative mean electron momenta.
-%Below, we consider two examples of electron optical elements that are being used, or are under consideration, for controlling and manipulating the space-time dynamics of electron pulses in DTEM and UED instrumentation; namely, magnetic lenses for spatial focusing and RF pulse compression cavities.
-
-For the case of a perfect lens, or in the longitudinal direction a perfect RF compression element, with a parabolic potential, the external force is linear with distance from the center of the electron pulse:
-\begin{equation} \label{eq:linear_force}
-\vec{F}_{ext} = - M_{ \smallT } r \hat{\operatorname{r}} - M_{z} z \hat{\operatorname{z}} \text{ .}
-\end{equation}
-Here, $M_{ \smallT }$ and $M_{z}$ characterize the strength of a lens or a compressor, respectively, and are chosen to be positive for focusing/compression elements.
-Substitution into the integral equation for the $ K^{ext} $ matrix elements,
-\begin{equation}
- \begin{split} 
-  K^{ext}_{ij} & \equiv \dfrac{1} {N} \int u_{i} u_{j} ( \vec{\nabla}_{p} f ) \cdot \vec{F}_{ext} d^{3} x d^{3} p \\ & = \dfrac{1} {N} \int u_{i} u_{j} \biggl[ x \left( \dfrac{ p_{x} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } x } { \sigma_{ \smallT } \eta_{ \smallT } } \right) + y \left( \dfrac{ p_{y} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } y } { \sigma_{ \smallT } \eta_{ \smallT } } \right) \\ & \quad + z \left( \dfrac{ p_{z} } { \eta_{z} } - \dfrac{ \gamma_{z} z } { \sigma_{z} \eta_{z} } \right) \biggr] f d^{3} x d^{3} p ,
- \end{split}
-\end{equation}
-leads to the $6\times6$ matrix
-\begin{equation}
-K^{ext} = 
-\begin{pmatrix}
-\hat{k}^{lin}_{\smallT} & \hat{0} & \hat{0} \\
-\hat{0} & \hat{k}^{lin}_{\smallT} & \hat{0} \\
-\hat{0} & \hat{0} & \hat{k}^{lin}_{z}
-\end{pmatrix} \text{ ,}
-\end{equation}
-where
-\begin{equation}
-\hat{k}^{lin}_{\alpha} =
-\begin{pmatrix}
-0 & M_{\alpha} \sigma_{\alpha} \\
-M_{\alpha} \sigma_{\alpha} & 2 M_{\alpha} \gamma_{\alpha} \text{.}
-\end{pmatrix} \text{ .}
-\end{equation}
-which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.
-This derivation has been for a single force, in practice however there may be multiple forces, therefore \ref{eq:ag_big_gamma_gamma} may be restated as
-\begin{equation}
-  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2(0) + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
-    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum M_{\alpha} 
-\end{equation}
-
-Since these terms are only added to $\dfrac{d \gamma_{\alpha}}{d t}$, the analysis of Liouville's Theorem, shown in Section \ref{sec:liouville}, is unaffected.
-The extended AG model still explicitly satisfies Liouvilles's Theorem.
-
-\subsection{Validity and Limitations}
-
-The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations.\cite{michalik_analytic_2006}
-As a result, we emphasize that the presented AG model extension is valid only for the action of linear forces (i.e., forces that have a linear dependence on distance in the pulse frame) on the electron pulse due to the model's self-similar Gaussian approximation.
-The inclusion of the action of nonlinear forces (e.g., due to space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations, which should be employed to gauge the accuracy (and hence limits) of our extended AG approach at some point.
-
-It should also be noted that these analyses are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
-Extension to include relativistic effects is conceptually straightforward using standard transformations between the laboratory and electron pulse reference frames.
-
-Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid.\cite{luiten_how_2004}
-This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
-Consequently, the AG approach is applicable to both the single electron per pulse limit,\cite{lobastov_four-dimensional_2005} where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate.\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}
-It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
-Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
 
@@
 
-In the transverse ($ \smallT $) direction, a magnetic lens may be modeled using one linear force contribution.
-Unfortunately, it is not practical to write an analytic form of the magnitude of this term; the field distribution is surprisingly complicated to describe \cite{montgomery_some_1961} and the ``lensing'' action is in fact a compound force (first rotating the beam, then lensing).
-Further discussion of the mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens}.
-We have found that in practice, characterizing a specific magnetic lens ($M_{\smallT}^{mag}(I)$ where $I$ is the current in the magnetic lens) is much easier than attempting to find a generic functional form.
+A simple magnetic lens is just a solenoid whose axis is collinear with the axis of the microscope column.
+%Initially one would think that electrons moving parallel to this axis would not be affected by the magnetic field lines in the solenoid, however the contribution of the fringing field must be taken into consideration.
+As electrons propagate through the solenoid the fringing magnetic field imparts an azimuthal velocity.
+This ``tumbling'' around the axis, along with the axial magnetic field, is what causes the inward-directed force.
+
+A full description of the motion of the electrons is easily derived \cite{el-kareh_electron_1970},
+\begin{subequations} \label{eq:lens_eq_of_motion}
+\begin{gather}
+  m \ddot{r} = -e B_z r \dot{\theta} + m r \dot{\theta}^2                 \\
+  \frac{d}{dt} ( m r^2 \dot{\theta} ) = e r \dot{r} B_z - e r B_r \dot{z} \\
+  m \ddot{z} = e B_r r \dot{\theta}
+\end{gather}
+\end{subequations}
+however, attempting to further seperate these equations quickly highlights the tightly coupled nature of the motion of the electron in this system.
+The full description of the off-axis field of a real (non-infinite) air-core solenoid is also surprisingly complicated.
+Montgomery and Terrell \cite{montgomery_some_1961} provide a full treatment for many lens configurations, including a tractable treatment for a single current loop of radius $a$ in cylindrical coordinates,
+\begin{subequations} \label{eq:field_of_loop}
+\begin{gather}
+  H_z = \frac{2I}{\sqrt{Q}} \left(   F(k) + \frac{ a^2 - r^2 - z^2 }{ (1-k^2) Q } E(k) \right) \\
+  H_r = \frac{2I}{\sqrt{Q}} \left( - F(k) + \frac{ a^2 + r^2 + z^2 }{ (1-k^2) Q } E(k) \right) \\
+  Q \equiv (a+r)^2 + z^2 \\
+  k \equiv \sqrt{ 4 a r / Q }
+\end{gather}
+\end{subequations}
+%%TODO I can't decide whether the Q and k definitions should be numbered or not
+%where
+%\begin{gather*}
+%  Q \equiv (a+r)^2 + z^2 \\
+%  k \equiv \sqrt{ 4 a r / Q }
+%\end{gather*}
+and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
+An even more accurate solenoid can be modeled by appropriately summing the fields of many such loops.
+
+For the AG model, a magnetic lens may be modeled using one linear force contribution in the transverse ($ \smallT $) direction.
+Unfortunately, it is not practical to write an analytic form of the magnitude of this term.
+As seen in \ref{eq:field_of_loop} the field distribution is surprisingly complicated to describe and the ``lensing'' action is in fact a compound force (first rotating the beam, then lensing).
+I have found that in practice, characterizing a specific magnetic lens ($M_{\smallT}^{mag}(I)$ where $I$ is the current in the magnetic lens) is much easier than attempting to find a generic functional form.
+This may be done in comparison to either experimental data or, as shown in \ref{fig:mag_lens_loops}, to numerical simulations based on \ref{eq:field_of_loop}.
+
+\begin{figure}
+  \centering
+  \input{mag_lens_plot}
+  \caption[Comparison of the AG model and a kinematic model of a magnetic lens]{
+    Comparison of electron rays tracked kinematically through a realistic magnetic lens simulation (green) and an AG model simulation of a Guassian electron beam (red) of a similar HW1/eM beam width $ w = 500 \mu\text{m} $.
+    The kinematic model employs ten individual current loops occupying a total length of 635$\mu$m at position $z=0$.
+    Each ray bunch consists of one ray parallel to the axis of propagation and one each having transverse velocity related to an excess photoemission energy $\Delta E \approx $ 0.1eV; the related transverse variance is used in the AG model.
+    The bunches are distributed evenly from $r=0$ to $r = 2 w$.
+  }
+  \label{fig:mag_lens_loops}
+\end{figure}
 
@@
 to the set of $M_{\smallT}$ parameters, where $z_{mag}^{\prime}$ and $L_{mag}$ are the position and length of the lens in question respectively.
+%TODO include discussion and plot of beam size versus aberrations
 
@@
 which demonstrates that the RF cavity acts both as a pulse acceleration (first term on RHS of \ref{eq:RF_Efield_approx}) and a pulse compression or expansion (second term on RHS of \eqref{eq:RF_Efield_approx}) device.
-The second term then gives the desired form (see \ref{eq:linear_force}) for the force manipulating the temporal dynamics and characteristics during electron pulse compression;
+%TODO clarify this sentance and the next
+The linear term (second) then gives the desired form (see \ref{eq:linear_force}) for the force manipulating the temporal dynamics and characteristics during electron pulse compression;
 \begin{equation} \label{eq:RF_Force_z}
-  \boldsymbol{F}_{ z , RF } = - \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] z \hat{\operatorname{ \boldsymbol{z} }} \text{.}
+  \boldsymbol{F}_{ z , RF }^{linear} = - \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] z \hat{\operatorname{ \boldsymbol{z} }} \text{.}
 \end{equation}

### 2013-03-07 inc/magnetic_lenses/maglens.tex
added magnetic lens loops comparison figure and moved some of the related content from the hardware section to the extension

@@
 
-\subsection{Introduction to Magnetic Lenses}
-
-A simple magnetic lens is just a solenoid.
-The axis the lens is collinear with the axis of the microscope column.
-Initially one would think that electrons moving parallel to this axis would not be affected by the magnetic field lines in the solenoid, however the contribution of the fringing field must be taken into consideration.
-As electrons propagate through the solenoid the fringing magnetic field imparts an azimuthal velocity.
-This ``tumbling'' around the axis, along with the axial magnetic field, is what causes the inward-directed force.
-
-A full description of the motion of the electrons is easily derived \cite{el-kareh_electron_1970},
-\begin{subequations} \label{eq:lens_eq_of_motion}
-\begin{gather}
-  m \ddot{r} = -e B_z r \dot{\theta} + m r \dot{\theta}^2                 \\
-  \frac{d}{dt} ( m r^2 \dot{\theta} ) = e r \dot{r} B_z - e r B_r \dot{z} \\
-  m \ddot{z} = e B_r r \dot{\theta}
-\end{gather}
-\end{subequations}
-however, attempting to further seperate these equations quickly highlights the tightly coupled nature of the motion of the electron in this system.
-The full description of the off-axis field of a real (non-infinite) air-core solenoid is also surprisingly complicated.
-Montgomery and Terrell \cite{montgomery_some_1961} provide a full treatment for many lens configurations, including a tractable treatment for a single current loop of radius $a$ in cylindrical coordinates,
-\begin{subequations} \label{eq:field_of_loop}
-\begin{gather}
-  H_z = \frac{2I}{\sqrt{Q}} \left(   F(k) + \frac{ a^2 - r^2 - z^2 }{ (1-k^2) Q } E(k) \right) \\
-  H_r = \frac{2I}{\sqrt{Q}} \left( - F(k) + \frac{ a^2 + r^2 + z^2 }{ (1-k^2) Q } E(k) \right) \\
-  Q \equiv (a+r)^2 + z^2 \\
-  k \equiv \sqrt{ 4 a r / Q }
-\end{gather}
-\end{subequations}
-%%TODO I can't decide whether the Q and k definitions should be numbered or not
-%where
-%\begin{gather*}
-%  Q \equiv (a+r)^2 + z^2 \\
-%  k \equiv \sqrt{ 4 a r / Q }
-%\end{gather*}
-and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
-An even more accurate solenoid can be modeled by appropriately summing the fields of many such loops.
-
 \subsection{Design}
 
+Discussion of the mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens}.
 While in principle one could design a magnetic lens using \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.

### 2013-03-07 inc/extension/external_forces/mag_lens_ag.pl

@@ -1,0 +1,57 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+use v5.10;
+
+use Physics::UEMColumn alias => [qw/:standard Pulse/];
+use Physics::UEMColumn::Auxiliary ':constants';
+
+use PDL;
+
+my $a = 1/4 * 0.001 * 25.4;
+my $num = 1;
+my $lens_position = 10 * $a;
+my $strength = 2.55e-10;
+
+my $column_length = 2 * $lens_position;
+
+my $pulse = Pulse->new(
+  number => $num,
+  velocity => vc / 3,
+  sigma_t => (500**2/2) . 'um^2',
+  sigma_z => ((0.5*1e8)**2/2) . ' (ps m / s)^2',
+  eta_t => (me * 0.114 / 3) . 'kg eV',
+);
+
+my $column = Column->new(
+  length => $column_length . 'm',
+);
+
+my $sim = Physics::UEMColumn->new(
+  column => $column,
+  pulse => $pulse,
+  steps => 200,
+);
+
+my $lens1 = MagneticLens->new(
+  location => $lens_position . 'm',
+  length   => $a / 10 . 'm',
+  strength => $strength,
+);
+
+$sim->add_element($lens1);
+
+my $result = pdl $sim->propagate;
+
+my $z = $result->slice('(1),') - ( $lens_position );
+my $w = sqrt( 2 * $result->slice('(3),') );
+
+use PDL::Graphics::Prima::Simple;
+line_plot $z, $w;
+
+wcols $z, $w, "mag_lens_ag.csv", { COLSEP => ',' };
+
+my $min_ind = $w->minimum_ind;
+say "\t" . $w->at($min_ind) . 'mm at z = ' . $z->at($min_ind) . 'f';
+

### 2013-03-07 inc/model_results/lens_charge/lens_charge.pl

@@ -8,7 +8,6 @@ use Physics::UEMColumn alias => [qw/:standard Pulse/];
 use Physics::UEMColumn::Auxiliary ':constants';
 
 use PDL;
-use PDL::Graphics::Prima::Simple;
 
 sub gen_sim {
   my ($lens_position, $strength) = @_;
@@ -16,16 +15,6 @@ sub gen_sim {
 
   return sub {
     my $num = shift;
-#    my $laser = Laser->new(
-#      width    => '1 mm',
-#      duration => '1 ps',
-#      energy   => '4.75 eV',
-#    );
-
-#    my $acc = DCAccelerator->new(
-#      length  => '20 mm',
-#      voltage => '20 kilovolts',
-#    );
 
     my $pulse = Pulse->new(
       number => $num,
@@ -37,15 +26,11 @@ sub gen_sim {
 
     my $column = Column->new(
       length       => $column_length . 'mm',
-#      laser        => $laser,
-#      accelerator  => $acc,
-#      photocathode => Photocathode->new( work_function => '4.25 eV' ), # Ta
     );
 
     my $sim = Physics::UEMColumn->new(
       column => $column,
       pulse => $pulse,
-#      number => $num,
       steps => 200,
     );
 
@@ -86,6 +71,4 @@ foreach my $set ( @sets ) {
   }
 
 }
-
-#line_plot $z, $st;
 

### 2013-03-07 inc/extension/external_forces/practical.tex
more tweaks to mag lens extension

@@
     Comparison of electron rays tracked kinematically through a realistic magnetic lens simulation (green) and an AG model simulation of a Guassian electron beam (red) of a similar HW1/eM beam width $ w = 500 \mu\text{m} $.
-    The kinematic model employs ten individual current loops occupying a total length of 635$\mu$m at position $z=0$.
-    Each ray bunch consists of one ray parallel to the axis of propagation and one each having transverse velocity related to an excess photoemission energy $\Delta E \approx $ 0.1eV; the related transverse variance is used in the AG model.
-    The bunches are distributed evenly from $r=0$ to $r = 2 w$.
   }
@@
 %TODO include discussion and plot of beam size versus aberrations
+
+\ref{fig:mag_lens_loops} shows electron rays tracked kinematically (by employing \ref{eq:lens_eq_of_motion}) through a realistic magnetic lens simulation (shown in green).
+The kinematic model employs ten individual current loops generating magnetic fields described by \ref{eq:field_of_loop}; these loops occupy a total length of 635$\mu$m at position $z=0$.
+Each ray bunch consists of one ray parallel to the axis of propagation and one each having an inward or outward transverse velocity related to an excess photoemission energy $\Delta E \approx $ 0.1eV.
+The bunches are distributed evenly from $r=0$ to $r = 1$mm.
+These are compared with an AG model simulation of a Guassian electron beam of HW1/eM beam width $ w = 500 \mu\text{m} $ having an initial transverse variance of the same $\Delta E$.
+
+The AG model very accurately mimics the behavior of the kinematic model for an appropriate choice of the free parameter $M_{\smallT}^{mag}$.
+In fact, when the rays are weighted by a Gaussian of width $w$ the minimum ``radius'' of the beam is 100$\mu$m, while the AG model has a minimum HW1/eM width of 73$\mu$m.
+(The kinematic model actually overestimates the width as there are two rays at the outer convergence zone for every one which converges at the axis.) %TODO footnote?
 

### 2013-03-07 inc/extension/external_forces/mag_lens_ag.pl

@@ -53,5 +53,5 @@ line_plot $z, $w;
 wcols $z, $w, "mag_lens_ag.csv", { COLSEP => ',' };
 
 my $min_ind = $w->minimum_ind;
-say "\t" . $w->at($min_ind) . 'mm at z = ' . $z->at($min_ind) . 'f';
+say "\t" . $w->at($min_ind) . 'm at z = ' . $z->at($min_ind) . 'f';
 

### 2013-03-08 inc/extension/external_forces/practical.tex
included a summary of RF cav resultant M/F terms

@@
 
+The mathematical formalism developed in Section \ref{sec:external_forces} is only an abstract mechanism for including external forces.
+To model electron pulses traveling through realistic electron microcope columns, these abstract mechanisms must be developed further to properly describe each specific electron-optical element.
+In the following section I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression and acceleration %TODO refs
+) and DC acceleration regions (i.e. electron gun dynamics).
+
 \subsection{Region of Influence}
 
-To specify the physical region of influence of the lenses, it is convenient to redefine $M_{i}$ to include a super-Gaussian envelope in $z^{\prime}$, the position of the peak of the pulse in the lab frame;
+Of course, physical electron-optical elements occupy only a finite space in the column, and thus their influence on the pulse should be limited to a region near the element itself.
+To localize the mathematical result shown in Section \ref{sec:external_forces}, it is convenient to redefine $M_{i}$ to include a super-Gaussian envelope in $z^{\prime}$, the position of the peak of the pulse in the lab frame;
 \begin{equation} \label{eq:reg_of_influence}
@@
 Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $,\cite{kim_rf_1989} where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\ref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\ref{eq:reg_of_influence}).
-In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures, as illustrated by $1 \frac{1}{2}$-cell RF photo-gun designs.\cite{mcdonald_design_1988}
+In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures (see for example $1 \frac{1}{2}$-cell RF photo-gun designs \cite{mcdonald_design_1988}).
 For the case of an evacuated cavity, these two contributions result in a net force,
@@
 
+To summarize, the additional terms are which are added to model internal pulse dynamics arising from a single RF cavity are
+\begin{gather}
+  M_{z}^{RF} = \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \\
+  \begin{aligned}
+  M_{\smallT}^{RF} = & e E_{0} \left [ \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \right.
+\\ \times & \left.
+\sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
+   \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ]
+  \end{aligned}
+\end{gather}
+Additionally, an acceleration force term
+\begin{equation}
+  F_{z}^{RF} = E_{0} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \, \text{,}
+\end{equation}
+which originates from the leading term in \ref{eq:RF_Efield_approx}, should be added to the dynamics of the motion of the overall pulse.
 % excluded example section from JAP paper as it will need new graphics

### 2013-03-08 inc/extension/implementation/implementation.tex
tweak implementation section
included a summary of RF cav resultant M/F terms

@@
 This module depends on the much larger Perl project PDL (\url{http://pdl.perl.org}), a matrix-oriented numerical library, making it slightly more difficult to install, however the speed and implementational simplicy benefit added by that dependency is marked.
-Usefully, it also comes with a subclass of \verb!Physics::UEMColumn::Photoemission! which uses this model to generate the pulse's $z$ direction initial conditions.
+
+Of particular interest to users of \verb!Physics::UEMColumn!, \verb!Physics::Photoemission!
+also includes a subclass of \verb!Physics::UEMColumn::Photoemission!
+which uses this binning model to generate the pulse's $z$ direction initial conditions.

### 2013-03-08 inc/extension/external_forces/practical.tex
tweaks

@@
 
-To summarize, the additional terms are which are added to model internal pulse dynamics arising from a single RF cavity are
+To summarize, the additional terms are which are added to model the internal pulse dynamics arising from a single RF cavity are
 \begin{gather}
@@
 \end{gather}
-Additionally, an acceleration force term
+Additionally, since an RF cavity can act as a pulse accelerator (as seen in the leading term of \ref{eq:RF_Efield_approx}) an additional force term
 \begin{equation}
@@
 \end{equation}
-which originates from the leading term in \ref{eq:RF_Efield_approx}, should be added to the dynamics of the motion of the overall pulse.
+should be added to the equation of motion governing the propagation of the pulse down the column.
 % excluded example section from JAP paper as it will need new graphics

### 2013-03-08 inc/gun/gun_model.tex
merge remaining comments in extension section

@@
 In the acceleration region of a UEM column, a constant electric field is applied to the newly created electron pulse, increasing its velocity until it achieves its ultimate velocity just upon leaving the region at the anode aperture.
-As seen in Section \ref{sec:const_force}, this constant electric field does not affect the intrapulse dynamics, however the electric field that nears this aperture turns outward, causing the anode itself to act as a negative lens.
+As seen in Section \ref{sec:const_force}, this constant electric field does not affect the intrapulse dynamics, however the electric field that nears this aperture turns outward, causing the anode itself to act as a negative transverse lens.
 %TODO reference
@@
 \end{equation}
-As noted before, the constant leading term does not contribute to the intrapulse dynamic, however the second term is of the proper form to say that in this region
+Again, the constant leading term does not contribute to the intrapulse dynamic, however the second term is of the proper form to say that in this region the anode's longitudinal compression strength $M_z$ is given by
 \begin{equation}
-  M_z = \frac{\partial F_z}{\partial z^{\prime}}
+  M_z = \frac{\partial F_z}{\partial z^{\prime}} \, \text{.}
 \end{equation}
 
-To model the radial action (negative lensing), we can expand for electrons that are a distance $r$ radially outward from the center of the pulse,
+To model the radial action (negative transverse lensing), we can expand the radial force $F_r$ for electrons that are a distance $r$ radially outward from the center of the pulse,
 \begin{equation}
@@
 \end{equation}
-without an analytic form for $F$, we can appeal to Gauss' law (in absense of external charge)
+Without an analytic form for $F$, we can appeal to Gauss' law (in absense of external charge)
+\begin{gather}
+  \frac{\partial F_x}{\partial x^{\prime}} + \frac{\partial F_y}{\partial y^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 2 \frac{\partial F_{\smallT}}{\partial r^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 0 \\
+  \Rightarrow \frac{\partial F_{\smallT}}{\partial r^{\prime}} = - \frac{1}{2} \frac{\partial F_z}{\partial z^{\prime}}
+\end{gather} 
+to see that the transverse lens traverse lens strength $M_{\smallT}$ from the anode is simply,
 \begin{equation}
-  \frac{\partial F_x}{\partial x^{\prime}} + \frac{\partial F_y}{\partial y^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 2 \frac{\partial F_{\smallT}}{\partial r^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 0
-\end{equation}
-to see that
-\begin{equation}
-  M_{\smallT} = -\frac{1}{2}\frac{\partial F_z}{\partial z^{\prime}}
+  M_{\smallT} = -\frac{1}{2}\frac{\partial F_z}{\partial z^{\prime}} \, \text{.}
 \end{equation}
@@
 We may choose to assume that the on-axis field is defined as a smoothed step down function
-\begin{equation}
+\begin{equation} \label{eq:anode_tanh}
   F_z(0,z^{\prime}) = \frac{qV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right)
@@
 Indeed, this form is analgous to the super-Gaussian envalope (\ref{eq:reg_of_influence}), however it is open ended on the cathode side, which is necessary for the electrons to be accelerated away from the photocathode.
-Given this form (demonstrated in Section \ref{sec:gun_design}) the lensing constants are then
+Given this form (demonstrated in Section \ref{sec:gun_design}) the ``lensing'' strengths are then
 \begin{gather}
   M_{\smallT} = -\frac{qVs}{4z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \\
-  M_z = \frac{qVs}{2z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right)
+  M_z = \frac{qVs}{2z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \, \text{.}
 \end{gather}
+As with the RF cavity, remember to include the on-axis force (\ref{eq:anode_tanh}) in the overall pulse propagation.
 

### 2013-03-11 inc/model_results/shapes/shapes.tex
some initial updates on model results

@@
 For simplicity, and unlike most simulations in this paper, the pulses in this simulation are not calculated from laser and photocathode parameters, but are rather defined directly.
+%TODO should I remove this? Should I just specify the creation types on each instance?
 The initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ while in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
-Since the initial pulse volumn is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} ($ N = 10^{6} $ electrons in the bunch) than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} ($ N = 10^{0} $).
+Since the initial pulse volume is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} ($ N = 10^{6} $ electrons in the bunch) than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} ($ N = 10^{0} $).
 The pulse velocity is $ v_{\smallzero} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.
@@
     Transverse (solid) and longitudinal (dashed) dynamics of three example pulses for three length/width ratios ($\xi$). 
-    The red lines represent a prolate pulse (a ``pancake'') with $\xi=1/3$, green lines for a spherical pulse with $\xi=1$, and blue for oblate (a ``cigar'') having $\xi=3$.
+    The red lines represent a prolate pulse (a ``pancake'') with $\xi=1/3$, green lines for a spherical pulse with $\xi=1$, and blue for an oblate pulse (a ``cigar'') having $\xi=3$.
+    The pulses shown in \ref{fig:compare_shape_N1} all have total charge $N=1$, while in \ref{fig:compare_shape_N1e6} the pulses have $N=10^6$.
   }
@@
 The bunch propagation dynamics for negligible pulse charge density ($ N = 10^{0} $) shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
-In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands roughly twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp ($ \gamma_{\alpha} = 0 $).
+In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands approximately twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp ($ \gamma_{\alpha} = 0 $).
 Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both disk- and cigar-like pulses are initially driven towards the spherical regime.
@@
 For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects can dominate the propagation dynamics.
-In \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, we display the predicted temporal dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
+\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted temporal dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
 The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
@@
 Even for nanosecond DTEM (the extreme cigar-like pulse ($\xi \gg 1 $)), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
-In this paper, however, we will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the disk-like regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}
+In this thesis, however, I will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the disk-like regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}
 
+%TODO perhaps all of this should be moved back or even cut?
 I have shown that a prolate spheroid pulse shape is preferred for UEM, both for generation in SECTION %TODO link to Child law discussion

### 2013-03-11 main.tex
consistent use of v_max vs v_0
added magnetic lens loops comparison figure and moved some of the related content from the hardware section to the extension
some initial updates on model results

@@
 \providecommand{\smallD}{\ensuremath{{\scriptscriptstyle D}}}
+\providecommand{\smallzero}{\ensuremath{{\scriptscriptstyle 0}}}
 \providecommand{\smallF}[0]{ { \scriptscriptstyle F } }
@@
 
-  \subimport{inc/extension/external_forces/}{external_forces}
+  \subimport{inc/extension/external_forces/}{formal}
+
+  \subimport{inc/extension/external_forces/}{practical}
 
@@
 \chapter{Model Results, Predictions and Implications}
+
+In Chapters \ref{chap:previous_models} and \ref{chap:extension}, I have introduced the extended Analytic Gaussian model.
+In the following chapter I use this model to make general observations about pulse propagation for Ultrafast Electron Microscopy.
 

### 2013-03-12 inc/model_results/shapes/shapes.tex
minor changes to shapes section

@@
 This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
-In fact, both the $ N = 10^{6} $ disk- and cigar-like pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
+%TODO the following sentance appears not to be true, still perhaps should include something like it
+%In fact, both the $ N = 10^{6} $ disk- and cigar-like pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
 The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.

### 2013-03-13 inc/model_results/spacecharge/spacecharge.pl
update spacecharge simulation to use accelerator or not depending on flag.
This should lead to using both in that section of the paper

@@ -3,34 +3,53 @@
 use strict;
 use warnings;
 
-use Physics::UEMColumn alias => ':standard';
-use Physics::UEMColumn::Auxiliary ':materials';
+use Getopt::Long;
+GetOptions( accelerator => \my $use_acc );
+
+use Physics::UEMColumn alias => [qw/:standard Pulse/];
+use Physics::UEMColumn::Auxiliary qw/:materials :constants/;
 
 sub for_number {
-  my $num = shift;
+  my ($num, $use_acc) = @_;
 
-  my $laser = Laser->new(
-    width    => '1 mm',
-    duration => '1 ps',
-    energy   => '4.75 eV',
-  );
+  my $sim;
+  if ( $use_acc ) {
+    my $laser = Laser->new(
+      width    => '1 mm',
+      duration => '1 ps',
+      energy   => '4.75 eV',
+    );
 
-  my $acc = DCAccelerator->new(
-    length  => '20 mm',
-    voltage => '20 kilovolts',
-  );
+    my $acc = DCAccelerator->new(
+      length  => '20 mm',
+      voltage => '30 kilovolts',
+    );
 
-  my $column = Column->new(
-    length       => '15 cm', 
-    laser        => $laser,
-    accelerator  => $acc,
-    photocathode => Photocathode->new(Ta),
-  );
+    my $column = Column->new(
+      length       => '15 cm', 
+      laser        => $laser,
+      accelerator  => $acc,
+      photocathode => Photocathode->new(Ta),
+    );
 
-  my $sim = Physics::UEMColumn->new(
-    column      => $column,
-    number      => $num,
-  );
+    $sim = Physics::UEMColumn->new(
+      column => $column,
+      number => $num,
+    );
+
+  } else {
+
+    $sim = Physics::UEMColumn->new(
+      column => Column->new( length => '15 cm' ),
+      pulse  => Pulse->new(
+        number => $num,
+        initial_width => '1 mm',
+        initial_length => ( vc / 3 * 1e-12 ) . ' m',
+        velocity => vc / 3,
+      ),
+    );
+
+  }
 
   my $result = $sim->propagate;
   return $result;
@@ -42,7 +61,7 @@ my @nums = map { ("1e$_", "2.2e$_", "4.6e$_") } (0..8);
 pop @nums for 1..2;
 
 for my $num ( @nums ) {
-  my $result = for_number($num);
[…]

### 2013-03-14 inc/extension/external_forces/formal.tex
included a summary of RF cav resultant M/F terms
merging comments from andreas on back half of extension chapter

@@
 
-While the free-space AG model does not contain any external forces, it does include an internal force.
-The Coulomb force is included to model the repulsion of the pulse's electrons from each other.
-Because this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identically ``Gaussian'' optics.\cite{michalik_analytic_2006}
+While the free-space AG model does not contain any external forces, it does include an internal force: the Coulomb force is included to model the repulsion of the pulse's electrons from each other.
+As this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics.\cite{michalik_analytic_2006}
 
 The internal Coulombic repulsion force enters the model in \ref{eq:potential_integral}.
-Since both potentials and integrals are linear, one may add additional potentials so that it becomes
+Since both potentials and integrals are linear, one may add additional potentials, so that it becomes
 \begin{equation}
@@
 \end{equation}
-In doing so, we rename the old potential, the internal repulsive Coulomb force, $V^{int}$.
+In doing so, we rename the potential due to the internal repulsive Coulomb force $V^{int}$.
 Following these new terms through the derivation, we see that
@@
 
-This new form of $\Phi(\vec{r};t)$ can be substituted into the definition of $K^{force}$ matrix (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector} 
+This new form of $\Phi(\vec{r};t)$ can be substituted into the definition of $K^{force}$ matrix (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector}, to give
 \begin{equation}
@@
 
-Noting that the spatial derivative of a potential is just the force, we can now define the matrix contributed by each external force is given by
+Noting that the spatial derivative of a potential is just the force, we can now define the matrix contributed by each external force as given by
 \begin{equation} \label{eq:Kint}
-  K^{ext}_{ij} = \frac{1}{N} \iint u_i u_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
+  K^{ext}_{ij} = \frac{1}{N} \iint u_i u_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p \,\text{.}
 \end{equation}
@@
 \begin{equation}
-  \frac{\partial}{\partial t} A^{-1}_{ij} = K^{flow}_{ij} + K^{int}_{ij} + \sum K^{ext}_{ij}
+  \frac{\partial}{\partial t} A^{-1}_{ij} = K^{flow}_{ij} + K^{int}_{ij} + \sum K^{ext}_{ij} \,\text{.}
 \end{equation}
@@
 For constant static forces having no dependence in $ x $, $ y $, $ z $, or $ t $, one can readily show that there is no effect on the pulse propagation parameters $ \sigma_{ \alpha } $, $ \eta_{ \alpha } $, and $ \gamma_{ \alpha } $.
-For example, consider a field consider a field $ \vec{E} = -E \hat{z} $ such that electrons are accelerated in the positive $ z $ direction by $ \vec{F}_{ext} = qE \hat{z} = q ( V / d ) \hat{z} $, so that from \ref{eq:Kint} we have
+For example, consider an acceleration electric field $ \vec{E} = -E \hat{z} $ such that electrons are accelerated in the positive $ z $ direction.
+%TODO e = |q| ?
+The force is then $ \vec{F}_{ext} = qE \hat{z} $, so that from \ref{eq:Kint} we have
 \begin{equation}
- \begin{split} 
-  K^{ext}_{ij} = \frac{1} {N} \dfrac{qV} {d} \int{ u_{i} u_{j} \left( \frac{p_{z}} {\eta_{z}} - \frac{\gamma_{z} z} {\sigma_{z} \eta_{z}} \right) f d^{3}x   d^{3}p } ;
- \end{split}
+  K^{ext}_{ij} = \frac{qE} {N} \int{ u_{i} u_{j} \left( \frac{p_{z}} {\eta_{z}} - \frac{\gamma_{z} z} {\sigma_{z} \eta_{z}} \right) f d^{3}x d^{3}p } \,\text{.}
 \end{equation}
-but it is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ (or equivalently $ K^{ext} = \hat{0} $) meaning that this static field has no effect on the parameters of the pulse.
+It is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ (or equivalently $ K^{ext} = \hat{0} $) meaning that this static field has no effect on the parameters of the pulse.
 Of course, this is only true in the pulse frame, the frame of the simulation, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m_{e} $ is unchanged.
@@
 
-For the case of a perfect lens, or in the longitudinal direction a perfect RF compression element, with a parabolic potential, the external force is linear with distance from the center of the electron pulse:
+For the case of a cylindrically symmetric perfect lens, or in the longitudinal direction a perfect RF compression element, with a parabolic potential, the external force is linear with distance from the center of the electron pulse:
 \begin{equation} \label{eq:linear_force}
@@
 \begin{equation}
-  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2(0) + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
+  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
     + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum M_{\alpha} 

### 2013-03-14 inc/extension/external_forces/formal.tex
merged all comments on formal extension

@@
 which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.
-This derivation has been for a single force, in practice however there may be multiple forces, therefore \ref{eq:ag_big_gamma_gamma} may be restated as
+This derivation has been for a single force, in practice however there may be multiple forces.
+Therefore if each force is labeled by $f$, \ref{eq:ag_big_gamma_gamma} may be restated as
 \begin{equation}
   \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
-    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum M_{\alpha} 
+    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} 
 \end{equation}
 
-Since these terms are only added to $\dfrac{d \gamma_{\alpha}}{d t}$, the analysis of Liouville's Theorem, shown in Section \ref{sec:liouville}, is unaffected.
-The extended AG model still explicitly satisfies Liouvilles's Theorem.
+Since these terms are only added to $\dfrac{d \gamma_{\alpha}}{d t}$, the analysis of Liouville's Theorem, shown in Section \ref{sec:liouville}, is unaffected: the \textit{extended} AG model still explicitly satisfies Liouvilles's Theorem.
 
@@
 As a result, we emphasize that the presented AG model extension is valid only for the action of linear forces (i.e., forces that have a linear dependence on distance in the pulse frame) on the electron pulse due to the model's self-similar Gaussian approximation.
-The inclusion of the action of nonlinear forces (e.g., due to space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations, which should be employed to gauge the accuracy (and hence limits) of our extended AG approach at some point.
+The inclusion of the action of nonlinear forces (e.g., due to higher-order space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations.%, which should be employed to gauge the accuracy (and hence limits) of our extended AG approach at some point.
+%TODO consider adding a replacement sentance like: "It would be interesting to employ ..." or "A future goal is to ..."
 

### 2013-03-14 inc/extension/external_forces/formal.tex
include most of the comments through RF cavity practical extension

@@
 \end{equation}
-which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.
+which by \ref{eq:dainvdt} this yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.
 This derivation has been for a single force, in practice however there may be multiple forces.

### 2013-03-14 inc/extension/external_forces/practical.tex
fix bug
include most of the comments through RF cavity practical extension

@@
 To model electron pulses traveling through realistic electron microcope columns, these abstract mechanisms must be developed further to properly describe each specific electron-optical element.
-In the following section I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression and acceleration %TODO refs
+In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression and acceleration %TODO refs
 ) and DC acceleration regions (i.e. electron gun dynamics).
@@
 
+%TODO perhaps this shouldn't be a redefinition, but some other function
 Of course, physical electron-optical elements occupy only a finite space in the column, and thus their influence on the pulse should be limited to a region near the element itself.
-To localize the mathematical result shown in Section \ref{sec:external_forces}, it is convenient to redefine $M_{i}$ to include a super-Gaussian envelope in $z^{\prime}$, the position of the peak of the pulse in the lab frame;
+To localize the mathematical result shown in Section \ref{sec:external_forces}, it is convenient to redefine a particular $M_{\alpha}$ to include a super-Gaussian envelope in $z^{\prime}$, the position of the peak of the pulse in the lab frame;
 \begin{equation} \label{eq:reg_of_influence}
-  M_{i}	\to M_{i} \exp \left [ - \left (  \frac{ z^{\prime} - z_{lens}^{\prime} }{ L_{lens} / 2 } \right )^{ 2 n } \right ] \text{ ,}
+  M_{\alpha} \to M_{\alpha} \exp \left [ - \left (  \frac{ z^{\prime} - z_{lens}^{\prime} }{ L_{lens} / 2 } \right )^{ 2 n } \right ] \text{ ,}
 \end{equation}
-where $z_{lens}^{\prime}$ is the position of the center of the lens, $L_{lens}$ is the length of the lens and $n$ is the super-Gaussian order parameter.
+where $z_{lens}^{\prime}$ is the position of the center of the lens, $L_{lens}$ is the axial length of the lens and $n$ is the super-Gaussian order parameter.
 The order parameter defines the ``sharpness'' of the edge of the region of influence with integer values ranging from 1 (a simple Gaussian) through $\infty$ (a top-hat).
@@
 
-A full description of the motion of the electrons is easily derived \cite{el-kareh_electron_1970},
+The full description of the motion of an electron (in cylindrical coordinates $r, \theta, z$) in an azimuthally symmetric magnetic field ($\vec{B} = B_r \hat{r} + B_z \hat{z}$) is easily derived \cite{el-kareh_electron_1970},
 \begin{subequations} \label{eq:lens_eq_of_motion}
 \begin{gather}
-  m \ddot{r} = -e B_z r \dot{\theta} + m r \dot{\theta}^2                 \\
-  \frac{d}{dt} ( m r^2 \dot{\theta} ) = e r \dot{r} B_z - e r B_r \dot{z} \\
-  m \ddot{z} = e B_r r \dot{\theta}
+  m \ddot{r} = -q B_z r \dot{\theta} + m r \dot{\theta}^2                 \\
+  \frac{d}{dt} ( m r^2 \dot{\theta} ) = q r \dot{r} B_z - q r B_r \dot{z} \\
+  m \ddot{z} = q B_r r \dot{\theta}
 \end{gather}
 \end{subequations}
-however, attempting to further seperate these equations quickly highlights the tightly coupled nature of the motion of the electron in this system.
-The full description of the off-axis field of a real (non-infinite) air-core solenoid is also surprisingly complicated.
-Montgomery and Terrell \cite{montgomery_some_1961} provide a full treatment for many lens configurations, including a tractable treatment for a single current loop of radius $a$ in cylindrical coordinates,
+where each dot denotes a time derivative.
+Attempting to further seperate these equations quickly highlights the tightly coupled nature of the motion of the electron in this system.
+Compounding the problem, the full description of the off-axis field of a real (non-infinite) air-core solenoid is also surprisingly complicated.
+Montgomery and Terrell \cite{montgomery_some_1961} provide a full treatment for many lens configurations, including a tractable treatment for a single current loop of radius $a$, again in cylindrical coordinates,
 \begin{subequations} \label{eq:field_of_loop}
 \begin{gather}
-  H_z = \frac{2I}{\sqrt{Q}} \left(   F(k) + \frac{ a^2 - r^2 - z^2 }{ (1-k^2) Q } E(k) \right) \\
-  H_r = \frac{2I}{\sqrt{Q}} \left( - F(k) + \frac{ a^2 + r^2 + z^2 }{ (1-k^2) Q } E(k) \right) \\
+  \frac{B_z}{\mu} = H_z = \frac{2I}{\sqrt{Q}} \left(   F(k) + \frac{ a^2 - r^2 - z^2 }{ (1-k^2) Q } E(k) \right) \\
+  \frac{B_r}{\mu} = H_r = \frac{2I}{\sqrt{Q}} \left( - F(k) + \frac{ a^2 + r^2 + z^2 }{ (1-k^2) Q } E(k) \right) \\
   Q \equiv (a+r)^2 + z^2 \\
@@
 %\end{gather*}
-and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
+where $\mu$ is the local magnetic permability, $H$ is the auxiliary field, and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
+In the vacuum, $\mu$ is simply the usual free space permeability $\mu_0$; in Section \ref{sec:mag_lens}, we will see how this permeability can be used to improve the performance of magnetic lenses.
 An even more accurate solenoid can be modeled by appropriately summing the fields of many such loops.
@@
 
-For localizing the region of influence of the lens, I have commonly modelled a thin magnetic lens as having a super-Gaussian order value $n=1$, a simple Gaussian; of course this may depend on the specifics of the lens.
-To simulate the effect of magnetic lenses in a UEM by the extended AG model, for each magnetic lens, add a term like
-\begin{equation} \label{eq:reg_of_influence}
+For localizing the region of influence of the lens, I have generally modelled a thin magnetic lens as having a super-Gaussian order value $n=1$, a simple Gaussian; of course this may depend on the specifics of the lens.
+To simulate the effect of magnetic lenses in a UEM by the extended AG model, for each magnetic lens, one adds a term like
+\begin{equation}
   M_{\smallT}^{mag} = M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \text{ ,}
@@
 The kinematic model employs ten individual current loops generating magnetic fields described by \ref{eq:field_of_loop}; these loops occupy a total length of 635$\mu$m at position $z=0$.
-Each ray bunch consists of one ray parallel to the axis of propagation and one each having an inward or outward transverse velocity related to an excess photoemission energy $\Delta E \approx $ 0.1eV.
+These loops each carry a current of 45A; this approximates a more realistic stack of 10 loops carrying 4.5A, thus simulating a  100 turn solenoid in a 10x10 array.
+Each ray bunch consists of one ray parallel to the axis of propagation and one each having a longitudinal velocity of $1 \times 10^{8}$ m/s and an inward or outward transverse velocity related to an excess photoemission energy $\Delta E \approx $ 0.1eV.
 The bunches are distributed evenly from $r=0$ to $r = 1$mm.
-These are compared with an AG model simulation of a Guassian electron beam of HW1/eM beam width $ w = 500 \mu\text{m} $ having an initial transverse variance of the same $\Delta E$.
+These are compared with an AG model simulation of a Guassian electron beam of HW1/eM beam width $ w = 500 \mu\text{m} $ having an initial transverse momentum variance associated with $\Delta E$.
 
@@
 \begin{equation}
-  \Omega = \frac{ c k }{a} \text{,}
+  \Omega = 2.405 \frac{ c }{a} \text{,}
 \end{equation}
-where $c$ is the speed of light in a vacuum and $ k = 2.405 $ is the first zero of the $\operatorname{ J_{0} }$ Bessel function.
+where $c$ is the speed of light in a vacuum and 2.405 is the first zero of the $\operatorname{ J_{0} }$ Bessel function.
 
@@
 In obtaining \ref{eq:RF_Force_r}, which is also of the form for use in \ref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \ref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
-For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $,\cite{kim_rf_1989} even though its magnitude is maximized for $\phi = 0$.
+%TODO examples shown below?
+For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $,\cite{kim_rf_1989} even though its magnitude is maximized for $\phi = 0$.
 Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.
 
-To summarize, the additional terms are which are added to model the internal pulse dynamics arising from a single RF cavity are
+To summarize, the additional terms are which are added to model the internal pulse dynamics arising from a single TM$_{010}$-mode cylindrical RF cavity are
 \begin{gather}
@@
 \begin{equation}
-  F_{z}^{RF} = E_{0} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \, \text{,}
+  F_{z}^{RF} = e E_{0} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \, \text{,}
 \end{equation}

### 2013-03-14 inc/gun/gun_model.tex
finished including andreas comments on extension chapter

@@
 In the acceleration region of a UEM column, a constant electric field is applied to the newly created electron pulse, increasing its velocity until it achieves its ultimate velocity just upon leaving the region at the anode aperture.
-As seen in Section \ref{sec:const_force}, this constant electric field does not affect the intrapulse dynamics, however the electric field that nears this aperture turns outward, causing the anode itself to act as a negative transverse lens.
+As seen in Section \ref{sec:const_force}, this constant electric field does not affect the intrapulse dynamics, however the distortion in the electric field near this aperture causes the anode itself to act as a negative transverse lens.
 %TODO reference
 Additionally, the longitudinal electric field stregth sharply decreases near the anode.
-Because the front of the pulse sees this change before the back, this changing field acts as a pulse compression element.
+As the front of the pulse sees this change before the back, this changing field also acts as a pulse compression element.
 
@@
 \begin{equation}
-  \vec{F}(r^{\prime}=0,z^{\prime}) = F_z (0,z^{\prime}) \hat{z}
+  \vec{F}(r^{\prime}=0,z^{\prime}) = F_z (0,z^{\prime}) \hat{z} = e E_z (0,z^{\prime}) \hat{z}
 \end{equation}
@@
 
-To first model the longitudinal (compression) action at the anode %TODO ref
-on electrons that are a distance $z$ ahead or behind of the center of the pulse feel a slightly different field, 
+To first model the longitudinal (compression) action at the anode, %TODO ref
+consider electrons that are a distance $z$ ahead or behind of the center of the pulse and so feel a slightly different force, 
 \begin{equation}
@@
 \begin{equation}
-  M_z = \frac{\partial F_z}{\partial z^{\prime}} \, \text{.}
+  M_z = \frac{\partial F_z}{\partial z^{\prime}} = e \frac{\partial E_z}{\partial z^{\prime}} \, \text{.}
 \end{equation}
@@
 \begin{equation} \label{eq:anode_tanh}
-  F_z(0,z^{\prime}) = \frac{qV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right)
+  F_z(0,z^{\prime}) = \frac{eV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right)
 \end{equation}
@@
 \begin{gather}
-  M_{\smallT} = -\frac{qVs}{4z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \\
-  M_z = \frac{qVs}{2z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \, \text{.}
+  M_{\smallT} = -\frac{eVs}{4z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \\
+  M_z = \frac{eVs}{2z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \, \text{.}
 \end{gather}
-As with the RF cavity, remember to include the on-axis force (\ref{eq:anode_tanh}) in the overall pulse propagation.
+As with the RF cavity, one needs to include the on-axis force (\ref{eq:anode_tanh}) in the overall pulse propagation kinematics.
 
@@
   }
-  \caption[An example of a tahn form acceleration region for three different ``sharpness'' values]{
+  \caption[An example of a tanh form acceleration region for three different ``sharpness'' values]{
     An example of a tahn form acceleration region (overall length 6.35mm) for three different ``sharpness'' values.
     The square data points in \protect\subref{fig:field_on_axis} represent the on-axis electric field data from FEM simulations, see Section \ref{sec:gun_design}.
-    The colored electric fields in \protect\subref{fig:field_on_axis} correspond to the pulse width (solid) and length (dashed) in \protect\subref{fig:tanh_sharpness}.
+    The colored electric fields in figure \protect\subref{fig:field_on_axis} correspond to the pulse width (solid) and length (dashed) in figure \protect\subref{fig:tanh_sharpness}.
   }
@@
 
-\ref{fig:gun-sharpness} shows an example acceleration region for three values of the ``sharpness'' $s$.
+\ref{fig:gun-sharpness} shows an example acceleration region for three values of the ``sharpness'' $s$ as well as longitudinal FEM field data which is explained further in Section \ref{sec:gun_design}.
+%TODO possibly remove the FEM data? it doesn't help much here and is confusing since it hasn't been introduced
 What is clear from this figure is that the dynamics are not highly coupled to the value of this hard-to-quantify parameter.
-For all values of $s$ the divergence of the beam is equal (thought the size may vary slightly).
+For all values of $s$ the divergence of the beam is equal (though the size may vary slightly).
 Interestingly the length of the pulse is stabilized on exiting the anode, coming very nearly to a constant length (not expanding nor contracting).

### 2013-03-15 inc/model_results/spacecharge/spacecharge.tex
expanded the spacecharge propagation section in model_results.
Added more realistic section which uses an accelerator too, though perhaps this should be moved to the model results with extension section, or maybe I should drop the distinction. hmmmm.

@@
 Much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements.
-
-Purely for illustrative purposes I am including \ref{fig:spacecharge} to provide an intuition about the order of magnitude of these effects.
+Simplistically stated, at low pulse charge densities (low number of electrons/pulse), the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
+On the other hand, space-charge effects are expected to dominate at large pulse charge densities.
 
@@
   \begin{tikzpicture}
-    \input{spacecharge_data}
+    \input{spacecharge_data_noacc}
   \end{tikzpicture}
-  \caption[Free-space pulse evolution vs charge density]{
-    The final HW1/eM width (solid) and length (dashed) of similar initial size pulses for different number of electrons in the pulse.
-    The pulse is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy. The pulse then traverses a 20mm acceleration region and a 15cm drift region.
+  \caption[Simplistic free-space pulse evolution vs charge density]{
+    The final HW1/eM width (solid) and length (dashed) of pulses which are intially similar except for the number of electrons in the pulse.
+    The pulse is initially 1 mm (HW1/eM) width and 1 ps in duration travelling at $c/3$ with 0.5eV excess photoemission energy.
+    It then traverses a 15cm drift region upon which the final length and width are recorded.
   }
-  \label{fig:spacecharge}
+  \label{fig:spacecharge_noacc}
 \end{figure}
+
+This trend is clearly evident in \ref{fig:spacecharge_noacc} which presents the predicted HW1/eM pulse width $\sqrt{2 \sigma_{\smallT}}$ (solid line) and longitudinal length $\sqrt{2 \sigma_{z}}$ (dashed line) as a function of the number of electrons in the pulse after a propagation distance of 15cm.
+The pulses are not accelerated but rather start with a velocity of $c/3$ (where $c$ is the speed of light in a vacuum), initial width of 1 mm, HW1/eM pulse duration 1 ps, have the initial momentum chirps $\gamma_{\smallT} = \gamma{z} = 0 $ and momentum uncertainties related to an excess photoemission energy of $\Delta E = $ 0.5 eV.
+These naive conditions are needed to isolate the behavior of a propagating pulse from its creation environment.
+The simulations, clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
+In the absence of space-charge effects, the presented prolate pulse shaped pulses should be maintained as such during unforced propagation because the electron pulse divergence and its temporal broadening rate are both proportional to $\Delta E^{1/2}$.
+
+%TODO highlight the following in the experimental section
+%In other words, the spatial and temporal dispersive effects due to the nonzero value of $\Delta E$ are expected to dominate the space-time dynamics of the electron pulse during its propagation
+%in our investigations because less than 10 5 electrons/pulse
+%are generated.
+
+As the number of electrons/pulse increases beyond $10^6$, both the pulse length and diameter rapidly increase.
+The model predicts that space-charge effects produce a spherical charge distribution of approximately five times larger width at $10^8$ electrons/pulse for these conditions.
+As noted in Section \ref{sec:initial_shapes}, pulses under the influence of the internal space-charge field expand preferentially in the dimension that was smallest initially; this causes the pulse to initially trend towards being spherical.
+
+\ref{fig:spacecharge_noacc} is specifically created to isolate the behavior of and ``ideal'' pulse which is independent of its creation environment.
+Of course in a more realistic system, the electron pulse is generated from a photocathode and then accelerated.
+These forces play an interesting role in the system.
+
+Consider now a simulation comparable to the one presented above, but which uses the initial conditions presented in \ref{eq:summary} and accelerated by an electron gun as modelled in Section \ref{sec:gun_model} which is 20mm in length.
+The results of this simulation are presented in \ref{fig:spacecharge_acc}.
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \input{spacecharge_data_acc}
+  \end{tikzpicture}
+  \caption[More realistic free-space pulse evolution vs charge density]{
+    The final HW1/eM width (solid) and length (dashed) of pulses which are intially similar except for the number of electrons in the pulse.
+    The pulse is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy.
+    It then traverses a 15cm column, a 20mm acceleration region followed by drift region.
+  }
+  \label{fig:spacecharge_acc}
+\end{figure}

### 2013-03-15 inc/model_results/spacecharge/spacecharge.pl

@@ -55,7 +55,8 @@ sub for_number {
   return $result;
 }
 
-open my $fh, '>', 'spacecharge.dat';
+my $outfilename = $use_acc ? 'spacecharge_acc.dat' : 'spacecharge_noacc.dat';
+open my $fh, '>', $outfilename or die "Cannot open $outfilename for writing";
 
 my @nums = map { ("1e$_", "2.2e$_", "4.6e$_") } (0..8);
 pop @nums for 1..2;

### 2013-03-16 inc/model_results/shapes/shapes.pl
use new Physics::UEMColumn::Pulse defaults to simplify shapes script

@@ -9,10 +9,10 @@ use Physics::UEMColumn::Auxiliary ':constants';
 my $n = shift or die "Please specify a number of electrons";
 
 my $length = 0.15; # 15 cm
-my $eta_t = me * qe * (0.5) / 3; # 0.5 eV excess energy
+my $dE = '0.5 eV';
 my $vol = (1e-3)**3;
 
-my $for_xi = gen_xi_series( $n, $vol, $eta_t, $length );
+my $for_xi = gen_xi_series( $n, $vol, $dE, $length );
 
 for my $spec ( [1/3, 'prolate' ], [1.0001, 'sphere'], [3, 'oblate'] ) {
   my ( $xi, $shape ) = @$spec;
@@ -28,7 +28,7 @@ for my $spec ( [1/3, 'prolate' ], [1.0001, 'sphere'], [3, 'oblate'] ) {
 }
 
 sub gen_xi_series {
-  my ($n, $vol, $eta_t, $length) = @_; 
+  my ($n, $vol, $dE, $length) = @_; 
 
   return sub {
     my $xi = shift;
@@ -39,12 +39,9 @@ sub gen_xi_series {
     my $pulse = Pulse->new(
       velocity => '1e8 m/s',
       number   => $n,
-      gamma_t  => 0,
-      gamma_z  => 0,
-      eta_t    => $eta_t,
-      eta_z    => $eta_t / 4,
       sigma_t  => $wt**2 / 2,
       sigma_z  => $wz**2 / 2,
+      excess_photoemission_energy => $dE,
     );
 
     my $column = Column->new(

### 2013-03-16 inc/model_results/shapes/shapes.pl
more improvements to shapes.pl

@@ -14,8 +14,8 @@ my $vol = (1e-3)**3;
 
 my $for_xi = gen_xi_series( $n, $vol, $dE, $length );
 
-for my $spec ( [1/3, 'prolate' ], [1.0001, 'sphere'], [3, 'oblate'] ) {
-  my ( $xi, $shape ) = @$spec;
+for my $spec ( [ prolate => 1/3 ], [ sphere => 1.0001 ], [ oblate => 3 ] ) {
+  my ( $shape, $xi ) = @$spec;
   my $file = "shape_${shape}_N$n.dat";
   open my $fh, '>', $file or die "Cannot open $file: $!\n";
 

### 2013-03-16 inc/model_results/lens_charge/lens_charge.pl
update lens_charge.pl to use new ::Pulse defaults and add $is_prolate flag to anticipate oblate output

@@ -10,7 +10,7 @@ use Physics::UEMColumn::Auxiliary ':constants';
 use PDL;
 
 sub gen_sim {
-  my ($lens_position, $strength) = @_;
+  my ($lens_position, $strength, $is_prolate) = @_;
   my $column_length = 4 * $lens_position;
 
   return sub {
@@ -19,13 +19,13 @@ sub gen_sim {
     my $pulse = Pulse->new(
       number => $num,
       velocity => vc / 3,
-      sigma_t => (500**2/2) . 'um^2',
-      sigma_z => ((0.5*1e8)**2/2) . ' (ps m / s)^2',
-      eta_t => (me * 0.5 / 3) . 'kg eV',
+      initial_width  => $is_prolate ? '500 um' : '50 um',
+      initial_length => $is_prolate ? '50 um'  : '500 um',
+      excess_photoemission_energy => '0.5 eV',
     );
 
     my $column = Column->new(
-      length       => $column_length . 'mm',
+      length => $column_length . 'mm',
     );
 
     my $sim = Physics::UEMColumn->new(
@@ -47,8 +47,8 @@ sub gen_sim {
 }
 
 my @sets = (
-  [ 6,  730e-12 ],
-  [ 60, 675e-13 ],
+  [ 6,  730e-12, 1 ],
+  [ 60, 675e-13, 1 ],
 );
 
 foreach my $set ( @sets ) {

### 2013-03-18 inc/model_results/spacecharge/spacecharge.tex
improve discussion of realistic spacecharge plot

@@
 
-Consider now a simulation comparable to the one presented above, but which uses the initial conditions presented in \ref{eq:summary} and accelerated by an electron gun as modelled in Section \ref{sec:gun_model} which is 20mm in length.
-The results of this simulation are presented in \ref{fig:spacecharge_acc}.
 \begin{figure}
@@
 \end{figure}
+
+Consider now a simulation comparable to the one presented above, but which uses the initial conditions presented in \ref{eq:summary} and accelerated by an electron gun as modelled in Section \ref{sec:gun_model} which is 20mm in length.
+The results of this simulation are presented in \ref{fig:spacecharge_acc}.
+It is immediately obvious that the final transverse width is larger than in the ideal case presented in \ref{fig:spacecharge_noacc}, even in the low charge-density regime; this is primarily due to the additional divergence imparted by the negative lensing at the anode (see Section \ref{sec:gun_model}).
+One can see that though the pulse size does increase beyond a certain total pulse charge, and while that trend appears to start at a higher charge-density, due to the complexity of the pulse dynamics it is rather difficult to compare the exact charge-density from this plot.
+What can be seen is that the transverse size increases more sharply with pulse charge at higher levels.
+Since the pulse is already diverging in the transverse direction, this is compounded by the increased pulse charge.
+This divergence then lowers the overall charge-density more quickly and thus the longitudinal width is not affected as strongly as was seen in \ref{fig:spacecharge_noacc}.

### 2013-03-18 inc/model_results/spacecharge/spacecharge.tex
tweaks to spacecharge result text

@@
 Since the pulse is already diverging in the transverse direction, this is compounded by the increased pulse charge.
-This divergence then lowers the overall charge-density more quickly and thus the longitudinal width is not affected as strongly as was seen in \ref{fig:spacecharge_noacc}.
+This divergence then acts to lower the overall charge-density more quickly, thus the longitudinal pulse length is not affected as strongly as was seen in \ref{fig:spacecharge_noacc}.

### 2013-03-18 inc/laser/laser.tex
initial clean-up of laser section, still need some paragraphs about connection to UEM

@@
 Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH \cite{website_limo}).
-The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm fullwidth half-maximum (FWHM) 980nm absorption line of Yb:KGW \cite{Paunescu_diode_2004,Biswal_thermo_optical_05,website_EKSPLA}.
+The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm full-width half-maximum (FWHM) 980nm absorption line of Yb:KGW \cite{Paunescu_diode_2004,Biswal_thermo_optical_05,website_EKSPLA}.
 Moreover, the two dual-axis collimated laser diodes have full-angle $1/e^2$ intensity beam divergences of $\theta_y \approx$ 4.4mrad and $\theta_x \approx$ 6.8mrad in the horizontal ($x$-direction parallel to the emitter array) and vertical ($y$) directions respectively.
@@
 Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry \cite{Rimington_thermal_lens_2004}.
-The specifications of our dual-axis collimated diodes \cite{website_limo} indicate that heat source can be written as
+The specifications of our dual-axis collimated diodes \cite{website_limo} indicate that heat source can be represented mathematically as
 \begin{equation}
@@
 The longer 75cm arm of the resonator is terminated by a concave ($R$ = 1m and angle of incidence $< 2^\circ$) high reflector
+%page 8636
+focusing the intracavity radiation on a saturable Bragg reflector (SBR) with a 1\% reflectivity modulation depth at 1040nm (BATOP GmbH \cite{website_BATOP}) positioned a distance $z$ from the focusing mirror.
+The shorter 63cm arm is terminated by a plane output coupler with a 7\% transmission at 1040nm.
+As a free-running oscillator (SBR focusing section replaced by flat high reflector at $d_2$ = 75cm), the laser produced over 6W of TEM$_{00}$ ($M^2 < 1.2$) output power.
 
@@
 
-%page 8636
-focusing the intracavity radiation on a saturable Bragg reflector (SBR) with a 1\% reflectivity modulation depth at 1040nm (BATOP GmbH \cite{website_BATOP}) positioned a distance $z$ from the focusing mirror.
-The shorter 63cm arm is terminated by a plane output coupler with a 7\% transmission at 1040nm.
-As a free-running oscillator (SBR focusing section replaced by flat high reflector at $d_2$ = 75cm), the laser produced over 6W of TEM$_{00}$ ($M^2 < 1.2$) output power.
-
-The stability analysis of our diode-pumped Yb:KGW laser (\ref{fig:laser-cavity}\subref{fig:laser_schematic}), clearly indicates that the resonator configuration is unstable for thermal focal lengths less than about $-6m^{-1}$ (diopters) and is unlikely to oscillate for negative thermal focal lengths as this requires a vertical TEM$_{00}$-mode size $w_y$ greater than 180$\mu$m in the Yb:KGW gain medium --- a value that would result in strong absorptive losses due to the restrictive $W_y = 200\mu$m super-Gaussian vertical pump beam size in the Yb-doped crystal.
+The stability analysis of our diode-pumped Yb:KGW laser (\ref{fig:laser-cavity}\subref{fig:laser_stability}), clearly indicates that the resonator configuration is unstable for thermal focal lengths less than about $-6m^{-1}$ (diopters) and is unlikely to oscillate for negative thermal focal lengths as this requires a vertical TEM$_{00}$-mode size $w_y$ greater than 180$\mu$m in the Yb:KGW gain medium --- a value that would result in strong absorptive losses due to the restrictive $W_y = 200\mu$m super-Gaussian vertical pump beam size in the Yb-doped crystal.
 The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
@@
 
+\begin{figure}
+  \centering
+  \includegraphics{profile.jpg}
+  \caption[Autocorrelation measurements of employed ultrafast laser's pulse duration]{
+Second harmonic autocorrelation measurements and output spectra of the ML Yb:KGW laser; (a) the 250fs laser pulse duration (and fit assuming a sech$^2$ pulse shape) for the cavity depicted in \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} at an output power of 3.5W, (b) the 347fs pulse duration (and fit assuming a sech$^2$ pulse shape) obtained with an additional -5600fs$^2$ of round-trip GVD at an output power of 5W, (c) the 4.7nm full-width half maximum spectrum associated with the 250fs ML pulse of (a), and (d) spectral output for the ML resonator of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} at a laser power of 3.7W.
+  }
+  \label{fig:laser-profile}
+\end{figure}
+
 The laser is capable of generating shorter ML pulse durations.
@@
 \ref{fig:laser-performance} shows the laser's output power as a function of diode current for the cavity configuration depicted in \ref{fig:laser-cavity}\subref{fig:laser_schematic}; a total negative GVD of $-5,200\text{fs}^2$ per round-trip and a 7\% transmission output coupler.
-Soliton mode-locked (ML) operation is observed for diode currents greater than 23A, corresponding to a continuous-wave (cw) output power of 1.8W.
-Between diode currents (output powers) of 23 and 25A (1.8 and 2.3W), the resonator operates mainly in the cw mode, but will transition to ML operation upon being subjected to a perturbation.
+Soliton mode-locked (ML) operation is observed for diode currents greater than 23A, corresponding to a continuous-wave (CW) output power of 1.8W.
+Between diode currents (output powers) of 23 and 25A (1.8 and 2.3W), the resonator operates mainly in the CW mode, but will transition to ML operation upon being subjected to a perturbation.
 The switching to ML operation is accompanied by a 250mW increase in output
@@
   \centering
-  \includegraphics{profile.jpg}
-  \caption[Autocorrelation measurements of employed ultrafast laser's pulse duration]{
-Second harmonic autocorrelation measurements and output spectra of the ML Yb:KGW laser; (a) the 250fs laser pulse duration (and fit assuming a sech$^2$ pulse shape) for the cavity depicted in \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} at an output power of 3.5W, (b) the 347fs pulse duration (and fit assuming a sech$^2$ pulse shape) obtained with an additional -5600fs$^2$ of round-trip GVD at an output power of 5W, (c) the 4.7nm full-width half maximum spectrum associated with the 250fs ML pulse of (a), and (d) spectral output for the ML resonator of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} at a laser power of 3.7W.
+  \includegraphics{performance.jpg}
+  \caption[Power measuments of employed ultrast laser]{
+Performance of the diode-pumped, TLS, Yb:KGW laser (in the cavity configuration of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} ) as a function of applied diode current: continuous-wave (CW) output power (black squares); mode locked (ML) output power (red circles); frequency doubled power (green triangles); and second harmonic generation (SHG) efficiency (line).
   }
-  \label{fig:laser-profile}
+  \label{fig:laser-performance}
 \end{figure}
@@
 Similar laser performance is observed with the insertion of an additional $-5,600\text{fs}^2$ of cavity GVD per round trip, which generates 347fs sech$^2$ pulses (\ref{fig:laser-profile}(b)); that is, a DC spectral component appears on the short wavelength side of a narrower $\sim$3.5nm ML spectrum at output powers above 5W.
-
-\begin{figure}
-  \centering
-  \includegraphics{performance.jpg}
-  \caption[Power measuments of employed ultrast laser]{
-Performance of the diode-pumped, TLS, Yb:KGW laser (in the cavity configuration of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} ) as a function of applied diode current: continuous-wave (CW) output power (black squares); mode locked (ML) output power (red circles); frequency doubled power (green triangles); and second harmonic generation (SHG) efficiency (line).
-  }
-  \label{fig:laser-performance}
-\end{figure}
 

### 2013-03-18 inc/hardware/column.tex
added aligner picture
added additional pictures to the hardware section

@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \section{Column Design and Construction}
@@
 This system is used to mount the accelerator anode and the the deflector plates which are both to be described later.
+
+  \begin{figure}
+    \centering
+    \includegraphics{chamber.jpg}
+    \caption[Side view of main chamber]{
+      Assembled side view of the main chamber section of the UEM column at UIC.
+      To the left are the high-voltage passthrough and ``port aligner'' to which the cathode is mounted.
+      Visible inside the window are the cathode Wehnelt and photocathode (left), the side of the anode, and the side of the deflector plates assembly just after the anode (right).
+      The anode and plate-assembly are mounted by the ``eV parts'' system attached to the main chamber.
+      The wires and connectors supplying the voltage to the plates are also clearly visible.
+    }
+    \label{fig:chamber-pic}
+  \end{figure}

### 2013-03-18 inc/hardware/deflector.tex
work on short titles for figures (continuing)
added additional pictures to the hardware section

@@
   \include{deflector_schematic}
-  \caption{
+  \caption[Schematic diagram of one pair of deflection plates]{
     Schematic diagram of one pair of deflection plates.
@@
 These plates therefore are used as a fine adjustment, as opposed to the coarse adjustment provided by aligning the acceleration gap.
+
+\begin{figure}
+  \centering
+  \includegraphics{deflector_plates.jpg}
+  \caption[Picture of custom deflector plates]{
+    Picture of the custom built deflector plates for use in the UEM at UIC.
+    A pair of vertical and horizontal plates are mounted in a non-conductive housing.
+    Each plate has a pin for connecting to voltage or ground.
+    Clearance drilled holes allow mounting onto the ``eV parts'' cage inside the chamber.
+  }
+  \label{fig:deflector-plates-pic}
+\end{figure}

### 2013-03-18 inc/hardware/introduction.tex
introduction to hardware is now its own file

--- 2013-03-01 main.tex
@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
-
-\documentclass{uicthesi}
-\usepackage{amsmath,amsfonts,amssymb}
-\usepackage{esint}
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage{ifthen}
-\usepackage{pgffor}
-
-%\usepackage{euler}
-%\usepackage{breqn}
-
-\usepackage[version=3]{mhchem}
-
-\usepackage{import}
-\usepackage{subfig}
-
-\usepackage{graphicx}
-\graphicspath{{./images/}}
-\DeclareGraphicsExtensions{.png,.pdf,.jpg}
-\input{mytikz}
-\setboolean{inputdata}{true}
-\usepackage{fancyvrb}
-\input{pygments}
-
-%% My convenience fuctions
-\providecommand{\vecprime}[1]{\vec{#1}\,^{\prime}}
-
-\providecommand{\smallT}{\ensuremath{{\scriptscriptstyle T}}}
-\providecommand{\smallD}{\ensuremath{{\scriptscriptstyle D}}}
-\providecommand{\smallF}[0]{ { \scriptscriptstyle F } }
-%\providecommand{\smallZ}{\ensuremath{{\scriptscriptstyle Z}}}
-
-\providecommand{\punct}[1]{\ensuremath{\, \text{#1}}}
-
-\providecommand{\abbr}[2]{\item[#1\hfill] #2}
-
-\usepackage{color}
-\providecommand{\fixme}[1]{\textcolor{red}{#1}}
-
-%% Special handling for using biblatex vs bibtex ==
-\newboolean{usebiblatex}
-\setboolean{usebiblatex}{false}
-
-\ifthenelse{\boolean{usebiblatex}}{
-  \usepackage{biblatex}
-  %\usepackage[backend=biber]{biblatex-chicago}
-  \addbibresource{bibliography.bib}
-  \addbibresource{biblaser.bib}
-  \addbibresource{bibeste.bib}
-}{}
-%% ==
-
-\title{Title}
-\author{Joel Berger}
-\pdegrees{of the\\University of Illinois at Chicago}
-\degree{Doctor of Philosophy in Physics}
-
-\begin{document}
-\maketitle
-
-%\copyrightpage
-\newpage
-\begin{figure}
-  \centering
-  \includegraphics{by-nc-nd}
-\end{figure}
-This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
-To view a copy of this license, visit \url{http://creativecommons.org/licenses/by-nc-nd/3.0/us/} or send a letter to
-\begin{verbatim}
-Creative Commons
-444 Castro Street, Suite 900
-Mountain View, CA, 94041, USA.
-\end{verbatim}
-
-\dedication
-The dedication.
-
-\acknowledgment
-The acknowledgment.
-\initials{JAB}
-
-\preface
-The preface.
-
-\tableofcontents
-\listoftables
-\listoffigures
-
-\input{abbreviations}
-
-\summary
-A summary is required.
-
-\chapter{Introduction}
-
-\chapter{Review of Foundational Modeling Efforts} \label{chap:previous_models}
-
-  \subimport{inc/ms_model/}{ms_model}
-
-\chapter{Extension of the Analytic Gaussian Model\\for UEM Column Modeling} \label{chap:extension}
-
-Modeling a system as complex as an Ultrafast Electron Microscope is an enourmous task.
-In Chapter \ref{chap:previous_models}, I have highlighted some previous efforts at modeling electron pulse evolution.
-While these are excellent works in their own right, none provide the complete picture of the evolution of an electron pulse through the entire column.
-In the following chapter, I will present a model which builds on and extends those individual pieces to form a single model; one which is generic enough to model many different columns and is computationally fast enough to be used to design and optimize a column.
-
-  \subimport{inc/extension/initial_conditions/}{initial_conditions}
-
-  \subimport{inc/extension/liouvilles_theorem/}{liouvilles_theorem}
-
-  %TODO refer from here to Free expansion results
-
-  \subimport{inc/extension/external_forces/}{external_forces}
-
-  \subimport{inc/gun/}{gun_model}
-
-  \subimport{inc/extension/implementation/}{implementation}
-
-\chapter{Model Results, Predictions and Implications}
-
-\section{Free Expansion Propagation}
-
-  \subimport{inc/model_results/shapes/}{shapes}
-
-  \subimport{inc/model_results/spacecharge/}{spacecharge}
-
-\section{Propagtion under External Forces}
-
-  \subimport{inc/model_results/lens_charge/}{lens_charge}
-
-\chapter{Prototype Instrument}
 
@@
 \end{figure}
-
-  \subimport{inc/laser/}{laser}
-
-  \subimport{inc/hardware/}{column}
-
-  \subimport{inc/gun/}{gun_design}
-
-  \subimport{inc/magnetic_lenses/}{maglens}
-
-  \subimport{inc/hardware/}{deflector}
-
-\chapter{Photocathode Engineering}
-
-  \subimport{inc/photocathode/setup/}{setup}
-
-\section{Flat Metal Photocathodes}
-
-  \subimport{inc/photocathode/plasmon/}{plasmon}
-
-  \subimport{inc/photocathode/este/}{este}
-
-\chapter{Experimental Results}
-
-%uncomment when multiple appedicies are used 
-%\appendices
-\appendix
-  \subimport{inc/extension/implementation/}{appendix-script}
-
-\newpage
-
-%% Special handling for using biblatex vs bibtex ==
-\ifthenelse{\boolean{usebiblatex}}{
-  \printbibliography
-}{
-  \bibformc
-  \bibliography{bibliography,biblaser,bibeste}
-}
-%% ==
-
-%\vita
-%This is where the vita goes.  Its organization is left as an exercise.
-%Hint: see the list of abbreviations.
-\end{document}

### 2013-03-18 main.tex
introduction to hardware is now its own file

@@
 
-  \begin{figure}
-    \centering
-    \includegraphics{inc/hardware/column.jpg}
-    \caption[Picture of the prototype UEM column at UIC]{
-      Picture of the majority of the prototype UEM column at UIC.
-      The main vacuum chamber is laid sideways, with the high voltage (gun) at the left and the imaging system on the right.
-      The magnetic levitated turbo pump hangs below the column, a scroll pump used for backing is not pictured.
-      The laser system is in the background.
-      All of the above is mounted on a custom shaped active vibration-canceling optical table.
-      The electronic control systems are not pictured.
-    }
-    \label{fig:column-pic}
-  \end{figure}
+  \subimport{inc/hardware/}{introduction}
 

### 2013-03-18 inc/hardware/introduction.tex
added some text the the hardware introduction

@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
+In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effect that column elements have upon it.
+However, microscopy is an experimental science, and designing a UEM is no less hands-on.
+
+It is our belief that a successful UEM project cannot simply stem from small modifications to the platform of a conventional electron microscope.
+In this chapter I describe the prototype UEM system that is being built at UIC.
+By necessity much of it is custom built, specifically designed to either allow flexibility or to meet some technical challenge.
 

### 2013-03-18 inc/laser/laser.tex
introduction to hardware is now its own file
tweaked laser intro

@@
 \section{Yb:KGW Laser System} \label{sec:laser}
-\subsection{Introduction}
+
 % Draws from:
@@
 
+In order to experiment using very short laser pulses, we have designed and built a custom Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) laser system.\cite{berger_high-power_2008}
+In practice, the flexibility afforded by having a custom-built system has outweighed its original performance benefit.
+I have shown in Section \ref{sec:initial_conditions} that the electron pulse duration is not a simple function of the laser pulse duration, yet, at the time the laser was built, this fact was not yet obvious.
+As such, our laser was designed to produce relatively high power output while generating ultrashort pulses, on the order of 250 fs for some conditions.
+
+\subsection{Laser Introduction}
 % page 8630
-Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
+Yb:KGW is becoming a widely used solid-state gain medium for ultrafast laser systems.
 As with other Yb-doped laser crystals \cite{Brenier_new_criteria}, this is primarily due to the gain medium's advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
@@
 %TODO is this first sentance correct or should I use "I" voice?
-In this section, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis \cite{Biswal_thermo_optical_05,mochalov_laser_1997,pujol_crystalline_1999}).
+In this chapter, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis \cite{Biswal_thermo_optical_05,mochalov_laser_1997,pujol_crystalline_1999}).
 The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media \cite{Rimington_thermal_lens_2004}.

### 2013-03-19 inc/laser/laser.tex
tweaks to the laser section

@@
 
+For most photoemission sources, the fundamental laser output's 1.19 eV photons do not provide enough energy to excite electrons above the vacuum level, either in single or two-photon emission.\cite{eastman_photoelectric_1970,johnson_optical_1972,dowell_quantum_2009}
+By appropriate application of crystals with strong non-linear properties, two photons at frequency $\omega$ may be converted into one photon at frequency $2 \omega$.\cite{shen_principles_2002}
+Since these higher harmonics will be necessary for stimulating photoemission it is important to optimize and characterize the efficiency of this process.
+
 The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
@@
 a peak green pulse power of $\sim$130kW.
-This femtosecond visible output will be used as the driving pulsed laser radiation source for an ultrafast electron microscope \cite{king_ultrafast_2005}.
+This laser is the short-pulse photon source used to photoemit electrons for all the experimental results shown in this thesis.
 

### 2013-03-19 inc/hardware/column.tex
added some content, but also restructured some of the folders related to hardware

@@
 
-The vacuum chamber itself is almost entirely 4.5 in standard ConFlat hardware.
-Much has been purchased from off-the-shelf hardware suppliers.
-More complicated chamber components have been purchased from Kimball Physics Inc.
-
-Kimball's line of Multi-CF hardware has been especially useful in building the prototype column.
-The main chamber is a 4.5 in ``spherical cube,'' having six 4.5 in ConFlat ports on a cube, with eight 1.33 in ports on the corners.
-All ports on Multi-CF hardware also contain a set of grooves, designed as mounting points for a system of erector-set-like equipment called ``eV Parts.''
-This system is used to mount the accelerator anode and the the deflector plates which are both to be described later.
-
 \begin{figure}
@@
     Visible inside the window are the cathode Wehnelt and photocathode (left), the side of the anode, and the side of the deflector plates assembly just after the anode (right).
-      The anode and plate-assembly are mounted by the ``eV parts'' system attached to the main chamber.
+    The anode and plate-assembly are mounted by the ``eV Parts'' system attached to the main chamber.
     The wires and connectors supplying the voltage to the plates are also clearly visible.
@@
 \end{figure}
+
+The vacuum chamber itself is almost entirely 4.5 in standard ConFlat hardware.
+Much has been purchased from off-the-shelf hardware suppliers.
+More complicated chamber components have been fabricated in-house or purchased from Kimball Physics Inc.
+
+Kimball's line of Multi-CF hardware has been especially useful in building the prototype column.
+The main chamber is a 4.5 in ``spherical cube,'' having six 4.5 in ConFlat ports on a cube, with eight 1.33 in ports on the corners.
+All ports on Multi-CF hardware also contain a set of grooves, designed as mounting points for a system of erector-set-like equipment called ``eV Parts.''
+This system is used to mount the accelerator anode and the the deflector plates which are both to be described later.

### 2013-03-19 inc/hardware/magnetic_lenses/maglens.tex
added some content, but also restructured some of the folders related to hardware

--- 2013-03-07 inc/magnetic_lenses/maglens.tex
@@
 
-\subsection{Motivation}
+%\subsection{Motivation}
 % montgomery_some_1961
@@
 Second, and more importantly, near a lens crossover (i.e. focus), the charge density can increase rapidly.
-This charge density can lead to an unacceptable amount of space-charge interaction (see section \ref{sec:free_spacecharge}).
+This charge density can lead to an unacceptable amount of space-charge interaction.
+These effects have been explored in Sections \ref{sec:free_spacecharge} and \ref{sec:mag_lens_charge}.
+The lenses I present in this section were designed, in part, using information computed by the extended AG model (Section \ref{sec:mag_lens_model}).
 
@@
 While in principle one could design a magnetic lens using \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
-Further --- as suggested by the use of the $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility.
+Further --- as suggested by the use of the auxiliary field $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility.
 These materials, called ``magnetic shielding'' since it captures the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.

### 2013-03-19 inc/design_considerations/considerations.tex
added new UEM design considerations chapter

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effect that column elements have upon it.
However, microscopy is an experimental science, and designing a UEM is no less hands-on.

When designing a UEM system, special care must be given to the creation and propagation of the electron beam.
In the case of single-shot UEM, several factors lead to two major considerations: first, that a large area photoemission source is preferred to smaller area or even single point emitters, and second, the transverse emittance of the beam must be minimized.

These considerations drive the remainder of this thesis.
In this chapter, I will first convince the reader of these considerations.
As a direct consequence of these considerations, it is our belief that a successful UEM project cannot simply stem from small modifications to the platform of a conventional electron microscope.
In the following chapters I will show how we are addressing these considerations for the prototype system at UIC.

%Ultrafast Electron Microscopy (UEM) is a subclass of a more general concept, Dynamic Transmission Electron Microscopy (DTEM).
%DTEM is a more general term for any electron microscope with higher time resolution, typically employing laser-stimulated photoemission.

\section{Pulse Charge Requirements}

It is easy to believe that in order to make a useful image in an electron microcope, the number of electrons collected must be above some minimum number.
The Rose criterion \cite{rose_television_1948} indicates that an average detector signal strength of $\sim$100 electrons/pixel is required for adequate gray scaling of the image.
For a 1k x 1k CCD camera, this implies an incident electron pulse with $N=10^8$ \cite{armstrong_prospects_2007}.

In a conventional electron microscope the electrons are generated by one of several processes, often by heating the cathode or applying a large field to the cathode.
Although these processes produce a very large number of electrons, far more than required by the Rose criterion, they are essentially non-directional.
A pinhole is used to eliminate most of the electrons with the exception of those which will travel down the column, reducing the beam current but producing a high-quality beam. %TODO ref(s)

In fact, for conventional electron microscopes without time resolution, the Rose criterion may be met for even very low beam current by simply allowing the image to acquire for a long enough time to collect sufficent electrons.
Further, high beam current may be damaging to the sample over time, so often a lower beam current is preferred.

By definition, for single-shot imaging with a DTEM, all of these electrons must be contained in a single pulse.
Although sample damage concerns may be neglected, destructive measurements will result in useful images, the use of a pinhole is likely to reduce the pulse charge below the useful level.
This means that the designer of a DTEM must endevor to keep all of the created electrons.

It should be noted that for diffraction studies with a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse \cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.

### 2013-03-19 inc/hardware/introduction.tex
added new UEM design considerations chapter

--- 2013-03-18 inc/hardware/introduction.tex
@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
+To satisfy the first restriction, many of the common elements of an electron microscope column must be redesigned to accomodate the larger beam area.
+I will then describe how we have addressed the necessity of the large-area source in building the prototype UEM system begin constructed at UIC.
+In order to allow flexibility or to meet some technical challenges presented by the above considerations, much of the system is custom built.
 

### 2013-03-19 inc/design_considerations/considerations.tex
rename some sections remove "experimental results" section (unused)
added large $\Delta x$ limitation to considerations

@@
 
-\section{Pulse Charge Requirements}
+\section{Pulse Charge Requirement}
 
@@
 The Rose criterion \cite{rose_television_1948} indicates that an average detector signal strength of $\sim$100 electrons/pixel is required for adequate gray scaling of the image.
-For a 1k x 1k CCD camera, this implies an incident electron pulse with $N=10^8$ \cite{armstrong_prospects_2007}.
+For a 1k $\times$ 1k CCD camera, this implies an incident electron pulse with $N=10^8$ \cite{armstrong_prospects_2007}.
 
@@
 It should be noted that for diffraction studies with a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse \cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.
+
+\section{Generation Current Limitations}
+
+Since we are now aware that there is a large minimum number of electrons needed per pulse, it is only natural to ask if there is the practical limit on how much charge (i.e., the number of electrons/pulse, $N$) can be extracted from a pulsed laser-driven DC photoelectron gun.
+When the incident laser pulse duration $\tau$ is longer than the time-of-flight (TOF) $t_{gun}$ of a single electron across the acceleration gap $d$ of the DC gun with a potential $V_{DC}$, $\tau > t_{gun} = d \sqrt{2m/eV_{DC}}$, then the maximum photocurrent density $J_{CL}$ is given by the well-known Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923}:
+\begin{equation}
+  J_{CL} = \frac{4 \varepsilon_0}{9} \sqrt{\frac{2e}{m}} \frac{V_{DC}^{3/2}}{d^2}
+\end{equation}
+where $e$ is the electron charge and $\varepsilon_0$ is the vacuum permittivity.
+
+This steady-state law for the maximum current density that can be achieved is not valid in the regime of interest for UEM (and UED) where the use of picosecond and femtosecond photocathode drive laser pulses implies that $t_{gun}$ is usually much greater than the electron pulse duration $\tau_e$.
+In fact, though there is no clear definition in the literature or the community, we propose that this is useful delimination between term UEM and the more general term DTEM; reserving the term UEM for when the pulse can be completely contained inside the accelerator, i.e.
+$t_{gun} > \tau_e$.
+In this short pulse case, an equivalent diode approximation \cite{valfells_effects_2002} allows for the definition of a critical current density associated with virtual cathode formation in terms of the steady-state Child-Langmuir limit:
+\begin{equation}
+  J_{crit} = \frac{27}{4} \left ( \frac{ 1 - \sqrt{1-\tfrac{1}{3} \chi^2} }{ \chi^3 } \right ) J_{CL}
+\end{equation}
+where $\chi = \tau_e / t_{gun}$.
+Thus, for the usual case in UEM, where $\chi \ll 1$, the critical current density is simply given by
+\begin{equation}
+  J_{crit} = \frac{\varepsilon_0 E_{DC}}{\tau_e}
+\end{equation}
+where the DC gun field $E_{DC} = V_{DC}/d$.
+This relation may be rewritten to yield a critical number of photoemitted electrons/pulse before virtual cathode formation:
+\begin{equation} \label{eq:n_crit}
+  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d}
+\end{equation}
+In obtaining the above equation, we have used an irradiated photocathode area of $\pi w^2$ associated with an incident Gaussian laser pulse of the form of \ref{eq:laser_form}.
+The relation clearly indicates that $N_{crit}$ is only dependent upon the initial DC surface field (i.e., $E_{DC}$) at the photocathode and the illuminated area --- as expected because a virtual cathode forms when the initial photocathode surface charge density associated with $E_{DC}$ is photoemitted, effectively screening the cathode from the acceleration field.
+
+For a typical DC gun with $E_{DC} \approx $ 30 kV/cm (e.g., $V_{DC} = $ 100 kV and $d \approx $ 3 cm), \ref{eq:n_crit} reveals that the generation of $10^8$ electrons/pulse needed for single-shot imaging applications in a DTEM with a 1k $\times$ 1k CCD detector requires an incident laser beam area $\pi w^2$ of about 1 mm$^2$ --- the virtual cathode being formed for the extraction of more than $1.7 \times 10^8$ electrons/mm$^2$ in this short pulse regime.
+We note that \ref{eq:n_crit} does not express a limit on the number of electrons/pulse $N$ that can be emitted from a pulsed laser-driven DC photoelectron gun, it merely indicates a critical number before virtual cathode formation occurs.
+More electrons can be produced, at the expense of a severely distorted temporal electron pulse profile \cite{valfells_effects_2002}.
+
+This finally leads to the first consideration for UEM design mentioned in the introduction: for the clean pulsed emission of sufficient to create a useful image, the generated beam should have a HW1/eM width
+\begin{equation}
+  w \gtrsim \sqrt{\frac{N e d}{\pi \varepsilon_0 V_{DC}}} \,\text{.}
+\end{equation}
+For our example case of a 1k $\times$ 1k CCD, $N \ge 10^8$ electrons/pulse, we require $w \gtrsim $ 0.56 mm.
+
+\section{Beam Quality and Resolution Limit}

### 2013-03-20 inc/hardware/column.tex
clean up the introductions to the hardware and consideration sections

@@
 Apart from the laser system, the majority of an Ultrafast Electron Microscope column is either vacuum hardware or contained within the vacuum system.
-It is common in the UEM field to modify an existing Transmission Electron Microscope column, adapting it for laser-stimulated electron generation.
-The system presented in the following section has been designed and built either from scatch or from commercially available parts.
-The reasons for this are two-fold, the ease of prototyping inherent in using replacable parts and the physical limitations our unique choices have imposed on our beam profile.
+The vacuum chamber itself is almost entirely 4.5 in standard ConFlat hardware.
+Much has been purchased from off-the-shelf hardware suppliers.
+Some of the more specialized chamber components have been fabricated in-house or purchased from Kimball Physics Inc.
 
@@
 
-The vacuum chamber itself is almost entirely 4.5 in standard ConFlat hardware.
-Much has been purchased from off-the-shelf hardware suppliers.
-More complicated chamber components have been fabricated in-house or purchased from Kimball Physics Inc.
-
 Kimball's line of Multi-CF hardware has been especially useful in building the prototype column.
-The main chamber is a 4.5 in ``spherical cube,'' having six 4.5 in ConFlat ports on a cube, with eight 1.33 in ports on the corners.
+The main chamber is a 4.5 in ``spherical cube,'' a large stainless-steel sphere which has six 4.5 in ConFlat ports oriented as a cube, with another eight 1.33 in ports on the corners of the cube.
 All ports on Multi-CF hardware also contain a set of grooves, designed as mounting points for a system of erector-set-like equipment called ``eV Parts.''

### 2013-03-20 inc/hardware/introduction.tex
clean up the introductions to the hardware and consideration sections

--- 2013-03-18 inc/hardware/introduction.tex
@@
 To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
+It is common in the DTEM field to modify an existing Transmission Electron Microscope column, adapting it for laser-stimulated electron generation.
+For the reasons covered in Section \ref{sec:childs_law}, the prototype UEM at UIC is being built to accomodate large width electron pulses.
+To satisfy this restriction, it has been designed and built from parts either fabricated in-house or from commercially available, both for the ease of prototyping inherent in using replacable parts and the accomodate the physical limitations our choice of a large beam width have imposed on the standard electron microscope column components.
+
+In this chapter, I will first describe the laser system which was custom built for the prototype UEM system being constructed at UIC.
+I will then describe how we have addressed the necessity of the large-area electron beam in each of the individual components of the UEM column.
 

### 2013-03-21 inc/gun/gun_design.tex
tweak gun design intro now that design considerations presents the argument for large area emission

@@
 The acceleration region of our prototype UEM system has several requirement which are unique in comparison to standard electron guns.
-Primarily, given the limitations on the beam discussed earlier, %TODO link to discussion of Childs law beam size limitation
-the employ laser pulse must be a relatively large area.
+As discussed in \ref{chap:considerations} notable requirements include the need to handle a large photoemission area and the inability to employ a pin-hole to selectively clean the electron beam.
 This concern leads to several additional criteria.
-To accomodate this large beam, the anode aperture must be similarly large; in fact, this will be larger than the emission source due to beam expansion during acceleration.
-Also, in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the accelerating electic field given that it will span a wider proportion of the accelerator than thin beams used in standard TEM.
+The system must first be able to accept the large laser spot size needed to generate the electron beam.
+Then, in order to accomodate this large electron pulse, both the cathode Wehnelt and anode aperture must be similarly large; in fact, the beam size at the anode aperture will typically be larger than at the emission source due to beam expansion during acceleration.
+Also, care must be taken to ensure a relatively uniform electric field for acceleration of the pulse.
+While in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the electic field given that it will span a wider proportion of the accelerator than the thin beams used in conventional TEM.
 

### 2013-03-21 inc/gun/gun_design.tex
tweaks in gun section

@@
 A schematic of this design, overlayed by a Finite Element Model (FEM) simulation of the electic field, is shown in \ref{fig:gun-field}.
-The schematic shows that the aperture concerns are easily satisfied.
+The simulation was performed using \texttt{FreeFEM++}, a free finite element partial differential equation solver and meshing toolkit (\url{http://www.freefem.org/ff++/index.htm}).
+The FEM simulation modeled the features of the anode and cathode, while emulating gap walls at infinity by placing a linear potential on gap edges at a large distance.
 From the field equipotentials, it is clear that this design has a consistently flat electric field for large displacements off the central axis.
+Since the shape of the field is scale independent, the schematic shows that the aperture width and laser acceptance concerns are easily satisfied for an appropriate choice of scale.
 
@@
   \caption[Schematic of the employed Togawa-inspired custom cathode-anode pair]{
-    Schematic of the employed Togawa-inspired cathode ($-V$) and anode (0V) pair.
+    Schematic of the employed Togawa-inspired cathode ($-V$) and anode (0V, ground) pair.
     The emission source is inserted behind the cathode Wehnelt.
+    The separation of the cathode and anode is adjustable to allow a greater range of input laser angles if needed. 
     Also shown are simulated equipotential lines generated by FEM modeling.
-    The separation of the cathode and anode is adjustable to allow a greater range of input laser angles if needed. 
+    The acceleration gap is modeled as having linear potential along an edge at a very large width. 
   }
@@
 
-Of course, as derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electic field.
+As derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electic field.
 This field will cause a divergence of the beam.
@@
 
-%TODO comment on alignment
-%TODO comment on high voltage
-%TODO comment on EV parts etc
-
-Re-alignment of cathode and anode is sometimes necessary to direct the emitted beam down the center axis of the column.
+A gross realignment of the cathode and anode is sometimes necessary to direct the emitted beam down the center axis of the column.
 To facilitate this, the prototype column has the photocathode and Wehnelt mounted on the end of the high voltage feedthrough, which is in turn mounted on a two axis alignment mechanism.

### 2013-03-22 inc/gun/gun_design.tex
move FEM data from figure in acc model to acc design and forward reference to it

@@
 
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}
+    \input{onaxis_FEM}
+  \end{tikzpicture}
+
+  \caption[An example of a tanh form acceleration region for three different ``sharpness'' values]{
+    An example of a tahn form acceleration region (overall length 6.35mm) for three different ``sharpness'' values, the same data as in \ref{fig:gun-sharpness}\subref{fig:field_on_axis}.
+    The square data points represent the on-axis electric field data from FEM simulations.
+  }
+  \label{fig:field_on_axis_fem}
+\end{figure}
+
+\ref{fig:field_on_axis_fem} compares the FEM computed electric field to the assumed model form (\label{eq:anode_tanh}) for several values of the sharpness parameter $s$; the lines are the same as in \ref{fig:gun-sharpness}\subref{fig:field_on_axis}.
 As derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electic field.

### 2013-03-22 inc/gun/gun_model.tex
move FEM data from figure in acc model to acc design and forward reference to it

@@
 Indeed, this form is analgous to the super-Gaussian envalope (\ref{eq:reg_of_influence}), however it is open ended on the cathode side, which is necessary for the electrons to be accelerated away from the photocathode.
-Given this form (demonstrated in Section \ref{sec:gun_design}) the ``lensing'' strengths are then
+Given this form the ``lensing'' strengths are then
 \begin{gather}
@@
       \label{fig:field_on_axis}
+      \begin{tikzpicture}
         \input{onaxis}
+      \end{tikzpicture}
     }
@@
     An example of a tahn form acceleration region (overall length 6.35mm) for three different ``sharpness'' values.
-    The square data points in \protect\subref{fig:field_on_axis} represent the on-axis electric field data from FEM simulations, see Section \ref{sec:gun_design}.
     The colored electric fields in figure \protect\subref{fig:field_on_axis} correspond to the pulse width (solid) and length (dashed) in figure \protect\subref{fig:tanh_sharpness}.
+    This form is compared to FEM simulated electric field data in \ref{fig:field_on_axis_fem}.
   }
@@
 
-\ref{fig:gun-sharpness} shows an example acceleration region for three values of the ``sharpness'' $s$ as well as longitudinal FEM field data which is explained further in Section \ref{sec:gun_design}.
-%TODO possibly remove the FEM data? it doesn't help much here and is confusing since it hasn't been introduced
+\ref{fig:gun-sharpness} shows an example acceleration region for three values of the ``sharpness'' $s$.
 What is clear from this figure is that the dynamics are not highly coupled to the value of this hard-to-quantify parameter.
@@
 This effect is due to the slower electrons remaining in the cavity longer and thus receiving an additional kick, as described in \cite{oudheusden_electron_2007}.
+This longitudinal electric field shape is compared to the electric field computed by finite-element (FEM) simulations of our acceleration cavity in Section \ref{sec:gun_design}.

### 2013-03-22 inc/hardware/magnetic_lenses/maglens.tex
added polepiece-design drawing, this included moving some labels on the original schematic. Note that the figure description still needs to be changed!

@@
 This gap can be much smaller than the total width of the coil turns, and yet contain in a field commensurate with the full coil.
+This geometry is presented schematically as \ref{fig:polepiece-schematic}.
 
@@
   }
+  \label{fig:polepiece-schematic}
 \end{figure}
 
-El-Kareh and El-Kareh \cite{el-kareh_electron_1970} provide an analysis of symmetric lenses with magnetic shielding and a pole-piece in chapter 8, with the most useful information being in section 8.7.
+El-Kareh and El-Kareh provide an analysis of symmetric lenses with magnetic shielding and a pole-piece (see Ref.
+\cite{el-kareh_electron_1970} chapter 8, with the most useful information being in section 8.7).
 They define
@@
 
-%TODO check S and D as I seem to remember we have S=0.1in
 %TODO need a description of the steel inner tube causing an effective diameter larger than the actual bore
-For the UEM, I chose a bore diameter $D=2R=$0.1in and a pole-piece gap $S$=0.5in, for an ratio $S/D=0.2$.
+For the UEM, I chose a bore diameter $D=2R=$0.5in and a pole-piece gap $S$=0.1in, for an ratio $S/D=0.2$.
 By their table 8.2a this results in $\beta=0.0146$. Then, to create a focal length of 4in ($f/R=16$) table 8.13 gives a value for $k^2 \approx 0.06$.
@@
 This result is easily confirmed manually; the oxygen-free copper wire coils remain cool to the touch under these operating conditions.
+
+\begin{figure}
+  \centering
+  \input{polepiece-design}
+  \caption[Side-view schematic of a shielded magnetic lens with a pole-piece]{
+    Side-view schematic of a shielded magnetic lens with a pole-piece.
+    The windings (current loops) are contained within the shielding.
+    The magnetic field created by these windings is contained inside the shielding except in the gap.
+  }
+  \label{fig:polepiece-design}
+\end{figure}

### 2013-03-22 inc/hardware/magnetic_lenses/maglens.tex
further tweaks to magnetic lenses

@@
   \input{polepiece-schematic}
-  \caption[Side-view schematic of a shielded magnetic lens with a pole-piece]{
-    Side-view schematic of a shielded magnetic lens with a pole-piece.
+  \caption[Schematic of a generic magnetic lens with a pole-piece]{
+    Side-view schematic of a generic magnetic lens with a pole-piece.
     The windings (current loops) are contained within the shielding.
@@
 
-%TODO need a description of the steel inner tube causing an effective diameter larger than the actual bore
-For the UEM, I chose a bore diameter $D=2R=$0.5in and a pole-piece gap $S$=0.1in, for an ratio $S/D=0.2$.
+\begin{figure}
+  \centering
+  \input{polepiece-design}
+  \caption[Design schematic of our magnetic lens with a pole-piece]{
+    Side-view schematic of the magnetic lens with a pole-piece designed for the UEM at UIC.
+    On the center axis is the inner drift tube, with pole-piece gap separation disk.
+    The end pieces are shown top and bottom.
+    One bridge piece is shown detached (left) and one attached (right).
+    Note that this figure is not drawn to the scale of the lenses at UIC, but rather to match the general schematic in \ref{fig:polepiece-schematic}.
+  }
+  \label{fig:polepiece-design}
+\end{figure}
+
+\ref{fig:polepiece-design} shows a simplistic design schematic of the lens designed and built for the UEM at UIC.
+The central bore is a tube of stainless steel having a magnetic permeability of 1; this behaves magnetically as if it is not involved in the field shaping.
+In this way the inner bore of the tube is in vacuum with minimal welding necessary.
+The outer diameter of tube is milled down to an outer diameter of $D=2R=$0.5in, though a disk of larger diameter, and length $S$=0.1in, is left at the midpoint, which will separate the shielding material and be the pole-piece gap.
+
+The high magnetic permeability shielding is in eleven pieces.
+There are two end pieces end pieces are revolution solid of ``L'' profile.
+These ends are slid together touching the disk at the midpoint of the drift tube; the disk is of the proper outer diameter to match the sheilding's smaller outer diameter.
+The combination of the two ends and this disk make a ``U'' shape which contains the windings.
+
+The top side of the end's larger outer side is flattend on eight sides.
+These are the mount points for eight bridge pieces which connect the ends, completing the field-capturing shielding loop.
+The gaps between the shielding bridges also allow for direct air cooling of the magnetic coils (see Section \ref{sec:mag_lens_thermal}).
+
+I chose drift tube outer diameter to be $D=$0.5in and a pole-piece gap $S$=0.1in, for an ratio $S/D=0.2$.
 By their table 8.2a this results in $\beta=0.0146$. Then, to create a focal length of 4in ($f/R=16$) table 8.13 gives a value for $k^2 \approx 0.06$.
@@
     Side-view of the two custom large-bore magnetic lenses designed and built at UIC, seen installed in the prototype UEM.
-    The coils are visible as is high permeability shielding used for the pole-piece.
+    The coils are visible in the gaps between the high permeability shielding ``bridges''.
+    The bridges are attached to the tops of the end pieces by bolts to complete the shielding loop.
     The outer (larger) plates and struts are for structural integrity only.
@@
 
-\subsection{Thermal Properties}
+\subsection{Thermal Properties} \label{sec:mag_lens_thermal}
 
@@
 This result is easily confirmed manually; the oxygen-free copper wire coils remain cool to the touch under these operating conditions.
-
-\begin{figure}
-  \centering
-  \input{polepiece-design}
-  \caption[Side-view schematic of a shielded magnetic lens with a pole-piece]{
-    Side-view schematic of a shielded magnetic lens with a pole-piece.
-    The windings (current loops) are contained within the shielding.
-    The magnetic field created by these windings is contained inside the shielding except in the gap.
-  }
-  \label{fig:polepiece-design}
-\end{figure}

### 2013-03-22 inc/gun/gun_design.tex
more tweaks

@@
 
-\ref{fig:field_on_axis_fem} compares the FEM computed electric field to the assumed model form (\label{eq:anode_tanh}) for several values of the sharpness parameter $s$; the lines are the same as in \ref{fig:gun-sharpness}\subref{fig:field_on_axis}.
+\ref{fig:field_on_axis_fem} compares the FEM computed electric field to the form asserted for the AG model (\ref{eq:anode_tanh}) for several values of the sharpness parameter $s$; the lines are the same as in \ref{fig:gun-sharpness}\subref{fig:field_on_axis}.
 As derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electic field.

### 2013-03-22 inc/hardware/magnetic_lenses/maglens.tex
more tweaks

@@
 
-Magnetic lenses present some challenges for Ultrafast Electron Microscopy.
-First, the beams are likely to be much larger and thus a UEM will need large-aperture lenses.
+Magnetic lenses built for use in Ultrafast Electron Microscopy will be subject to the some unique design criteria.
+First, because of the considerations presented in Chapter \ref{chap:considerations}, the beams are likely to be much larger than in conventional electron microscopy and thus a UEM will need large-aperture lenses.
 Second, and more importantly, near a lens crossover (i.e. focus), the charge density can increase rapidly.
@@
 These effects have been explored in Sections \ref{sec:free_spacecharge} and \ref{sec:mag_lens_charge}.
-The lenses I present in this section were designed, in part, using information computed by the extended AG model (Section \ref{sec:mag_lens_model}).
+The lenses I present in this section were designed, in part, using information computed by the extended AG model (Section \ref{sec:mag_lens_model}) in order to address these concerns.
 
@@
 
-Discussion of the mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens}.
+Discussion of the formal mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens_model}.
 While in principle one could design a magnetic lens using \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
@@
 
+\begin{figure}
+  \centering
+  \input{polepiece-design}
+  \caption[Design schematic of our magnetic lens with a pole-piece]{
+    Side-view schematic of the magnetic lens with a pole-piece designed for the UEM at UIC.
+    On the center axis is the inner drift tube, with pole-piece gap separation disk.
+    The end pieces are shown top and bottom.
+    One bridge piece is shown detached (left) and one attached (right).
+    Note that this figure is not drawn to the scale of the lenses at UIC, but rather to match the general schematic in \ref{fig:polepiece-schematic}.
+  }
+  \label{fig:polepiece-design}
+\end{figure}
+
+\ref{fig:polepiece-design} shows a simplistic design schematic of the lens designed and built for the UEM at UIC.
+The central drift tube is made of stainless steel having a magnetic permeability of 1, thus magnetically not involved in the field shaping.
+In this way the inner bore of the tube is in vacuum with minimal welding necessary.
+The outer diameter of tube is machined down to $D=2R=$0.5in, though a disk of larger diameter, and length $S$=0.1in, is left at the midpoint, which will separate the shielding material and be the pole-piece gap.
+
+The high magnetic permeability shielding is in eleven pieces.
+There are two end pieces which are revolution solids of an ``L'' shaped profile.
+These ends slide together over the drift tube until touching the disk at the midpoint; the disk is of the proper outer diameter to match the sheilding's smaller outer diameter.
+The combination of the two ends and this disk make a ``U'' shape which contains the windings.
+
+The top of the each end's large outer disk (the walls of the winding container) are flattend on eight sides.
+These are the mount points for eight bridge pieces which connect the ends to complete the field-capturing shielding loop.
+The gaps between the shielding bridges also allow for direct air cooling of the magnetic coils (see Section \ref{sec:mag_lens_thermal}).
+
 El-Kareh and El-Kareh provide an analysis of symmetric lenses with magnetic shielding and a pole-piece (see Ref.
@@
 
-\begin{figure}
-  \centering
-  \input{polepiece-design}
-  \caption[Design schematic of our magnetic lens with a pole-piece]{
-    Side-view schematic of the magnetic lens with a pole-piece designed for the UEM at UIC.
-    On the center axis is the inner drift tube, with pole-piece gap separation disk.
-    The end pieces are shown top and bottom.
-    One bridge piece is shown detached (left) and one attached (right).
-    Note that this figure is not drawn to the scale of the lenses at UIC, but rather to match the general schematic in \ref{fig:polepiece-schematic}.
-  }
-  \label{fig:polepiece-design}
-\end{figure}
-
-\ref{fig:polepiece-design} shows a simplistic design schematic of the lens designed and built for the UEM at UIC.
-The central bore is a tube of stainless steel having a magnetic permeability of 1; this behaves magnetically as if it is not involved in the field shaping.
-In this way the inner bore of the tube is in vacuum with minimal welding necessary.
-The outer diameter of tube is milled down to an outer diameter of $D=2R=$0.5in, though a disk of larger diameter, and length $S$=0.1in, is left at the midpoint, which will separate the shielding material and be the pole-piece gap.
-
-The high magnetic permeability shielding is in eleven pieces.
-There are two end pieces end pieces are revolution solid of ``L'' profile.
-These ends are slid together touching the disk at the midpoint of the drift tube; the disk is of the proper outer diameter to match the sheilding's smaller outer diameter.
-The combination of the two ends and this disk make a ``U'' shape which contains the windings.
-
-The top side of the end's larger outer side is flattend on eight sides.
-These are the mount points for eight bridge pieces which connect the ends, completing the field-capturing shielding loop.
-The gaps between the shielding bridges also allow for direct air cooling of the magnetic coils (see Section \ref{sec:mag_lens_thermal}).
-
-I chose drift tube outer diameter to be $D=$0.5in and a pole-piece gap $S$=0.1in, for an ratio $S/D=0.2$.
+Since the drift tube has an outer diameter $D=$0.5in and the pole-piece gap length is $S$=0.1in, the lens has a ratio $S/D=0.2$.
 By their table 8.2a this results in $\beta=0.0146$. Then, to create a focal length of 4in ($f/R=16$) table 8.13 gives a value for $k^2 \approx 0.06$.
@@
 
-To manage the divergence created by an RF cavity, two lenses are employed (one before and one after the cavity), forming a lens system of a similar effective focal length.
-For the purposes of design, one can make a reasonable approximation that the two lenses together may together provide the total number of needed amp-turns.
-As each has 100 turns, a current of only about 2A is needed to drive the two lenses in series.
-
+For the UEM system at UIC, two of these lenses are employed to manage the divergence created by an RF cavity and focus the pulses to a focal spot. 
+These lenses are positioned one before and one after the cavity, forming a lens system of a combined effective focal length.
+For the purposes of design, one can make a reasonable approximation that the two lenses together may together provide the total necessary number of amp-turns.
+As each lens has 100 turns, a current of only about 2A is needed to drive the two lenses in series.
 A picture of this system is shown in \ref{fig:maglens-pic}.
-In that picture the two lenses are shown, however a placeholder is inserted between them rather than the RF cavity.
+In that picture the two lenses are shown, however a placeholder is shown inserted between them rather than an actual RF cavity.
 
@@
 %TODO need reference for equation and \alpha_{Cu}
-Since the lenses have no cooling devices, it was prudent to characterize the thermal response to a moderate current load.
+Since the lenses have no active cooling system, it was prudent to characterize the thermal response to a moderate current load.
 Resistance of metals changes as a function of temperature by $ R = R_0 [ 1 + \alpha ( T - T_0 ) ] $, where $\alpha$ is a material contant; for copper, $\alpha = 3.9 * 10^{-3} \text{K}^{-1}$. 

### 2013-03-22 inc/hardware/deflector/deflector.tex
added some content, but also restructured some of the folders related to hardware
minor tweaks to deflector section

--- 2013-03-18 inc/hardware/deflector.tex
@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \section{Deflection Plates}
 
-In order to redirect the beam down the column, in conjunction with the accelerator alignment mechanisms, custom large area deflector plates were employed.
+In order to redirect the beam down the column, custom large area deflector plates were employed in conjunction with the gross accelerator alignment mechanism described earlier.
 Four plates are arranged on a hollow tube with a square profile, one on each side.
@@
 Using this result, when $L \sim d$ and $v_z = c/3$ a voltage of 1kV will deflect the beam approximately one degree.
-These plates therefore are used as a fine adjustment, as opposed to the coarse adjustment provided by aligning the acceleration gap.
+These plates are, therefore, useful as a fine beam adjustment, as contrast to the coarse adjustment provided by manually aligning the acceleration gap.
 
@@
     Each plate has a pin for connecting to voltage or ground.
-    Clearance drilled holes allow mounting onto the ``eV parts'' cage inside the chamber.
+    Clearance drilled holes allow mounting onto the ``eV Parts'' cage inside the chamber.
   }

### 2013-03-24 inc/design_considerations/considerations.tex
clean up the introductions to the hardware and consideration sections
began section on transverse emittance limitations

@@
 In the following chapters I will show how we are addressing these considerations for the prototype system at UIC.
-
-%Ultrafast Electron Microscopy (UEM) is a subclass of a more general concept, Dynamic Transmission Electron Microscopy (DTEM).
-%DTEM is a more general term for any electron microscope with higher time resolution, typically employing laser-stimulated photoemission.
 
@@
 
-\section{Generation Current Limitations}
+\section{Generation Current Limitations} \label{sec:childs_law}
 
@@
 \section{Beam Quality and Resolution Limit}
+
+A common figure of merit for electron beams is its normalized transverse emittance \cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, defined as 
+\begin{equation} \label{eq:emittance_definition}
+  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > } \,\text{,}
+\end{equation}
+where $m$ is the electron mass, $c$ is the speed of light in vacuum and $x$ and $p_x$ are the transverse position and momentum respectively.
+By Liouville's Theorem (Section \ref{sec:liouville}), we know that this quantity is conserved throught the propagation of the pulse, and therefore is related directly to the initial photoemission properties of the system.
+For a laser of the form of \ref{eq:laser_form} we have simply
+\begin{equation} \label{eq:emittance_definition_w}
+  \varepsilon_{\smallT} = \frac{w}{m c} \sqrt{ \left < p_x^2 \right > } \,\text{.}
+\end{equation}
+
+This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred.
+In Ref. \ref{berger_dc_2009}, we explore these limitations, eventually arriving at the following result.
+For a desired spatial resolution per pixel $\Delta X$ (related to the magnification) using a $N_p$ x $N_p$ detector, in the M$^{th}$ diffraction order, one may only resolve dynamics arising from a diffraction peak shift $\Delta a / a$ where the Bragg plane spacing is $a$ if
+\begin{equation}
+  \varepsilon_{\smallT} \le \frac{5 \hbar}{m c} N_p \, \Delta X \frac{M}{a} \left ( \frac{\Delta a}{a} \right ) \,\text{.}
+\end{equation}

### 2013-03-24 inc/design_considerations/considerations.tex
tweak the emittance section

@@
 This finally leads to the first consideration for UEM design mentioned in the introduction: for the clean pulsed emission of sufficient to create a useful image, the generated beam should have a HW1/eM width
-\begin{equation}
+\begin{equation} \label{eq:minimum_w}
   w \gtrsim \sqrt{\frac{N e d}{\pi \varepsilon_0 V_{DC}}} \,\text{.}
@@
 \begin{equation} \label{eq:emittance_definition_w}
-  \varepsilon_{\smallT} = \frac{w}{m c} \sqrt{ \left < p_x^2 \right > } \,\text{.}
+  \varepsilon_{\smallT} = \frac{w}{m c} \sqrt{ \frac{\left < p_x^2 \right >}{2} } \,\text{.}
 \end{equation}
@@
 This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred.
-In Ref. \ref{berger_dc_2009}, we explore these limitations, eventually arriving at the following result.
-For a desired spatial resolution per pixel $\Delta X$ (related to the magnification) using a $N_p$ x $N_p$ detector, in the M$^{th}$ diffraction order, one may only resolve dynamics arising from a diffraction peak shift $\Delta a / a$ where the Bragg plane spacing is $a$ if
+Since \ref{eq:minimum_w} specifies a lower bound on the emission spot size $w$, knowing a UEM will exhibit higher resolution by employing an electron source with the minimum possible transverse emittance directly implies that every effort must be made to reduce the initial transverse momentum spread.
+Though deriving an absolute limit on the allowable transverse emittance for imaging in a UEM would depend on many column-specific factors, since diffraction is expected to impose less stringent requirements on the beam than imaging, we may consider it when searching for an upper bound on transverse emittance.
+
+In Ref. \cite{berger_dc_2009}, we explore the impact of these limitations for an Ultrafast Electron Diffraction (UED) experiment.
+We find that for a desired spatial resolution per pixel $\Delta X$ (related to the magnification) using a $N_p$ x $N_p$ detector, one may only resolve a fractional change $\Delta a / a$ in a crystalline material's Bragg plane spacing $a$ in the M$^{th}$ diffraction order if
 \begin{equation}
@@
 \end{equation}
+However, because of lower limit that exists on $w$ from \ref{eq:minimum_w}, we can say that these requirements impose an upper limit on the initial transverse momentum spread
+\begin{equation}
+  \sqrt{ \left < p_x^2 \right > } = \sqrt{ \eta_{\smallT} (0) } \le \frac{5\sqrt{2} \hbar}{w} N_p \, \Delta X \frac{M}{a} \left ( \frac{\Delta a}{a} \right ) \,\text{.}
+\end{equation}

### 2013-03-25 inc/design_considerations/considerations.tex
rather large number of tweaks to the considerations section

@@
 
-In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effect that column elements have upon it.
+In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effects that column elements have upon it.
 However, microscopy is an experimental science, and designing a UEM is no less hands-on.
@@
 As a direct consequence of these considerations, it is our belief that a successful UEM project cannot simply stem from small modifications to the platform of a conventional electron microscope.
-In the following chapters I will show how we are addressing these considerations for the prototype system at UIC.
+In the following chapters I will show how we are addressing these considerations for the prototype system at UIC, both in column design (Chapter \ref{chap:prototype}) and photocathode engineering (Chapter \ref{chap:photocathode}).
 
@@
 
-In a conventional electron microscope the electrons are generated by one of several processes, often by heating the cathode or applying a large field to the cathode.
+In a conventional electron microscope the electrons are generated by one of several processes, often by heating the cathode or applying a large field to the cathode, or combinations thereof.
 Although these processes produce a very large number of electrons, far more than required by the Rose criterion, they are essentially non-directional.
@@
 In fact, for conventional electron microscopes without time resolution, the Rose criterion may be met for even very low beam current by simply allowing the image to acquire for a long enough time to collect sufficent electrons.
-Further, high beam current may be damaging to the sample over time, so often a lower beam current is preferred.
+Further, high beam current may be damaging to the sample over time, so a lower beam current is often preferred.
 
@@
 Although sample damage concerns may be neglected, destructive measurements will result in useful images, the use of a pinhole is likely to reduce the pulse charge below the useful level.
-This means that the designer of a DTEM must endevor to keep all of the created electrons.
+This means that the designer of a DTEM must endeavor to retain and utilize all of the generated electrons.
 
-It should be noted that for diffraction studies with a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse \cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.
+It should be noted that for diffraction studies performed in a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse \cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.
 
@@
 This steady-state law for the maximum current density that can be achieved is not valid in the regime of interest for UEM (and UED) where the use of picosecond and femtosecond photocathode drive laser pulses implies that $t_{gun}$ is usually much greater than the electron pulse duration $\tau_e$.
-In fact, though there is no clear definition in the literature or the community, we propose that this is useful delimination between term UEM and the more general term DTEM; reserving the term UEM for when the pulse can be completely contained inside the accelerator, i.e.
+Though there is no clear delineation in the literature or the community between moniker of Ultrafast Electron Microscopy (UEM) and the more general term Dynamic Transmission Electron Microcopy (DTEM), we propose that this is useful delimination, reserving the term UEM for those systems where the pulse can be completely contained inside the accelerator, i.e.
 $t_{gun} > \tau_e$.
+Pulses of this type are subject to a different ``short pulse analog'' of the Child-Langmuir limit.
+
 In this short pulse case, an equivalent diode approximation \cite{valfells_effects_2002} allows for the definition of a critical current density associated with virtual cathode formation in terms of the steady-state Child-Langmuir limit:
@@
 where $\chi = \tau_e / t_{gun}$.
-Thus, for the usual case in UEM, where $\chi \ll 1$, the critical current density is simply given by
+When $\chi \ll 1$, the usual case in UEM, the critical current density is simply given by
 \begin{equation}
@@
 \end{equation}
-For our example case of a 1k $\times$ 1k CCD, $N \ge 10^8$ electrons/pulse, we require $w \gtrsim $ 0.56 mm.
+For our example case above, to cleanly extract $N \ge 10^8$ electrons/pulse we require $w \gtrsim 0.56 \text{ mm}$.
+I will discuss the effect that accomodating a large pulse size has on designing an Ultrafast Electron Microscope in Chapter \ref{chap:prototype}.
 
@@
 
-This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred.
+This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred \cite{berger_dc_2009}.
 Since \ref{eq:minimum_w} specifies a lower bound on the emission spot size $w$, knowing a UEM will exhibit higher resolution by employing an electron source with the minimum possible transverse emittance directly implies that every effort must be made to reduce the initial transverse momentum spread.
@@
 \end{equation}
+
+To have $\Delta X = $1 nm spatial resolution, and be able to measure a 10\% change ($\Delta a / a = 0.1$) in a Bragg spacing $a \sim$ 1-2 \AA{} in the 1$^{st}$ diffraction order, the example DTEM from Section \ref{sec:childs_law}, equipped with a 1k x 1k CCD will require $\varepsilon_{\smallT} \lesssim$ 2 nm.
+For comparison, using the simplistic initial conditions in \ref{eq:summary}, exciting carriers from a HW1/eM $w = 0.56 \text{ mm}$ Gaussian spot-size on a Tantalum metal photocathode (work function $\Phi =$ 4.25 eV) using our laser (photon energy $\hbar \omega =$ 4.75 eV, further described in Section \ref{sec:laser}), for an excess photoemission energy $\Delta E = \hbar \omega - \Phi =$ 0.5 eV, the normalized transverse emittance is 226 nm.
+
+One might think to narrow this discrepancy by simply better matching the photocathode work function and the laser photon energy, reducing the excess photoemission energy $\Delta E$.
+Unfortunately, since the photoemission efficiency is proportional to $\Delta E^2$ \cite{shalaev_electron_1994}, though this would indeed reduce the transverse emittance, it would do so at the cost of a much reduced pulse charge.
+In Chapter \ref{chap:photocathode}, I will explore several more exotic photoemission processes in an attempt to reduce $\varepsilon_{\smallT}$ without this penalty.

### 2013-03-26 inc/model_results/lens_charge/lens_charge.pl
simplify the file structure related to the lens_charge plots

@@ -46,6 +46,8 @@ sub gen_sim {
   }
 }
 
+use Tie::Array::CSV;
+
 my @sets = (
   [ 6,  730e-12, 1 ],
   [ 60, 675e-13, 1 ],
@@ -53,6 +55,10 @@ my @sets = (
 
 foreach my $set ( @sets ) {
   my $lens = $set->[0];
+
+  my $filename = "lens_${lens}mm_" . ( $set->[2] ? 'prolate' : 'oblate' ) . '.dat';
+  tie my @out, 'Tie::Array::CSV', $filename, sep_char => ' ';
+  my $column = 0;
 
   say "$lens mm";
   my $sim = gen_sim( @$set );
@@ -64,10 +70,16 @@ foreach my $set ( @sets ) {
     my $z  = $result->slice('(1),') / ( $lens * 1e-3 ) - 1;
     my $w = sqrt( 2 * $result->slice('(3),') ) * 1000;
 
-    wcols $z, $w, "lens_${lens}mm_n$_.dat";
+    #wcols $z, $w, "lens_${lens}mm_n$_.dat";
+    my $i = 0;
+    $out[$i++][$column] = $_ for $z->list;
+    $i = 0;
+    $out[$i++][$column+1] = $_ for $w->list;
+    $column += 2;
 
     my $min_ind = $w->minimum_ind;
     say "\t" . $w->at($min_ind) . 'mm at z = ' . $z->at($min_ind) . 'f';
+
   }
 
 }

### 2013-03-26 inc/model_results/lens_charge/lens_charge.tex
added some content, but also restructured some of the folders related to hardware
lens_charge is now all four permutations of long/short oblate/prolate

@@
-\subsection{Space-charge Effects in Magnetic Lenses}
+\subsection{Space-charge Effects in Magnetic Lenses} \label{sec:mag_lens_charge}
 
@@
 When focusing the pulse it can be shown that prolate pulses will also have markedly better performance.
-Therefore, in the sake of simplicity, only the prolate case will be demonstrated here.
-For more information see reference \cite{berger_semi-analytic_2010}.
 
-\begin{figure}
-  \centering
-  \centerline{
-    \subfloat[][] {
-      \label{fig:focus_lens_charge_long}
-      \input{lens_60mm}
-    }
-    \subfloat[][] { 
-      \label{fig:focus_lens_charge_short}
-      \input{lens_6mm}
-    }
-  }
-  \caption[AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$]{
-    AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$.
-    Pulses begin at the lens' front focal plane with HW1/eM width $ 500 \mu \text{m}$ and duration 500 fs at 20 kV velocity ($c/3$) for an ellipticiy $ \xi ( 0 ) = 0.1 $ (prolate).
-    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth, \subref{fig:focus_lens_charge_long} $f = $ 60mm and \subref{fig:focus_lens_charge_short} $ f = $ 6.0mm.
-    A logarithmic scale is used for comparative clarity.
-  }
-  \label{fig:focus_lens_charge}
-\end{figure}
+\input{lens_charge_plot}
 

### 2013-03-26 inc/model_results/lens_charge/lens_charge_plot.tex
lens_charge is now all four permutations of long/short oblate/prolate

%TODO fix axes to be more consistent
%TODO add N=1e8 for prolate?

\begin{figure}
  \centering
  \centerline{
    \subfloat[][] {
      \label{fig:focus_long_prolate}
      \begin{tikzpicture}
        \input{lens_60mm_prolate_data}
        \node at (5.9,1.65) {$N=10^5$};
        \node at (4,3) {$N=10^6$};
        \node at (3,4) {$N=10^7$};
      \end{tikzpicture}
    }
    \subfloat[][] {
      \label{fig:focus_long_oblate}
      \begin{tikzpicture}
        \input{lens_60mm_oblate_data}
        \node at (4.2,2.3) {$N=10^5$};
        \node at (3.3,3.5) {$N=10^6$};
      \end{tikzpicture}
    }
  }
  \centerline{
    \subfloat[][] { 
      \label{fig:focus_short_prolate}
      \begin{tikzpicture}
        \input{lens_6mm_prolate_data}
        \node at (6.5,3) {$N=10^7$};
      \end{tikzpicture}
    }
    \subfloat[][] { 
      \label{fig:focus_short_oblate}
      \begin{tikzpicture}
        \input{lens_6mm_oblate_data}
        \node at (6.9,2.6) {$N=10^7$};
        \node at (6,2) {$N=10^6$};
      \end{tikzpicture}
    }
  }
  \caption[AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$]{
    AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$.
    All plot's pulses begin at the lens' front focal plane at 20 kV velocity ($c/3$) and have the same volume.
    The left plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are prolate shaped, initially HW1/eM width $ 500 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 0.1 $.
    The right plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are oblate shaped, initially HW1/eM width $ 107.7 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 10 $.
    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth.
    The top plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_long_oblate}) have a longer focal length $f = $ 60mm compared to the bottom plots (\subref{fig:focus_short_prolate} and \subref{fig:focus_short_oblate}) $ f = $ 6.0mm.
    A logarithmic scale is used for comparative clarity.
    Clearly the best performance is achieved for shorter focal lengths and prolate pulses even at higher initial charge densities.
  }
  \label{fig:focus_lens_charge}
\end{figure}

### 2013-03-26 inc/model_results/lens_charge/lens_charge.pl

@@ -19,8 +19,8 @@ sub gen_sim {
     my $pulse = Pulse->new(
       number => $num,
       velocity => vc / 3,
-      initial_width  => $is_prolate ? '500 um' : '50 um',
-      initial_length => $is_prolate ? '50 um'  : '500 um',
+      initial_width  => $is_prolate ? '500 um' : '107.7 um',
+      initial_length => $is_prolate ? '50 um'  : '1077 um',
       excess_photoemission_energy => '0.5 eV',
     );
 
@@ -48,13 +48,18 @@ sub gen_sim {
 
 use Tie::Array::CSV;
 
+my @charges = qw/ 1e0 1e4 1e5 1e6 1e7 /;
+
 my @sets = (
-  [ 6,  730e-12, 1 ],
-  [ 60, 675e-13, 1 ],
+  [ 6,  730e-12, 1, \@charges ],
+  [ 60, 675e-13, 1, \@charges ],
+  [ 6,  730e-12, 0, \@charges ],
+  [ 60, 920e-13, 0, [ @charges[0..$#charges-1] ] ],
 );
 
 foreach my $set ( @sets ) {
   my $lens = $set->[0];
+  my $charges = pop @$set;
 
   my $filename = "lens_${lens}mm_" . ( $set->[2] ? 'prolate' : 'oblate' ) . '.dat';
   tie my @out, 'Tie::Array::CSV', $filename, sep_char => ' ';
@@ -63,7 +68,7 @@ foreach my $set ( @sets ) {
   say "$lens mm";
   my $sim = gen_sim( @$set );
 
-  for ( qw/ 1e0 1e4 1e5 1e6 1e7 / ) {
+  for ( @$charges ) {
     print "\t$_:";
 
     my $result = $sim->($_);

### 2013-03-27 inc/model_results/lens_charge/lens_charge.tex
fix oblate/prolate gaff (ooops)

@@
 
-In the preceding sections it has already been established that prolate eccentricities are preferred for UEM in general.
-When focusing the pulse it can be shown that prolate pulses will also have markedly better performance.
+In the preceding sections it has already been established that oblate eccentricities are preferred for UEM in general.
+When focusing the pulse it can be shown that oblate pulses will also have markedly better performance.
 
@@
 
-\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for prolate pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_long}) and a `short' focal length of 6.0mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_short}).
+\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for oblate pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_long}) and a `short' focal length of 6.0mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_short}).
 For both lens strengths, the deleterious effects of space-charge are evident as the number of electrons/pulse $ N $ is increased.

### 2013-03-27 inc/model_results/lens_charge/lens_charge_plot.tex
fix oblate/prolate gaff (ooops)

@@
 %TODO fix axes to be more consistent
-%TODO add N=1e8 for prolate?
+%TODO add N=1e8 for oblate?
 
@@
     \subfloat[][] {
-      \label{fig:focus_long_prolate}
+      \label{fig:focus_long_oblate}
       \begin{tikzpicture}
-        \input{lens_60mm_prolate_data}
+        \input{lens_60mm_oblate_data}
         \node at (5.9,1.65) {$N=10^5$};
@@
     \subfloat[][] {
-      \label{fig:focus_long_oblate}
+      \label{fig:focus_long_prolate}
       \begin{tikzpicture}
-        \input{lens_60mm_oblate_data}
+        \input{lens_60mm_prolate_data}
         \node at (4.2,2.3) {$N=10^5$};
@@
     \subfloat[][] { 
-      \label{fig:focus_short_prolate}
+      \label{fig:focus_short_oblate}
       \begin{tikzpicture}
-        \input{lens_6mm_prolate_data}
+        \input{lens_6mm_oblate_data}
         \node at (6.5,3) {$N=10^7$};
@@
     \subfloat[][] { 
-      \label{fig:focus_short_oblate}
+      \label{fig:focus_short_prolate}
       \begin{tikzpicture}
-        \input{lens_6mm_oblate_data}
+        \input{lens_6mm_prolate_data}
         \node at (6.9,2.6) {$N=10^7$};
@@
     All plot's pulses begin at the lens' front focal plane at 20 kV velocity ($c/3$) and have the same volume.
-    The left plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are prolate shaped, initially HW1/eM width $ 500 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 0.1 $.
-    The right plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are oblate shaped, initially HW1/eM width $ 107.7 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 10 $.
+    The left plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are oblate shaped, initially HW1/eM width $ 500 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 0.1 $.
+    The right plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are prolate shaped, initially HW1/eM width $ 107.7 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 10 $.
     Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth.
-    The top plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_long_oblate}) have a longer focal length $f = $ 60mm compared to the bottom plots (\subref{fig:focus_short_prolate} and \subref{fig:focus_short_oblate}) $ f = $ 6.0mm.
+    The top plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_long_prolate}) have a longer focal length $f = $ 60mm compared to the bottom plots (\subref{fig:focus_short_oblate} and \subref{fig:focus_short_prolate}) $ f = $ 6.0mm.
     A logarithmic scale is used for comparative clarity.
-    Clearly the best performance is achieved for shorter focal lengths and prolate pulses even at higher initial charge densities.
+    Clearly the best performance is achieved for shorter focal lengths and oblate pulses even at higher initial charge densities.
   }

### 2013-03-27 inc/model_results/shapes/shapes.tex
fix oblate/prolate gaff (ooops)

@@
     Transverse (solid) and longitudinal (dashed) dynamics of three example pulses for three length/width ratios ($\xi$). 
-    The red lines represent a prolate pulse (a ``pancake'') with $\xi=1/3$, green lines for a spherical pulse with $\xi=1$, and blue for an oblate pulse (a ``cigar'') having $\xi=3$.
+    The red lines represent a oblate pulse (a ``pancake'') with $\xi=1/3$, green lines for a spherical pulse with $\xi=1$, and blue for an prolate pulse (a ``cigar'') having $\xi=3$.
     The pulses shown in \ref{fig:compare_shape_N1} all have total charge $N=1$, while in \ref{fig:compare_shape_N1e6} the pulses have $N=10^6$.
@@
 %TODO perhaps all of this should be moved back or even cut?
-I have shown that a prolate spheroid pulse shape is preferred for UEM, both for generation in SECTION %TODO link to Child law discussion
+I have shown that a oblate spheroid pulse shape is preferred for UEM, both for generation in SECTION %TODO link to Child law discussion
 and in this section for propagation.
@@
 that electron pulses shorter than the laser pulse duration are likely occurances.
-Henceforth I shall only consider prolate pulses when a choice of initial pulse shape is possible.
+Henceforth I shall only consider oblate pulses when a choice of initial pulse shape is possible.

### 2013-03-27 inc/model_results/spacecharge/spacecharge.tex
fix oblate/prolate gaff (ooops)

@@
 The simulations, clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
-In the absence of space-charge effects, the presented prolate pulse shaped pulses should be maintained as such during unforced propagation because the electron pulse divergence and its temporal broadening rate are both proportional to $\Delta E^{1/2}$.
+In the absence of space-charge effects, the presented oblate shaped pulses should be maintained as such during unforced propagation because the electron pulse divergence and its temporal broadening rate are both proportional to $\Delta E^{1/2}$.
 

### 2013-03-27 inc/model_results/lens_charge/lens_charge.pl

@@ -10,7 +10,7 @@ use Physics::UEMColumn::Auxiliary ':constants';
 use PDL;
 
 sub gen_sim {
-  my ($lens_position, $strength, $is_prolate) = @_;
+  my ($lens_position, $strength, $is_oblate) = @_;
   my $column_length = 4 * $lens_position;
 
   return sub {
@@ -19,8 +19,8 @@ sub gen_sim {
     my $pulse = Pulse->new(
       number => $num,
       velocity => vc / 3,
-      initial_width  => $is_prolate ? '500 um' : '107.7 um',
-      initial_length => $is_prolate ? '50 um'  : '1077 um',
+      initial_width  => $is_oblate ? '500 um' : '107.7 um',
+      initial_length => $is_oblate ? '50 um'  : '1077 um',
       excess_photoemission_energy => '0.5 eV',
     );
 
@@ -61,7 +61,7 @@ foreach my $set ( @sets ) {
   my $lens = $set->[0];
   my $charges = pop @$set;
 
-  my $filename = "lens_${lens}mm_" . ( $set->[2] ? 'prolate' : 'oblate' ) . '.dat';
+  my $filename = "lens_${lens}mm_" . ( $set->[2] ? 'oblate' : 'prolate' ) . '.dat';
   tie my @out, 'Tie::Array::CSV', $filename, sep_char => ' ';
   my $column = 0;
 

### 2013-03-27 inc/model_results/shapes/shapes.pl

@@ -14,7 +14,7 @@ my $vol = (1e-3)**3;
 
 my $for_xi = gen_xi_series( $n, $vol, $dE, $length );
 
-for my $spec ( [ prolate => 1/3 ], [ sphere => 1.0001 ], [ oblate => 3 ] ) {
+for my $spec ( [ oblate => 1/3 ], [ sphere => 1.0001 ], [ prolate => 3 ] ) {
   my ( $shape, $xi ) = @$spec;
   my $file = "shape_${shape}_N$n.dat";
   open my $fh, '>', $file or die "Cannot open $file: $!\n";

### 2013-03-27 inc/model_results/lens_charge/lens_charge.tex
tweaks for lens_charge section

@@
 
-\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for oblate pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_long}) and a `short' focal length of 6.0mm (\ref{fig:focus_lens_charge}\subref{fig:focus_lens_charge_short}).
+\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for intially oblate ($ \xi ( 0 ) = 0.1 $) or prolate ($ \xi ( 0 ) = 10 $) pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm and a `short' focal length of 6.0mm.
 For both lens strengths, the deleterious effects of space-charge are evident as the number of electrons/pulse $ N $ is increased.
@@
 A consequent further effect is a shift in the focal position to $ z > f $.
-%A more subtle effect is that disk-like pulses are more readily focused than cigar-like pulses --- the rate of transverse pulse broadening due to space-charge effects already being intrinsically greater in the latter (\Fig \ref{fig:compare}).
+A more subtle effect is that disk-like pulses are more readily focused than cigar-like pulses --- the rate of transverse pulse broadening due to space-charge effects already being intrinsically greater in the latter (\ref{fig:compare_shape}).
 

### 2013-03-27 inc/model_results/lens_charge/lens_charge_plot.tex
tweaks for lens_charge section

@@
     All plot's pulses begin at the lens' front focal plane at 20 kV velocity ($c/3$) and have the same volume.
-    The left plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are oblate shaped, initially HW1/eM width $ 500 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 0.1 $.
-    The right plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are prolate shaped, initially HW1/eM width $ 107.7 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 10 $.
+    The left plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are initially oblate shaped, HW1/eM width $ 500 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 0.1 $.
+    The right plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are initially prolate shaped, HW1/eM width $ 107.7 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 10 $.
     Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth.

### 2013-03-27 inc/model_results/shapes/shapes.tex
changed disk-like and cigar-like in several places in shapes section

@@
 The initial `space-time' eccentricity \ref{eq:define_xi}, $\xi ( 0 ) = v_{\smallzero} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
-\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and `cigar-like' ($ \sigma_{ z } ( 0 ) > \sigma_{ \smallT } ( 0 ) $) pulses.
+\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), oblate or `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and prolate or `cigar-like' ($ \sigma_{ z } ( 0 ) > \sigma_{ \smallT } ( 0 ) $) pulses.
 %TODO rewrite using w?
@@
 In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands approximately twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp ($ \gamma_{\alpha} = 0 $).
-Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both disk- and cigar-like pulses are initially driven towards the spherical regime.
+Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both oblate and prolate pulses are initially driven towards the spherical regime.
 In all cases, this intrinsic electron pulse broadening due to non-zero $ \eta_{\alpha} $ generates a momentum chirp across the electron bunch; in the longitudinal dimension the faster electrons photoemitted with larger excess energy leading the slower electrons.
@@
 %TODO the following sentance appears not to be true, still perhaps should include something like it
-%In fact, both the $ N = 10^{6} $ disk- and cigar-like pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
+%In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
 The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.
@@
 In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers,\cite{lobastov_four-dimensional_2005} deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
-Even for nanosecond DTEM (the extreme cigar-like pulse ($\xi \gg 1 $)), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
-In this thesis, however, I will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the disk-like regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}
+Even for nanosecond DTEM (the extreme prolate pulse ($\xi \gg 1 $)), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
+In this thesis, however, I will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}
 

### 2013-03-27 inc/model_results/shapes/shapes.tex
tweaks

@@
 In this thesis, however, I will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}
-
-%TODO perhaps all of this should be moved back or even cut?
-I have shown that a oblate spheroid pulse shape is preferred for UEM, both for generation in SECTION %TODO link to Child law discussion
-and in this section for propagation.
-Further I have shown in %TODO link to section where I show that initial conditions create shorter pulses than laser
-that electron pulses shorter than the laser pulse duration are likely occurances.
-Henceforth I shall only consider oblate pulses when a choice of initial pulse shape is possible.

### 2013-03-27 inc/model_results/spacecharge/spacecharge.tex
tweaks

@@
 These naive conditions are needed to isolate the behavior of a propagating pulse from its creation environment.
-The simulations, clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
+The simulations clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
 In the absence of space-charge effects, the presented oblate shaped pulses should be maintained as such during unforced propagation because the electron pulse divergence and its temporal broadening rate are both proportional to $\Delta E^{1/2}$.

### 2013-03-28 inc/model_results/compound/compound.tex
added compound element section (heading) and demonstration figure

--- 2013-03-27 inc/model_results/spacecharge/spacecharge.tex
@@
-\subsection{Effect of Space Charge} \label{sec:free_spacecharge}
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
 
-A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
-Much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements.
-Simplistically stated, at low pulse charge densities (low number of electrons/pulse), the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
-On the other hand, space-charge effects are expected to dominate at large pulse charge densities.
+\subsection{Compound Element Systems} \label{sec:compound}
 
@@
   \begin{tikzpicture}
-    \input{spacecharge_data_noacc}
+    \input{compound_data}
   \end{tikzpicture}
-  \caption[Simplistic free-space pulse evolution vs charge density]{
-    The final HW1/eM width (solid) and length (dashed) of pulses which are intially similar except for the number of electrons in the pulse.
-    The pulse is initially 1 mm (HW1/eM) width and 1 ps in duration travelling at $c/3$ with 0.5eV excess photoemission energy.
-    It then traverses a 15cm drift region upon which the final length and width are recorded.
+  \caption[Example model of a compound system of electron-optical elements]{
+    Example model of a compound system of electron-optical elements.
   }
-  \label{fig:spacecharge_noacc}
+  \label{fig:compound}
 \end{figure}
-
-This trend is clearly evident in \ref{fig:spacecharge_noacc} which presents the predicted HW1/eM pulse width $\sqrt{2 \sigma_{\smallT}}$ (solid line) and longitudinal length $\sqrt{2 \sigma_{z}}$ (dashed line) as a function of the number of electrons in the pulse after a propagation distance of 15cm.
-The pulses are not accelerated but rather start with a velocity of $c/3$ (where $c$ is the speed of light in a vacuum), initial width of 1 mm, HW1/eM pulse duration 1 ps, have the initial momentum chirps $\gamma_{\smallT} = \gamma{z} = 0 $ and momentum uncertainties related to an excess photoemission energy of $\Delta E = $ 0.5 eV.
-These naive conditions are needed to isolate the behavior of a propagating pulse from its creation environment.
-The simulations clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
-In the absence of space-charge effects, the presented oblate shaped pulses should be maintained as such during unforced propagation because the electron pulse divergence and its temporal broadening rate are both proportional to $\Delta E^{1/2}$.
-
-%TODO highlight the following in the experimental section
-%In other words, the spatial and temporal dispersive effects due to the nonzero value of $\Delta E$ are expected to dominate the space-time dynamics of the electron pulse during its propagation
-%in our investigations because less than 10 5 electrons/pulse
-%are generated.
-
-As the number of electrons/pulse increases beyond $10^6$, both the pulse length and diameter rapidly increase.
-The model predicts that space-charge effects produce a spherical charge distribution of approximately five times larger width at $10^8$ electrons/pulse for these conditions.
-As noted in Section \ref{sec:initial_shapes}, pulses under the influence of the internal space-charge field expand preferentially in the dimension that was smallest initially; this causes the pulse to initially trend towards being spherical.
-
-\ref{fig:spacecharge_noacc} is specifically created to isolate the behavior of and ``ideal'' pulse which is independent of its creation environment.
-Of course in a more realistic system, the electron pulse is generated from a photocathode and then accelerated.
-These forces play an interesting role in the system.
-
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \input{spacecharge_data_acc}
-  \end{tikzpicture}
-  \caption[More realistic free-space pulse evolution vs charge density]{
-    The final HW1/eM width (solid) and length (dashed) of pulses which are intially similar except for the number of electrons in the pulse.
-    The pulse is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy.
-    It then traverses a 15cm column, a 20mm acceleration region followed by drift region.
-  }
-  \label{fig:spacecharge_acc}
-\end{figure}
-
-Consider now a simulation comparable to the one presented above, but which uses the initial conditions presented in \ref{eq:summary} and accelerated by an electron gun as modelled in Section \ref{sec:gun_model} which is 20mm in length.
-The results of this simulation are presented in \ref{fig:spacecharge_acc}.
-It is immediately obvious that the final transverse width is larger than in the ideal case presented in \ref{fig:spacecharge_noacc}, even in the low charge-density regime; this is primarily due to the additional divergence imparted by the negative lensing at the anode (see Section \ref{sec:gun_model}).
-One can see that though the pulse size does increase beyond a certain total pulse charge, and while that trend appears to start at a higher charge-density, due to the complexity of the pulse dynamics it is rather difficult to compare the exact charge-density from this plot.
-What can be seen is that the transverse size increases more sharply with pulse charge at higher levels.
-Since the pulse is already diverging in the transverse direction, this is compounded by the increased pulse charge.
-This divergence then acts to lower the overall charge-density more quickly, thus the longitudinal pulse length is not affected as strongly as was seen in \ref{fig:spacecharge_noacc}.

### 2013-03-28 inc/model_results/compound/compound.pl

@@ -1,0 +1,111 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+use Physics::UEMColumn alias => ':standard';
+
+use PDL;
+use PDL::Graphics::Prima::Simple [700,500];
+
+my $laser = Laser->new(
+  width    => '100 um',
+  duration => '0.1 ps',
+  energy   => '4.75 eV',
+);
+
+my $acc = DCAccelerator->new(
+  length  => '20 mm',
+  voltage => '20 kilovolts',
+);
+
+my $column = Column->new(
+  length       => '350 mm', 
+  laser        => $laser,
+  accelerator  => $acc,
+  photocathode => Photocathode->new( work_function => '4.25 eV' ), # Ta
+);
+
+my $sim = Physics::UEMColumn->new(
+  column => $column,
+  number => 1e6,
+  steps  => 200,
+  solver_opts => {
+    h_max => 5e-13,
+    h_init => 5e-14,
+  },
+);
+
+my $z_rf       = 20; #cm
+my $l_mag_lens = '1 in';
+my $cooke_sep  = 5; #cm
+my $str_mag    = 43e-13;
+
+my $lens1 = MagneticLens->new(
+  location => ($z_rf - $cooke_sep) . 'cm',
+  length   => $l_mag_lens,
+  strength => $str_mag,
+);
+my $lens2 = MagneticLens->new(
+  location => ($z_rf + $cooke_sep) . 'cm',
+  length   => $l_mag_lens,
+  strength => $str_mag,
+);
+$sim->add_element($lens1);
+$sim->add_element($lens2);
+
+my $rf_cav = RFCavity->new(
+  location  => $z_rf . 'cm',
+  length    => '2 cm',
+  strength  => '335 kilovolts / m',
+  frequency => '3 gigahertz',
+);
+$sim->add_element($rf_cav);
+
+my $result = pdl( $sim->propagate );
+
+my $z  = $result->slice('(1),');
+my $st = $result->slice('(3),');
+my $sz = $result->slice('(4),');
+
+plot(
+  -st => ds::Pair( 
+    $z, sqrt( $st / maximum($st) ),
+    colors => pdl(255,0,0)->rgb_to_color,
+    plotType => ppair::Lines,
+    lineWidths => 3,
+  ),
+  -sz => ds::Pair( 
+    $z, sqrt( $sz / maximum($sz) ),
[…]

### 2013-03-28 inc/model_results/compound/compound.tex
Ported all of the compound lens system from JAP paper. Still probably needs cleanup and tweaking

@@
 
-\subsection{Compound Element Systems} \label{sec:compound}
+\subsection{UEM Column Example} \label{sec:compound}
+
+Through the use of \ref{eq:reg_of_influence} to specify the position and influence of electron optical components such as magnetic lenses and RF cavities, the extended AG model can be employed to design and assess a full UEM column from pulse generation to sample stage.
+The design of such a compound system is, of course, a cornerstone of modern electron microscope development, although space-charge effects are generally not included, since, on average, there is usually no more than one electron in the microscope column at a time.
+However, for a nanosecond DTEM,\cite{lagrange_nanosecond_2008,reed_evolution_2009} the microscope column must be designed to compensate for transverse ($\smallT$) space-charge effects, which dominate the propagation dynamics for the prolate electron pulses employed in such instruments (see \ref{fig:compare_shape}).
+For the ultrashort (i.e., picosecond and femtosecond) electron pulse durations used in UEMs\cite{park_direct_2009} and UED experiments,\cite{oudheusden_electron_2007} the longitudinal ($z$) pulse dynamics due to both dispersion and space-charge effects must also be considered (as indicated in \ref{fig:compare_shape}).
 
@@
     Example model of a compound system of electron-optical elements.
+    The length (green) of the pulse is compressed to a ``focus'' 10 cm after the RF cavity at 20 cm.
+    The width (red) is affected by two magnetic lenses (5 cm before and after the RF cavity) as well as by the cavity itself.
+    The lenses are tuned to produce a simultaneous focus with the pulse length, creating a small spatial and temporal probe should this focal point represent a sample holder.
   }
@@
 \end{figure}
+
+%Mask starts at 20.67mm
+%w after anode: 579.33um
+%Min Width: 29.50um at 301.49mm
+%Min Length: 22.58um (269.30fs) at 299.45mm
+
+\ref{fig:compound} shows an example of a compound element system for the delivery of an ultrashort electron pulse to the specimen area; in this case, optimized for $ N = 10^{ 6 } $ --- the number of electrons per pulse required for single shot diffraction measurements.\cite{armstrong_practical_2007}
+The displayed electron pulse dynamics are for a $\tau = $ 100fs electron pulse with $w = $ 0.1mm generated from a laser driven Ta photocathode, with $\Delta E = $ 0.5eV using $ \hbar \omega = $ 4.75eV photons, then accelerated by 20 kV to a velocity of $8.4 \times 10^{7} \text{m}/\text{s}$.
+In this simple system, an optimum RF cavity ($ \Omega = \pi v_{{ \scriptscriptstyle 0}} / d $) is surrounded by two magnetic lenses (positioned a realistic 5cm to either side of the RF cavity) that compensate for the electron beam divergence due to the non-zero $ \eta_{\smallT} (0) $ and the negative spatial lensing of the RF pulse compression cavity.
+For the transverse dimension, this arrangement of electron optical elements is similar to the Cooke triplet in optics.
+The magnetic lenses are assumed to be identical (but oppositely wound to allow for the compensation of rotational effects in the magnetic focusing through the use of oppositely directed axial magnetic fields).
+The three lensing elements are positioned at a distance of 15cm from the cathode to the front of the first magnetic lens, adjusted to space-time image the electron pulse 5cm behind the second magnetic lens.
+The dependencies of the the spatial variances, $\sigma_{\alpha}$, shown in \ref{fig:compound} as the electron pulse propagates through the multi-element system clearly illustrate the effect of the magnetic lenses and RF cavity in the transverse ($\smallT$) and longitudinal ($z$) dimensions.
+As the image-to-object distance ratio is less than unity in this case, both the transverse and longitudinal imaged pulse dimensions are reduced relative to their initial values. %TODO comment futher
+
+As seen in Section \ref{sec:free_spacecharge}, with increasing pulse charge, $ N e $, space-charge effects in general will act to degrade the space-time imaging by increasing the three-dimensional pulse size delivered to the reference plane positioned 5cm behind the second magnetic lens.
+At $ N = 10^{6} $, space-charge effects moderate the image demagnification due to the $\sim$2 object-to-image distance ratio in \ref{fig:compound}. 
+In other words, for the displayed compound element system, which could be employed in UED experiments,\cite{oudheusden_electron_2007} the initial space-time resolution associated with the $\sim$0.1mm spot size and $\sim$100fs pulse duration can be maintained (or bettered) for $N \lesssim 10^{6} $.
+For a UEM, on the other hand, it may be necessary to separate longitudinally the space and time images produced by the compound element column.
+This is because the inclusion of the microscope's objective lens will likely require the pulse's time focus to be at the objective's back focal plane (i.e., in the specimen) while the column's spatial image may need to be located at the objective's front focal plane (i.e., entrance aperture) for efficient beam transfer with minimal spherical aberration.
+Fortunately, a compound component system of the type shown in \ref{fig:compound} with a RF cavity can meet this challenge since the longitudinal positions of the two foci can be manipulated independently.

### 2013-03-29 inc/model_results/introduction.tex
Expand the model_result introduction to include a discussion of employed initial conditions.
Also moves the introduction to its own file.

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

In Chapters \ref{chap:previous_models} and \ref{chap:extension}, I have introduced the extended Analytic Gaussian model.
In the following chapter I use this model to make general observations about pulse propagation for Ultrafast Electron Microscopy.

These examples will each use one of two initialization schemes.
The first represents an idealized pulse which is not generated by photocathode and then accelerated, but rather springs into existence at its initial position with its velocity already established.
This type of pulse is useful to demonstrate general pulse dynamics or the ideal influence of electron-optical elements.
Simulations using pulses of this type will provide an initial HW1/eM length and width, a velocity, and an excess photoemission energy.
The length and width will be used directly to calculate initial spatial variances $\sigma_{\alpha}$.
Both chirp terms, $\gamma_{\alpha}$, will be initially zero, but since pulses must still be subject to Liouville's Theorem (discussed in Section \ref{sec:liouville}), \ref{eq:summary} will still be used to calculate meaningful local momentum widths $\eta_{\alpha}$.
This idealization only applies to the initialization, therefore these pulses will still be subject to internal space-charge repulsion which might cause pulse expansion before or during the demonstration, especially for high charge densities.

Other examples in this section will require or are markedly different when using a more realistic pulse, one resulting from photoemission and acceleration.
Simulations using pulses of this type will provide and HW1/eM width and duration of the laser, an acceleration voltage, and an excess photoemission energy.
In these cases, all initial parameters will be calculated from \ref{eq:summary}.
The acceleration process, and more specifically the anode (see Section \ref{sec:gun_model}), will have an impact on the beam dynamics which will be evident in the pulse's evolution.

### 2013-03-29 main.tex
added some text the the hardware introduction
added some content, but also restructured some of the folders related to hardware
added new UEM design considerations chapter
rename some sections remove "experimental results" section (unused)
clean up the introductions to the hardware and consideration sections
rather large number of tweaks to the considerations section
added compound element section (heading) and demonstration figure
Expand the model_result introduction to include a discussion of employed initial conditions.
Also moves the introduction to its own file.

@@
 
-\chapter{Model Results, Predictions and Implications}
+\chapter{Model Results, Predictions and Implications} \label{chap:model_results}
 
-In Chapters \ref{chap:previous_models} and \ref{chap:extension}, I have introduced the extended Analytic Gaussian model.
-In the following chapter I use this model to make general observations about pulse propagation for Ultrafast Electron Microscopy.
+  \subimport{inc/model_results/}{introduction}
 
@@
 
-\chapter{Prototype Instrument}
+  \subimport{inc/model_results/compound/}{compound}
+
+\chapter{UEM Design Considerations} \label{chap:considerations}
+
+  \subimport{inc/design_considerations/}{considerations}
+
+\chapter{Prototype Instrument} \label{chap:prototype}
 
@@
 
-  \subimport{inc/laser/}{laser}
+  \subimport{inc/hardware/laser/}{laser}
 
@@
 
-  \subimport{inc/magnetic_lenses/}{maglens}
+  \subimport{inc/hardware/magnetic_lenses/}{maglens}
 
-  \subimport{inc/hardware/}{deflector}
+  \subimport{inc/hardware/deflector/}{deflector}
 
-\chapter{Photocathode Engineering}
+\chapter{Photocathode Engineering} \label{chap:photocathode}
 
@@
   \subimport{inc/photocathode/este/}{este}
-
-\chapter{Experimental Results}
 

### 2013-03-30 inc/model_results/shapes/shapes.tex
added cc notices to the tops of several files
flattened model_results section headings, removing free-space versus forced
began correction on model_results sections

@@
-\subsection{Trends for Initial Shapes} \label{sec:initial_shapes}
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
 
-The initial `space-time' eccentricity \ref{eq:define_xi}, $\xi ( 0 ) = v_{\smallzero} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
-\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), oblate or `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and prolate or `cigar-like' ($ \sigma_{ z } ( 0 ) > \sigma_{ \smallT } ( 0 ) $) pulses.
+\section{Effect of Pulse Eccentricity on Dynamics} \label{sec:initial_shapes}
+
+The initial `space-time' eccentricity (defined in \ref{eq:define_xi}), $\xi ( 0 ) = v_{\smallzero} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
+\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), oblate or `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and prolate or `cigar-like' ($ \sigma_{ \smallT } ( 0 ) < \sigma_{ z } ( 0 ) $) pulses.
 %TODO rewrite using w?
-For simplicity, and unlike most simulations in this paper, the pulses in this simulation are not calculated from laser and photocathode parameters, but are rather defined directly.
-%TODO should I remove this? Should I just specify the creation types on each instance?
-The initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ while in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
-Since the initial pulse volume is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} ($ N = 10^{6} $ electrons in the bunch) than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} ($ N = 10^{0} $).
+This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
+Pulses used in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} have just one electron per pulse (i.e.
+$ N = 10^{0} $), while \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} have a much higher $ N = 10^{6} $ electrons in the bunch.
+Since the initial pulse volume is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1}.
 The pulse velocity is $ v_{\smallzero} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.
@@
     The red lines represent a oblate pulse (a ``pancake'') with $\xi=1/3$, green lines for a spherical pulse with $\xi=1$, and blue for an prolate pulse (a ``cigar'') having $\xi=3$.
-    The pulses shown in \ref{fig:compare_shape_N1} all have total charge $N=1$, while in \ref{fig:compare_shape_N1e6} the pulses have $N=10^6$.
+    The pulses shown in \subref{fig:compare_shape_N1} all have total charge $N=1$, while in \subref{fig:compare_shape_N1e6} the pulses have $N=10^6$.
   }

### 2013-03-30 inc/photocathode/setup/fig1.tex
remove (essentially) unused \inputdata command, which was meant to ignore loading tikz data

\begin{tikzpicture}
  \fill [blue!20] 
    (1.3,0.95)
    -- ++(0.7,0)
    -- ++(0,7.5)
    -- ++(-0.7,0)
      node [black, above, pos=0.5, align=center] {Acceleration\\Region}
    -- cycle
  ;
  \fill [blue!20]
    (4.4,0.95)
    -- ++(2,0)
    -- ++(0,7.5)
    -- ++(-2,0)
      node [black, pos=0.5, above, align=center] {Magnetic\\Lenses}
    -- cycle
  ;
  \fill [blue!20] 
    (10.5,0.95)
    -- ++(0.7,0)
    -- ++(0,7.5)
    -- ++(-0.7,0)
      node [black, above, pos=0.5] {Detector}
    -- cycle
  ;
  \input{fig1data}
\end{tikzpicture}

### 2013-03-30 inc/model_results/shapes/shapes.tex
more tweaks to shapes section

@@
 
-The bunch propagation dynamics for negligible pulse charge density ($ N = 10^{0} $) shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
-In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands approximately twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp ($ \gamma_{\alpha} = 0 $).
+The bunch propagation dynamics for negligible pulse charge density case shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
+In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands approximately twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp.
 Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both oblate and prolate pulses are initially driven towards the spherical regime.
@@
 For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects can dominate the propagation dynamics.
-\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted temporal dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
+\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
 The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
 %TODO revisit this next sentance, the data doesn't seem to bear it out
-In the spherical case, space-charge induced bunch broadening overwhelms the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing an almost uniform three-dimensional `Coulomb explosion' of the electron pulse with a one-dimensional bunch broadening rate roughly one order of magnitude greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
+In the spherical case, space-charge induced bunch broadening begins to overwhelm the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing a more uniform three-dimensional `Coulomb explosion' of the electron pulse with the overall bunch broadening rate much greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
 For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation.
@@
 In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers,\cite{lobastov_four-dimensional_2005} deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
-Even for nanosecond DTEM (the extreme prolate pulse ($\xi \gg 1 $)), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
+Even for nanosecond DTEM (the extreme prolate pulse case, $\xi \gg 1 $), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
 In this thesis, however, I will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}

### 2013-04-01 inc/model_results/shapes/shapes.tex
clear some TODOs from shapes
tweaks to shapes

@@
 The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
-%TODO revisit this next sentance, the data doesn't seem to bear it out
 In the spherical case, space-charge induced bunch broadening begins to overwhelm the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing a more uniform three-dimensional `Coulomb explosion' of the electron pulse with the overall bunch broadening rate much greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
@@
 This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
-%TODO the following sentance appears not to be true, still perhaps should include something like it
+% the following sentance is no longer true, still perhaps should include something like it
 %In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
@@
 Even for nanosecond DTEM (the extreme prolate pulse case, $\xi \gg 1 $), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
-In this thesis, however, I will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}
+In this thesis, however, I will often restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be easily generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}

### 2013-04-01 inc/model_results/introduction.tex
tweaks to spacecharge section

@@
 Other examples in this section will require or are markedly different when using a more realistic pulse, one resulting from photoemission and acceleration.
-Simulations using pulses of this type will provide and HW1/eM width and duration of the laser, an acceleration voltage, and an excess photoemission energy.
+Simulations using pulses of this type will provide and HW1/eM width and duration of the laser, the length and voltage of the acceleration region, and an excess photoemission energy.
 In these cases, all initial parameters will be calculated from \ref{eq:summary}.

### 2013-04-01 inc/model_results/spacecharge/spacecharge.tex
added cc notices to the tops of several files
flattened model_results section headings, removing free-space versus forced
tweaks to spacecharge section

@@
-\subsection{Effect of Space Charge} \label{sec:free_spacecharge}
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
 
-A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
-Much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements.
+\section{Effect of Space Charge on Dynamics} \label{sec:free_spacecharge}
+
 Simplistically stated, at low pulse charge densities (low number of electrons/pulse), the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
 On the other hand, space-charge effects are expected to dominate at large pulse charge densities.
+A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
+While much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a typical example.
 
@@
   \end{tikzpicture}
-  \caption[Simplistic free-space pulse evolution vs charge density]{
-    The final HW1/eM width (solid) and length (dashed) of pulses which are intially similar except for the number of electrons in the pulse.
-    The pulse is initially 1 mm (HW1/eM) width and 1 ps in duration travelling at $c/3$ with 0.5eV excess photoemission energy.
-    It then traverses a 15cm drift region upon which the final length and width are recorded.
+  \caption[Idealistic free-space pulse evolution vs charge density]{
+    The final HW1/eM pulse width (solid) and length (dashed) after a 15 cm propagation. 
+    Pulses are intially similar except for the number of electrons in the pulse, each having 1 mm (HW1/eM) width and 1 ps in duration travelling at $c/3$ with 0.5eV excess photoemission energy.
   }
@@
 
-This trend is clearly evident in \ref{fig:spacecharge_noacc} which presents the predicted HW1/eM pulse width $\sqrt{2 \sigma_{\smallT}}$ (solid line) and longitudinal length $\sqrt{2 \sigma_{z}}$ (dashed line) as a function of the number of electrons in the pulse after a propagation distance of 15cm.
-The pulses are not accelerated but rather start with a velocity of $c/3$ (where $c$ is the speed of light in a vacuum), initial width of 1 mm, HW1/eM pulse duration 1 ps, have the initial momentum chirps $\gamma_{\smallT} = \gamma{z} = 0 $ and momentum uncertainties related to an excess photoemission energy of $\Delta E = $ 0.5 eV.
+The expected trend is clearly evident in \ref{fig:spacecharge_noacc} which presents the predicted HW1/eM pulse width $\sqrt{2 \sigma_{\smallT}}$ (solid line) and longitudinal length $\sqrt{2 \sigma_{z}}$ (dashed line) after a propagation distance of 15cm as a function of the number of electrons in the pulse.
+Once again, this simulation employs the idealistic pulse described earlier, starting with a velocity of $c/3$ (where $c$ is the speed of light in a vacuum), initial width of 1 mm, HW1/eM pulse duration 1 ps, have the initial momentum chirps $\gamma_{\smallT} = \gamma{z} = 0 $ and momentum uncertainties related to an excess photoemission energy of $\Delta E = $ 0.5 eV.
 These naive conditions are needed to isolate the behavior of a propagating pulse from its creation environment.
 The simulations clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
-In the absence of space-charge effects, the presented oblate shaped pulses should be maintained as such during unforced propagation because the electron pulse divergence and its temporal broadening rate are both proportional to $\Delta E^{1/2}$.
+In the absence of space-charge effects, the presented oblate shaped pulses should be maintained as such during unforced propagation because the initial values of the electron pulse' divergence and temporal broadening rate are both proportional to $\Delta E^{1/2}$ but not to $N$.
 
@@
 
-As the number of electrons/pulse increases beyond $10^6$, both the pulse length and diameter rapidly increase.
+In this example, as the number of electrons/pulse increases beyond $10^6$, both the pulse length and diameter rapidly increase.
 The model predicts that space-charge effects produce a spherical charge distribution of approximately five times larger width at $10^8$ electrons/pulse for these conditions.
@@
 
-\ref{fig:spacecharge_noacc} is specifically created to isolate the behavior of and ``ideal'' pulse which is independent of its creation environment.
+\ref{fig:spacecharge_noacc} employed the idealistic conditions specifically to isolate the behavior of an ``ideal'' pulse which is independent of its creation environment.
 Of course in a more realistic system, the electron pulse is generated from a photocathode and then accelerated.
-These forces play an interesting role in the system.
+Now that we have seen how the ideal case behaves, it is important to consider the interesting role that these more realistic factors play in the system.
 
@@
   \caption[More realistic free-space pulse evolution vs charge density]{
-    The final HW1/eM width (solid) and length (dashed) of pulses which are intially similar except for the number of electrons in the pulse.
-    The pulse is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy.
-    It then traverses a 15cm column, a 20mm acceleration region followed by drift region.
+    The final HW1/eM pulse width (solid) and length (dashed) after a 15 cm propagation. 
+    Pulses are intially similar except for the number of electrons in the pulse, each is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy and then traverses a 20mm acceleration region followed by drift region, accelerating to a velocity of $c/3$.
   }
@@
 
-Consider now a simulation comparable to the one presented above, but which uses the initial conditions presented in \ref{eq:summary} and accelerated by an electron gun as modelled in Section \ref{sec:gun_model} which is 20mm in length.
+Consider now a simulation comparable to the one presented above, but which uses the realistic initial conditions set.
+The pulse is generated by a 1 mm width 1 ps laser (HW1/eM) and accelerated by an electron gun which is 20mm in length and which has an accelerating voltage of 30 kV to speed $c/3$, the same velocity in the idealistic case.
 The results of this simulation are presented in \ref{fig:spacecharge_acc}.

### 2013-04-01 inc/model_results/lens_charge/lens_charge.tex
added cc notices to the tops of several files
flattened model_results section headings, removing free-space versus forced
tweaks to the lens_charge section

@@
-\subsection{Space-charge Effects in Magnetic Lenses} \label{sec:mag_lens_charge}
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
+\section{Space-charge Effects in Magnetic Lenses} \label{sec:mag_lens_charge}
 
@@
 For both lens strengths, the deleterious effects of space-charge are evident as the number of electrons/pulse $ N $ is increased.
-To make an effective comparison, highlighting the effect of space-charge alone, the pulses used are at 20kV velocity with $\Delta E = $ 0.5eV but are not simulated using an accelerator.
-The length of the accelerator and the anode lensing would make useful comparison difficult; as such the simulation presented approximates a beam which has been perfectly collimated at the front-focal plane of the lens.
+To make an effective comparison, highlighting the effect of space-charge alone, this simulation again uses the idealistic initial conditions set; the pulses used are at 20kV velocity with $\Delta E = $ 0.5eV but are not simulated using an accelerator.
+As such the simulation presented approximates a beam which has been perfectly collimated at the front-focal plane of the lens.
 
-The single-electron ($ N = 1 $) focal spot sizes of $\sim$150 microns for the $ f = 60 \text{mm} $ lens and $\sim$15 microns for the shorter $ f = 6.0 \text{mm} $ lens show marked increase starting around $ N \approx 10^{ 5 } $ and $ N \approx 10^{ 6 }$, respectively, and worsening dramatically above these thresholds.
+The single-electron ($ N = 1 $) focal spot sizes of $\sim$150 microns for the $ f = 60 \text{mm} $ lens and $\sim$15 microns for the shorter $ f = 6.0 \text{mm} $ lens show marked increase starting around $ N \approx 10^{ 5 } $ and $ N \approx 10^{ 6 }$, respectively, and worsen dramatically above these thresholds.
 This order of magnitude difference is fundamentally related to the greater time of flight before the focus of the $ f = 60\text{mm} $ lens; it increases the impulse that the internal Coulomb force can exert on the pulse as its charge density is increased under focusing.

### 2013-04-01 inc/model_results/rf_cav/rf_cav.tex
added rf cavity comparison plot (like figs 5(a)/6(a) from JAP paper)

--- 2013-03-29 inc/model_results/introduction.tex
@@
 
-In Chapters \ref{chap:previous_models} and \ref{chap:extension}, I have introduced the extended Analytic Gaussian model.
-In the following chapter I use this model to make general observations about pulse propagation for Ultrafast Electron Microscopy.
+\section{Space-charge Effects in RF Cavities} \label{sec:rf_cav_charge}
 
-These examples will each use one of two initialization schemes.
-The first represents an idealized pulse which is not generated by photocathode and then accelerated, but rather springs into existence at its initial position with its velocity already established.
-This type of pulse is useful to demonstrate general pulse dynamics or the ideal influence of electron-optical elements.
-Simulations using pulses of this type will provide an initial HW1/eM length and width, a velocity, and an excess photoemission energy.
-The length and width will be used directly to calculate initial spatial variances $\sigma_{\alpha}$.
-Both chirp terms, $\gamma_{\alpha}$, will be initially zero, but since pulses must still be subject to Liouville's Theorem (discussed in Section \ref{sec:liouville}), \ref{eq:summary} will still be used to calculate meaningful local momentum widths $\eta_{\alpha}$.
-This idealization only applies to the initialization, therefore these pulses will still be subject to internal space-charge repulsion which might cause pulse expansion before or during the demonstration, especially for high charge densities.
-
-Other examples in this section will require or are markedly different when using a more realistic pulse, one resulting from photoemission and acceleration.
-Simulations using pulses of this type will provide and HW1/eM width and duration of the laser, an acceleration voltage, and an excess photoemission energy.
-In these cases, all initial parameters will be calculated from \ref{eq:summary}.
-The acceleration process, and more specifically the anode (see Section \ref{sec:gun_model}), will have an impact on the beam dynamics which will be evident in the pulse's evolution.
+\input{rf_cav_plot}

### 2013-04-01 inc/model_results/rf_cav/rf_cav.pl

@@ -1,0 +1,75 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+use v5.10;
+
+use Physics::UEMColumn alias => [qw/:standard Pulse/];
+use Physics::UEMColumn::Auxiliary ':constants';
+
+use PDL;
+#use PDL::Graphics::Prima::Simple;
+
+my $vel  = sqrt( 2 * qe * 20_000 / me );
+my $freq = 3e9; # 3 GHz
+my $d = $vel / (2 * $freq);
+
+my @conditions = (
+  ['rf_cav_N1.dat', 1, '450 kilovolts / m'],
+  ['rf_cav_N1e5.dat', 1e5, '494 kilovolts / m'],
+);
+
+for my $cond (@conditions) {
+  my ($file, $num, $strength) = @$cond;
+
+  my $pulse = Pulse->new(
+    number => $num,
+    velocity => $vel,
+    initial_width  => '100 um',
+    initial_length => ($vel * 0.1) . 'ps m / s',
+    excess_photoemission_energy => '0.5 eV',
+  );
+
+  my $column = Column->new(
+    length => '250 mm',
+  );
+
+  my $sim = Physics::UEMColumn->new(
+    column => $column,
+    pulse => $pulse,
+    steps => 200,
+  );
+
+  my $lens1 = RFCavity->new(
+    location => '100 mm',
+    length   => $d . 'm',
+    frequency => $freq . 'Hz',
+    strength => $strength,
+  );
+
+  $sim->add_element($lens1);
+
+  my $result = pdl $sim->propagate;
+
+  my $z  = $result->slice('(1),') - 100e-3;
+  my $w = sqrt( 2 * $result->slice('(3),') );
+  my $vt = sqrt( 2 * $result->slice('(4),') );
+
+  #line_plot $z, $w;
+  #line_plot $z, $vt;
+
+  my $mask = $result->slice('1,') > 0.02;
+  my $later = $result->transpose->whereND( $mask->transpose )->transpose;
+
+  printf "\nNum: %d\n", $num;
+  printf "Mask starts at %.2fmm\n", $later->at(1,0)/1e-3;
+
+  # length
+  my $min_sz_ind = $later->slice('(4),')->minimum_ind;
+  my $min_length = sqrt( 2 * $later->at(4,$min_sz_ind) );
+  my $min_duration = $min_length / $later->at(2,$min_sz_ind);
[…]

### 2013-04-01 inc/model_results/rf_cav/rf_cav.tex
porting rf_cav model_results text from JAP paper

@@
 \input{rf_cav_plot}
+
+\ref{fig:rf_cav_num} illustrates the operation of a $ \Omega / 2 \pi = $ 3GHz $\text{TM}_{010}$ RF cavity of axial length $ d = \pi v_{\smallzero} / \Omega $ as an electron pulse compressor tuned to produce the shortest pulse at a distance $ z^{\prime} = $ 10cm behind the cavity for $N=1$ and middle $N=10^5$ pulse charge.
+The mathematical model of the RF cavity (see Section \ref{rf_cav_model}) employs $L_{RF} = d \approx $ 14mm, $ z^{\prime}_{RF} = 0 $, and $n=2$ (in \ref{eq:reg_of_influence}) to reflect the sharper end to the region-of-influence of the RF cavity.
+In both cases, the idealized initial conditions are used, starting the pulse at $ z^{\prime} = -10$ cm, an excess photoemission energy $\Delta E = 0.5$ eV, velocity related to a 20kV acceleration potential, HW1/eM $ w = $ 0.1mm and duration $\tau = $ 100fs ($v_{\smallzero} \tau \approx 8.4 \mu$m).
+This electron pulse is therefore initially in the preferred highly oblate regime ($ \sigma_{\smallT}(0) \gg \sigma_{z} (0) $), for which space-charge effects predominantly act in the longitudinal direction, broadening the pulse in time (see \ref{fig:compare_shape}).
+
+%\ref{fig:RFCav_N1} depicts the optimized RF cavity performance with $N=1$ electrons in the pulse; conditions under which the internal pulse space-charge effects are negligible, so that the RF cavity only compensates for the `dispersive' electron pulse broadening to $\sim$1.3ps at the RF cavity entrance (\ref{fig:RFCav_N1}\subref{fig:RFCav_N1_sigma}) that is due to the initial longitudinal momentum spread caused by the $\Delta E = $ 0.5eV excess photoemission energy.
+For an RF field amplitude $E_{0} = $ 423.5kV/m and phase $\phi = 0$, the dispersion-generated momentum chirp $\gamma_{z}$ is exactly reversed (\ref{fig:RFCav_N1}\subref{fig:RFCav_N1_gamma}) upon propagation through the RF cavity to produce a compressed pulse at $ z^{\prime} = $ 10cm behind the RF cavity (\ref{fig:RFCav_N1}\subref{fig:RFCav_N1_sigma}).
+In this case, the pulse is returned to its original temporal duration --- an exact 1-to-1 `temporal imaging' of the photo-generated electron pulse at $z' = $--10cm by the RF cavity to $ z' = $ 10cm.
+On the other hand, they dynamics of the spatial size of the electron pulse are clearly dominated by the negative compound transverse lens of the RF cavity, increasing at the temporal focus to a factor of 50$\times$ its initial value of $\sqrt{\sigma_{\smallT}} = 71\mu\text{m}$.
+
+%In analogy to optical imaging, the compressed pulse duration in this dispersive limit is determined by the image-to-object distance (magnification) ratio, with pulses shorter than the original laser pulse duration being produced when the distance from the RF cavity to the temporal focus is less than that from the effective photocathode to the RF cavity; specifically, the ratio of the time-of-flight between the RF cavity and the temporal focus to that between the photo-gun and the RF cavity.
+
+%The RF electric field amplitude $E_{0}$ of about 400kV/m required to achieve the temporal focusing 10cm behind the RF cavity in the dispersive limit is also consistent with expectations.
+By equating the differential impulse applied by the RF cavity across the electron pulse duration to that required to compensate for dispersive pulse broadening before the RF cavity and then compress the pulse thereafter, one can show that the required RF field amplitude for a $\text{TM}_{010}$ cavity with $ d = \pi v_{\smallzero} / \Omega $ is given to a good approximation by
+%\begin{equation} \label{eq:RF_field_required}
+%  E_{0} \approx V_{DC} \left ( \frac{1}{ L_{1} } + \frac{1}{ L_{2} } \right ) \text{,}
+%\end{equation}
+%where $L_{1}$ is the distance from the pulse source (the DC photoelectron gun) to the RF cavity (the object distance) and $L_{2}$ is the distance to the temporal pulse focus behind the RF cavity (the image distance).
+For the considered case, $V_{DC} = $ 20kV and $ L_{1} = L_{2} = $ 10cm, giving $ E_{0} \approx $ 400kV/m, which is in good agreement with the optimum value of 423.5kV/m employed for the simulation displayed in \ref{fig:RFCav_N1}\subref{fig:RFCav_N1_gamma}.
+\eqref{eq:RF_field_required} also clearly indicates that the DC photoelectron gun and $\text{TM}_{010}$ RF pulse compression cavity should be considered as a single system, since the temporal `focal length' of the RF cavity, $ V_{DC} / E_{0} $, is dependent upon the DC photo-gun voltage.
+
+%The optimized RF cavity performance for the temporal compression of a pulse with the same initial photoemission conditions, but with $N = 10^{5}$ electrons, is shown in \ref{fig:RFCav_N105}.
+In this case, intra-pulse space-charge effects act to further broaden the initial 100fs pulse duration to $\sim$6ps after the $\sim$1ns time-of-flight over the 10cm distance to the RF cavity.
+The fact that almost the same RF field strength is required to generate a temporal focus at $z^{\prime} = $ 10cm is a direct consequence of the linear dependence of the longitudinal RF cavity force on $z$ (\eqref{eq:RF_Force_z}), which means that the required restorative force is not dependent on $\sigma_{z}$ --- just as the focal length of a perfect optical lens is independent of its aperture size.
+However, unlike in the dispersion limited case (Figure \ref{fig:RFCav_N1}), the compressed pulse duration is now a factor of $\sim$4 less than the original 100fs laser pulse duration.
+This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression.\cite{strickland_compression_1985} In this case, the $\sim$100-fold increase in the transverse pulse size (corresponding to a $10^{4}\times$ reduced peak pulse charge density) aids the temporal compression by greatly diminishing Coulomb effects that counteract the reduction in electron pulse duration.
+Again, a stronger (and suitably phased) RF field will result in an even shorter compressed pulse duration as the distance to the temporal focus behind the RF cavity is reduced.
+
+%It is important to note that the example illustrated in \ref{fig:RFCav_N105} approaches the limit imposed in obtaining \eqref{eq:RF_Efield_approx}; namely that the longitudinal electron pulse length be much shorter than the axial length $d$ of the RF cavity, or $ \Omega z \ll v_{\smallzero} $.
+For a $\sim$10ps electron pulse entering the 3GHz RF cavity we have $ \Omega z / v_{\smallzero} = \Omega \tau \approx 0.19 $, for which the small angle approximation $ \sin ( \Omega z / v_{\smallzero} ) = \Omega z / v_{\smallzero} $ is 0.6\% inaccurate.
+Again, in principle, the AG model could simulate the action of the RF cavity on longer pulse by including additional terms $O \left ( z^{3} \right )$ and higher in \eqref{eq:RF_Force_z} resulting from the expansion of $ \sin \left ( \Omega z / v_{0} \right ) $, but at the expense of voiding its self-similar Gaussian approximation.

### 2013-04-01 inc/model_results/rf_cav/rf_cav.tex
done porting text of rf_cav model_results from JAP, still need figure caption

--- 2013-04-01 inc/model_results/rf_cav/rf_cav.tex
@@
 \input{rf_cav_plot}
+
+\ref{fig:rf_cav_num} illustrates the operation of a $ \Omega / 2 \pi = $ 3GHz $\text{TM}_{010}$ RF cavity of axial length $ d = \pi v_{\smallzero} / \Omega $ as an electron pulse compressor tuned to produce the shortest pulse at a distance $ z^{\prime} = $ 10cm behind the cavity for $N=1$ and middle $N=10^5$ pulse charge.
+The mathematical model of the RF cavity (see Section \ref{rf_cav_model}) employs $L_{RF} = d \approx $ 14mm, $ z^{\prime}_{RF} = 0 $, and $n=2$ (in \ref{eq:reg_of_influence}) to reflect the sharper end to the region-of-influence of the RF cavity.
+In both cases, the idealized initial conditions are used, starting the pulse at $ z^{\prime} = -10$ cm, an excess photoemission energy $\Delta E = 0.5$ eV, velocity related to a $V_{DC} = $ 20kV acceleration potential, HW1/eM $ w = $ 0.1mm and duration $\tau = $ 100fs ($v_{\smallzero} \tau \approx 8.4 \mu$m).
+This electron pulse is therefore initially in the preferred highly oblate regime ($ \sigma_{\smallT}(0) \gg \sigma_{z} (0) $), for which space-charge effects predominantly act in the longitudinal direction, broadening the pulse in time (see \ref{fig:compare_shape}).
+
+\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1} depicts the optimized RF cavity performance with $N=1$ electrons in the pulse.
+The internal pulse space-charge effects are negligible under these conditions, so that the RF cavity only compensates for the `dispersive' electron pulse broadening to $\sim$1.67ps at the RF cavity entrance that is due to the initial longitudinal momentum spread caused by the $\Delta E = $ 0.5eV excess photoemission energy.
+For an RF field amplitude $E_{0} = $ 450 kV/m and phase $\phi = 0$, the dispersion-generated momentum chirp $\gamma_{z}$ is exactly reversed upon propagation through the RF cavity to produce a compressed pulse at $ z^{\prime} = $ 10cm behind the RF cavity.
+In this case, the pulse is returned to its original temporal duration --- an exact 1-to-1 `temporal imaging' of the photo-generated electron pulse at $z' = -10$cm by the RF cavity to $ z' = $ 10cm.
+On the other hand, the dynamics of the spatial size of the electron pulse are clearly affected by the negative compound transverse lens of the RF cavity, increasing at the temporal focus to a factor of $\sim 10\times$ its initial value.
+
+In analogy to optical imaging, the compressed pulse duration in this dispersive limit is determined by the image-to-object distance (magnification) ratio, with pulses shorter than the original laser pulse duration being produced when the distance from the RF cavity to the temporal focus is less than that from the effective photocathode to the RF cavity; specifically, the ratio of the time-of-flight between the RF cavity and the temporal focus to that between the photo-gun and the RF cavity.
+
+%TODO why is this even worse than before?
+The RF electric field amplitude $E_{0}$ of about 450kV/m required to achieve the temporal focusing 10cm behind the RF cavity in the dispersive limit is also consistent with expectations.
+By equating the differential impulse applied by the RF cavity across the electron pulse duration to that required to compensate for dispersive pulse broadening before the RF cavity and then compress the pulse thereafter, one can show that the required RF field amplitude for a $\text{TM}_{010}$ cavity with $ d = \pi v_{\smallzero} / \Omega $ is given to a good approximation by
+\begin{equation} \label{eq:RF_field_required}
+  E_{0} \approx V_{DC} \left ( \frac{1}{ L_{1} } + \frac{1}{ L_{2} } \right ) \text{,}
+\end{equation}
+where $L_{1}$ is the distance from the pulse source (the DC photoelectron gun) to the RF cavity (the object distance) and $L_{2}$ is the distance to the temporal pulse focus behind the RF cavity (the image distance).
+For the considered case, $V_{DC} = $ 20kV and $ L_{1} = L_{2} = $ 10cm, giving $ E_{0} \approx $ 400kV/m, which is in reasonably good agreement with the optimum value of 450kV/m employed for the simulation displayed in \ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}.
+\ref{eq:RF_field_required} also clearly indicates that the DC photoelectron gun and $\text{TM}_{010}$ RF pulse compression cavity should be considered as a single system, since the temporal `focal length' of the RF cavity, $ V_{DC} / E_{0} $, is dependent upon the acceleration voltage.
+
+The optimized RF cavity performance for the temporal compression of a pulse with the same initial photoemission conditions, but with $N = 10^{5}$ electrons, is shown in \ref{fig:rf_cav_num}\subref{fig:rf_cav_N1e5}.
+In this case, intra-pulse space-charge effects act to further broaden the initial 100fs pulse duration to $\sim$6ps after the $\sim$1ns time-of-flight over the 10cm distance to the RF cavity.
+The fact that almost the same RF field strength (495 kV/m) is required to generate a temporal focus at $z^{\prime} = $ 10cm is a direct consequence of the linear dependence of the longitudinal RF cavity force on $z$ (\ref{eq:RF_Force_z}), which means that the required restorative force is not dependent on $\sigma_{z}$ --- just as the focal length of a perfect optical lens is independent of its aperture size.
+However, unlike in the dispersion limited case (\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}), the compressed pulse duration is now a factor of $\sim$4 less than the original 100fs laser pulse duration.
+This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression.\cite{strickland_compression_1985}
+In this case, the $\sim$100-fold increase in the transverse pulse size (corresponding to a $10^{4}\times$ reduced peak pulse charge density) aids the temporal compression by greatly diminishing Coulomb effects that counteract the reduction in electron pulse duration.
+Again, a stronger (and suitably phased) RF field will result in an even shorter compressed pulse duration as the distance to the temporal focus behind the RF cavity is reduced.
+
+It is important to note that the example illustrated in \ref{fig:rf_cav_num}\subref{fig:rf_cav_N1e5} approaches the limit imposed in obtaining \ref{eq:RF_Efield_approx}; namely that the longitudinal electron pulse length be much shorter than the axial length $d$ of the RF cavity, or $ \Omega z \ll v_{\smallzero} $.
+For a $\sim$10ps electron pulse entering the 3GHz RF cavity we have $ \Omega z / v_{\smallzero} = \Omega \tau \approx 0.19 $, for which the small angle approximation $ \sin ( \Omega z / v_{\smallzero} ) = \Omega z / v_{\smallzero} $ is 0.6\% inaccurate.
+In principle, the AG model could simulate the action of the RF cavity on longer pulse by including additional terms $O \left ( z^{3} \right )$ and higher in \ref{eq:RF_Force_z} resulting from the expansion of $ \sin \left ( \Omega z / v_{0} \right ) $, but at the expense of voiding its self-similar Gaussian approximation.

### 2013-04-01 inc/model_results/rf_cav/rf_cav.pl

@@ -51,6 +51,8 @@ for my $cond (@conditions) {
 
   my $result = pdl $sim->propagate;
 
+  printf "\nNum: %d\n", $num;
+
   my $z  = $result->slice('(1),') - 100e-3;
   my $w = sqrt( 2 * $result->slice('(3),') );
   my $vt = sqrt( 2 * $result->slice('(4),') );
@@ -61,7 +63,6 @@ for my $cond (@conditions) {
   my $mask = $result->slice('1,') > 0.02;
   my $later = $result->transpose->whereND( $mask->transpose )->transpose;
 
-  printf "\nNum: %d\n", $num;
   printf "Mask starts at %.2fmm\n", $later->at(1,0)/1e-3;
 
   # length
@@ -69,6 +70,7 @@ for my $cond (@conditions) {
   my $min_length = sqrt( 2 * $later->at(4,$min_sz_ind) );
   my $min_duration = $min_length / $later->at(2,$min_sz_ind);
   printf "Min Length: %.2fum (%.2ffs) at %.2fmm\n", $min_length/1e-6, $min_duration/1e-15, $later->at(1,$min_sz_ind)/1e-3;
+  printf "Width at min length: %.2fmm\n", sqrt( 2 * $later->at(3,$min_sz_ind) )/1e-3;
 
   wcols $z, $w, $vt, $file;
 }

### 2013-04-01 inc/model_results/rf_cav/rf_cav.tex
tweaks to the rf_cav model_results section, including a figure caption

@@
 The fact that almost the same RF field strength (495 kV/m) is required to generate a temporal focus at $z^{\prime} = $ 10cm is a direct consequence of the linear dependence of the longitudinal RF cavity force on $z$ (\ref{eq:RF_Force_z}), which means that the required restorative force is not dependent on $\sigma_{z}$ --- just as the focal length of a perfect optical lens is independent of its aperture size.
-However, unlike in the dispersion limited case (\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}), the compressed pulse duration is now a factor of $\sim$4 less than the original 100fs laser pulse duration.
+However, unlike in the dispersion limited case (\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}), the compressed pulse duration is now a factor of $\sim$2 less than the original 100fs laser pulse duration ($\sim 48$ fs).
 This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression.\cite{strickland_compression_1985}
-In this case, the $\sim$100-fold increase in the transverse pulse size (corresponding to a $10^{4}\times$ reduced peak pulse charge density) aids the temporal compression by greatly diminishing Coulomb effects that counteract the reduction in electron pulse duration.
-Again, a stronger (and suitably phased) RF field will result in an even shorter compressed pulse duration as the distance to the temporal focus behind the RF cavity is reduced.
+In this case, the $\sim$20-fold increase in the transverse pulse size (corresponding to a $\sim400\times$ reduced peak pulse charge density) aids the temporal compression by greatly diminishing Coulomb effects that counteract the reduction in electron pulse duration.
+Again, in analogy to optical imaging, a stronger (and suitably phased) RF field will result in an even shorter compressed pulse duration as the distance to the temporal focus behind the RF cavity is reduced.
 

### 2013-04-01 inc/model_results/rf_cav/rf_cav_plot.tex
added rf cavity comparison plot (like figs 5(a)/6(a) from JAP paper)
tweaks to the rf_cav model_results section, including a figure caption

--- 2013-01-10 inc/extension/initial_conditions/bin_model_figure.tex
@@
     \subfloat[][]{
+      \label{fig:rf_cav_N1}
       \begin{tikzpicture}
-        \input{Acc}
+        \input{rf_cav_N1}
       \end{tikzpicture}
@@
     \subfloat[][]{
+      \label{fig:rf_cav_N1e5}
       \begin{tikzpicture}
-        \input{Vmax}
-      \end{tikzpicture}
-    }
-    \\
-  }
-  \centerline{
-    \subfloat[][]{
-      \begin{tikzpicture}
-        \input{Equal}
+        \input{rf_cav_N1e5}
       \end{tikzpicture}
@@
   }
+  \caption[Dynamics of the pulse length and width under the influence of an RF cavity]{
+    Dynamics of the pulse length (green) and width (red) under the influence of an RF cavity placed at $z' = 0$.
+    The employed idealized pulses start at $z'= -100$ cm with velocity $v_{\smallzero} = \sim 84 \times 10^{6}$ m/s (20kV) and initial HW1/eM width of 100 $\mu$m and length $\sim 8.4 \mu$m ($v_{\smallzero} \cdot $ 100fs).
+    \subref{fig:rf_cav_N1} has $N=1$ electron and demonstrates one-to-one imaging of the initial pulse length at the focus at $ z' = 100 $ cm.
+    \subref{fig:rf_cav_N1e5} has $ N = 1 \times 10^5 $ electrons, whose added broadening allows for the factor of two reduction in pulse length at the focus ($\sim 48$ fs).
+  }
+  \label{fig:rf_cav_num}
 \end{figure}

### 2013-04-02 inc/model_results/introduction.tex
more tweaks to model_results introduction

@@
 The first represents an idealized pulse which is not generated by photocathode and then accelerated, but rather springs into existence at its initial position with its velocity already established.
-This type of pulse is useful to demonstrate general pulse dynamics or the ideal influence of electron-optical elements.
-Simulations using pulses of this type will provide an initial HW1/eM length and width, a velocity, and an excess photoemission energy.
-The length and width will be used directly to calculate initial spatial variances $\sigma_{\alpha}$.
-Both chirp terms, $\gamma_{\alpha}$, will be initially zero, but since pulses must still be subject to Liouville's Theorem (discussed in Section \ref{sec:liouville}), \ref{eq:summary} will still be used to calculate meaningful local momentum widths $\eta_{\alpha}$.
-This idealization only applies to the initialization, therefore these pulses will still be subject to internal space-charge repulsion which might cause pulse expansion before or during the demonstration, especially for high charge densities.
+This type of pulse is useful to demonstrate general pulse dynamics or the ideal influence of electron-optical elements independent of the pulse's creation conditions.
+Simulations using pulses of this type will provide an initial velocity $v_{\smallzero}$ (or equivalent acceleration voltage $V_{DC}$), an excess photoemission energy $\Delta E$, and HW1/eM dimensions width $w$ and duration $\tau$ or length $ w_z = v_{\smallzero} \tau$.
+The width and length will be used directly to calculate initial spatial variances $\sigma_{\alpha}$.
+Both momentum chirp terms, $\gamma_{\alpha}$, will be initially zero, but since pulses must still be subject to Liouville's Theorem (discussed in Section \ref{sec:liouville}), \ref{eq:summary} will still be used to calculate meaningful local momentum widths $\eta_{\alpha}$.
+This idealization only applies to the initialization, therefore these pulses will still be subject to intrinsic pulse broadening and internal space-charge repulsion which might cause pulse expansion before or during the demonstration, especially for high charge densities.
 
-Other examples in this section will require or are markedly different when using a more realistic pulse, one resulting from photoemission and acceleration.
+Other examples in this section will require --- or are markedly different when using --- a more realistic pulse resulting from photoemission and subsequent acceleration.
 Simulations using pulses of this type will provide and HW1/eM width and duration of the laser, the length and voltage of the acceleration region, and an excess photoemission energy.

### 2013-04-02 inc/model_results/shapes/shapes.tex
more tweaks to shapes section

@@
 
-The initial `space-time' eccentricity (defined in \ref{eq:define_xi}), $\xi ( 0 ) = v_{\smallzero} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
-\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), oblate or `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and prolate or `cigar-like' ($ \sigma_{ \smallT } ( 0 ) < \sigma_{ z } ( 0 ) $) pulses.
+The initial `space-time' eccentricity (defined in \ref{eq:define_xi}),
+\begin{equation}
+  \xi ( 0 ) = \sqrt{ \frac{ \sigma_{z}(0) }{ \sigma_{\smallT}(0) } } = \frac{ w_z(0) }{ w(0) } = \frac{ v_{\smallzero} \tau(0) }{ w(0) } \,\text{,}
+\end{equation}
+of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
+\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \xi(0) = 0 $), oblate or `disk-like' ($ \xi(0) < 1 $), and prolate or `cigar-like' ($ \xi(0) > 1 $) pulses.
 %TODO rewrite using w?
-This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
+This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w^2 w_z = $ 1mm$^3$.
 Pulses used in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} have just one electron per pulse (i.e.
@@
 The bunch propagation dynamics for negligible pulse charge density case shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
-In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands approximately twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp.
+In particular, in the initially spherical case ($ \xi(0) = 1 $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands approximately twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp.
 Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both oblate and prolate pulses are initially driven towards the spherical regime.
@@
 
-For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects can dominate the propagation dynamics.
+For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects may dominate the propagation dynamics.
 \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}

### 2013-04-02 inc/model_results/spacecharge/spacecharge.tex
tweaks to the spacecharge section

@@
 
-Simplistically stated, at low pulse charge densities (low number of electrons/pulse), the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
+As seen in Section \ref{sec:initial_shapes}, at low pulse charge densities (low number of electrons/pulse), the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
 On the other hand, space-charge effects are expected to dominate at large pulse charge densities.
 A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
-While much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a typical example.
+While much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a progressive series of examples.
 
@@
 
-The expected trend is clearly evident in \ref{fig:spacecharge_noacc} which presents the predicted HW1/eM pulse width $\sqrt{2 \sigma_{\smallT}}$ (solid line) and longitudinal length $\sqrt{2 \sigma_{z}}$ (dashed line) after a propagation distance of 15cm as a function of the number of electrons in the pulse.
-Once again, this simulation employs the idealistic pulse described earlier, starting with a velocity of $c/3$ (where $c$ is the speed of light in a vacuum), initial width of 1 mm, HW1/eM pulse duration 1 ps, have the initial momentum chirps $\gamma_{\smallT} = \gamma{z} = 0 $ and momentum uncertainties related to an excess photoemission energy of $\Delta E = $ 0.5 eV.
-These naive conditions are needed to isolate the behavior of a propagating pulse from its creation environment.
-The simulations clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
+The expected trend is evident in \ref{fig:spacecharge_noacc} which presents the predicted HW1/eM pulse width $\sqrt{2 \sigma_{\smallT}}$ (solid line) and longitudinal length $\sqrt{2 \sigma_{z}}$ (dashed line) after a propagation distance of 15cm as a function of the number of electrons in the pulse.
+Once again, this simulation employs the idealistic pulse described earlier, starting with a velocity of $c/3$ (where $c$ is the speed of light in a vacuum), initial width of 1 mm, HW1/eM pulse duration 1 ps, and an excess photoemission energy $\Delta E = $ 0.5 eV.
+For these conditions, used to isolate the behavior of a propagating pulse from its creation environment, the simulations clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
 In the absence of space-charge effects, the presented oblate shaped pulses should be maintained as such during unforced propagation because the initial values of the electron pulse' divergence and temporal broadening rate are both proportional to $\Delta E^{1/2}$ but not to $N$.
@@
 
-In this example, as the number of electrons/pulse increases beyond $10^6$, both the pulse length and diameter rapidly increase.
+In this example, as the number of electrons/pulse increases beyond $10^6$, both the pulse's final length and width rapidly increase.
 The model predicts that space-charge effects produce a spherical charge distribution of approximately five times larger width at $10^8$ electrons/pulse for these conditions.
@@
 Consider now a simulation comparable to the one presented above, but which uses the realistic initial conditions set.
-The pulse is generated by a 1 mm width 1 ps laser (HW1/eM) and accelerated by an electron gun which is 20mm in length and which has an accelerating voltage of 30 kV to speed $c/3$, the same velocity in the idealistic case.
+The pulse is generated by a 1 mm width 1 ps laser (HW1/eM) and accelerated by an electron gun which is 20mm in length and which accelerates the pulse through a voltage of 30 kV to speed $c/3$, the same velocity as in the idealistic case.
 The results of this simulation are presented in \ref{fig:spacecharge_acc}.
 It is immediately obvious that the final transverse width is larger than in the ideal case presented in \ref{fig:spacecharge_noacc}, even in the low charge-density regime; this is primarily due to the additional divergence imparted by the negative lensing at the anode (see Section \ref{sec:gun_model}).
-One can see that though the pulse size does increase beyond a certain total pulse charge, and while that trend appears to start at a higher charge-density, due to the complexity of the pulse dynamics it is rather difficult to compare the exact charge-density from this plot.
+One can see that though the pulse size does increase beyond a certain total pulse charge, and while that trend appears to start at a higher charge, due to the complexity of the pulse dynamics it is rather difficult to compare the exact charge-density from this plot to that in the idealistic case.
 What can be seen is that the transverse size increases more sharply with pulse charge at higher levels.

### 2013-04-02 inc/model_results/lens_charge/lens_charge.tex
tweaks to lens_charge section

@@
 A consequent further effect is a shift in the focal position to $ z > f $.
-A more subtle effect is that disk-like pulses are more readily focused than cigar-like pulses --- the rate of transverse pulse broadening due to space-charge effects already being intrinsically greater in the latter (\ref{fig:compare_shape}).
+A more subtle effect is that oblate pulses are more readily focused than prolate pulses --- the rate of transverse pulse broadening due to space-charge effects already being intrinsically greater in the latter (\ref{fig:compare_shape}).
 

### 2013-04-02 inc/model_results/lens_charge/lens_charge_plot.tex
tweaks to lens_charge section

@@
     AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$.
-    All plot's pulses begin at the lens' front focal plane at 20 kV velocity ($c/3$) and have the same volume.
+    Each plot's pulses begin at the lens' front focal plane at 20 kV velocity ($c/3$) and have the same volume.
     The left plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are initially oblate shaped, HW1/eM width $ 500 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 0.1 $.
@@
     The top plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_long_prolate}) have a longer focal length $f = $ 60mm compared to the bottom plots (\subref{fig:focus_short_oblate} and \subref{fig:focus_short_prolate}) $ f = $ 6.0mm.
-    A logarithmic scale is used for comparative clarity.
-    Clearly the best performance is achieved for shorter focal lengths and oblate pulses even at higher initial charge densities.
+    A logarithmic scale is used for comparative clarity near the foci.
+    Clearly the best performance is achieved for shorter focal lengths and oblate pulses \subref{fig:focus_short_oblate} even at higher initial charge densities.
   }

### 2013-04-02 inc/model_results/rf_cav/rf_cav.tex
tweaks to the rf_cav model_results section

@@
 
-\input{rf_cav_plot}
-
-\ref{fig:rf_cav_num} illustrates the operation of a $ \Omega / 2 \pi = $ 3GHz $\text{TM}_{010}$ RF cavity of axial length $ d = \pi v_{\smallzero} / \Omega $ as an electron pulse compressor tuned to produce the shortest pulse at a distance $ z^{\prime} = $ 10cm behind the cavity for $N=1$ and middle $N=10^5$ pulse charge.
-The mathematical model of the RF cavity (see Section \ref{rf_cav_model}) employs $L_{RF} = d \approx $ 14mm, $ z^{\prime}_{RF} = 0 $, and $n=2$ (in \ref{eq:reg_of_influence}) to reflect the sharper end to the region-of-influence of the RF cavity.
+In comparison to using magnetic lenses to compensate for transverse pulse broadening by using magnetic lenses, RF cavities are being employed to compensate for longitudinal broadening in the ultrashort pulses used by UEM.\cite{oudheusden_electron_2007}
+\ref{fig:rf_cav_num} illustrates the operation of a $ \Omega / 2 \pi = $ 3GHz $\text{TM}_{010}$ RF cavity of axial length $ d = \pi v_{\smallzero} / \Omega $ as an electron pulse compressor tuned to produce the shortest pulse at a distance $ z^{\prime} = $ 10cm behind the cavity for $N=1$ and $N=10^5$ pulse charges.
+The mathematical model of the RF cavity (see Section \ref{rf_cav_model}) employs $L_{RF} = d \approx $ 14mm, $ z^{\prime}_{RF} = 0 $, and a region of influence super-Gaussian of order 2 ($n=2$ in \ref{eq:reg_of_influence}) to reflect the sharper end to the region-of-influence of the RF cavity.
 In both cases, the idealized initial conditions are used, starting the pulse at $ z^{\prime} = -10$ cm, an excess photoemission energy $\Delta E = 0.5$ eV, velocity related to a $V_{DC} = $ 20kV acceleration potential, HW1/eM $ w = $ 0.1mm and duration $\tau = $ 100fs ($v_{\smallzero} \tau \approx 8.4 \mu$m).
 This electron pulse is therefore initially in the preferred highly oblate regime ($ \sigma_{\smallT}(0) \gg \sigma_{z} (0) $), for which space-charge effects predominantly act in the longitudinal direction, broadening the pulse in time (see \ref{fig:compare_shape}).
+
+\input{rf_cav_plot}
 
@@
 where $L_{1}$ is the distance from the pulse source (the DC photoelectron gun) to the RF cavity (the object distance) and $L_{2}$ is the distance to the temporal pulse focus behind the RF cavity (the image distance).
-For the considered case, $V_{DC} = $ 20kV and $ L_{1} = L_{2} = $ 10cm, giving $ E_{0} \approx $ 400kV/m, which is in reasonably good agreement with the optimum value of 450kV/m employed for the simulation displayed in \ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}.
+For the considered case, $V_{DC} = $ 20kV and $ L_{1} = L_{2} = $ 10cm, giving $ E_{0} \approx $ 400kV/m, which is in reasonably good agreement with the simulation-determined optimum value of 450kV/m employed for the simulation displayed in \ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}.
 \ref{eq:RF_field_required} also clearly indicates that the DC photoelectron gun and $\text{TM}_{010}$ RF pulse compression cavity should be considered as a single system, since the temporal `focal length' of the RF cavity, $ V_{DC} / E_{0} $, is dependent upon the acceleration voltage.

### 2013-04-02 inc/model_results/compound/compound.tex
flattened model_results section headings, removing free-space versus forced
tweaks for compound (including more plot labels)

@@
 
-\subsection{UEM Column Example} \label{sec:compound}
+\section{UEM Column Example} \label{sec:compound}
 
-Through the use of \ref{eq:reg_of_influence} to specify the position and influence of electron optical components such as magnetic lenses and RF cavities, the extended AG model can be employed to design and assess a full UEM column from pulse generation to sample stage.
+Through the use of multiple electron optical components such as magnetic lenses and RF cavities and accelerators (properly region-of-influence limited by application of \ref{eq:reg_of_influence}), the extended AG model can be employed to design and assess a full UEM column from pulse generation to sample stage.
 The design of such a compound system is, of course, a cornerstone of modern electron microscope development, although space-charge effects are generally not included, since, on average, there is usually no more than one electron in the microscope column at a time.
@@
   \centering
-  \begin{tikzpicture}
-    \input{compound_data}
-  \end{tikzpicture}
+  \input{compound_plot}
   \caption[Example model of a compound system of electron-optical elements]{
@@
 
-%Mask starts at 20.67mm
-%w after anode: 579.33um
-%Min Width: 29.50um at 301.49mm
-%Min Length: 22.58um (269.30fs) at 299.45mm
-
 \ref{fig:compound} shows an example of a compound element system for the delivery of an ultrashort electron pulse to the specimen area; in this case, optimized for $ N = 10^{ 6 } $ --- the number of electrons per pulse required for single shot diffraction measurements.\cite{armstrong_practical_2007}
-The displayed electron pulse dynamics are for a $\tau = $ 100fs electron pulse with $w = $ 0.1mm generated from a laser driven Ta photocathode, with $\Delta E = $ 0.5eV using $ \hbar \omega = $ 4.75eV photons, then accelerated by 20 kV to a velocity of $8.4 \times 10^{7} \text{m}/\text{s}$.
+The displayed dynamics are for an electron pulse generated by a $\tau = $ 100fs $w = $ 0.1mm laser, driving a photocathode, with $\Delta E = $ 0.5eV, then accelerated by 20 kV to a velocity of $8.4 \times 10^{7} \text{m}/\text{s}$.
 In this simple system, an optimum RF cavity ($ \Omega = \pi v_{{ \scriptscriptstyle 0}} / d $) is surrounded by two magnetic lenses (positioned a realistic 5cm to either side of the RF cavity) that compensate for the electron beam divergence due to the non-zero $ \eta_{\smallT} (0) $ and the negative spatial lensing of the RF pulse compression cavity.

### 2013-04-02 inc/model_results/compound/compound_plot.tex
tweaks for compound (including more plot labels)

\begin{tikzpicture}
  \draw [dashed,fill=blue!20]
    (1.5,8.375)
    -- ++(0,-7.4)
    -- ++(0.6,0)
    -- ++(0,7.4)
    -- ++(-0.6,0)
      node [pos=0.5,rotate=30,above right] {Acceleration Region}
    -- cycle
  ;
  \draw [dashed]
    (5.73,8.375)
    -- ++(0,-7.4)
      node [pos=0,rotate=30,above right] {Magnetic Lens}
  ;
  \draw [dashed]
    (7.16,8.375)
    -- ++(0,-7.4)
      node [pos=0,rotate=30,above right] {RF Cavity}
  ;
  \draw [dashed]
    (8.55,8.375)
    -- ++(0,-7.4)
      node [pos=0,rotate=30,above right] {Magnetic Lens}
  ;
  \draw [dashed]
    (9.975,8.375)
    -- ++(0,-7.4)
      node [pos=0,rotate=30,above right] {Focus}
  ;
  \input{compound_data}
\end{tikzpicture}

### 2013-04-02 inc/design_considerations/considerations.tex
begin merging Andreas' comments on considerations section

@@
 However, microscopy is an experimental science, and designing a UEM is no less hands-on.
-
+%TODO avoid repitition? "When doing so, special ..."?
 When designing a UEM system, special care must be given to the creation and propagation of the electron beam.
-In the case of single-shot UEM, several factors lead to two major considerations: first, that a large area photoemission source is preferred to smaller area or even single point emitters, and second, the transverse emittance of the beam must be minimized.
+%TODO define single-shot and transverse emittance here or rephrase?
+In the case of single-shot UEM, several factors lead to two major considerations: first, that a large area photoemission source is required to smaller area or even single point emitters, and second, the transverse emittance of the beam must be minimized.
 
@@
 In this chapter, I will first convince the reader of these considerations.
-As a direct consequence of these considerations, it is our belief that a successful UEM project cannot simply stem from small modifications to the platform of a conventional electron microscope.
-In the following chapters I will show how we are addressing these considerations for the prototype system at UIC, both in column design (Chapter \ref{chap:prototype}) and photocathode engineering (Chapter \ref{chap:photocathode}).
+As a direct consequence thereof, it is our belief that a successful UEM project cannot simply stem from small modifications to the platform of a conventional electron microscope.
+In the following chapters, I will show how we are addressing these considerations for the prototype system at UIC, both in column design (Chapter \ref{chap:prototype}) and photocathode engineering (Chapter \ref{chap:photocathode}).
 
@@
 In a conventional electron microscope the electrons are generated by one of several processes, often by heating the cathode or applying a large field to the cathode, or combinations thereof.
-Although these processes produce a very large number of electrons, far more than required by the Rose criterion, they are essentially non-directional.
-A pinhole is used to eliminate most of the electrons with the exception of those which will travel down the column, reducing the beam current but producing a high-quality beam. %TODO ref(s)
-
-In fact, for conventional electron microscopes without time resolution, the Rose criterion may be met for even very low beam current by simply allowing the image to acquire for a long enough time to collect sufficent electrons.
+Although these processes produce a very large number of electrons, far more than required by the Rose criterion, they are essentially not highly directional.
+A pinhole is used to eliminate most of the electrons, with the exception of those which will travel down the column axis, thereby reducing the beam current but producing a high-quality beam.
+%TODO ref(s)
+For conventional electron microscopes without time resolution, the Rose criterion may be met for arbitrarily low beam current by simply allowing the image to acquire for a long enough time to collect sufficent electrons.
 Further, high beam current may be damaging to the sample over time, so a lower beam current is often preferred.
 
-By definition, for single-shot imaging with a DTEM, all of these electrons must be contained in a single pulse.
-Although sample damage concerns may be neglected, destructive measurements will result in useful images, the use of a pinhole is likely to reduce the pulse charge below the useful level.
+By definition, for single-shot imaging with a DTEM, all of the electrons used to form an image must be contained in a single pulse.
+In a single-shot DTEM, destructive measurements will usually result in useful images, in fact, the dynamics being studied are often destructive, regardless of the knock-on damage cause by the probe.
+However, although sample damage concerns may be neglected, the use of a pinhole is likely to reduce the pulse charge below the useful level.
 This means that the designer of a DTEM must endeavor to retain and utilize all of the generated electrons.

### 2013-04-02 inc/design_considerations/considerations.tex
merge Andreas' comments through child's law section

@@
 
-It is easy to believe that in order to make a useful image in an electron microcope, the number of electrons collected must be above some minimum number.
+It is straightforward to show that in order to make a useful image in an electron microcope, the number of electrons collected must be above some minimum number.
 The Rose criterion \cite{rose_television_1948} indicates that an average detector signal strength of $\sim$100 electrons/pixel is required for adequate gray scaling of the image.
@@
 By definition, for single-shot imaging with a DTEM, all of the electrons used to form an image must be contained in a single pulse.
-In a single-shot DTEM, destructive measurements will usually result in useful images, in fact, the dynamics being studied are often destructive, regardless of the knock-on damage cause by the probe.
-However, although sample damage concerns may be neglected, the use of a pinhole is likely to reduce the pulse charge below the useful level.
+In a single-shot DTEM, the dynamics being studied are often destructive, regardless of the knock-on damage cause by the probe pulse; what matters is that the system be able to probe the sample before it is destroyed.
+Nevertheless, although sample damage concerns may be neglected, the use of a pinhole is likely to reduce the pulse charge below the useful level.
 This means that the designer of a DTEM must endeavor to retain and utilize all of the generated electrons.
@@
 Since we are now aware that there is a large minimum number of electrons needed per pulse, it is only natural to ask if there is the practical limit on how much charge (i.e., the number of electrons/pulse, $N$) can be extracted from a pulsed laser-driven DC photoelectron gun.
-When the incident laser pulse duration $\tau$ is longer than the time-of-flight (TOF) $t_{gun}$ of a single electron across the acceleration gap $d$ of the DC gun with a potential $V_{DC}$, $\tau > t_{gun} = d \sqrt{2m/eV_{DC}}$, then the maximum photocurrent density $J_{CL}$ is given by the well-known Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923}:
+When the incident laser pulse duration $\tau$ is longer than the time-of-flight (TOF) $t_{gun}$ of a single electron across the acceleration gap $d$ of the DC gun with a potential $V_{DC}$, $\tau > t_{gun} = d \sqrt{2m/eV_{DC}} \approx 0.1-1$ns, typically, then the maximum photocurrent density $J_{CL}$ is given by the well-known Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923}:
 \begin{equation}
@@
 In obtaining the above equation, we have used an irradiated photocathode area of $\pi w^2$ associated with an incident Gaussian laser pulse of the form of \ref{eq:laser_form}.
-The relation clearly indicates that $N_{crit}$ is only dependent upon the initial DC surface field (i.e., $E_{DC}$) at the photocathode and the illuminated area --- as expected because a virtual cathode forms when the initial photocathode surface charge density associated with $E_{DC}$ is photoemitted, effectively screening the cathode from the acceleration field.
+The relation clearly indicates that $N_{crit}$ is only dependent upon the initial DC surface field (i.e., $E_{DC}$) at the photocathode and the illuminated area --- as expected, because a virtual cathode forms when the initial photocathode surface charge density associated with $E_{DC}$ is photoemitted, effectively screening the cathode from the acceleration field.
 
@@
 For our example case above, to cleanly extract $N \ge 10^8$ electrons/pulse we require $w \gtrsim 0.56 \text{ mm}$.
-I will discuss the effect that accomodating a large pulse size has on designing an Ultrafast Electron Microscope in Chapter \ref{chap:prototype}.
+The effect that accomodating a large transverse pulse size has on designing an Ultrafast Electron Microscope will be discussed in Chapter \ref{chap:prototype}.
 

### 2013-04-03 inc/design_considerations/considerations.tex
merged remaining comments from andreas to considerations
* added emittance growth to liouville and formal extension sections
* added explicit mention of \Delta p_T to initial conditions

@@
 \begin{equation} \label{eq:emittance_definition}
-  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > } \,\text{,}
+  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > - \left < x p_x \right >^2 } \,\text{,}
 \end{equation}
 where $m$ is the electron mass, $c$ is the speed of light in vacuum and $x$ and $p_x$ are the transverse position and momentum respectively.
-By Liouville's Theorem (Section \ref{sec:liouville}), we know that this quantity is conserved throught the propagation of the pulse, and therefore is related directly to the initial photoemission properties of the system.
-For a laser of the form of \ref{eq:laser_form} we have simply
+For the usual case of axially symmetric electron pulses, we may use the simplified form
+\begin{equation} \label{eq:emittance_definition}
+  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > } \,\text{.}
+\end{equation}
+By Liouville's Theorem (Section \ref{sec:liouville}), we know that this quantity is conserved throughout pulse propagation (for free-space and perfect electron-optical elements), and therefore is related directly to the initial photoemission properties of the system.
+For a laser pulse of the form of \ref{eq:laser_form} we have simply
 \begin{equation} \label{eq:emittance_definition_w}
@@
 \end{equation}
+Note that imperfect electron-optical elements may cause ``emittance growth'' \cite{oshea_reversible_1998}, increasing the pulse's emittance after these elements.
 
@@
 
-In Ref. \cite{berger_dc_2009}, we explore the impact of these limitations for an Ultrafast Electron Diffraction (UED) experiment.
-We find that for a desired spatial resolution per pixel $\Delta X$ (related to the magnification) using a $N_p$ x $N_p$ detector, one may only resolve a fractional change $\Delta a / a$ in a crystalline material's Bragg plane spacing $a$ in the M$^{th}$ diffraction order if
+In Ref. \cite{berger_dc_2009}, we explore the impact of these limitations for an ultrafast electron diffraction (UED) experiment.
+We find that for a desired spatial resolution per pixel $\Delta X$ (related to the magnification) using a $N_p$ x $N_p$ detector, one may only resolve a fractional change $\Delta a / a$ in a crystalline material's Bragg plane spacing $a$ in the $M^{th}$ diffraction order if
 \begin{equation}
@@
 To have $\Delta X = $1 nm spatial resolution, and be able to measure a 10\% change ($\Delta a / a = 0.1$) in a Bragg spacing $a \sim$ 1-2 \AA{} in the 1$^{st}$ diffraction order, the example DTEM from Section \ref{sec:childs_law}, equipped with a 1k x 1k CCD will require $\varepsilon_{\smallT} \lesssim$ 2 nm.
-For comparison, using the simplistic initial conditions in \ref{eq:summary}, exciting carriers from a HW1/eM $w = 0.56 \text{ mm}$ Gaussian spot-size on a Tantalum metal photocathode (work function $\Phi =$ 4.25 eV) using our laser (photon energy $\hbar \omega =$ 4.75 eV, further described in Section \ref{sec:laser}), for an excess photoemission energy $\Delta E = \hbar \omega - \Phi =$ 0.5 eV, the normalized transverse emittance is 226 nm.
+For comparison, using the simplistic initial conditions in \ref{eq:summary}, exciting carriers from a HW1/eM $w = 0.56 \text{ mm}$ Gaussian spot-size on a Tantalum metal photocathode (work function $\Phi =$ 4.25 eV) using our laser (photon energy $\hbar \omega =$ 4.75 eV, further described in Section \ref{sec:laser}), for an excess photoemission energy $\Delta E = \hbar \omega - \Phi =$ 0.5 eV, the normalized transverse emittance is about 200 nm.
 
-One might think to narrow this discrepancy by simply better matching the photocathode work function and the laser photon energy, reducing the excess photoemission energy $\Delta E$.
+One might think to narrow this discrepancy by simply better matching the photocathode work function and the laser photon energy, thereby reducing the excess photoemission energy $\Delta E$.
 Unfortunately, since the photoemission efficiency is proportional to $\Delta E^2$ \cite{shalaev_electron_1994}, though this would indeed reduce the transverse emittance, it would do so at the cost of a much reduced pulse charge.

### 2013-04-03 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
merging comments from andreas on back half of extension chapter
merged remaining comments from andreas to considerations
* added emittance growth to liouville and formal extension sections
* added explicit mention of \Delta p_T to initial conditions

@@
 Further, when viewed in comparison to the chirps $\gamma_{\alpha}$, one can now see that these $\Gamma_{\alpha}$ constants can be thought of as some minimum/intrinsic action of the pulse.
-This will become more apparent when in future sections
+This will become more apparent when, in future sections,
 %TODO include internal reference
@@
 %TODO include more commentary on $\varepsilon_{\smallT}$?
+Additionally, though the AG model explicitly satisfies Liouville's Theorem, there are cases where the pulse's phase space product may change in a realistic system.
+This effect is called ``emittance growth'' and is most commonly seen in TEM as a result of aberations in lens systems \cite{oshea_reversible_1998}.
+As will be seen in Section \ref{extension_limitations}, the AG model will, of necessity, only consider perfect lenses.

### 2013-04-03 inc/hardware/introduction.tex
merge andreas' comments on hardware introduction

@@
 
-It is common in the DTEM field to modify an existing Transmission Electron Microscope column, adapting it for laser-stimulated electron generation.
-For the reasons covered in Section \ref{sec:childs_law}, the prototype UEM at UIC is being built to accomodate large width electron pulses.
-To satisfy this restriction, it has been designed and built from parts either fabricated in-house or from commercially available, both for the ease of prototyping inherent in using replacable parts and the accomodate the physical limitations our choice of a large beam width have imposed on the standard electron microscope column components.
+It is common in the DTEM field to modify an existing transmission electron microscope column, adapting it for laser-stimulated electron generation.
+For the reasons covered in Section \ref{sec:childs_law}, the prototype UEM at UIC is being built to accomodate electron pulses with a large transverse size.
+To satisfy this restriction, it has been designed and built from parts either fabricated in-house or from commercially available sources, both for the ease of prototyping inherent in using replacable parts and to accomodate the physical limitations that our choice of a large beam width have imposed on the standard electron microscope column components.
 

### 2013-04-03 inc/hardware/laser/laser.tex
merge andreas' comments until laser frequency doubling section (not inclusive)

--- 2013-03-19 inc/laser/laser.tex
@@
 The spectroscopic properties of the biaxial Yb:KGW laser crystal \cite{Biswal_thermo_optical_05} indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength \cite{Brenier_new_criteria}.
-However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that along the optical $N_m$-axis \cite{Holtom_mode_locked_2006,Biswal_thermo_optical_05}.
+However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that of the absorption along the optical $N_m$-axis \cite{Holtom_mode_locked_2006,Biswal_thermo_optical_05}.
 This optimal requirement for orthogonal linear pump and laser polarizations led Holtom \cite{Holtom_mode_locked_2006} to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
@@
   \caption[Laser crystal and anti-reflection coating]{
-    \protect\subref{fig:crystal_geo} Top view of the 3x3x10mm AR coated Yb:KGW laser crystal showing the crystal cut, the direction of propagation of p-polarized laser radiation at 1040nm, and the $x$ (tangential) and $y$ (saggital) directions (in blue) employed in the thermal analysis;
+    \protect\subref{fig:crystal_geo} Top view of the 3x3x10mm AR coated Yb:KGW laser crystal showing the crystal cut, the direction of propagation of p-polarized laser radiation at 1040nm, and the $x$ (tangential) and $y$ (saggital) directions (in blue) employed in the thermal analysis.
+    Pump laser radiation (at $\sim 20^{\circ}$ to the crystal face) is not pictured.
     \protect\subref{fig:crystal_coating} Calculated reflectivity of the 193nm \ce{SiO2} AR coating as a function of incidence angle for $p$-polarized 1040nm laser radiation (solid line) and $s$-polarized 980nm pump radiation (dashed line).
@@
 Moreover, the two dual-axis collimated laser diodes have full-angle $1/e^2$ intensity beam divergences of $\theta_y \approx$ 4.4mrad and $\theta_x \approx$ 6.8mrad in the horizontal ($x$-direction parallel to the emitter array) and vertical ($y$) directions respectively.
-With 100mm focal length lenses, this allows the pump radiation to be focused to a pump spot diameter 2W $\approx$ 400$\mu$m, yielding a maximum incident pump irradiance of $\sim$28kW/cm$^2$ from one laser diode array.
+With 100mm focal length lenses, this allows the pump radiation to be focused to a pump spot diameter $2W \approx 400\mu$m, yielding a maximum incident pump irradiance of $\sim$28kW/cm$^2$ from one laser diode array.
 The Yb concentration was chosen to be 2\% atomic doping to ensure that more than 90\% of the pump radiation was absorbed over the 3mm Yb:KGW crystal thickness.
@@
   \frac{B_\smallT}{A_\smallT} = 
-  \frac{ \cos^2 \theta_2 \left( \dfrac{dn}{dT} \right) + 2 \Delta \alpha_\smallT \left( n \cos\theta_2 - \cos\theta_1 \right) }
-       { \cos^2 \theta_1 \left[ \left( \dfrac{dn}{dT} \right) + 2 \Delta \alpha_\smallT \left( n \cos\theta_2 - \cos\theta_1 \right) \right] }
+  \frac{ \cos^2 \theta_2 \left( \dfrac{dn}{dT} \right) + 2 \alpha_{\smallT} \Delta \left( n \cos\theta_2 - \cos\theta_1 \right) }
+       { \cos^2 \theta_1 \left[ \left( \dfrac{dn}{dT} \right) + 2 \alpha_{\smallT} \Delta \left( n \cos\theta_2 - \cos\theta_1 \right) \right] }
   \quad\text{.}
@@
 expansion with coefficient $\alpha T$ \cite{Rimington_thermal_lens_2004}.
-Generally, $\Delta \approx W l$ , where $W$ is the transverse size (radius) of the pumped region \cite{Koechner_thermal_1970}.
+Generally, $\Delta \approx W / l$ , where $W$ is the transverse size (radius) of the pumped region \cite{Koechner_thermal_1970}.
 In our case, $l \approx$ 3mm and $W \approx$ 0.2mm, so that $\Delta < 0.1$, implying that the thermal duct due to a non-zero $\frac{dn}{dT}$ primarily determines the required $B_T$/$A_T$ ratio.
@@
 Also shown are the parabolic approximations to the pump-induced temperature distribution (dashed lines) and representative TEM$_{00}$ Gaussian laser modes (shaded area) in the Yb:KGW crystal with a HW1/eM field mode sizes of $w_y = 160\mu$m and $w_x = 182 \mu$m (14\% larger due to refraction (\ref{fig:crystal}\subref{fig:crystal_geo}).
-This mode size ensures efficient pump-probe overlap while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium \cite{Brenier_new_criteria}; in particular, in the vertical direction (\ref{fig:laser-thermal}(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
+This mode size ensures efficient pump-probe overlap for the TEM$_{00}$-mode while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium \cite{Brenier_new_criteria}; in particular, in the vertical direction (\ref{fig:laser-thermal}(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
 
@@
 For the employed crystal cut in our TLS geometry (\ref{fig:crystal}\subref{fig:crystal_geo}), the anisotropy in $\kappa$ \cite{mochalov_laser_1997,website_EKSPLA} is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
-The refractive index
 %page 8635
-changes due to stress in the pumped gain medium are difficult to quantify since the substantial variations in the reported values of the thermo-optic coefficients in KGW make it impossible to extract either the sign or magnitude of the stress dependence of the refractive index \cite{website_time_bandwidth}.
-Moreover, nearly all analyses of thermal lensing effects Nd:KGW and Yb:KGW lasers have been performed for laser radiation polarized along the optical $N_m$-axis, which accesses the largest emission cross-section for both Nd- and Yb-doping \cite{Yumashev_laser_2007,Hellstrom_efficient_2006}.
-Consequently, for our laser polarized along the optical $N_p$-axis \cite{Holtom_mode_locked_2006}, there is insufficient information to include accurately the effects of stress-induced refractive index changes.
+We see no observable effect from changes due to stress in the pumped gain medium, due both to the much stronger effect of the natural birefringence of the crystal and the laser polarization being along the optical $N_p$-axis \cite{Holtom_mode_locked_2006}.
 
@@
 This is consistent both with the laser accessing more of the available inhomogeneous gain bandwidth and with our resonator design favoring operation with a strong Kerr lens in the Yb:KGW gain medium.
-Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal's optical $N_p$-axis \cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
+Given a value of nonlinear refractive index $n_2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal's optical $N_p$-axis \cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^{-1}$.
 From \ref{fig:laser-cavity}\subref{fig:laser_stability}, we see that the resulting increase in the strength of the lensing in the Yb:KGW crystal will result in a reduction of the laser mode size in the gain medium, which produces a better spatial match to gain duct produced by the laser diode pumps (less absorptive loss), thus generating an increase in output power.

### 2013-04-03 inc/hardware/laser/laser.tex
merged remaining comments from andreas on laser section. This included adding the section on long-term performance

@@
 
-For most photoemission sources, the fundamental laser output's 1.19 eV photons do not provide enough energy to excite electrons above the vacuum level, either in single or two-photon emission.\cite{eastman_photoelectric_1970,johnson_optical_1972,dowell_quantum_2009}
+For most photoemission sources, the fundamental laser photon energy of 1.19 eV does not provide enough energy to excite electrons above the vacuum level, either for single or two-photon emission.\cite{eastman_photoelectric_1970,johnson_optical_1972,dowell_quantum_2009}
 By appropriate application of crystals with strong non-linear properties, two photons at frequency $\omega$ may be converted into one photon at frequency $2 \omega$.\cite{shen_principles_2002}
@@
 
-The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
+The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was intially frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
 The crystal is uncoated and cut for frequency doubling with the fundamental $p$-polarized radiation at 1046nm incident at Brewster's angle --- a geometry that ensures both minimum loss for the fundamental radiation and, more importantly, a high damage threshold.
@@
 
-\subsection{Summary}
+\subsection{Long-term Performance Enhancements}
 
-A robust, high-power ($\sim$5W), diode-pumped and thermal-lens-shaped, femtosecond (220-350fs) Yb:KGW oscillator is presented.
-The ML operation of the laser is self-starting above a critical output power of $\sim$2W, being initiated by a 1\% modulation depth SBR and subsequently supported in the soliton regime through a balance of self-phase modulation (Kerr lensing) in the gain medium and negative GVD (introduced using dispersion compensating mirrors).
-The use of the TLS technique in the diode-pumped laser head provides for astigmatism-compensated TEM$_{00}$-mode operation (a spatial beam quality $M^2 <1.2$), while proper cavity design accommodates the strong Kerr lens (up to $\sim4\text{m}^{-1}$) in the Yb:KGW crystal under ML operation.
-A full analysis of the thermal conduction in our 3x3x10mm Yb:KGW crystal is used to determine the correct TLS pump conditions, which lead to both excellent pump-laser beam overlap for the near collinear pump geometry and a near perfect pump-induced thermal lens (i.e., parabolic temperature distribution) experienced by the intracavity laser beam in the gain medium.
+Since the initial construction and characterization of this laser system, described above, several improvements have been made, primarily to improve the long-term stability of the laser.
+Due to the laser's role as a component of the larger UEM system, its stability is a tremendously important factor, both for repeatability of measurements and suitability for use in a user instrument.
 
-The laser head design also employs a simple single layer \ce{SiO2} anti-reflection coating on the gain medium to provide simultaneously (i) minimal loss for the intracavity 1040nm laser radiation $p$-polarized along the optical $N_p$-axis (crystallographic $b$-axis) associated with the largest gain bandwidth and (ii) less than 0.5\% surface reflection for the orthogonally $s$-polarized 980nm laser diode radiation (polarized parallel to the crystallographic $a$-axis) to enable efficient pumping of the strongest absorption line along the optical $N_m$-axis (rotated by $\sim20^\circ$ from the crystal $a$-axis).
-In this near optimal pump-emission Yb:KGW crystal geometry for ML operation, the pump-induced thermal lens in the gain medium is observed to be positive, indicating that the sum of the thermo-optical coefficient and the stress-generated refractive index change with temperature is positive for 1$\mu$m radiation polarized parallel to the optical $N_p$-axis (crystallographic $b$-axis).
+The second harmonic generation system now employs a 3mm non-critically phase matched LBO crystal.
+This has two benefits, in this geometry the walk-off angle between the fundamental and second harmonic is zero while the group velocity mismatch of 150fs ($\sim$ 50fs/mm) is much less than the pulse duration 250fs.
+Second, because the non-critically phase matched crystal is constantly held at a temperature above the boiling point of water, this prevents the deleterious absorption of water which is inherent in the hygroscopic crystal.
+In the place of the curved mirrors originally employed, this second harmonic system uses 75mm lenses to focus the laser into the crystal and collimate the output.
 
-The 250fs ML pulses generated at 3.5W of average output power and a 63.5MHz repetition rate are frequency doubled in a 2mm-thick, Brewster-cut and critically phase-matched Lithium triborate (LBO) crystal.
-The resultant efficient frequency doubling, which exceeds 60\% at a ML fundamental power of $\sim$3W, yields 1.65W of average power at 523nm ---
+Although the laser system had been relatively robust, maintaining mode-locking was initially problematic.
+To prevent dust from entering interrupting mode locking, a dust cover was added above the laser cavity.
+More importantly, a 1.4mm aperture (approximately four times the HW1/eM beam width at the output coupler) was added to suppress higher-order transverse modes.
+As this is an additional loss in the cavity, the output coupler was changed from 93\% to 95\% reflectivity to maintain the soliton modelocking power.
+
+These changes have greatly improved the mode locking stability.
+The laser now regularly operates at $1.9 \pm 0.1$W in mode locked operation with 250fs pulse durations, with minimal regular maintenance.
+The 40-50\% efficient second harmonic generation yields 0.8-1W of green output.
+The green radiation can be again doubled to the ultraviolet (UV) in a 6mm $\beta$-Barium borate (BBO) crystal, though the group velocity mismatch results in a 4ps pulse duration.
+All data presented in this thesis has been acquired subsequent to these performance enhancements.
+
+%\subsection{Summary}
+
+%A robust, high-power ($\sim$5W), diode-pumped and thermal-lens-shaped, femtosecond (220-350fs) Yb:KGW oscillator is presented.
+%The ML operation of the laser is self-starting above a critical output power of $\sim$2W, being initiated by a 1\% modulation depth SBR and subsequently supported in the soliton regime through a balance of self-phase modulation (Kerr lensing) in the gain medium and negative GVD (introduced using dispersion compensating mirrors).
+%The use of the TLS technique in the diode-pumped laser head provides for astigmatism-compensated TEM$_{00}$-mode operation (a spatial beam quality $M^2 <1.2$), while proper cavity design accommodates the strong Kerr lens (up to $\sim4\text{m}^{-1}$) in the Yb:KGW crystal under ML operation.
+%A full analysis of the thermal conduction in our 3x3x10mm Yb:KGW crystal is used to determine the correct TLS pump conditions, which lead to both excellent pump-laser beam overlap for the near collinear pump geometry and a near perfect pump-induced thermal lens (i.e., parabolic temperature distribution) experienced by the intracavity laser beam in the gain medium.
+%
+%The laser head design also employs a simple single layer \ce{SiO2} anti-reflection coating on the gain medium to provide simultaneously (i) minimal loss for the intracavity 1040nm laser radiation $p$-polarized along the optical $N_p$-axis (crystallographic $b$-axis) associated with the largest gain bandwidth and (ii) less than 0.5\% surface reflection for the orthogonally $s$-polarized 980nm laser diode radiation (polarized parallel to the crystallographic $a$-axis) to enable efficient pumping of the strongest absorption line along the optical $N_m$-axis (rotated by $\sim20^\circ$ from the crystal $a$-axis).
+%In this near optimal pump-emission Yb:KGW crystal geometry for ML operation, the pump-induced thermal lens in the gain medium is observed to be positive, indicating that the sum of the thermo-optical coefficient and the stress-generated refractive index change with temperature is positive for 1$\mu$m radiation polarized parallel to the optical $N_p$-axis (crystallographic $b$-axis).
+%
+%The 250fs ML pulses generated at 3.5W of average output power and a 63.5MHz repetition rate are frequency doubled in a 2mm-thick, Brewster-cut and critically phase-matched Lithium triborate (LBO) crystal.
+%The resultant efficient frequency doubling, which exceeds 60\% at a ML fundamental power of $\sim$3W, yields 1.65W of average power at 523nm ---
 %page 8640
-a peak green pulse power of $\sim$130kW.
-This laser is the short-pulse photon source used to photoemit electrons for all the experimental results shown in this thesis.
+%a peak green pulse power of $\sim$130kW.
+%This laser is the short-pulse photon source used to photoemit electrons for all the experimental results shown in this thesis.
 

### 2013-04-03 inc/hardware/laser/laser.tex
tweaks to the laser section

@@
 The second harmonic generation system now employs a 3mm non-critically phase matched LBO crystal.
-This has two benefits, in this geometry the walk-off angle between the fundamental and second harmonic is zero while the group velocity mismatch of 150fs ($\sim$ 50fs/mm) is much less than the pulse duration 250fs.
-Second, because the non-critically phase matched crystal is constantly held at a temperature above the boiling point of water, this prevents the deleterious absorption of water which is inherent in the hygroscopic crystal.
-In the place of the curved mirrors originally employed, this second harmonic system uses 75mm lenses to focus the laser into the crystal and collimate the output.
+This change has two benefits. 
+First, in this geometry, the walk-off angle between the fundamental and second harmonic is zero, while the group velocity mismatch of 150fs ($\sim$ 50 fs/mm) is much less than the pulse duration 250fs.
+%TODO cite Koechner?
+Second, because in this case, non-critically phase matching requires that the crystal is held at an elevated temperature, well above the boiling point of water, the deleterious absorption of water in the hygroscopic LBO crystal is prevented.
+Note that additionally, in the place of the curved mirrors originally employed, the second harmonic system now uses 75mm lenses to focus the laser into the crystal and collimate the output.
 
-Although the laser system had been relatively robust, maintaining mode-locking was initially problematic.
-To prevent dust from entering interrupting mode locking, a dust cover was added above the laser cavity.
+Although the laser system had been relatively robust, maintaining mode-locking operation was initially problematic.
+A cover was added above the laser cavity to prevent dust and air currents from interrupting mode locking.
 More importantly, a 1.4mm aperture (approximately four times the HW1/eM beam width at the output coupler) was added to suppress higher-order transverse modes.

### 2013-04-04 inc/hardware/column.tex
merge comments on column

@@
 Apart from the laser system, the majority of an Ultrafast Electron Microscope column is either vacuum hardware or contained within the vacuum system.
-The vacuum chamber itself is almost entirely 4.5 in standard ConFlat hardware.
+The vacuum chamber itself is almost entirely standard 4.5 in ConFlat (CF) hardware.
 Much has been purchased from off-the-shelf hardware suppliers.
@@
 Kimball's line of Multi-CF hardware has been especially useful in building the prototype column.
-The main chamber is a 4.5 in ``spherical cube,'' a large stainless-steel sphere which has six 4.5 in ConFlat ports oriented as a cube, with another eight 1.33 in ports on the corners of the cube.
+The main chamber is a 4.5 in ``spherical cube''; a large stainless-steel sphere which has six 4.5 in ConFlat ports oriented as a cube, with another eight 1.33 in ports on the corners of the cube.
 All ports on Multi-CF hardware also contain a set of grooves, designed as mounting points for a system of erector-set-like equipment called ``eV Parts.''
-This system is used to mount the accelerator anode and the the deflector plates which are both to be described later.
+This system is used to mount the accelerator anode and the deflector plates which are both to be described later.

### 2013-04-04 inc/gun/gun_design.tex
merge comments from gun design section

@@
 
-The acceleration region of our prototype UEM system has several requirement which are unique in comparison to standard electron guns.
-As discussed in \ref{chap:considerations} notable requirements include the need to handle a large photoemission area and the inability to employ a pin-hole to selectively clean the electron beam.
-This concern leads to several additional criteria.
+The acceleration region of our prototype UEM system has several features which are unique in comparison to standard electron guns.
+As discussed in \ref{chap:considerations} notable requirements for a UEM electron gun include the need to handle a large photoemission area and yet to retain all of the electrons, it may not employ a pin-hole to selectively clean the electron beam.
+These necessities lead to several additional criteria.
 The system must first be able to accept the large laser spot size needed to generate the electron beam.
-Then, in order to accomodate this large electron pulse, both the cathode Wehnelt and anode aperture must be similarly large; in fact, the beam size at the anode aperture will typically be larger than at the emission source due to beam expansion during acceleration.
+Then, in order to accomodate this large electron pulse, both the cathode Wehnelt and anode aperture must be similarly large; in fact, the beam size at the anode aperture will typically be larger than at the emission source due to transverse beam expansion during acceleration.
 Also, care must be taken to ensure a relatively uniform electric field for acceleration of the pulse.
-While in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the electic field given that it will span a wider proportion of the accelerator than the thin beams used in conventional TEM.
+While in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the electic field given that it will span a wider proportion %TODO andreas wrote "larger <<something unreadable>>"
+of the acceleration region than the thin beams used in conventional TEM.
 
@@
 
-These requirements lead to the fabrication of cathode/Wehnelt and anode pair based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007} (pictured in \ref{fig:togawa-pic}).
+These requirements led to the fabrication of cathode/Wehnelt and anode pair based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007} (pictured in \ref{fig:togawa-pic}).
 A schematic of this design, overlayed by a Finite Element Model (FEM) simulation of the electic field, is shown in \ref{fig:gun-field}.
@@
 \ref{fig:field_on_axis_fem} compares the FEM computed electric field to the form asserted for the AG model (\ref{eq:anode_tanh}) for several values of the sharpness parameter $s$; the lines are the same as in \ref{fig:gun-sharpness}\subref{fig:field_on_axis}.
-As derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electic field.
-This field will cause a divergence of the beam.
+As derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electic field which, in turn, will cause a divergence of the beam.
 Future work will include investigating the designs of Butler %TODO reference
@@
 A gross realignment of the cathode and anode is sometimes necessary to direct the emitted beam down the center axis of the column.
-To facilitate this, the prototype column has the photocathode and Wehnelt mounted on the end of the high voltage feedthrough, which is in turn mounted on a two axis alignment mechanism.
-Although this mechanism, called a ``port aligner,'' is typically used to couple two non-colinear vacuum ports, for use in this instrument it has been modified to be able to manually adjust the aligner during use.
-In contrast to this flexibility, the anode is mounted to column using Kimball Physics's ``eV Parts'' mounting systems.
+To facilitate this in the prototype column, the photocathode and Wehnelt are mounted on the end of the high voltage feedthrough which is, in turn, mounted on a two axis alignment mechanism.
+Although this mechanism, called a ``port aligner,'' is typically used to couple two non-colinear vacuum ports, for use in this instrument it has been modified to be able to manually adjust the aligner (i.e.
+cathode position) during use.
+In contrast, the anode is mounted to column using Kimball Physics's ``eV Parts'' fixed mounting systems.
 These strictly manufactured mounting hardwares ensure that the anode is properly aligned with the remainder of the microscope column.
-Thus, by iteratively repositioning the input laser and then the gun alignment, the beam may be walked to pass as near as possible down the axis of the column.
+Thus, by iteratively repositioning the input laser on the cathode and then the gun cathode alignment, the beam may be walked to pass as near as possible down the axis of the column.
 

### 2013-04-04 inc/gun/gun_design.tex
fixed typo in gun design

@@
 The acceleration region of our prototype UEM system has several features which are unique in comparison to standard electron guns.
-As discussed in \ref{chap:considerations} notable requirements for a UEM electron gun include the need to handle a large photoemission area and yet to retain all of the electrons, it may not employ a pin-hole to selectively clean the electron beam.
+As discussed in Chapter \ref{chap:considerations}, notable requirements for a UEM electron gun include the need to handle a large photoemission area and yet to retain all of the electrons, it may not employ a pin-hole to selectively clean the electron beam.
 These necessities lead to several additional criteria.

### 2013-04-04 inc/hardware/magnetic_lenses/maglens.tex
merging andreas comments on magnetic lens design section

@@
 
-In many ways, an electron beam can be thought of in terms of an optical equivalent system. 
+In many ways, an electron beams and electron microscope columns can be thought of in terms of an optical equivalents. 
 Of course there are notable exceptions; photons don't repel each other; photon energies are hard to change.
-Electron optical systems do have an analogous system to the lens, in this case using magnetic fields rather than glass. 
+Electron optical systems do have an element which is analogous to the optical lens, in this case using magnetic fields rather than glass. 
 The magnetic lens is a common element in electron microscopes and other electron-optical systems.
@@
 Further --- as suggested by the use of the auxiliary field $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility.
-These materials, called ``magnetic shielding'' since it captures the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.
+These materials, called ``magnetic shielding'' since they capture the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.
 The field lines exit the shielding in the gap and it is this field that acts as the lens.
-This gap can be much smaller than the total width of the coil turns, and yet contain in a field commensurate with the full coil.
+This gap can be much smaller than the total width of the coil turns, and yet contain a field strength commensurate with that of the full coil.
 This geometry is presented schematically as \ref{fig:polepiece-schematic}.
@@
 
-\ref{fig:polepiece-design} shows a simplistic design schematic of the lens designed and built for the UEM at UIC.
+\ref{fig:polepiece-design} shows a schematic of the lens designed and built for the UEM at UIC.
 The central drift tube is made of stainless steel having a magnetic permeability of 1, thus magnetically not involved in the field shaping.
-In this way the inner bore of the tube is in vacuum with minimal welding necessary.
-The outer diameter of tube is machined down to $D=2R=$0.5in, though a disk of larger diameter, and length $S$=0.1in, is left at the midpoint, which will separate the shielding material and be the pole-piece gap.
+In this way only the inner bore of the tube may be in vacuum thus minimizing necessary welding.
+The outer diameter of tube is machined down to $ D = 2R = 0.5 \text{in}$, though a disk of larger diameter, and length $ S = $ 0.1in, is left at the midpoint, which will separate the shielding material and be the pole-piece gap.
 
-The high magnetic permeability shielding is in eleven pieces.
+The high magnetic permeability shielding is made of eleven pieces.
 There are two end pieces which are revolution solids of an ``L'' shaped profile.
@@
 The combination of the two ends and this disk make a ``U'' shape which contains the windings.
-
 The top of the each end's large outer disk (the walls of the winding container) are flattend on eight sides.
@@
 
-El-Kareh and El-Kareh provide an analysis of symmetric lenses with magnetic shielding and a pole-piece (see Ref.
+El-Kareh and El-Kareh provide an analysis for symmetric lenses with magnetic shielding and a pole-piece (see Ref.
 \cite{el-kareh_electron_1970} chapter 8, with the most useful information being in section 8.7).
-They define
-\begin{subequations}
-\begin{gather}
-  k^2 \equiv \frac{e}{8 m V_r} B_0^2 R^2 \\
-  \beta \equiv k^2 V_r / (NI)^2 \label{eq:elkareh_k2}
-\end{gather}
-\end{subequations}
+They define two parameters,
+\begin{equation}
+  k^2 \equiv \frac{e}{8 m V_r} B_0^2 R^2
+\end{equation}
+and
+\begin{equation} \label{eq:elkareh_k2}
+  \beta \equiv k^2 V_r / (NI)^2 \,\text{,}
+\end{equation}
 where $N$ is the number of loops, $I$ is the current in a single loop, $B_0$ is the maximum axial magnetic field strength, $R$ is the bore radius and $V_r$ is the relativity corrected acceleration potential, given by
 \begin{equation}
-  V_r = V ( 1 + \frac{0.978 \cdot 10^{-6}}{\text{Volt}} V )
+  V_r = V ( 1 + \frac{0.978 \cdot 10^{-6}}{\text{Volt}} V ) \,\text{.}
 \end{equation}

### 2013-04-04 inc/hardware/magnetic_lenses/maglens.tex
merge more comments on magentic lens design

@@
 \end{equation}
-They then derive a number of formulae and tabulate the results for a variety of parameters.
-Using these tabulated results, and for a choice of parameters one can determine the number of ampere-turns ($NI$) to create a certain focal length.
-
+Using the provided tabulated results, and for a choice of parameters one can determine the number of ampere-turns ($NI$) to create a certain focal length.
 Since the drift tube has an outer diameter $D=$0.5in and the pole-piece gap length is $S$=0.1in, the lens has a ratio $S/D=0.2$.
-By their table 8.2a this results in $\beta=0.0146$. Then, to create a focal length of 4in ($f/R=16$) table 8.13 gives a value for $k^2 \approx 0.06$.
+By their table 8.2a this results in $\beta=0.0146$.
+Then, to create a focal length of 4in ($f/R=16$) table 8.13 in Ref. \cite{el-kareh_electron_1970} gives a value for $k^2 \approx 0.06$.
 Finally substituting this value into \ref{eq:elkareh_k2}, with $V=30$kV ($V_r\approx30.88$kV), one can deduce that $NI\approx356$A.

### 2013-04-04 inc/gun/gun_design.tex
tweak on gun_design

@@
 Also, care must be taken to ensure a relatively uniform electric field for acceleration of the pulse.
-While in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the electic field given that it will span a wider proportion %TODO andreas wrote "larger <<something unreadable>>"
-of the acceleration region than the thin beams used in conventional TEM.
+While in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the electic field given that it will span a larger volume of the acceleration region than the thin beams used in conventional TEM.
 

### 2013-04-04 inc/hardware/magnetic_lenses/maglens.tex
merge remaining comments on mag_lens design. still need to tweak build schematic

@@
 
-For the UEM system at UIC, two of these lenses are employed to manage the divergence created by an RF cavity and focus the pulses to a focal spot. 
-These lenses are positioned one before and one after the cavity, forming a lens system of a combined effective focal length.
-For the purposes of design, one can make a reasonable approximation that the two lenses together may together provide the total necessary number of amp-turns.
-As each lens has 100 turns, a current of only about 2A is needed to drive the two lenses in series.
+For the UEM system at UIC, two of these lenses are employed to manage the divergence created by an RF cavity and the anode aperture, then focus the pulses to a focal spot. 
+These lenses are positioned one before and one after the RF cavity, forming a lens system of a combined effective focal length; the two lenses are wound oppositely to counteract beam rotation.
+For the purposes of design, one can make a reasonable approximation that the two lenses together provide the total necessary number of amp-turns.
+As each lens has 100 turns, a current of only about 2A is needed to drive the two lenses in series to attain a 4in focal length.
 A picture of this system is shown in \ref{fig:maglens-pic}.
@@
 %TODO need reference for equation and \alpha_{Cu}
-Since the lenses have no active cooling system, it was prudent to characterize the thermal response to a moderate current load.
+As the lenses have no active cooling system, it was prudent to characterize the thermal response to a moderate current load.
 Resistance of metals changes as a function of temperature by $ R = R_0 [ 1 + \alpha ( T - T_0 ) ] $, where $\alpha$ is a material contant; for copper, $\alpha = 3.9 * 10^{-3} \text{K}^{-1}$. 

### 2013-04-04 inc/hardware/deflector/deflector.tex
merge andreas' comments on the deflector

@@
 
-In order to redirect the beam down the column, custom large area deflector plates were employed in conjunction with the gross accelerator alignment mechanism described earlier.
+In order to direct the beam down the column, custom large area deflector plates are employed in conjunction with the gross accelerator alignment mechanism described earlier.
 Four plates are arranged on a hollow tube with a square profile, one on each side.

### 2013-04-04 inc/hardware/laser/laser.tex
more tweaks to laser section
merge additional comments on laser section

@@
 
-Since the initial construction and characterization of this laser system, described above, several improvements have been made, primarily to improve the long-term stability of the laser.
+Since the initial construction and characterization of this laser system, described above, several improvements have been made, primarily to improve the long-term mode-locking stability of the laser.
 Due to the laser's role as a component of the larger UEM system, its stability is a tremendously important factor, both for repeatability of measurements and suitability for use in a user instrument.
@@
 This change has two benefits. 
-First, in this geometry, the walk-off angle between the fundamental and second harmonic is zero, while the group velocity mismatch of 150fs ($\sim$ 50 fs/mm) is much less than the pulse duration 250fs.
-%TODO cite Koechner?
+First, in this geometry, the walk-off angle between the fundamental and second harmonic is zero, while the group velocity mismatch of 150fs ($\sim$ 50 fs/mm) is still less than the pulse duration 250fs.
+%TODO cite Newlight Photonics, Canada
 Second, because in this case, non-critically phase matching requires that the crystal is held at an elevated temperature, well above the boiling point of water, the deleterious absorption of water in the hygroscopic LBO crystal is prevented.
-Note that additionally, in the place of the curved mirrors originally employed, the second harmonic system now uses 75mm lenses to focus the laser into the crystal and collimate the output.
+%TODO replace above with once oven temperature is known
+%Second, because in this case, non-critically phase matching requires that the crystal is held at an elevated temperature of XXX$^{\circ}$C, the deleterious absorption of water in the hygroscopic LBO crystal is prevented.
+Additionally, in the place of the curved mirrors originally employed, the second harmonic system now uses 75mm lenses to focus the laser into the crystal and collimate the output.
 
@@
 As this is an additional loss in the cavity, the output coupler was changed from 93\% to 95\% reflectivity to maintain the soliton modelocking power.
+These changes have greatly improved the mode locking stability.
 
-These changes have greatly improved the mode locking stability.
-The laser now regularly operates at $1.9 \pm 0.1$W in mode locked operation with 250fs pulse durations, with minimal regular maintenance.
+The laser now regularly operates at $1.9 (\pm 0.1)$W in mode locked operation with 250fs pulse durations, with minimal regular maintenance.
 The 40-50\% efficient second harmonic generation yields 0.8-1W of green output.
-The green radiation can be again doubled to the ultraviolet (UV) in a 6mm $\beta$-Barium borate (BBO) crystal, though the group velocity mismatch results in a 4ps pulse duration.
+The green radiation can be again doubled to yield 261nm ($\hbar \omega = 4.75$eV) ultraviolet (UV) radiation in a 6mm $\beta$-Barium borate (BBO) crystal, though the group velocity mismatch results in a 4ps pulse duration.
 All data presented in this thesis has been acquired subsequent to these performance enhancements.

### 2013-04-05 inc/model_results/introduction.tex
merge more comments from andreas, now on to model_results chapter

@@
 In Chapters \ref{chap:previous_models} and \ref{chap:extension}, I have introduced the extended Analytic Gaussian model.
-In the following chapter I use this model to make general observations about pulse propagation for Ultrafast Electron Microscopy.
+In this chapter, I use this model to make general observations about pulse propagation for Ultrafast Electron Microscopy.
 
-These examples will each use one of two initialization schemes.
+The examples presented will each use one of two initialization schemes.
 The first represents an idealized pulse which is not generated by photocathode and then accelerated, but rather springs into existence at its initial position with its velocity already established.
 This type of pulse is useful to demonstrate general pulse dynamics or the ideal influence of electron-optical elements independent of the pulse's creation conditions.
-Simulations using pulses of this type will provide an initial velocity $v_{\smallzero}$ (or equivalent acceleration voltage $V_{DC}$), an excess photoemission energy $\Delta E$, and HW1/eM dimensions width $w$ and duration $\tau$ or length $ w_z = v_{\smallzero} \tau$.
+Simulations using pulses of this type will provide an initial velocity $v_{\smallzero}$ (or equivalent acceleration voltage $V_{DC}$), an excess photoemission energy $\Delta E$, and HW1/eM dimensions; width $w$ and duration $\tau$ or length $ w_z = v_{\smallzero} \tau$.
 The width and length will be used directly to calculate initial spatial variances $\sigma_{\alpha}$.

### 2013-04-05 inc/model_results/shapes/shapes.tex
merge more comments from andreas, now on to model_results chapter

@@
 of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
-\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \xi(0) = 0 $), oblate or `disk-like' ($ \xi(0) < 1 $), and prolate or `cigar-like' ($ \xi(0) > 1 $) pulses.
+\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \xi(0) = 1 $), oblate or `disk-like' ($ \xi(0) < 1 $), and prolate or `cigar-like' ($ \xi(0) > 1 $) pulses.
 %TODO rewrite using w?
@@
 Since the initial pulse volume is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1}.
-The pulse velocity is $ v_{\smallzero} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.
+The pulse velocity is $ v_{\smallzero} \approx c/3 $, which is equivalent to a DC photogun with $V_{DC} = 30\text{kV} $.
 
@@
 For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation.
-This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
+This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce the greatest differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
 % the following sentance is no longer true, still perhaps should include something like it

### 2013-04-05 inc/model_results/spacecharge/spacecharge.tex
merge more comments from andreas, now on to model_results chapter

@@
 
-As seen in Section \ref{sec:initial_shapes}, at low pulse charge densities (low number of electrons/pulse), the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
+As seen in Section \ref{sec:initial_shapes}, at low pulse charge densities, the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
 On the other hand, space-charge effects are expected to dominate at large pulse charge densities.

### 2013-04-05 inc/model_results/shapes/shapes.tex
ok I have decided to go with w_{\smallT} rather than just assume the _{\smallT}

@@
 \begin{equation}
-  \xi ( 0 ) = \sqrt{ \frac{ \sigma_{z}(0) }{ \sigma_{\smallT}(0) } } = \frac{ w_z(0) }{ w(0) } = \frac{ v_{\smallzero} \tau(0) }{ w(0) } \,\text{,}
+  \xi ( 0 ) = \sqrt{ \frac{ \sigma_{z}(0) }{ \sigma_{\smallT}(0) } } = \frac{ w_z(0) }{ w_{\smallT}(0) } = \frac{ v_{\smallzero} \tau(0) }{ w_{\smallT}(0) } \,\text{,}
 \end{equation}
@@
 %TODO rewrite using w?
-This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w^2 w_z = $ 1mm$^3$.
+This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
 Pulses used in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} have just one electron per pulse (i.e.

### 2013-04-05 inc/photocathode/plasmon/plasmon.tex
tweaking plasmon section

@@
 In the search for alternative mechanisms to reduce the intrinsic divergence of the electron beam upon photoemission, a possible area of interest was Plasmon-Assisted Photoemission (PAPE). 
+As the name suggests, a plasmon is the quasiparticle of plasma oscillation.
+When this occurs at the surface of a material, the resultant quasiparticles are called ``surface plasmon-polaritons.''
+Since these will be the only plasmons discussed in this paper, the word ``plasmon'' will be used interchangeably for the name ``surface plasmon-polariton.''
+%TODO should this be formalized, either one way (SPP) or the other
+Electrons in a material can couple with electromagnetic fields, in this case a laser field, and oscillate accordingly.
+This overall oscillation of electrons near the surface of the material creates strong oscillatory local surface electric field \cite{cottam_introduction_2004,concepts_2002}.
+In PAPE, the goal is use this field to both enhance the emission and `control' the divergence of photoemitted electrons through the combined action of photoemission barrier suppression and ponderomotive acceleration.
+
 This avenue of investigation was motivated by Zawadzka et al \cite{zawadzka_evanescent_2001} who indicate that when driving a surface plasmon on gold they witnessed a reduction in the emission angle of photoemitted electrons.
-In their paper, the authors mount a gold foil on a prism; the laser is incident on the gold through the prism.
-This geometry is called the Kretschmann geometry.
-
-While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, that UEM applications should explore the ``grating coupling'' geometry.
+In their paper, the authors mount a thin gold foil on a prism; the laser then ``back illuminates'' the gold through the prism, ejecting electrons on vacuum side of the foil.
+This arrangement is called the Kretschmann geometry.
+While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, we felt that for UEM applications we should explore the ``grating coupling'' geometry.
 In this geometry, the plasmon is driven on a periodic surface rather than a flat surface attached to glass.
@@
 
-%\subsection{Driving the Surface Plasmon}
-
-As the name suggests, a plasmon is the quasiparticle of plasma oscillation.
-Electrons in a material can couple with electromagnetic fields, in this case a laser field, and oscillate accordingly.
-When this occurs at the surface of a material, the resultant quasiparticles are called ``surface plasmon-polaritons.''
-Since these will be the only plasmons discussed in this paper, the word ``plasmon'' will be used interchangeably for the name ``surface plasmon-polaritons.''
-%TODO should this be formalized, either one way (SPP) or the other
-
-The mathematical formalism of surface plasmon-polaritons is a lengthy subject and well described elsewhere \cite{cottam_introduction_2004,concepts_2002}.
+\subsection{The Grating Coupling Geometry}
 
@@
 \begin{equation}
-  \omega_{P} = \sqrt{\frac{\rho_{e} e^2}{\varepsilon_{0} m^*}}
+  \omega_{P} = \sqrt{\frac{\rho_{e} e^2}{\varepsilon_{0} m^*}} \,\text{,}
 \end{equation}
 where $\rho_{e}$ is the electron density and $m*$ is the effective mass of the material.
-And in the most simple (undamped) approximations, the dielectric function of the material may be written in terms of this constant.
+And in the most simple (undamped) approximations, the dielectric function of the material may be written in terms of this constant
 \begin{equation} \label{eq:dielectric_as_omega}
-  \varepsilon(\omega) = 1 - \frac{ \omega_{P}^2 }{ \omega^2 }
+  \varepsilon(\omega) = 1 - \frac{ \omega_{P}^2 }{ \omega^2 } \,\text{.}
 \end{equation}
@@
 \begin{equation}
-  \omega_{SP} = \omega_{P} / \sqrt{2}
+  \omega_{SP} = \omega_{P} / \sqrt{2} \,\text{.}
 \end{equation}
-
 Further, plasmons carry momentum whose magnitude is given by
 \begin{equation}
-  k_{P}^2 = \frac{\omega^2}{c^2} \left( \frac{\varepsilon_{1} \varepsilon_{2}}{\varepsilon_{1} + \varepsilon_{2}} \right)
+  k_{P}^2 = \frac{\omega^2}{c^2} \left( \frac{\varepsilon_{1} \varepsilon_{2}}{\varepsilon_{1} + \varepsilon_{2}} \right) \,\text{,}
 \end{equation}
@@
 \begin{equation}
-  k_{SP}^2 = \frac{\omega^2}{c^2} \left( \frac{ \varepsilon }{ 1 + \varepsilon } \right)
+  k_{SP}^2 = \frac{\omega^2}{c^2} \left( \frac{ \varepsilon }{ 1 + \varepsilon } \right) \,\text{.}
 \end{equation}
@@
 \begin{equation}
-  k_{SP} = k_{Lx} + n k_{G}
+  k_{SP} = k_{Lx} + n k_{G} \,\text{,}
 \end{equation}
@@
 \begin{equation} \label{eq:plasmon_condition}
-  \sqrt{ \frac{ \varepsilon }{ 1 + \varepsilon } } = \sin \theta + \frac{ n \lambda }{ d }
+  \sqrt{ \frac{ \varepsilon }{ 1 + \varepsilon } } = \sin \theta + \frac{ n \lambda }{ d } \,\text{.}
 \end{equation}
@@
 At the instant that the plasmon started, a bright flash was visible on the scintillator screen then disappeared.
-
 After this, the grating showed visual signs of damage, both on the surface and in the reflected light pattern.
@@
 
-Though this conclusively shows that a plasmon was driven successfully, the heat generated by the plasmon oscillation destroyed the grating.
-Further this experiment validated the decision not to pursue a Kretschmann geometry for fear of heat damage at these high powers.
+There are several conclusions that can be drawn from this experiment.
+Most importantly, it shows that a plasmon was driven successfully given the angular dependence and large response, few other processes could have occurred.
+The heat generated by the plasmon oscillation destroyed the grating, therefore it validated the decision not to pursue a Kretschmann geometry over fears of heat damage at these high powers.
+Finally, it suggests that if using a thermally conductive and structurally sound photoemission material and mount for the plasmonic photocathode, PAPE might generate a large number of electrons (as evident by the bright flash).
 
@@
 Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
-Given a laser wavelength of 523nm ($\varepsilon_{Au} \approx -3.95$ \cite{johnson_optical_1972}), this choice of periodicity corresponds by \ref{eq:plasmon_condition} to an incidence angle of $\sim39^{\circ}$, which is within the acceptance range of our experimental setup (35-75$^{\circ}$).
-%TODO point to Togawa section
+By \ref{eq:plasmon_condition}, given a laser wavelength of 523nm ($\varepsilon_{Au} \approx -3.95$ \cite{johnson_optical_1972}), this choice of periodicity corresponds to an incidence angle of $\sim39^{\circ}$, which is within the acceptance range of our experimental setup (35-75$^{\circ}$) (see Section \ref{sec:gun_design}).
 An SEM image of one such region is shown in \ref{fig:fib-si-sem}.

### 2013-04-05 inc/photocathode/plasmon/plasmon.tex
more tweaks to plasmonic section

@@
 
-Using the Varian e-beam deposition chamber at the UIC Nano Core Facility (NCF), I deposited a thin layer of chromium as a binding agent, then a 300nm film of gold.
-This thickness of gold is several times thicker than the skin depth of green light on gold, therefore the laser radiation cannot interact with the silicon.
+A 300nm film of gold was deposited onto the silicon using the Varian e-beam deposition chamber at the UIC Nano Core Facility (NCF); a thin layer of chromium was deposited between the silicon and gold as a binding agent.
+As the gold film is several times thicker than the penetration depth of green light on gold, the laser radiation cannot interact with the silicon.
 %TODO should I comment on the coating acting to smooth the surface to a proper sinusoid?
@@
 % Content taken from 2010 ANL report
-Note that for this measurement, relative emission efficiencies are calculated from the integration of the luminosity detected on the YAG scintillator screen.
+Note that for the following measurement, relative emission efficiencies are calculated from the integration of the luminosity detected on the YAG scintillator screen.
 The scintillator and CCD camera detection scheme was also shown to be linear (i.e., integrated signal proportional to number of electrons per pulse) for at least $10^4$ electrons per pulse, using the known linearity of single-photon photoemission from Ta irradiated by our 261nm ultraviolet laser source.
@@
 
-Given the geometry of the previous measurement, notably the colinear arrangement of plasmon vector and the grating vector, the true cause is difficult to determine.
-However, if instead the grating is rotated --- and the laser incidence angle is corrected to account for the longer effective periodicity of the surface --- the directionality of these effects may be separated.
-
 \begin{figure}
@@
 
+Given the geometry of the previous measurement, notably the colinear arrangement of plasmon vector and the grating vector, the true cause is difficult to determine.
+However, if instead the grating is rotated --- and the laser incidence angle is corrected to account for the longer effective periodicity of the surface --- the directionality of these effects may be separated.
 \ref{fig:rotated} shows the comparison of the grating vector geometries for the original scheme (top) and one in which the grating is rotated by $45^{\circ}$ (bottom).
@@
 As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
-
 This result is commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
@@
 They are very intolerant to misalignment and would therefore be problematic for users.
-Finally, the cannot be shown to produce the performance needed to justify the previous concerns.
+Finally, they cannot be shown to produce the performance (that is, a reduced transverse emittance) needed to justify the previous concerns.
 

### 2013-04-05 inc/model_results/introduction.tex
more tweaks to model_results

@@
 This type of pulse is useful to demonstrate general pulse dynamics or the ideal influence of electron-optical elements independent of the pulse's creation conditions.
-Simulations using pulses of this type will provide an initial velocity $v_{\smallzero}$ (or equivalent acceleration voltage $V_{DC}$), an excess photoemission energy $\Delta E$, and HW1/eM dimensions; width $w$ and duration $\tau$ or length $ w_z = v_{\smallzero} \tau$.
+Simulations using pulses of this type will provide an initial velocity $v_{\smallzero}$ (or equivalent acceleration voltage $V_{DC}$), an excess photoemission energy $\Delta E$, and HW1/eM dimensions; width $w_{\smallT}$ and duration $\tau$ or length $ w_z = v_{\smallzero} \tau$.
 The width and length will be used directly to calculate initial spatial variances $\sigma_{\alpha}$.

### 2013-04-05 inc/model_results/spacecharge/spacecharge.tex
more tweaks to model_results

@@
 A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
-While much discussion could be had on the degree of distortion allowable as well as considerations of the initial shape of the pulse and length and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a progressive series of examples.
+While much discussion could be had on the degree of distortion allowable, as well as considerations of the initial shape of the pulse and length, and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a progressive series of examples.
 
@@
 The expected trend is evident in \ref{fig:spacecharge_noacc} which presents the predicted HW1/eM pulse width $\sqrt{2 \sigma_{\smallT}}$ (solid line) and longitudinal length $\sqrt{2 \sigma_{z}}$ (dashed line) after a propagation distance of 15cm as a function of the number of electrons in the pulse.
-Once again, this simulation employs the idealistic pulse described earlier, starting with a velocity of $c/3$ (where $c$ is the speed of light in a vacuum), initial width of 1 mm, HW1/eM pulse duration 1 ps, and an excess photoemission energy $\Delta E = $ 0.5 eV.
+Once again, this simulation employs the idealistic pulse described earlier, starting with a velocity of $c/3$ where $c$ is the speed of light in a vacuum (1.5ns total propagation time), initial width of 1 mm, HW1/eM pulse duration 1 ps, and an excess photoemission energy $\Delta E = $ 0.5 eV.
 For these conditions, used to isolate the behavior of a propagating pulse from its creation environment, the simulations clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.

### 2013-04-05 inc/hardware/laser/laser.tex

@@
 
-The second harmonic generation system now employs a 3mm non-critically phase matched LBO crystal.
+The second harmonic generation system now employs a 3mm non-critically phase matched LBO crystal \cite{website_newlight}.
 This change has two benefits. 
 First, in this geometry, the walk-off angle between the fundamental and second harmonic is zero, while the group velocity mismatch of 150fs ($\sim$ 50 fs/mm) is still less than the pulse duration 250fs.
-%TODO cite Newlight Photonics, Canada
-Second, because in this case, non-critically phase matching requires that the crystal is held at an elevated temperature, well above the boiling point of water, the deleterious absorption of water in the hygroscopic LBO crystal is prevented.
-%TODO replace above with once oven temperature is known
-%Second, because in this case, non-critically phase matching requires that the crystal is held at an elevated temperature of XXX$^{\circ}$C, the deleterious absorption of water in the hygroscopic LBO crystal is prevented.
+Second, because in this case, non-critically phase matching requires that the crystal is held at an elevated temperature of 176$^{\circ}$C, the deleterious absorption of water in the hygroscopic LBO crystal is prevented.
 Additionally, in the place of the curved mirrors originally employed, the second harmonic system now uses 75mm lenses to focus the laser into the crystal and collimate the output.

### 2013-04-05 inc/photocathode/setup/setup.tex
work on short titles for figures (continuing)

@@
 The spatial quality of the sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
+This quantity is conserved under free-space propagation and perfect lensing, and may increase due to lens aberrations \cite{oshea_reversible_1998}.
+As I have discussed in Chapter \ref{chap:considerations}, a electron beam (or in the case of UEM, a pulse) with a lower transverse emittance will result in higher quality images, however since this quantity is conserved, the only chance to minimize this quantity occurs upon emission.
 As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field)\cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness\cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source\cite{dowell_quantum_2009,jensen_emittance_2010}.
@@
   \input{fig1}
-  \caption{
+  \caption[Beam profile simulation of experimental setup]{
     Simulations of the electron pulse propagation through the apparatus using the extended AG model.  
@@
 \ref{fig:transverse-measurement} depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
-The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate (see Section \ref{sec:laser}).
-This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV) (see Section \ref{sec:laser}).
+Though many of the experimental components have been described in detail in Chapter \ref{chap:prototype}, for convenience I will briefly summarize them again here.
+The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate.
+This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV).
 Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
@@
 A YAG scintillator placed 27.5cm after the second magnetic lens detects the spatial electron beam profile, and its visible fluorescence is 1:1 imaged onto a CCD detector with 5.4$\mu$m pixels.
-%TODO internal cross-references to appropriate sections above
 

### 2013-04-05 inc/photocathode/este/este.tex
typo and added este paper to bibtex file

@@
 For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_x = \sqrt{m^* ( \hbar \omega - \phi ) / 3 }$.
-Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$ps, 261nm laser radiation source.
+Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.

### 2013-04-06 inc/model_results/spacecharge/spacecharge.tex
merging more comments into spacecharge results

@@
 The expected trend is evident in \ref{fig:spacecharge_noacc} which presents the predicted HW1/eM pulse width $\sqrt{2 \sigma_{\smallT}}$ (solid line) and longitudinal length $\sqrt{2 \sigma_{z}}$ (dashed line) after a propagation distance of 15cm as a function of the number of electrons in the pulse.
-Once again, this simulation employs the idealistic pulse described earlier, starting with a velocity of $c/3$ where $c$ is the speed of light in a vacuum (1.5ns total propagation time), initial width of 1 mm, HW1/eM pulse duration 1 ps, and an excess photoemission energy $\Delta E = $ 0.5 eV.
-For these conditions, used to isolate the behavior of a propagating pulse from its creation environment, the simulations clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
-In the absence of space-charge effects, the presented oblate shaped pulses should be maintained as such during unforced propagation because the initial values of the electron pulse' divergence and temporal broadening rate are both proportional to $\Delta E^{1/2}$ but not to $N$.
+Once again, this simulation employs the idealistic pulse described earlier, starting with a velocity of $c/3$ where $c$ is the speed of light in a vacuum (1.5ns total propagation time), initial width $w_{\smallT} = $ 1 mm, HW1/eM pulse duration 1 ps ($w_{z} = $ 0.1 mm), and an excess photoemission energy $\Delta E = $ 0.5 eV.
+For these conditions, that are used to isolate the behavior of a propagating pulse from its creation environment, the simulations clearly indicate that both the electron beam width and length at the end of the simulation will remain unchanged below about $10^5$ electrons/pulse; that is, global space-charge effects only significantly influence the pulse propagation for pulses having more than $10^5$ electrons/pulse.
+In the absence of space-charge effects, the presented oblate shaped pulses should be maintained as such during free-space propagation, because the initial values of the electron pulse's divergence and temporal broadening rate are both proportional to $\Delta E^{1/2}$ but not to $N$.
 
@@
 
-Consider now a simulation comparable to the one presented above, but which uses the realistic initial conditions set.
+Consider now a simulation comparable to the one presented above, but which uses the realistic initial conditions set (\ref{eq:summary}).
 The pulse is generated by a 1 mm width 1 ps laser (HW1/eM) and accelerated by an electron gun which is 20mm in length and which accelerates the pulse through a voltage of 30 kV to speed $c/3$, the same velocity as in the idealistic case.

### 2013-04-06 inc/model_results/lens_charge/lens_charge.tex
merge comments on lens_charge section

@@
 
-\section{Space-charge Effects in Magnetic Lenses} \label{sec:mag_lens_charge}
+\section{Space-charge Effects and Spatial Focusing} \label{sec:mag_lens_charge}
 
@@
 In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry.\cite{berger_dc_2009}
-The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity).
+The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity), while still subject to Liouville's Theorem (see Section \ref{sec:liouville}).
+
+\input{lens_charge_plot}
 
@@
 When focusing the pulse it can be shown that oblate pulses will also have markedly better performance.
-
-\input{lens_charge_plot}
-
 \ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for intially oblate ($ \xi ( 0 ) = 0.1 $) or prolate ($ \xi ( 0 ) = 10 $) pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm and a `short' focal length of 6.0mm.
+The pulse's initial volume is constant in all cases ($w_{\smallT}^2 w_{z} = (232 \mu\text{m})^3$).
 For both lens strengths, the deleterious effects of space-charge are evident as the number of electrons/pulse $ N $ is increased.
-To make an effective comparison, highlighting the effect of space-charge alone, this simulation again uses the idealistic initial conditions set; the pulses used are at 20kV velocity with $\Delta E = $ 0.5eV but are not simulated using an accelerator.
-As such the simulation presented approximates a beam which has been perfectly collimated at the front-focal plane of the lens.
+To make an effective comparison, highlighting the effect of space-charge alone, this simulation again uses the idealistic initial conditions set; the pulses used are at velocity associated with a 30kV energy and have momentum variance associated with $\Delta E = $ 0.5eV, but are not simulated using an emitter and accelerator.
+As such the simulation presented approximates a beam which has been nearly perfectly collimated at the focus of a long focal-length lens.
 

### 2013-04-06 inc/model_results/rf_cav/rf_cav.tex
merge comments on rf cav model_results section

--- 2013-04-01 inc/model_results/rf_cav/rf_cav.tex
@@
 
-\section{Space-charge Effects in RF Cavities} \label{sec:rf_cav_charge}
+\section{Compensation of Temporal Space-charge Effects: RF Cavities} \label{sec:rf_cav_charge}
+
+In comparison to using magnetic lenses to compensate for transverse pulse broadening by using magnetic lenses, RF cavities are being employed in UED systems \cite{oudheusden_electron_2007}, and are expected to be employed in UEM systems as well, to compensate for longitudinal broadening in the ultrashort pulses.
+\ref{fig:rf_cav_num} illustrates the operation of a $ \Omega / 2 \pi = $ 3GHz $\text{TM}_{010}$ RF cavity of axial length $ d = \pi v_{\smallzero} / \Omega $ as an electron pulse compressor tuned to produce the shortest pulse at a distance $ z^{\prime} = $ 10cm behind the cavity for $N=1$ and $N=10^5$ pulse charges.
+The mathematical model of the RF cavity (see Section \ref{rf_cav_model}) employs $L_{RF} = d \approx $ 14mm, $ z^{\prime}_{RF} = 0 $, and a region of influence super-Gaussian of order 2 ($n=2$ in \ref{eq:reg_of_influence}) to reflect the sharper ends of the RF cavity.
+In both cases, the idealized initial conditions are used, starting the pulse at $ z^{\prime} = -10$ cm, an excess photoemission energy $\Delta E = 0.5$ eV, velocity related to a $V_{DC} = $ 20kV acceleration potential, HW1/eM $ w = $ 0.1mm and duration $\tau = $ 100fs ($v_{\smallzero} \tau \approx 8.4 \mu$m).
+This electron pulse is therefore initially in the preferred highly oblate regime ($ \sigma_{\smallT}(0) \gg \sigma_{z} (0) $), for which space-charge effects predominantly act in the longitudinal direction, broadening the pulse in time (see \ref{fig:compare_shape}).
 
 \input{rf_cav_plot}
-
-\ref{fig:rf_cav_num} illustrates the operation of a $ \Omega / 2 \pi = $ 3GHz $\text{TM}_{010}$ RF cavity of axial length $ d = \pi v_{\smallzero} / \Omega $ as an electron pulse compressor tuned to produce the shortest pulse at a distance $ z^{\prime} = $ 10cm behind the cavity for $N=1$ and middle $N=10^5$ pulse charge.
-The mathematical model of the RF cavity (see Section \ref{rf_cav_model}) employs $L_{RF} = d \approx $ 14mm, $ z^{\prime}_{RF} = 0 $, and $n=2$ (in \ref{eq:reg_of_influence}) to reflect the sharper end to the region-of-influence of the RF cavity.
-In both cases, the idealized initial conditions are used, starting the pulse at $ z^{\prime} = -10$ cm, an excess photoemission energy $\Delta E = 0.5$ eV, velocity related to a $V_{DC} = $ 20kV acceleration potential, HW1/eM $ w = $ 0.1mm and duration $\tau = $ 100fs ($v_{\smallzero} \tau \approx 8.4 \mu$m).
-This electron pulse is therefore initially in the preferred highly oblate regime ($ \sigma_{\smallT}(0) \gg \sigma_{z} (0) $), for which space-charge effects predominantly act in the longitudinal direction, broadening the pulse in time (see \ref{fig:compare_shape}).
 
@@
 In this case, the pulse is returned to its original temporal duration --- an exact 1-to-1 `temporal imaging' of the photo-generated electron pulse at $z' = -10$cm by the RF cavity to $ z' = $ 10cm.
-On the other hand, the dynamics of the spatial size of the electron pulse are clearly affected by the negative compound transverse lens of the RF cavity, increasing at the temporal focus to a factor of $\sim 10\times$ its initial value.
+On the other hand, the dynamics of the spatial size of the electron pulse are clearly affected by the negative compound transverse lens of the RF cavity (the magnetic contribution of the TM$_{010}$-mode plus the axial entrance and exit aperture electric fields, see Section \ref{rf_cav_model}), increasing at the temporal focus to a factor of $\sim 10\times$ its initial value.
 
@@
 The fact that almost the same RF field strength (495 kV/m) is required to generate a temporal focus at $z^{\prime} = $ 10cm is a direct consequence of the linear dependence of the longitudinal RF cavity force on $z$ (\ref{eq:RF_Force_z}), which means that the required restorative force is not dependent on $\sigma_{z}$ --- just as the focal length of a perfect optical lens is independent of its aperture size.
-However, unlike in the dispersion limited case (\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}), the compressed pulse duration is now a factor of $\sim$2 less than the original 100fs laser pulse duration ($\sim 48$ fs).
+However, unlike in the dispersion limited case (\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}), the compressed pulse duration is now a factor of $\sim$2 less ($\sim 48$ fs) than the original 100fs laser pulse duration.
 This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression.\cite{strickland_compression_1985}
@@
 For a $\sim$10ps electron pulse entering the 3GHz RF cavity we have $ \Omega z / v_{\smallzero} = \Omega \tau \approx 0.19 $, for which the small angle approximation $ \sin ( \Omega z / v_{\smallzero} ) = \Omega z / v_{\smallzero} $ is 0.6\% inaccurate.
-In principle, the AG model could simulate the action of the RF cavity on longer pulse by including additional terms $O \left ( z^{3} \right )$ and higher in \ref{eq:RF_Force_z} resulting from the expansion of $ \sin \left ( \Omega z / v_{0} \right ) $, but at the expense of voiding its self-similar Gaussian approximation.
+In principle, the AG model could simulate the action of the RF cavity on longer pulse by including additional terms $O( z^{3} )$ and higher in \ref{eq:RF_Force_z} resulting from the expansion of $ \sin \left ( \Omega z / v_{0} \right ) $, but at the expense of voiding its self-similar Gaussian approximation.

### 2013-04-06 inc/model_results/compound/compound.tex
merge remaining comments from andreas on the model_results section (compound systems)

@@
 For the transverse dimension, this arrangement of electron optical elements is similar to the Cooke triplet in optics.
-The magnetic lenses are assumed to be identical (but oppositely wound to allow for the compensation of rotational effects in the magnetic focusing through the use of oppositely directed axial magnetic fields).
-The three lensing elements are positioned at a distance of 15cm from the cathode to the front of the first magnetic lens, adjusted to space-time image the electron pulse 5cm behind the second magnetic lens.
-The dependencies of the the spatial variances, $\sigma_{\alpha}$, shown in \ref{fig:compound} as the electron pulse propagates through the multi-element system clearly illustrate the effect of the magnetic lenses and RF cavity in the transverse ($\smallT$) and longitudinal ($z$) dimensions.
+The magnetic lenses are assumed to be identical and wired in series, and thus have equal magnetic field strengths.
+The three lensing elements are positioned at a distance of 15cm from the cathode to the front of the first magnetic lens and are adjusted so as to space-time image the electron pulse 5cm behind the second magnetic lens.
+The dependencies of the the spatial variances, $\sigma_{\alpha}$, shown in \ref{fig:compound} as the electron pulse propagates through the multi-element system, clearly illustrate the effect of the magnetic lenses and RF cavity in the transverse ($\smallT$) and longitudinal ($z$) dimensions.
 As the image-to-object distance ratio is less than unity in this case, both the transverse and longitudinal imaged pulse dimensions are reduced relative to their initial values. %TODO comment futher

### 2013-04-06 inc/extension/external_forces/formal.tex
merged remaining comments from andreas to considerations
* added emittance growth to liouville and formal extension sections
* added explicit mention of \Delta p_T to initial conditions
move punctuation to after citation

@@
 This is because there is no allowance for external forces, such as those imparted by the ``optical'' elements of an electron microscope column.
-In this section, I show that the AG model can be extended to include external forces on the pulse, such as the influence of magnetic electron lenses and RF pulse compression cavities.\cite{oudheusden_electron_2007,veisz_hybrid_2007}
+In this section, I show that the AG model can be extended to include external forces on the pulse, such as the influence of magnetic electron lenses and RF pulse compression cavities \cite{oudheusden_electron_2007,veisz_hybrid_2007}.
 
@@
 While the free-space AG model does not contain any external forces, it does include an internal force: the Coulomb force is included to model the repulsion of the pulse's electrons from each other.
-As this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics.\cite{michalik_analytic_2006}
+As this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics \cite{michalik_analytic_2006}.
 
@@
 
-\subsection{Validity and Limitations}
+\subsection{Validity and Limitations} \label{extension_limitations}
 
-The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations.\cite{michalik_analytic_2006}
+The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}.
 As a result, we emphasize that the presented AG model extension is valid only for the action of linear forces (i.e., forces that have a linear dependence on distance in the pulse frame) on the electron pulse due to the model's self-similar Gaussian approximation.
+Since nonlinear forces are not considered, ``emittance growth'' effects arising from imperfect lenses cannot be modeled \cite{oshea_reversible_1998}.
 The inclusion of the action of nonlinear forces (e.g., due to higher-order space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations.%, which should be employed to gauge the accuracy (and hence limits) of our extended AG approach at some point.
@@
 
-Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid.\cite{luiten_how_2004}
+Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes \cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid \cite{luiten_how_2004}.
 This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
-Consequently, the AG approach is applicable to both the single electron per pulse limit,\cite{lobastov_four-dimensional_2005} where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate.\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}
+Consequently, the AG approach is applicable to both the single electron per pulse limit \cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate \cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
 It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.

### 2013-04-06 inc/extension/external_forces/practical.tex
added some content, but also restructured some of the folders related to hardware
porting rf_cav model_results text from JAP paper
move punctuation to after citation

@@
 
-\subsection{Magnetic Lenses}
+\subsection{Magnetic Lenses} \label{sec:mag_lens_model}
 
@@
 
-\subsection{RF Cavities}
+\subsection{RF Cavities} \label{rf_cav_model}
 
@@
 
-Although there are many possible resonant cavity structures,\cite{oudheusden_electron_2007,humphries_principles_1986} for simplicity, this analysis will consider only the evacuated cylindrical (or `pillbox') $\text{TM}_{010}$ cavity for which the electrons propagate down the axis.\cite{fill_sub-fs_2006,humphries_principles_1986}
-The known mathematical forms of the electric and magnetic components of this resonant mode together with a super-Gaussian envelope function (\ref{eq:reg_of_influence}) allow the three main effects of a pill-box RF cavity element on the electron pulse to be described within the approximation of the linear force extension to the AG model: the primary action of the pulse compression (or acceleration and expansion dependent upon the relative RF phase), and the transverse lens effects due to the magnetic component of the $\text{TM}_{010}$ resonance and the electric RF cavity fringe fields at its entrance and exit apertures.\cite{kim_rf_1989}
+Although there are many possible resonant cavity structures \cite{oudheusden_electron_2007,humphries_principles_1986}, for simplicity, this analysis will consider only the evacuated cylindrical (or `pillbox') $\text{TM}_{010}$ cavity for which the electrons propagate down the axis \cite{fill_sub-fs_2006,humphries_principles_1986}.
+The known mathematical forms of the electric and magnetic components of this resonant mode together with a super-Gaussian envelope function (\ref{eq:reg_of_influence}) allow the three main effects of a pill-box RF cavity element on the electron pulse to be described within the approximation of the linear force extension to the AG model: the primary action of the pulse compression (or acceleration and expansion dependent upon the relative RF phase), and the transverse lens effects due to the magnetic component of the $\text{TM}_{010}$ resonance and the electric RF cavity fringe fields at its entrance and exit apertures \cite{kim_rf_1989}.
 
@@
 The azimuthal magnetic component of the resonant mode oscillates 90$^{\circ}$ out of phase with the electric component and with a $\operatorname{ J_{1} }$ Bessel radial form; i.e., $ \operatorname{ J_{1} }( k r / a ) \sim k r / a $ for small $r$.
-Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $,\cite{kim_rf_1989} where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\ref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\ref{eq:reg_of_influence}).
+Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $ \cite{kim_rf_1989}, where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\ref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\ref{eq:reg_of_influence}).
 In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures (see for example $1 \frac{1}{2}$-cell RF photo-gun designs \cite{mcdonald_design_1988}).
@@
 %TODO examples shown below?
-For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $,\cite{kim_rf_1989} even though its magnitude is maximized for $\phi = 0$.
+For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $ \cite{kim_rf_1989}, even though its magnitude is maximized for $\phi = 0$.
 Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.

### 2013-04-06 inc/extension/initial_conditions/initial_conditions.tex
merged remaining comments from andreas to considerations
* added emittance growth to liouville and formal extension sections
* added explicit mention of \Delta p_T to initial conditions
move punctuation to after citation

@@
 \begin{equation} \label{eq:jensen_eta}
-  \eta_{ \smallT } ( 0 ) \approx \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \punct{,}
+  \eta_{ \smallT } ( 0 )  = (\Delta p_{\smallT})^2 \approx \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \punct{,}
 \end{equation}
@@
 
-We recognize that these initial conditions for an electron pulse generated by a laser-driven DC gun are idealized in that they neglect several important effects; in particular, the interplay of dispersive ($ \eta_{i}(0) \neq 0 $) and space-charge effects in the acceleration region of the photoelectron gun and transverse and longitudinal effects due to the divergence of the DC field in the vicinity of the gun anode.\cite{berger_dc_2009,togawa_ceb6_2007}
+We recognize that these initial conditions for an electron pulse generated by a laser-driven DC gun are idealized in that they neglect several important effects; in particular, the interplay of dispersive ($ \eta_{i}(0) \neq 0 $) and space-charge effects in the acceleration region of the photoelectron gun and transverse and longitudinal effects due to the divergence of the DC field in the vicinity of the gun anode \cite{berger_dc_2009,togawa_ceb6_2007}.
 Such effects and the limits of the validity of the longitudinal terms of \ref{eq:summary}, within the context of the AG model, will be presented in Section \ref{sec:bin_model}.
 However, \ref{eq:summary} also represent a defined set of initial conditions that allows the effect of electron optics (i.e., magnetic lenses and RF cavities) on electron pulse propagation to be more readily understood.
-Moreover, they reflect the fact that the majority of DC photoelectron guns are driven by laser pulses with spatial and temporal profiles that are close to Gaussian.\cite{williamson_clocking_1997,sciaini_electronic_2009}
-Only for specialized cases, such as the generation of uniform ellipsoid electron bunch shapes, is spatial and temporal manipulation of the incident laser pulse considered.\cite{luiten_how_2004,li_generating_2008}
+Moreover, they reflect the fact that the majority of DC photoelectron guns are driven by laser pulses with spatial and temporal profiles that are close to Gaussian \cite{williamson_clocking_1997,sciaini_electronic_2009}.
+Only for specialized cases, such as the generation of uniform ellipsoid electron bunch shapes, is spatial and temporal manipulation of the incident laser pulse considered \cite{luiten_how_2004,li_generating_2008}.
 

### 2013-04-06 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
move punctuation to after citation

@@
 
-We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed et al.\cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the trasverse spatial emittance of an electron beam \cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
+We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed \textit{et al.} \cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the trasverse spatial emittance of an electron beam \cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
 %TODO include more commentary on $\varepsilon_{\smallT}$?

### 2013-04-06 inc/hardware/laser/laser.tex
move punctuation to after citation

@@
 
-In order to experiment using very short laser pulses, we have designed and built a custom Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) laser system.\cite{berger_high-power_2008}
+In order to experiment using very short laser pulses, we have designed and built a custom Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) laser system \cite{berger_high-power_2008}.
 In practice, the flexibility afforded by having a custom-built system has outweighed its original performance benefit.
@@
 
-For most photoemission sources, the fundamental laser photon energy of 1.19 eV does not provide enough energy to excite electrons above the vacuum level, either for single or two-photon emission.\cite{eastman_photoelectric_1970,johnson_optical_1972,dowell_quantum_2009}
-By appropriate application of crystals with strong non-linear properties, two photons at frequency $\omega$ may be converted into one photon at frequency $2 \omega$.\cite{shen_principles_2002}
+For most photoemission sources, the fundamental laser photon energy of 1.19 eV does not provide enough energy to excite electrons above the vacuum level, either for single or two-photon emission \cite{eastman_photoelectric_1970,johnson_optical_1972,dowell_quantum_2009}.
+By appropriate application of crystals with strong non-linear properties, two photons at frequency $\omega$ may be converted into one photon at frequency $2 \omega$ \cite{shen_principles_2002}.
 Since these higher harmonics will be necessary for stimulating photoemission it is important to optimize and characterize the efficiency of this process.

### 2013-04-06 inc/model_results/compound/compound.tex
move punctuation to after citation

@@
 The design of such a compound system is, of course, a cornerstone of modern electron microscope development, although space-charge effects are generally not included, since, on average, there is usually no more than one electron in the microscope column at a time.
-However, for a nanosecond DTEM,\cite{lagrange_nanosecond_2008,reed_evolution_2009} the microscope column must be designed to compensate for transverse ($\smallT$) space-charge effects, which dominate the propagation dynamics for the prolate electron pulses employed in such instruments (see \ref{fig:compare_shape}).
-For the ultrashort (i.e., picosecond and femtosecond) electron pulse durations used in UEMs\cite{park_direct_2009} and UED experiments,\cite{oudheusden_electron_2007} the longitudinal ($z$) pulse dynamics due to both dispersion and space-charge effects must also be considered (as indicated in \ref{fig:compare_shape}).
+However, for a nanosecond DTEM \cite{lagrange_nanosecond_2008,reed_evolution_2009}, the microscope column must be designed to compensate for transverse ($\smallT$) space-charge effects, which dominate the propagation dynamics for the prolate electron pulses employed in such instruments (see \ref{fig:compare_shape}).
+For the ultrashort (i.e., picosecond and femtosecond) electron pulse durations used in UEMs\cite{park_direct_2009} and UED experiments \cite{oudheusden_electron_2007}, the longitudinal ($z$) pulse dynamics due to both dispersion and space-charge effects must also be considered (as indicated in \ref{fig:compare_shape}).
 
@@
 
-\ref{fig:compound} shows an example of a compound element system for the delivery of an ultrashort electron pulse to the specimen area; in this case, optimized for $ N = 10^{ 6 } $ --- the number of electrons per pulse required for single shot diffraction measurements.\cite{armstrong_practical_2007}
+\ref{fig:compound} shows an example of a compound element system for the delivery of an ultrashort electron pulse to the specimen area; in this case, optimized for $ N = 10^{ 6 } $ --- the number of electrons per pulse required for single shot diffraction measurements \cite{armstrong_practical_2007}.
 The displayed dynamics are for an electron pulse generated by a $\tau = $ 100fs $w = $ 0.1mm laser, driving a photocathode, with $\Delta E = $ 0.5eV, then accelerated by 20 kV to a velocity of $8.4 \times 10^{7} \text{m}/\text{s}$.
@@
 At $ N = 10^{6} $, space-charge effects moderate the image demagnification due to the $\sim$2 object-to-image distance ratio in \ref{fig:compound}. 
-In other words, for the displayed compound element system, which could be employed in UED experiments,\cite{oudheusden_electron_2007} the initial space-time resolution associated with the $\sim$0.1mm spot size and $\sim$100fs pulse duration can be maintained (or bettered) for $N \lesssim 10^{6} $.
+In other words, for the displayed compound element system, which could be employed in UED experiments \cite{oudheusden_electron_2007}, the initial space-time resolution associated with the $\sim$0.1mm spot size and $\sim$100fs pulse duration can be maintained (or bettered) for $N \lesssim 10^{6} $.
 For a UEM, on the other hand, it may be necessary to separate longitudinally the space and time images produced by the compound element column.

### 2013-04-06 inc/model_results/lens_charge/lens_charge.tex
move punctuation to after citation

@@
 
-In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses.\cite{oudheusden_electron_2007,lagrange_nanosecond_2008}
-In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry.\cite{berger_dc_2009}
+In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses \cite{oudheusden_electron_2007,lagrange_nanosecond_2008}.
+In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry \cite{berger_dc_2009}.
 The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity), while still subject to Liouville's Theorem (see Section \ref{sec:liouville}).

### 2013-04-06 inc/model_results/rf_cav/rf_cav.tex
move punctuation to after citation

@@
 However, unlike in the dispersion limited case (\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}), the compressed pulse duration is now a factor of $\sim$2 less ($\sim 48$ fs) than the original 100fs laser pulse duration.
-This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression.\cite{strickland_compression_1985}
+This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression \cite{strickland_compression_1985}.
 In this case, the $\sim$20-fold increase in the transverse pulse size (corresponding to a $\sim400\times$ reduced peak pulse charge density) aids the temporal compression by greatly diminishing Coulomb effects that counteract the reduction in electron pulse duration.

### 2013-04-06 inc/model_results/shapes/shapes.tex
move punctuation to after citation

@@
 For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects may dominate the propagation dynamics.
-\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
+\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM \cite{berger_dc_2009,armstrong_practical_2007}.
 The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
@@
 
-The data presented in \ref{fig:compare_shape} using the AG model of Michalik and Sipe \cite{michalik_analytic_2006} are entirely consistent with prior works.\cite{reed_femtosecond_2006,siwick_ultrafast_2002}
-In femtosecond UED experiments, significant efforts are made to offset space-charge induced temporal (longitudinal, $z$) pulse broadening by reducing the time of flight from the gun photocathode to the sample.\cite{siwick_ultrafast_2002,reed_evolution_2009}
-Moreover, efforts are now underway to compensate for temporal electron pulse broadening using RF cavity pulse compression techniques.\cite{oudheusden_electron_2007}
-In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers,\cite{lobastov_four-dimensional_2005} deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
-Even for nanosecond DTEM (the extreme prolate pulse case, $\xi \gg 1 $), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
-In this thesis, however, I will often restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be easily generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}
+The data presented in \ref{fig:compare_shape} using the AG model of Michalik and Sipe \cite{michalik_analytic_2006} are entirely consistent with prior works \cite{reed_femtosecond_2006,siwick_ultrafast_2002}.
+In femtosecond UED experiments, significant efforts are made to offset space-charge induced temporal (longitudinal, $z$) pulse broadening by reducing the time of flight from the gun photocathode to the sample \cite{siwick_ultrafast_2002,reed_evolution_2009}.
+Moreover, efforts are now underway to compensate for temporal electron pulse broadening using RF cavity pulse compression techniques \cite{oudheusden_electron_2007}.
+In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers \cite{lobastov_four-dimensional_2005}, deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
+Even for nanosecond DTEM (the extreme prolate pulse case, $\xi \gg 1 $), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $ \cite{lagrange_nanosecond_2008}.
+In this thesis, however, I will often restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be easily generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening \cite{veisz_hybrid_2007}.

### 2013-04-06 inc/ms_model/ms_model.tex
move punctuation to after citation

@@
 The ``self-similar'' nature of the model, the fact that the overall shape of the pulse cannot change, prevents inclusion of these distortions.
-Even so, the AG model of charge bunch dynamics has been benchmarked against particle tracking simulations for a number of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid which explicitly features a linear internal space-charge field.\cite{luiten_how_2004}
+Even so, the AG model of charge bunch dynamics has been benchmarked against particle tracking simulations for a number of electron pulse shapes \cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid which explicitly features a linear internal space-charge field \cite{luiten_how_2004}.
 %TODO add the error found in these benchmarks

### 2013-04-06 inc/photocathode/setup/setup.tex
move punctuation to after citation

@@
 
-%Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED)\cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers\cite{dowell_cathode_2010}.
-The spatial quality of the sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
+%Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED) \cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers\cite{dowell_cathode_2010}.
+The spatial quality of the sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance \cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
 This quantity is conserved under free-space propagation and perfect lensing, and may increase due to lens aberrations \cite{oshea_reversible_1998}.
@@
 Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface\cite{berger_dc_2009}.
-The electron gun design is based on the that of Togawa et al.\cite{berger_dc_2009,togawa_ceb6_2007} and has an acceleration gap of 25($\pm$2)mm.
+The electron gun design is based on the that of Togawa \textit{et al.} \cite{berger_dc_2009,togawa_ceb6_2007} and has an acceleration gap of 25($\pm$2)mm.
 After acceleration in the DC photo-gun, the electrons pass through a set of electrostatic deflection plates to ensure that the electron beam passes, as close as possible, axially through the center of two, 6.35mm-bore, thin and round magnetic lenses with counter-propagating currents to minimize image rotation effects.

### 2013-04-06 inc/extension/external_forces/formal.tex
added space before \cite where needed

@@
 
-Simulation of electron pulse propagation in UEM and modern UED apparatus\cite{siwick_ultrafast_2002,oudheusden_electron_2007,reed_femtosecond_2006} requires the ability to model electron-optical elements designed to control the three-dimensional space-charge and dispersion induced broadening of the pulse, and hence deliver electron pulses of the desired quality to the specimen.
+Simulation of electron pulse propagation in UEM and modern UED apparatus \cite{siwick_ultrafast_2002,oudheusden_electron_2007,reed_femtosecond_2006} requires the ability to model electron-optical elements designed to control the three-dimensional space-charge and dispersion induced broadening of the pulse, and hence deliver electron pulses of the desired quality to the specimen.
 The Analytic Gaussian model as presented by Michalik and Sipe (see Section \ref{sec:ms_model}) is only useful for modeling free-space propagation.

### 2013-04-06 inc/model_results/compound/compound.tex
added space before \cite where needed

@@
 However, for a nanosecond DTEM \cite{lagrange_nanosecond_2008,reed_evolution_2009}, the microscope column must be designed to compensate for transverse ($\smallT$) space-charge effects, which dominate the propagation dynamics for the prolate electron pulses employed in such instruments (see \ref{fig:compare_shape}).
-For the ultrashort (i.e., picosecond and femtosecond) electron pulse durations used in UEMs\cite{park_direct_2009} and UED experiments \cite{oudheusden_electron_2007}, the longitudinal ($z$) pulse dynamics due to both dispersion and space-charge effects must also be considered (as indicated in \ref{fig:compare_shape}).
+For the ultrashort (i.e., picosecond and femtosecond) electron pulse durations used in UEMs \cite{park_direct_2009} and UED experiments \cite{oudheusden_electron_2007}, the longitudinal ($z$) pulse dynamics due to both dispersion and space-charge effects must also be considered (as indicated in \ref{fig:compare_shape}).
 

### 2013-04-06 inc/model_results/shapes/shapes.tex
added space before \cite where needed

@@
 %In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
-The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.
+The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width \cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.
 

### 2013-04-06 inc/photocathode/este/este.tex
added space before \cite where needed

@@
 The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
-This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water\cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface\cite{eastman_photoelectric_1970}.
-Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function\cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
-Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds\cite{chen_semiclassical_2006}.
+This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water \cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface \cite{eastman_photoelectric_1970}.
+Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function \cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
+Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds \cite{chen_semiclassical_2006}.
 The solid line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
-Using the known optical properties of gold\cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm\cite{seah_quantitative_1979}.
+Using the known optical properties of gold \cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm \cite{seah_quantitative_1979}.
 This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
@@
 Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
-Prior studies\cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included\cite{dowell_quantum_2009}.
+Prior studies \cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included \cite{dowell_quantum_2009}.
 Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
 
-The known optical properties of the two semiconductors\cite{aspnes_dielectric_1983} indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb\cite{chelikowsky_nonlocal_1976} and InSb\cite{chelikowsky_erratum_1984}, is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
+The known optical properties of the two semiconductors \cite{aspnes_dielectric_1983} indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb \cite{chelikowsky_nonlocal_1976} and InSb \cite{chelikowsky_erratum_1984}, is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
 Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m_0$ (0.5$m_0$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV).
 For the $\sim$0.1nJ incident UV pulse energies, $\sim40\mu$m laser spot size, and $p$-polarized reflectance of $\sim$40\% for both semiconductors at our 60$^{\circ}$ incidence angle, the photo-injected electron density is $\sim10^{18}cm^{-3}$.
-At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization\cite{portella_k-space_1992} and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
+At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization \cite{portella_k-space_1992} and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
 
 Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above $\phi$, thereby curtailing the observed ESTE.
-By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band\cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
+By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band \cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
 This means that even without population decay mechanisms the temperature of the exited electron distribution $T_e$ drops at a rate of $\sim$1,600K/ps in both semiconductors.
@@
 \ref{fig:este-semicond} also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
-More significantly, using $\Delta p_x = \sqrt{m_0 k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m_0$ in the semiconductor), simulation of the experiment with the extended AG model\cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\phi/(k_B T_e)] \sim 10^{-15}$).
+More significantly, using $\Delta p_x = \sqrt{m_0 k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m_0$ in the semiconductor), simulation of the experiment with the extended AG model \cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\phi/(k_B T_e)] \sim 10^{-15}$).
 On the other hand, employing a mass of 0.3$m_0$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
 In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (\ref{fig:este-semicond}).
-We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy\cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.
+We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy \cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.
 
@@
 The increased rms transverse momentum for the same ESTE process is consistent with the larger $\sim$0.5$m_0$ effective mass of the upper $\Gamma_8$ conduction band in InSb, the expected higher initial temperature of the electron distribution, and the marginally slower cooling process associated with the lower 24meV LO phonon energy.
-The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation\cite{aspnes_dielectric_1983}.
+The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation \cite{aspnes_dielectric_1983}.
 
@@
 As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_x$ of the electron source clearly must scale with $\sqrt{m^*}$ as observed in the experiment.
-This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies\cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
-We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes\cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m_0$)  to allow for direct emission of electrons excited into the conduction band.
+This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies \cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
+We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes \cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m_0$)  to allow for direct emission of electrons excited into the conduction band.
 For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_x = \sqrt{m^* ( \hbar \omega - \phi ) / 3 }$.

### 2013-04-06 inc/photocathode/setup/setup.tex
added space before \cite where needed

@@
 
-%Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED) \cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers\cite{dowell_cathode_2010}.
+%Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED) \cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers \cite{dowell_cathode_2010}.
 The spatial quality of the sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance \cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
@@
 As I have discussed in Chapter \ref{chap:considerations}, a electron beam (or in the case of UEM, a pulse) with a lower transverse emittance will result in higher quality images, however since this quantity is conserved, the only chance to minimize this quantity occurs upon emission.
-As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field)\cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness\cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source\cite{dowell_quantum_2009,jensen_emittance_2010}.
+As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) \cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness \cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source \cite{dowell_quantum_2009,jensen_emittance_2010}.
 
@@
 Though many of the experimental components have been described in detail in Chapter \ref{chap:prototype}, for convenience I will briefly summarize them again here.
-The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate.
+The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator \cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate.
 This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV).
 Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
-Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface\cite{berger_dc_2009}.
+Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface \cite{berger_dc_2009}.
 The electron gun design is based on the that of Togawa \textit{et al.} \cite{berger_dc_2009,togawa_ceb6_2007} and has an acceleration gap of 25($\pm$2)mm.

### 2013-04-07 inc/photocathode/single_photon/single_photon.tex
added single photon section and basic plot

\section{Single-photon Photoemission from Flat Metal Photocathodes}

\begin{figure}
  \centering
  \input{w_plot}
  \caption[Single-photon photoemission from Tungsten]{
    Single-photon photoemission from Tungsten.
  }
  \label{fig:single_photon_tungsten}
\end{figure}

### 2013-04-07 inc/photocathode/single_photon/analyze.pl

@@ -1,0 +1,214 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+use PDL;
+use PDL::Image2D;
+use PDL::Fit::Gaussian;
+
+use Prima;
+use PDL::Graphics::Prima::Simple;# -sequential;
+
+use File::chdir;
+
+use List::Util;
+
+use constant { 
+  pi => 4*atan2(1,1),
+  wconv => 1 / ( 2 * sqrt( log(2) ) ),
+};
+
+my $folder = shift;
+local $CWD = $folder;
+
+my $cache_name = 'data';
+my $tag = 'fourier';
+my $nd_tag = qr/nd(\d+)/;
+my $handle_backgrounds = 0;
+my $too_large = 900;
+my $laser = sub{ cos(pi / 180 * 2 * shift)**2 };
+
+my $files;
+if ( not -e $cache_name ) {
+  $files = analyze_folder();
+
+  # cache data
+  require Data::Dumper;
+  open my $fh, '>', $cache_name;
+  print $fh Data::Dumper::Dumper($files);
+} else {
+  $files = do $cache_name;
+}
+
+# print for humans
+use Data::Printer;
+p $files;
+
+line_plot(make_piddles($files));
+
+sub analyze_folder {
+
+  # find all fit(s) files
+  my %files = map { $_, { file => $_ } } do {
+    opendir( my $dh, $CWD );
+
+    grep { /\.fits?/ }
+    readdir $dh;
+  };
+
+  # keep only certain tags (e.g. fourier)
+  if ($tag) {
+    my @wrong_tag = grep { not /$tag/i } keys %files;
+    delete @files{@wrong_tag};
+  }
+
+  # remove background images
+  unless ($handle_backgrounds) {
+    my @bg = grep { /bg/ } keys %files;
+    delete @files{@bg};
+  }
+
+  # parse filenames
+  foreach (keys %files) {
+    # polarizer info
+    my $re_angle = /deg/ ? qr/(m?)(\d+)deg/ : qr/(m?)(\d{2,})/;
+
+    if ( $_ =~ $re_angle ) {
+      $files{$_}{polarizer} = $2 * ($1 ? -1 : 1); # m indicates negative number
+      $files{$_}{laser} = $laser->($files{$_}{polarizer});
+    } 
+
+    # nd info
+    if ( $_ =~ $nd_tag ) {
+      $files{$_}{nd} = $1;
+    } else {
[…]

### 2013-04-07 inc/photocathode/single_photon/format_data.pl

@@ -1,0 +1,26 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+my $file = 'data';
+my $data = do $file or die "Cannot load $file";
+
+my $output = 'w_plot_data.dat';
+open my $fh, '>', $output or die "Cannot open $output";
+
+local $, = " ";
+
+my $bg;
+for my $res (sort {$a->{laser} <=> $b->{laser}} values %$data ) {
+
+  # lowest power considered background
+  unless ($bg) {
+    $bg = $res->{yield};
+    next;
+  }
+
+  $res->{yield} -= $bg;
+  print $fh "@{$res}{qw/laser w_av yield/}\n";  
+}
+

### 2013-04-07 inc/photocathode/single_photon/single_photon.tex
more work on single_photon photoemission

@@
 \section{Single-photon Photoemission from Flat Metal Photocathodes}
+
+Single-photon photoemission is the most common photoemission process and the easiest to explain.
+As described by Einstein in his famous paper on ``the photoelectric effect,'' if an electron absorbs a photon whose energy is sufficient to promote the electron above the local vacuum level, it become free to leave the host material.
+In Section \ref{sec:must_reduce_transverse_momentum}, we have already seen that flat metal photocathodes employing single-photon photoemission are not likely to be useful for UEM since the transverse momentum variance $\Delta p_{\smallT} = \sqrt{ m \Delta E / 3 } $ \cite{dowell_quantum_2009} cannot be reduced without also reducing the emission efficiency which is proportional to $\Delta E^2$ \cite{shalaev_electron_1994}.
 
@@
 \end{figure}
+
+An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.55 eV, %TODO ref
+is used as a photocathode.
+When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E = \hbar \omega - \Phi \approx $ 0.2 eV.

### 2013-04-08 inc/photocathode/future.tex
more work on single_photon section

\section{Future Prospects} \label{sec:photocathode_future}

### 2013-04-08 inc/photocathode/single_photon/single_photon.tex
more work on single_photon section

--- 2013-04-07 inc/photocathode/single_photon/single_photon.tex
@@
 \section{Single-photon Photoemission from Flat Metal Photocathodes}
+
+Single-photon photoemission is the most common photoemission process and the easiest to explain.
+As described by Einstein in his famous paper on ``the photoelectric effect'' \cite{einstein_uber_1905}, as it is colloquially known, if an electron absorbs a photon whose energy is sufficient to promote the electron above the local vacuum level, it become free to leave the host material.
 
@@
     Single-photon photoemission from Tungsten.
+    Circles indicate relative electron yield (per pulse), the linear fit line indicates a single-photon process.
+    The Fourier plane spot size (squares), related to the initial transverse momentum variance, does not depend on laser pulse energy, as expected.
   }
@@
 \end{figure}
+
+An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.3-4.5 eV \cite{yen_thermally_1980}, is used as a photocathode.
+When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E = \hbar \omega - \Phi \approx $ 0.2 eV.
+The linear dependence of the electron yield (circles) on the laser pulse energy is indicative of a single photon process, as expected.
+The fact that the Fourier plane spot-size, directly related the initial transverse momentum variance $\Delta p_{\smallT}$, is unchanged over the full range of laser power is also expected, since by $\Delta p_{\smallT} = \sqrt{ m \Delta E / 3 } $ \cite{dowell_quantum_2009}, $\Delta p_{\smallT}$ only depends on $\Delta E$.
+
+In order to improve performance, one must be able to minimize $\Delta p_{\smallT}$ without negatively affecting the photoemission efficiency, and thus electron yield.
+As it stands currently, and as we have already seen in Section \ref{sec:must_reduce_transverse_momentum}, flat metal photocathodes employing single-photon photoemission are not likely to be useful for UEM since the transverse momentum variance cannot be reduced without also reducing the emission efficiency which is proportional to $\Delta E^2$ \cite{shalaev_electron_1994}.
+However, as I will show in Section \ref{sec:photocathode_future}, this does not quite tell the complete story.

### 2013-04-08 inc/photocathode/este/este.tex
Began to split out 2PTE from ESTE section

@@
 % the content of this section is a modification of our APL paper on ESTE
-\section{Excited State Themionic Emission Photocathodes}
+\section{Excited State Thermionic Emission}
 
@@
 Simulations of the experiment using our extended analytical Gaussian (AG) model of electron pulse propagation (see Section \ref{sec:external_forces}) are in close agreement with the data.
-
-\begin{figure}
-  \centering
-  \input{fig2}
-  \caption[Measured Fourier plane electron beam size and yield vs incident laser pulse energy on gold photocathode]{
-    Measured Fourier plane electron beam size (HWe$^{-1}$M; squares) and electron yield (circles) as a function of the incident $\sim$200fs, 523nm laser pulse energy for the 300nm-thick gold photocathode.
-    The electron yield is proportional to the square of laser pulse energy (solid line) and the dependence of the beam size with the laser pulse energy fits that predicted by a zero free parameter electron heating model (solid line; dashed lines represent $\pm$10\% error in the laser spot size).
-    A representative raw Fourier plane beam image is also shown.
-  }
-  \label{fig:gold-emission}
-\end{figure}
-
-To verify the efficacy of the experimental technique (see Section \ref{sec:photocathode-method}) for these purposes, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
-Due to the 60$^{\circ}$ angle of incidence, the circular laser beam is focused by the 30cm focal length lens to an elliptical 50x100$\mu$m (half-width 1/e maximum (HWe$^{-1}$M) of the field) spot size on the photocathode surface.
-The results, the measured Fourier plane beam size and electron yield, are displayed in \ref{fig:gold-emission} as a function of the incident laser pulse energy, corrected to account for 10\% optical loss (mainly from the uncoated vacuum system windows).
-
-The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
-This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water \cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface \cite{eastman_photoelectric_1970}.
-Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function \cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
-Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds \cite{chen_semiclassical_2006}.
-The solid line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
-Using the known optical properties of gold \cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm \cite{seah_quantitative_1979}.
-This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
-The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
-The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m_0$\cite{johnson_optical_1972}.
 

### 2013-04-08 inc/photocathode/two_photon_thermionic/two_photon_thermionic.tex
Began to split out 2PTE from ESTE section

--- 2012-11-14 inc/este/este.tex
@@
 
-% the content of this section is a modification of our APL paper on ESTE
-\section{Excited State Themionic Emission Photocathodes}
-
-%TODO consider moving this paragraph to chapter intro 
-Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED)\cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers\cite{dowell_cathode_2010}.
-The spatial quality of these sources is often characterized in terms of their normalized transverse rms emittance\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
-As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field)\cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness\cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source\cite{dowell_quantum_2009,jensen_emittance_2010}.
-
-In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_x$ developed in Refs.
-4 and 5 for thermionic electron emission should be rewritten to read $\sqrt{m^* k_B T_e}$, where $T_e$ is the temperature of the emitting electron distribution and $k_B$ is Boltzmann's constant.
-Direct observation of the transverse momentum distribution for excited-state thermionic emission (ESTE) from GaSb and InSb, two similar III-V zincblende semiconductors, demonstrates the $m^*$ dependence.
-A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m_0$\cite{johnson_optical_1972}, confirms the temperature dependence of $\Delta p_x$.
-Simulations of the experiment using our extended analytical Gaussian (AG) model of electron pulse propagation (see Section \ref{sec:external_forces}) are in close agreement with the data.
+\section{Two-photon Assisted Thermionic Photoemission} \label{sec:two_photon}
 
@@
   \centering
-  \input{fig1}
-  \caption{
-    Simulations of the electron pulse propagation through the apparatus using the extended AG model.  
-    Solid line: The spatial beam width (normalized to that at the photocathode $\Delta x_0$) with the magnetic lens strengths adjusted to ensure that the detector is at their back focal plane (Fourier plane).
-    Dashed line: The case when the strength of the magnetic lenses is increased by a factor of two to focus the electron beam on the YAG scintillator detector.
-  }
-  \label{fig:este-sim}
-\end{figure}
-
-\ref{fig:este-sim} depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
-The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate (see Section \ref{sec:laser}).
-This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV).
-Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
-Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface\cite{berger_dc_2009}.
-The electron gun design is based on the that of Togawa et al.\cite{berger_dc_2009,togawa_ceb6_2007} and has an acceleration gap of 25($\pm$2)mm.
-After acceleration in the DC photo-gun, the electrons pass through a set of electrostatic deflection plates to ensure that the electron beam passes, as close as possible, axially through the center of two, 6.35mm-bore, thin and round magnetic lenses with counter-propagating currents to minimize image rotation effects.
-A YAG scintillator placed 27.5cm after the second magnetic lens detects the spatial electron beam profile, and its visible fluorescence is 1:1 imaged onto a CCD detector with 5.4$\mu$m pixels.
-
-To ensure that the transverse momentum distribution is observed, the YAG scintillator must be positioned at the back focal plane (the Fourier plane) of the lens system.
-This is readily achieved by appropriate adjustment of the current $I$ in the magnetic lens coils; specifically, after determining the current required to focus the electron beam onto the scintillator, optical imaging relations together with the $I^2$ dependence of the magnetic lens strength are used to find the current that ensures that the average lens to scintillator distance equals the focal length of the lens system.
-Simulation of the experimental apparatus (employing the extended AG model, see Section \ref{sec:external_forces} ), also shown in \ref{fig:este-sim}, confirms that this procedure places the YAG scintillator at the back focal plane of the lens system to within a $\pm$0.5cm error.
-This simulation also confirms that for our experimental conditions (less than 5,000 electrons/pulse and $\Delta x > $ 30$\mu$m) space-charge effects do not influence the observations.
-Additionally, and as expected from optical considerations, the modeling shows that the spatial beam size in the Fourier plane is independent of the incident laser spot size on the photocathode (i.e., the electron source size, $\Delta x_0$).
-
-\begin{figure}
-  \centering
-  \input{fig2}
-  \caption{
+  \input{au_plot}
+  \caption[Measured Fourier plane electron beam size and yield vs incident laser pulse energy on gold photocathode]{
     Measured Fourier plane electron beam size (HWe$^{-1}$M; squares) and electron yield (circles) as a function of the incident $\sim$200fs, 523nm laser pulse energy for the 300nm-thick gold photocathode.
@@
 
-To verify the efficacy of the experimental technique, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
+To verify the efficacy of the experimental technique (see Section \ref{sec:photocathode-method}) for these purposes, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
 Due to the 60$^{\circ}$ angle of incidence, the circular laser beam is focused by the 30cm focal length lens to an elliptical 50x100$\mu$m (half-width 1/e maximum (HWe$^{-1}$M) of the field) spot size on the photocathode surface.
@@
 The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
-This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water\cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface\cite{eastman_photoelectric_1970}.
-Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function\cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
-Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds\cite{chen_semiclassical_2006}.
+This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water \cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface \cite{eastman_photoelectric_1970}.
+Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function \cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
+Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds \cite{chen_semiclassical_2006}.
 The solid line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
-Using the known optical properties of gold\cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm\cite{seah_quantitative_1979}.
-This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:este-sim}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
+Using the known optical properties of gold \cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm \cite{seah_quantitative_1979}.
+This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
 The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
 The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m_0$\cite{johnson_optical_1972}.
-
-Figure 3 depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
-Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
-Prior studies\cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included\cite{dowell_quantum_2009}.
-Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
-
-The known optical properties of the two semiconductors\cite{aspnes_dielectric_1983} indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb\cite{chelikowsky_nonlocal_1976} and InSb\cite{chelikowsky_erratum_1984}, is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
-Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m_0$ (0.5$m_0$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV).
-For the $\sim$0.1nJ incident UV pulse energies, $\sim40\mu$m laser spot size, and $p$-polarized reflectance of $\sim$40\% for both semiconductors at our 60$^{\circ}$ incidence angle, the photo-injected electron density is $\sim10^{18}cm^{-3}$.
-At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization\cite{portella_k-space_1992} and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
-
-Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above $\phi$, thereby curtailing the observed ESTE.
-By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band\cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
-This means that even without population decay mechanisms the temperature of the exited electron distribution $T_e$ drops at a rate of $\sim$1,600K/ps in both semiconductors.
-We note that the close proximity of the lower and unconfined (in momentum space) $\Gamma_7$ upper conduction band to the $\Gamma_8$ band will likely result in a fast population decay, so that the observed ESTE process should have an intrinsic latency of much less than the $\sim$4ps UV laser pulse duration.
-
-Figure 3 also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
-More significantly, using $\Delta p_x = \sqrt{m_0 k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m_0$ in the semiconductor), simulation of the experiment with the extended AG model\cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\phi/(k_B T_e)] \sim 10^{-15}$).
-On the other hand, employing a mass of 0.3$m_0$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
-In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (Figure 3).
-We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy\cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.
-
-Very similar results are observed with the $p$-type InSb photocathode except that the Fourier plane spot size is 30-40\%  larger and the electron yield is a factor of 1.7 higher.
-The increased rms transverse momentum for the same ESTE process is consistent with the larger $\sim$0.5$m_0$ effective mass of the upper $\Gamma_8$ conduction band in InSb, the expected higher initial temperature of the electron distribution, and the marginally slower cooling process associated with the lower 24meV LO phonon energy.
-The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation\cite{aspnes_dielectric_1983}.
-
-The dependence of $\Delta p_x$ on $m^*$ is readily explained through a consideration of energy and momentum conservation in transmission across a boundary with a potential step associated with the effective photoelectric work function $\phi$.
-Inside the material, an electron with an energy $E$ greater than $\phi$ has a maximum momentum of $\sqrt{2 m^* (E-\phi) }$ parallel to the boundary if it is to be emitted.
-As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_x$ of the electron source clearly must scale with $\sqrt{m_*}$ as observed in the experiment.
-This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies\cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
-We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes\cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m_0$)  to allow for direct emission of electrons excited into the conduction band.
-For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_x = \sqrt{m^* ( \hbar \omega - \phi ) / 3 }$.
-Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$ps, 261nm laser radiation source.

### 2013-04-08 inc/photocathode/future.tex
working on cleaning up the splitting out of parts of este section

@@
 \section{Future Prospects} \label{sec:photocathode_future}
+
+A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m_0$\cite{johnson_optical_1972}, confirms the temperature dependence of $\Delta p_x$.
+ and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m_0$\cite{johnson_optical_1972}

### 2013-04-08 inc/photocathode/two_photon_thermionic/two_photon_thermionic.tex
working on cleaning up the splitting out of parts of este section

@@
 
-\section{Two-photon Assisted Thermionic Photoemission} \label{sec:two_photon}
+\section{Two-photon Assisted Thermionic Photoemission} \label{sec:two_photon_thermionic}
 
@@
 
-To verify the efficacy of the experimental technique (see Section \ref{sec:photocathode-method}) for these purposes, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
+Again using the experimental technique outlined in Section \ref{sec:photocathode-method}, we monitored the momentum distribution (Fourier plane spot size) for electrons emitted by 2PTE from gold using the $\sim$200fs green (523nm) $p$-polarized laser pulses.
 Due to the 60$^{\circ}$ angle of incidence, the circular laser beam is focused by the 30cm focal length lens to an elliptical 50x100$\mu$m (half-width 1/e maximum (HWe$^{-1}$M) of the field) spot size on the photocathode surface.
@@
 The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
-The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m_0$\cite{johnson_optical_1972}.
+The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism.

### 2013-04-08 inc/photocathode/este/este.tex
working on cleaning up the splitting out of parts of este section
improve symbolic consistency between este paper derived sections and the rest of the paper

@@
 
-In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_x$ developed in Refs.
+In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_{\smallT}$ developed in Refs.
 \cite{dowell_quantum_2009} and \cite{jensen_emittance_2010} for thermionic electron emission should be rewritten to read $\sqrt{m^* k_B T_e}$, where $T_e$ is the temperature of the emitting electron distribution and $k_B$ is Boltzmann's constant.
 Direct observation of the transverse momentum distribution for excited-state thermionic emission (ESTE) from GaSb and InSb, two similar III-V zincblende semiconductors, demonstrates the $m^*$ dependence.
-A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m_0$\cite{johnson_optical_1972}, confirms the temperature dependence of $\Delta p_x$.
 Simulations of the experiment using our extended analytical Gaussian (AG) model of electron pulse propagation (see Section \ref{sec:external_forces}) are in close agreement with the data.
@@
 The known optical properties of the two semiconductors \cite{aspnes_dielectric_1983} indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb \cite{chelikowsky_nonlocal_1976} and InSb \cite{chelikowsky_erratum_1984}, is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
-Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m_0$ (0.5$m_0$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV).
+Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m$ (0.5$m$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV).
 For the $\sim$0.1nJ incident UV pulse energies, $\sim40\mu$m laser spot size, and $p$-polarized reflectance of $\sim$40\% for both semiconductors at our 60$^{\circ}$ incidence angle, the photo-injected electron density is $\sim10^{18}cm^{-3}$.
-At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization \cite{portella_k-space_1992} and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
+At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization \cite{portella_k-space_1992} and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\Phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\Phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
 
-Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above $\phi$, thereby curtailing the observed ESTE.
+Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above $\Phi$, thereby curtailing the observed ESTE.
 By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band \cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
@@
 \ref{fig:este-semicond} also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
-More significantly, using $\Delta p_x = \sqrt{m_0 k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m_0$ in the semiconductor), simulation of the experiment with the extended AG model \cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\phi/(k_B T_e)] \sim 10^{-15}$).
-On the other hand, employing a mass of 0.3$m_0$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
-In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (\ref{fig:este-semicond}).
-We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\phi$ = 4.31eV and assuming $\Delta p_x = \sqrt{m_0 ( \hbar \omega - \phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy \cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.
+More significantly, using $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m$ in the semiconductor), simulation of the experiment with the extended AG model \cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\Phi/(k_B T_e)] \sim 10^{-15}$).
+On the other hand, employing a mass of 0.3$m$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
+In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\Phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (\ref{fig:este-semicond}).
+We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\Phi$ = 4.31eV and assuming $\Delta p_{\smallT} = \sqrt{m ( \hbar \omega - \Phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy \cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.
 
 Very similar results are observed with the $p$-type InSb photocathode except that the Fourier plane spot size is 30-40\%  larger and the electron yield is a factor of 1.7 higher.
-The increased rms transverse momentum for the same ESTE process is consistent with the larger $\sim$0.5$m_0$ effective mass of the upper $\Gamma_8$ conduction band in InSb, the expected higher initial temperature of the electron distribution, and the marginally slower cooling process associated with the lower 24meV LO phonon energy.
+The increased rms transverse momentum for the same ESTE process is consistent with the larger $\sim$0.5$m$ effective mass of the upper $\Gamma_8$ conduction band in InSb, the expected higher initial temperature of the electron distribution, and the marginally slower cooling process associated with the lower 24meV LO phonon energy.
 The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation \cite{aspnes_dielectric_1983}.
 
-The dependence of $\Delta p_x$ on $m^*$ is readily explained through a consideration of energy and momentum conservation in transmission across a boundary with a potential step associated with the effective photoelectric work function $\phi$.
-Inside the material, an electron with an energy $E$ greater than $\phi$ has a maximum momentum of $\sqrt{2 m^* (E-\phi) }$ parallel to the boundary if it is to be emitted.
-As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_x$ of the electron source clearly must scale with $\sqrt{m^*}$ as observed in the experiment.
+The dependence of $\Delta p_{\smallT}$ on $m^*$ is readily explained through a consideration of energy and momentum conservation in transmission across a boundary with a potential step associated with the effective photoelectric work function $\Phi$.
+Inside the material, an electron with an energy $E$ greater than $\Phi$ has a maximum momentum of $\sqrt{2 m^* (E-\Phi) }$ parallel to the boundary if it is to be emitted.
+As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_{\smallT}$ of the electron source clearly must scale with $\sqrt{m^*}$ as observed in the experiment.
 This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies \cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
-We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes \cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m_0$)  to allow for direct emission of electrons excited into the conduction band.
-For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_x = \sqrt{m^* ( \hbar \omega - \phi ) / 3 }$.
+We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes \cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m$)  to allow for direct emission of electrons excited into the conduction band.
+For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
 Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.

### 2013-04-08 inc/photocathode/future.tex
improve symbolic consistency between este paper derived sections and the rest of the paper

--- 2013-04-08 inc/photocathode/future.tex
@@
 \section{Future Prospects} \label{sec:photocathode_future}
+
+A further experiment employing two-photon assisted thermionic emission (2PTE) from gold, for which $m^* = m$ \cite{johnson_optical_1972}, confirms the temperature dependence of $\Delta p_{\smallT}$.
+ and, as expected, that the effective mass of a free electron in Au is equal to the rest mass $m$ \cite{johnson_optical_1972}

### 2013-04-08 inc/photocathode/setup/setup.tex
improve symbolic consistency between este paper derived sections and the rest of the paper

@@
 %Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED) \cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers \cite{dowell_cathode_2010}.
-The spatial quality of the sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance \cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_x \Delta x / (m_0 c)$, where $\Delta p_x = \sqrt{ \langle p_x^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m_0$ is the electron rest mass and $c$ is the speed of light in vacuum.
+The spatial quality of the sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance \cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_{\smallT} \Delta x / (m c)$, where $\Delta p_{\smallT} = \sqrt{ \langle p_{\smallT}^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m$ is the electron rest mass and $c$ is the speed of light in vacuum.
 This quantity is conserved under free-space propagation and perfect lensing, and may increase due to lens aberrations \cite{oshea_reversible_1998}.
 As I have discussed in Chapter \ref{chap:considerations}, a electron beam (or in the case of UEM, a pulse) with a lower transverse emittance will result in higher quality images, however since this quantity is conserved, the only chance to minimize this quantity occurs upon emission.
-As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) \cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness \cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_x$ of the electron source --- an intrinsic property of the emission source \cite{dowell_quantum_2009,jensen_emittance_2010}.
+As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) \cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness \cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_{\smallT}$ of the electron source --- an intrinsic property of the emission source \cite{dowell_quantum_2009,jensen_emittance_2010}.
 
@@
 
-\ref{fig:transverse-measurement} depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_x$) of the laser-driven electron sources.
+\ref{fig:transverse-measurement} depicts the experimental technique used to determine directly the transverse momentum distributions (and hence $\Delta p_{\smallT}$) of the laser-driven electron sources.
 Though many of the experimental components have been described in detail in Chapter \ref{chap:prototype}, for convenience I will briefly summarize them again here.

### 2013-04-08 inc/photocathode/two_photon_thermionic/two_photon_thermionic.tex
improve symbolic consistency between este paper derived sections and the rest of the paper

@@
 The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
-This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water \cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface \cite{eastman_photoelectric_1970}.
-Moreover, when $2\hbar \omega \approx \phi$, the dependence of the yield on the square of the electron's excess energy above the work function \cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
-Consequently, the observed increase in $\Delta p_x$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds \cite{chen_semiclassical_2006}.
+This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\Phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water \cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface \cite{eastman_photoelectric_1970}.
+Moreover, when $2\hbar \omega \approx \Phi$, the dependence of the yield on the square of the electron's excess energy above the work function \cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
+Consequently, the observed increase in $\Delta p_{\smallT}$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds \cite{chen_semiclassical_2006}.
 The solid line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
 Using the known optical properties of gold \cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm \cite{seah_quantitative_1979}.
-This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_x = \sqrt{m_0 k_B T_e}$.
+This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_{\smallT} = \sqrt{m k_B T_e}$.
 The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.

### 2013-04-08 inc/photocathode/este/este.tex
moved more mentions of non-este m* from other sections to future

@@
 % the content of this section is a modification of our APL paper on ESTE
-\section{Excited State Thermionic Emission}
+\section{Excited State Thermionic Emission} \label{sec:este}
 
@@
 We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes \cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m$)  to allow for direct emission of electrons excited into the conduction band.
-For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
-Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.

### 2013-04-08 inc/photocathode/future.tex
moved more mentions of non-este m* from other sections to future

--- 2013-04-08 inc/photocathode/future.tex
@@
 \section{Future Prospects} \label{sec:photocathode_future}
+
+The philosophical argument presented in Section \ref{sec:este} for using the effective mass $m^*$ in the expressions for the transverse momentum variance $\Delta p_{\smallT}$ simply follows from conservation of energy and momentum at the barrier and actually does not depend on the ``excited state'' nature nor the thermionic final emission process.
+The earlier experiment employing two-photon assisted thermionic emission (2PTE) from gold (Section \ref{sec:two_photon_thermionic}), the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to the rest mass $m$ \cite{johnson_optical_1972}.
+
+For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
+Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.
+In an upcoming publication, we will show that the dependence of $\Delta p_{\smallT}$ on the electron effective mass may be extended to simple planar metal photocathodes --- thus potentially allowing for the future development or discovery of a robust, high brightness photocathode material.
+The simulation is in good agreement with our direct measurements of $\Delta p_{\smallT}$ for Bi(111) and polycrystalline Ag, Mo, Ta, Be, and W photocathodes.
+We will also discuss the effect of laser-induced heatingon $\Delta p_{\smallT}$ of the electron distribution in the photocathode during photoemission, which may limit the attainable pulsed electron source brightness at the desired high peak pulse current densities.

### 2013-04-08 inc/photocathode/single_photon/format_data.pl
added errorbars to plot, though they aren't very representative

@@ -5,6 +5,8 @@ use warnings;
 
 my $file = 'data';
 my $data = do $file or die "Cannot load $file";
+my $pulse_energy = 0.2;
+my $err = 0.1;
 
 my $output = 'w_plot_data.dat';
 open my $fh, '>', $output or die "Cannot open $output";
@@ -21,6 +23,9 @@ for my $res (sort {$a->{laser} <=> $b->{laser}} values %$data ) {
   }
 
   $res->{yield} -= $bg;
-  print $fh "@{$res}{qw/laser w_av yield/}\n";  
+  my $w = $res->{w_av};
+  my $yield = $res->{yield};
+
+  print $fh join( " ", $res->{laser} * $pulse_energy, $w, $w*$err, $yield, $yield*$err ) . "\n";  
 }
 

### 2013-04-08 inc/photocathode/single_photon/format_data.pl
added error bars to yield which are a quarter of the background

@@ -2,30 +2,32 @@
 
 use strict;
 use warnings;
+use List::Util 'max';
 
 my $file = 'data';
 my $data = do $file or die "Cannot load $file";
 my $pulse_energy = 0.2;
-my $err = 0.1;
 
 my $output = 'w_plot_data.dat';
 open my $fh, '>', $output or die "Cannot open $output";
 
 local $, = " ";
 
-my $bg;
+my $w_err = 0.1;
+
+my ($bg, $yield_err);
 for my $res (sort {$a->{laser} <=> $b->{laser}} values %$data ) {
 
   # lowest power considered background
   unless ($bg) {
     $bg = $res->{yield};
+    $yield_err = $bg / 4;
     next;
   }
 
-  $res->{yield} -= $bg;
   my $w = $res->{w_av};
-  my $yield = $res->{yield};
+  my $yield = $res->{yield} - $bg;
 
-  print $fh join( " ", $res->{laser} * $pulse_energy, $w, $w*$err, $yield, $yield*$err ) . "\n";  
+  print $fh join( " ", $res->{laser} * $pulse_energy, $w, $w*$w_err, $yield, $yield_err ) . "\n";  
 }
 

### 2013-04-08 inc/photocathode/setup/setup.tex
tweak the introduction to photocathode engineering

@@
 The spatial quality of the sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance \cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_{\smallT} \Delta x / (m c)$, where $\Delta p_{\smallT} = \sqrt{ \langle p_{\smallT}^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m$ is the electron rest mass and $c$ is the speed of light in vacuum.
-This quantity is conserved under free-space propagation and perfect lensing, and may increase due to lens aberrations \cite{oshea_reversible_1998}.
-As I have discussed in Chapter \ref{chap:considerations}, a electron beam (or in the case of UEM, a pulse) with a lower transverse emittance will result in higher quality images, however since this quantity is conserved, the only chance to minimize this quantity occurs upon emission.
-As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) \cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness \cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_{\smallT}$ of the electron source --- an intrinsic property of the emission source \cite{dowell_quantum_2009,jensen_emittance_2010}.
+This quantity is conserved under free-space propagation and perfect lensing, but may increase due to lens aberrations \cite{oshea_reversible_1998}.
+As I have discussed in Chapter \ref{chap:considerations}, an electron beam (or in the case of UEM, a pulse) with a lower transverse emittance will result in higher quality images, however since this quantity is conserved, the only chance to minimize this quantity occurs upon emission.
+As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited incident laser pulse focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) \cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness \cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_{\smallT}$ of the electron source --- an intrinsic property of the emission source \cite{dowell_quantum_2009,jensen_emittance_2010}.
+
+To address this need, I have experimented with several candidate photocathodes, which belong to several different categories of photoemission processes.
+In this chapter, I will present an example of each of these different processes, and discuss its prospects for reducing the transverse emittance and therefore utility as an electron source for UEM.
 

### 2013-04-08 inc/photocathode/setup/setup.tex
tweaks to the methodology section of pe chapter

@@
 %Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED) \cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers \cite{dowell_cathode_2010}.
-The spatial quality of the sources for UEM and UED are often characterized in terms of their normalized transverse rms emittance \cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_{\smallT} \Delta x / (m c)$, where $\Delta p_{\smallT} = \sqrt{ \langle p_{\smallT}^2 \rangle }$ is the rms transverse momentum of the source and $\Delta x = \sqrt{ \langle x^2 \rangle }$  is its size (cylindrical symmetry assumed), and $m$ is the electron rest mass and $c$ is the speed of light in vacuum.
+The spatial quality of the sources for UEM and UED are often characterized in terms of their emitted electron beam's normalized transverse rms emittance \cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_{\smallT} \Delta x / (m c)$, where $\Delta p_{\smallT} = \sqrt{ \langle p_{\smallT}^2 \rangle }$ is the rms transverse momentum of the beam, $\Delta x = \sqrt{ \langle x^2 \rangle }$ is its size (cylindrical symmetry assumed), and $m$ is the electron rest mass and $c$ is the speed of light in vacuum.
 This quantity is conserved under free-space propagation and perfect lensing, but may increase due to lens aberrations \cite{oshea_reversible_1998}.
 As I have discussed in Chapter \ref{chap:considerations}, an electron beam (or in the case of UEM, a pulse) with a lower transverse emittance will result in higher quality images, however since this quantity is conserved, the only chance to minimize this quantity occurs upon emission.
-As a reduction in the source size $\Delta x$ is limited either by the attainable diffraction-limited incident laser pulse focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) \cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness \cite{berger_dc_2009}) may only be possible by decreasing the rms transverse momentum $\Delta p_{\smallT}$ of the electron source --- an intrinsic property of the emission source \cite{dowell_quantum_2009,jensen_emittance_2010}.
+As a reduction in the source size $\Delta x_0$ is limited either by the attainable diffraction-limited incident laser pulse focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) \cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness \cite{berger_dc_2009}) may only be possible by decreasing the initial rms transverse momentum $\Delta p_{\smallT}$ of the electron source --- an intrinsic property of the emission source \cite{dowell_quantum_2009,jensen_emittance_2010}.
 
@@
 This is readily achieved by appropriate adjustment of the current $I$ in the magnetic lens coils; specifically, after determining the current required to focus the electron beam onto the scintillator, optical imaging relations together with the $I^2$ dependence of the magnetic lens strength are used to find the current that ensures that the average lens to scintillator distance equals the focal length of the lens system.
-Simulation of the experimental apparatus (employing the extended AG model, see Section \ref{sec:external_forces} ), also shown in \ref{fig:transverse-measurement}, confirms that this procedure places the YAG scintillator at the back focal plane of the lens system to within a $\pm$0.5cm error.
-This simulation also confirms that for our experimental conditions (less than 5,000 electrons/pulse and $\Delta x > $ 30$\mu$m) space-charge effects do not influence the observations.
-Additionally, and as expected from optical considerations, the modeling shows that the spatial beam size in the Fourier plane is independent of the incident laser spot size on the photocathode (i.e., the electron source size, $\Delta x_0$).
+Simulation of the experimental apparatus (employing the extended AG model, see Section \ref{sec:external_forces} ), confirms that this procedure places the YAG scintillator at the back focal plane of the lens system to within a $\pm$0.5cm error (\ref{fig:transverse-measurement}).
+This simulation also confirms that for our experimental conditions (less than 5,000 electrons/pulse and $\Delta x_0 > $ 30$\mu$m) space-charge effects do not influence the observations.
+Additionally, and as expected from optical considerations, the modeling shows that the spatial beam size in the Fourier plane is independent of $\Delta x_0$, the rms laser spot size incident on the photocathode (i.e., the electron source size).

### 2013-04-08 inc/photocathode/single_photon/single_photon.tex
include a note on relative yield
Began to split out 2PTE from ESTE section
tweaks to single_photon section

@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \section{Single-photon Photoemission from Flat Metal Photocathodes}
@@
 As described by Einstein in his famous paper on ``the photoelectric effect'' \cite{einstein_uber_1905}, as it is colloquially known, if an electron absorbs a photon whose energy is sufficient to promote the electron above the local vacuum level, it become free to leave the host material.
+The material dependent minimum energy needed to eject an electron is termed its ``work function'' $\Phi$.
+Any additional energy above the work function remains with the ejected electron, usually in the form of kinetic energy, and is called the ``excess photoemission energy.''
 
@@
 An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.3-4.5 eV \cite{yen_thermally_1980}, is used as a photocathode.
-When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E = \hbar \omega - \Phi \approx $ 0.2 eV.
-The linear dependence of the electron yield (circles) on the laser pulse energy is indicative of a single photon process, as expected.
-The fact that the Fourier plane spot-size, directly related the initial transverse momentum variance $\Delta p_{\smallT}$, is unchanged over the full range of laser power is also expected, since by $\Delta p_{\smallT} = \sqrt{ m \Delta E / 3 } $ \cite{dowell_quantum_2009}, $\Delta p_{\smallT}$ only depends on $\Delta E$.
+When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E = \hbar \omega - \Phi \approx $ 0.35 eV.
+The linear dependence of the electron yield (circles) on the laser pulse energy is indicative of a single photon process, as expected since each absorbed photon may eject one electron.
+In this experiment, the yield is determined by integrating the total counts recorded on the CCD detector and therefore is only a relative measurement.
+The fact that the Fourier plane spot-size (squares), directly related the initial transverse momentum variance $\Delta p_{\smallT}$, is unchanged over the full range of laser power is also expected, since by $\Delta p_{\smallT} = \sqrt{ m \Delta E / 3 } $ \cite{dowell_quantum_2009}, $\Delta p_{\smallT}$ only depends on $\Delta E$.
 

### 2013-04-08 inc/photocathode/single_photon/analyze.pl
clean-up binning problem for single_photon plot

@@ -169,13 +169,14 @@ sub xy_fit {
   my $x_lineout = $pdl->slice(",($y)");
   my $y_lineout = $pdl->slice("($x),");
 
+  # pixels are binned in twos
   my (undef, undef, $fwhm_x) = fitgauss1d(
-    $x_lineout->sequence * 5.4,
+    $x_lineout->sequence * 2 * 5.4,
     $x_lineout,
   );
 
   my (undef, undef, $fwhm_y) = fitgauss1d(
-    $y_lineout->sequence * 5.4,
+    $y_lineout->sequence * 2* 5.4,
     $y_lineout,
   );
 

### 2013-04-08 inc/photocathode/single_photon/format_data.pl

@@ -25,7 +25,7 @@ for my $res (sort {$a->{laser} <=> $b->{laser}} values %$data ) {
     next;
   }
 
-  my $w = $res->{w_av};
+  my $w = $res->{w_av} / 1000;
   my $yield = $res->{yield} - $bg;
 
   print $fh join( " ", $res->{laser} * $pulse_energy, $w, $w*$w_err, $yield, $yield_err ) . "\n";  

### 2013-04-08 inc/photocathode/single_photon/single_photon.tex
clean-up binning problem for single_photon plot
more tweaks to single_photon

@@
     Single-photon photoemission from Tungsten.
-    Circles indicate relative electron yield (per pulse), the linear fit line indicates a single-photon process.
-    The Fourier plane spot size (squares), related to the initial transverse momentum variance, does not depend on laser pulse energy, as expected.
+    The good linear fit (green) to the relative electron yield (squares) indicates a single-photon process.
+    The Fourier plane spot size (squares), related to the initial transverse momentum variance, do not depend on laser pulse energy, as expected.
   }

### 2013-04-08 inc/photocathode/este/este.tex
fixed binning problem with este plot (and cleaned things up a bit while I was at it)

@@
   \centering
-  \input{fig3}
+  \input{este_plot}
   \caption[Electron emission vs UV laser pulse energy for GaSb and InSb]{
-    Number of electrons emitted per pulse as a function of the incident 261nm, UV laser pulse energy for GaSb (open squares) and InSb (filled circles): Linear efficiency dependences are shown by the solid lines.
-    The laser pulse energy invariance of the Fourier plane electron beam size (HWe$^{-1}$M right axis; open squares) for GaSb and a representative raw Fourier plane beam image is also shown.
+    Number of electrons emitted per pulse as a function of the incident 261nm, UV laser pulse energy for GaSb (filled circles) and InSb (open circles): Linear efficiency dependences are shown by the solid lines.
+    The laser pulse energy invariance of the HWe$^{-1}$M Fourier plane electron beam size (filled squares) for GaSb and a representative raw Fourier plane beam image is also shown.
   }

### 2013-04-08 inc/photocathode/two_photon_thermionic/two_photon_thermionic.tex
fix binning factor of two in 2PTE plot
began tweaking 2PTE section

@@
 
-\section{Two-photon Assisted Thermionic Photoemission} \label{sec:two_photon_thermionic}
+\section{Two-photon Assisted Thermionic Emission} \label{sec:two_photon_thermionic}
+
+Although single-photon photoemission is more efficient than multi-photon photoemission, the concept remains the same; an electron in the photocathode must absorb enough energy to exceed the vacuum level.
+In multi-photon photoemission this energy is attained by absorbing multiple photons.
+As this is more likely to occur when more photons are present, the efficiency, and thus the yield, is proportional to the laser intensity to the power of the number of photons required.
+Two-photon photoemission, therefore, will have yield that is quadratic in laser intensity ($I^2$).
+
+Two-photon assisted thermionic emission (2PTE) occurs when two photons provide nearly enough energy to emit a cold electron, and therefore a hotter electron, in the tail of the Fermi distribution, may have enough energy to be emitted \cite{yen_thermally_1980}.
+Of course this is even less efficient than two-photon photoemission, but as three-photon photoemission goes as $I^3$, 2PTE will still be the dominant photoemission mechanism.
+For thermionic emission from electrons at a temperature $T_e$, the rms transverse momentum $\Delta p_{\smallT}$ is given by 
+\begin{equation}
+  \Delta p_{\smallT} = \sqrt{m k_B T_e} \,\text{,}
+\end{equation}
+where $k_B$ is the Boltzmann constant \cite{dowell_quantum_2009,jensen_emittance_2010}.
+Since a high laser intensity is required to drive 2PTE, the laser itself will heat the electrons, therefore we expect the measured Fourier plane spot size to grow as the square root of the laser pulse energy.
 

### 2013-04-09 inc/photocathode/este/este.tex
more tweaks to the photocathode sections

@@
 Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
+Note that due to the surprisingly large yield, we repeated the measurements, this time replacing the the scintillator screen with a Faraday cup and repeating the laser energy sequence.
+In this way, the yields for these photocathodes were measured directly and absolutely, unlike the results presented in Sections \ref{sec:single_photon} and \ref{sec:two_photon_thermionic}, for which relative measurements were sufficient.
 
@@
 On the other hand, employing a mass of 0.3$m$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
-In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to 108 electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\Phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (\ref{fig:este-semicond}).
+In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to $10^8$ electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\Phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (\ref{fig:este-semicond}).
 We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\Phi$ = 4.31eV and assuming $\Delta p_{\smallT} = \sqrt{m ( \hbar \omega - \Phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy \cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.

### 2013-04-09 inc/photocathode/future.tex
more tweaks to the photocathode sections

@@
 
-The philosophical argument presented in Section \ref{sec:este} for using the effective mass $m^*$ in the expressions for the transverse momentum variance $\Delta p_{\smallT}$ simply follows from conservation of energy and momentum at the barrier and actually does not depend on the ``excited state'' nature nor the thermionic final emission process.
-The earlier experiment employing two-photon assisted thermionic emission (2PTE) from gold (Section \ref{sec:two_photon_thermionic}), the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to the rest mass $m$ \cite{johnson_optical_1972}.
+The philosophical argument presented in Section \ref{sec:este} for using the effective mass $m^*$ in the expressions for the transverse momentum variance $\Delta p_{\smallT}$ simply follows from conservation of energy and momentum at the barrier and actually does not depend on the ``excited state'' nature nor the thermionic emission process.
+For the example two-photon assisted thermionic emission (2PTE) from gold presented in Section \ref{sec:two_photon_thermionic}, the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to the rest mass $m$ \cite{johnson_optical_1972}.
 
@@
 In an upcoming publication, we will show that the dependence of $\Delta p_{\smallT}$ on the electron effective mass may be extended to simple planar metal photocathodes --- thus potentially allowing for the future development or discovery of a robust, high brightness photocathode material.
-The simulation is in good agreement with our direct measurements of $\Delta p_{\smallT}$ for Bi(111) and polycrystalline Ag, Mo, Ta, Be, and W photocathodes.
+For a selection of single-crystal and polycrystalline photocathodes, we will show that the simulation, incorporating $m^*$, is in good agreement with our direct measurements of $\Delta p_{\smallT}$.
 We will also discuss the effect of laser-induced heatingon $\Delta p_{\smallT}$ of the electron distribution in the photocathode during photoemission, which may limit the attainable pulsed electron source brightness at the desired high peak pulse current densities.

### 2013-04-09 inc/photocathode/two_photon_thermionic/two_photon_thermionic.tex
more tweaks to the photocathode sections

@@
     Measured Fourier plane electron beam size (HWe$^{-1}$M; squares) and electron yield (circles) as a function of the incident $\sim$200fs, 523nm laser pulse energy for the 300nm-thick gold photocathode.
-    The electron yield is proportional to the square of laser pulse energy (solid line) and the dependence of the beam size with the laser pulse energy fits that predicted by a zero free parameter electron heating model (solid line; dashed lines represent $\pm$10\% error in the laser spot size).
+    The electron yield is proportional to the square of laser pulse energy (solid green line) and the dependence of the beam size with the laser pulse energy fits that predicted by a zero free parameter electron heating model (solid red line; dashed lines represent $\pm$10\% error in the laser spot size).
     A representative raw Fourier plane beam image is also shown.
@@
 Consequently, the observed increase in $\Delta p_{\smallT}$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds \cite{chen_semiclassical_2006}.
-The solid line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
+The solid red line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
 Using the known optical properties of gold \cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm \cite{seah_quantitative_1979}.
 This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_{\smallT} = \sqrt{m k_B T_e}$.
-The dashed lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
+The dashed red lines indicate the expected trends for a $\pm$10\% error in the laser spot size $\Delta x_0$ incident on the gold cathode surface.
 The close agreement between the data and the simulation strongly supports our interpretation of the emission mechanism.

### 2013-04-09 inc/photocathode/plasmon/plasmon.tex
merging comments from andreas on PAPE section

@@
 As the name suggests, a plasmon is the quasiparticle of plasma oscillation.
-When this occurs at the surface of a material, the resultant quasiparticles are called ``surface plasmon-polaritons.''
-Since these will be the only plasmons discussed in this paper, the word ``plasmon'' will be used interchangeably for the name ``surface plasmon-polariton.''
-%TODO should this be formalized, either one way (SPP) or the other
-Electrons in a material can couple with electromagnetic fields, in this case a laser field, and oscillate accordingly.
-This overall oscillation of electrons near the surface of the material creates strong oscillatory local surface electric field \cite{cottam_introduction_2004,concepts_2002}.
+When this occurs at the surface of a conducting material, the resultant quasiparticles are called ``surface plasmon-polaritons.''
+Since these will be the only plasmons discussed in this thesis, the word ``plasmon'' will be used interchangeably for the name ``surface plasmon-polariton.''
+Electrons in a conductor can couple with electromagnetic fields, in this case a laser field, and oscillate accordingly.
+This overall oscillation of electrons near the surface of the conductor creates strong oscillatory local surface electric field \cite{cottam_introduction_2004,concepts_2002}.
 In PAPE, the goal is use this field to both enhance the emission and `control' the divergence of photoemitted electrons through the combined action of photoemission barrier suppression and ponderomotive acceleration.
+On the other hand, these oscillating electrons are nothing but an oscillating current, and thus if a large plasmon response is established, resistive heating is a major concern even for small the small resistivities typical of conductors.
 
@@
 This arrangement is called the Kretschmann geometry.
-While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, we felt that for UEM applications we should explore the ``grating coupling'' geometry.
+While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, we felt that for UEM applications we should explore the ``grating coupling'' geometry \cite{kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
 In this geometry, the plasmon is driven on a periodic surface rather than a flat surface attached to glass.
-Futher, in this geometry the back surface is now free to be attached to a heat sink.
+Futher, in this front-side illuminated geometry the back surface is now free to be attached to a heat sink.
 
@@
 
-Every material has an intrinsic plasma frequency given by 
+Every conducting material has an intrinsic plasma frequency given by 
 \begin{equation}
@@
 \end{equation}
-where $\rho_{e}$ is the electron density and $m*$ is the effective mass of the material.
+where $\rho_{e}$ is the free-electron density and $m^*$ is the electron effective mass in the material.
 And in the most simple (undamped) approximations, the dielectric function of the material may be written in terms of this constant
@@
 
-At the interface of a vacuum and such a simple material material, contstraints on the material parameters which may allow surface plasmons result in the condition that the dielectric constant at the oscillation frequency of the surface plasmon $ \varepsilon(\omega_{SP}) = -1 $.
-Substituting this result into \ref{eq:dielectric_as_omega} show that this frequency of oscillation is simply related to intrinsic plasma frequency by
+At the interface of a vacuum and such a simple material, contstraints on the material parameters which may allow surface plasmons, result in the condition that the dielectric constant at the oscillation frequency of the surface plasmon $ \varepsilon(\omega_{SP}) = -1 $ \cite{cottam_introduction_2004}.
+Substituting this result into \ref{eq:dielectric_as_omega} shows that this frequency of oscillation is simply related to intrinsic plasma frequency by
 \begin{equation}
@@
 \end{equation}
-Further, plasmons carry momentum whose magnitude is given by
-\begin{equation}
-  k_{P}^2 = \frac{\omega^2}{c^2} \left( \frac{\varepsilon_{1} \varepsilon_{2}}{\varepsilon_{1} + \varepsilon_{2}} \right) \,\text{,}
-\end{equation}
-but of course for a vacuum interface this is simplified to 
+Further, plasmons propagating along a vacuum interface carry momentum whose magnitude is given by 
 \begin{equation}
@@
 In the grating coupling scheme (shown in \ref{fig:plasmon_schematic}) a laser of wavelength $\lambda$ is incident on a material with a grating of period $d$, making an angle $\theta$ from the surface normal.
-A surface plasmon is driven when the portion of the laser wave-vector parallel to the plasmon ($k_{Lx}$) and a fourier component ($k_{G} = 2 \pi / d$) of the grating matches the wave-vector of the surface plasmon
+A surface plasmon is driven when the sum of the component of the laser wave-vector parallel to the material surface ($k_{Lx}$) and a fourier component ($k_{G} = 2 \pi / d$) of the grating matches the wave-vector of the surface plasmon
 \begin{equation}
@@
 where $n$ here is an integer.
-Which reduces to 
+This expression reduces to 
 \begin{equation} \label{eq:plasmon_condition}
-  \sqrt{ \frac{ \varepsilon }{ 1 + \varepsilon } } = \sin \theta + \frac{ n \lambda }{ d } \,\text{.}
+  \sqrt{ \frac{ \varepsilon }{ 1 + \varepsilon } } = \sin \theta + \frac{ n \lambda }{ d } \,\text{,}
 \end{equation}
+where $\lambda$ is the incident laser wavelength.
 For a given grating period, \ref{eq:plasmon_condition} determines the angle of incidence that the laser must make with the surface in order to drive the plasmon.
-Further, it is used to design the grating, allowing for angles that the microscope column will accept.
+Alternatively, it may be used to design the periodic surface structure, for example allowing for angles that the microscope column will accept, different driving orders ($n$) and even other Fourier components of the periodic surface structure.
 
@@
 Like the Kretschmann geometry, a foil on glass grating is vulnerable to heating problems; a plasmon creates a large amount of heat but the glass is a poor heat conductor.
-The aforementioned laser (emitting $\sim$200fs pulses of green light at 30MW/cm$^2$) was used to drive the plasmon.
-
-The input laser angle was slowly changed until the plasmon oscillation is visible.
-At the instant that the plasmon started, a bright flash was visible on the scintillator screen then disappeared.
-After this, the grating showed visual signs of damage, both on the surface and in the reflected light pattern.
-\ref{fig:grating-damage} shows the results of Scanning Electron Microscopy (SEM) on the grating.
-The damage pattern is round and of the size of the laser spot used to drive the plasmon.
-The extent of the damage indicates a large temperature increase, which given the other evidence of the test, can be assumed to be from the sudden appearance of a surface plasmon.
-
-There are several conclusions that can be drawn from this experiment.
-Most importantly, it shows that a plasmon was driven successfully given the angular dependence and large response, few other processes could have occurred.
-The heat generated by the plasmon oscillation destroyed the grating, therefore it validated the decision not to pursue a Kretschmann geometry over fears of heat damage at these high powers.
-Finally, it suggests that if using a thermally conductive and structurally sound photoemission material and mount for the plasmonic photocathode, PAPE might generate a large number of electrons (as evident by the bright flash).
+The aforementioned laser (emitting $\sim$200fs pulses of green light at a peak pulse intensity of 30MW/cm$^2$) was used to drive the plasmon.
 
@@
 
+The input laser angle was slowly changed until the plasmon oscillation is visible.
+At the instant that the plasmon started, a bright flash was visible on the scintillator screen and then disappeared.
+After this, the grating showed visual signs of damage, both on the surface and in the reflected light pattern.
+\ref{fig:grating-damage} shows the results of Scanning Electron Microscopy (SEM) on the grating.
+The damage pattern is round and of the size of the laser spot used to drive the plasmon.
+The extent of the damage indicates a large temperature increase, which given the other evidence of the test, can be assumed to be from the sudden appearance of a surface plasmon.
+
+There are several conclusions that can be drawn from this experiment.
+Most importantly, it shows that a plasmon was driven successfully given the angular dependence and large response, few other processes could have occurred.
+The heat generated by the plasmon oscillation destroyed the grating, therefore it validated the decision not to pursue a Kretschmann geometry over fears of heat damage at the high heating rates of the laser-plasmon interaction.
+Finally, it suggests that if using a thermally conductive and structurally sound photoemission material and mount for the plasmonic photocathode, PAPE might generate a large number of electrons (as evident by the bright flash).
+
 \subsection{Sinusoidal Grating}
 
-After the destruction of the gold foil on glass grating, new gratings were designed to have a higher thermal conductivity.
+After the destruction of the gold foil on glass grating, new gratings were designed to have a higher subtrate thermal conductivity.
 The second generation grating was constructed from a silicon substract then coated with a gold film.
 The thermal conductivity of silicon is approximately 100 times that of glass. %TODO reference needed?
-
-Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division employed a defocused Gallium focused ion beam (FIB) (Zeiss 1540XB) to iteratively mill an approximately sinusoidal shape into the surface of a silicon wafer.
-Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
-By \ref{eq:plasmon_condition}, given a laser wavelength of 523nm ($\varepsilon_{Au} \approx -3.95$ \cite{johnson_optical_1972}), this choice of periodicity corresponds to an incidence angle of $\sim39^{\circ}$, which is within the acceptance range of our experimental setup (35-75$^{\circ}$) (see Section \ref{sec:gun_design}).
-An SEM image of one such region is shown in \ref{fig:fib-si-sem}.
 
@@
 
+Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division employed a defocused Gallium focused ion beam (FIB) (Zeiss 1540XB) to iteratively mill an approximately sinusoidal shape into the surface of a silicon wafer.
+Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
+By \ref{eq:plasmon_condition}, given a laser wavelength of 523nm ($\varepsilon_{Au} \approx -3.95$ \cite{johnson_optical_1972}), this choice of periodicity corresponds to an incidence angle of $\sim39^{\circ}$, which is within the acceptance range of our experimental setup (35-75$^{\circ}$) (see Section \ref{sec:gun_design}).
+An SEM image of one such region is shown in \ref{fig:fib-si-sem}.
 A 300nm film of gold was deposited onto the silicon using the Varian e-beam deposition chamber at the UIC Nano Core Facility (NCF); a thin layer of chromium was deposited between the silicon and gold as a binding agent.

### 2013-04-09 inc/photocathode/este/este.tex
merge most of the remaining comments from andreas on the PAPE section

@@
 Note that due to the surprisingly large yield, we repeated the measurements, this time replacing the the scintillator screen with a Faraday cup and repeating the laser energy sequence.
-In this way, the yields for these photocathodes were measured directly and absolutely, unlike the results presented in Sections \ref{sec:single_photon} and \ref{sec:two_photon_thermionic}, for which relative measurements were sufficient.
+In this way, the yields for these photocathodes were measured directly and absolutely, unlike the results presented in the preceeding sections, for which relative measurements were sufficient.
 

### 2013-04-09 inc/photocathode/plasmon/plasmon.tex
merge most of the remaining comments from andreas on the PAPE section

@@
 % Content taken from 2010 ANL report
-Note that for the following measurement, relative emission efficiencies are calculated from the integration of the luminosity detected on the YAG scintillator screen.
-The scintillator and CCD camera detection scheme was also shown to be linear (i.e., integrated signal proportional to number of electrons per pulse) for at least $10^4$ electrons per pulse, using the known linearity of single-photon photoemission from Ta irradiated by our 261nm ultraviolet laser source.
-As a result, the detection scheme will accurately determine any laser power law dependence to the electron emission from the photocathode.
-
+\ref{fig:pape} summarizes the photoemission results obtained from this gold-coated periodic photocathode under green (523nm) laser irradiation.
 Strong enhanced phototemission was observed for only a narrow $\pm$5mrad range of incidence angles $\theta$ around the expected plasmon resonance angle of 39$^{\circ}$.
 Moreover, this enhanced laser-driven electron emission was only observed for $p$-polarized incident 523nm femtosecond laser radiation.
-These two properties are known characteristics of plasmon-assisted photoemission.
+These two properties are known characteristics of plasmon-assisted photoemission \cite{kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
 
-As seen in \ref{fig:pape}, at low incident laser intensities, below about 25MW/cm$^2$, the emission from the nano-patterned photocathode at the plasmon resonance condition is very similar to that observed from a flat gold surface neighboring the FIB-milled region.
+As seen in \ref{fig:pape}, at low incident laser intensities, below a laser intensity of about 25MW/cm$^2$, the emission from the nano-patterned photocathode at the plasmon resonance condition is very similar to that observed from a flat gold surface neighboring the FIB-milled region.
 Specifically, the strength of both emissions have a quadratic dependence on the incident laser intensity.
@@
 
-Above $\sim$32MW/cm$^2$, there is a rapid and dramatic increase in the emission efficiency for the nano-patterned photocathode.
+I note here that for these measurements, the relative emission efficiencies are calculated from the integration of the luminosity detected on the YAG scintillator screen.
+The scintillator and CCD camera detection scheme was shown to be linear (i.e., integrated signal proportional to number of electrons per pulse) for at least $10^4$ electrons per pulse, using the known linearity of single-photon photoemission from Ta irradiated by our 261nm ultraviolet laser source.
+%TODO introduce this linearity earlier?
+As a result, the detection scheme will accurately determine any laser power law dependence to the electron emission from the photocathode.
+
+Above an incident laser pulse intensity of $\sim$32MW/cm$^2$, there is a rapid and dramatic increase in the emission efficiency for the nano-patterned photocathode.
 This can be explained by a suppression of the barrier to photoemission below the two-photon excitation energy, thus allowing direct two-photon photoemission.
@@
 
-In principle this unexpected increase in transverse momentum could be caused by any of three effects.
+In principle, this unexpected increase in transverse momentum could be caused by any of three effects.
 The first, and most easily ruled out, is that the electric field of the laser itself could affect the electrons, causing the divergence.
@@
   \caption[Comparison of the colinear and rotated plasmonic excitation geometries]{
-    Comparison of the colinear (above) and rotated (below) plasmonic excitation geometries.
+    Comparison of the collinear (above) and rotated (below) plasmonic excitation geometries.
     This rotation separates the directionality of effects caused by the laser, plasmon field and grating.
@@
 
-Given the geometry of the previous measurement, notably the colinear arrangement of plasmon vector and the grating vector, the true cause is difficult to determine.
+Given the geometry of the previous measurement, notably the collinear arrangement of plasmon vector and the grating vector, the true cause is difficult to determine.
 However, if instead the grating is rotated --- and the laser incidence angle is corrected to account for the longer effective periodicity of the surface --- the directionality of these effects may be separated.
-\ref{fig:rotated} shows the comparison of the grating vector geometries for the original scheme (top) and one in which the grating is rotated by $45^{\circ}$ (bottom).
+The left side of \ref{fig:rotated} shows the comparison of the grating vector geometries for the original scheme (top) and one in which the grating is rotated by $45^{\circ}$ (bottom).
 For this rotation, the laser angle must be increased from $39^{\circ}$ to $47^{\circ}$.
 
-The results are shown to the right in \ref{fig:rotated}.
+The experimental results are shown to the right in \ref{fig:rotated}.
 As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
 This result is commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
-As that geometry does not employ a patterned surface, it would is not subject to such local field enhancements of the accelerating electric field.
+As that geometry does not employ a patterned surface, it would not be subject to such local field enhancements of the accelerating electric field.
 
@@
 Though the cause of the failure may never be known for certain, several causes are suspected.
-Now that the surface is not purely sinusoidal, the surface has multiple Fourier components.
+Now that the surface is not purely sinusoidal, so that the surface has multiple Fourier components.
 This may mean that the laser may not couple as efficiently to the plasmon, requiring increased laser power for oscillation. 
-Additionally, the non-smooth surface may cause increaded difficulty for propagating the plasmon for large distances.
-Finally, upon further investigation of the surface by SEM, evidence seen of incomplete filling of the gold into the trenches, which may have interrupted creation or propagation of the plasmon.
+Additionally, the non-smooth surface may cause increased difficulty for propagating the plasmon for large distances.
+Finally, upon further investigation of the surface by SEM, evidence was observed of incomplete filling of the gold coating into the trenches, which may have interrupted creation or propagation of the plasmon.
 
@@
 Although plasmonic photocathodes are still an interesting field of research in photocathode physics, we have concluded that they are not a promising candidate for an electron source for UEM at this time.
-They have been difficult to manufacture and thus would be expensive for owners.
+They have been difficult to manufacture and thus would likely be expensive.
 They are very intolerant to misalignment and would therefore be problematic for users.
@@
 
-Other nano-engineered surfaces, including quantum-wells, are still being considered for future work.
+Other nano-engineered surfaces are still being considered for future work.
+Various other more exotic nano-structures are under examination for PAPE \cite{li_surface_2013}.
+Bi-metallic layers are known to enhance PAPE, which may allow for reduced incident laser power for driving the plasmon \cite{kupersztych_anomalous_2005}.
+Photoemission from nano-arrays of quantum-wells also offers the tantalizing hope of reduced rms transverse momentum of the generated electrons.

### 2013-04-09 inc/photocathode/plasmon/plasmon.tex
Merge remaining comments from andreas on PAPE

@@
 This can be explained by a suppression of the barrier to photoemission below the two-photon excitation energy, thus allowing direct two-photon photoemission.
-This is likely caused by the laser-excited surface plasmon since a similar increase from the neighboring flat gold surface is not present.
-The observed distinct changes in the momentum distributions (\ref{fig:pape} B and C) in this laser intensity range support this argument.
+The well-known Schottky Effect states that for an applied electric field $E$ the work function will be suppressed by an amount
+\begin{equation}
+  \Delta \Phi = - e\sqrt{\frac{e E}{4 \pi \varepsilon_{\smallzero}}} \,\text{.}
+\end{equation}
+This suppression is likely caused by the laser-excited surface plasmon, since on the neighboring flat gold surface, where the suppression due to the laser (20MV/m) and the 20kV acceleration potential (0.8MV/m) is $\Delta \Phi \approx $ 0.18eV, no similar increase in efficiency is observed.
+If this is indeed the case, and the surface plasmon field futher suppresses the barrier to a total $ \Delta \Phi \approx $ 0.35eV to allow direct two-photon photoemission, the field must be of the order of 60MV/m or three times that of the laser field.
 In \ref{fig:pape} B, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves).
@@
 
-In principle, this unexpected increase in transverse momentum could be caused by any of three effects.
+In principle, this unexpected increase in transverse momentum (\ref{fig:pape} B and C) could be caused by any of three effects.
 The first, and most easily ruled out, is that the electric field of the laser itself could affect the electrons, causing the divergence.
@@
 As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
-This result is commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
+For a mathematical treatment of the transverse electric field of such patterned surfaces, the reader is referred to Ref. \cite{watts_sharp_1997}. 
+Our results are commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
 As that geometry does not employ a patterned surface, it would not be subject to such local field enhancements of the accelerating electric field.
@@
 Other nano-engineered surfaces are still being considered for future work.
-Various other more exotic nano-structures are under examination for PAPE \cite{li_surface_2013}.
+Various other more exotic nano-structures are under examination for PAPE \cite{li_surface_2013,polyakov_plasmon_2013}.
 Bi-metallic layers are known to enhance PAPE, which may allow for reduced incident laser power for driving the plasmon \cite{kupersztych_anomalous_2005}.

### 2013-04-09 inc/photocathode/single_photon/single_photon.tex
more tweaks to the photocathode sections
Merge remaining comments from andreas on PAPE

@@
 
-\section{Single-photon Photoemission from Flat Metal Photocathodes}
+\section{Single-photon Photoemission from Flat Metal Photocathodes} \label{sec:single_photon}
 
@@
 The material dependent minimum energy needed to eject an electron is termed its ``work function'' $\Phi$.
-Any additional energy above the work function remains with the ejected electron, usually in the form of kinetic energy, and is called the ``excess photoemission energy.''
+Any additional energy above the work function remains with the ejected electron, usually in the form of kinetic energy, and is called the ``excess photoemission energy'' $\Delta E$, for single photon photoemission, using a laser of frequency $\omega$ (photon energy $\hbar \omega$), this is simply,
+\begin{equation}
+  \Delta E = \hbar \omega - \Phi \,\text{.}
+\end{equation}
+The efficiency of photoemission $\eta_{{\scriptscriptstyle PE}}$, when $\Delta E$ is small compared to $\Phi$ may be written as,
+\begin{equation}
+  \eta_{{\scriptscriptstyle PE}} = A ( 1 - R ) \Delta E^2 \,\text{,}
+\end{equation}
+where $A$ is a material paramenter, and R is the photocathode surface reflectivity.
 
@@
 An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.3-4.5 eV \cite{yen_thermally_1980}, is used as a photocathode.
-When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E = \hbar \omega - \Phi \approx $ 0.35 eV.
+When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E \approx $ 0.35 eV.
 The linear dependence of the electron yield (circles) on the laser pulse energy is indicative of a single photon process, as expected since each absorbed photon may eject one electron.
@@
 In order to improve performance, one must be able to minimize $\Delta p_{\smallT}$ without negatively affecting the photoemission efficiency, and thus electron yield.
-As it stands currently, and as we have already seen in Section \ref{sec:must_reduce_transverse_momentum}, flat metal photocathodes employing single-photon photoemission are not likely to be useful for UEM since the transverse momentum variance cannot be reduced without also reducing the emission efficiency which is proportional to $\Delta E^2$ \cite{shalaev_electron_1994}.
+For single-photon photoemission, although the polarization-dependent reflectivity can be tweaked to improve performance slightly \cite{berger_dc_2009}, due to the quadratic dependence of the efficiency on excess photoemission energy \cite{shalaev_electron_1994}, any reduction in $\Delta p_{\smallT}$ gained by reducing $\Delta E$ will be mitigated by a large drop in electron yield.
+For this reason, as it stands currently and as we have already seen in Section \ref{sec:must_reduce_transverse_momentum}, flat metal photocathodes employing single-photon photoemission are not likely to be useful for UEM.
 However, as I will show in Section \ref{sec:photocathode_future}, this does not quite tell the complete story.

### 2013-04-09 inc/photocathode/este/este.tex
more tweaks to photocathode section

@@
 
+Recent experiments in our group have come to change our understanding of several photoemission processes, once thought to be well understood \cite{berger_excited_2012}.
+Indeed while this knowledge has been commonplace in other fields, such as angle-resolved photoemission spectroscopy (ARPES) \cite{himpsel_angle-resolved_1983}, to our knowledge the effect has not been considered for electron beam generation.
 In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_{\smallT}$ developed in Refs.

### 2013-04-09 inc/photocathode/future.tex
more tweaks to photocathode section

@@
 
-For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
+For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum (\ref{eq:jensen_eta}, Refs.
+\cite{dowell_quantum_2009} and \cite{jensen_emittance_2010}) to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
 Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.

### 2013-04-09 inc/photocathode/setup/setup.tex
more tweaks to photocathode section

@@
 
-To address this need, I have experimented with several candidate photocathodes, which belong to several different categories of photoemission processes.
-In this chapter, I will present an example of each of these different processes, and discuss its prospects for reducing the transverse emittance and therefore utility as an electron source for UEM.
+To address this need, I have experimented with many candidate photocathodes, which implement a variety of different photoemission processes.
+In this chapter, I will present an example of several of these different processes, discuss its prospects for reducing the transverse emittance and finally assess its utility as an electron source for UEM.
+I will then outline some future research directions that the group at UIC will continue to pursue.
+Since this is the most active current field in our group's reseach, some very new results are presented near the end of the chapter and future plans are evolving.
 

### 2013-04-09 inc/ms_model/ms_model.tex
killing TODOs

@@
 
-This section introduces the Analytic Gaussian model of Michalik and Sipe \cite{michalik_analytic_2006}.
-%TODO more here (obviously), comparison with GPT, extended in ch. 3, etc
+This section introduces the Analytic Gaussian (AG) model of Michalik and Sipe \cite{michalik_analytic_2006}.
+In the AG model the dynamics of a Gaussian electon pulse are computed using a mean-field self-similar approximation.
+The result is a set of equations that can be solved very quickly using numerical methods on a modern computer.
+When checked against N-body simulations, the authors find that the model is also surprisingly accurate, even for a large set of non-Gaussian pulses \cite{michalik_analytic_2006,michalik_evolution_2009}.
+
+The only force included in the AG model is that of the internal repulsive space-charge field.
+In Chapter \ref{chap:extension}, I will extend this model, adding the influence of external forces, making it more generally applicable for UEM.
 
@@
 Even so, the AG model of charge bunch dynamics has been benchmarked against particle tracking simulations for a number of electron pulse shapes \cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid which explicitly features a linear internal space-charge field \cite{luiten_how_2004}.
-%TODO add the error found in these benchmarks

### 2013-04-10 inc/design_considerations/considerations.tex
bugfix
more work on single_photon photoemission
killing more TODOs

@@
 In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effects that column elements have upon it.
-However, microscopy is an experimental science, and designing a UEM is no less hands-on.
-%TODO avoid repitition? "When doing so, special ..."?
-When designing a UEM system, special care must be given to the creation and propagation of the electron beam.
-%TODO define single-shot and transverse emittance here or rephrase?
-In the case of single-shot UEM, several factors lead to two major considerations: first, that a large area photoemission source is required to smaller area or even single point emitters, and second, the transverse emittance of the beam must be minimized.
+However, since microscopy is an experimental science, we are also designing and building a prototype UEM system at UIC.
+In doing so we have decided that special care must be given to the creation and propagation of the electron beam.
+In the case of single-shot UEM, in which the image of the sample is acquired using only one electron pulse, several factors lead to two major considerations: first, that a large area photoemission source is required to smaller area or even single point emitters, and second, that the ``normalized transverse emittance,'' a beam quality figure of merit which will be defined later, must be minimized.
 
@@
 
-\section{Beam Quality and Resolution Limit}
+\section{Beam Quality and Resolution Limit} \label{sec:must_reduce_transverse_momentum}
 
 A common figure of merit for electron beams is its normalized transverse emittance \cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, defined as 
-\begin{equation} \label{eq:emittance_definition}
+\begin{equation}
   \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > - \left < x p_x \right >^2 } \,\text{,}

### 2013-04-10 inc/extension/external_forces/formal.tex
killing more TODOs

@@
 Since nonlinear forces are not considered, ``emittance growth'' effects arising from imperfect lenses cannot be modeled \cite{oshea_reversible_1998}.
-The inclusion of the action of nonlinear forces (e.g., due to higher-order space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations.%, which should be employed to gauge the accuracy (and hence limits) of our extended AG approach at some point.
-%TODO consider adding a replacement sentance like: "It would be interesting to employ ..." or "A future goal is to ..."
+The inclusion of the action of nonlinear forces (e.g., due to higher-order space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations.
 

### 2013-04-10 inc/extension/external_forces/practical.tex
killing more TODOs

@@
 To model electron pulses traveling through realistic electron microcope columns, these abstract mechanisms must be developed further to properly describe each specific electron-optical element.
-In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression and acceleration %TODO refs
+In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression \cite{oudheusden_electron_2007} and acceleration %TODO refs
 ) and DC acceleration regions (i.e. electron gun dynamics).
@@
 
-%TODO perhaps this shouldn't be a redefinition, but some other function
 Of course, physical electron-optical elements occupy only a finite space in the column, and thus their influence on the pulse should be limited to a region near the element itself.
@@
 A simple magnetic lens is just a solenoid whose axis is collinear with the axis of the microscope column.
-%Initially one would think that electrons moving parallel to this axis would not be affected by the magnetic field lines in the solenoid, however the contribution of the fringing field must be taken into consideration.
 As electrons propagate through the solenoid the fringing magnetic field imparts an azimuthal velocity.
@@
 \end{subequations}
-%%TODO I can't decide whether the Q and k definitions should be numbered or not
-%where
-%\begin{gather*}
-%  Q \equiv (a+r)^2 + z^2 \\
-%  k \equiv \sqrt{ 4 a r / Q }
-%\end{gather*}
 where $\mu$ is the local magnetic permability, $H$ is the auxiliary field, and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
@@
 to the set of $M_{\smallT}$ parameters, where $z_{mag}^{\prime}$ and $L_{mag}$ are the position and length of the lens in question respectively.
-%TODO include discussion and plot of beam size versus aberrations
 
@@
 In fact, when the rays are weighted by a Gaussian of width $w$ the minimum ``radius'' of the beam is 100$\mu$m, while the AG model has a minimum HW1/eM width of 73$\mu$m.
-(The kinematic model actually overestimates the width as there are two rays at the outer convergence zone for every one which converges at the axis.) %TODO footnote?
+(The kinematic model actually overestimates the width as there are two rays at the outer convergence zone for every one which converges at the axis.)
 
@@
 In obtaining \ref{eq:RF_Force_r}, which is also of the form for use in \ref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \ref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
-%TODO examples shown below?
-For the examples shown below under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $ \cite{kim_rf_1989}, even though its magnitude is maximized for $\phi = 0$.
+As will be seen in Section \ref{sec:rf_cav_charge}, under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $ \cite{kim_rf_1989}, even though its magnitude is maximized for $\phi = 0$.
 Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.
@@
 should be added to the equation of motion governing the propagation of the pulse down the column.
-% excluded example section from JAP paper as it will need new graphics

### 2013-04-10 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
killing more TODOs

@@
 Further, when viewed in comparison to the chirps $\gamma_{\alpha}$, one can now see that these $\Gamma_{\alpha}$ constants can be thought of as some minimum/intrinsic action of the pulse.
-This will become more apparent when, in future sections,
-%TODO include internal reference
-it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.
+%This will become more apparent when, in future sections, it is shown that at the focal point of a lens, $\gamma_{\alpha}$ goes through zero.
+This becomes apparent when noting that if a momentum chirp $\gamma_{\alpha}$ should ever reach zero (e.g.
+as it does at the focal point of a lens) there is still a non-Coulombic term driving the $\frac{d\gamma_{\alpha}}{dt}$, as there surely must be when the local momentum variance $\eta_{\alpha}$ is not zero.
 
 We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed \textit{et al.} \cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the trasverse spatial emittance of an electron beam \cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
-%TODO include more commentary on $\varepsilon_{\smallT}$?
 Additionally, though the AG model explicitly satisfies Liouville's Theorem, there are cases where the pulse's phase space product may change in a realistic system.

### 2013-04-10 inc/ms_model/ms_model.tex
adding lots of punctuation around equations

@@
 \begin{equation}
-N = \int\limits_{\text{all space}} \!\!\! f(\vec{x}) \; d^{n}x
+  N = \int\limits_{\text{all space}} \!\!\! f(\vec{x}) \; d^{n}x \,\text{,}
 \end{equation}
@@
 \begin{equation} \label{eq:normalization}
-N = \int f (\vec{r}, \vec{p}; t) \; d^{3}x\,d^{3}p
+  N = \int f (\vec{r}, \vec{p}; t) \; d^{3}x\,d^{3}p \,\text{.}
 \end{equation}
@@
   f_{\smallD}(\vec{r}, \vec{p}; t) = 
-  \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p})
+  \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p}) \,\text{,}
 \end{equation}
@@
   \frac{d\vec{r}_{i}(t)}{dt} = \frac{\vec{p}_{i}(t)}{m}\\
-  \frac{d\vec{p}_{i}(t)}{dt} = -\sum\limits_{j} \frac{d}{d\vec{r}_{i}} V(\vec{r}_{i} - \vec{r}_{j}) \label{eq:std_motion_v}
+  \frac{d\vec{p}_{i}(t)}{dt} = -\sum\limits_{j} \frac{d}{d\vec{r}_{i}} V(\vec{r}_{i} - \vec{r}_{j}) \label{eq:std_motion_v} \,\text{.}
 \end{gather}
@@
   \frac{d\vec{r}}{dt} \frac{\partial f_{\smallD}}{\partial \vec{r}} 
-  + \frac{\partial f_{\smallD}}{\partial \vec{p}} \frac{d\vec{p}}{dt}
+  + \frac{\partial f_{\smallD}}{\partial \vec{p}} \frac{d\vec{p}}{dt} \,\text{,}
 \end{equation}
@@
   + \frac{\partial}{\partial \vec{p}} f_{\smallD} (\vec{r}, \vec{p}; t)
-  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f_{\smallD}(\vecprime{r},\vecprime{p};t)
+  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f_{\smallD}(\vecprime{r},\vecprime{p};t) \,\text{,}
 \end{equation}
-where the extra minus sign comes from the change of perspective of \ref{eq:std_motion_v} from the frame of the individual particle $i$ to the observer frame.
+where the extra negation comes from the change of perspective of \ref{eq:std_motion_v} from the frame of the individual particle $i$ to the observer frame.
 
@@
 \begin{equation}
-  \left < f_{\smallD} (\vec{r}, \vec{p}; t) f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right > = \left < f_{\smallD} (\vec{r}, \vec{p}; t) \right > \left < f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right >
+  \left < f_{\smallD} (\vec{r}, \vec{p}; t) f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right > = \left < f_{\smallD} (\vec{r}, \vec{p}; t) \right > \left < f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right > \,\text{,}
 \end{equation}
@@
   + \frac{\partial}{\partial \vec{p}} f (\vec{r}, \vec{p}; t)
-  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f(\vecprime{r},\vecprime{p};t)
+  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f(\vecprime{r},\vecprime{p};t) \,\text{.}
 \end{equation}
@@
 \begin{equation}
-  f(\vec{r}, \vec{p}; t) = C(t)\exp \left[ - G(\vec{r}, \vec{p}; t) \right]
+  f(\vec{r}, \vec{p}; t) = C(t)\exp \left[ - G(\vec{r}, \vec{p}; t) \right] \,\text{.}
 \end{equation}
@@
 \begin{equation} \label{eq:gaussian_exponent}
-  G(\vec{r}, \vec{p}; t) =
-  \frac{x^2 + y^2}{2 \sigma_{\smallT}} + \frac{z^2}{2 \sigma_{z}}
-  + \frac{
+\begin{split}
+  G(\vec{r}, \vec{p}; t) &= \frac{x^2 + y^2}{2 \sigma_{\smallT}} + \frac{z^2}{2 \sigma_{z}} \\
+  &+\frac{
         [p_x - (\gamma_{\smallT}/\sigma_{\smallT}) x ]^2 
@@
     }{2 \eta_{\smallT}}
-  + \frac{ [p_z - (\gamma_{z}/\sigma_{z}) z ]^2 }{2 \eta_{z}}
+  + \frac{ [p_z - (\gamma_{z}/\sigma_{z}) z ]^2 }{2 \eta_{z}} \,\text{.}
+\end{split}
 \end{equation}
@@
     \frac{1}{\sigma_{\smallT}^2\eta_{\smallT}^2\sigma_{z}\eta_{z}}
-  \right)^{\frac{1}{2}}
+  \right)^{\frac{1}{2}} \text{.}
 \end{equation}
@@
 \begin{equation} \label{eq:coordinate_vector}
-u_i = \{x, p_x, y, p_y, z, p_z\}\text{ .}
+u_i = \{x, p_x, y, p_y, z, p_z\} \,\text{.}
 \end{equation}
@@
 \begin{equation}
-  G(\vec{r}, \vec{p}; t) = \frac{1}{2}\sum\limits_{ij} A_{ij} u_i u_j
+  G(\vec{r}, \vec{p}; t) = \frac{1}{2}\sum\limits_{ij} A_{ij} u_i u_j \,\text{,}
 \end{equation}
@@
     \gamma_{\alpha}/(\eta_{\alpha} \sigma_{\alpha}) & 1/\eta_{\alpha}
-  \end{pmatrix}
+  \end{pmatrix} \,\text{,}
 \end{equation}
-and again ${\scriptstyle \alpha} = \smallT \text{ or } z$. 
+again with $\alpha = {\scriptstyle T} \text{ or } z$. 
 A nice property of Gaussians of this form is the convenient ``Multidimensional Gaussian Integral'' given by
 \begin{equation} \label{eq:gaussian_integral}
-  (A^{-1})_{ij} = \frac{1}{N} \int u_i u_j f(\vec{r}, \vec{p}; t) d\vec{r} d\vec{p}
+  (A^{-1})_{ij} = \frac{1}{N} \int u_i u_j f(\vec{r}, \vec{p}; t) d\vec{r} d\vec{p} \,\text{,}
 \end{equation}
@@
     \gamma_{\alpha} & \eta_{\alpha} + \gamma_{\alpha}^2/\sigma_{\alpha}
-  \end{pmatrix}
+  \end{pmatrix} \,\text{.}
 \end{equation}
@@
 \begin{equation} \label{eq:gaussian_integral_dt}
-  \frac{d}{dt}(A^{-1})_{ij} = \frac{1}{N} \int u_i u_j \frac{df(\vec{r}, \vec{p}; t)}{dt} d\vec{r} d\vec{p}
+  \frac{d}{dt}(A^{-1})_{ij} = \frac{1}{N} \int u_i u_j \frac{df(\vec{r}, \vec{p}; t)}{dt} d\vec{r} d\vec{p} \,\text{,}
 \end{equation}
@@
     \dfrac{d\gamma_{\alpha}}{dt} & \dfrac{d\eta_{\alpha}}{dt} + 2\dfrac{\gamma_{\alpha}}{\sigma_{\alpha}}\dfrac{d\gamma_{\alpha}}{dt}- \dfrac{\gamma^{2}_{\alpha}}{\sigma^{2}_{\alpha}}\dfrac{d\sigma_{\alpha}}{dt}
-  \end{pmatrix}
+  \end{pmatrix} \,\text{,}
 \end{equation}
@@
 Inserting \ref{eq:formal_evolution} into \ref{eq:gaussian_integral_dt} yields two integral terms, one related to the propagation of a non-charged Gaussian beam, the other adds the influence of the internal force due to Coulomb repulsion for an electron pulse.
-These terms will be denoted $K^{flow}$ and $K^{force}$ respectively.
+These terms will be denoted $K^{flow}$ and $K^{force}$ respectively, 
 \begin{equation} \label{eq:gaussian_integral_dt_k}
-  \frac{d}{dt}(A^{-1})_{ij} = K^{flow}_{ij} + K^{int}_{ij}
+  \frac{d}{dt}(A^{-1})_{ij} = K^{flow}_{ij} + K^{int}_{ij} \,\text{.}
 \end{equation}
@@
     \label{eq:potential_integral}
-    \Phi(\vec{r};t) \equiv \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t)
+    \Phi(\vec{r};t) \equiv \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) \,\text{.}
 \end{gather}
@@
     \dfrac{1}{m} \left(\eta_{\alpha} + \dfrac{\gamma_{\alpha}^2}{\sigma_{\alpha}} \right) & 0
-  \end{pmatrix}
+  \end{pmatrix} \,\text{.}
 \end{equation}
@@
     0 & 0 & \hat{k}^{force}_{z}
-  \end{pmatrix}
+  \end{pmatrix} \,\text{,}
 \end{equation}
@@
     1 & \dfrac{2 \gamma_{\alpha}}{\sigma_{\alpha}}
-  \end{pmatrix}
+  \end{pmatrix} \,\text{.}
 \end{equation}
@@
 \begin{equation} \label{eq:define_xi}
-  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}}
+  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}} \,\text{.}
 \end{equation}
@@
     L_{z}(\xi) = \frac{ 3 \xi^2 }{ \xi^2 - 1} \left( \xi L(\xi) - 1 \right) \\
-    L(\xi) = \frac{1}{2} \int_0^{\pi} \frac{1}{1+\xi\sin(\theta)} d\theta \label{eq:ag_l}
+    L(\xi) = \frac{1}{2} \int_0^{\pi} \frac{1}{1+\xi\sin(\theta)} d\theta \,\text{.} \label{eq:ag_l}
   \end{gather}
@@
     \frac{ \arcsin \sqrt{1-\xi^2} }{ \sqrt{1-\xi^2} } & 0 \le \xi \le 1 \\
-    \frac{ \ln\left( \xi + \sqrt{\xi^2 - 1} \right) }{ \sqrt{\xi^2 - 1} } & \xi \ge 1
+    \frac{ \ln\left( \xi + \sqrt{\xi^2 - 1} \right) }{ \sqrt{\xi^2 - 1} } & \xi \ge 1 \,\text{,}
   \end{cases}

### 2013-04-10 inc/extension/external_forces/formal.tex
more punctuation near equations and some equation source reformmatting

@@
 \begin{equation}
-  V(\vec{r} - \vecprime{r}) \rightarrow V^{int}(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r})
+  V(\vec{r} - \vecprime{r}) \rightarrow V^{int}(\vec{r} - \vecprime{r}) + \sum_{k} V_{k}^{ext}(\vec{r}) \,\text{.}
 \end{equation}
@@
     = & \int d\vecprime{r} V^{int}(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) + \sum_{k} \int d\vecprime{r} V_{k}^{ext}(\vec{r}) n(\vecprime{r}; t) \\
-    \equiv & \Phi^{int}(\vec{r};t) + \sum_{k} \Phi_{k}^{ext}(\vec{r};t)
+    \equiv & \Phi^{int}(\vec{r};t) + \sum_{k} \Phi_{k}^{ext}(\vec{r};t) \,\text{.}
   \end{split}
@@
 \begin{equation}
-  \Phi^{ext}(\vec{r};t) \equiv \int d\vecprime{r} V^{ext}(\vec{r}) n(\vecprime{r}; t)
+  \Phi^{ext}(\vec{r};t) \equiv \int d\vecprime{r} V^{ext}(\vec{r}) n(\vecprime{r}; t) \,\text{.}
 \end{equation}
@@
     K^{force}_{ij} = & \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial}{\partial \vec{r}} \left( \Phi(\vec{r};t) + \sum_{k} \Phi_{k}^{ext}(\vec{r};t) \right) d\vec{r} d\vec{p} \\
-    \equiv & K^{int}_{ij} + \sum_{k} (K_{k}^{ext})_{ij}
+    \equiv & K^{int}_{ij} + \sum_{k} (K_{k}^{ext})_{ij} \,\text{.}
   \end{split}
@@
 For example, consider an acceleration electric field $ \vec{E} = -E \hat{z} $ such that electrons are accelerated in the positive $ z $ direction.
-%TODO e = |q| ?
-The force is then $ \vec{F}_{ext} = qE \hat{z} $, so that from \ref{eq:Kint} we have
+The force is then $ \vec{F}_{ext} = e E \hat{z} $, so that from \ref{eq:Kint} we have
 \begin{equation}
-  K^{ext}_{ij} = \frac{qE} {N} \int{ u_{i} u_{j} \left( \frac{p_{z}} {\eta_{z}} - \frac{\gamma_{z} z} {\sigma_{z} \eta_{z}} \right) f d^{3}x d^{3}p } \,\text{.}
+  K^{ext}_{ij} = \frac{e E} {N} \int{ u_{i} u_{j} \left( \frac{p_{z}} {\eta_{z}} - \frac{\gamma_{z} z} {\sigma_{z} \eta_{z}} \right) f d^{3}x d^{3}p } \,\text{.}
 \end{equation}
 It is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ (or equivalently $ K^{ext} = \hat{0} $) meaning that this static field has no effect on the parameters of the pulse.
-Of course, this is only true in the pulse frame, the frame of the simulation, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m_{e} $ is unchanged.
-In the lab frame however, where  $ \Delta E = p \Delta p / m_{e} $, there is an increase in the momentum spread, as the pulse now has the large momentum $p$ imparted by the accelerating field.
+Of course, this is only true in the pulse frame, the frame of the simulation, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m $ is unchanged.
+In the lab frame however, where  $ \Delta E = p \Delta p / m $, there is an increase in the momentum spread, as the pulse now has the large momentum $p$ imparted by the accelerating field.
 
@@
 
-%External forces that vary with position in the pulse reference frame (i.e., are not constant) act to change the electron pulse propagation dynamics, most often by affecting $\gamma_{\alpha}$ --- the spatial dependence of the local relative mean electron momenta.
-%Below, we consider two examples of electron optical elements that are being used, or are under consideration, for controlling and manipulating the space-time dynamics of electron pulses in DTEM and UED instrumentation; namely, magnetic lenses for spatial focusing and RF pulse compression cavities.
-
 For the case of a cylindrically symmetric perfect lens, or in the longitudinal direction a perfect RF compression element, with a parabolic potential, the external force is linear with distance from the center of the electron pulse:
 \begin{equation} \label{eq:linear_force}
-\vec{F}_{ext} = - M_{ \smallT } r \hat{\operatorname{r}} - M_{z} z \hat{\operatorname{z}} \text{ .}
+\vec{F}_{ext} = - M_{ \smallT } r \hat{\operatorname{r}} - M_{z} z \hat{\operatorname{z}} \,\text{.}
 \end{equation}
@@
   \begin{split} 
-  K^{ext}_{ij} & \equiv \dfrac{1} {N} \int u_{i} u_{j} ( \vec{\nabla}_{p} f ) \cdot \vec{F}_{ext} d^{3} x d^{3} p \\ & = \dfrac{1} {N} \int u_{i} u_{j} \biggl[ x \left( \dfrac{ p_{x} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } x } { \sigma_{ \smallT } \eta_{ \smallT } } \right) + y \left( \dfrac{ p_{y} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } y } { \sigma_{ \smallT } \eta_{ \smallT } } \right) \\ & \quad + z \left( \dfrac{ p_{z} } { \eta_{z} } - \dfrac{ \gamma_{z} z } { \sigma_{z} \eta_{z} } \right) \biggr] f d^{3} x d^{3} p ,
+    K^{ext}_{ij} & \equiv \dfrac{1} {N} \int u_{i} u_{j} ( \vec{\nabla}_{p} f ) \cdot \vec{F}_{ext} d^{3} x d^{3} p \\ 
+    & = \dfrac{1} {N} \int u_{i} u_{j} \biggl[ x \left( \dfrac{ p_{x} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } x } { \sigma_{ \smallT } \eta_{ \smallT } } \right) + y \left( \dfrac{ p_{y} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } y } { \sigma_{ \smallT } \eta_{ \smallT } } \right) \\
+    & \quad + z \left( \dfrac{ p_{z} } { \eta_{z} } - \dfrac{ \gamma_{z} z } { \sigma_{z} \eta_{z} } \right) \biggr] f d^{3} x d^{3} p \,\text{,}
   \end{split}
@@
     \hat{0} & \hat{0} & \hat{k}^{lin}_{z}
-\end{pmatrix} \text{ ,}
+  \end{pmatrix} \text{,}
 \end{equation}
@@
     0 & M_{\alpha} \sigma_{\alpha} \\
-M_{\alpha} \sigma_{\alpha} & 2 M_{\alpha} \gamma_{\alpha} \text{.}
-\end{pmatrix} \text{ .}
+    M_{\alpha} \sigma_{\alpha} & 2 M_{\alpha} \gamma_{\alpha}
+  \end{pmatrix} \text{.}
 \end{equation}
@@
   \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
-    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} 
+    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} \,\text{.}
 \end{equation}

### 2013-04-10 inc/hardware/deflector/deflector.tex
merged remaining corrections from Carolyn

@@
 A schematic description of such a system is shown in \ref{fig:deflector_schematic}.
-To estimate the required voltage on the plates $V$, consider the applied impulse provided to the beam by the plate's electric field
+To estimate the required voltage on the plates $V$ needed to correct for a lateral velocity $v_x$, consider the applied impulse provided to the beam by the plate's force $F$ of the electric field,
 \begin{gather}
   F \Delta t = m \Delta v \\
-  \frac{ e V }{ d } \frac{ L }{ v_z } = m_e \Delta v_x
+  \frac{ e V }{ d } \frac{ L }{ v_z } = m \Delta v_x \,\text{,}
 \end{gather}
 where (as seen in the figure) $L$ is the length of the plate (in the direction of travel $\hat{z}$) and $d$ is the separation of the plates.
-Then in the small angle approximation, the angle that the plate may correct for is approximately
+If the beam has a velocity $v_z$ down the axis of the column, then in the small angle approximation, the angle that the plate may correct for is approximately
 \begin{equation}
-  \theta = \frac{ \Delta v_x }{ v_z } = \frac{ e }{ m_e } \frac{ L }{ d } \frac{ V }{ v_z^2 }
+  \theta = \frac{ \Delta v_x }{ v_z } = \frac{ e }{ m } \frac{ L }{ d } \frac{ V }{ v_z^2 } \,\text{.}
 \end{equation}

### 2013-04-10 inc/conclusions.tex
began working on a conclusion section

%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.

Ultrafast Electron Microscopy is an exciting new field of intrumentation research which promises to open our eyes to the dynamics occurring in nanoscale processes.

During my time at UIC I have worked to advance this technology, on both the theoretical (simulation/modeling) and experimental (prototyping and photocathode engineering) fronts.

\section{Modeling}

In this thesis I have presented the extended Analytic Gaussian (AG) model, an effiecient numerical model based on the work of Michalik and Sipe \cite{michalik_analytic_2006}, but extended so as to be useful in modeling the dynamics of ultrafast electron pulses in microscope columns.
For pulses containing $N$ carriers of charge $e$ and mass $m$, the model is simply these four equations,
\begin{subequations}
\begin{gather}
  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} \,\text{.}
\end{gather}
\end{subequations}
where the subscript $\alpha$ denotes the directionality of the dynamics, either $\smallT$ for the transverse or $z$ for the longitudinal.
At any time $t$, the $\sigma_{\alpha}$ quantities characterize the spatial variance of the pulse and $\gamma_{\alpha}$ represent the momentum chirp across the pulse.
A third parameter, the local momentum variance $\eta_{\alpha}$, is also encoded in these equations.
By Liouville's theorem, which the AG model explicitly satisfies, we know that the product of the spatial variance $\sigma_{\alpha}$ and the local momentum variance $\eta_{\alpha}$ is constant throughout propagation (for perfect lensing and without clipping the puse), so we define the constant,
\begin{equation} \label{eq:define_big_gamma_conc}
  \Gamma_{\alpha}^2 \equiv \sigma_{\alpha} \eta_{\alpha} \,\text{.}
\end{equation}
By evaluating this constant once, presumably from initial conditions, and later applying \ref{eq:define_big_gamma_conc} we can get the extract these $\eta_{\alpha}$ parameters while only needing to compute the other four.
The parameter $\xi$ characterizes the ellipticity of the pulse,
\begin{equation}
  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}} \,\text{,}
\end{equation}
and the family of $L$ functions (defined in \ref{eq:ag_l_family}) are smooth and well behaved functions for all physically meaningful $\xi$.
Finally the $M_{\alpha}^{\lbrace f \rbrace}$ quantities define the $\alpha$ diretion influence of the $f^{\text{th}}$ external force.

In UEM, the initial pulse conditions are of notable importance for the performance of the instrument.
Therefore I have presented a simplistic but useful set of initial conditions for the AG model.
First lets assume we photocathode which has a work function $\Phi$, the energy an electron must absorb before it is freed to leave the material.
Lets assume that the photocathode is in an acceleration region, having an applied electric field $E_{DC}$.
Finally, lets assume we have a laser pulse whose intensity $I$ is of the form,
\begin{equation}\label{eq:laser_form}
  I ( x , y , t ) = I_{0} \exp \left ( - \frac{ ( x^{2} +y^{2} ) }{ w^{2} } - \frac{ t^{ 2 } }{ \tau^{ 2 } } \right ) \punct{,}
\end{equation}
where $w$ and $\tau$ are the HW1/eM laser pulse width and duration respectively, and the photon energy $\hbar \omega$ is greater than $\Phi$, thus permitting single-photon photoemission.
Ejected electrons will then carry an excess photoemission energy $\Delta E = \hbar \omega - \Phi$, so that the maximum velocity that an ejected electron may have is $ v_{max} = \sqrt{ 2 \Delta E / m } $.
Given these generic assumptions, the resulting initial conditions are simply,
\begin{subequations}
  \begin{gather}
    \sigma_{ \smallT } ( 0 ) \approx \frac{ w^{ 2 } }{ 2 } \\
    \sigma_{ z } ( 0 ) \approx \frac{ ( v_{max} \tau )^{ 2 } }{ 2 } + \left ( \frac{ e E_{ DC } \tau^{ 2 } }{ 4 m } \right )^{ 2 } \\
    \gamma_{ \smallT } ( 0 ) \approx 0 \\
    \gamma_{ z } ( 0 ) \approx \sqrt{ \sigma_{ z } ( 0 ) } \left ( m v_{max} + \frac{ e E_{DC} \tau }{ \sqrt{ 2 } } \right ) \\
    \eta_{ \smallT } ( 0 ) \approx \frac{ m }{ 3 } ( \hbar \omega - \Phi ) \\
    \eta_{ z } ( 0 ) \approx \frac{ \eta_{ \smallT } ( 0 ) }{ 4 } \punct{.}
  \end{gather}
\end{subequations}
In Section \ref{sec:bin_model} I have also presented a binning model which, when fully investigated, may eventually be employed rather than this simplisitc set of initial conditions.

### 2013-04-10 inc/conclusions.tex
more work on conclusions and a bugfix on RF cav M terms

@@
   \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
-    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} \,\text{.}
+    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} \label{eq:ag_model_gamma_conc} \,\text{.}
 \end{gather}
@@
 In Section \ref{sec:bin_model} I have also presented a binning model which, when fully investigated, may eventually be employed rather than this simplisitc set of initial conditions.
+
+Most UEM systems will employ magnetic lenses and DC accelerators; there is also increasing interest in using RF cavities for pulse compression.
+The AG model can model the influence of these external forces via the $M$ parameters in \ref{eq:ag_model_gamma_conc}, which are in terms of the position of the center of the pulse $z^{\prime}$ and its velocity $v_{\smallzero}$.
+The each magnetic lens, positioned at $z_{mag}^{\prime}$ and of length $L_{mag}$, can be represented simply as 
+\begin{equation}
+  M_{\smallT}^{mag} = M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \,\text{,}
+\end{equation}
+where $M_{\smallT}^{mag}(I)$ is a lens-specific characterization of the magnetic field strength (versus current $I$).
+A single TM$_{010}$-mode cylindrical RF cavity, positioned at $z_{RF}^{\prime}$, having length $L_{RF}$, and oscillating at frequency $\Omega$ with field strength $E_{0}$ and phase shift $\phi$, may be expressed as
+\begin{gather}
+  M_{z}^{RF} = \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \\
+  \begin{aligned}
+  M_{\smallT}^{RF} &= e E_{0} \left [ \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{L_{RF}} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{L_{RF} / 2} \right )^{2 n - 1} \right.
+\\ & \times \left.
+\sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
+   \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \,\text{.}
+  \end{aligned}
+\end{gather}
+Additionally, since an RF cavity can act as a pulse accelerator an additional force term
+\begin{equation}
+  F_{z}^{RF} = e E_{0} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \, \text{,}
+\end{equation}
+should be added to the equation of motion governing the propagation of the pulse down the column.
+Finally, if we assume that an acceleration region, with cathode at $z^{\prime} = 0$ and anode at $z^{\prime}_A$ and acceleration potential $V$, has an on-axis force of the form
+\begin{equation} \label{eq:anode_tanh}
+  F_z(0,z^{\prime}) = \frac{eV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right) \,\text{,}
+\end{equation}
+where $s$ is a parameter that quantifies the sharpness of the field drop at the anode, then accelerator's influence on the pulse dynamics are given by
+\begin{gather}
+  M_{\smallT} = -\frac{eVs}{4z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \\
+  M_z = \frac{eVs}{2z^{\prime 2}_A} \operatorname{sech}^2 \left( \frac{z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \, \text{.}
+\end{gather}

### 2013-04-10 inc/extension/external_forces/practical.tex
more corrections and punctuation
more work on conclusions and a bugfix on RF cav M terms

@@
 \begin{equation} \label{eq:reg_of_influence}
-  M_{\alpha} \to M_{\alpha} \exp \left [ - \left (  \frac{ z^{\prime} - z_{lens}^{\prime} }{ L_{lens} / 2 } \right )^{ 2 n } \right ] \text{ ,}
+  M_{\alpha} \to M_{\alpha} \exp \left [ - \left (  \frac{ z^{\prime} - z_{lens}^{\prime} }{ L_{lens} / 2 } \right )^{ 2 n } \right ] \,\text{,}
 \end{equation}
@@
 \begin{gather}
-  m \ddot{r} = -q B_z r \dot{\theta} + m r \dot{\theta}^2                 \\
-  \frac{d}{dt} ( m r^2 \dot{\theta} ) = q r \dot{r} B_z - q r B_r \dot{z} \\
-  m \ddot{z} = q B_r r \dot{\theta}
+  m \ddot{r} = -e B_z r \dot{\theta} + m r \dot{\theta}^2                 \\
+  \frac{d}{dt} ( m r^2 \dot{\theta} ) = e r \dot{r} B_z - e r B_r \dot{z} \\
+  m \ddot{z} = e B_r r \dot{\theta} \,\text{,}
 \end{gather}
@@
   Q \equiv (a+r)^2 + z^2 \\
-  k \equiv \sqrt{ 4 a r / Q }
+  k \equiv \sqrt{ 4 a r / Q } \,\text{,}
 \end{gather}
@@
 \begin{equation}
-  M_{\smallT}^{mag} = M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \text{ ,}
+  M_{\smallT}^{mag} = M_{\smallT}^{mag}(I) \exp \left [ - \left (  \frac{ z^{\prime} - z_{mag}^{\prime} }{ L_{mag} / 2 } \right )^{ 2 } \right ] \text{,}
 \end{equation}
@@
   \begin{aligned}
-  M_{\smallT}^{RF} = & e E_{0} \left [ \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \right.
-\\ \times & \left.
+  M_{\smallT}^{RF} &= e E_{0} \left [ \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{L_{RF}} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{L_{RF} / 2} \right )^{2 n - 1} \right.
+\\ & \times \left.
 \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
-   \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ]
+   \right ] \exp \left [ - \left (  \frac{ z^{\prime} - z_{RF}^{\prime} }{ L_{RF} / 2 } \right )^{ 2 } \right ] \,\text{.}
   \end{aligned}

### 2013-04-10 inc/conclusions.tex
more work on conclusion, possibly complete through the model section (not model results)

@@
 \end{gather}
+
+I have written an implementation of this model for the Perl programming language and released to the Comprehensive Perl Archive Network (CPAN) under the name \verb!Physics::UEMColumn!
+for use by the UEM community (\url{https://metacpan.org/module/Physics::UEMColumn}).
+This model is fully object-oriented, which makes writing simulations remarkably easy for a user to read, write and change scripts.
+Under the hood, though, it uses a custom set of C level bindings to the GNU Scientific Library (GSL), so that the numerical equation solving is fast and accurate.
+I have found that this speed, combined with the dynamic flexibility offered by object-oriented Perl, has allowed me to easily write optimization scripts and generate figures which might have been nearly impossible previously; several of these have been shown in this thesis.

### 2013-04-10 inc/photocathode/plasmon/plasmon.tex
merging comments from andreas on photocathode chapter

@@
 In PAPE, the goal is use this field to both enhance the emission and `control' the divergence of photoemitted electrons through the combined action of photoemission barrier suppression and ponderomotive acceleration.
-On the other hand, these oscillating electrons are nothing but an oscillating current, and thus if a large plasmon response is established, resistive heating is a major concern even for small the small resistivities typical of conductors.
+On the other hand, these oscillating electrons are nothing but an oscillating current, and thus if a large plasmon response is established, resistive heating is a major concern even for the small resistivities typical of conductors.
 
 This avenue of investigation was motivated by Zawadzka et al \cite{zawadzka_evanescent_2001} who indicate that when driving a surface plasmon on gold they witnessed a reduction in the emission angle of photoemitted electrons.
-In their paper, the authors mount a thin gold foil on a prism; the laser then ``back illuminates'' the gold through the prism, ejecting electrons on vacuum side of the foil.
+In their paper, the authors coat a thin gold film on a prism; the laser then ``back illuminates'' the gold through the prism, ejecting electrons on vacuum side of the foil.
 This arrangement is called the Kretschmann geometry.
-While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, we felt that for UEM applications we should explore the ``grating coupling'' geometry \cite{kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
+While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, we felt that for UEM applications we should explore the alternative ``grating coupling'' geometry \cite{kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
 In this geometry, the plasmon is driven on a periodic surface rather than a flat surface attached to glass.
@@
 where $\rho_{e}$ is the free-electron density and $m^*$ is the electron effective mass in the material.
-And in the most simple (undamped) approximations, the dielectric function of the material may be written in terms of this constant
+And in the most simple (undamped) approximations, $\varepsilon$, the real part of the dielectric function of the material, may be written in the following form:
 \begin{equation} \label{eq:dielectric_as_omega}
@@
 In the grating coupling scheme (shown in \ref{fig:plasmon_schematic}) a laser of wavelength $\lambda$ is incident on a material with a grating of period $d$, making an angle $\theta$ from the surface normal.
-A surface plasmon is driven when the sum of the component of the laser wave-vector parallel to the material surface ($k_{Lx}$) and a fourier component ($k_{G} = 2 \pi / d$) of the grating matches the wave-vector of the surface plasmon
+A surface plasmon is driven when the sum of the component of the laser wave-vector parallel to the material surface ($k_{Lx}$) and a Fourier component ($k_{G} = 2 \pi / d$) of the grating matches the wave-vector of the surface plasmon
 \begin{equation}
@@
 For a given grating period, \ref{eq:plasmon_condition} determines the angle of incidence that the laser must make with the surface in order to drive the plasmon.
-Alternatively, it may be used to design the periodic surface structure, for example allowing for angles that the microscope column will accept, different driving orders ($n$) and even other Fourier components of the periodic surface structure.
+Alternatively, it may be used to design the periodic surface structure, for example allowing for angles that the microscope column will accept, different driving orders ($n$), and even other Fourier components of the periodic surface structure.
 
@@
 
-For proof-of-concept purposes the first experiment was run on a commercially available gold foil on glass holographic grating (750 lines/mm).
+For proof-of-concept purposes, the first experiment was performed on a commercially available gold foil on glass holographic grating (750 lines/mm).
 Like the Kretschmann geometry, a foil on glass grating is vulnerable to heating problems; a plasmon creates a large amount of heat but the glass is a poor heat conductor.
@@
 There are several conclusions that can be drawn from this experiment.
-Most importantly, it shows that a plasmon was driven successfully given the angular dependence and large response, few other processes could have occurred.
+Most importantly, it shows that a plasmon was driven successfully; given the angular dependence and large response, few other processes could have occurred.
 The heat generated by the plasmon oscillation destroyed the grating, therefore it validated the decision not to pursue a Kretschmann geometry over fears of heat damage at the high heating rates of the laser-plasmon interaction.
-Finally, it suggests that if using a thermally conductive and structurally sound photoemission material and mount for the plasmonic photocathode, PAPE might generate a large number of electrons (as evident by the bright flash).
+Finally, it suggests that through the use of a thermally conductive and structurally sound photoemission material and mount for the plasmonic photocathode, PAPE might generate a large number of electrons (as evident by the bright flash).
 
@@
 
-Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division employed a defocused Gallium focused ion beam (FIB) (Zeiss 1540XB) to iteratively mill an approximately sinusoidal shape into the surface of a silicon wafer.
+Collaborators at Argonne National Laboratory's Electron Microscopy Center, Material Science Division employed a defocused Gallium focused ion beam (FIB) (Zeiss 1540XB) to mill an approximately sinusoidal shape into the surface of a silicon wafer.
 Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
@@
 A 300nm film of gold was deposited onto the silicon using the Varian e-beam deposition chamber at the UIC Nano Core Facility (NCF); a thin layer of chromium was deposited between the silicon and gold as a binding agent.
-As the gold film is several times thicker than the penetration depth of green light on gold, the laser radiation cannot interact with the silicon.
+As the gold film is many times thicker than the penetration depth of green light on gold and several times thicker than the penetration depth of the plasmon field, neither the laser radiation nor the plasmon field will couple with the silicon.
 %TODO should I comment on the coating acting to smooth the surface to a proper sinusoid?
@@
 Specifically, the strength of both emissions have a quadratic dependence on the incident laser intensity.
-This is consistent with a two-photon field emission process --- it requires the absorption of two 2.37eV (green) photons for an electron to acquire sufficient energy to allow any significant field emission as gold has a work function of 5.1eV.
+This is consistent with a two-photon field emission process --- it requires the absorption of two 2.37eV (green) photons for an electron to acquire sufficient energy to allow any significant field emission as gold has a work function of $\sim$5eV.
 Moreover, both emissions have similar momentum distributions (\ref{fig:pape} A and E), again indicating a similar emission process.

### 2013-04-10 inc/photocathode/setup/setup.tex
merging comments from andreas on photocathode chapter

@@
 
-To ensure that the transverse momentum distribution is observed, the YAG scintillator must be positioned at the back focal plane (the Fourier plane) of the lens system.
+To ensure that the transverse momentum distribution is monitored, the YAG scintillator must be positioned at the back focal plane (the Fourier plane) of the lens system.
 This is readily achieved by appropriate adjustment of the current $I$ in the magnetic lens coils; specifically, after determining the current required to focus the electron beam onto the scintillator, optical imaging relations together with the $I^2$ dependence of the magnetic lens strength are used to find the current that ensures that the average lens to scintillator distance equals the focal length of the lens system.
@@
 Additionally, and as expected from optical considerations, the modeling shows that the spatial beam size in the Fourier plane is independent of $\Delta x_0$, the rms laser spot size incident on the photocathode (i.e., the electron source size).
+All of the results in this chapter use the Fourier plane beam size as a measure of the initial rms transverse momentum $\Delta p_{\smallT}$.

### 2013-04-10 inc/photocathode/single_photon/single_photon.tex
merging comments from andreas on photocathode chapter

@@
 Single-photon photoemission is the most common photoemission process and the easiest to explain.
-As described by Einstein in his famous paper on ``the photoelectric effect'' \cite{einstein_uber_1905}, as it is colloquially known, if an electron absorbs a photon whose energy is sufficient to promote the electron above the local vacuum level, it become free to leave the host material.
+As described by Einstein in his famous paper on ``the photoelectric effect'' \cite{einstein_uber_1905}, as it is colloquially known, if an electron absorbs a photon whose energy is sufficient to promote the electron above the local vacuum level, it becomes free to leave the host material.
 The material dependent minimum energy needed to eject an electron is termed its ``work function'' $\Phi$.
-Any additional energy above the work function remains with the ejected electron, usually in the form of kinetic energy, and is called the ``excess photoemission energy'' $\Delta E$, for single photon photoemission, using a laser of frequency $\omega$ (photon energy $\hbar \omega$), this is simply,
+Any additional energy above the work function remains with the ejected electron, usually in the form of kinetic energy, and is called the ``excess photoemission energy'' $\Delta E$; for single photon photoemission, using a laser of frequency $\omega$ (photon energy $\hbar \omega$), this is simply,
 \begin{equation}
@@
 \end{equation}
-The efficiency of photoemission $\eta_{{\scriptscriptstyle PE}}$, when $\Delta E$ is small compared to $\Phi$ may be written as,
+The efficiency of photoemission $\eta_{{\scriptscriptstyle PE}}$, when $\Delta E$ is small compared to $\Phi$ may be written as, %TODO ref
 \begin{equation}
@@
 \end{equation}
-where $A$ is a material paramenter, and R is the photocathode surface reflectivity.
+where $A$ is a material paramenter (typically of the order of $\sim 10^{-4} \text{eV}^{-2}$ %TODO ref
+), and $R$ is the photocathode surface reflectivity, which is itself a function of the polarization, frequency and angle of the incident light.
+%TODO calculation of $\eta_{PE}$?
 

### 2013-04-10 inc/photocathode/two_photon_thermionic/two_photon_thermionic.tex
merging comments from andreas on photocathode chapter

@@
 Two-photon assisted thermionic emission (2PTE) occurs when two photons provide nearly enough energy to emit a cold electron, and therefore a hotter electron, in the tail of the Fermi distribution, may have enough energy to be emitted \cite{yen_thermally_1980}.
-Of course this is even less efficient than two-photon photoemission, but as three-photon photoemission goes as $I^3$, 2PTE will still be the dominant photoemission mechanism.
+Of course this is even less efficient than direct two-photon photoemission, but as three-photon photoemission goes as $I^3$, 2PTE will still be the dominant photoemission mechanism.
 For thermionic emission from electrons at a temperature $T_e$, the rms transverse momentum $\Delta p_{\smallT}$ is given by 
@@
 where $k_B$ is the Boltzmann constant \cite{dowell_quantum_2009,jensen_emittance_2010}.
-Since a high laser intensity is required to drive 2PTE, the laser itself will heat the electrons, therefore we expect the measured Fourier plane spot size to grow as the square root of the laser pulse energy.
+Since a high laser intensity is required to drive 2PTE, the laser itself will heat the electrons, therefore, if the laser pulse duration is less than the thermalization time of the electron distribution, %TODO check this!
+we expect the measured Fourier plane spot size to grow as incident laser pulse energy increases.
 

### 2013-04-11 inc/photocathode/este/este.tex
merge most of the remaining comments from andreas on photocathode

--- 2013-04-09 inc/photocathode/este/este.tex
@@
 
+Recent experiments in our group have come to change our understanding of several photoemission processes, once thought to be well understood \cite{berger_excited_2012}.
+Indeed, while this knowledge has been commonplace in other fields, such as angle-resolved photoemission spectroscopy (ARPES) \cite{himpsel_angle-resolved_1983}, to our knowledge the effect has not been considered for electron beam generation.
 In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_{\smallT}$ developed in Refs.
@@
 
-\ref{fig:este-semicond} depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique.
+\ref{fig:este-semicond} depicts the results obtained for GaSb and InSb photocathode materials under pulsed UV (261nm) laser irradiation using the same experimental technique employed for PAPE, 2PTE, and single-photon PE studies.
 Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
@@
 The known optical properties of the two semiconductors \cite{aspnes_dielectric_1983} indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb \cite{chelikowsky_nonlocal_1976} and InSb \cite{chelikowsky_erratum_1984}, is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
-Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m$ (0.5$m$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV).
+Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m$ (0.5$m$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV); that is, an initial electron temperature $T_e$ of $\sim$4,200K ($\sim$4,900K).
 For the $\sim$0.1nJ incident UV pulse energies, $\sim40\mu$m laser spot size, and $p$-polarized reflectance of $\sim$40\% for both semiconductors at our 60$^{\circ}$ incidence angle, the photo-injected electron density is $\sim10^{18}cm^{-3}$.
@@
 
-Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above $\Phi$, thereby curtailing the observed ESTE.
+Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above the effective $\sim$1eV work function, thereby curtailing the observed ESTE.
 By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band \cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
@@
 \ref{fig:este-semicond} also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
-More significantly, using $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m$ in the semiconductor), simulation of the experiment with the extended AG model \cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.45($\pm$0.03)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\Phi/(k_B T_e)] \sim 10^{-15}$).
+More significantly, using $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m$ in the semiconductor), simulation of the experiment with the extended AG model \cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.90($\pm$0.06)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\Phi/(k_B T_e)] \sim 10^{-15}$).
 On the other hand, employing a mass of 0.3$m$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
@@
 The dependence of $\Delta p_{\smallT}$ on $m^*$ is readily explained through a consideration of energy and momentum conservation in transmission across a boundary with a potential step associated with the effective photoelectric work function $\Phi$.
-Inside the material, an electron with an energy $E$ greater than $\Phi$ has a maximum momentum of $\sqrt{2 m^* (E-\Phi) }$ parallel to the boundary if it is to be emitted.
+Inside the material, an electron with an energy $E$ above the Fermi energy which is greater than $\Phi$ has a maximum momentum of $\sqrt{2 m^* (E-\Phi) }$ parallel to the boundary if it is to be emitted.
 As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_{\smallT}$ of the electron source clearly must scale with $\sqrt{m^*}$ as observed in the experiment.

### 2013-04-11 inc/photocathode/future.tex
merge most of the remaining comments from andreas on photocathode

@@
 The philosophical argument presented in Section \ref{sec:este} for using the effective mass $m^*$ in the expressions for the transverse momentum variance $\Delta p_{\smallT}$ simply follows from conservation of energy and momentum at the barrier and actually does not depend on the ``excited state'' nature nor the thermionic emission process.
-For the example two-photon assisted thermionic emission (2PTE) from gold presented in Section \ref{sec:two_photon_thermionic}, the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to the rest mass $m$ \cite{johnson_optical_1972}.
+For the example of two-photon assisted thermionic emission (2PTE) from gold presented in Section \ref{sec:two_photon_thermionic}, the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to its rest mass $m$ \cite{johnson_optical_1972} --- the value employed in that analysis.
 
@@
 For a selection of single-crystal and polycrystalline photocathodes, we will show that the simulation, incorporating $m^*$, is in good agreement with our direct measurements of $\Delta p_{\smallT}$.
-We will also discuss the effect of laser-induced heatingon $\Delta p_{\smallT}$ of the electron distribution in the photocathode during photoemission, which may limit the attainable pulsed electron source brightness at the desired high peak pulse current densities.
+We will also discuss the effect of laser-induced heating on $\Delta p_{\smallT}$ of the electron distribution in the photocathode during photoemission, which may limit the attainable pulsed electron source brightness at the desired high peak pulse current densities.

### 2013-04-11 inc/photocathode/plasmon/plasmon.tex
merge most of the remaining comments from andreas on photocathode

@@
 In an attempt to mitigate the effect of the surface peaks on the accelerating field, a new pattern was commissioned and prepared in a similar manner to the original.
-This new photocathode featured long trenches of an approximately square profile, arranged periodically across the surface.
-This periodicity should have been sufficient to facilitate the generation of the plasmon.
-Although the structure should be less disruptive to the accelerating electric field, this photocathode performed markedly worse than its sinusoidal predecessor.
+This new photocathode featured long trenches of an approximately square profile, arranged periodically across the surface; the trench spacing was five times the trench width in an attempt to present a ``flatter'' surface, which would should be less disruptive to the accelerating electric field.
+Although this periodicity should have been sufficient to facilitate the generation of the plasmon, the photocathode performed markedly worse than its sinusoidal predecessor.
 Indeed, no evidence of sucessful plasmon oscillation was demonstrated after repeated attempts.
@@
 Though the cause of the failure may never be known for certain, several causes are suspected.
-Now that the surface is not purely sinusoidal, so that the surface has multiple Fourier components.
-This may mean that the laser may not couple as efficiently to the plasmon, requiring increased laser power for oscillation. 
-Additionally, the non-smooth surface may cause increased difficulty for propagating the plasmon for large distances.
+Now that the surface is not purely sinusoidal, the surface has multiple Fourier components.
+This may mean that the laser may not couple as efficiently to the plasmon, requiring significantly increased laser power for PAPE. 
+Additionally, the non-smooth surface may cause increased difficulty for propagating the plasmon over large distances.
 Finally, upon further investigation of the surface by SEM, evidence was observed of incomplete filling of the gold coating into the trenches, which may have interrupted creation or propagation of the plasmon.

### 2013-04-11 inc/conclusions.tex
continuing to work on conclusion

@@
 
-During my time at UIC I have worked to advance this technology, on both the theoretical (simulation/modeling) and experimental (prototyping and photocathode engineering) fronts.
+During my time at UIC I have worked to advance this technology, I have developed a model and implemented it, and I have  both the theoretical (simulation/modeling) and experimental (prototyping and photocathode engineering) fronts.
 
@@
 Finally, lets assume we have a laser pulse whose intensity $I$ is of the form,
-\begin{equation}\label{eq:laser_form}
+\begin{equation}
   I ( x , y , t ) = I_{0} \exp \left ( - \frac{ ( x^{2} +y^{2} ) }{ w^{2} } - \frac{ t^{ 2 } }{ \tau^{ 2 } } \right ) \punct{,}
@@
 should be added to the equation of motion governing the propagation of the pulse down the column.
-Finally, if we assume that an acceleration region, with cathode at $z^{\prime} = 0$ and anode at $z^{\prime}_A$ and acceleration potential $V$, has an on-axis force of the form
-\begin{equation} \label{eq:anode_tanh}
+Finally, if we assume that an acceleration region, with cathode at $z^{\prime} = 0$, anode at $z^{\prime}_A$, and acceleration potential $V$, has an on-axis force of the form
+\begin{equation}
   F_z(0,z^{\prime}) = \frac{eV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right) \,\text{,}
@@
 I have found that this speed, combined with the dynamic flexibility offered by object-oriented Perl, has allowed me to easily write optimization scripts and generate figures which might have been nearly impossible previously; several of these have been shown in this thesis.
+
+\section{UEM Design}
+
+I have expounded on several factors that we feel are key concepts in designing a UEM.
+I have used the above model to show that an electron pulse, whether oblate (disk-like) or prolate (cigar-like), its initial expansion will be primarily in the the shorter dimension, i.e.
+an electron pulse ``wants'' to be spherical.
+The model predicts that the use of shorter focal-length magnetic lenses will result in higher fidelity focusing; such lenses limit the impulse delivered to the pulse by reducing the time of flight that the pulse experiences the higher charge densities inherent during focusing.
+Additionally, we have seen that oblate pulses are preferred to prolate pulses in terms of focus fidelity.
+The model predicts that when properly designed and driven, RF cavities can effectively recompress electron pulses which have expanded longitudinally.
+Interestingly, the model suggests that the additional bandwidth generated by space-charge repulsion (especially in higher charge-density pulses) can allow for additional pulse compression even at one-to-one image to object distances.
+
+The Rose criterion \cite{rose_television_1948} states that in order to have proper gray-scaling in a single-shot DTEM employing a 1k X 1k CCD camera, the pulse must contain at least $10^8$ electrons.
+We have proposed that the term ``ultrafast'' be used in the case where an electron pulse may be completely contained inside the acceleration region of the column.
+The reason for this designation is that such pulses are subject to a different form of Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, which limits its the number of electrons $N_{crit}$ that may be generated by a laser of HW1/eM width $w$ before the pulse is subject to deleterious distortions,
+\begin{equation}
+  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d} \,\text{,}
+\end{equation}
+where $V_{DC}$ and $d$ are the accelerator voltage and length respectively.
+In order to satisfy the both the Rose criterion and this ultrafast Child-Langmuir restriction, we expect that UEM instruments will need to employ a large-area emission source.
+This has necessitated building a custom acceleration region based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007} which boasts a moderately uniform acceleration field near the axis of the accelerator, as well as other large-aperture electron-optical elements for our prototype UEM column.
+We have also designed and built a custom laser source capable of generating these electron pulses.
+
+A common figure of merit for electron beams is its normalized transverse emittance \cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, which for the usual case of axially symmetric electron pulses has the form,
+\begin{equation}
+  \varepsilon_{\smallT} = \frac{1}{m c} \Delta x \Delta p_{\smallT} \,\text{,}
+\end{equation}
+where $\Delta x$ and $\Delta p_{\smallT}$ are the transverse spatial uncertainty and transverse momentum uncertainty, respectively.
+This quantity is conserved under propagation through free space and perfect electron-optical, though it may increase for imperfect ones \cite{oshea_reversible_1998}.
+The normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred \cite{berger_dc_2009}.
+Since we have already established that a large area emission source must be used, the only avenue to reduce $\varepsilon_{\smallT}$ is to reduce the transverse momentum uncertainty, which is dependent on the photocathode and photoemission process employed.
+
+\section{Photocathode Engineering}
+
+The desire to reduce the transverse momentum uncertainty $\Delta p_{\smallT}$ --- in an effort to reduce the rms transverse emittance $\varepsilon_{\smallT}$ and thereby improve the resolving power of the electron pulse~\cite{berger_dc_2009} --- has driven us to explore many photocathodes and even several different photoemission processes.

### 2013-04-11 inc/design_considerations/considerations.tex
a few more corrections
continuing to work on conclusion

@@
 \begin{equation}
-  J_{CL} = \frac{4 \varepsilon_0}{9} \sqrt{\frac{2e}{m}} \frac{V_{DC}^{3/2}}{d^2}
+  J_{CL} = \frac{4 \varepsilon_0}{9} \sqrt{\frac{2e}{m}} \frac{V_{DC}^{3/2}}{d^2} \,\text{,}
 \end{equation}
@@
 \begin{equation}
-  J_{crit} = \frac{27}{4} \left ( \frac{ 1 - \sqrt{1-\tfrac{1}{3} \chi^2} }{ \chi^3 } \right ) J_{CL}
+  J_{crit} = \frac{27}{4} \left ( \frac{ 1 - \sqrt{1-\tfrac{1}{3} \chi^2} }{ \chi^3 } \right ) J_{CL} \,\text{,}
 \end{equation}
@@
 \begin{equation}
-  J_{crit} = \frac{\varepsilon_0 E_{DC}}{\tau_e}
+  J_{crit} = \frac{\varepsilon_0 E_{DC}}{\tau_e} \,\text{,}
 \end{equation}
@@
 \begin{equation} \label{eq:n_crit}
-  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d}
+  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d} \,\text{.}
 \end{equation}
@@
 \begin{equation}
-  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > - \left < x p_x \right >^2 } \,\text{,}
+  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_{\smallT}^2 \right > - \left < x p_{\smallT} \right >^2 } \,\text{,}
 \end{equation}
-where $m$ is the electron mass, $c$ is the speed of light in vacuum and $x$ and $p_x$ are the transverse position and momentum respectively.
+where $m$ is the electron mass, $c$ is the speed of light in vacuum and $x$ and $p_{\smallT}$ are the transverse position and momentum respectively.
 For the usual case of axially symmetric electron pulses, we may use the simplified form
 \begin{equation} \label{eq:emittance_definition}
-  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_x^2 \right > } \,\text{.}
+  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_{\smallT}^2 \right > } \,\text{.}
 \end{equation}
@@
 \begin{equation} \label{eq:emittance_definition_w}
-  \varepsilon_{\smallT} = \frac{w}{m c} \sqrt{ \frac{\left < p_x^2 \right >}{2} } \,\text{.}
+  \varepsilon_{\smallT} = \frac{w}{m c} \sqrt{ \frac{\left < p_{\smallT}^2 \right >}{2} } \,\text{.}
 \end{equation}
@@
 \begin{equation}
-  \sqrt{ \left < p_x^2 \right > } = \sqrt{ \eta_{\smallT} (0) } \le \frac{5\sqrt{2} \hbar}{w} N_p \, \Delta X \frac{M}{a} \left ( \frac{\Delta a}{a} \right ) \,\text{.}
+  \sqrt{ \left < p_{\smallT}^2 \right > } = \sqrt{ \eta_{\smallT} (0) } \le \frac{5\sqrt{2} \hbar}{w} N_p \, \Delta X \frac{M}{a} \left ( \frac{\Delta a}{a} \right ) \,\text{.}
 \end{equation}

### 2013-04-11 inc/conclusions.tex
finished major work on conclusions

@@
 The desire to reduce the transverse momentum uncertainty $\Delta p_{\smallT}$ --- in an effort to reduce the rms transverse emittance $\varepsilon_{\smallT}$ and thereby improve the resolving power of the electron pulse~\cite{berger_dc_2009} --- has driven us to explore many photocathodes and even several different photoemission processes.
+I have demonstrated that single photon photoemission and two-photon thermionic photoemission were not likely candidates to reduce the $\Delta p_{\smallT}$ as those processes were tied to intrinsic properties of the material which couldn't provide high-yield and low emittance simultaneously.
+Recently, however, we have discovered that for those processes, the effective electron mass $m^*$ should have been considered, and therefore we are currently revisiting many of the materials that fall into these categories, a publication on this topic is in preparation.
+
+I have also presented our efforts to employ periodic nanostructured photocathode surfaces, designed to allow the incident laser to drive a surface plasmon, in hopes that a surface plasmon field might reduce the transverse emittance \cite{zawadzka_evanescent_2001,kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
+Though we had success in driving the plasmon, we were unable to demonstrate reduction in $\Delta p_{\smallT}$.
+We believe the problems were due to disturbances in the acceleration field near these nanostructures, though efforts to mitigate those disturbances has yet been unsuccessful.
+
+Our most promising results have come from excited-state thermionic emission (ESTE) sources, GaSb and InSb.
+In this process the laser promotes carriers from the valence band to an upper conduction band close to the vacuum level, reducing the effective work function.
+The electron temperature, elevated by the laser, causes the tail of the Fermi distribution to exceed the effective work function and thus emission occurs.
+Importantly, the results we observed were inconsistent with predictions until we considered the effective mass of these upper conduction bands.
+Not only did this allow our prediction to match the experimental results, it opened our eyes to the possibilities of low effective mass materials, even for photoemission processes other than ESTE.
+As I have mentioned earlier, this work is ongoing and is a very exciting avenue for future research.

### 2013-04-11 inc/introduction.tex
began introduction

The development of laboratory-scale instrumentation for the direct visualization of fundamental ultrafast (i.e., sub-nanosecond) events with nano-scale (even atomic) spatial resolution is a current active area of research \cite{king_ultrafast_2005}.
In many areas of physics, chemistry and materials science a space-time resolution in the sub-1nm$\cdot$ps range is ideally required to allow for the observation of the fastest structural events.
A promising approach is to employ ultrashort pulses of electrons generated by a laser-driven 10-100keV photoelectron gun to probe the dynamics of the specimen that has been perturbed by a synchronized laser pulse \cite{king_ultrafast_2005}.
Ultrafast electron diffraction (UED) or femtosecond electron diffraction (FED) \cite{srinivasan_ultrafast_2003,williamson_clocking_1997,cao_femtosecond_2003} is an example of this type of technique and has been used successfully to elucidate atomic-scale dynamics in melting,\cite{cao_femtosecond_2003,sciaini_electronic_2009,siwick_atomic-level_2003} chemical bond breaking,\cite{zewail_laser_1988} etc.
Direct time-resolved imaging of similar events can be accomplished using transmission electron microscopy (TEM).
To date, such dynamic transmission electron microscopes (DTEMs) \cite{bostanjoglo_tracing_1980,domer_high-speed_2003} have been developed for both single-shot imaging with $\sim$$10^{5}$nm$\cdot$ps space-time resolution (e.g., 30ns electron pulses and 5-10nm spatial resolution \cite{lagrange_single-shot_2006,armstrong_practical_2007}) and $\sim$1nm$\cdot$ps space-time resolution ($\sim$100fs with a few nanometer resolution\cite{park_direct_2009}) in the multi-shot data acquisition mode of operation.
Both of these instruments employ retro-fitted standard electron microscope columns and are limited in performance mainly by space-charge effects.

Improvements in the performance of DTEMs, especially ultrafast electron microscopes (UEMs), and UED systems will likely require the implementation of techniques that offset or compensate for deleterious space-charge effects \cite{armstrong_prospects_2007,lobastov_four-dimensional_2005}.
Such performance optimization efforts will require efficient models of electron pulse propagation dynamics that include the ability to simulate the action of electron optical elements (including the photoelectron gun itself) on the pulse.
Indeed $2^{\text{nd}}$ generation UED instruments are now beginning to employ both magnetic lenses and RF cavities \cite{oudheusden_electron_2007,veisz_hybrid_2007} to optimize the electron pulse delivery to the specimen; that is, compensate for global space-charge effects in both the transverse and longitudinal pulse dimensions, respectively.
Similarly, future UEM column designs are expected to use RF pulse compression cavities while circumventing high charge density beam cross-overs (i.e., beam foci).

In this thesis, I will present an extension of the work of Michalik and Sipe \cite{michalik_analytic_2006,michalik_erratum:_2008}, describing electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment, to include the influence of linear external forces due to electron optical elements.
The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
For completeness, in Section \ref{sec:ms_model} I will introduce, in MKS units, the essential features of the AG model.
In Chapter \ref{chap:extension}, I present some optimization to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
I then present my extensions to the Michalik and Sipe AG model which permit the inclusion of generic external forces.
In order to adapt this extended AG model for simulating pulses in a realistic UEM column, I then present specific contributions that arise from magnetic lenses, RF cavities and DC accelerators.
I would like to note that the analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.

Since my extension to the AG model is valid only within the limits of the analytical method itself, in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid \cite{luiten_how_2004}.
This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
Consequently, the AG approach is applicable to both the single electron per pulse limit \cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate \cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.

### 2013-04-11 inc/introduction.tex
continuing with the introduction

@@
 A promising approach is to employ ultrashort pulses of electrons generated by a laser-driven 10-100keV photoelectron gun to probe the dynamics of the specimen that has been perturbed by a synchronized laser pulse \cite{king_ultrafast_2005}.
-Ultrafast electron diffraction (UED) or femtosecond electron diffraction (FED) \cite{srinivasan_ultrafast_2003,williamson_clocking_1997,cao_femtosecond_2003} is an example of this type of technique and has been used successfully to elucidate atomic-scale dynamics in melting,\cite{cao_femtosecond_2003,sciaini_electronic_2009,siwick_atomic-level_2003} chemical bond breaking,\cite{zewail_laser_1988} etc.
+Ultrafast electron diffraction (UED) or femtosecond electron diffraction (FED) \cite{srinivasan_ultrafast_2003,williamson_clocking_1997,cao_femtosecond_2003} is an example of this type of technique and has been used successfully to elucidate atomic-scale dynamics in melting \cite{cao_femtosecond_2003,sciaini_electronic_2009,siwick_atomic-level_2003}, chemical bond breaking \cite{zewail_laser_1988}, etc.
 Direct time-resolved imaging of similar events can be accomplished using transmission electron microscopy (TEM).
@@
 For completeness, in Section \ref{sec:ms_model} I will introduce, in MKS units, the essential features of the AG model.
-In Chapter \ref{chap:extension}, I present some optimization to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
+In Chapter \ref{chap:extension}, I present some optimizations to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
 I then present my extensions to the Michalik and Sipe AG model which permit the inclusion of generic external forces.
@@
 
-Since my extension to the AG model is valid only within the limits of the analytical method itself, in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
-Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes,\cite{michalik_analytic_2006,michalik_evolution_2009} including the uniform ellipsoid \cite{luiten_how_2004}.
+Since the extension to the AG model is valid only within the limits of the analytical method itself, in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
+Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes \cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid \cite{luiten_how_2004}.
 This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
@@
 It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
-Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
+%Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
+
+In Chapter \ref{chap:model_results}, I apply the AG model to draw some initial insight into the behaviors of ultrafast electron pulses in a UEM column.
+Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion \cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance \cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.
+First, that the electron pulses should be emitted from a large-area source, and second, that every effort must be made to reduce the initial rms transverse momentum (i.e., the emission cone angle).
+
+All of these considerations motivate the construction of the prototype UEM column being built by our group at UIC.
+In Chapter \ref{chap:prototype} I detail the design and fabrication of large-aperture electron-optical elements, which can readily admit these large width electron pulses; included among these are the accelerator and magnetic lenses.
+The, before I conclude, in \ref{chap:photocathode}, I highlight the photocathode engineering work which has been undertaken in an attempt to reduce the rms transverse momentum.
+I will present several different photoemission processes, some of which are yielding very promising results.
+In fact, our recent work may even challenge our current understanding of photoemission.

### 2013-04-11 inc/conclusions.tex
citations are now preceeded by non-breaking space (~)

@@
 
-In this thesis I have presented the extended Analytic Gaussian (AG) model, an effiecient numerical model based on the work of Michalik and Sipe \cite{michalik_analytic_2006}, but extended so as to be useful in modeling the dynamics of ultrafast electron pulses in microscope columns.
+In this thesis I have presented the extended Analytic Gaussian (AG) model, an effiecient numerical model based on the work of Michalik and Sipe~\cite{michalik_analytic_2006}, but extended so as to be useful in modeling the dynamics of ultrafast electron pulses in microscope columns.
 For pulses containing $N$ carriers of charge $e$ and mass $m$, the model is simply these four equations,
@@
 
-The Rose criterion \cite{rose_television_1948} states that in order to have proper gray-scaling in a single-shot DTEM employing a 1k X 1k CCD camera, the pulse must contain at least $10^8$ electrons.
+The Rose criterion~\cite{rose_television_1948} states that in order to have proper gray-scaling in a single-shot DTEM employing a 1k X 1k CCD camera, the pulse must contain at least $10^8$ electrons.
 We have proposed that the term ``ultrafast'' be used in the case where an electron pulse may be completely contained inside the acceleration region of the column.
-The reason for this designation is that such pulses are subject to a different form of Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, which limits its the number of electrons $N_{crit}$ that may be generated by a laser of HW1/eM width $w$ before the pulse is subject to deleterious distortions,
+The reason for this designation is that such pulses are subject to a different form of Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, which limits its the number of electrons $N_{crit}$ that may be generated by a laser of HW1/eM width $w$ before the pulse is subject to deleterious distortions,
 \begin{equation}
@@
 In order to satisfy the both the Rose criterion and this ultrafast Child-Langmuir restriction, we expect that UEM instruments will need to employ a large-area emission source.
-This has necessitated building a custom acceleration region based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007} which boasts a moderately uniform acceleration field near the axis of the accelerator, as well as other large-aperture electron-optical elements for our prototype UEM column.
+This has necessitated building a custom acceleration region based on the design of Togawa \textit{et al.}~\cite{togawa_ceb6_2007} which boasts a moderately uniform acceleration field near the axis of the accelerator, as well as other large-aperture electron-optical elements for our prototype UEM column.
 We have also designed and built a custom laser source capable of generating these electron pulses.
 
-A common figure of merit for electron beams is its normalized transverse emittance \cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, which for the usual case of axially symmetric electron pulses has the form,
+A common figure of merit for electron beams is its normalized transverse emittance~\cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, which for the usual case of axially symmetric electron pulses has the form,
 \begin{equation}
@@
 where $\Delta x$ and $\Delta p_{\smallT}$ are the transverse spatial uncertainty and transverse momentum uncertainty, respectively.
-This quantity is conserved under propagation through free space and perfect electron-optical, though it may increase for imperfect ones \cite{oshea_reversible_1998}.
-The normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred \cite{berger_dc_2009}.
+This quantity is conserved under propagation through free space and perfect electron-optical, though it may increase for imperfect ones~\cite{oshea_reversible_1998}.
+The normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
 Since we have already established that a large area emission source must be used, the only avenue to reduce $\varepsilon_{\smallT}$ is to reduce the transverse momentum uncertainty, which is dependent on the photocathode and photoemission process employed.
@@
 
-I have also presented our efforts to employ periodic nanostructured photocathode surfaces, designed to allow the incident laser to drive a surface plasmon, in hopes that a surface plasmon field might reduce the transverse emittance \cite{zawadzka_evanescent_2001,kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
+I have also presented our efforts to employ periodic nanostructured photocathode surfaces, designed to allow the incident laser to drive a surface plasmon, in hopes that a surface plasmon field might reduce the transverse emittance~\cite{zawadzka_evanescent_2001,kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
 Though we had success in driving the plasmon, we were unable to demonstrate reduction in $\Delta p_{\smallT}$.

### 2013-04-11 inc/design_considerations/considerations.tex
citations are now preceeded by non-breaking space (~)

@@
 It is straightforward to show that in order to make a useful image in an electron microcope, the number of electrons collected must be above some minimum number.
-The Rose criterion \cite{rose_television_1948} indicates that an average detector signal strength of $\sim$100 electrons/pixel is required for adequate gray scaling of the image.
-For a 1k $\times$ 1k CCD camera, this implies an incident electron pulse with $N=10^8$ \cite{armstrong_prospects_2007}.
+The Rose criterion~\cite{rose_television_1948} indicates that an average detector signal strength of $\sim$100 electrons/pixel is required for adequate gray scaling of the image.
+For a 1k $\times$ 1k CCD camera, this implies an incident electron pulse with $N=10^8$~\cite{armstrong_prospects_2007}.
 
@@
 
-It should be noted that for diffraction studies performed in a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse \cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.
+It should be noted that for diffraction studies performed in a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse~\cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.
 
@@
 Since we are now aware that there is a large minimum number of electrons needed per pulse, it is only natural to ask if there is the practical limit on how much charge (i.e., the number of electrons/pulse, $N$) can be extracted from a pulsed laser-driven DC photoelectron gun.
-When the incident laser pulse duration $\tau$ is longer than the time-of-flight (TOF) $t_{gun}$ of a single electron across the acceleration gap $d$ of the DC gun with a potential $V_{DC}$, $\tau > t_{gun} = d \sqrt{2m/eV_{DC}} \approx 0.1-1$ns, typically, then the maximum photocurrent density $J_{CL}$ is given by the well-known Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923}:
+When the incident laser pulse duration $\tau$ is longer than the time-of-flight (TOF) $t_{gun}$ of a single electron across the acceleration gap $d$ of the DC gun with a potential $V_{DC}$, $\tau > t_{gun} = d \sqrt{2m/eV_{DC}} \approx 0.1-1$ns, typically, then the maximum photocurrent density $J_{CL}$ is given by the well-known Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923}:
 \begin{equation}
@@
 
-In this short pulse case, an equivalent diode approximation \cite{valfells_effects_2002} allows for the definition of a critical current density associated with virtual cathode formation in terms of the steady-state Child-Langmuir limit:
+In this short pulse case, an equivalent diode approximation~\cite{valfells_effects_2002} allows for the definition of a critical current density associated with virtual cathode formation in terms of the steady-state Child-Langmuir limit:
 \begin{equation}
@@
 We note that \ref{eq:n_crit} does not express a limit on the number of electrons/pulse $N$ that can be emitted from a pulsed laser-driven DC photoelectron gun, it merely indicates a critical number before virtual cathode formation occurs.
-More electrons can be produced, at the expense of a severely distorted temporal electron pulse profile \cite{valfells_effects_2002}.
+More electrons can be produced, at the expense of a severely distorted temporal electron pulse profile~\cite{valfells_effects_2002}.
 
@@
 
-A common figure of merit for electron beams is its normalized transverse emittance \cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, defined as 
+A common figure of merit for electron beams is its normalized transverse emittance~\cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, defined as 
 \begin{equation}
@@
 \end{equation}
-Note that imperfect electron-optical elements may cause ``emittance growth'' \cite{oshea_reversible_1998}, increasing the pulse's emittance after these elements.
+Note that imperfect electron-optical elements may cause ``emittance growth''~\cite{oshea_reversible_1998}, increasing the pulse's emittance after these elements.
 
-This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred \cite{berger_dc_2009}.
+This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
 Since \ref{eq:minimum_w} specifies a lower bound on the emission spot size $w$, knowing a UEM will exhibit higher resolution by employing an electron source with the minimum possible transverse emittance directly implies that every effort must be made to reduce the initial transverse momentum spread.
@@
 
-In Ref. \cite{berger_dc_2009}, we explore the impact of these limitations for an ultrafast electron diffraction (UED) experiment.
+In Ref.~\cite{berger_dc_2009}, we explore the impact of these limitations for an ultrafast electron diffraction (UED) experiment.
 We find that for a desired spatial resolution per pixel $\Delta X$ (related to the magnification) using a $N_p$ x $N_p$ detector, one may only resolve a fractional change $\Delta a / a$ in a crystalline material's Bragg plane spacing $a$ in the $M^{th}$ diffraction order if
@@
 One might think to narrow this discrepancy by simply better matching the photocathode work function and the laser photon energy, thereby reducing the excess photoemission energy $\Delta E$.
-Unfortunately, since the photoemission efficiency is proportional to $\Delta E^2$ \cite{shalaev_electron_1994}, though this would indeed reduce the transverse emittance, it would do so at the cost of a much reduced pulse charge.
+Unfortunately, since the photoemission efficiency is proportional to $\Delta E^2$~\cite{shalaev_electron_1994}, though this would indeed reduce the transverse emittance, it would do so at the cost of a much reduced pulse charge.
 In Chapter \ref{chap:photocathode}, I will explore several more exotic photoemission processes in an attempt to reduce $\varepsilon_{\smallT}$ without this penalty.

### 2013-04-11 inc/extension/external_forces/formal.tex
citations are now preceeded by non-breaking space (~)

@@
 
-Simulation of electron pulse propagation in UEM and modern UED apparatus \cite{siwick_ultrafast_2002,oudheusden_electron_2007,reed_femtosecond_2006} requires the ability to model electron-optical elements designed to control the three-dimensional space-charge and dispersion induced broadening of the pulse, and hence deliver electron pulses of the desired quality to the specimen.
+Simulation of electron pulse propagation in UEM and modern UED apparatus~\cite{siwick_ultrafast_2002,oudheusden_electron_2007,reed_femtosecond_2006} requires the ability to model electron-optical elements designed to control the three-dimensional space-charge and dispersion induced broadening of the pulse, and hence deliver electron pulses of the desired quality to the specimen.
 The Analytic Gaussian model as presented by Michalik and Sipe (see Section \ref{sec:ms_model}) is only useful for modeling free-space propagation.
 This is because there is no allowance for external forces, such as those imparted by the ``optical'' elements of an electron microscope column.
-In this section, I show that the AG model can be extended to include external forces on the pulse, such as the influence of magnetic electron lenses and RF pulse compression cavities \cite{oudheusden_electron_2007,veisz_hybrid_2007}.
+In this section, I show that the AG model can be extended to include external forces on the pulse, such as the influence of magnetic electron lenses and RF pulse compression cavities~\cite{oudheusden_electron_2007,veisz_hybrid_2007}.
 
@@
 While the free-space AG model does not contain any external forces, it does include an internal force: the Coulomb force is included to model the repulsion of the pulse's electrons from each other.
-As this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics \cite{michalik_analytic_2006}.
+As this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics~\cite{michalik_analytic_2006}.
 
@@
 
-The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}.
+The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations~\cite{michalik_analytic_2006}.
 As a result, we emphasize that the presented AG model extension is valid only for the action of linear forces (i.e., forces that have a linear dependence on distance in the pulse frame) on the electron pulse due to the model's self-similar Gaussian approximation.
-Since nonlinear forces are not considered, ``emittance growth'' effects arising from imperfect lenses cannot be modeled \cite{oshea_reversible_1998}.
+Since nonlinear forces are not considered, ``emittance growth'' effects arising from imperfect lenses cannot be modeled~\cite{oshea_reversible_1998}.
 The inclusion of the action of nonlinear forces (e.g., due to higher-order space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations.
@@
 
-Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes \cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid \cite{luiten_how_2004}.
+Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
 This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
-Consequently, the AG approach is applicable to both the single electron per pulse limit \cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate \cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
+Consequently, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
 It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.

### 2013-04-11 inc/extension/external_forces/practical.tex
citations are now preceeded by non-breaking space (~)

@@
 To model electron pulses traveling through realistic electron microcope columns, these abstract mechanisms must be developed further to properly describe each specific electron-optical element.
-In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression \cite{oudheusden_electron_2007} and acceleration %TODO refs
+In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression~\cite{oudheusden_electron_2007} and acceleration %TODO refs
 ) and DC acceleration regions (i.e. electron gun dynamics).
@@
 
-The full description of the motion of an electron (in cylindrical coordinates $r, \theta, z$) in an azimuthally symmetric magnetic field ($\vec{B} = B_r \hat{r} + B_z \hat{z}$) is easily derived \cite{el-kareh_electron_1970},
+The full description of the motion of an electron (in cylindrical coordinates $r, \theta, z$) in an azimuthally symmetric magnetic field ($\vec{B} = B_r \hat{r} + B_z \hat{z}$) is easily derived~\cite{el-kareh_electron_1970},
 \begin{subequations} \label{eq:lens_eq_of_motion}
@@
 Compounding the problem, the full description of the off-axis field of a real (non-infinite) air-core solenoid is also surprisingly complicated.
-Montgomery and Terrell \cite{montgomery_some_1961} provide a full treatment for many lens configurations, including a tractable treatment for a single current loop of radius $a$, again in cylindrical coordinates,
+Montgomery and Terrell~\cite{montgomery_some_1961} provide a full treatment for many lens configurations, including a tractable treatment for a single current loop of radius $a$, again in cylindrical coordinates,
 \begin{subequations} \label{eq:field_of_loop}
@@
 
-In the longitudinal ($ z $) direction, a resonant radio-frequency (RF) cavity can be used in a UEM column or UED experiment \cite{oudheusden_electron_2007,fill_sub-fs_2006} to act as a pulse compression or `temporal focusing' element.
+In the longitudinal ($ z $) direction, a resonant radio-frequency (RF) cavity can be used in a UEM column or UED experiment~\cite{oudheusden_electron_2007,fill_sub-fs_2006} to act as a pulse compression or `temporal focusing' element.
 The model of the RF cavity is different from the magnetic lens in two distinct ways.
@@
 
-Although there are many possible resonant cavity structures \cite{oudheusden_electron_2007,humphries_principles_1986}, for simplicity, this analysis will consider only the evacuated cylindrical (or `pillbox') $\text{TM}_{010}$ cavity for which the electrons propagate down the axis \cite{fill_sub-fs_2006,humphries_principles_1986}.
-The known mathematical forms of the electric and magnetic components of this resonant mode together with a super-Gaussian envelope function (\ref{eq:reg_of_influence}) allow the three main effects of a pill-box RF cavity element on the electron pulse to be described within the approximation of the linear force extension to the AG model: the primary action of the pulse compression (or acceleration and expansion dependent upon the relative RF phase), and the transverse lens effects due to the magnetic component of the $\text{TM}_{010}$ resonance and the electric RF cavity fringe fields at its entrance and exit apertures \cite{kim_rf_1989}.
+Although there are many possible resonant cavity structures~\cite{oudheusden_electron_2007,humphries_principles_1986}, for simplicity, this analysis will consider only the evacuated cylindrical (or `pillbox') $\text{TM}_{010}$ cavity for which the electrons propagate down the axis~\cite{fill_sub-fs_2006,humphries_principles_1986}.
+The known mathematical forms of the electric and magnetic components of this resonant mode together with a super-Gaussian envelope function (\ref{eq:reg_of_influence}) allow the three main effects of a pill-box RF cavity element on the electron pulse to be described within the approximation of the linear force extension to the AG model: the primary action of the pulse compression (or acceleration and expansion dependent upon the relative RF phase), and the transverse lens effects due to the magnetic component of the $\text{TM}_{010}$ resonance and the electric RF cavity fringe fields at its entrance and exit apertures~\cite{kim_rf_1989}.
 
@@
 The azimuthal magnetic component of the resonant mode oscillates 90$^{\circ}$ out of phase with the electric component and with a $\operatorname{ J_{1} }$ Bessel radial form; i.e., $ \operatorname{ J_{1} }( k r / a ) \sim k r / a $ for small $r$.
-Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $ \cite{kim_rf_1989}, where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\ref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\ref{eq:reg_of_influence}).
-In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures (see for example $1 \frac{1}{2}$-cell RF photo-gun designs \cite{mcdonald_design_1988}).
+Symmetry and Gauss' Law dictate that in the vicinity of the axial center of the entrance and exit apertures of the RF cavity there exists a radial component of the oscillating electric field determined by $ E_{r} = -\left ( r / 2 \right ) \left ( \partial E_{z} / \partial z \right )_{r=0} $~\cite{kim_rf_1989}, where the form of the axial electric field for $\text{TM}_{010}$ oscillation (\ref{eq:RF_Efield}) implies that, in the context of this analysis, its axial gradient is solely dependent upon the envelope function for the RF cavity (\ref{eq:reg_of_influence}).
+In practice, of course, the exact form of $E_{r}(z)$ is strongly tied to the size and shape of the RF cavity apertures (see for example $1 \frac{1}{2}$-cell RF photo-gun designs~\cite{mcdonald_design_1988}).
 For the case of an evacuated cavity, these two contributions result in a net force,
@@
 In obtaining \ref{eq:RF_Force_r}, which is also of the form for use in \ref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \ref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
-As will be seen in Section \ref{sec:rf_cav_charge}, under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $ \cite{kim_rf_1989}, even though its magnitude is maximized for $\phi = 0$.
+As will be seen in Section \ref{sec:rf_cav_charge}, under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $~\cite{kim_rf_1989}, even though its magnitude is maximized for $\phi = 0$.
 Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.

### 2013-04-11 inc/extension/initial_conditions/initial_conditions.tex
killing more TODOs
citations are now preceeded by non-breaking space (~)

@@
 \section{Determining Appropriate Initial Conditions} \label{sec:initial_conditions}
-For a DTEM or UED apparatus, where the electron pulse is produced in a laser-driven photo-electron gun, the implementation of the AG model of Michalik and Sipe \cite{michalik_analytic_2006} requires knowledge of the initial ($ t = 0 $) bunch parameters immediately after generation at the photocathode surface inside the acceleration region of the gun.
+For a DTEM or UED apparatus, where the electron pulse is produced in a laser-driven photo-electron gun, the implementation of the AG model of Michalik and Sipe~\cite{michalik_analytic_2006} requires knowledge of the initial ($ t = 0 $) bunch parameters immediately after generation at the photocathode surface inside the acceleration region of the gun.
 For the purposes of this thesis, we will make a simple approximation to the behavior of the gun in order to provide generic initial conditions for the electron pulse.
@@
 For spatially-uniform photoemission from a planar photocathode, from symmetry we have $ \gamma_{\smallT} ( 0 ) = 0 $; i.e., there is no initial transverse spatial momentum chirp across the electron bunch.
-Finally, thanks to the careful analysis of Dowell and Schmerge \cite{dowell_quantum_2009} (confirmed by Jensen \textit{et.
-al.} \cite{jensen_emittance_2010}), we know that metallic photocathode's initial transverse momentum variance at $ t = 0 $ is given by
+Finally, thanks to the careful analysis of Dowell and Schmerge~\cite{dowell_quantum_2009} (confirmed by Jensen \textit{et.
+al.}~\cite{jensen_emittance_2010}), we know that metallic photocathode's initial transverse momentum variance at $ t = 0 $ is given by
 \begin{equation} \label{eq:jensen_eta}
@@
 in the limit where the maximum excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ is significantly greater than the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
-%TODO include TE value for $\Delta E < 0$  is $\eta_{\smallT} = m k_B T$.
+For completeness we note that for thermionic emission $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ where $k_B$ is the Boltzmann constant and $T_e$ is the electron temperature~\cite{dowell_quantum_2009,jensen_emittance_2010}..
 
@@
 Since the electric field has no time to influence the front of the pulse before the back is also generated we can say that the front of the pulse has a velocity equal to $v_{max}$ and the back is essentially stationary just after generation.
-Further, since the photoemission efficiency is proportional to $\Delta E^2$ \cite{shalaev_electron_1994} the distribution of emitted electrons vs velocity is proportional to $v^4$; therefore to a first approximation it is reasonable to say that in this case $\Delta v \approx v_{max}$.
+Further, since the photoemission efficiency is proportional to $\Delta E^2$~\cite{shalaev_electron_1994} the distribution of emitted electrons vs velocity is proportional to $v^4$; therefore to a first approximation it is reasonable to say that in this case $\Delta v \approx v_{max}$.
 The pulse's initial spatial variance $\Delta z$ is therefore approximately to the spread of distances traveled during the laser pulse duration $\Delta t_L$,
@@
 
-We recognize that these initial conditions for an electron pulse generated by a laser-driven DC gun are idealized in that they neglect several important effects; in particular, the interplay of dispersive ($ \eta_{i}(0) \neq 0 $) and space-charge effects in the acceleration region of the photoelectron gun and transverse and longitudinal effects due to the divergence of the DC field in the vicinity of the gun anode \cite{berger_dc_2009,togawa_ceb6_2007}.
+We recognize that these initial conditions for an electron pulse generated by a laser-driven DC gun are idealized in that they neglect several important effects; in particular, the interplay of dispersive ($ \eta_{i}(0) \neq 0 $) and space-charge effects in the acceleration region of the photoelectron gun and transverse and longitudinal effects due to the divergence of the DC field in the vicinity of the gun anode~\cite{berger_dc_2009,togawa_ceb6_2007}.
 Such effects and the limits of the validity of the longitudinal terms of \ref{eq:summary}, within the context of the AG model, will be presented in Section \ref{sec:bin_model}.
 However, \ref{eq:summary} also represent a defined set of initial conditions that allows the effect of electron optics (i.e., magnetic lenses and RF cavities) on electron pulse propagation to be more readily understood.
-Moreover, they reflect the fact that the majority of DC photoelectron guns are driven by laser pulses with spatial and temporal profiles that are close to Gaussian \cite{williamson_clocking_1997,sciaini_electronic_2009}.
-Only for specialized cases, such as the generation of uniform ellipsoid electron bunch shapes, is spatial and temporal manipulation of the incident laser pulse considered \cite{luiten_how_2004,li_generating_2008}.
+Moreover, they reflect the fact that the majority of DC photoelectron guns are driven by laser pulses with spatial and temporal profiles that are close to Gaussian~\cite{williamson_clocking_1997,sciaini_electronic_2009}.
+Only for specialized cases, such as the generation of uniform ellipsoid electron bunch shapes, is spatial and temporal manipulation of the incident laser pulse considered~\cite{luiten_how_2004,li_generating_2008}.
 
@@
 This limitation on functional form severely limits the accuracy with which a determined spatially constant value of $\eta_z(0)$ represents the actual distribution of momentum uncertainties computed by the binning model.
-The self-similar theory of Michalik and Sipe \cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch; with a more realistic fit to the simulated initial data, the pulse would eventually evolve to be non-Gaussian.
-%Consequently, we will assume that $ \eta_{ z } $ is a constant over the bunch with a value given by $ \eta_{\smallT } ( 0 ) \cong 3.85 \eta{ z } ( 0 ) $: indeed in this short pulse regime, one can show that over the central $ \sim 75 \% $ of the electrons in the inital photo-generated bunch the variation in $ \sqrt{ \eta_{ z } ( 0 ) } $ is only about $ \pm 25 \% $.
+The self-similar theory of Michalik and Sipe~\cite{michalik_analytic_2006} is not equipped to handle such a variation as it assumes a constant $ \eta_{ z } $ over the electron bunch; with a more realistic fit to the simulated initial data, the pulse would eventually evolve to be non-Gaussian.

### 2013-04-11 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
more punctuation near equations and some equation source reformmatting
citations are now preceeded by non-breaking space (~)

@@
 \begin{equation}
-  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0
+  \frac{d}{dt} (\sigma_{\alpha} \eta_{\alpha}) = \sigma_{\alpha} \frac{d \eta_{\alpha}}{dt} + \eta_{\alpha} \frac{d \sigma_{\alpha}}{dt} = 0 \,\text{,}
 \end{equation}
@@
 
-We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed \textit{et al.} \cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the trasverse spatial emittance of an electron beam \cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
+We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed \textit{et al.}~\cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the trasverse spatial emittance of an electron beam~\cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
 Additionally, though the AG model explicitly satisfies Liouville's Theorem, there are cases where the pulse's phase space product may change in a realistic system.
-This effect is called ``emittance growth'' and is most commonly seen in TEM as a result of aberations in lens systems \cite{oshea_reversible_1998}.
+This effect is called ``emittance growth'' and is most commonly seen in TEM as a result of aberations in lens systems~\cite{oshea_reversible_1998}.
 As will be seen in Section \ref{extension_limitations}, the AG model will, of necessity, only consider perfect lenses.

### 2013-04-11 inc/gun/gun_design.tex
citations are now preceeded by non-breaking space (~)

@@
 
-These requirements led to the fabrication of cathode/Wehnelt and anode pair based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007} (pictured in \ref{fig:togawa-pic}).
+These requirements led to the fabrication of cathode/Wehnelt and anode pair based on the design of Togawa \textit{et al.}~\cite{togawa_ceb6_2007} (pictured in \ref{fig:togawa-pic}).
 A schematic of this design, overlayed by a Finite Element Model (FEM) simulation of the electic field, is shown in \ref{fig:gun-field}.

### 2013-04-11 inc/gun/gun_model.tex
added cc notices to the tops of several files
more corrections and punctuation
citations are now preceeded by non-breaking space (~)

@@
+%This work is licensed under the Creative Commons Attribution-NonCommercial-NoDerivs 3.0 United States License.
+To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-nd/3.0/us/ or send a letter to Creative Commons, 444 Castro Street, Suite 900, Mountain View, California, 94041, USA.
+
 \subsection{Acceleration Region} \label{sec:gun_model}
@@
 \begin{equation}
-  \vec{F}(r^{\prime}=0,z^{\prime}) = F_z (0,z^{\prime}) \hat{z} = e E_z (0,z^{\prime}) \hat{z}
+  \vec{F}(r^{\prime}=0,z^{\prime}) = F_z (0,z^{\prime}) \hat{z} = e E_z (0,z^{\prime}) \hat{z} \,\text{,}
 \end{equation}
@@
 \begin{equation}
-  F_z(0,z^{\prime} - z) \approx F_z(0,z^{\prime} - z) + \frac{\partial F_z}{\partial z^{\prime}} z
+  F_z(0,z^{\prime} - z) \approx F_z(0,z^{\prime} - z) + \frac{\partial F_z}{\partial z^{\prime}} z \,\text{,}
 \end{equation}
@@
 \begin{equation}
-  F_r(r,z^{\prime}) \approx F_r(r,z^{\prime}) + \frac{\partial F_r}{\partial r^{\prime}} r
+  F_r(r,z^{\prime}) \approx F_r(r,z^{\prime}) + \frac{\partial F_r}{\partial r^{\prime}} r \,\text{.}
 \end{equation}
-Without an analytic form for $F$, we can appeal to Gauss' law (in absense of external charge)
+Without an analytic form for $F$, we can appeal to Gauss' law (in absense of external charge),
 \begin{gather}
   \frac{\partial F_x}{\partial x^{\prime}} + \frac{\partial F_y}{\partial y^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 2 \frac{\partial F_{\smallT}}{\partial r^{\prime}} + \frac{\partial F_z}{\partial z^{\prime}} = 0 \\
-  \Rightarrow \frac{\partial F_{\smallT}}{\partial r^{\prime}} = - \frac{1}{2} \frac{\partial F_z}{\partial z^{\prime}}
+  \Rightarrow \frac{\partial F_{\smallT}}{\partial r^{\prime}} = - \frac{1}{2} \frac{\partial F_z}{\partial z^{\prime}} \,\text{,}
 \end{gather} 
@@
 \begin{equation} \label{eq:anode_tanh}
-  F_z(0,z^{\prime}) = \frac{eV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right)
+  F_z(0,z^{\prime}) = \frac{eV}{2z^{\prime}_A} \left( 1 - \tanh \left( \frac{ z^{\prime} - z^{\prime}_A }{ z^{\prime}_A / s } \right) \right) \,\text{,}
 \end{equation}
@@
 Interestingly the length of the pulse is stabilized on exiting the anode, coming very nearly to a constant length (not expanding nor contracting).
-This effect is due to the slower electrons remaining in the cavity longer and thus receiving an additional kick, as described in \cite{oudheusden_electron_2007}.
+This effect is due to the slower electrons remaining in the cavity longer and thus receiving an additional kick, as described in~\cite{oudheusden_electron_2007}.
 This longitudinal electric field shape is compared to the electric field computed by finite-element (FEM) simulations of our acceleration cavity in Section \ref{sec:gun_design}.

### 2013-04-11 inc/hardware/laser/laser.tex
killing TODOs
merged remaining corrections from Carolyn
citations are now preceeded by non-breaking space (~)

@@
 
-In order to experiment using very short laser pulses, we have designed and built a custom Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) laser system \cite{berger_high-power_2008}.
+In order to experiment using very short laser pulses, we have designed and built a custom Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) laser system~\cite{berger_high-power_2008}.
 In practice, the flexibility afforded by having a custom-built system has outweighed its original performance benefit.
@@
 Yb:KGW is becoming a widely used solid-state gain medium for ultrafast laser systems.
-As with other Yb-doped laser crystals \cite{Brenier_new_criteria}, this is primarily due to the gain medium's advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
-Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated \cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially \cite{website_amplitude,website_solar}.
-Such lasers have used dichroic mirrors \cite{Brunner_diode_pumped,Major_femtosecond_2006,Paunescu_diode_2004,Major_extended_2006}, polarization-coupling \cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers \cite{website_amplitude,Brunner_pulses_2002}) to pump the laser crystal in an efficient manner.
+As with other Yb-doped laser crystals~\cite{Brenier_new_criteria}, this is primarily due to the gain medium's advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load.
+Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated~\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially~\cite{website_amplitude,website_solar}.
+Such lasers have used dichroic mirrors~\cite{Brunner_diode_pumped,Major_femtosecond_2006,Paunescu_diode_2004,Major_extended_2006}, polarization-coupling~\cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers~\cite{website_amplitude,Brunner_pulses_2002}) to pump the laser crystal in an efficient manner.
 
 % page 8631
-The spectroscopic properties of the biaxial Yb:KGW laser crystal \cite{Biswal_thermo_optical_05} indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength \cite{Brenier_new_criteria}.
-However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW) \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that of the absorption along the optical $N_m$-axis \cite{Holtom_mode_locked_2006,Biswal_thermo_optical_05}.
-This optimal requirement for orthogonal linear pump and laser polarizations led Holtom \cite{Holtom_mode_locked_2006} to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
+The spectroscopic properties of the biaxial Yb:KGW laser crystal~\cite{Biswal_thermo_optical_05} indicate that the optimum diode pumping arrangement is for 980nm radiation polarized parallel to the optical $N_m$-axis since the crystal has the largest absorption cross-section in this configuration and, hence, the smallest absorption saturation intensity to drive the gain medium into a quasi 4-level system or to achieve transparency at the emission wavelength~\cite{Brenier_new_criteria}.
+However, as is the case for Yb-doped potassium yttrium tungstate (Yb:KYW)~\cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, mode-locked Yb:KGW lasers should ideally oscillate with radiation polarized parallel to the optical $N_p$-axis to access the broadest gain bandwidth around 1040nm, where the emission cross-section is comparable to that of the absorption along the optical $N_m$-axis~\cite{Holtom_mode_locked_2006,Biswal_thermo_optical_05}.
+This optimal requirement for orthogonal linear pump and laser polarizations led Holtom~\cite{Holtom_mode_locked_2006} to employ a polarization-coupling scheme to diode pump longitudinally a 10W, sub-500fs, Yb:KGW laser.
 
-%TODO is this first sentance correct or should I use "I" voice?
-In this chapter, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis \cite{Biswal_thermo_optical_05,mochalov_laser_1997,pujol_crystalline_1999}).
-The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media \cite{Rimington_thermal_lens_2004}.
+In this chapter, a high-power, femtosecond Yb:KGW oscillator is presented that also accesses the broader emission bandwidth parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis), yet employs a simpler geometry for polarized diode-pumping on the strongest 980nm absorption line parallel to the optical $N_m$-axis (rotated by $\sim\,20^\circ$ from the crystallographic $a$-axis~\cite{Biswal_thermo_optical_05,mochalov_laser_1997,pujol_crystalline_1999}).
+The laser head design is based on the thermal lens shaping (TLS) technique developed in-house for astigmatism compensation in diode-pumped Nd-doped lasers with Brewster-cut gain media~\cite{Rimington_thermal_lens_2004}.
 It employs a simple single layer \ce{SiO2} anti-reflection coating on the gain medium to provide minimal loss for $p$-polarized intracavity 1040nm laser radiation and less than 0.5\% surface reflection for orthogonal $s$-polarized 980nm pump light.
@@
 The Watt-level green output power from the frequency-doubled sub-picosecond Yb:KGW oscillator corresponds to a visible peak laser pulse power in excess of 100kW, which is well-suited for further frequency conversion through harmonic generation or pumping of an ultrashort pulse optical parametric oscillator.
-%TODO change voice/check ->
-To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk \cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped \cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators \cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media \cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.
-%TODO replace to "This output will then be use to ..." and link to something from previous
-%In our case, the visible femtosecond pulses will be used as the driving laser radiation source for an ultrafast electron microscope \cite{king_ultrafast_2005} --- a technology that seeks to combine the sub-nanometer spatial resolution of electron microscopy with the sub-picosecond temporal resolution afforded by today's ultrashort pulse laser systems.
-%Critical to this technology is likely to be the development of a femtosecond photo-electron gun with a low spatial emittance \cite{zawadzka_evanescent_2001}.
-%The coherent control of photoemission through visible laser-driven plasmon excitations on large-area (a few mm2), gold or silver, nano-patterned photocathodes may provide a mechanism by which this can be achieved \cite{zawadzka_evanescent_2001}.
+To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk~\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped~\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators~\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media~\cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.
 
@@
 
-The design of the laser head follows the example of Holtom \cite{Holtom_mode_locked_2006}, which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism \cite{Rimington_thermal_lens_2004}.
+The design of the laser head follows the example of Holtom~\cite{Holtom_mode_locked_2006}, which ensures efficient pumping and access to the broadest emission bandwidth of Yb:KGW, while also employing the thermal lens shaping (TLS) technique to compensate for astigmatism~\cite{Rimington_thermal_lens_2004}.
 To ensure efficient operation of a Yb:KGW laser, the dominant absorption feature at 980nm for radiation polarized parallel to the optical $N_m$-axis ($\sim\,20^\circ$ from the crystallographic a-axis) should be pumped using high brightness diode lasers.
-On the other hand, as with \ce{Yb$:$KY(WO4)2} \cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis \cite{Holtom_mode_locked_2006}) in this biaxial material.
+On the other hand, as with \ce{Yb$:$KY(WO4)2}~\cite{Liu_diode_pumped_2001,Killi_high_peak_2005}, the broadest emission bandwidth for ultrashort pulse generation occurs for radiation polarized parallel to the optical $N_p$-axis (parallel to the crystallographic $b$-axis~\cite{Holtom_mode_locked_2006}) in this biaxial material.
 %page 8632
-Instead of using a polarization-coupled pumping geometry \cite{Holtom_mode_locked_2006} to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in \ref{fig:crystal}\subref{fig:crystal_geo}, where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
-At an angle of incidence of 32.3$^\circ$, this coating theoretically has zero reflection loss for $p$-polarized 1040nm laser radiation parallel to the $b$-axis propagating perpendicular to the $a$-$b$ plane of the crystal (\ref{fig:crystal}\subref{fig:crystal_coating}); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98 \cite{Biswal_thermo_optical_05,pujol_crystalline_1999}.
+Instead of using a polarization-coupled pumping geometry~\cite{Holtom_mode_locked_2006} to ensure these optimum pump-lasing conditions, we employ the non-Brewster crystal geometry and cut shown in \ref{fig:crystal}\subref{fig:crystal_geo}, where a single 193nm-thick \ce{SiO2} anti-reflection coating is applied to both polished 3x10mm Yb:KGW crystal faces.
+At an angle of incidence of 32.3$^\circ$, this coating theoretically has zero reflection loss for $p$-polarized 1040nm laser radiation parallel to the $b$-axis propagating perpendicular to the $a$-$b$ plane of the crystal (\ref{fig:crystal}\subref{fig:crystal_coating}); i.e., the preferred optical $N_p$-axis with refractive index $n_p$ = 1.98~\cite{Biswal_thermo_optical_05,pujol_crystalline_1999}.
 The \ce{SiO2} coating also has an appropriate bandwidth for femtosecond laser operation in excess of 100nm (to the
@@
 
-Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH \cite{website_limo}).
-The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm full-width half-maximum (FWHM) 980nm absorption line of Yb:KGW \cite{Paunescu_diode_2004,Biswal_thermo_optical_05,website_EKSPLA}.
+Efficient pumping is achieved with two, 35W, dual-axis collimated and TM-polarized laser diode arrays operating at 980nm (HLU35C10x5-980 from LIMO GmbH~\cite{website_limo}).
+The $\sim$3nm emission linewidth of the two laser diode arrays is well matched to the 3.7nm full-width half-maximum (FWHM) 980nm absorption line of Yb:KGW~\cite{Paunescu_diode_2004,Biswal_thermo_optical_05,website_EKSPLA}.
 Moreover, the two dual-axis collimated laser diodes have full-angle $1/e^2$ intensity beam divergences of $\theta_y \approx$ 4.4mrad and $\theta_x \approx$ 6.8mrad in the horizontal ($x$-direction parallel to the emitter array) and vertical ($y$) directions respectively.
@@
 The Yb concentration was chosen to be 2\% atomic doping to ensure that more than 90\% of the pump radiation was absorbed over the 3mm Yb:KGW crystal thickness.
-With the employed counter-propagating pump geometry, the resultant maximum average pump irradiance over the crystal length is then roughly 3 times the $\sim$6kW/cm$^2$ saturation intensity of the dominant 980nm absorption line \cite{Brunner_powerful_2004}.
-The laser crystal was obtained from NovaPhase \cite{website_nova}.
+With the employed counter-propagating pump geometry, the resultant maximum average pump irradiance over the crystal length is then roughly 3 times the $\sim$6kW/cm$^2$ saturation intensity of the dominant 980nm absorption line~\cite{Brunner_powerful_2004}.
+The laser crystal was obtained from NovaPhase~\cite{website_nova}.
 
-To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique \cite{Rimington_thermal_lens_2004} to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal.
+To ensure a stigmatic laser resonator, we also employ the thermal lens shaping (TLS) technique~\cite{Rimington_thermal_lens_2004} to compensate for the astigmatism induced by the non-normal incidence geometry of the laser crystal.
 For the employed non-Brewster crystal cut (\ref{fig:crystal}\subref{fig:crystal_geo}), a ray
@@
 is the on-axis temperature, is given by
-%TODO check W symbol consistent with the rest of the paper
 \begin{equation} \label{eq:laser_ratio_ba}
@@
   \frac{ \cos^2 \theta_2 \left( \dfrac{dn}{dT} \right) + 2 \alpha_{\smallT} \Delta \left( n \cos\theta_2 - \cos\theta_1 \right) }
-       { \cos^2 \theta_1 \left[ \left( \dfrac{dn}{dT} \right) + 2 \alpha_{\smallT} \Delta \left( n \cos\theta_2 - \cos\theta_1 \right) \right] }
-  \quad\text{.}
+       { \cos^2 \theta_1 \left[ \left( \dfrac{dn}{dT} \right) + 2 \alpha_{\smallT} \Delta \left( n \cos\theta_2 - \cos\theta_1 \right) \right] } \,\text{.}
 \end{equation}
-Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98 \cite{Biswal_thermo_optical_05,pujol_crystalline_1999}), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
-expansion with coefficient $\alpha T$ \cite{Rimington_thermal_lens_2004}.
-Generally, $\Delta \approx W / l$ , where $W$ is the transverse size (radius) of the pumped region \cite{Koechner_thermal_1970}.
+Here, $\frac{dn}{dT}$ and $n$ are the thermo-optic coefficient and refractive index of the medium respectively (in this case for radiation polarized along the optical $N_p$-axis), $\theta_1$ is the 32.3$^\circ$ angle of incidence, $\theta_2$ = 15.7$^\circ$ is the angle of refraction (as $n_p$ = 1.98~\cite{Biswal_thermo_optical_05,pujol_crystalline_1999}), and $\Delta$ is the fraction of the crystal length $l$ contributing to the bowing of the crystal faces under thermal
+expansion with coefficient $\alpha T$~\cite{Rimington_thermal_lens_2004}.
+Generally, $\Delta \approx W / l$ , where $W$ is the transverse size (radius) of the pumped region~\cite{Koechner_thermal_1970}.
 In our case, $l \approx$ 3mm and $W \approx$ 0.2mm, so that $\Delta < 0.1$, implying that the thermal duct due to a non-zero $\frac{dn}{dT}$ primarily determines the required $B_T$/$A_T$ ratio.
@@
 \begin{equation} \label{eq:heat_eqn}
-  \kappa \nabla^2 T(x,y) = Q(x,y)
-  \quad\text{,}
+  \kappa \nabla^2 T(x,y) = Q(x,y) \,\text{,}
 \end{equation}
 where $\kappa$ is the thermal diffusion coefficient.
-Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry \cite{Rimington_thermal_lens_2004}.
-The specifications of our dual-axis collimated diodes \cite{website_limo} indicate that heat source can be represented mathematically as
+Thermal diffusion in the third $z$-direction has been neglected due to the relatively uniform longitudinal heat deposition resulting from our counter-propagating pump geometry~\cite{Rimington_thermal_lens_2004}.
+The specifications of our dual-axis collimated diodes~\cite{website_limo} indicate that heat source can be represented mathematically as
 \begin{equation}
-  Q(x,y) = Q_0 \exp \left[ - \frac{x^2}{W_x^2} - \left( \frac{y^2}{W_y^2} \right)^3 \right]
-  \quad\text{;}
+  Q(x,y) = Q_0 \exp \left[ - \frac{x^2}{W_x^2} - \left( \frac{y^2}{W_y^2} \right)^3 \right] \,\text{;}
 \end{equation}
 that is, the focused pump radiation is described accurately by a Gaussian irradiance distribution in the horizontal ($x$) direction with half-width 1/e maximum (HW1/eM) spot size $W_x$ and a super-Gaussian of order 3 in the vertical ($y$) direction with spot size $W_y$.
-\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique \cite{Rimington_thermal_lens_2004}; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
+\ref{eq:heat_eqn} can then be solved using the appropriate boundary conditions for our 3x10mm crystal cross-section for the employed TLS technique~\cite{Rimington_thermal_lens_2004}; namely, heat removal only through the top crystal face implies $T(x,y = \pm 1.5\text{mm}) = 0$ (or arbitrary constant) and $\dfrac{\partial T}{\partial x} = 0$ at $ x = \pm 5.0$mm for any $y$.
 For a fixed vertical pump spot size of $W_y = 0.2$mm, our analysis indicates that in order to achieve the required ratio $B_T / A_T = 1.3$ in the parabolic approximation for $T(x,y)$ about $(x,y) = (0,0)$ a horizontal pump spot size $W_x = 0.3$mm is needed; in other words, the required $Q(x,y)$ heat source spot size ratio (or ellipticity) $W_x / W_y = 1.5$.
@@
 Also shown are the parabolic approximations to the pump-induced temperature distribution (dashed lines) and representative TEM$_{00}$ Gaussian laser modes (shaded area) in the Yb:KGW crystal with a HW1/eM field mode sizes of $w_y = 160\mu$m and $w_x = 182 \mu$m (14\% larger due to refraction (\ref{fig:crystal}\subref{fig:crystal_geo}).
-This mode size ensures efficient pump-probe overlap for the TEM$_{00}$-mode while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium \cite{Brenier_new_criteria}; in particular, in the vertical direction (\ref{fig:laser-thermal}(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
+This mode size ensures efficient pump-probe overlap for the TEM$_{00}$-mode while minimizing the loss due to ground state absorption in the unpumped regions of the Yb-doped gain medium~\cite{Brenier_new_criteria}; in particular, in the vertical direction (\ref{fig:laser-thermal}(b)) where $w_y$ greater than about $0.8 W_y$ leads to significant absorptive loss as the extrema of the laser mode extend beyond the sharply-bounded super-Gaussian pump region.
 
@@
 Clearly, this can be readily accommodated by the $W_x = 0.3$mm horizontal pump spot size when $w_x \approx 180\mu$m.
-However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium \cite{Brenier_new_criteria}.
+However, even though $W_x > w_x$, the Gaussian (rather than super-Gaussian) horizontal pump mode profile is still expected to assist in ensuring TEM$_{00}$-mode operation since the net horizontal gain profile (gain minus absorptive loss) will fall off more rapidly than the pump profile for a quasi-three level Yb-doped gain medium~\cite{Brenier_new_criteria}.
 
-We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2}) \cite{Biswal_thermo_optical_05,mochalov_laser_1997} and stress-induced refractive index changes \cite{Yumashev_laser_2007}.
-For the employed crystal cut in our TLS geometry (\ref{fig:crystal}\subref{fig:crystal_geo}), the anisotropy in $\kappa$ \cite{mochalov_laser_1997,website_EKSPLA} is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
+We note that the above thermal lens analysis neglects the anisotropy of the thermal conductivity $\kappa$ in potassium gadolinium tungstate (\ce{KGd(WO4)2})~\cite{Biswal_thermo_optical_05,mochalov_laser_1997} and stress-induced refractive index changes~\cite{Yumashev_laser_2007}.
+For the employed crystal cut in our TLS geometry (\ref{fig:crystal}\subref{fig:crystal_geo}), the anisotropy in $\kappa$~\cite{mochalov_laser_1997,website_EKSPLA} is not expected to have a significant effect; the laser head geometry dictates that the dominant pump-induced heat conduction will be in the saggital ($y$) direction along the crystal $a$-axis where $\kappa \approx 2.6$W/(mK) (heat extraction only through the top and bottom 3x10mm crystal faces), so that the 46\% larger value of $\kappa$ along the $b$-axis ($x$) direction at 3.8W/(mK) should not have a large influence on the heat conduction.
 %page 8635
-We see no observable effect from changes due to stress in the pumped gain medium, due both to the much stronger effect of the natural birefringence of the crystal and the laser polarization being along the optical $N_p$-axis \cite{Holtom_mode_locked_2006}.
+We see no observable effect from changes due to stress in the pumped gain medium, due both to the much stronger effect of the natural birefringence of the crystal and the laser polarization being along the optical $N_p$-axis~\cite{Holtom_mode_locked_2006}.
 
@@
 The diode-pumped TLS Yb:KGW laser head is placed in the 2.35m-long, asymmetric z-fold cavity shown in \ref{fig:laser-cavity}\subref{fig:laser_schematic}.
-The two focusing gain section mirrors each have a group velocity dispersion (GVD) of $-1300 ( \pm 150 ) \text{fs}^2$ (Layertec GmbH \cite{website_layertec}), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
+The two focusing gain section mirrors each have a group velocity dispersion (GVD) of $-1300 ( \pm 150 ) \text{fs}^2$ (Layertec GmbH~\cite{website_layertec}), a radius of curvature $R$ = 50cm, and are positioned 29.5cm from the gain medium.
 The 5$^\circ$ angle of incidence on the gain section mirrors results in minimal astigmatism, although any net round-trip cavity astigmatism can be compensated for by the TLS technique.
@@
 %page 8636
-focusing the intracavity radiation on a saturable Bragg reflector (SBR) with a 1\% reflectivity modulation depth at 1040nm (BATOP GmbH \cite{website_BATOP}) positioned a distance $z$ from the focusing mirror.
+focusing the intracavity radiation on a saturable Bragg reflector (SBR) with a 1\% reflectivity modulation depth at 1040nm (BATOP GmbH~\cite{website_BATOP}) positioned a distance $z$ from the focusing mirror.
 The shorter 63cm arm is terminated by a plane output coupler with a 7\% transmission at 1040nm.
@@
 The clear implication is therefore that the net thermal lens due to the dominant thermal duct is positive; that is, the sum of the refractive index change due to the thermo-optic coefficient and stress for 1040nm laser radiation polarized along the optical $N_p$-axis (crystal $b$-axis) is positive.
-This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW \cite{Biswal_thermo_optical_05}, but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis \cite{Holtom_mode_locked_2006,Hellstrom_efficient_2006}.
+This result is in apparent contradiction to some recent measurements of relatively large negative thermo-optic coefficients in Yb:KGW~\cite{Biswal_thermo_optical_05}, but is consistent with other determinations of the pump-induced thermal focal length for laser oscillation polarized parallel to the optical $N_p$-axis~\cite{Holtom_mode_locked_2006,Hellstrom_efficient_2006}.
 
@@
 The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (\ref{fig:laser-profile}(a)).
-The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH \cite{website_layertec}) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
+The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH~\cite{website_layertec}) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
 For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (\ref{fig:laser-profile}(b)) at an optimum mode-locked output power of $5.0$W.
-This trend is consistent with the expected `soliton' modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity \cite{Brunner_diode_pumped}.
+This trend is consistent with the expected `soliton' modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity~\cite{Brunner_diode_pumped}.
 
@@
 We also note that in this case the optimum SBR position $z$ is larger than that for higher power 3.5W ML operation with a negative cavity GVD of $-5,200\text{fs}^2$ per round-trip \ref{fig:laser-profile}(a)), but is always less than the 50cm focal length of the concave high reflector.
-This is due to the need to match the absorption-induced thermal bowing of the SBR (i.e., the radius of curvature of its front surface mirror) to the radius of curvature of the intra-cavity laser mode \cite{Schieffer_dual_passive_2006}.
+This is due to the need to match the absorption-induced thermal bowing of the SBR (i.e., the radius of curvature of its front surface mirror) to the radius of curvature of the intra-cavity laser mode~\cite{Schieffer_dual_passive_2006}.
 In general, good ML operation is achieved with the SBR positioned within $\pm$1cm from the optimum $z$ distance.
@@
 This is consistent both with the laser accessing more of the available inhomogeneous gain bandwidth and with our resonator design favoring operation with a strong Kerr lens in the Yb:KGW gain medium.
-Given a value of nonlinear refractive index $n_2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal's optical $N_p$-axis \cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^{-1}$.
+Given a value of nonlinear refractive index $n_2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal's optical $N_p$-axis~\cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^{-1}$.
 From \ref{fig:laser-cavity}\subref{fig:laser_stability}, we see that the resulting increase in the strength of the lensing in the Yb:KGW crystal will result in a reduction of the laser mode size in the gain medium, which produces a better spatial match to gain duct produced by the laser diode pumps (less absorptive loss), thus generating an increase in output power.
@@
 
-For most photoemission sources, the fundamental laser photon energy of 1.19 eV does not provide enough energy to excite electrons above the vacuum level, either for single or two-photon emission \cite{eastman_photoelectric_1970,johnson_optical_1972,dowell_quantum_2009}.
-By appropriate application of crystals with strong non-linear properties, two photons at frequency $\omega$ may be converted into one photon at frequency $2 \omega$ \cite{shen_principles_2002}.
+For most photoemission sources, the fundamental laser photon energy of 1.19 eV does not provide enough energy to excite electrons above the vacuum level, either for single or two-photon emission~\cite{eastman_photoelectric_1970,johnson_optical_1972,dowell_quantum_2009}.
+By appropriate application of crystals with strong non-linear properties, two photons at frequency $\omega$ may be converted into one photon at frequency $2 \omega$~\cite{shen_principles_2002}.
 Since these higher harmonics will be necessary for stimulating photoemission it is important to optimize and characterize the efficiency of this process.
@@
 At the 3.5W ML output power, with a pulse duration of 250fs, the estimated focal irradiance for the fundamental is $\sim50\text{GW}/\text{cm}^2$ in the LBO doubling crystal.
-Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH \cite{website_layertec}), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
+Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH~\cite{website_layertec}), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
 After the collimating
@@
 
-This frequency doubling geometry for LBO is chosen to be close to the optimum for efficient short pulse doubling \cite{wang_efficiency_2003,Saltiel_second_harmonic_2004} defined by limits imposed by group-velocity walk-off between the fundamental and second harmonic pulses, spatial walk-off in the critical phase-matching, and the ratio of the confocal parameter of the focused fundamental wave to the LBO crystal length.
+This frequency doubling geometry for LBO is chosen to be close to the optimum for efficient short pulse doubling~\cite{wang_efficiency_2003,Saltiel_second_harmonic_2004} defined by limits imposed by group-velocity walk-off between the fundamental and second harmonic pulses, spatial walk-off in the critical phase-matching, and the ratio of the confocal parameter of the focused fundamental wave to the LBO crystal length.
 In addition, the $z$-fold configuration allows for astigmatism compensation (e.g., due to the Brewster LBO crystal) of the second harmonic output by adjustment of the angle of incidence on the second collimating mirror.
@@
 
-The second harmonic generation system now employs a 3mm non-critically phase matched LBO crystal \cite{website_newlight}.
+The second harmonic generation system now employs a 3mm non-critically phase matched LBO crystal~\cite{website_newlight}.
 This change has two benefits. 

### 2013-04-11 inc/hardware/magnetic_lenses/maglens.tex
tweak assembly view of mag lens design schematic
citations are now preceeded by non-breaking space (~)

@@
     On the center axis is the inner drift tube, with pole-piece gap separation disk.
-    The end pieces are shown top and bottom.
-    One bridge piece is shown detached (left) and one attached (right).
+    The end pieces are shown top and bottom, bridges are on the left and right.
+    The left side shows a bridge piece and one end before assembly.
     Note that this figure is not drawn to the scale of the lenses at UIC, but rather to match the general schematic in \ref{fig:polepiece-schematic}.
@@
 
-El-Kareh and El-Kareh provide an analysis for symmetric lenses with magnetic shielding and a pole-piece (see Ref.
-\cite{el-kareh_electron_1970} chapter 8, with the most useful information being in section 8.7).
+El-Kareh and El-Kareh provide an analysis for symmetric lenses with magnetic shielding and a pole-piece (see Ref.~\cite{el-kareh_electron_1970} chapter 8, with the most useful information being in section 8.7).
 They define two parameters,
@@
 By their table 8.2a this results in $\beta=0.0146$.
-Then, to create a focal length of 4in ($f/R=16$) table 8.13 in Ref. \cite{el-kareh_electron_1970} gives a value for $k^2 \approx 0.06$.
+Then, to create a focal length of 4in ($f/R=16$) table 8.13 in Ref.~\cite{el-kareh_electron_1970} gives a value for $k^2 \approx 0.06$.
 Finally substituting this value into \ref{eq:elkareh_k2}, with $V=30$kV ($V_r\approx30.88$kV), one can deduce that $NI\approx356$A.

### 2013-04-11 inc/introduction.tex
citations are now preceeded by non-breaking space (~)

@@
-The development of laboratory-scale instrumentation for the direct visualization of fundamental ultrafast (i.e., sub-nanosecond) events with nano-scale (even atomic) spatial resolution is a current active area of research \cite{king_ultrafast_2005}.
+The development of laboratory-scale instrumentation for the direct visualization of fundamental ultrafast (i.e., sub-nanosecond) events with nano-scale (even atomic) spatial resolution is a current active area of research~\cite{king_ultrafast_2005}.
 In many areas of physics, chemistry and materials science a space-time resolution in the sub-1nm$\cdot$ps range is ideally required to allow for the observation of the fastest structural events.
-A promising approach is to employ ultrashort pulses of electrons generated by a laser-driven 10-100keV photoelectron gun to probe the dynamics of the specimen that has been perturbed by a synchronized laser pulse \cite{king_ultrafast_2005}.
-Ultrafast electron diffraction (UED) or femtosecond electron diffraction (FED) \cite{srinivasan_ultrafast_2003,williamson_clocking_1997,cao_femtosecond_2003} is an example of this type of technique and has been used successfully to elucidate atomic-scale dynamics in melting \cite{cao_femtosecond_2003,sciaini_electronic_2009,siwick_atomic-level_2003}, chemical bond breaking \cite{zewail_laser_1988}, etc.
+A promising approach is to employ ultrashort pulses of electrons generated by a laser-driven 10-100keV photoelectron gun to probe the dynamics of the specimen that has been perturbed by a synchronized laser pulse~\cite{king_ultrafast_2005}.
+Ultrafast electron diffraction (UED) or femtosecond electron diffraction (FED)~\cite{srinivasan_ultrafast_2003,williamson_clocking_1997,cao_femtosecond_2003} is an example of this type of technique and has been used successfully to elucidate atomic-scale dynamics in melting~\cite{cao_femtosecond_2003,sciaini_electronic_2009,siwick_atomic-level_2003}, chemical bond breaking~\cite{zewail_laser_1988}, etc.
 Direct time-resolved imaging of similar events can be accomplished using transmission electron microscopy (TEM).
-To date, such dynamic transmission electron microscopes (DTEMs) \cite{bostanjoglo_tracing_1980,domer_high-speed_2003} have been developed for both single-shot imaging with $\sim$$10^{5}$nm$\cdot$ps space-time resolution (e.g., 30ns electron pulses and 5-10nm spatial resolution \cite{lagrange_single-shot_2006,armstrong_practical_2007}) and $\sim$1nm$\cdot$ps space-time resolution ($\sim$100fs with a few nanometer resolution\cite{park_direct_2009}) in the multi-shot data acquisition mode of operation.
+To date, such dynamic transmission electron microscopes (DTEMs)~\cite{bostanjoglo_tracing_1980,domer_high-speed_2003} have been developed for both single-shot imaging with $\sim$$10^{5}$nm$\cdot$ps space-time resolution (e.g., 30ns electron pulses and 5-10nm spatial resolution~\cite{lagrange_single-shot_2006,armstrong_practical_2007}) and $\sim$1nm$\cdot$ps space-time resolution ($\sim$100fs with a few nanometer resolution~\cite{park_direct_2009}) in the multi-shot data acquisition mode of operation.
 Both of these instruments employ retro-fitted standard electron microscope columns and are limited in performance mainly by space-charge effects.
 
-Improvements in the performance of DTEMs, especially ultrafast electron microscopes (UEMs), and UED systems will likely require the implementation of techniques that offset or compensate for deleterious space-charge effects \cite{armstrong_prospects_2007,lobastov_four-dimensional_2005}.
+Improvements in the performance of DTEMs, especially ultrafast electron microscopes (UEMs), and UED systems will likely require the implementation of techniques that offset or compensate for deleterious space-charge effects~\cite{armstrong_prospects_2007,lobastov_four-dimensional_2005}.
 Such performance optimization efforts will require efficient models of electron pulse propagation dynamics that include the ability to simulate the action of electron optical elements (including the photoelectron gun itself) on the pulse.
-Indeed $2^{\text{nd}}$ generation UED instruments are now beginning to employ both magnetic lenses and RF cavities \cite{oudheusden_electron_2007,veisz_hybrid_2007} to optimize the electron pulse delivery to the specimen; that is, compensate for global space-charge effects in both the transverse and longitudinal pulse dimensions, respectively.
+Indeed $2^{\text{nd}}$ generation UED instruments are now beginning to employ both magnetic lenses and RF cavities~\cite{oudheusden_electron_2007,veisz_hybrid_2007} to optimize the electron pulse delivery to the specimen; that is, compensate for global space-charge effects in both the transverse and longitudinal pulse dimensions, respectively.
 Similarly, future UEM column designs are expected to use RF pulse compression cavities while circumventing high charge density beam cross-overs (i.e., beam foci).
 
-In this thesis, I will present an extension of the work of Michalik and Sipe \cite{michalik_analytic_2006,michalik_erratum:_2008}, describing electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment, to include the influence of linear external forces due to electron optical elements.
+In this thesis, I will present an extension of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment, to include the influence of linear external forces due to electron optical elements.
 The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
@@
 
-Since the extension to the AG model is valid only within the limits of the analytical method itself, in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
-Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes \cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid \cite{luiten_how_2004}.
+Since the extension to the AG model is valid only within the limits of the analytical method itself, in particular, its mean internal space-charge field and self-similar Gaussian approximations~\cite{michalik_analytic_2006}, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
+Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
 This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
-Consequently, the AG approach is applicable to both the single electron per pulse limit \cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate \cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
+Consequently, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
 It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
@@
 In Chapter \ref{chap:model_results}, I apply the AG model to draw some initial insight into the behaviors of ultrafast electron pulses in a UEM column.
-Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion \cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance \cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.
+Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion~\cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance~\cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.
 First, that the electron pulses should be emitted from a large-area source, and second, that every effort must be made to reduce the initial rms transverse momentum (i.e., the emission cone angle).

### 2013-04-11 inc/model_results/compound/compound.tex
killing TODOs
citations are now preceeded by non-breaking space (~)

--- 2013-04-06 inc/model_results/compound/compound.tex
@@
 The design of such a compound system is, of course, a cornerstone of modern electron microscope development, although space-charge effects are generally not included, since, on average, there is usually no more than one electron in the microscope column at a time.
-However, for a nanosecond DTEM \cite{lagrange_nanosecond_2008,reed_evolution_2009}, the microscope column must be designed to compensate for transverse ($\smallT$) space-charge effects, which dominate the propagation dynamics for the prolate electron pulses employed in such instruments (see \ref{fig:compare_shape}).
-For the ultrashort (i.e., picosecond and femtosecond) electron pulse durations used in UEMs\cite{park_direct_2009} and UED experiments \cite{oudheusden_electron_2007}, the longitudinal ($z$) pulse dynamics due to both dispersion and space-charge effects must also be considered (as indicated in \ref{fig:compare_shape}).
+However, for a nanosecond DTEM~\cite{lagrange_nanosecond_2008,reed_evolution_2009}, the microscope column must be designed to compensate for transverse ($\smallT$) space-charge effects, which dominate the propagation dynamics for the prolate electron pulses employed in such instruments (see \ref{fig:compare_shape}).
+For the ultrashort (i.e., picosecond and femtosecond) electron pulse durations used in UEMs~\cite{park_direct_2009} and UED experiments~\cite{oudheusden_electron_2007}, the longitudinal ($z$) pulse dynamics due to both dispersion and space-charge effects must also be considered (as indicated in \ref{fig:compare_shape}).
 
@@
 
-\ref{fig:compound} shows an example of a compound element system for the delivery of an ultrashort electron pulse to the specimen area; in this case, optimized for $ N = 10^{ 6 } $ --- the number of electrons per pulse required for single shot diffraction measurements \cite{armstrong_practical_2007}.
+\ref{fig:compound} shows an example of a compound element system for the delivery of an ultrashort electron pulse to the specimen area; in this case, optimized for $ N = 10^{ 6 } $ --- the number of electrons per pulse required for single shot diffraction measurements~\cite{armstrong_practical_2007}.
 The displayed dynamics are for an electron pulse generated by a $\tau = $ 100fs $w = $ 0.1mm laser, driving a photocathode, with $\Delta E = $ 0.5eV, then accelerated by 20 kV to a velocity of $8.4 \times 10^{7} \text{m}/\text{s}$.
@@
 The dependencies of the the spatial variances, $\sigma_{\alpha}$, shown in \ref{fig:compound} as the electron pulse propagates through the multi-element system, clearly illustrate the effect of the magnetic lenses and RF cavity in the transverse ($\smallT$) and longitudinal ($z$) dimensions.
-As the image-to-object distance ratio is less than unity in this case, both the transverse and longitudinal imaged pulse dimensions are reduced relative to their initial values. %TODO comment futher
+As the image-to-object distance ratio is less than unity in this case, both the transverse and longitudinal imaged pulse dimensions are reduced relative to their initial values.
 
@@
 At $ N = 10^{6} $, space-charge effects moderate the image demagnification due to the $\sim$2 object-to-image distance ratio in \ref{fig:compound}. 
-In other words, for the displayed compound element system, which could be employed in UED experiments \cite{oudheusden_electron_2007}, the initial space-time resolution associated with the $\sim$0.1mm spot size and $\sim$100fs pulse duration can be maintained (or bettered) for $N \lesssim 10^{6} $.
+In other words, for the displayed compound element system, which could be employed in UED experiments~\cite{oudheusden_electron_2007}, the initial space-time resolution associated with the $\sim$0.1mm spot size and $\sim$100fs pulse duration can be maintained (or bettered) for $N \lesssim 10^{6} $.
 For a UEM, on the other hand, it may be necessary to separate longitudinally the space and time images produced by the compound element column.

### 2013-04-11 inc/model_results/lens_charge/lens_charge.tex
citations are now preceeded by non-breaking space (~)

@@
 
-In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses \cite{oudheusden_electron_2007,lagrange_nanosecond_2008}.
-In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry \cite{berger_dc_2009}.
+In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses~\cite{oudheusden_electron_2007,lagrange_nanosecond_2008}.
+In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry~\cite{berger_dc_2009}.
 The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity), while still subject to Liouville's Theorem (see Section \ref{sec:liouville}).

### 2013-04-11 inc/model_results/rf_cav/rf_cav.tex
killing TODOs
citations are now preceeded by non-breaking space (~)

@@
 
-In comparison to using magnetic lenses to compensate for transverse pulse broadening by using magnetic lenses, RF cavities are being employed in UED systems \cite{oudheusden_electron_2007}, and are expected to be employed in UEM systems as well, to compensate for longitudinal broadening in the ultrashort pulses.
+In comparison to using magnetic lenses to compensate for transverse pulse broadening by using magnetic lenses, RF cavities are being employed in UED systems~\cite{oudheusden_electron_2007}, and are expected to be employed in UEM systems as well, to compensate for longitudinal broadening in the ultrashort pulses.
 \ref{fig:rf_cav_num} illustrates the operation of a $ \Omega / 2 \pi = $ 3GHz $\text{TM}_{010}$ RF cavity of axial length $ d = \pi v_{\smallzero} / \Omega $ as an electron pulse compressor tuned to produce the shortest pulse at a distance $ z^{\prime} = $ 10cm behind the cavity for $N=1$ and $N=10^5$ pulse charges.
@@
 
-%TODO why is this even worse than before?
 The RF electric field amplitude $E_{0}$ of about 450kV/m required to achieve the temporal focusing 10cm behind the RF cavity in the dispersive limit is also consistent with expectations.
@@
 However, unlike in the dispersion limited case (\ref{fig:rf_cav_num}\subref{fig:rf_cav_N1}), the compressed pulse duration is now a factor of $\sim$2 less ($\sim 48$ fs) than the original 100fs laser pulse duration.
-This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression \cite{strickland_compression_1985}.
+This additional pulse compression for the employed 1-to-1 image-to-object distance (magnification) ratio is made possible by the additional pulse bandwidth (and momentum chirp) generated by the space-charge effects --- in analogy to self-phase modulation (and dispersion) in fiber-grating laser pulse compression~\cite{strickland_compression_1985}.
 In this case, the $\sim$20-fold increase in the transverse pulse size (corresponding to a $\sim400\times$ reduced peak pulse charge density) aids the temporal compression by greatly diminishing Coulomb effects that counteract the reduction in electron pulse duration.

### 2013-04-11 inc/model_results/shapes/shapes.tex
killing TODOs
citations are now preceeded by non-breaking space (~)

--- 2013-04-06 inc/model_results/shapes/shapes.tex
@@
 \ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \xi(0) = 1 $), oblate or `disk-like' ($ \xi(0) < 1 $), and prolate or `cigar-like' ($ \xi(0) > 1 $) pulses.
-%TODO rewrite using w?
 This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
@@
 For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects may dominate the propagation dynamics.
-\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM \cite{berger_dc_2009,armstrong_practical_2007}.
+\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM~\cite{berger_dc_2009,armstrong_practical_2007}.
 The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
@@
 %In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
-The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.
+The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width~\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.
 
-The data presented in \ref{fig:compare_shape} using the AG model of Michalik and Sipe \cite{michalik_analytic_2006} are entirely consistent with prior works \cite{reed_femtosecond_2006,siwick_ultrafast_2002}.
-In femtosecond UED experiments, significant efforts are made to offset space-charge induced temporal (longitudinal, $z$) pulse broadening by reducing the time of flight from the gun photocathode to the sample \cite{siwick_ultrafast_2002,reed_evolution_2009}.
-Moreover, efforts are now underway to compensate for temporal electron pulse broadening using RF cavity pulse compression techniques \cite{oudheusden_electron_2007}.
-In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers \cite{lobastov_four-dimensional_2005}, deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
-Even for nanosecond DTEM (the extreme prolate pulse case, $\xi \gg 1 $), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $ \cite{lagrange_nanosecond_2008}.
-In this thesis, however, I will often restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be easily generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening \cite{veisz_hybrid_2007}.
+The data presented in \ref{fig:compare_shape} using the AG model of Michalik and Sipe~\cite{michalik_analytic_2006} are entirely consistent with prior works~\cite{reed_femtosecond_2006,siwick_ultrafast_2002}.
+In femtosecond UED experiments, significant efforts are made to offset space-charge induced temporal (longitudinal, $z$) pulse broadening by reducing the time of flight from the gun photocathode to the sample~\cite{siwick_ultrafast_2002,reed_evolution_2009}.
+Moreover, efforts are now underway to compensate for temporal electron pulse broadening using RF cavity pulse compression techniques~\cite{oudheusden_electron_2007}.
+In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers~\cite{lobastov_four-dimensional_2005}, deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
+Even for nanosecond DTEM (the extreme prolate pulse case, $\xi \gg 1 $), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $~\cite{lagrange_nanosecond_2008}.
+In this thesis, however, I will often restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry~\cite{king_ultrafast_2005} and (ii) can be easily generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening~\cite{veisz_hybrid_2007}.

### 2013-04-11 inc/ms_model/ms_model.tex
began working on a conclusion section
citations are now preceeded by non-breaking space (~)

@@
 
-This section introduces the Analytic Gaussian (AG) model of Michalik and Sipe \cite{michalik_analytic_2006}.
+This section introduces the Analytic Gaussian (AG) model of Michalik and Sipe~\cite{michalik_analytic_2006}.
 In the AG model the dynamics of a Gaussian electon pulse are computed using a mean-field self-similar approximation.
 The result is a set of equations that can be solved very quickly using numerical methods on a modern computer.
-When checked against N-body simulations, the authors find that the model is also surprisingly accurate, even for a large set of non-Gaussian pulses \cite{michalik_analytic_2006,michalik_evolution_2009}.
+When checked against N-body simulations, the authors find that the model is also surprisingly accurate, even for a large set of non-Gaussian pulses~\cite{michalik_analytic_2006,michalik_evolution_2009}.
 
@@
 The $K^{force}$ integrals, related to the internal force of the pulse's Coulombic repulsion, are far more complicated to evaluate.
-For more information about the derivation, see \cite{michalik_analytic_2006}. %TODO link to an appendix with Bessel etc for Kforce
+For more information about the derivation, see~\cite{michalik_analytic_2006}. %TODO link to an appendix with Bessel etc for Kforce
 The final matrices are
@@
 \end{equation}
-[Note that the original paper of Michalik and Sipe (Reference \cite{michalik_analytic_2006}) contained an error which incorrectly defined $\xi$; this was corrected in an erratum (Reference \cite{michalik_erratum:_2008}).]
+[Note that the original paper of Michalik and Sipe (Reference~\cite{michalik_analytic_2006}) contained an error which incorrectly defined $\xi$; this was corrected in an erratum (Reference~\cite{michalik_erratum:_2008}).]
 The family of $L$ functions are
-\begin{subequations}
+\begin{subequations} \label{eq:ag_l_family}
   \begin{gather}
@@
 
-This AG formalism of Michalik and Sipe \cite{michalik_analytic_2006} employs a mean-field approximation in the evolution of the dynamical effect of the electron pulse's internal space-charge field.
+This AG formalism of Michalik and Sipe~\cite{michalik_analytic_2006} employs a mean-field approximation in the evolution of the dynamical effect of the electron pulse's internal space-charge field.
 This is of course not strictly accurate for a Gaussian charge distribution and the error leads to a distortion of the pulse shape.
 The ``self-similar'' nature of the model, the fact that the overall shape of the pulse cannot change, prevents inclusion of these distortions.
-Even so, the AG model of charge bunch dynamics has been benchmarked against particle tracking simulations for a number of electron pulse shapes \cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid which explicitly features a linear internal space-charge field \cite{luiten_how_2004}.
+Even so, the AG model of charge bunch dynamics has been benchmarked against particle tracking simulations for a number of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid which explicitly features a linear internal space-charge field~\cite{luiten_how_2004}.

### 2013-04-11 inc/photocathode/este/este.tex
citations are now preceeded by non-breaking space (~)

@@
 
-Recent experiments in our group have come to change our understanding of several photoemission processes, once thought to be well understood \cite{berger_excited_2012}.
-Indeed, while this knowledge has been commonplace in other fields, such as angle-resolved photoemission spectroscopy (ARPES) \cite{himpsel_angle-resolved_1983}, to our knowledge the effect has not been considered for electron beam generation.
-In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_{\smallT}$ developed in Refs.
-\cite{dowell_quantum_2009} and \cite{jensen_emittance_2010} for thermionic electron emission should be rewritten to read $\sqrt{m^* k_B T_e}$, where $T_e$ is the temperature of the emitting electron distribution and $k_B$ is Boltzmann's constant.
+Recent experiments in our group have come to change our understanding of several photoemission processes, once thought to be well understood~\cite{berger_excited_2012}.
+Indeed, while this knowledge has been commonplace in other fields, such as angle-resolved photoemission spectroscopy (ARPES)~\cite{himpsel_angle-resolved_1983}, to our knowledge the effect has not been considered for electron beam generation.
+In this section, we show that, at least for thermionic emission, the effective mass $m^*$ of the electronic state from which the electron was emitted affects the rms transverse momentum; that is, the expression for $\Delta p_{\smallT}$ developed in Refs.~\cite{dowell_quantum_2009} and~\cite{jensen_emittance_2010} for thermionic electron emission should be rewritten to read $\sqrt{m^* k_B T_e}$, where $T_e$ is the temperature of the emitting electron distribution and $k_B$ is Boltzmann's constant.
 Direct observation of the transverse momentum distribution for excited-state thermionic emission (ESTE) from GaSb and InSb, two similar III-V zincblende semiconductors, demonstrates the $m^*$ dependence.
@@
 Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
-Prior studies \cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included \cite{dowell_quantum_2009}.
+Prior studies~\cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included~\cite{dowell_quantum_2009}.
 Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.
@@
 
-The known optical properties of the two semiconductors \cite{aspnes_dielectric_1983} indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb \cite{chelikowsky_nonlocal_1976} and InSb \cite{chelikowsky_erratum_1984}, is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
+The known optical properties of the two semiconductors~\cite{aspnes_dielectric_1983} indicate a strong absorption at 261nm, with an optical absorption depth of only 7-8nm, which, based on the band structures of GaSb~\cite{chelikowsky_nonlocal_1976} and InSb~\cite{chelikowsky_erratum_1984}, is primarily due to promotion of electrons near the $\Gamma$ point from the valence band (heavy-hole, light-hole, and split-off bands) directly into the upper $\Gamma_8$ conduction band located at 3.77eV and 3.59eV above the valence band maximum in GaSb and InSb, respectively.
 Assuming parabolic bands and an estimated effective mass $m^*$ of about 0.3$m$ (0.5$m$) for the $\Gamma_8$ conduction band in GaSb (InSb), we determine that these electrons are excited with an average excess energy (above the $\Gamma_8$ conduction band minimum) of $\sim$0.35eV ($\sim$0.41eV); that is, an initial electron temperature $T_e$ of $\sim$4,200K ($\sim$4,900K).
 For the $\sim$0.1nJ incident UV pulse energies, $\sim40\mu$m laser spot size, and $p$-polarized reflectance of $\sim$40\% for both semiconductors at our 60$^{\circ}$ incidence angle, the photo-injected electron density is $\sim10^{18}cm^{-3}$.
-At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization \cite{portella_k-space_1992} and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\Phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\Phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
+At these carrier densities, which are non-degenerate for the upper $\Gamma_8$ conduction bands, we expect rapid thermalization~\cite{portella_k-space_1992} and, consequently, an electron distribution with a well-populated Boltzmann tail extending above the vacuum level located at an effective work function $\Phi$ of 0.99eV (1.18eV) above the $\Gamma_8$ conduction band minimum in GaSb (InSb); thus allowing for thermionic emission from this excited state as initially $\exp[\Phi/(k_B T_e)] \approx 0.06$ in both semiconductors.
 
 Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above the effective $\sim$1eV work function, thereby curtailing the observed ESTE.
-By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band \cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
+By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band~\cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
 This means that even without population decay mechanisms the temperature of the exited electron distribution $T_e$ drops at a rate of $\sim$1,600K/ps in both semiconductors.
@@
 \ref{fig:este-semicond} also shows that the measured Fourier plane electron beam spot size for GaSb is independent of the UV laser pulse energy, as would be expected for the proposed emission process.
-More significantly, using $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m$ in the semiconductor), simulation of the experiment with the extended AG model \cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.90($\pm$0.06)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\Phi/(k_B T_e)] \sim 10^{-15}$).
+More significantly, using $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ (i.e., assuming the emitted electrons have a mass $m$ in the semiconductor), simulation of the experiment with the extended AG model~\cite{berger_semi-analytic_2010} indicates that the observed GaSb Fourier plane spot size of 0.90($\pm$0.06)mm (HWe$^{-1}$M) would be due to an electron temperature $T_e \approx$ 360K which is associated with negligible thermionic emission ($\exp[-\Phi/(k_B T_e)] \sim 10^{-15}$).
 On the other hand, employing a mass of 0.3$m$ for the electrons in GaSb allows an average electron temperature (over the $\sim$4ps laser pulse duration) of about 1,200K to be extracted using the AG model simulation --- a value much more consistent with the proposed ESTE mechanism and the expected cooling rate by LO phonon emission.
 In fact, at the maximum $\sim$0.2nJ incident UV laser pulse energy up to $10^8$ electrons/pulse could be excited into the upper $\Gamma_8$ conduction band, of which one in $\exp[\Phi/(k_B T_e)] \approx$ 20,000 are above the effective 0.99eV work function, thus agreeing with the observed yield of $\sim$10$^3$ electrons/pulse (\ref{fig:este-semicond}).
-We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\Phi$ = 4.31eV and assuming $\Delta p_{\smallT} = \sqrt{m ( \hbar \omega - \Phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy \cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.
+We also note that the normalized transverse rms emittance $\varepsilon_{\smallT}$ of this UV laser-driven ultrafast GaSb electron source is more than a factor of two less than that expected from a standard Cu photocathode ($\Phi$ = 4.31eV and assuming $\Delta p_{\smallT} = \sqrt{m ( \hbar \omega - \Phi ) / 3 }$, where $\hbar \omega$ is the incident photon energy~\cite{dowell_quantum_2009,jensen_emittance_2010}) irradiated at the same 261nm wavelength.
 
@@
 The increased rms transverse momentum for the same ESTE process is consistent with the larger $\sim$0.5$m$ effective mass of the upper $\Gamma_8$ conduction band in InSb, the expected higher initial temperature of the electron distribution, and the marginally slower cooling process associated with the lower 24meV LO phonon energy.
-The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation \cite{aspnes_dielectric_1983}.
+The resultant higher average electron temperature $T_e$ also contributes to the increased electron yield that is likely further enhanced by a larger absorption into the upper $\Gamma_8$ conduction band (due to its larger effective mass) and a 4\% lower $p$-polarized reflectance for the 60$^{\circ}$ incident UV laser radiation~\cite{aspnes_dielectric_1983}.
 
@@
 As this momentum component is conserved in emission from the material surface into the vacuum, the rms transverse momentum $\Delta p_{\smallT}$ of the electron source clearly must scale with $\sqrt{m^*}$ as observed in the experiment.
-This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies \cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
-We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes \cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m$)  to allow for direct emission of electrons excited into the conduction band.
+This is, of course, the \textit{sine qua non} of angle-resolved photoemission studies~\cite{himpsel_angle-resolved_1983} where determination of the electron emission angle and energy allows the effective mass parallel to the sample surface to be determined.
+We note that the electron effective mass also causes the narrow emission cone reported for $p$-type GaAs(100) negative electron affinity photocathodes~\cite{liu_narrow_2005}, where surface cesiation lowers the vacuum level below the $\Gamma_6$-valley minimum ($m^* = 0.067m$)  to allow for direct emission of electrons excited into the conduction band.

### 2013-04-11 inc/photocathode/future.tex
citations are now preceeded by non-breaking space (~)

@@
 The philosophical argument presented in Section \ref{sec:este} for using the effective mass $m^*$ in the expressions for the transverse momentum variance $\Delta p_{\smallT}$ simply follows from conservation of energy and momentum at the barrier and actually does not depend on the ``excited state'' nature nor the thermionic emission process.
-For the example of two-photon assisted thermionic emission (2PTE) from gold presented in Section \ref{sec:two_photon_thermionic}, the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to its rest mass $m$ \cite{johnson_optical_1972} --- the value employed in that analysis.
+For the example of two-photon assisted thermionic emission (2PTE) from gold presented in Section \ref{sec:two_photon_thermionic}, the close agreement between the data and the simulation is supported by the fact that the effective mass of a free electron in Au is equal to its rest mass $m$~\cite{johnson_optical_1972} --- the value employed in that analysis.
 
-For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum (\ref{eq:jensen_eta}, Refs.
-\cite{dowell_quantum_2009} and \cite{jensen_emittance_2010}) to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
+For direct single-photon photoemission, we would therefore also expect to require that the expression describing the rms transverse momentum (\ref{eq:jensen_eta}, Refs.~\cite{dowell_quantum_2009} and~\cite{jensen_emittance_2010}) to be written as $\Delta p_{\smallT} = \sqrt{m^* ( \hbar \omega - \Phi ) / 3 }$.
 Efforts are currently underway to determine if this is indeed the case using a variety of metal photocathodes and the $\sim$4ps, 261nm laser radiation source.

### 2013-04-11 inc/photocathode/plasmon/plasmon.tex
citations are now preceeded by non-breaking space (~)

@@
 Electrons in a conductor can couple with electromagnetic fields, in this case a laser field, and oscillate accordingly.
-This overall oscillation of electrons near the surface of the conductor creates strong oscillatory local surface electric field \cite{cottam_introduction_2004,concepts_2002}.
+This overall oscillation of electrons near the surface of the conductor creates strong oscillatory local surface electric field~\cite{cottam_introduction_2004,concepts_2002}.
 In PAPE, the goal is use this field to both enhance the emission and `control' the divergence of photoemitted electrons through the combined action of photoemission barrier suppression and ponderomotive acceleration.
@@
 
-This avenue of investigation was motivated by Zawadzka et al \cite{zawadzka_evanescent_2001} who indicate that when driving a surface plasmon on gold they witnessed a reduction in the emission angle of photoemitted electrons.
+This avenue of investigation was motivated by Zawadzka et al~\cite{zawadzka_evanescent_2001} who indicate that when driving a surface plasmon on gold they witnessed a reduction in the emission angle of photoemitted electrons.
 In their paper, the authors coat a thin gold film on a prism; the laser then ``back illuminates'' the gold through the prism, ejecting electrons on vacuum side of the foil.
 This arrangement is called the Kretschmann geometry.
-While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, we felt that for UEM applications we should explore the alternative ``grating coupling'' geometry \cite{kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
+While this geometry is common, due to fears about the lack of heat conduction through the glass and the power needed to generate a sufficient electron beam current, we felt that for UEM applications we should explore the alternative ``grating coupling'' geometry~\cite{kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
 In this geometry, the plasmon is driven on a periodic surface rather than a flat surface attached to glass.
@@
 
-At the interface of a vacuum and such a simple material, contstraints on the material parameters which may allow surface plasmons, result in the condition that the dielectric constant at the oscillation frequency of the surface plasmon $ \varepsilon(\omega_{SP}) = -1 $ \cite{cottam_introduction_2004}.
+At the interface of a vacuum and such a simple material, contstraints on the material parameters which may allow surface plasmons, result in the condition that the dielectric constant at the oscillation frequency of the surface plasmon $ \varepsilon(\omega_{SP}) = -1 $~\cite{cottam_introduction_2004}.
 Substituting this result into \ref{eq:dielectric_as_omega} shows that this frequency of oscillation is simply related to intrinsic plasma frequency by
@@
 Several regions were created each 100x100$\mu$m in area, having a $\sim$1$\mu$m period and $\sim$10\% modulation depth.
-By \ref{eq:plasmon_condition}, given a laser wavelength of 523nm ($\varepsilon_{Au} \approx -3.95$ \cite{johnson_optical_1972}), this choice of periodicity corresponds to an incidence angle of $\sim39^{\circ}$, which is within the acceptance range of our experimental setup (35-75$^{\circ}$) (see Section \ref{sec:gun_design}).
+By \ref{eq:plasmon_condition}, given a laser wavelength of 523nm ($\varepsilon_{Au} \approx -3.95$~\cite{johnson_optical_1972}), this choice of periodicity corresponds to an incidence angle of $\sim39^{\circ}$, which is within the acceptance range of our experimental setup (35-75$^{\circ}$) (see Section \ref{sec:gun_design}).
 An SEM image of one such region is shown in \ref{fig:fib-si-sem}.
@@
 Moreover, this enhanced laser-driven electron emission was only observed for $p$-polarized incident 523nm femtosecond laser radiation.
-These two properties are known characteristics of plasmon-assisted photoemission \cite{kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
+These two properties are known characteristics of plasmon-assisted photoemission~\cite{kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
 
@@
 As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
-For a mathematical treatment of the transverse electric field of such patterned surfaces, the reader is referred to Ref. \cite{watts_sharp_1997}. 
-Our results are commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
+For a mathematical treatment of the transverse electric field of such patterned surfaces, the reader is referred to Ref.~\cite{watts_sharp_1997}. 
+Our results are commensurate with the results shown by Zawadzka et al~\cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
 As that geometry does not employ a patterned surface, it would not be subject to such local field enhancements of the accelerating electric field.
@@
 Other nano-engineered surfaces are still being considered for future work.
-Various other more exotic nano-structures are under examination for PAPE \cite{li_surface_2013,polyakov_plasmon_2013}.
-Bi-metallic layers are known to enhance PAPE, which may allow for reduced incident laser power for driving the plasmon \cite{kupersztych_anomalous_2005}.
+Various other more exotic nano-structures are under examination for PAPE~\cite{li_surface_2013,polyakov_plasmon_2013}.
+Bi-metallic layers are known to enhance PAPE, which may allow for reduced incident laser power for driving the plasmon~\cite{kupersztych_anomalous_2005}.
 Photoemission from nano-arrays of quantum-wells also offers the tantalizing hope of reduced rms transverse momentum of the generated electrons.

### 2013-04-11 inc/photocathode/setup/setup.tex
citations are now preceeded by non-breaking space (~)

@@
 
-%Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED) \cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers \cite{dowell_cathode_2010}.
-The spatial quality of the sources for UEM and UED are often characterized in terms of their emitted electron beam's normalized transverse rms emittance \cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_{\smallT} \Delta x / (m c)$, where $\Delta p_{\smallT} = \sqrt{ \langle p_{\smallT}^2 \rangle }$ is the rms transverse momentum of the beam, $\Delta x = \sqrt{ \langle x^2 \rangle }$ is its size (cylindrical symmetry assumed), and $m$ is the electron rest mass and $c$ is the speed of light in vacuum.
-This quantity is conserved under free-space propagation and perfect lensing, but may increase due to lens aberrations \cite{oshea_reversible_1998}.
+%Sub-nanosecond (ultrafast) pulsed laser-driven electron sources are now being employed in a wide variety of scientific studies: in particular, ultrafast electron diffraction (UED)~\cite{sciaini_femtosecond_2011}, ultrafast transmission electron microscopy (UEM)~\cite{king_ultrafast_2005}, and pulsed X-ray free electron lasers~\cite{dowell_cathode_2010}.
+The spatial quality of the sources for UEM and UED are often characterized in terms of their emitted electron beam's normalized transverse rms emittance~\cite{dowell_quantum_2009,jensen_emittance_2010}; $\varepsilon_{\smallT} = \Delta p_{\smallT} \Delta x / (m c)$, where $\Delta p_{\smallT} = \sqrt{ \langle p_{\smallT}^2 \rangle }$ is the rms transverse momentum of the beam, $\Delta x = \sqrt{ \langle x^2 \rangle }$ is its size (cylindrical symmetry assumed), and $m$ is the electron rest mass and $c$ is the speed of light in vacuum.
+This quantity is conserved under free-space propagation and perfect lensing, but may increase due to lens aberrations~\cite{oshea_reversible_1998}.
 As I have discussed in Chapter \ref{chap:considerations}, an electron beam (or in the case of UEM, a pulse) with a lower transverse emittance will result in higher quality images, however since this quantity is conserved, the only chance to minimize this quantity occurs upon emission.
-As a reduction in the source size $\Delta x_0$ is limited either by the attainable diffraction-limited incident laser pulse focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field) \cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness \cite{berger_dc_2009}) may only be possible by decreasing the initial rms transverse momentum $\Delta p_{\smallT}$ of the electron source --- an intrinsic property of the emission source \cite{dowell_quantum_2009,jensen_emittance_2010}.
+As a reduction in the source size $\Delta x_0$ is limited either by the attainable diffraction-limited incident laser pulse focusing or by the short-pulse Child's Law (screening of the electron gun's acceleration field)~\cite{valfells_effects_2002}, significant improvements in the performance of scientific instruments employing pulsed electron sources through a reduction in their spatial emittance $\varepsilon_{\smallT}$ (and consequent increase in brightness~\cite{berger_dc_2009}) may only be possible by decreasing the initial rms transverse momentum $\Delta p_{\smallT}$ of the electron source --- an intrinsic property of the emission source~\cite{dowell_quantum_2009,jensen_emittance_2010}.
 
@@
 Though many of the experimental components have been described in detail in Chapter \ref{chap:prototype}, for convenience I will briefly summarize them again here.
-The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator \cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate.
+The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator~\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate.
 This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV).
 Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
-Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface \cite{berger_dc_2009}.
-The electron gun design is based on the that of Togawa \textit{et al.} \cite{berger_dc_2009,togawa_ceb6_2007} and has an acceleration gap of 25($\pm$2)mm.
+Both the green and UV laser pulses are focused with 30cm focal lenses onto the photocathode in a 20kV electron gun at an angle of incidence of $60(\pm5)^{\circ}$, which improves the coupling of the $p$-polarized laser radiation into the photocathode surface~\cite{berger_dc_2009}.
+The electron gun design is based on the that of Togawa \textit{et al.}~\cite{berger_dc_2009,togawa_ceb6_2007} and has an acceleration gap of 25($\pm$2)mm.
 After acceleration in the DC photo-gun, the electrons pass through a set of electrostatic deflection plates to ensure that the electron beam passes, as close as possible, axially through the center of two, 6.35mm-bore, thin and round magnetic lenses with counter-propagating currents to minimize image rotation effects.

### 2013-04-11 inc/photocathode/single_photon/single_photon.tex
citations are now preceeded by non-breaking space (~)

@@
 Single-photon photoemission is the most common photoemission process and the easiest to explain.
-As described by Einstein in his famous paper on ``the photoelectric effect'' \cite{einstein_uber_1905}, as it is colloquially known, if an electron absorbs a photon whose energy is sufficient to promote the electron above the local vacuum level, it becomes free to leave the host material.
+As described by Einstein in his famous paper on ``the photoelectric effect''~\cite{einstein_uber_1905}, as it is colloquially known, if an electron absorbs a photon whose energy is sufficient to promote the electron above the local vacuum level, it becomes free to leave the host material.
 The material dependent minimum energy needed to eject an electron is termed its ``work function'' $\Phi$.
@@
 
-An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.3-4.5 eV \cite{yen_thermally_1980}, is used as a photocathode.
+An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.3-4.5 eV~\cite{yen_thermally_1980}, is used as a photocathode.
 When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E \approx $ 0.35 eV.
@@
 In this experiment, the yield is determined by integrating the total counts recorded on the CCD detector and therefore is only a relative measurement.
-The fact that the Fourier plane spot-size (squares), directly related the initial transverse momentum variance $\Delta p_{\smallT}$, is unchanged over the full range of laser power is also expected, since by $\Delta p_{\smallT} = \sqrt{ m \Delta E / 3 } $ \cite{dowell_quantum_2009}, $\Delta p_{\smallT}$ only depends on $\Delta E$.
+The fact that the Fourier plane spot-size (squares), directly related the initial transverse momentum variance $\Delta p_{\smallT}$, is unchanged over the full range of laser power is also expected, since by $\Delta p_{\smallT} = \sqrt{ m \Delta E / 3 } $~\cite{dowell_quantum_2009}, $\Delta p_{\smallT}$ only depends on $\Delta E$.
 
 In order to improve performance, one must be able to minimize $\Delta p_{\smallT}$ without negatively affecting the photoemission efficiency, and thus electron yield.
-For single-photon photoemission, although the polarization-dependent reflectivity can be tweaked to improve performance slightly \cite{berger_dc_2009}, due to the quadratic dependence of the efficiency on excess photoemission energy \cite{shalaev_electron_1994}, any reduction in $\Delta p_{\smallT}$ gained by reducing $\Delta E$ will be mitigated by a large drop in electron yield.
+For single-photon photoemission, although the polarization-dependent reflectivity can be tweaked to improve performance slightly~\cite{berger_dc_2009}, due to the quadratic dependence of the efficiency on excess photoemission energy~\cite{shalaev_electron_1994}, any reduction in $\Delta p_{\smallT}$ gained by reducing $\Delta E$ will be mitigated by a large drop in electron yield.
 For this reason, as it stands currently and as we have already seen in Section \ref{sec:must_reduce_transverse_momentum}, flat metal photocathodes employing single-photon photoemission are not likely to be useful for UEM.

### 2013-04-11 inc/photocathode/two_photon_thermionic/two_photon_thermionic.tex
citations are now preceeded by non-breaking space (~)

@@
 
-Two-photon assisted thermionic emission (2PTE) occurs when two photons provide nearly enough energy to emit a cold electron, and therefore a hotter electron, in the tail of the Fermi distribution, may have enough energy to be emitted \cite{yen_thermally_1980}.
+Two-photon assisted thermionic emission (2PTE) occurs when two photons provide nearly enough energy to emit a cold electron, and therefore a hotter electron, in the tail of the Fermi distribution, may have enough energy to be emitted~\cite{yen_thermally_1980}.
 Of course this is even less efficient than direct two-photon photoemission, but as three-photon photoemission goes as $I^3$, 2PTE will still be the dominant photoemission mechanism.
@@
 \end{equation}
-where $k_B$ is the Boltzmann constant \cite{dowell_quantum_2009,jensen_emittance_2010}.
+where $k_B$ is the Boltzmann constant~\cite{dowell_quantum_2009,jensen_emittance_2010}.
 Since a high laser intensity is required to drive 2PTE, the laser itself will heat the electrons, therefore, if the laser pulse duration is less than the thermalization time of the electron distribution, %TODO check this!
@@
 The quadratic dependence of the electron yield on pulse energy clearly indicates a two-photon emission process.
-This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\Phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water \cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface \cite{eastman_photoelectric_1970}.
-Moreover, when $2\hbar \omega \approx \Phi$, the dependence of the yield on the square of the electron's excess energy above the work function \cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
-Consequently, the observed increase in $\Delta p_{\smallT}$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds \cite{chen_semiclassical_2006}.
+This is expected since an excitation energy of two green photons ($\hbar \omega$ = 2.37eV) will be required to overcome the reported $\Phi$ = 4.69eV effective work function of a thick gold film contaminated with adsorbed water~\cite{monjushiro_ultraviolet_1991} or ensure that at least the tail of the electron Fermi distribution has sufficient energy to overcome the 5.1($\pm$0.1)eV work function of a clean Au surface~\cite{eastman_photoelectric_1970}.
+Moreover, when $2\hbar \omega \approx \Phi$, the dependence of the yield on the square of the electron's excess energy above the work function~\cite{monjushiro_ultraviolet_1991} ensures that predominantly only the high energy (Boltzmann) tail of the Fermi distribution contributes to the emission.
+Consequently, the observed increase in $\Delta p_{\smallT}$ with incident pulse energy must be due to a heating effect; specifically, laser heating of the free electron Fermi gas as two-photon excitation is instantaneous and the $\sim$200fs green laser pulse duration excludes any coupling to the lattice which occurs on the time scale of a few picoseconds~\cite{chen_semiclassical_2006}.
 The solid red line in \ref{fig:gold-emission} is the predicted variation derived using a simple zero free parameter model of this effect.
-Using the known optical properties of gold \cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm \cite{seah_quantitative_1979}.
+Using the known optical properties of gold~\cite{johnson_optical_1972}, which give a reflectance $R_p$ = 0.53 and an absorption depth of 20nm, and the temperature-dependent heat capacity of a free electron Fermi gas, the model evaluates the average temperature of the laser-heated electron gas that may be two-photon excited assuming that only electrons within a few nanometers of the surface can be emitted, as the mean free path for electrons in Au is $\sim$4nm~\cite{seah_quantitative_1979}.
 This average electron temperature $T_e$ is then used in the AG model simulation of the experiment (\ref{fig:transverse-measurement}) to determine the expected Fourier plane spot size on the YAG scintillator; that is, using an initial $\Delta p_{\smallT} = \sqrt{m k_B T_e}$.

### 2013-04-11 inc/design_considerations/considerations.tex
as I have suspected, the only previous model I need to introduce is the AG model.
This commit changes the introductions of several chapters and lowers the headings by one in the new AG model chapter

@@
 
-In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effects that column elements have upon it.
+In Chapters \ref{chap:ms_model} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effects that column elements have upon it.
 However, since microscopy is an experimental science, we are also designing and building a prototype UEM system at UIC.

### 2013-04-11 inc/extension/introduction.tex
as I have suspected, the only previous model I need to introduce is the AG model.
This commit changes the introductions of several chapters and lowers the headings by one in the new AG model chapter

In Chapter \ref{chap:ms_model}, I have introduced the AG model, originally presented by Michalik and Sipe~\cite{michalik_analytic_2006}.
Although it is a very impressive model and a very useful starting point, because it cannot include the influence of external forces on the pulse dynamics, it does not provide the complete picture of the evolution of an electron pulse through the entire column.
Futher, it does not contain a set of initial conditions which reflect a realistic photoemission process.

In the following chapter, I will derive a set of initial conditions which should be sufficient to model single-photon photoemission from a metal, and which may serve as a basis for future implementations employing other photoemission processes.
I will then present a model which builds on and extends the AG model to include the influence of external forces on pulses.
I will also include the mathematical representations of magnetic lenses, TM$_{010}$ RF cavities, and DC accelerators.
With these computational tool it is possible to form one model which has the ability to simulate a UEM column.
Indeed I have implemeted this model in the Perl programming language, in a form generic enough to model many different columns and yet is computationally fast enough to be used to design and optimize a column.

### 2013-04-11 inc/ms_model/ms_model.tex
as I have suspected, the only previous model I need to introduce is the AG model.
This commit changes the introductions of several chapters and lowers the headings by one in the new AG model chapter

@@
 
-\section{The Analytic Gaussian Model} \label{sec:ms_model}
-
-This section introduces the Analytic Gaussian (AG) model of Michalik and Sipe~\cite{michalik_analytic_2006}.
+This chapter introduces the Analytic Gaussian (AG) model of Michalik and Sipe~\cite{michalik_analytic_2006}.
 In the AG model the dynamics of a Gaussian electon pulse are computed using a mean-field self-similar approximation.
@@
 
-The only force included in the AG model is that of the internal repulsive space-charge field.
+A drawback of this model is that the only force included is that of the internal repulsive space-charge field.
 In Chapter \ref{chap:extension}, I will extend this model, adding the influence of external forces, making it more generally applicable for UEM.
 
-\subsection{Formal Evolution} \label{sec:formal_evolution}
+\section{Formal Evolution} \label{sec:formal_evolution}
 
@@
 
-\subsection{Evolution of a Gaussian Distribution}
+\section{Evolution of a Gaussian Distribution}
 
@@
 
-\subsection{Generating the Differential Equations}
+\section{Generating the Differential Equations}
 
@@
 
-\subsection{The Differential Equation System}
+\section{The Differential Equation System}
 
@@
 
-\subsection{Validity and Limitations}
+\section{Validity and Limitations}
 

### 2013-04-11 main.tex
flattened model_results section headings, removing free-space versus forced
remove (essentially) unused \inputdata command, which was meant to ignore loading tikz data
reconstitute mytikz as a package, with option external to handle this stuff
be more careful with default external handling
added rf cavity comparison plot (like figs 5(a)/6(a) from JAP paper)
I have a title!
added single photon section and basic plot
more work on single_photon section
improve symbolic consistency between este paper derived sections and the rest of the paper
killing more TODOs
began working on a conclusion sectio

@@
 \DeclareGraphicsExtensions{.png,.pdf,.jpg}
-\input{mytikz}
-\setboolean{inputdata}{true}
+
+\usepackage[noexternal]{mytikz}
+
 \usepackage{fancyvrb}
@@
 
-\title{Title}
+\title{Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling}
 \author{Joel Berger}
@@
 \tableofcontents
-\listoftables
+%\listoftables
 \listoffigures
@@
 
-\chapter{Review of Foundational Modeling Efforts} \label{chap:previous_models}
+  \subimport{inc/}{introduction}
+
+\chapter{The Analytic Gaussian Model} \label{chap:ms_model}
 
@@
 
-Modeling a system as complex as an Ultrafast Electron Microscope is an enourmous task.
-In Chapter \ref{chap:previous_models}, I have highlighted some previous efforts at modeling electron pulse evolution.
-While these are excellent works in their own right, none provide the complete picture of the evolution of an electron pulse through the entire column.
-In the following chapter, I will present a model which builds on and extends those individual pieces to form a single model; one which is generic enough to model many different columns and is computationally fast enough to be used to design and optimize a column.
+  \subimport{inc/extension/}{introduction}
 
@@
   \subimport{inc/extension/liouvilles_theorem/}{liouvilles_theorem}
-
-  %TODO refer from here to Free expansion results
 
@@
 
-\section{Free Expansion Propagation}
-
   \subimport{inc/model_results/shapes/}{shapes}
@@
 
-\section{Propagtion under External Forces}
+  \subimport{inc/model_results/lens_charge/}{lens_charge}
 
-  \subimport{inc/model_results/lens_charge/}{lens_charge}
+  \subimport{inc/model_results/rf_cav/}{rf_cav}
 
@@
 
-\section{Flat Metal Photocathodes}
+  \subimport{inc/photocathode/single_photon/}{single_photon}
+
+  \subimport{inc/photocathode/two_photon_thermionic/}{two_photon_thermionic}
 
@@
   \subimport{inc/photocathode/este/}{este}
+
+  \subimport{inc/photocathode/}{future}
+
+\chapter{Conclusions}
+
+  \subimport{inc/}{conclusions}
 

### 2013-04-11 inc/extension/introduction.tex
tweaks to extension section introduction

@@
 In Chapter \ref{chap:ms_model}, I have introduced the AG model, originally presented by Michalik and Sipe~\cite{michalik_analytic_2006}.
-Although it is a very impressive model and a very useful starting point, because it cannot include the influence of external forces on the pulse dynamics, it does not provide the complete picture of the evolution of an electron pulse through the entire column.
+Although it is a very impressive model and a very useful starting point, because it cannot include the influence of external forces on the pulse dynamics, it does not provide the complete picture of the evolution of an electron pulse through an entire microscope column.
 Futher, it does not contain a set of initial conditions which reflect a realistic photoemission process.
 
-In the following chapter, I will derive a set of initial conditions which should be sufficient to model single-photon photoemission from a metal, and which may serve as a basis for future implementations employing other photoemission processes.
+In the following chapter, I will derive a set of initial conditions which, though simplistic, should be sufficient to simulate single-photon photoemission from a metal, and which may additionally serve as a basis for future sets of initial conditions for other photoemission processes.
 I will then present a model which builds on and extends the AG model to include the influence of external forces on pulses.
 I will also include the mathematical representations of magnetic lenses, TM$_{010}$ RF cavities, and DC accelerators.
-With these computational tool it is possible to form one model which has the ability to simulate a UEM column.
-Indeed I have implemeted this model in the Perl programming language, in a form generic enough to model many different columns and yet is computationally fast enough to be used to design and optimize a column.
+With these computational tools it is possible to form one model which has the ability to simulate a UEM column.
+Indeed I have implemeted this model in the Perl programming language, in a form generic enough to easily model many different electron column configurations and yet is computationally fast enough to be used to design and optimize a column.

### 2013-04-12 inc/photocathode/este/este.tex
tweaks

@@
 Both samples are cut from [100]-oriented polished wafers; the GaSb is undoped and the InSb is moderately $p$-type.
-Prior studies~\cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included~\cite{dowell_quantum_2009}.
+Prior studies~\cite{gobeli_photoelectric_1965} indicate that single-photon photoemission should not possible for either semiconductor --- the effective photoemission work functions (about 4.8eV for undoped GaSb and 4.89eV for $p$-type InSb (Fermi level pinned at valence band maximum)) being greater than the 4.75eV photon energy even when the 34meV Schottky barrier suppression due to the applied 8kV/cm DC field is included~\cite{dowell_quantum_2009,berger_excited_2012}.
 Nonetheless, for both zincblende semiconductors, significant laser-driven emission (more than expected for a single-photon field emission process) is observed with a yield nearly linearly proportional to the $\sim$4ps UV laser pulse energy.

### 2013-04-12 inc/photocathode/two_photon_thermionic/two_photon_thermionic.tex
tweaks

@@
 where $k_B$ is the Boltzmann constant~\cite{dowell_quantum_2009,jensen_emittance_2010}.
-Since a high laser intensity is required to drive 2PTE, the laser itself will heat the electrons, therefore, if the laser pulse duration is less than the thermalization time of the electron distribution, %TODO check this!
-we expect the measured Fourier plane spot size to grow as incident laser pulse energy increases.
+Since a high laser intensity is required to drive 2PTE, the laser itself will heat the electrons, therefore, we expect the measured Fourier plane spot size to grow as incident laser pulse energy increases, assuming that the laser pulse duration is less than the cooling time of the electron distribution so that the laser is only heating the electrons and not the lattice, which is true in this case.
 

### 2013-04-12 inc/conclusions.tex
massive spellchecking

@@
 
-Ultrafast Electron Microscopy is an exciting new field of intrumentation research which promises to open our eyes to the dynamics occurring in nanoscale processes.
+Ultrafast Electron Microscopy is an exciting new field of instrumentation research which promises to open our eyes to the dynamics occurring in nanoscale processes.
 
@@
 
-In this thesis I have presented the extended Analytic Gaussian (AG) model, an effiecient numerical model based on the work of Michalik and Sipe~\cite{michalik_analytic_2006}, but extended so as to be useful in modeling the dynamics of ultrafast electron pulses in microscope columns.
+In this thesis I have presented the extended Analytic Gaussian (AG) model, an efficient numerical model based on the work of Michalik and Sipe~\cite{michalik_analytic_2006}, but extended so as to be useful in modeling the dynamics of ultrafast electron pulses in microscope columns.
 For pulses containing $N$ carriers of charge $e$ and mass $m$, the model is simply these four equations,
@@
 A third parameter, the local momentum variance $\eta_{\alpha}$, is also encoded in these equations.
-By Liouville's theorem, which the AG model explicitly satisfies, we know that the product of the spatial variance $\sigma_{\alpha}$ and the local momentum variance $\eta_{\alpha}$ is constant throughout propagation (for perfect lensing and without clipping the puse), so we define the constant,
+By Liouville's theorem, which the AG model explicitly satisfies, we know that the product of the spatial variance $\sigma_{\alpha}$ and the local momentum variance $\eta_{\alpha}$ is constant throughout propagation (for perfect lensing and without clipping the pulse), so we define the constant,
 \begin{equation} \label{eq:define_big_gamma_conc}
@@
 and the family of $L$ functions (defined in \ref{eq:ag_l_family}) are smooth and well behaved functions for all physically meaningful $\xi$.
-Finally the $M_{\alpha}^{\lbrace f \rbrace}$ quantities define the $\alpha$ diretion influence of the $f^{\text{th}}$ external force.
+Finally the $M_{\alpha}^{\lbrace f \rbrace}$ quantities define the $\alpha$ direction influence of the $f^{\text{th}}$ external force.
 
@@
 \end{subequations}
-In Section \ref{sec:bin_model} I have also presented a binning model which, when fully investigated, may eventually be employed rather than this simplisitc set of initial conditions.
+In Section \ref{sec:bin_model} I have also presented a binning model which, when fully investigated, may eventually be employed rather than this simplistic set of initial conditions.
 
@@
 This quantity is conserved under propagation through free space and perfect electron-optical, though it may increase for imperfect ones~\cite{oshea_reversible_1998}.
-The normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
+The normalized transverse emittance limits the resolving capability of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
 Since we have already established that a large area emission source must be used, the only avenue to reduce $\varepsilon_{\smallT}$ is to reduce the transverse momentum uncertainty, which is dependent on the photocathode and photoemission process employed.

### 2013-04-12 inc/design_considerations/considerations.tex
massive spellchecking

@@
 
-It is straightforward to show that in order to make a useful image in an electron microcope, the number of electrons collected must be above some minimum number.
+It is straightforward to show that in order to make a useful image in an electron microscope, the number of electrons collected must be above some minimum number.
 The Rose criterion~\cite{rose_television_1948} indicates that an average detector signal strength of $\sim$100 electrons/pixel is required for adequate gray scaling of the image.
@@
 %TODO ref(s)
-For conventional electron microscopes without time resolution, the Rose criterion may be met for arbitrarily low beam current by simply allowing the image to acquire for a long enough time to collect sufficent electrons.
+For conventional electron microscopes without time resolution, the Rose criterion may be met for arbitrarily low beam current by simply allowing the image to acquire for a long enough time to collect sufficient electrons.
 Further, high beam current may be damaging to the sample over time, so a lower beam current is often preferred.
@@
 This steady-state law for the maximum current density that can be achieved is not valid in the regime of interest for UEM (and UED) where the use of picosecond and femtosecond photocathode drive laser pulses implies that $t_{gun}$ is usually much greater than the electron pulse duration $\tau_e$.
-Though there is no clear delineation in the literature or the community between moniker of Ultrafast Electron Microscopy (UEM) and the more general term Dynamic Transmission Electron Microcopy (DTEM), we propose that this is useful delimination, reserving the term UEM for those systems where the pulse can be completely contained inside the accelerator, i.e.
+Though there is no clear delineation in the literature or the community between moniker of Ultrafast Electron Microscopy (UEM) and the more general term Dynamic Transmission Electron Microscopy (DTEM), we propose that this is useful delimitation, reserving the term UEM for those systems where the pulse can be completely contained inside the accelerator, i.e.
 $t_{gun} > \tau_e$.
@@
 For our example case above, to cleanly extract $N \ge 10^8$ electrons/pulse we require $w \gtrsim 0.56 \text{ mm}$.
-The effect that accomodating a large transverse pulse size has on designing an Ultrafast Electron Microscope will be discussed in Chapter \ref{chap:prototype}.
+The effect that accommodating a large transverse pulse size has on designing an Ultrafast Electron Microscope will be discussed in Chapter \ref{chap:prototype}.
 
@@
 
-This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
+This normalized transverse emittance limits the resolving capability of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
 Since \ref{eq:minimum_w} specifies a lower bound on the emission spot size $w$, knowing a UEM will exhibit higher resolution by employing an electron source with the minimum possible transverse emittance directly implies that every effort must be made to reduce the initial transverse momentum spread.

### 2013-04-12 inc/extension/external_forces/formal.tex
as I have suspected, the only previous model I need to introduce is the AG model.
This commit changes the introductions of several chapters and lowers the headings by one in the new AG model chapter
massive spellchecking

@@
 Simulation of electron pulse propagation in UEM and modern UED apparatus~\cite{siwick_ultrafast_2002,oudheusden_electron_2007,reed_femtosecond_2006} requires the ability to model electron-optical elements designed to control the three-dimensional space-charge and dispersion induced broadening of the pulse, and hence deliver electron pulses of the desired quality to the specimen.
-The Analytic Gaussian model as presented by Michalik and Sipe (see Section \ref{sec:ms_model}) is only useful for modeling free-space propagation.
+The Analytic Gaussian model as presented by Michalik and Sipe (see Chapter \ref{chap:ms_model}) is only useful for modeling free-space propagation.
 This is because there is no allowance for external forces, such as those imparted by the ``optical'' elements of an electron microscope column.
@@
 While the free-space AG model does not contain any external forces, it does include an internal force: the Coulomb force is included to model the repulsion of the pulse's electrons from each other.
-As this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics~\cite{michalik_analytic_2006}.
+As this force is self-influencing, it is the major source of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics~\cite{michalik_analytic_2006}.
 
@@
 
-Since these terms are only added to $\dfrac{d \gamma_{\alpha}}{d t}$, the analysis of Liouville's Theorem, shown in Section \ref{sec:liouville}, is unaffected: the \textit{extended} AG model still explicitly satisfies Liouvilles's Theorem.
+Since these terms are only added to $\dfrac{d \gamma_{\alpha}}{d t}$, the analysis of Liouville's Theorem, shown in Section \ref{sec:liouville}, is unaffected: the \textit{extended} AG model still explicitly satisfies Liouville's Theorem.
 

### 2013-04-12 inc/extension/external_forces/practical.tex
massive spellchecking

@@
 The mathematical formalism developed in Section \ref{sec:external_forces} is only an abstract mechanism for including external forces.
-To model electron pulses traveling through realistic electron microcope columns, these abstract mechanisms must be developed further to properly describe each specific electron-optical element.
+To model electron pulses traveling through realistic electron microscope columns, these abstract mechanisms must be developed further to properly describe each specific electron-optical element.
 In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression~\cite{oudheusden_electron_2007} and acceleration %TODO refs
@@
 where each dot denotes a time derivative.
-Attempting to further seperate these equations quickly highlights the tightly coupled nature of the motion of the electron in this system.
+Attempting to further separate these equations quickly highlights the tightly coupled nature of the motion of the electron in this system.
 Compounding the problem, the full description of the off-axis field of a real (non-infinite) air-core solenoid is also surprisingly complicated.
@@
 \end{subequations}
-where $\mu$ is the local magnetic permability, $H$ is the auxiliary field, and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
+where $\mu$ is the local magnetic permeability, $H$ is the auxiliary field, and $F(k)$ and $E(k)$ are the complete elliptic integrals of the first and second kind respectively.
 In the vacuum, $\mu$ is simply the usual free space permeability $\mu_0$; in Section \ref{sec:mag_lens}, we will see how this permeability can be used to improve the performance of magnetic lenses.
@@
   \caption[Comparison of the AG model and a kinematic model of a magnetic lens]{
-    Comparison of electron rays tracked kinematically through a realistic magnetic lens simulation (green) and an AG model simulation of a Guassian electron beam (red) of a similar HW1/eM beam width $ w = 500 \mu\text{m} $.
+    Comparison of electron rays tracked kinematically through a realistic magnetic lens simulation (green) and an AG model simulation of a Gaussian electron beam (red) of a similar HW1/eM beam width $ w = 500 \mu\text{m} $.
   }
@@
 
-For localizing the region of influence of the lens, I have generally modelled a thin magnetic lens as having a super-Gaussian order value $n=1$, a simple Gaussian; of course this may depend on the specifics of the lens.
+For localizing the region of influence of the lens, I have generally modeled a thin magnetic lens as having a super-Gaussian order value $n=1$, a simple Gaussian; of course this may depend on the specifics of the lens.
 To simulate the effect of magnetic lenses in a UEM by the extended AG model, for each magnetic lens, one adds a term like
@@
 The bunches are distributed evenly from $r=0$ to $r = 1$mm.
-These are compared with an AG model simulation of a Guassian electron beam of HW1/eM beam width $ w = 500 \mu\text{m} $ having an initial transverse momentum variance associated with $\Delta E$.
+These are compared with an AG model simulation of a Gaussian electron beam of HW1/eM beam width $ w = 500 \mu\text{m} $ having an initial transverse momentum variance associated with $\Delta E$.
 
@@
 In obtaining \ref{eq:RF_Force_r}, which is also of the form for use in \ref{eq:linear_force}, we have again used an arbitrary super-Gaussian of order $n$ as a region of influence envelope (see \ref{eq:reg_of_influence}) with $L_{RF} \equiv d$, the RF cavity length.
-As will be seen in Section \ref{sec:rf_cav_charge}, under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocusing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $~\cite{kim_rf_1989}, even though its magnitude is maximized for $\phi = 0$.
+As will be seen in Section \ref{sec:rf_cav_charge}, under the pulse compression conditions ($\phi = 0$ when the pulse is at $z^{\prime} = 0$), both effects are defocussing for the electron pulse and the magnetic contribution is generally smaller than the transverse lenses associated with the RF cavity apertures when $ d = \pi v_{\smallzero} / \Omega $~\cite{kim_rf_1989}, even though its magnitude is maximized for $\phi = 0$.
 Other effects, such as magnetic fringe field aperture effects, could also be included, but are significantly weaker.

### 2013-04-12 inc/extension/implementation/implementation.tex
finished including andreas comments on extension chapter
citations are now preceeded by non-breaking space (~)
massive spellchecking

@@
 %The model presented in this chapter is deceptively hard to implement.
-%As noted by Michalik and Sipe in their original paper \cite{michalik_analytic_2006}, in the absence of space-charge, the solution of the system of equations in \ref{eq:ag_original} may be solved analytically and result in the standard equations of Gaussian optics.
+%As noted by Michalik and Sipe in their original paper~\cite{michalik_analytic_2006}, in the absence of space-charge, the solution of the system of equations in \ref{eq:ag_original} may be solved analytically and result in the standard equations of Gaussian optics.
 %With the inclusion of space-charge (that is the inclusion of electron-electron repulsion) the system must be solved by numerical methods.
@@
 Recently the model was rewritten in Perl, and is called \verb!Physics::UEMColumn!.
-Perl is a dynamic scripting language known for its versitility and flexibility.
+Perl is a dynamic scripting language known for its versatility and flexibility.
 In this implementation, the model is implemented as a decoupled set of high-level object-oriented classes.
@@
 \verb!Physics::UEMColumn! has been released as open source software under the terms of the Artistic License (a derivative of the GNU General Public License), which allows for free use of the software.
-The current state of development of the simulation toolkit is publically available at \url{https://github.com/jberger/Physics-UEMColumn}.
+The current state of development of the simulation toolkit is publicly available at \url{https://github.com/jberger/Physics-UEMColumn}.
 Releases of the toolkit are available via the Comprehensive Perl Archive Network (CPAN), which is a central repository of software developed for Perl, a standard tool for the language.
@@
 These objects act as containers for the information about the physical object that they represent.
-Further they manage progammatic tasks related to those objects.
+Further they manage programmatic tasks related to those objects.
 For example, electron-optical element objects manage the Gaussian region-of-influence (\ref{eq:reg_of_influence}) and calculate the force that their fields apply to the pulse.
-This allows for clean, managable and extensible code.
+This allows for clean, manageable and extensible code.
 
@@
 
-Many scripts which use \verb!Physics::UEMColumn! were need to generate data used in the writing of this paper.
+Many scripts which use \verb!Physics::UEMColumn! were need to generate data used in the writing of this thesis.
 I have included most of them in the source repository which is available at \url{https://github.com/jberger/Thesis}.
@@
 Additionally, a development version of the bin model presented in Section \ref{sec:bin_model} is available at \url{https://github.com/jberger/Physics-Photoemission}.
-Once I am sufficently confident in its implementation, and more importantly its voracity, I intend to release it to the CPAN as well, most likely under the name \verb!Physics::Photoemission!.
-This module depends on the much larger Perl project PDL (\url{http://pdl.perl.org}), a matrix-oriented numerical library, making it slightly more difficult to install, however the speed and implementational simplicy benefit added by that dependency is marked.
+Once I am sufficiently confident in its implementation, and more importantly its voracity, I intend to release it to the CPAN as well, most likely under the name \verb!Physics::Photoemission!.
+This module depends on the much larger Perl project PDL (\url{http://pdl.perl.org}), a matrix-oriented numerical library, making it slightly more difficult to install, however the speed and implementational simplicity benefit added by that dependency is marked.
 
@@
 which uses this binning model to generate the pulse's $z$ direction initial conditions.
+
+%TODO compare to luiten RF compressed pulse system as example (or might go in results section)

### 2013-04-12 inc/extension/initial_conditions/bin_model_figure.tex
massive spellchecking

@@
     Compared are: \subref{fig:bin_model_acc} the DC acceleration dominated case, \subref{fig:bin_model_vmax} the photoemission energy dominated case, and \subref{fig:bin_model_equal} an intermediate case.
-    For the Gaussian spatial distribution whose variance is $\sigma_z$ (red) and the momemtum chirp $\gamma_z$ (green) the simulated data (solid) and the fit to the AG allowed form (dashed) are decently well matched in all cases.
+    For the Gaussian spatial distribution whose variance is $\sigma_z$ (red) and the momentum chirp $\gamma_z$ (green) the simulated data (solid) and the fit to the AG allowed form (dashed) are decently well matched in all cases.
     Unfortunately, the model only allows for a constant local momentum variance $\eta_z$ (green) and in all but the extreme acceleration dominated case \subref{fig:bin_model_acc} the model cannot well represent the simulated data.

### 2013-04-12 inc/extension/initial_conditions/initial_conditions.tex
massive spellchecking

@@
 
-The initial tranverse properties of the emitted electron beam are easily understood.
+The initial transverse properties of the emitted electron beam are easily understood.
 The electron pulse's initial spatial variance is simply
@@
 In the longitudinal direction ($z$), the interplay between acceleration, $ a = e E_{DC}/m $, in the DC gun and the physics of the photoemission process determine the initial values of $\sigma_z$, the longitudinal spatial variance, and $\gamma_z$, the longitudinal momentum chirp, of an electron bunch.
-Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) when the incident drive laser pulse duration is sufficently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than the distance traveled by electrons moving with a constant velocity $ v_{max} = \sqrt{ 2 \Delta E / m } $, the maximum photoemission velocity, over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
+Two limiting regimes for the initial longitudinal variances $ \sigma_{i} ( 0 ) $ are readily identified: (i) when the incident drive laser pulse duration is sufficiently long to ensure that the distance the electrons travel due to acceleration in $ E_{DC} $ over the laser pulse duration is much greater than the distance traveled by electrons moving with a constant velocity $ v_{max} = \sqrt{ 2 \Delta E / m } $, the maximum photoemission velocity, over the same time period; and (ii) the opposite regime for sufficiently short laser pulse durations.
 
 % dc dominated
-In the long-pulse regime, the DC field dominated case, we assume that the intrinsic velocity of the emitted electrons is negligable.
+In the long-pulse regime, the DC field dominated case, we assume that the intrinsic velocity of the emitted electrons is negligible.
 The emitted electrons travel an average distance $ \Delta z $, during the $ \Delta t_{ L } $ rms laser pulse duration given by,
@@
 away from the the photocathode, and the the electrons created in the last slice will be at $z=0$.
-By counting the electrons in each spatial bin and fitting to a gaussian, we can determine $\sigma_z(0)$, the inital longitudinal spatial variance of the pulse.
+By counting the electrons in each spatial bin and fitting to a Gaussian, we can determine $\sigma_z(0)$, the initial longitudinal spatial variance of the pulse.
 Taking a weighted average of the velocity of the electrons in each spatial bin and fitting to a linear function across the pulse (as defined by the AG model), we can determine the initial longitudinal velocity (momentum) chirp $\gamma_z(0)$.
@@
 Finally, the velocity uncertainty in each bin can be computed, however unlike the previous parameters, the AG model only allows for a constant value of $\eta_z(0)$, the local momentum variance across the pulse.
-Unfortunately, \ref{fig:bin_model} clearly demonstrates that the local momentum variance is not constant over the inital photo-generated electron bunch; in fact it varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch.
-The narrowest momentum distribution are clearly at the front of the bunch where only the earliest generated electrons with initial velocity near $ v_{max} $ initially reside; the largest distribution is at the back where the last electrons generated still haven't had time to seperate from one another.
+Unfortunately, \ref{fig:bin_model} clearly demonstrates that the local momentum variance is not constant over the initial photo-generated electron bunch; in fact it varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch.
+The narrowest momentum distribution are clearly at the front of the bunch where only the earliest generated electrons with initial velocity near $ v_{max} $ initially reside; the largest distribution is at the back where the last electrons generated still haven't had time to separate from one another.
 

### 2013-04-12 inc/extension/introduction.tex
massive spellchecking

@@
 Although it is a very impressive model and a very useful starting point, because it cannot include the influence of external forces on the pulse dynamics, it does not provide the complete picture of the evolution of an electron pulse through an entire microscope column.
-Futher, it does not contain a set of initial conditions which reflect a realistic photoemission process.
+Further, it does not contain a set of initial conditions which reflect a realistic photoemission process.
 
@@
 With these computational tools it is possible to form one model which has the ability to simulate a UEM column.
-Indeed I have implemeted this model in the Perl programming language, in a form generic enough to easily model many different electron column configurations and yet is computationally fast enough to be used to design and optimize a column.
+Indeed I have implemented this model in the Perl programming language, in a form generic enough to easily model many different electron column configurations and yet is computationally fast enough to be used to design and optimize a column.

### 2013-04-12 inc/extension/liouvilles_theorem/liouvilles_theorem.tex
massive spellchecking

@@
 
-We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed \textit{et al.}~\cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the trasverse spatial emittance of an electron beam~\cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
+We note that, in the transverse directions ($\smallT$), this is consistent with the definition of `coherent fluence' employed by Reed \textit{et al.}~\cite{reed_evolution_2009} for time-resolved electron microscopy and, of course, the transverse spatial emittance of an electron beam~\cite{jensen_theoretical_2006,siwick_ultrafast_2002}.
 Additionally, though the AG model explicitly satisfies Liouville's Theorem, there are cases where the pulse's phase space product may change in a realistic system.
-This effect is called ``emittance growth'' and is most commonly seen in TEM as a result of aberations in lens systems~\cite{oshea_reversible_1998}.
+This effect is called ``emittance growth'' and is most commonly seen in TEM as a result of aberrations in lens systems~\cite{oshea_reversible_1998}.
 As will be seen in Section \ref{extension_limitations}, the AG model will, of necessity, only consider perfect lenses.

### 2013-04-12 inc/gun/gun_design.tex
massive spellchecking

@@
 The system must first be able to accept the large laser spot size needed to generate the electron beam.
-Then, in order to accomodate this large electron pulse, both the cathode Wehnelt and anode aperture must be similarly large; in fact, the beam size at the anode aperture will typically be larger than at the emission source due to transverse beam expansion during acceleration.
+Then, in order to accommodate this large electron pulse, both the cathode Wehnelt and anode aperture must be similarly large; in fact, the beam size at the anode aperture will typically be larger than at the emission source due to transverse beam expansion during acceleration.
 Also, care must be taken to ensure a relatively uniform electric field for acceleration of the pulse.
-While in transit from the cathode to the anode, this large beam is more susceptable to non-uniformities in the electic field given that it will span a larger volume of the acceleration region than the thin beams used in conventional TEM.
+While in transit from the cathode to the anode, this large beam is more susceptible to non-uniformities in the electric field given that it will span a larger volume of the acceleration region than the thin beams used in conventional TEM.
 
@@
 These requirements led to the fabrication of cathode/Wehnelt and anode pair based on the design of Togawa \textit{et al.}~\cite{togawa_ceb6_2007} (pictured in \ref{fig:togawa-pic}).
-A schematic of this design, overlayed by a Finite Element Model (FEM) simulation of the electic field, is shown in \ref{fig:gun-field}.
+A schematic of this design, overlaid by a Finite Element Model (FEM) simulation of the electric field, is shown in \ref{fig:gun-field}.
 The simulation was performed using \texttt{FreeFEM++}, a free finite element partial differential equation solver and meshing toolkit (\url{http://www.freefem.org/ff++/index.htm}).
@@
   \caption[An example of a tanh form acceleration region for three different ``sharpness'' values]{
-    An example of a tahn form acceleration region (overall length 6.35mm) for three different ``sharpness'' values, the same data as in \ref{fig:gun-sharpness}\subref{fig:field_on_axis}.
+    An example of a tanh form acceleration region (overall length 6.35mm) for three different ``sharpness'' values, the same data as in \ref{fig:gun-sharpness}\subref{fig:field_on_axis}.
     The square data points represent the on-axis electric field data from FEM simulations.
@@
 \ref{fig:field_on_axis_fem} compares the FEM computed electric field to the form asserted for the AG model (\ref{eq:anode_tanh}) for several values of the sharpness parameter $s$; the lines are the same as in \ref{fig:gun-sharpness}\subref{fig:field_on_axis}.
-As derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electic field which, in turn, will cause a divergence of the beam.
+As derived in Section \ref{sec:gun_model}, the change of electric field strength near the anode will cause some divergence of the electric field which, in turn, will cause a divergence of the beam.
 Future work will include investigating the designs of Butler %TODO reference

### 2013-04-12 inc/gun/gun_model.tex
massive spellchecking

@@
 %TODO reference
-Additionally, the longitudinal electric field stregth sharply decreases near the anode.
+Additionally, the longitudinal electric field strength sharply decreases near the anode.
 As the front of the pulse sees this change before the back, this changing field also acts as a pulse compression element.
@@
 \end{equation}
-Without an analytic form for $F$, we can appeal to Gauss' law (in absense of external charge),
+Without an analytic form for $F$, we can appeal to Gauss' law (in absence of external charge),
 \begin{gather}
@@
 where $V$ is the accelerating potential, $z^{\prime}_A$ is the position of the anode (the cathode assumed to be at $z^{\prime} = 0$) and $s$ is a constant quantifying the ``sharpness'' of the fall-off of the electric field.
-Indeed, this form is analgous to the super-Gaussian envalope (\ref{eq:reg_of_influence}), however it is open ended on the cathode side, which is necessary for the electrons to be accelerated away from the photocathode.
+Indeed, this form is analogous to the super-Gaussian envelope (\ref{eq:reg_of_influence}), however it is open ended on the cathode side, which is necessary for the electrons to be accelerated away from the photocathode.
 Given this form the ``lensing'' strengths are then
@@
   \caption[An example of a tanh form acceleration region for three different ``sharpness'' values]{
-    An example of a tahn form acceleration region (overall length 6.35mm) for three different ``sharpness'' values.
+    An example of a tanh form acceleration region (overall length 6.35mm) for three different ``sharpness'' values.
     The colored electric fields in figure \protect\subref{fig:field_on_axis} correspond to the pulse width (solid) and length (dashed) in figure \protect\subref{fig:tanh_sharpness}.

### 2013-04-12 inc/hardware/introduction.tex
massive spellchecking

@@
 It is common in the DTEM field to modify an existing transmission electron microscope column, adapting it for laser-stimulated electron generation.
-For the reasons covered in Section \ref{sec:childs_law}, the prototype UEM at UIC is being built to accomodate electron pulses with a large transverse size.
-To satisfy this restriction, it has been designed and built from parts either fabricated in-house or from commercially available sources, both for the ease of prototyping inherent in using replacable parts and to accomodate the physical limitations that our choice of a large beam width have imposed on the standard electron microscope column components.
+For the reasons covered in Section \ref{sec:childs_law}, the prototype UEM at UIC is being built to accommodate electron pulses with a large transverse size.
+To satisfy this restriction, it has been designed and built from parts either fabricated in-house or from commercially available sources, both for the ease of prototyping inherent in using replaceable parts and to accommodate the physical limitations that our choice of a large beam width have imposed on the standard electron microscope column components.
 

### 2013-04-12 inc/hardware/laser/laser.tex
massive spellchecking

@@
 The Watt-level green output power from the frequency-doubled sub-picosecond Yb:KGW oscillator corresponds to a visible peak laser pulse power in excess of 100kW, which is well-suited for further frequency conversion through harmonic generation or pumping of an ultrashort pulse optical parametric oscillator.
-To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk~\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped~\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators~\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media~\cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.
+To our knowledge, the generation of sub-picosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk~\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped~\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators~\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media~\cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.
 
@@
   \includegraphics{performance.jpg}
-  \caption[Power measuments of employed ultrast laser]{
+  \caption[Power measurements of employed ultrafast laser]{
 Performance of the diode-pumped, TLS, Yb:KGW laser (in the cavity configuration of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} ) as a function of applied diode current: continuous-wave (CW) output power (black squares); mode locked (ML) output power (red circles); frequency doubled power (green triangles); and second harmonic generation (SHG) efficiency (line).
@@
 
-The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was intially frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
+The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was initially frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
 The crystal is uncoated and cut for frequency doubling with the fundamental $p$-polarized radiation at 1046nm incident at Brewster's angle --- a geometry that ensures both minimum loss for the fundamental radiation and, more importantly, a high damage threshold.

### 2013-04-12 inc/hardware/magnetic_lenses/maglens.tex
massive spellchecking

@@
 While in principle one could design a magnetic lens using \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
-Further --- as suggested by the use of the auxiliary field $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility.
+Further --- as suggested by the use of the auxiliary field $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeability.
 These materials, called ``magnetic shielding'' since they capture the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.
@@
 There are two end pieces which are revolution solids of an ``L'' shaped profile.
-These ends slide together over the drift tube until touching the disk at the midpoint; the disk is of the proper outer diameter to match the sheilding's smaller outer diameter.
+These ends slide together over the drift tube until touching the disk at the midpoint; the disk is of the proper outer diameter to match the shielding's smaller outer diameter.
 The combination of the two ends and this disk make a ``U'' shape which contains the windings.
-The top of the each end's large outer disk (the walls of the winding container) are flattend on eight sides.
+The top of the each end's large outer disk (the walls of the winding container) are flattened on eight sides.
 These are the mount points for eight bridge pieces which connect the ends to complete the field-capturing shielding loop.
@@
 As the lenses have no active cooling system, it was prudent to characterize the thermal response to a moderate current load.
-Resistance of metals changes as a function of temperature by $ R = R_0 [ 1 + \alpha ( T - T_0 ) ] $, where $\alpha$ is a material contant; for copper, $\alpha = 3.9 * 10^{-3} \text{K}^{-1}$. 
+Resistance of metals changes as a function of temperature by $ R = R_0 [ 1 + \alpha ( T - T_0 ) ] $, where $\alpha$ is a material constant; for copper, $\alpha = 3.9 * 10^{-3} \text{K}^{-1}$. 
 One can calculate the temperature increase of the system by monitoring the change in resistance.

### 2013-04-12 inc/model_results/lens_charge/lens_charge.tex
massive spellchecking

@@
 In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses~\cite{oudheusden_electron_2007,lagrange_nanosecond_2008}.
-In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry~\cite{berger_dc_2009}.
+In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pix elation determining the spatial resolution in a TEM geometry~\cite{berger_dc_2009}.
 The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity), while still subject to Liouville's Theorem (see Section \ref{sec:liouville}).
@@
 When focusing the pulse it can be shown that oblate pulses will also have markedly better performance.
-\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for intially oblate ($ \xi ( 0 ) = 0.1 $) or prolate ($ \xi ( 0 ) = 10 $) pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm and a `short' focal length of 6.0mm.
+\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for initially oblate ($ \xi ( 0 ) = 0.1 $) or prolate ($ \xi ( 0 ) = 10 $) pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm and a `short' focal length of 6.0mm.
 The pulse's initial volume is constant in all cases ($w_{\smallT}^2 w_{z} = (232 \mu\text{m})^3$).

### 2013-04-12 inc/model_results/lens_charge/lens_charge_plot.tex
improve lens_charge multiplot
killing TODOs
more corrections and punctuation
massive spellchecking

@@
-%TODO fix axes to be more consistent
-%TODO add N=1e8 for oblate?
-
 \begin{figure}
@@
         \input{lens_60mm_oblate_data}
+        \node at (2.7,4.9) [align=left] {$\xi=0.1$\\$f=$ 60mm};
         \node at (5.9,1.65) {$N=10^5$};
-        \node at (4,3) {$N=10^6$};
-        \node at (3,4) {$N=10^7$};
+        \node at (4.5,2.7) {$N=10^6$};
+        \node at (4,3.9) {$N=10^7$};
       \end{tikzpicture}
@@
         \input{lens_60mm_prolate_data}
+        \node at (2.7,4.9) [align=left] {$\xi=10$\\$f=$ 60mm};
         \node at (4.2,2.3) {$N=10^5$};
@@
         \input{lens_6mm_oblate_data}
+        \node at (2.7,4.9) [align=left] {$\xi=0.1$\\$f=$ 6mm};
         \node at (6.5,3) {$N=10^7$};
@@
         \input{lens_6mm_prolate_data}
+        \node at (2.7,4.9) [align=left] {$\xi=10$\\$f=$ 6mm};
         \node at (6.9,2.6) {$N=10^7$};
@@
     AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$.
-    Each plot's pulses begin at the lens' front focal plane at 20 kV velocity ($c/3$) and have the same volume.
-    The left plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are initially oblate shaped, HW1/eM width $ 500 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 0.1 $.
-    The right plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are initially prolate shaped, HW1/eM width $ 107.7 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 10 $.
-    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth.
+    Each plot's pulses begin at the lens' front focal plane at velocity $c/3$ (related to 30 kV energy) and have the same volume.
+    The left plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are initially oblate shaped, HW1/eM width $ 500 \mu \text{m}$ and ellipticity $ \xi ( 0 ) = 0.1 $.
+    The right plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are initially prolate shaped, HW1/eM width $ 107.7 \mu \text{m}$ and ellipticity $ \xi ( 0 ) = 10 $.
+    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal length.
     The top plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_long_prolate}) have a longer focal length $f = $ 60mm compared to the bottom plots (\subref{fig:focus_short_oblate} and \subref{fig:focus_short_prolate}) $ f = $ 6.0mm.

### 2013-04-12 inc/model_results/spacecharge/spacecharge.tex
killing TODOs
massive spellchecking

@@
 On the other hand, space-charge effects are expected to dominate at large pulse charge densities.
-A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
+A full analysis of how high the charge density may be before deleterious space-charge become unmanageable would be a sizable task.
 While much discussion could be had on the degree of distortion allowable, as well as considerations of the initial shape of the pulse and length, and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a progressive series of examples.
@@
     The final HW1/eM pulse width (solid) and length (dashed) after a 15 cm propagation. 
-    Pulses are intially similar except for the number of electrons in the pulse, each having 1 mm (HW1/eM) width and 1 ps in duration travelling at $c/3$ with 0.5eV excess photoemission energy.
+    Pulses are initially similar except for the number of electrons in the pulse, each having 1 mm (HW1/eM) width and 1 ps in duration traveling at $c/3$ with 0.5eV excess photoemission energy.
   }
@@
 In the absence of space-charge effects, the presented oblate shaped pulses should be maintained as such during free-space propagation, because the initial values of the electron pulse's divergence and temporal broadening rate are both proportional to $\Delta E^{1/2}$ but not to $N$.
-
-%TODO highlight the following in the experimental section
-%In other words, the spatial and temporal dispersive effects due to the nonzero value of $\Delta E$ are expected to dominate the space-time dynamics of the electron pulse during its propagation
-%in our investigations because less than 10 5 electrons/pulse
-%are generated.
 
@@
     The final HW1/eM pulse width (solid) and length (dashed) after a 15 cm propagation. 
-    Pulses are intially similar except for the number of electrons in the pulse, each is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy and then traverses a 20mm acceleration region followed by drift region, accelerating to a velocity of $c/3$.
+    Pulses are initially similar except for the number of electrons in the pulse, each is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy and then traverses a 20mm acceleration region followed by drift region, accelerating to a velocity of $c/3$.
   }

### 2013-04-12 inc/ms_model/ms_model.tex
massive spellchecking

@@
 This chapter introduces the Analytic Gaussian (AG) model of Michalik and Sipe~\cite{michalik_analytic_2006}.
-In the AG model the dynamics of a Gaussian electon pulse are computed using a mean-field self-similar approximation.
+In the AG model the dynamics of a Gaussian electron pulse are computed using a mean-field self-similar approximation.
 The result is a set of equations that can be solved very quickly using numerical methods on a modern computer.
@@
 
-For a classical set of $N$ discreet particles (labelled by the index $i$), distributed in 6D space (3 position, 3 momentum) the distribution function is 
+For a classical set of $N$ discreet particles (labeled by the index $i$), distributed in 6D space (3 position, 3 momentum) the distribution function is 
 \begin{equation}
@@
 \end{equation}
-Although this equation looks very similar to \ref{eq:evolution_discreet}, notice that it contains approximations, mean-field that has been seen and in practice the self-similarity of the distribution function, that the discreet discription did not.
+Although this equation looks very similar to \ref{eq:evolution_discreet}, notice that it contains approximations, mean-field that has been seen and in practice the self-similarity of the distribution function, that the discreet description did not.
 
@@
 The formal evolution explored in Section \ref{sec:formal_evolution} allows for any distribution function.
-For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian in space and time, mimicing the laser pulse which generated it.
+For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian in space and time, mimicking the laser pulse which generated it.
 How correct this assumption is will be explored in \ref{sec:initial_conditions}.
@@
 The analysis presented here will consider a cylindrically symmetric pulse.
-For the remainder of this analysis, the subscripts $\smallT$ (transverse) and $z$ (longitudinal) will accompany many quantites, reflecting the congruence of $x$ and $y$.
+For the remainder of this analysis, the subscripts $\smallT$ (transverse) and $z$ (longitudinal) will accompany many quantities, reflecting the congruence of $x$ and $y$.
 The subscript $\alpha$ will refer to either $\smallT$ or $z$.

### 2013-04-12 inc/photocathode/este/este.tex
massive spellchecking

@@
 Note that due to the surprisingly large yield, we repeated the measurements, this time replacing the the scintillator screen with a Faraday cup and repeating the laser energy sequence.
-In this way, the yields for these photocathodes were measured directly and absolutely, unlike the results presented in the preceeding sections, for which relative measurements were sufficient.
+In this way, the yields for these photocathodes were measured directly and absolutely, unlike the results presented in the preceding sections, for which relative measurements were sufficient.
 
@@
 Carrier cooling, primarily by longitudinal optical (LO) phonon emission, and decay out of the upper conduction band to lower states will rapidly deplete the Boltzmann tail above the effective $\sim$1eV work function, thereby curtailing the observed ESTE.
-By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ condunction band~\cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
+By comparison with the $\Gamma_6$ conduction band Fr\"ohlich coupling constant in GaAs, where the characteristic LO phonon emission time is 165fs for a hot electron in the $\Gamma_6$ conduction band~\cite{kash_subpicosecond_1985}, we estimate that hot electrons in the upper $\Gamma_8$ conduction band of both GaSb and InSb emit a LO phonon every $\sim$200fs with an energy of 29 and 24meV, respectively.
 This means that even without population decay mechanisms the temperature of the exited electron distribution $T_e$ drops at a rate of $\sim$1,600K/ps in both semiconductors.

### 2013-04-12 inc/photocathode/plasmon/plasmon.tex
massive spellchecking

@@
 In this geometry, the plasmon is driven on a periodic surface rather than a flat surface attached to glass.
-Futher, in this front-side illuminated geometry the back surface is now free to be attached to a heat sink.
+Further, in this front-side illuminated geometry the back surface is now free to be attached to a heat sink.
 
@@
 
-At the interface of a vacuum and such a simple material, contstraints on the material parameters which may allow surface plasmons, result in the condition that the dielectric constant at the oscillation frequency of the surface plasmon $ \varepsilon(\omega_{SP}) = -1 $~\cite{cottam_introduction_2004}.
+At the interface of a vacuum and such a simple material, constraints on the material parameters which may allow surface plasmons, result in the condition that the dielectric constant at the oscillation frequency of the surface plasmon $ \varepsilon(\omega_{SP}) = -1 $~\cite{cottam_introduction_2004}.
 Substituting this result into \ref{eq:dielectric_as_omega} shows that this frequency of oscillation is simply related to intrinsic plasma frequency by
@@
 
-\subsection{Commerical Grating on Glass}
+\subsection{Commercial Grating on Glass}
 
@@
   \includegraphics{damage.png}
-  \caption{Damage to commerical gold-on-glass holographic grating}
+  \caption{Damage to commercial gold-on-glass holographic grating}
   \label{fig:grating-damage}
@@
 
-After the destruction of the gold foil on glass grating, new gratings were designed to have a higher subtrate thermal conductivity.
-The second generation grating was constructed from a silicon substract then coated with a gold film.
+After the destruction of the gold foil on glass grating, new gratings were designed to have a higher substrate thermal conductivity.
+The second generation grating was constructed from a silicon substrate then coated with a gold film.
 The thermal conductivity of silicon is approximately 100 times that of glass. %TODO reference needed?
@@
       % with parabolic fit to initial dependence (dashed curve)
-      and comparative emission from a plane gold surface under the same irradiantion conditions.
+      and comparative emission from a plane gold surface under the same irradiation conditions.
     Bottom: Two-dimensional electron momentum distributions for plasmon-assisted emission (A-D) and two-photon field emission from gold (E) at the laser intensities indicated on the top graph.
@@
 \ref{fig:pape} summarizes the photoemission results obtained from this gold-coated periodic photocathode under green (523nm) laser irradiation.
-Strong enhanced phototemission was observed for only a narrow $\pm$5mrad range of incidence angles $\theta$ around the expected plasmon resonance angle of 39$^{\circ}$.
+Strong enhanced photoemission was observed for only a narrow $\pm$5mrad range of incidence angles $\theta$ around the expected plasmon resonance angle of 39$^{\circ}$.
 Moreover, this enhanced laser-driven electron emission was only observed for $p$-polarized incident 523nm femtosecond laser radiation.
@@
 This suppression is likely caused by the laser-excited surface plasmon, since on the neighboring flat gold surface, where the suppression due to the laser (20MV/m) and the 20kV acceleration potential (0.8MV/m) is $\Delta \Phi \approx $ 0.18eV, no similar increase in efficiency is observed.
-If this is indeed the case, and the surface plasmon field futher suppresses the barrier to a total $ \Delta \Phi \approx $ 0.35eV to allow direct two-photon photoemission, the field must be of the order of 60MV/m or three times that of the laser field.
+If this is indeed the case, and the surface plasmon field further suppresses the barrier to a total $ \Delta \Phi \approx $ 0.35eV to allow direct two-photon photoemission, the field must be of the order of 60MV/m or three times that of the laser field.
 In \ref{fig:pape} B, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves).
@@
   \input{rotated}
-  \caption[Comparison of the colinear and rotated plasmonic excitation geometries]{
+  \caption[Comparison of the collinear and rotated plasmonic excitation geometries]{
     Comparison of the collinear (above) and rotated (below) plasmonic excitation geometries.
@@
     Each vector diagrams (left) is shown next to a representative observed momentum distribution image (right).
-    The increased divergence is shown to be rotated by the same $45^{\circ}$ as the grating (the original colinear line is shown dashed); evidence that the grating is the cause of the deleterious effect.
+    The increased divergence is shown to be rotated by the same $45^{\circ}$ as the grating (the original collinear line is shown dashed); evidence that the grating is the cause of the deleterious effect.
   }
@@
 The experimental results are shown to the right in \ref{fig:rotated}.
-As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
+As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the collinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
 For a mathematical treatment of the transverse electric field of such patterned surfaces, the reader is referred to Ref.~\cite{watts_sharp_1997}. 
@@
 Although this periodicity should have been sufficient to facilitate the generation of the plasmon, the photocathode performed markedly worse than its sinusoidal predecessor.
-Indeed, no evidence of sucessful plasmon oscillation was demonstrated after repeated attempts.
+Indeed, no evidence of successful plasmon oscillation was demonstrated after repeated attempts.
 

### 2013-04-12 inc/photocathode/setup/setup.tex
massive spellchecking

@@
 I will then outline some future research directions that the group at UIC will continue to pursue.
-Since this is the most active current field in our group's reseach, some very new results are presented near the end of the chapter and future plans are evolving.
+Since this is the most active current field in our group's research, some very new results are presented near the end of the chapter and future plans are evolving.
 

### 2013-04-12 inc/photocathode/single_photon/single_photon.tex
massive spellchecking

@@
 \end{equation}
-where $A$ is a material paramenter (typically of the order of $\sim 10^{-4} \text{eV}^{-2}$ %TODO ref
+where $A$ is a material parameter (typically of the order of $\sim 10^{-4} \text{eV}^{-2}$ %TODO ref
 ), and $R$ is the photocathode surface reflectivity, which is itself a function of the polarization, frequency and angle of the incident light.

### 2013-04-12 inc/extension/external_forces/practical.tex
finding refs and fixing bugs

@@
 To model electron pulses traveling through realistic electron microscope columns, these abstract mechanisms must be developed further to properly describe each specific electron-optical element.
-In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression~\cite{oudheusden_electron_2007} and acceleration %TODO refs
-) and DC acceleration regions (i.e. electron gun dynamics).
+In the following section, I demonstrate concrete applications of the extended AG model, including magnetic lenses, radio-frequency (RF) cavities (useful for both pulse compression~\cite{oudheusden_electron_2007} and acceleration~\cite{kim_rf_1989}) and DC acceleration regions (i.e.
+electron gun dynamics).
 
@@
 \end{equation}
-which demonstrates that the RF cavity acts both as a pulse acceleration (first term on RHS of \ref{eq:RF_Efield_approx}) and a pulse compression or expansion (second term on RHS of \eqref{eq:RF_Efield_approx}) device.
+which demonstrates that the RF cavity acts both as a pulse acceleration (first term on RHS of \ref{eq:RF_Efield_approx}) and a pulse compression or expansion (second term on RHS of \ref{eq:RF_Efield_approx}) device.
 %TODO clarify this sentance and the next

### 2013-04-12 inc/photocathode/single_photon/single_photon.tex
finding refs and fixing bugs

@@
 \end{equation}
-The efficiency of photoemission $\eta_{{\scriptscriptstyle PE}}$, when $\Delta E$ is small compared to $\Phi$ may be written as, %TODO ref
+The efficiency of photoemission $\eta_{{\scriptscriptstyle PE}}$, when $\Delta E$ is small compared to $\Phi$ may be written as~\cite{fowler_analysis_1931,dowell_quantum_2009,maldonado_experimental_2012},
 \begin{equation}
@@
 \end{equation}
-where $A$ is a material parameter (typically of the order of $\sim 10^{-4} \text{eV}^{-2}$ %TODO ref
+where $A$ is a material parameter (typically of the order of $\sim 10^{-4} \text{eV}^{-2}$
 ), and $R$ is the photocathode surface reflectivity, which is itself a function of the polarization, frequency and angle of the incident light.
-%TODO calculation of $\eta_{PE}$?
 

### 2013-04-12 inc/introduction.tex
as I have suspected, the only previous model I need to introduce is the AG model.
This commit changes the introductions of several chapters and lowers the headings by one in the new AG model chapter
tweaks and dedication

@@
 The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
-For completeness, in Section \ref{sec:ms_model} I will introduce, in MKS units, the essential features of the AG model.
+For completeness, in Chapter \ref{chap:ms_model} I will introduce, in MKS units, the essential features of the AG model.
 In Chapter \ref{chap:extension}, I present some optimizations to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
@@
 
-In Chapter \ref{chap:model_results}, I apply the AG model to draw some initial insight into the behaviors of ultrafast electron pulses in a UEM column.
+In Chapter \ref{chap:model_results}, I apply the AG model to draw some initial insight into the dynamics of ultrafast electron pulses traversing a UEM column.
+One of the conclusions that I will draw is that oblate, or ``disk-like'', electron pulses will be generally preferred to prolate ``cigar-like;'' for UEM, the electron pulses are likely to be oblate for reasonably sized incident laser spot sizes.
 Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion~\cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance~\cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.
@@
 All of these considerations motivate the construction of the prototype UEM column being built by our group at UIC.
-In Chapter \ref{chap:prototype} I detail the design and fabrication of large-aperture electron-optical elements, which can readily admit these large width electron pulses; included among these are the accelerator and magnetic lenses.
-The, before I conclude, in \ref{chap:photocathode}, I highlight the photocathode engineering work which has been undertaken in an attempt to reduce the rms transverse momentum.
+In Chapter \ref{chap:prototype}, I detail the design and fabrication of large-aperture electron-optical elements, which can readily admit these large width electron pulses; included among these are the accelerator and magnetic lenses.
+Then, before I conclude, I will in Chapter \ref{chap:photocathode} highlight the photocathode engineering work which has been undertaken in an attempt to reduce the rms transverse momentum.
 I will present several different photoemission processes, some of which are yielding very promising results.

### 2013-04-12 inc/appendix/compound.tex
use real scripts from the paper for the appendix. text to come later

--- 2012-12-12 inc/extension/implementation/Cooke.tex
@@
-\begin{Verbatim}[commandchars=\\\{\}]
+\begin{Verbatim}[commandchars=\\\{\},numbers=left,firstnumber=1,stepnumber=1]
 \PY{c+c1}{\PYZsh{}!/usr/bin/env perl}
@@
 
-\PY{k}{use} \PY{n+nn}{Physics::UEMColumn} \PY{n}{alias} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{':standard'}\PY{p}{;}
-\PY{k}{use} \PY{n+nn}{Physics::UEMColumn::Auxiliary} \PY{l+s}{':materials'}\PY{p}{;}
+\PY{k}{use} \PY{n}{Physics::}\PY{n}{UEMColumn} \PY{n}{alias} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{':standard'}\PY{p}{;}
+
+\PY{k}{use} \PY{n}{PDL}\PY{p}{;}
+\PY{k}{use} \PY{n}{PDL::}\PY{n}{Graphics::}\PY{n}{Prima::}\PY{n}{Simple} \PY{p}{[}\PY{l+m+mi}{700}\PY{p}{,}\PY{l+m+mi}{500}\PY{p}{]}\PY{p}{;}
 
 \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{laser} \PY{o}{=} \PY{n}{Laser}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
-  \PY{n}{width}    \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'1 mm'}\PY{p}{,}
-  \PY{n}{duration} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'1 ps'}\PY{p}{,}
+  \PY{n}{width}    \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'100 um'}\PY{p}{,}
+  \PY{n}{duration} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'0.1 ps'}\PY{p}{,}
   \PY{n}{energy}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'4.75 eV'}\PY{p}{,}
@@
 \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{column} \PY{o}{=} \PY{n}{Column}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
-  \PY{n}{length}       \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'400 mm'}\PY{p}{,} 
+  \PY{n}{length}       \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'350 mm'}\PY{p}{,} 
   \PY{n}{laser}        \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{laser}\PY{p}{,}
   \PY{n}{accelerator}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{acc}\PY{p}{,}
-  \PY{n}{photocathode} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{Photocathode}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}\PY{n}{Ta}\PY{p}{)}\PY{p}{,}
+  \PY{n}{photocathode} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{Photocathode}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(} \PY{n}{work\PYZus{}function} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'4.25 eV'} \PY{p}{)}\PY{p}{,} \PY{c+c1}{\PYZsh{} Ta}
 \PY{p}{)}\PY{p}{;}
 
-\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim} \PY{o}{=} \PY{n+nn}{Physics::}\PY{n}{UEMColumn}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
+\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim} \PY{o}{=} \PY{n}{Physics::}\PY{n}{UEMColumn}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
   \PY{n}{column} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{column}\PY{p}{,}
-  \PY{n}{number}      \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mi}{1}\PY{p}{,}
-  \PY{n}{debug}       \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mi}{1}\PY{p}{,}
+  \PY{n}{number} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mf}{1e6}\PY{p}{,}
+  \PY{n}{steps}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mi}{200}\PY{p}{,}
+  \PY{n}{solver\PYZus{}opts} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{p}{\PYZob{}}
+    \PY{n}{h\PYZus{}max} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mf}{5e-13}\PY{p}{,}
+    \PY{n}{h\PYZus{}init} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mf}{5e-14}\PY{p}{,}
+  \PY{p}{\PYZcb{}}\PY{p}{,}
 \PY{p}{)}\PY{p}{;}
@@
 \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{cooke\PYZus{}sep}  \PY{o}{=} \PY{l+m+mi}{5}\PY{p}{;} \PY{c+c1}{\PYZsh{}cm}
-\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{str\PYZus{}mag}    \PY{o}{=} \PY{l+m+mf}{33e-13}\PY{p}{;}
+\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{str\PYZus{}mag}    \PY{o}{=} \PY{l+m+mf}{43e-13}\PY{p}{;}
 
@@
   \PY{n}{length}    \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'2 cm'}\PY{p}{,}
-  \PY{n}{strength}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'230 kilovolts / m'}\PY{p}{,}
+  \PY{n}{strength}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'335 kilovolts / m'}\PY{p}{,}
   \PY{n}{frequency} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'3 gigahertz'}\PY{p}{,}
@@
 
-\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result} \PY{o}{=} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{propagate}\PY{p}{;}
+\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result} \PY{o}{=} \PY{n}{pdl}\PY{p}{(} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{propagate} \PY{p}{)}\PY{p}{;}
+
+\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{z}  \PY{o}{=} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{slice}\PY{p}{(}\PY{l+s}{'(1),'}\PY{p}{)}\PY{p}{;}
+\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{st} \PY{o}{=} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{slice}\PY{p}{(}\PY{l+s}{'(3),'}\PY{p}{)}\PY{p}{;}
+\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sz} \PY{o}{=} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{slice}\PY{p}{(}\PY{l+s}{'(4),'}\PY{p}{)}\PY{p}{;}
+
+\PY{n}{plot}\PY{p}{(}
+  \PY{o}{-}\PY{n}{st} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{ds::}\PY{n}{Pair}\PY{p}{(} 
+    \PY{n+nv}{\PYZdl{}}\PY{n+nv}{z}\PY{p}{,} \PY{n+nb}{sqrt}\PY{p}{(} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{st} \PY{o}{/} \PY{n}{maximum}\PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{st}\PY{p}{)} \PY{p}{)}\PY{p}{,}
+    \PY{n}{colors} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{pdl}\PY{p}{(}\PY{l+m+mi}{255}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{)}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{rgb\PYZus{}to\PYZus{}color}\PY{p}{,}
+    \PY{n}{plotType} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{ppair::}\PY{n}{Lines}\PY{p}{,}
+    \PY{n}{lineWidths} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mi}{3}\PY{p}{,}
+  \PY{p}{)}\PY{p}{,}
+  \PY{o}{-}\PY{n}{sz} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{ds::}\PY{n}{Pair}\PY{p}{(} 
+    \PY{n+nv}{\PYZdl{}}\PY{n+nv}{z}\PY{p}{,} \PY{n+nb}{sqrt}\PY{p}{(} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sz} \PY{o}{/} \PY{n}{maximum}\PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{sz}\PY{p}{)} \PY{p}{)}\PY{p}{,}
+    \PY{n}{colors} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{pdl}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,}\PY{l+m+mi}{255}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{)}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{rgb\PYZus{}to\PYZus{}color}\PY{p}{,}
+    \PY{n}{plotType} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{ppair::}\PY{n}{Lines}\PY{p}{,}
+    \PY{n}{lineWidths} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+m+mi}{3}\PY{p}{,}
+  \PY{p}{)}\PY{p}{,}
+  \PY{n}{x} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{p}{\PYZob{}} \PY{n}{label} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'Position (m)'} \PY{p}{\PYZcb{}}\PY{p}{,}
+\PY{p}{)}\PY{p}{;}
 \end{Verbatim}

### 2013-04-12 inc/appendix/script.tex
use real scripts from the paper for the appendix. text to come later

\chapter{Example Simulation Script} \label{sec:script}

\begin{singlespace}
  \scriptsize
  \input{compound}
\end{singlespace}

\begin{singlespace}
  \scriptsize
  \input{spacecharge}
\end{singlespace}

### 2013-04-12 inc/appendix/spacecharge.tex
use real scripts from the paper for the appendix. text to come later

\begin{Verbatim}[commandchars=\\\{\},numbers=left,firstnumber=1,stepnumber=1]
\PY{c+c1}{\PYZsh{}!/usr/bin/env perl}

\PY{k}{use} \PY{n}{strict}\PY{p}{;}
\PY{k}{use} \PY{n}{warnings}\PY{p}{;}

\PY{k}{use} \PY{n}{Getopt::}\PY{n}{Long}\PY{p}{;}
\PY{n}{GetOptions}\PY{p}{(} \PY{n}{accelerator} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{o}{\PYZbs{}}\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{use\PYZus{}acc} \PY{p}{)}\PY{p}{;}

\PY{k}{use} \PY{n}{Physics::}\PY{n}{UEMColumn} \PY{n}{alias} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{p}{[}\PY{l+s+sx}{qw/:standard Pulse/}\PY{p}{]}\PY{p}{;}
\PY{k}{use} \PY{n}{Physics::}\PY{n}{UEMColumn::}\PY{n}{Auxiliary} \PY{l+s+sx}{qw/:materials :constants/}\PY{p}{;}

\PY{k}{sub }\PY{n+nf}{for\PYZus{}number} \PY{p}{\PYZob{}}
  \PY{k}{my} \PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{num}\PY{p}{,} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{use\PYZus{}acc}\PY{p}{)} \PY{o}{=} \PY{n+nv}{@}\PY{n+nv}{\PYZus{}}\PY{p}{;}

  \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{p}{;}
  \PY{k}{if} \PY{p}{(} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{use\PYZus{}acc} \PY{p}{)} \PY{p}{\PYZob{}}
    \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{laser} \PY{o}{=} \PY{n}{Laser}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
      \PY{n}{width}    \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'1 mm'}\PY{p}{,}
      \PY{n}{duration} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'1 ps'}\PY{p}{,}
      \PY{n}{energy}   \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'4.75 eV'}\PY{p}{,}
    \PY{p}{)}\PY{p}{;}

    \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{acc} \PY{o}{=} \PY{n}{DCAccelerator}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
      \PY{n}{length}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'20 mm'}\PY{p}{,}
      \PY{n}{voltage} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'30 kilovolts'}\PY{p}{,}
    \PY{p}{)}\PY{p}{;}

    \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{column} \PY{o}{=} \PY{n}{Column}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
      \PY{n}{length}       \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'15 cm'}\PY{p}{,} 
      \PY{n}{laser}        \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{laser}\PY{p}{,}
      \PY{n}{accelerator}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{acc}\PY{p}{,}
      \PY{n}{photocathode} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{Photocathode}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}\PY{n}{Ta}\PY{p}{)}\PY{p}{,}
    \PY{p}{)}\PY{p}{;}

    \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim} \PY{o}{=} \PY{n}{Physics::}\PY{n}{UEMColumn}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
      \PY{n}{column} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{column}\PY{p}{,}
      \PY{n}{number} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{num}\PY{p}{,}
    \PY{p}{)}\PY{p}{;}

  \PY{p}{\PYZcb{}} \PY{k}{else} \PY{p}{\PYZob{}}

    \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim} \PY{o}{=} \PY{n}{Physics::}\PY{n}{UEMColumn}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
      \PY{n}{column} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{Column}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(} \PY{n}{length} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'15 cm'} \PY{p}{)}\PY{p}{,}
      \PY{n}{pulse}  \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{Pulse}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{k}{new}\PY{p}{(}
        \PY{n}{number} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{num}\PY{p}{,}
        \PY{n}{initial\PYZus{}width} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{l+s}{'1 mm'}\PY{p}{,}
        \PY{n}{initial\PYZus{}length} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{p}{(} \PY{n}{vc} \PY{o}{/} \PY{l+m+mi}{3} \PY{o}{*} \PY{l+m+mf}{1e-12} \PY{p}{)} \PY{o}{.} \PY{l+s}{' m'}\PY{p}{,}
        \PY{n}{velocity} \PY{o}{=}\PY{o}{\PYZgt{}} \PY{n}{vc} \PY{o}{/} \PY{l+m+mi}{3}\PY{p}{,}
      \PY{p}{)}\PY{p}{,}
    \PY{p}{)}\PY{p}{;}

  \PY{p}{\PYZcb{}}

  \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result} \PY{o}{=} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{sim}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{propagate}\PY{p}{;}
  \PY{k}{return} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result}\PY{p}{;}
\PY{p}{\PYZcb{}}

\PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{outfilename} \PY{o}{=} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{use\PYZus{}acc} \PY{p}{?} \PY{l+s}{'spacecharge\PYZus{}acc.dat'} \PY{p}{:} \PY{l+s}{'spacecharge\PYZus{}noacc.dat'}\PY{p}{;}
\PY{n+nb}{open} \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{fh}\PY{p}{,} \PY{l+s}{'\PYZgt{}'}\PY{p}{,} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{outfilename} \PY{o+ow}{or} \PY{n+nb}{die} \PY{l+s}{"Cannot open \PYZdl{}outfilename for writing"}\PY{p}{;}

\PY{k}{my} \PY{n+nv}{@}\PY{n+nv}{nums} \PY{o}{=} \PY{n+nb}{map} \PY{p}{\PYZob{}} \PY{p}{(}\PY{l+s}{"1e\PYZdl{}\PYZus{}"}\PY{p}{,} \PY{l+s}{"2.2e\PYZdl{}\PYZus{}"}\PY{p}{,} \PY{l+s}{"4.6e\PYZdl{}\PYZus{}"}\PY{p}{)} \PY{p}{\PYZcb{}} \PY{p}{(}\PY{l+m+mi}{0}\PY{o}{..}\PY{l+m+mi}{8}\PY{p}{)}\PY{p}{;}
\PY{n+nb}{pop} \PY{n+nv}{@}\PY{n+nv}{nums} \PY{k}{for} \PY{l+m+mi}{1}\PY{o}{..}\PY{l+m+mi}{2}\PY{p}{;}

\PY{k}{for} \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{num} \PY{p}{(} \PY{n+nv}{@}\PY{n+nv}{nums} \PY{p}{)} \PY{p}{\PYZob{}}
  \PY{k}{my} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{result} \PY{o}{=} \PY{n}{for\PYZus{}number}\PY{p}{(}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{num}\PY{p}{,} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{use\PYZus{}acc}\PY{p}{)}\PY{p}{;}
  \PY{k}{for} \PY{p}{(}\PY{n+nv}{@\PYZdl{}}\PY{n+nv}{result}\PY{p}{)} \PY{p}{\PYZob{}}
    \PY{k}{next} \PY{k}{unless} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{\PYZus{}}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]} \PY{o}{\PYZgt{}=} \PY{l+m+mf}{0.15}\PY{p}{;}
    \PY{k}{print} \PY{p}{\PYZob{}}\PY{n+nv}{\PYZdl{}}\PY{n+nv}{fh}\PY{p}{\PYZcb{}} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{num} \PY{o}{.} \PY{l+s}{' '} \PY{o}{.} \PY{n+nb}{sqrt}\PY{p}{(} \PY{l+m+mi}{2} \PY{o}{*} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{\PYZus{}}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{p}{[}\PY{l+m+mi}{3}\PY{p}{]} \PY{p}{)} \PY{o}{*} \PY{l+m+mf}{1e3} \PY{o}{.} \PY{l+s}{' '} \PY{o}{.} \PY{n+nb}{sqrt}\PY{p}{(} \PY{l+m+mi}{2} \PY{o}{*} \PY{n+nv}{\PYZdl{}}\PY{n+nv}{\PYZus{}}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{p}{[}\PY{l+m+mi}{4}\PY{p}{]} \PY{p}{)} \PY{o}{*} \PY{l+m+mf}{1e3} \PY{o}{.} \PY{l+s}{"\PYZbs{}n"}\PY{p}{;}
    \PY{k}{last}\PY{p}{;}
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}
\end{Verbatim}

### 2013-04-12 inc/appendix/x-compound.pl

@@ -4,11 +4,13 @@ use strict;
 use warnings;
 
 use Physics::UEMColumn alias => ':standard';
-use Physics::UEMColumn::Auxiliary ':materials';
+
+use PDL;
+use PDL::Graphics::Prima::Simple [700,500];
 
 my $laser = Laser->new(
-  width    => '1 mm',
-  duration => '1 ps',
+  width    => '100 um',
+  duration => '0.1 ps',
   energy   => '4.75 eV',
 );
 
@@ -18,22 +20,26 @@ my $acc = DCAccelerator->new(
 );
 
 my $column = Column->new(
-  length       => '400 mm', 
+  length       => '350 mm', 
   laser        => $laser,
   accelerator  => $acc,
-  photocathode => Photocathode->new(Ta),
+  photocathode => Photocathode->new( work_function => '4.25 eV' ), # Ta
 );
 
 my $sim = Physics::UEMColumn->new(
-  column      => $column,
-  number      => 1,
-  debug       => 1,
+  column => $column,
+  number => 1e6,
+  steps  => 200,
+  solver_opts => {
+    h_max => 5e-13,
+    h_init => 5e-14,
+  },
 );
 
 my $z_rf       = 20; #cm
 my $l_mag_lens = '1 in';
 my $cooke_sep  = 5; #cm
-my $str_mag    = 33e-13;
+my $str_mag    = 43e-13;
 
 my $lens1 = MagneticLens->new(
   location => ($z_rf - $cooke_sep) . 'cm',
@@ -51,10 +57,30 @@ $sim->add_element($lens2);
 my $rf_cav = RFCavity->new(
   location  => $z_rf . 'cm',
   length    => '2 cm',
-  strength  => '230 kilovolts / m',
+  strength  => '335 kilovolts / m',
   frequency => '3 gigahertz',
 );
 $sim->add_element($rf_cav);
 
-my $result = $sim->propagate;
+my $result = pdl( $sim->propagate );
+
+my $z  = $result->slice('(1),');
+my $st = $result->slice('(3),');
+my $sz = $result->slice('(4),');
+
+plot(
+  -st => ds::Pair( 
+    $z, sqrt( $st / maximum($st) ),
+    colors => pdl(255,0,0)->rgb_to_color,
+    plotType => ppair::Lines,
+    lineWidths => 3,
+  ),
+  -sz => ds::Pair( 
+    $z, sqrt( $sz / maximum($sz) ),
[…]

### 2013-04-12 inc/appendix/x-spacecharge.pl

@@ -1,0 +1,72 @@
+#!/usr/bin/env perl
+
+use strict;
+use warnings;
+
+use Getopt::Long;
+GetOptions( accelerator => \my $use_acc );
+
+use Physics::UEMColumn alias => [qw/:standard Pulse/];
+use Physics::UEMColumn::Auxiliary qw/:materials :constants/;
+
+sub for_number {
+  my ($num, $use_acc) = @_;
+
+  my $sim;
+  if ( $use_acc ) {
+    my $laser = Laser->new(
+      width    => '1 mm',
+      duration => '1 ps',
+      energy   => '4.75 eV',
+    );
+
+    my $acc = DCAccelerator->new(
+      length  => '20 mm',
+      voltage => '30 kilovolts',
+    );
+
+    my $column = Column->new(
+      length       => '15 cm', 
+      laser        => $laser,
+      accelerator  => $acc,
+      photocathode => Photocathode->new(Ta),
+    );
+
+    $sim = Physics::UEMColumn->new(
+      column => $column,
+      number => $num,
+    );
+
+  } else {
+
+    $sim = Physics::UEMColumn->new(
+      column => Column->new( length => '15 cm' ),
+      pulse  => Pulse->new(
+        number => $num,
+        initial_width => '1 mm',
+        initial_length => ( vc / 3 * 1e-12 ) . ' m',
+        velocity => vc / 3,
+      ),
+    );
+
+  }
+
+  my $result = $sim->propagate;
+  return $result;
+}
+
+my $outfilename = $use_acc ? 'spacecharge_acc.dat' : 'spacecharge_noacc.dat';
+open my $fh, '>', $outfilename or die "Cannot open $outfilename for writing";
+
+my @nums = map { ("1e$_", "2.2e$_", "4.6e$_") } (0..8);
+pop @nums for 1..2;
+
+for my $num ( @nums ) {
+  my $result = for_number($num, $use_acc);
+  for (@$result) {
+    next unless $_->[1] >= 0.15;
+    print {$fh} $num . ' ' . sqrt( 2 * $_->[3] ) * 1e3 . ' ' . sqrt( 2 * $_->[4] ) * 1e3 . "\n";
+    last;
+  }
+}
+

### 2013-04-12 main.tex
remove unnecessary import and function declaration
update pdegrees
tweaks and dedication
tweak dedication
use real scripts from the paper for the appendix. text to come later
added summary and beginnings of acknowledgements

@@
 
-\usepackage{color}
-\providecommand{\fixme}[1]{\textcolor{red}{#1}}
+%\usepackage{color}
 
@@
 \author{Joel Berger}
-\pdegrees{of the\\University of Illinois at Chicago}
+\pdegrees{B.S., University of Illinois at Chicago, 2005}
 \degree{Doctor of Philosophy in Physics}
@@
 \dedication
-The dedication.
+This thesis is dedicated to my parents, who in love and infinite patience always encouraged me, no matter how many questions I asked; and to my amazing wife, who has believed in me since the day we met.
 
 \acknowledgment
-The acknowledgment.
+I would like to thank my advisor, W. Andreas Schroeder, and all of my committee members, Nestor Zaluzec, %of the Electron Microscopy Center at Argonne National Laboratory
+Alan Nicholls, % of Research Resources Center at UIC
+Christoph Grein and Robert Klie.
+It has been my honor to learn from and work with each one of you.
+
+I would like to thank all of my research group members, past and present, Stephanie Schieffer, Benjamin Rickman, John Hogan, Tuo Li, Daniel Brajkovic, Michael Greco, Michael Szotec, %TODO sp?
+and Robin. %TODO last name!
+I also want to thank former member Nathan Rimmington %TODO sp!
+for laying the groundwork for the presented laser.
+Through your professional efforts and personal interactions, you have made my time at UIC most enjoyable.
+
+I want to thank my collaborators, Nigel Browning, %TODO argonne FIB guys,
+
+Finally, I want to thank the funding agencies which have made this research possible.
+It has been supported by: %TODO list grants
+
 \initials{JAB}
 
-\preface
-The preface.
+%\preface
+%The preface.
 
@@
 \summary
-A summary is required.
+Dynamic Transmission Electron Microscopy, and its picosecond/femtosecond subclass Ultrafast Electron Microscopy, is an emerging field in intrumentation science.
+It attempts to combine the nanoscale spatial resolution of transmission electron microscopes with the temporal resolution of modern ultrafast lasers.
+In this thesis I present my contributions to this young field. 
+These include a novel model for simulating the dynamics os ultrafast electron pulses in electron microscope systems,
+design critera for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron generation.
+I also present the prototype ultrafast electron microscope being built at UIC.
 
@@
 \appendix
-  \subimport{inc/extension/implementation/}{appendix-script}
+  \subimport{inc/appendix/}{script}
 

### 2013-04-12 main.tex
improved acknowledgements

@@
 
-I would like to thank all of my research group members, past and present, Stephanie Schieffer, Benjamin Rickman, John Hogan, Tuo Li, Daniel Brajkovic, Michael Greco, Michael Szotec, %TODO sp?
-and Robin. %TODO last name!
-I also want to thank former member Nathan Rimmington %TODO sp!
-for laying the groundwork for the presented laser.
+I would like to thank all of my research group members, past and present, Stephanie Schieffer, Benjamin Rickman, John Hogan, Tuo Li, Daniel Brajkovic, Michael Greco, Michael Szotek, and Robin Taylor.
 Through your professional efforts and personal interactions, you have made my time at UIC most enjoyable.
+I also want to thank former member Nathan Rimington for laying the groundwork for the presented laser.
 
-I want to thank my collaborators, Nigel Browning, %TODO argonne FIB guys,
+I want to thank my collaborators, Nigel Browning, now of Pacific Nortwest National Laboratory, and Dean Miller and Jon Hiller of the Electron Microscopy Center at Argonne National Laboratory.
 

### 2013-04-12 inc/conclusions.tex
more work on intro and conclusion, with more to come

@@
 
-During my time at UIC I have worked to advance this technology, I have developed a model and implemented it, and I have  both the theoretical (simulation/modeling) and experimental (prototyping and photocathode engineering) fronts.
+My contributions to this field have worked to advance this technology, I have developed a model and implemented it, and I have  both the theoretical (simulation/modeling) and experimental (prototyping and photocathode engineering) fronts.
 
@@
 
-In this thesis I have presented the extended Analytic Gaussian (AG) model, an efficient numerical model based on the work of Michalik and Sipe~\cite{michalik_analytic_2006}, but extended so as to be useful in modeling the dynamics of ultrafast electron pulses in microscope columns.
+In this thesis, I have presented the extended Analytic Gaussian (AG) model, an efficient mean-field numerical model based on the work of Michalik and Sipe~\cite{michalik_analytic_2006}, but extended so as to be useful in modeling the dynamics of ultrafast electron pulses in microscope columns.
 For pulses containing $N$ carriers of charge $e$ and mass $m$, the model is simply these four equations,
@@
 \begin{equation} \label{eq:define_big_gamma_conc}
-  \Gamma_{\alpha}^2 \equiv \sigma_{\alpha} \eta_{\alpha} \,\text{.}
+  \Gamma_{\alpha}^2 \equiv \sigma_{\alpha} \eta_{\alpha} \,\text{,}
 \end{equation}
-By evaluating this constant once, presumably from initial conditions, and later applying \ref{eq:define_big_gamma_conc} we can get the extract these $\eta_{\alpha}$ parameters while only needing to compute the other four.
-The parameter $\xi$ characterizes the ellipticity of the pulse,
+which in the transverse direction is explicitly related to the the normalized transverse emittance (for cylindrically symmetric pulses),
+\begin{equation}
+  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_{\smallT}^2 \right > } = \frac{1}{m c} \sqrt{ \sigma_{\smallT} \eta_{\smallT} } \,\text{.}
+\end{equation}
+By evaluating this constant once, presumably from initial conditions, and later applying \ref{eq:define_big_gamma_conc} one may extract these $\eta_{\alpha}$ parameters, having only needed to compute the other four.
+The parameter $\xi$ characterizes the ellipticity of the cylindrically symmetric pulse,
 \begin{equation}
@@
 
-In UEM, the initial pulse conditions are of notable importance for the performance of the instrument.
-Therefore I have presented a simplistic but useful set of initial conditions for the AG model.
-First lets assume we photocathode which has a work function $\Phi$, the energy an electron must absorb before it is freed to leave the material.
-Lets assume that the photocathode is in an acceleration region, having an applied electric field $E_{DC}$.
-Finally, lets assume we have a laser pulse whose intensity $I$ is of the form,
+In UEM, the initial pulse conditions are of notable importance for the performance of the instrument, therefore having a realistic description of these intial conditions is essential for accurately modeling the pulse dynamics.
+To this end, I have presented a simplistic but useful set of initial conditions for the extended AG model.
+This analysis assumes a photocathode with a work function $\Phi$, the energy an electron must absorb before it is freed to leave the material, placed in a DC gun which accelerates the pulse by an electric field $E_{DC}$.
+It also assumes a laser pulse whose intensity $I$ is of the form,
 \begin{equation}
@@
 \end{subequations}
-In Section \ref{sec:bin_model} I have also presented a binning model which, when fully investigated, may eventually be employed rather than this simplistic set of initial conditions.
+In Section \ref{sec:bin_model} I have also presented a binning model which is being investigated as a possible improvement to this simplistic set of initial conditions.
 
 Most UEM systems will employ magnetic lenses and DC accelerators; there is also increasing interest in using RF cavities for pulse compression.
-The AG model can model the influence of these external forces via the $M$ parameters in \ref{eq:ag_model_gamma_conc}, which are in terms of the position of the center of the pulse $z^{\prime}$ and its velocity $v_{\smallzero}$.
+The extended AG model can simulate the influence of the external forces imparted by these column elements on the electron pulse via the $M$ parameters in \ref{eq:ag_model_gamma_conc}, which are in terms of the position of the center of the pulse $z^{\prime}$ (in the lab frame) and its velocity $v_{\smallzero}$.
 The each magnetic lens, positioned at $z_{mag}^{\prime}$ and of length $L_{mag}$, can be represented simply as 
@@
 
-I have written an implementation of this model for the Perl programming language and released to the Comprehensive Perl Archive Network (CPAN) under the name \verb!Physics::UEMColumn!
+Since the extension to the AG model is valid only within the limits of the analytical method itself, in particular, its mean internal space-charge field and self-similar Gaussian approximations~\cite{michalik_analytic_2006}, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
+Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
+This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
+Consequently, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
+It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
+%Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
+
+I have written an implementation of this model for the Perl programming language and released it to the Comprehensive Perl Archive Network (CPAN) under the name \verb!Physics::UEMColumn!
 for use by the UEM community (\url{https://metacpan.org/module/Physics::UEMColumn}).
-This model is fully object-oriented, which makes writing simulations remarkably easy for a user to read, write and change scripts.
-Under the hood, though, it uses a custom set of C level bindings to the GNU Scientific Library (GSL), so that the numerical equation solving is fast and accurate.
-I have found that this speed, combined with the dynamic flexibility offered by object-oriented Perl, has allowed me to easily write optimization scripts and generate figures which might have been nearly impossible previously; several of these have been shown in this thesis.
+This module is fully object-oriented, which makes it remarkably easy for an end-user to read, write, and change simulation scripts which use it.
+Under the hood, though, it uses a custom set of C level bindings to the GNU Scientific Library (GSL, \url{http://www.gnu.org/software/gsl/}), so that the numerical equation solving is fast and accurate.
+This speed, combined with the dynamic flexibility offered by object-oriented Perl, allows one to easily write optimization scripts and generate figures which might have been nearly impossible previously; several of these have been shown in this thesis.
 
@@
 
-I have expounded on several factors that we feel are key concepts in designing a UEM.
-I have used the above model to show that an electron pulse, whether oblate (disk-like) or prolate (cigar-like), its initial expansion will be primarily in the the shorter dimension, i.e.
-an electron pulse ``wants'' to be spherical.
+This thesis expounds on several factors that are key concepts in designing a UEM.
+I have used the above model to show that for a given electron pulse, whether oblate (disk-like) or prolate (cigar-like), its initial expansion will be primarily in the same direction as its shorter dimension, i.e., an electron pulse ``wants'' to be spherical.
 The model predicts that the use of shorter focal-length magnetic lenses will result in higher fidelity focusing; such lenses limit the impulse delivered to the pulse by reducing the time of flight that the pulse experiences the higher charge densities inherent during focusing.
 Additionally, we have seen that oblate pulses are preferred to prolate pulses in terms of focus fidelity.
-The model predicts that when properly designed and driven, RF cavities can effectively recompress electron pulses which have expanded longitudinally.
+The model also demonstrates that when properly designed and driven, RF cavities can effectively recompress electron pulses which have expanded longitudinally.
 Interestingly, the model suggests that the additional bandwidth generated by space-charge repulsion (especially in higher charge-density pulses) can allow for additional pulse compression even at one-to-one image to object distances.
 
-The Rose criterion~\cite{rose_television_1948} states that in order to have proper gray-scaling in a single-shot DTEM employing a 1k X 1k CCD camera, the pulse must contain at least $10^8$ electrons.
-We have proposed that the term ``ultrafast'' be used in the case where an electron pulse may be completely contained inside the acceleration region of the column.
-The reason for this designation is that such pulses are subject to a different form of Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, which limits its the number of electrons $N_{crit}$ that may be generated by a laser of HW1/eM width $w$ before the pulse is subject to deleterious distortions,
+The Rose criterion~\cite{rose_television_1948} states that in order to have sufficient gray-scaling in a single-shot DTEM employing a 1k X 1k CCD camera for imaging, the pulse must contain at least $10^8$ electrons.
+Additionally, the Child-Langmuir Law~\cite{child_discharge_1911,langmuir_effect_1923} governs the upper-limit of charge extraction in an electron gun. 
+Pulses which short enough to be fully contained inside the acceleration region are subject to a modified form of Child-Langmuir Law~\cite{valfells_effects_2002}, which limits its the number of electrons $N_{crit}$ that may be generated by a laser of HW1/eM width $w$ before the pulse is subject to deleterious distortions,
 \begin{equation}
@@
 where $V_{DC}$ and $d$ are the accelerator voltage and length respectively.
-In order to satisfy the both the Rose criterion and this ultrafast Child-Langmuir restriction, we expect that UEM instruments will need to employ a large-area emission source.
+Due to this distinction, it is proposed that the term ``ultrafast'' should therefore be reserved for pulses of this type.
+In order to satisfy the both the Rose criterion and this ultrafast Child-Langmuir restriction, we expect that UEM instruments will need to employ a relatively large-area emission source, typically of the order of $100\mu\text{m}$ HW1/eM width or greater.
 This has necessitated building a custom acceleration region based on the design of Togawa \textit{et al.}~\cite{togawa_ceb6_2007} which boasts a moderately uniform acceleration field near the axis of the accelerator, as well as other large-aperture electron-optical elements for our prototype UEM column.
@@
 
+\section{Photocathode Engineering}
+
 A common figure of merit for electron beams is its normalized transverse emittance~\cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, which for the usual case of axially symmetric electron pulses has the form,
 \begin{equation}
-  \varepsilon_{\smallT} = \frac{1}{m c} \Delta x \Delta p_{\smallT} \,\text{,}
+  \varepsilon_{\smallT} = \frac{1}{m c} \Delta x \Delta p_{\smallT} = \frac{\Gamma_{\smallT}}{m c} \,\text{,}
 \end{equation}
@@
 The normalized transverse emittance limits the resolving capability of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
-Since we have already established that a large area emission source must be used, the only avenue to reduce $\varepsilon_{\smallT}$ is to reduce the transverse momentum uncertainty, which is dependent on the photocathode and photoemission process employed.
-
-\section{Photocathode Engineering}
+Since we have already established that a large area emission source must be used in UEM, the only avenue to reduce $\varepsilon_{\smallT}$ is to reduce the transverse momentum uncertainty, which is dependent on the photocathode and photoemission process employed.
 
@@
 Though we had success in driving the plasmon, we were unable to demonstrate reduction in $\Delta p_{\smallT}$.
-We believe the problems were due to disturbances in the acceleration field near these nanostructures, though efforts to mitigate those disturbances has yet been unsuccessful.
+We believe the problems were due to disturbances in the acceleration field near these nanostructures, though efforts to mitigate those disturbances have yet been unsuccessful.
 
@@
 In this process the laser promotes carriers from the valence band to an upper conduction band close to the vacuum level, reducing the effective work function.
-The electron temperature, elevated by the laser, causes the tail of the Fermi distribution to exceed the effective work function and thus emission occurs.
+The electron temperature, elevated by the laser-driven photoemission process to $\sim10^3$K, causes the tail of the Fermi distribution to exceed the effective work function and thus emission occurs.
 Importantly, the results we observed were inconsistent with predictions until we considered the effective mass of these upper conduction bands.

### 2013-04-12 inc/introduction.tex
more work on intro and conclusion, with more to come

@@
 
-In this thesis, I will present an extension of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment, to include the influence of linear external forces due to electron optical elements.
-The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
-For completeness, in Chapter \ref{chap:ms_model} I will introduce, in MKS units, the essential features of the AG model.
+In Chapter \ref{chap:ms_model} I will introduce the of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment.
+For free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
+Due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
+ to include the influence of linear external forces due to electron optical elements.
 In Chapter \ref{chap:extension}, I present some optimizations to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
@@
 In order to adapt this extended AG model for simulating pulses in a realistic UEM column, I then present specific contributions that arise from magnetic lenses, RF cavities and DC accelerators.
+The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
 I would like to note that the analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
 
-Since the extension to the AG model is valid only within the limits of the analytical method itself, in particular, its mean internal space-charge field and self-similar Gaussian approximations~\cite{michalik_analytic_2006}, the extension reflects a first-order (i.e., linear force) analysis of the effects of electron optics upon electron pulse propagation.
-Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
-This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
-Consequently, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
-It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
-%Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
-
-In Chapter \ref{chap:model_results}, I apply the AG model to draw some initial insight into the dynamics of ultrafast electron pulses traversing a UEM column.
-One of the conclusions that I will draw is that oblate, or ``disk-like'', electron pulses will be generally preferred to prolate ``cigar-like;'' for UEM, the electron pulses are likely to be oblate for reasonably sized incident laser spot sizes.
+In Chapter \ref{chap:model_results}, I apply the extended AG model to draw some initial insight into the dynamics of ultrafast electron pulses traversing a UEM column.
+One of the conclusions that I will make is that oblate, or ``disk-like'', electron pulses will be generally preferred to prolate ``cigar-like;'' for UEM, the electron pulses are likely to be oblate for reasonably sized incident laser spot sizes.
 Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion~\cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance~\cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.

### 2013-04-12 inc/conclusions.tex
included grant award numbers and other tweaks

@@
 where the subscript $\alpha$ denotes the directionality of the dynamics, either $\smallT$ for the transverse or $z$ for the longitudinal.
-At any time $t$, the $\sigma_{\alpha}$ quantities characterize the spatial variance of the pulse and $\gamma_{\alpha}$ represent the momentum chirp across the pulse.
+At any time $t$, the $\sigma_{\alpha}$ quantities characterize the spatial variance of the pulse and $\gamma_{\alpha}$ represent the momentum chirps across the pulse.
 A third parameter, the local momentum variance $\eta_{\alpha}$, is also encoded in these equations.

### 2013-04-12 main.tex
included grant award numbers and other tweaks

@@
 
-Finally, I want to thank the funding agencies which have made this research possible.
-It has been supported by: %TODO list grants
+Finally, I want to thank the funding agencies which have made this research possible: 
+the National Science Foundation (DMR-0619573) and the Department of Energy (DE-FG52-06NA26213 and DE-FG52-09NA29451), as well as a Department of Education GAANN fellowship (DED P200A070409).
 
@@
 In this thesis I present my contributions to this young field. 
-These include a novel model for simulating the dynamics os ultrafast electron pulses in electron microscope systems,
+These include a novel model for simulating the dynamics of ultrafast electron pulses in electron microscope systems,
 design critera for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron generation.
-I also present the prototype ultrafast electron microscope being built at UIC.
+I also present the prototype ultrafast electron microscope system being built at UIC.
 

### 2013-04-12 inc/introduction.tex
tweak introduction

@@
 
-In Chapter \ref{chap:ms_model} I will introduce the of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment.
+In Chapter \ref{chap:ms_model}, I will introduce the of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing free-space electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment.
 For free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
 Due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
- to include the influence of linear external forces due to electron optical elements.
+Unfortunately, the model lacks the ability to model the dynamics of electron pulses under influence of external forces; this prevents the AG model from being used to simulate the behavior of electron pulses in realistic microscope columns.
+Additionally, to be useful for UEM, for which initial pulse conditions are of notable importance for the performance of the instrument, a realistic set of intial conditions is essential for accurately modeling the pulse dynamics.
+
 In Chapter \ref{chap:extension}, I present some optimizations to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
-I then present my extensions to the Michalik and Sipe AG model which permit the inclusion of generic external forces.
+I then present my extensions to the Michalik and Sipe AG model which permit the inclusion of generic linear external forces, such as those due to perfect electron-optical elements in a microscope column.
 In order to adapt this extended AG model for simulating pulses in a realistic UEM column, I then present specific contributions that arise from magnetic lenses, RF cavities and DC accelerators.
 The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
-I would like to note that the analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
+Note that the analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
 
@@
 In Chapter \ref{chap:prototype}, I detail the design and fabrication of large-aperture electron-optical elements, which can readily admit these large width electron pulses; included among these are the accelerator and magnetic lenses.
-Then, before I conclude, I will in Chapter \ref{chap:photocathode} highlight the photocathode engineering work which has been undertaken in an attempt to reduce the rms transverse momentum.
-I will present several different photoemission processes, some of which are yielding very promising results.
-In fact, our recent work may even challenge our current understanding of photoemission.
+Then, before I conclude, I will in Chapter \ref{chap:photocathode} highlight the photocathode engineering work which has been undertaken in an attempt to reduce the rms transverse momentum of the emitted electron pulses.
+In fact, the search for a low transverse momentum electron source has applications beyond, UEM and UED; the fields of X-ray Free Electron Lasers (XFELs,~\cite{nemeth_high_2010}) and Compact X-ray Light Sources (CXLS) [ref] will also benefit from such a source.
+I will present investigations of several different photoemission processes, some of which are yielding very promising results.
+In fact, our recent work on excited-state thermionic emission~\cite{berger_excited_2012} may even challenge long-standing ideas about which materials and processes may hold the keys to low-emittance (or high-brightness) electron sources.

### 2013-04-12 inc/conclusions.tex
more tweaks including adding a laser section to the conclusions

@@
 This has necessitated building a custom acceleration region based on the design of Togawa \textit{et al.}~\cite{togawa_ceb6_2007} which boasts a moderately uniform acceleration field near the axis of the accelerator, as well as other large-aperture electron-optical elements for our prototype UEM column.
+
 We have also designed and built a custom laser source capable of generating these electron pulses.
+The primary laser radiation source for the studies is a home-built, diode-pumped and thermal-lens-shaped, femtosecond Yb:KGW oscillator~\cite{berger_high-power_2008} delivering 250fs duration pulses at 1047nm and a 63MHz repetition rate.
+This 2W average power laser is frequency doubled in a 3mm lithium triborate (LBO) crystal with 50\% efficiency to yield $\sim$200fs green pulses at 523nm (photon energy, $\hbar \omega$ = 2.37eV).
+Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
+As an ancillary benefit of using an in-house built laser source, the cavity length may be changed; for soliton mode-locked lasers, this length is directly related to the repitition rate.
+By tuning the length, we are able to frequency match a high harmonic of the repitition rate and the natural oscillation of the RF cavity.
+This is done to synchronize the laser and the RF cavity in an essentially passive manner, obviating the need for expensive electronic devices which would be add deleterious electronic phase noise (jitter) to the system.
 

### 2013-04-12 inc/introduction.tex
more tweaks including adding a laser section to the conclusions

@@
 Then, before I conclude, I will in Chapter \ref{chap:photocathode} highlight the photocathode engineering work which has been undertaken in an attempt to reduce the rms transverse momentum of the emitted electron pulses.
-In fact, the search for a low transverse momentum electron source has applications beyond, UEM and UED; the fields of X-ray Free Electron Lasers (XFELs,~\cite{nemeth_high_2010}) and Compact X-ray Light Sources (CXLS) [ref] will also benefit from such a source.
+In fact, the search for a low transverse momentum electron source has applications beyond UEM and UED; the fields of X-ray Free Electron Lasers (XFELs,~\cite{nemeth_high_2010}) and Compact X-ray Light Sources (CXLS) \cite{cxls_workshop_report_2012} will also benefit from such a source.
 I will present investigations of several different photoemission processes, some of which are yielding very promising results.

### 2013-04-12 inc/conclusions.tex
Some last-minute tweaks

@@
 Ultrafast Electron Microscopy is an exciting new field of instrumentation research which promises to open our eyes to the dynamics occurring in nanoscale processes.
-
-My contributions to this field have worked to advance this technology, I have developed a model and implemented it, and I have  both the theoretical (simulation/modeling) and experimental (prototyping and photocathode engineering) fronts.
+Through my contributions to this field, I have worked to advance this new technology.
+These include a developing and implementing a model for simulating the dynamics of ultrafast electron pulses in electron microscope systems, deriving necessary design critera for constructing such a system, and laying the theoretical and experimental groundwork towards selecting a useful photocathode for electron pulse generation.
+These advances have guided the design and construction of the prototype ultrafast electron microscope system being built at UIC.
 
@@
 The model also demonstrates that when properly designed and driven, RF cavities can effectively recompress electron pulses which have expanded longitudinally.
-Interestingly, the model suggests that the additional bandwidth generated by space-charge repulsion (especially in higher charge-density pulses) can allow for additional pulse compression even at one-to-one image to object distances.
+Interestingly, the model suggests that the additional bandwidth generated by space-charge repulsion (especially in higher charge-density pulses) can allow for additional pulse compression, to durations shorter than expected, even at one-to-one image to object distances.
 
@@
 Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
-As an ancillary benefit of using an in-house built laser source, the cavity length may be changed; for soliton mode-locked lasers, this length is directly related to the repitition rate.
-By tuning the length, we are able to frequency match a high harmonic of the repitition rate and the natural oscillation of the RF cavity.
-This is done to synchronize the laser and the RF cavity in an essentially passive manner, obviating the need for expensive electronic devices which would be add deleterious electronic phase noise (jitter) to the system.
+As an ancillary benefit of using an in-house built laser source, the cavity length may be changed; for soliton mode-locked lasers, this length is directly related to the repetition rate.
+By tuning the length, one may frequency match a high harmonic of the repetition rate and the natural oscillation of the RF cavity.
+As a result, the laser and the RF cavity may be synchronized in an essentially passive manner, obviating the need for expensive electronic devices which would be add deleterious electronic phase noise (jitter) to the system.
 
@@
 
-Our most promising results have come from excited-state thermionic emission (ESTE) sources, GaSb and InSb.
+The most promising results have come from excited-state thermionic emission (ESTE) sources, GaSb and InSb.
 In this process the laser promotes carriers from the valence band to an upper conduction band close to the vacuum level, reducing the effective work function.

### 2013-04-12 inc/introduction.tex
Some last-minute tweaks

@@
 To date, such dynamic transmission electron microscopes (DTEMs)~\cite{bostanjoglo_tracing_1980,domer_high-speed_2003} have been developed for both single-shot imaging with $\sim$$10^{5}$nm$\cdot$ps space-time resolution (e.g., 30ns electron pulses and 5-10nm spatial resolution~\cite{lagrange_single-shot_2006,armstrong_practical_2007}) and $\sim$1nm$\cdot$ps space-time resolution ($\sim$100fs with a few nanometer resolution~\cite{park_direct_2009}) in the multi-shot data acquisition mode of operation.
-Both of these instruments employ retro-fitted standard electron microscope columns and are limited in performance mainly by space-charge effects.
+Both of these instruments, which operate with vastly different pulse charge densities, employ retro-fitted standard electron microscope columns and are limited in performance mainly by space-charge effects.
 
@@
 For free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
-Due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
-Unfortunately, the model lacks the ability to model the dynamics of electron pulses under influence of external forces; this prevents the AG model from being used to simulate the behavior of electron pulses in realistic microscope columns.
-Additionally, to be useful for UEM, for which initial pulse conditions are of notable importance for the performance of the instrument, a realistic set of intial conditions is essential for accurately modeling the pulse dynamics.
+Due primarily to the versatility of the AG model, which results from its use of transverse and longitudinal pulse position and momentum variances, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
+Unfortunately, the AG model as laid out by Michalik and Sipe lacks the ability to model the dynamics of electron pulses under influence of external forces; this prevents the AG model from being used to simulate the behavior of electron pulses in realistic microscope columns.
+Additionally, to be useful for UEM, for which the initial electron pulse conditions are of notable importance for the performance of the instrument, a realistic set of intial conditions is essential for accurately modeling the pulse dynamics.
 
@@
 The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
-Note that the analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
+These analyses presented are performed in the non-relativistic limit, which is a reasonable approximation for the typical 20-200keV electron energies employed in UED and electron microscopy.
 
 In Chapter \ref{chap:model_results}, I apply the extended AG model to draw some initial insight into the dynamics of ultrafast electron pulses traversing a UEM column.
-One of the conclusions that I will make is that oblate, or ``disk-like'', electron pulses will be generally preferred to prolate ``cigar-like;'' for UEM, the electron pulses are likely to be oblate for reasonably sized incident laser spot sizes.
+One of the conclusions that I will make is that oblate, or ``disk-like'', electron pulses will be generally preferred to prolate ``cigar-like.'' 
+Indeed for UEM, with typical laser pulse durations in the femtosecond to picosecond timescales, the electron pulses are likely to be oblate for reasonably sized incident laser spot sizes.
 Then in Chapter \ref{chap:considerations}, I apply some well known results from the field of electron microcopy, namely the Rose criterion~\cite{rose_television_1948}, an ultrafast analog to the Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, and the conservation of transverse emittance~\cite{jensen_emittance_2010} (each of which will be introduced) to arrive at two design goals for building a UEM which will produce high-resolution images.
-First, that the electron pulses should be emitted from a large-area source, and second, that every effort must be made to reduce the initial rms transverse momentum (i.e., the emission cone angle).
+First, that the electron pulses should be emitted from a large-area source, and second, that every effort must be made to reduce the initial rms transverse momentum (i.e., the emission cone angle) of the electron pulse as generated from the photocathode.
 

### 2013-04-12 main.tex
Some last-minute tweaks

@@
 These include a novel model for simulating the dynamics of ultrafast electron pulses in electron microscope systems,
-design critera for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron generation.
+design critera for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron pulse generation.
 I also present the prototype ultrafast electron microscope system being built at UIC.

### 2013-04-12 inc/appendix/script.tex
Added text to the appendix demonstrating the scripts.

@@
 \chapter{Example Simulation Script} \label{sec:script}
+
+In this appendix, I want to introduce the Perl implementation of the extended AG model, called \verb!Physics::UEMColumn!.
+This distribution has been released as open source software under the terms of the Artistic License (a derivative of the GNU General Public License), which allows for free use of the software.
+The current state of development is available at \url{https://github.com/jberger/Physics-UEMColumn}, while official releases are available via the Comprehensive Perl Archive Network (CPAN) and its many tracker websites, see for example \url{https://metacpan.org/module/Physics::UEMColumn}.
+
+The module is implemented as fully object-oriented Perl.
+For ordinary differential equation solving, it employs a custom designed, closure-based interface to the GNU Scientific Library (GSL, \url{http://www.gnu.org/software/gsl/}), which uses a Runge-Kutta Prince-Dormand (8,9) step function (RK8PD), so that the numerical equation solving is both fast and accurate (see \url{https://metacpan.org/module/PerlGSL::DiffEq}).
+I have also developed a system by which the object framework can use input units of the user's choosing for convenience (see \url{https://metacpan.org/module/MooseX::Types::NumUnit}).
+
+To write a simulation script, the end-user need only define physical parameters of the system as instances of certain classes, and appropriately add them to the overall simulation object of class \verb!Physics::UEMColumn!.
+In the following script, which produces \ref{fig:compound} (the compound element system), this occurs in lines 11-63.
+The \verb!propagate!
+method call (line 65) then performs all of the simulations given the current state of the overall \verb!Physics::UEMColumn!
+object, which returns a two-dimensional Perl array of results.
+The columns are, in this order: elapsed time $t$, pulse center position $z^{\prime}$, velocity of the pulse center $v_{\smallzero}$, $\sigma_{\smallT}$, $\sigma_{z}$, $\gamma_{\smallT}$, $\gamma_{z}$, $\eta_{\smallT}$, $\eta_{z}$.
+In line 65 I immediately load the returned data into the Perl numerical library PDL (see \url{http://pdl.perl.org}), which is rather like MatLab.
+Using this PDL object, I extract the pulse position (column 1, zero indexed), and the spatial variances (columns 3 and 4).
+I then plot the the data in terms of the normalized width and length of the pulse.
 
@@
 
+As a further example, I also include the script used to generate \ref{fig:spacecharge_noacc} and \ref{fig:spacecharge_acc}.
+In this example, in lines 12-56 I define a function, which simulates the pulse dynamics for a given number of electrons with the option of having a realistic photocathode and acceleration region or an idealized pulse.
+From there, I loop over a set of predefined number of charges per pulse, simulate each sequentially and extract the final pulse dimensions after a propagation distance of 15cm.
+I write this data to a file to be plotted elsewhere.
+
 \begin{singlespace}
@@
 \end{singlespace}
+
+I present these examples to demonstrate the utility of the Perl implementation of the AG model; I hope it will be useful to other members of the UEM community.
+Other examples may be seen in the source repository for this thesis, which is available at \url{https://github.com/jberger/Thesis}.

### 2013-04-12 inc/extension/implementation/implementation.tex
Added text to the appendix demonstrating the scripts.

@@
 For example, electron-optical element objects manage the Gaussian region-of-influence (\ref{eq:reg_of_influence}) and calculate the force that their fields apply to the pulse.
-This allows for clean, manageable and extensible code.
+These objects are managed by a central ``simulation'' class (called \verb!Physics::UEMColumn!) which handles the time evolution of the system.
+This arrangement allows for clean, manageable and extensible code.
 
-These objects are managed by a central ``simulation'' class (called \verb!Physics::UEMColumn!) which handles the time evolution of the system.
-The source package contains an example script analogous to the optical Cooke triplet, containing a magnetic lens, RF cavity and another magnetic lens.
-This script is included as Appendix \ref{sec:script}.
-
-Many scripts which use \verb!Physics::UEMColumn! were need to generate data used in the writing of this thesis.
+Many scripts which use \verb!Physics::UEMColumn! were needed to generate data and figures used for the preparation of this thesis.
 I have included most of them in the source repository which is available at \url{https://github.com/jberger/Thesis}.
+In Appendix \ref{sec:script}, I present the simulation scripts used to create \ref{fig:spacecharge_noacc}, \ref{fig:spacecharge_acc} and \ref{fig:compound}, as examples of this implementation.
 
@@
 This module depends on the much larger Perl project PDL (\url{http://pdl.perl.org}), a matrix-oriented numerical library, making it slightly more difficult to install, however the speed and implementational simplicity benefit added by that dependency is marked.
-
 Of particular interest to users of \verb!Physics::UEMColumn!, \verb!Physics::Photoemission!
@@
 which uses this binning model to generate the pulse's $z$ direction initial conditions.
-
-%TODO compare to luiten RF compressed pulse system as example (or might go in results section)

### 2013-04-12 inc/conclusions.tex
final tweaks and spell check! submission coming soon!

@@
 Through my contributions to this field, I have worked to advance this new technology.
-These include a developing and implementing a model for simulating the dynamics of ultrafast electron pulses in electron microscope systems, deriving necessary design critera for constructing such a system, and laying the theoretical and experimental groundwork towards selecting a useful photocathode for electron pulse generation.
+These include developing and implementing a model for simulating the dynamics of ultrafast electron pulses in electron microscope systems, deriving necessary design criteria for constructing such a system, and laying the theoretical and experimental groundwork towards selecting a useful photocathode for electron pulse generation.
 These advances have guided the design and construction of the prototype ultrafast electron microscope system being built at UIC.
@@
 \end{equation}
-By evaluating this constant once, presumably from initial conditions, and later applying \ref{eq:define_big_gamma_conc} one may extract these $\eta_{\alpha}$ parameters, having only needed to compute the other four.
+By evaluating this constant once, presumably from initial conditions, and later applying \ref{eq:define_big_gamma_conc}, one may extract these $\eta_{\alpha}$ parameters, having only needed to compute the other four.
 The parameter $\xi$ characterizes the ellipticity of the cylindrically symmetric pulse,
@@
 
-In UEM, the initial pulse conditions are of notable importance for the performance of the instrument, therefore having a realistic description of these intial conditions is essential for accurately modeling the pulse dynamics.
+In UEM, the initial pulse conditions are of notable importance for the performance of the instrument, therefore having a realistic description of these initial conditions is essential for accurately modeling the pulse dynamics.
 To this end, I have presented a simplistic but useful set of initial conditions for the extended AG model.
@@
 The extended AG model can simulate the influence of the external forces imparted by these column elements on the electron pulse via the $M$ parameters in \ref{eq:ag_model_gamma_conc}, which are in terms of the position of the center of the pulse $z^{\prime}$ (in the lab frame) and its velocity $v_{\smallzero}$.
-The each magnetic lens, positioned at $z_{mag}^{\prime}$ and of length $L_{mag}$, can be represented simply as 
+Each magnetic lens, positioned at $z_{mag}^{\prime}$ and of length $L_{mag}$, can be represented simply as 
 \begin{equation}
@@
 should be added to the equation of motion governing the propagation of the pulse down the column.
-Finally, if we assume that an acceleration region, with cathode at $z^{\prime} = 0$, anode at $z^{\prime}_A$, and acceleration potential $V$, has an on-axis force of the form
+Finally, if an acceleration region with cathode at $z^{\prime} = 0$, anode at $z^{\prime}_A$, and acceleration potential $V$, has an on-axis force of the form
 \begin{equation}
@@
 for use by the UEM community (\url{https://metacpan.org/module/Physics::UEMColumn}).
+It is released as free software under the terms of the Artistic License (a derivative of the GNU General Public License).
 This module is fully object-oriented, which makes it remarkably easy for an end-user to read, write, and change simulation scripts which use it.
@@
 I have used the above model to show that for a given electron pulse, whether oblate (disk-like) or prolate (cigar-like), its initial expansion will be primarily in the same direction as its shorter dimension, i.e., an electron pulse ``wants'' to be spherical.
-The model predicts that the use of shorter focal-length magnetic lenses will result in higher fidelity focusing; such lenses limit the impulse delivered to the pulse by reducing the time of flight that the pulse experiences the higher charge densities inherent during focusing.
+The model predicts that the use of shorter focal-length magnetic lenses will result in higher fidelity focusing; such lenses limit the impulse delivered to the pulse by reducing the time of flight during which the pulse experiences the higher charge densities inherent while focusing.
 Additionally, we have seen that oblate pulses are preferred to prolate pulses in terms of focus fidelity.
@@
 Additionally, the Child-Langmuir Law~\cite{child_discharge_1911,langmuir_effect_1923} governs the upper-limit of charge extraction in an electron gun. 
-Pulses which short enough to be fully contained inside the acceleration region are subject to a modified form of Child-Langmuir Law~\cite{valfells_effects_2002}, which limits its the number of electrons $N_{crit}$ that may be generated by a laser of HW1/eM width $w$ before the pulse is subject to deleterious distortions,
+Pulses which are short enough to be fully contained inside the acceleration region are subject to a modified form of Child-Langmuir Law~\cite{valfells_effects_2002}, which limits its the number of electrons $N_{crit}$ that may be generated by a laser of HW1/eM width $w$ before the pulse is subject to deleterious distortions,
 \begin{equation}
@@
 Further doubling with a cylindrical focusing geometry in a 7mm $\beta$-barium borate (BBO) crystal yields 261nm ($\hbar \omega$ = 4.75eV) UV pulses with a duration of $\sim$4ps (due to the 600fs/mm group velocity mismatch between the green and generated UV).
-As an ancillary benefit of using an in-house built laser source, the cavity length may be changed; for soliton mode-locked lasers, this length is directly related to the repetition rate.
-By tuning the length, one may frequency match a high harmonic of the repetition rate and the natural oscillation of the RF cavity.
+An ancillary benefit of using an in-house built laser source is that the cavity length may be easily changed; for soliton mode-locked lasers, this length is directly related to the repetition rate.
+By tuning the length, one may frequency match a high harmonic of the repetition rate and the natural oscillation of an RF cavity.
 As a result, the laser and the RF cavity may be synchronized in an essentially passive manner, obviating the need for expensive electronic devices which would be add deleterious electronic phase noise (jitter) to the system.
@@
 This quantity is conserved under propagation through free space and perfect electron-optical, though it may increase for imperfect ones~\cite{oshea_reversible_1998}.
-The normalized transverse emittance limits the resolving capability of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
+The normalized transverse emittance limits the resolving capability of the beam, so this quantity can be thought of as a measure of the beam quality, where a smaller value is preferred~\cite{berger_dc_2009}.
 Since we have already established that a large area emission source must be used in UEM, the only avenue to reduce $\varepsilon_{\smallT}$ is to reduce the transverse momentum uncertainty, which is dependent on the photocathode and photoemission process employed.

### 2013-04-12 inc/introduction.tex
final tweaks and spell check! submission coming soon!

@@
 
-In Chapter \ref{chap:ms_model}, I will introduce the of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing free-space electron bunch propagation in an mean-field self-similar Analytic Gaussian (AG) pulse treatment.
+In Chapter \ref{chap:ms_model}, I will introduce the of the work of Michalik and Sipe~\cite{michalik_analytic_2006,michalik_erratum:_2008}, describing free-space electron bunch propagation in a mean-field self-similar Analytic Gaussian (AG) pulse treatment.
 For free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
@@
 Unfortunately, the AG model as laid out by Michalik and Sipe lacks the ability to model the dynamics of electron pulses under influence of external forces; this prevents the AG model from being used to simulate the behavior of electron pulses in realistic microscope columns.
-Additionally, to be useful for UEM, for which the initial electron pulse conditions are of notable importance for the performance of the instrument, a realistic set of intial conditions is essential for accurately modeling the pulse dynamics.
+Additionally, to be useful for UEM, for which the initial electron pulse conditions are of notable importance for the performance of the instrument, a realistic set of initial conditions is essential for accurately modeling the pulse dynamics.
 

### 2013-04-14 main.tex
typos (Colin Hayes++)

--- 2013-04-12 main.tex
@@
 
-I want to thank my collaborators, Nigel Browning, now of Pacific Nortwest National Laboratory, and Dean Miller and Jon Hiller of the Electron Microscopy Center at Argonne National Laboratory.
+I want to thank my collaborators, Nigel Browning, now of Pacific Northwest National Laboratory, and Dean Miller and Jon Hiller of the Electron Microscopy Center at Argonne National Laboratory.
 
@@
 \summary
-Dynamic Transmission Electron Microscopy, and its picosecond/femtosecond subclass Ultrafast Electron Microscopy, is an emerging field in intrumentation science.
+Dynamic Transmission Electron Microscopy, and its picosecond/femtosecond subclass Ultrafast Electron Microscopy, is an emerging field in instrumentation science.
 It attempts to combine the nanoscale spatial resolution of transmission electron microscopes with the temporal resolution of modern ultrafast lasers.
@@
 These include a novel model for simulating the dynamics of ultrafast electron pulses in electron microscope systems,
-design critera for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron generation.
+design criteria for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron pulse generation.
 I also present the prototype ultrafast electron microscope system being built at UIC.

### 2013-04-14 main.tex
Fixed grammar in Summary

--- 2013-04-12 main.tex
@@
 It attempts to combine the nanoscale spatial resolution of transmission electron microscopes with the temporal resolution of modern ultrafast lasers.
-In this thesis I present my contributions to this young field. 
+In this thesis, I present my contributions to this young field. 
 These include a novel model for simulating the dynamics of ultrafast electron pulses in electron microscope systems,
-design critera for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron generation.
+design criteria for constructing such a system, and theoretical and experimental groundwork geared towards selecting a useful photocathode for electron pulse generation.
 I also present the prototype ultrafast electron microscope system being built at UIC.

### 2013-04-15 inc/ms_model/ms_model.tex
more tweaks including adding a laser section to the conclusions
typos (Jeff Berger++)

@@
 When checked against N-body simulations, the authors find that the model is also surprisingly accurate, even for a large set of non-Gaussian pulses~\cite{michalik_analytic_2006,michalik_evolution_2009}.
-
-A drawback of this model is that the only force included is that of the internal repulsive space-charge field.
+A drawback is that the only force included in this model is that of the internal repulsive space-charge field.
 In Chapter \ref{chap:extension}, I will extend this model, adding the influence of external forces, making it more generally applicable for UEM.
@@
 
-For a classical set of $N$ discreet particles (labeled by the index $i$), distributed in 6D space (3 position, 3 momentum) the distribution function is 
+For a classical set of $N$ discrete particles (labeled by the index $i$), distributed in 6D space (3 position, 3 momentum) the distribution function is 
 \begin{equation}
@@
 \end{equation}
-but with \ref{eq:std_motion} the final discreet evolution equation is
-\begin{equation} \label{eq:evolution_discreet}
+but with \ref{eq:std_motion} the final discrete evolution equation is
+\begin{equation} \label{eq:evolution_discrete}
   \frac{d}{dt} f_{\smallD}(\vec{r}, \vec{p}; t) =
@@
 
-This analysis can be extended to continuous distributions by taking the ensemble average over \ref{eq:evolution_discreet}.
+This analysis can be extended to continuous distributions by taking the ensemble average over \ref{eq:evolution_discrete}.
 In order to simplify the resulting equation to a useful result, one applies the usual mean-field approximation
@@
 \end{equation}
-Although this equation looks very similar to \ref{eq:evolution_discreet}, notice that it contains approximations, mean-field that has been seen and in practice the self-similarity of the distribution function, that the discreet description did not.
+Although this equation looks very similar to \ref{eq:evolution_discrete}, notice that it contains approximations, mean-field that has been seen and in practice the self-similarity of the distribution function, that the discrete description did not.
 

### 2013-04-15 inc/extension/initial_conditions/initial_conditions.tex
grammar (Dan Brajkovic++)

@@
 \end{equation}
-away from the the photocathode, and the the electrons created in the last slice will be at $z=0$.
+away from the the photocathode and the the electrons created in the last slice will be at $z=0$.
 By counting the electrons in each spatial bin and fitting to a Gaussian, we can determine $\sigma_z(0)$, the initial longitudinal spatial variance of the pulse.
@@
 
-Finally, the velocity uncertainty in each bin can be computed, however unlike the previous parameters, the AG model only allows for a constant value of $\eta_z(0)$, the local momentum variance across the pulse.
+Finally, the velocity uncertainty in each bin can be computed, however, unlike the previous parameters, the AG model only allows for a constant value of $\eta_z(0)$, the local momentum variance across the pulse.
 Unfortunately, \ref{fig:bin_model} clearly demonstrates that the local momentum variance is not constant over the initial photo-generated electron bunch; in fact it varies by almost an order of magnitude over the $ 1/e^{2} $ maximum full-width of the electron bunch.

### 2013-04-15 inc/extension/implementation/implementation.tex
typos (Jeff Berger++)

@@
 Additionally, a development version of the bin model presented in Section \ref{sec:bin_model} is available at \url{https://github.com/jberger/Physics-Photoemission}.
-Once I am sufficiently confident in its implementation, and more importantly its voracity, I intend to release it to the CPAN as well, most likely under the name \verb!Physics::Photoemission!.
+Once I am sufficiently confident in its implementation, and more importantly its veracity, I intend to release it to the CPAN as well, most likely under the name \verb!Physics::Photoemission!.
 This module depends on the much larger Perl project PDL (\url{http://pdl.perl.org}), a matrix-oriented numerical library, making it slightly more difficult to install, however the speed and implementational simplicity benefit added by that dependency is marked.

### 2013-04-15 inc/model_results/lens_charge/lens_charge.tex
typos (Jeff Berger++)

@@
 In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses~\cite{oudheusden_electron_2007,lagrange_nanosecond_2008}.
-In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pix elation determining the spatial resolution in a TEM geometry~\cite{berger_dc_2009}.
+In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry~\cite{berger_dc_2009}.
 The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity), while still subject to Liouville's Theorem (see Section \ref{sec:liouville}).

### 2013-05-03 inc/introduction.tex
typo fixes in introduction material

@@
 Such performance optimization efforts will require efficient models of electron pulse propagation dynamics that include the ability to simulate the action of electron optical elements (including the photoelectron gun itself) on the pulse.
-Indeed $2^{\text{nd}}$ generation UED instruments are now beginning to employ both magnetic lenses and RF cavities~\cite{oudheusden_electron_2007,veisz_hybrid_2007} to optimize the electron pulse delivery to the specimen; that is, compensate for global space-charge effects in both the transverse and longitudinal pulse dimensions, respectively.
+Indeed $2^{\text{nd}}$ generation UED instruments are now beginning to employ both magnetic lenses and radio-frequency (RF) resonant cavities~\cite{oudheusden_electron_2007,veisz_hybrid_2007} to optimize the electron pulse delivery to the specimen; that is, compensate for global space-charge effects in both the transverse and longitudinal pulse dimensions, respectively.
 Similarly, future UEM column designs are expected to use RF pulse compression cavities while circumventing high charge density beam cross-overs (i.e., beam foci).
@@
 Due primarily to the versatility of the AG model, which results from its use of transverse and longitudinal pulse position and momentum variances, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
-Unfortunately, the AG model as laid out by Michalik and Sipe lacks the ability to model the dynamics of electron pulses under influence of external forces; this prevents the AG model from being used to simulate the behavior of electron pulses in realistic microscope columns.
-Additionally, to be useful for UEM, for which the initial electron pulse conditions are of notable importance for the performance of the instrument, a realistic set of initial conditions is essential for accurately modeling the pulse dynamics.
+Unfortunately, the AG model as laid out by Michalik and Sipe lacks the ability to model the dynamics of electron pulses under the influence of external forces; this prevents the AG model from being used to simulate the behavior of electron pulses in realistic microscope columns.
+Additionally, to be useful for UEM, for which the electron pulse's initial conditions are of notable importance for the performance of the instrument, a realistic set of initial conditions is essential for accurately modeling the pulse dynamics.
 
-In Chapter \ref{chap:extension}, I present some optimizations to the AG model, and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
-I then present my extensions to the Michalik and Sipe AG model which permit the inclusion of generic linear external forces, such as those due to perfect electron-optical elements in a microscope column.
-In order to adapt this extended AG model for simulating pulses in a realistic UEM column, I then present specific contributions that arise from magnetic lenses, RF cavities and DC accelerators.
+In Chapter \ref{chap:extension}, I present some optimizations to the AG model and develop useful initial conditions representing single-photon photoemission from a typical metal photocathode.
+I then present extensions to the Michalik and Sipe AG model which permit the inclusion of generic linear external forces, such as those due to perfect electron-optical elements in a microscope column.
+In order to adapt this extended AG model for simulating pulses in a realistic UEM column, I then present specific contributions that arise from magnetic lenses, RF cavities, and DC accelerators.
 The resulting computationally efficient propagation analysis can then be used to model, and hence design, UEM columns and UED systems in a straightforward manner.
@@
 In Chapter \ref{chap:model_results}, I apply the extended AG model to draw some initial insight into the dynamics of ultrafast electron pulses traversing a UEM column.
-One of the conclusions that I will make is that oblate, or ``disk-like'', electron pulses will be generally preferred to prolate ``cigar-like.'' 
+One of the conclusions that I will make is that electron pulses which are oblate or ``disk-like'' in shape will be generally preferred to those that are prolate or ``cigar-like.'' 
 Indeed for UEM, with typical laser pulse durations in the femtosecond to picosecond timescales, the electron pulses are likely to be oblate for reasonably sized incident laser spot sizes.
@@
 I will present investigations of several different photoemission processes, some of which are yielding very promising results.
-In fact, our recent work on excited-state thermionic emission~\cite{berger_excited_2012} may even challenge long-standing ideas about which materials and processes may hold the keys to low-emittance (or high-brightness) electron sources.
+Our recent work on excited-state thermionic emission~\cite{berger_excited_2012} may even challenge long-standing ideas about which materials and processes may hold the keys to low-emittance (or high-brightness) electron sources.

### 2013-05-03 main.tex
typo fixes in introduction material

--- 2013-04-14 main.tex
@@
 Finally, I want to thank the funding agencies which have made this research possible: 
-the National Science Foundation (DMR-0619573) and the Department of Energy (DE-FG52-06NA26213 and DE-FG52-09NA29451), as well as a Department of Education GAANN fellowship (DED P200A070409).
+the National Science Foundation (DMR-0619573) and the Department of Energy (DE-FG52-06NA26213 and DE-FG52-09NA29451).
+I have been honored to be personally funded by a Department of Education GAANN fellowship (DED P200A070409), a UIC Dean's Scholar Award fellowship, and a Physics Department James Kouvel Fellowship.
 
@@
 It attempts to combine the nanoscale spatial resolution of transmission electron microscopes with the temporal resolution of modern ultrafast lasers.
-In this thesis I present my contributions to this young field. 
+In this thesis, I present my contributions to this young field. 
 These include a novel model for simulating the dynamics of ultrafast electron pulses in electron microscope systems,

### 2013-05-03 inc/ms_model/ms_model.tex
tweaks to ms_model chapter

@@
 The result is a set of equations that can be solved very quickly using numerical methods on a modern computer.
-When checked against N-body simulations, the authors find that the model is also surprisingly accurate, even for a large set of non-Gaussian pulses~\cite{michalik_analytic_2006,michalik_evolution_2009}.
+When checked against N-body simulations, the authors find that the model is also surprisingly accurate, even for a large set of non-Gaussian pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}.
 A drawback is that the only force included in this model is that of the internal repulsive space-charge field.
@@
 \end{equation}
-Although this equation looks very similar to \ref{eq:evolution_discrete}, notice that it contains approximations, mean-field that has been seen and in practice the self-similarity of the distribution function, that the discrete description did not.
+Although this equation looks very similar to \ref{eq:evolution_discrete}, notice that it contains the mean-field approximation that has already been seen and the distribution function is now implicitly self-similar, an approximation that the discrete description did not employ.
 
@@
 For Ultrafast Electron Microscopy, the electron pulse is generated by a laser pulse, therefore the electron pulse can be thought to be a Gaussian in space and time, mimicking the laser pulse which generated it.
-How correct this assumption is will be explored in \ref{sec:initial_conditions}.
-In this section we consider a Gaussian distribution function, generally of the form
+How correct this assumption is will be explored in Section \ref{sec:initial_conditions}.
+Consider a Gaussian distribution function, generally of the form
 \begin{equation}
@@
 \begin{equation}
-  A = 
+  A^{-1} = 
   \begin{pmatrix}
@@
   K^{force}_{ij} \equiv \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial \Phi(\vec{r};t)}{\partial \vec{r}} d\vec{r} d\vec{p} \\
+  \Phi(\vec{r};t) \equiv \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) \label{eq:potential_integral} \\ 
   n(\vec{r};t) \equiv \int d\vec{p} f(\vec{r},\vec{p};t) 
-    = \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right) \\
-    \label{eq:potential_integral}
-    \Phi(\vec{r};t) \equiv \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) \,\text{.}
+    = \frac{N}{(2\pi)^{3/2} \sigma_{\smallT} \sqrt{\sigma_{z}}} \exp \left( -\frac{x^2+y^2}{2\sigma_{T}} - \frac{z^2}{2\sigma_{z}} \right) \,\text{.}
 \end{gather}
@@
 The $K^{force}$ integrals, related to the internal force of the pulse's Coulombic repulsion, are far more complicated to evaluate.
-For more information about the derivation, see~\cite{michalik_analytic_2006}. %TODO link to an appendix with Bessel etc for Kforce
+For more information about the derivation, see Reference~\cite{michalik_analytic_2006}. 
 The final matrices are
@@
   \begin{pmatrix}
-    \hat{k}^{force}_{\smallT} & 0 & 0 \\
-    0 & \hat{k}^{force}_{\smallT} & 0 \\
-    0 & 0 & \hat{k}^{force}_{z}
+    k^{force}_{\smallT} & 0 & 0 \\
+    0 & k^{force}_{\smallT} & 0 \\
+    0 & 0 & k^{force}_{z}
   \end{pmatrix} \,\text{,}
@@
 \begin{equation}
-  \hat{k}^{force}_{\alpha} = 
+  k^{force}_{\alpha} = 
   \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{2\pi^{1/2}\sigma_{\alpha}^{1/2}} L_{\alpha}(\xi)

### 2013-05-03 inc/extension/external_forces/formal.tex
fixed some operator/matrix symbology

@@
 \end{equation}
-It is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ (or equivalently $ K^{ext} = \hat{0} $) meaning that this static field has no effect on the parameters of the pulse.
+It is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ meaning that this static field has no effect on the parameters of the pulse.
 Of course, this is only true in the pulse frame, the frame of the simulation, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m $ is unchanged.
@@
 \begin{equation} \label{eq:linear_force}
-\vec{F}_{ext} = - M_{ \smallT } r \hat{\operatorname{r}} - M_{z} z \hat{\operatorname{z}} \,\text{.}
+\vec{F}_{ext} = - M_{ \smallT } r \hat{r} - M_{z} z \hat{z} \,\text{.}
 \end{equation}
@@
   \begin{pmatrix}
-    \hat{k}^{lin}_{\smallT} & \hat{0} & \hat{0} \\
-    \hat{0} & \hat{k}^{lin}_{\smallT} & \hat{0} \\
-    \hat{0} & \hat{0} & \hat{k}^{lin}_{z}
+    k^{lin}_{\smallT} & 0 & 0 \\
+    0 & k^{lin}_{\smallT} & 0 \\
+    0 & 0 & k^{lin}_{z}
   \end{pmatrix} \text{,}
@@
 \begin{equation}
-  \hat{k}^{lin}_{\alpha} =
+  k^{lin}_{\alpha} =
   \begin{pmatrix}

### 2013-05-10 inc/extension/external_forces/formal.tex
tweaks to chapter 3

@@
 While the free-space AG model does not contain any external forces, it does include an internal force: the Coulomb force is included to model the repulsion of the pulse's electrons from each other.
-As this force is self-influencing, it is the major source of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics~\cite{michalik_analytic_2006}.
+As this force is self-influencing, it is the major source of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to so-called ``Gaussian optics''~\cite{michalik_analytic_2006}.
 

### 2013-05-10 inc/extension/external_forces/practical.tex
fixed some operator/matrix symbology
tweaks to chapter 3

@@
 
-\section{Practical Extension to Include Magnetic Lenses, RF Cavities and DC Accelerators}
+\section{Extension to Include Magnetic Lenses, RF Cavities, and DC Accelerators}
 
@@
 The order parameter defines the ``sharpness'' of the edge of the region of influence with integer values ranging from 1 (a simple Gaussian) through $\infty$ (a top-hat).
-We connect $z^{\prime}$ to $t$ through $ v_{\smallzero} $; the kinetic equations are simply defined for the position of the peak of the pulse.
+We connect $z^{\prime}$ to $t$ through $ v_{\smallzero} $, the speed of the pulse; the kinematic equations are simply defined for the position of the peak of the pulse.
 
@@
 \begin{equation} \label{eq:RF_Force_z}
-  \boldsymbol{F}_{ z , RF }^{linear} = - \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] z \hat{\operatorname{ \boldsymbol{z} }} \text{.}
+  F_{ z , RF }^{linear} = - \left [ \frac{ e \Omega E_{0} }{ v_{\smallzero} } \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) \right ] z \hat{ z } \text{.}
 \end{equation}
@@
 \begin{equation} \label{eq:RF_Force_r}
-  \boldsymbol{F}_{ r , RF } = e E_{0} \left [ 
+  F_{ r , RF } = e E_{0} \left [ 
     \frac{ v_{\smallzero} \Omega }{c^{2}} \cos \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right ) + \frac{2 n}{d} \left ( \frac{z^{\prime} - z^{\prime}_{RF}}{d / 2} \right )^{2 n - 1} \sin \left ( \frac{ \Omega z^{\prime} }{ v_{\smallzero} } + \phi \right )
-   \right ] r \hat{\operatorname{ \boldsymbol{r} }} \text{.}
+   \right ] r \hat{ r } \text{.}
 \end{equation}

### 2013-05-10 inc/extension/implementation/implementation.tex
tweaks to chapter 3

@@
 For example, electron-optical element objects manage the Gaussian region-of-influence (\ref{eq:reg_of_influence}) and calculate the force that their fields apply to the pulse.
+While the pulse object may be created manually, the toolkit contains the class \verb!Physics::UEMColumn::Photocathode!
+which implements the simple initial conditions described in Section \ref{sec:initial_conditions}.
 These objects are managed by a central ``simulation'' class (called \verb!Physics::UEMColumn!) which handles the time evolution of the system.
@@
 Of particular interest to users of \verb!Physics::UEMColumn!, \verb!Physics::Photoemission!
-also includes a subclass of \verb!Physics::UEMColumn::Photoemission!
+also includes a subclass of \verb!Physics::UEMColumn::Photocathode!
 which uses this binning model to generate the pulse's $z$ direction initial conditions.

### 2013-05-10 inc/extension/initial_conditions/initial_conditions.tex
tweaks to chapter 3

@@
 in the limit where the maximum excess photoemission energy, $ \Delta E = \hbar \omega - \Phi $ is significantly greater than the thermal energy $ k_{B} T $ of the electrons in the photocathode -- a condition that is generally met in laser-driven DC photoelectron guns.
-For completeness we note that for thermionic emission $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ where $k_B$ is the Boltzmann constant and $T_e$ is the electron temperature~\cite{dowell_quantum_2009,jensen_emittance_2010}..
+For completeness we note that for thermionic emission $\Delta p_{\smallT} = \sqrt{m k_B T_e}$ where $k_B$ is the Boltzmann constant and $T_e$ is the electron temperature~\cite{dowell_quantum_2009,jensen_emittance_2010}.
 
@@
 \begin{equation}
-  \gamma_{ z } ( 0 ) \approx \Delta v \Delta z = m a \Delta t_{ L } \sqrt{ \sigma_{ z } ( 0 ) } = \frac{ e E_{ DC } \tau }{ \sqrt{ 2 } }\sqrt{ \sigma_{ z } ( 0 ) } \punct{.}
+  \gamma_{ z } ( 0 ) \approx m \Delta v \Delta z = m a \Delta t_{ L } \sqrt{ \sigma_{ z } ( 0 ) } = \frac{ e E_{ DC } \tau }{ \sqrt{ 2 } }\sqrt{ \sigma_{ z } ( 0 ) } \punct{.}
 \end{equation}
@@
 Such effects and the limits of the validity of the longitudinal terms of \ref{eq:summary}, within the context of the AG model, will be presented in Section \ref{sec:bin_model}.
-However, \ref{eq:summary} also represent a defined set of initial conditions that allows the effect of electron optics (i.e., magnetic lenses and RF cavities) on electron pulse propagation to be more readily understood.
+However, \ref{eq:summary} also represents a defined set of initial conditions that allows the effect of electron optics (i.e., magnetic lenses and RF cavities) on electron pulse propagation to be more readily understood.
 Moreover, they reflect the fact that the majority of DC photoelectron guns are driven by laser pulses with spatial and temporal profiles that are close to Gaussian~\cite{williamson_clocking_1997,sciaini_electronic_2009}.

### 2013-05-14 inc/model_results/introduction.tex
as I have suspected, the only previous model I need to introduce is the AG model.
This commit changes the introductions of several chapters and lowers the headings by one in the new AG model chapter
tweaks in chap 4 introduction

--- 2013-04-05 inc/model_results/introduction.tex
@@
 
-In Chapters \ref{chap:previous_models} and \ref{chap:extension}, I have introduced the extended Analytic Gaussian model.
+In Chapters \ref{chap:ms_model} and \ref{chap:extension}, I have introduced the extended Analytic Gaussian model.
 In this chapter, I use this model to make general observations about pulse propagation for Ultrafast Electron Microscopy.
@@
 This type of pulse is useful to demonstrate general pulse dynamics or the ideal influence of electron-optical elements independent of the pulse's creation conditions.
-Simulations using pulses of this type will provide an initial velocity $v_{\smallzero}$ (or equivalent acceleration voltage $V_{DC}$), an excess photoemission energy $\Delta E$, and HW1/eM dimensions; width $w$ and duration $\tau$ or length $ w_z = v_{\smallzero} \tau$.
+Simulations using pulses of this type will specify an initial velocity $v_{\smallzero}$ (or equivalent acceleration voltage $V_{DC}$), an excess photoemission energy $\Delta E$, and HW1/eM dimensions; width $w_{\smallT}$ and duration $\tau$ or length $ w_z = v_{\smallzero} \tau$.
 The width and length will be used directly to calculate initial spatial variances $\sigma_{\alpha}$.
@@
 
-Other examples in this section will require --- or are markedly different when using --- a more realistic pulse resulting from photoemission and subsequent acceleration.
-Simulations using pulses of this type will provide and HW1/eM width and duration of the laser, the length and voltage of the acceleration region, and an excess photoemission energy.
+Other examples in this section will require a more realistic pulse resulting from photoemission and subsequent acceleration.
+Indeed, some of the examples shown with the idealistic pulses display markedly different behavior in light of these realistic conditions.
+Simulations using pulses of this type will specify the HW1/eM width and duration of the laser, the length and voltage of the acceleration region, and an excess photoemission energy.
 In these cases, all initial parameters will be calculated from \ref{eq:summary}.

### 2013-05-14 inc/model_results/shapes/shapes.tex
tweaks to shapes section

@@
 
-The bunch propagation dynamics for negligible pulse charge density case shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
+The bunch propagation dynamics for the negligible pulse charge density case shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
 In particular, in the initially spherical case ($ \xi(0) = 1 $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands approximately twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp.
 Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both oblate and prolate pulses are initially driven towards the spherical regime.
-In all cases, this intrinsic electron pulse broadening due to non-zero $ \eta_{\alpha} $ generates a momentum chirp across the electron bunch; in the longitudinal dimension the faster electrons photoemitted with larger excess energy leading the slower electrons.
+In all cases, this intrinsic electron pulse broadening due to non-zero $ \eta_{\alpha} $ generates a momentum chirp across the electron bunch; in the longitudinal dimension the faster electrons, those photoemitted with larger excess energy, lead the slower electrons.
 

### 2013-05-22 inc/model_results/spacecharge/spacecharge.tex
tweaks in ch 4 (spacecharge and lens_charge sections)

@@
 What can be seen is that the transverse size increases more sharply with pulse charge at higher levels.
-Since the pulse is already diverging in the transverse direction, this is compounded by the increased pulse charge.
-This divergence then acts to lower the overall charge-density more quickly, thus the longitudinal pulse length is not affected as strongly as was seen in \ref{fig:spacecharge_noacc}.
+Recall that the pulse is already diverging in the transverse direction, the increased pulse charge now serves to compound this action.
+This increased divergence then acts to lower the overall charge-density more quickly, thus the longitudinal pulse length is not affected as strongly as was seen in \ref{fig:spacecharge_noacc}.

### 2013-05-22 inc/model_results/rf_cav/rf_cav.tex
tweaks to model_results rf_cav section

@@
 The mathematical model of the RF cavity (see Section \ref{rf_cav_model}) employs $L_{RF} = d \approx $ 14mm, $ z^{\prime}_{RF} = 0 $, and a region of influence super-Gaussian of order 2 ($n=2$ in \ref{eq:reg_of_influence}) to reflect the sharper ends of the RF cavity.
-In both cases, the idealized initial conditions are used, starting the pulse at $ z^{\prime} = -10$ cm, an excess photoemission energy $\Delta E = 0.5$ eV, velocity related to a $V_{DC} = $ 20kV acceleration potential, HW1/eM $ w = $ 0.1mm and duration $\tau = $ 100fs ($v_{\smallzero} \tau \approx 8.4 \mu$m).
+In both cases, the idealized initial conditions are used, starting the pulse at $ z^{\prime} = -10$ cm, an excess photoemission energy $\Delta E = 0.5$ eV, velocity related to a $V_{DC} = $ 20kV acceleration potential, HW1/eM width $ w = $ 0.1mm and duration $\tau = $ 100fs ($v_{\smallzero} \tau \approx 8.4 \mu$m).
 This electron pulse is therefore initially in the preferred highly oblate regime ($ \sigma_{\smallT}(0) \gg \sigma_{z} (0) $), for which space-charge effects predominantly act in the longitudinal direction, broadening the pulse in time (see \ref{fig:compare_shape}).
@@
 In this case, the pulse is returned to its original temporal duration --- an exact 1-to-1 `temporal imaging' of the photo-generated electron pulse at $z' = -10$cm by the RF cavity to $ z' = $ 10cm.
-On the other hand, the dynamics of the spatial size of the electron pulse are clearly affected by the negative compound transverse lens of the RF cavity (the magnetic contribution of the TM$_{010}$-mode plus the axial entrance and exit aperture electric fields, see Section \ref{rf_cav_model}), increasing at the temporal focus to a factor of $\sim 10\times$ its initial value.
+On the other hand, the dynamics of the spatial size of the electron pulse are clearly affected by the negative compound transverse lens of the RF cavity (the magnetic contribution of the TM$_{010}$-mode plus the axial entrance and exit aperture electric fields, see Section \ref{rf_cav_model}), increasing the pulse's width at the temporal focus to a factor of $\sim 10\times$ its initial value.
 

### 2013-05-22 inc/model_results/rf_cav/rf_cav_plot.tex
tweaks to model_results rf_cav section

@@
   \caption[Dynamics of the pulse length and width under the influence of an RF cavity]{
-    Dynamics of the pulse length (green) and width (red) under the influence of an RF cavity placed at $z' = 0$.
-    The employed idealized pulses start at $z'= -100$ cm with velocity $v_{\smallzero} = \sim 84 \times 10^{6}$ m/s (20kV) and initial HW1/eM width of 100 $\mu$m and length $\sim 8.4 \mu$m ($v_{\smallzero} \cdot $ 100fs).
-    \subref{fig:rf_cav_N1} has $N=1$ electron and demonstrates one-to-one imaging of the initial pulse length at the focus at $ z' = 100 $ cm.
-    \subref{fig:rf_cav_N1e5} has $ N = 1 \times 10^5 $ electrons, whose added broadening allows for the factor of two reduction in pulse length at the focus ($\sim 48$ fs).
+    Dynamics of the pulse length (green) and width (red) under the influence of an RF cavity placed at $z^{\prime} = 0$.
+    The employed idealized pulses start at $z^{\prime} = -100\text{cm}$ with velocity $v_{\smallzero} \approx 84 \times 10^{6} \text{m/s}$ ($20\text{kV}$) and initial HW1/eM width of 100 $\mu$m and length $\sim 8.4 \mu\text{m}$ ($v_{\smallzero} \cdot 100\text{fs}$).
+    \subref{fig:rf_cav_N1} has $N=1$ electron and demonstrates one-to-one imaging of the initial pulse length at the focus at $ z^{\prime} = 100 \text{cm}$.
+    \subref{fig:rf_cav_N1e5} has $ N = 1 \times 10^5 $ electrons, whose added broadening allows for the factor of two reduction in pulse length at the focus ($\sim 48\text{fs}$).
   }

### 2013-05-22 inc/model_results/compound/compound.tex
tweak model_results compound system section

@@
 
-Through the use of multiple electron optical components such as magnetic lenses and RF cavities and accelerators (properly region-of-influence limited by application of \ref{eq:reg_of_influence}), the extended AG model can be employed to design and assess a full UEM column from pulse generation to sample stage.
+Through the use of multiple electron optical components such as magnetic lenses, RF cavities, and accelerators (properly region-of-influence limited by application of \ref{eq:reg_of_influence}), the extended AG model can be employed to design and assess a full UEM column from pulse generation to sample stage.
 The design of such a compound system is, of course, a cornerstone of modern electron microscope development, although space-charge effects are generally not included, since, on average, there is usually no more than one electron in the microscope column at a time.
@@
 \ref{fig:compound} shows an example of a compound element system for the delivery of an ultrashort electron pulse to the specimen area; in this case, optimized for $ N = 10^{ 6 } $ --- the number of electrons per pulse required for single shot diffraction measurements~\cite{armstrong_practical_2007}.
-The displayed dynamics are for an electron pulse generated by a $\tau = $ 100fs $w = $ 0.1mm laser, driving a photocathode, with $\Delta E = $ 0.5eV, then accelerated by 20 kV to a velocity of $8.4 \times 10^{7} \text{m}/\text{s}$.
+The displayed dynamics are for an electron pulse generated by a laser having HW1/eM duration $\tau = $ 100fs and width $w = $ 0.1mm, driving a photocathode, with $\Delta E = $ 0.5eV, then accelerated by 20 kV to a velocity of $8.4 \times 10^{7} \text{m}/\text{s}$.
 In this simple system, an optimum RF cavity ($ \Omega = \pi v_{{ \scriptscriptstyle 0}} / d $) is surrounded by two magnetic lenses (positioned a realistic 5cm to either side of the RF cavity) that compensate for the electron beam divergence due to the non-zero $ \eta_{\smallT} (0) $ and the negative spatial lensing of the RF pulse compression cavity.

### 2013-05-23 inc/design_considerations/considerations.tex
tweak into to considerations chapter

@@
 In doing so we have decided that special care must be given to the creation and propagation of the electron beam.
-In the case of single-shot UEM, in which the image of the sample is acquired using only one electron pulse, several factors lead to two major considerations: first, that a large area photoemission source is required to smaller area or even single point emitters, and second, that the ``normalized transverse emittance,'' a beam quality figure of merit which will be defined later, must be minimized.
+In the case of single-shot UEM, in which the image of the sample is acquired using only one electron pulse, several factors lead to two major considerations: first, that a large area photoemission source is required versus smaller area or even single point emitters, and second, that the ``normalized transverse emittance,'' a beam quality figure of merit which will be defined later, must be minimized.
 

### 2013-05-23 inc/design_considerations/considerations.tex
tweak charge requirement section

@@
 In a conventional electron microscope the electrons are generated by one of several processes, often by heating the cathode or applying a large field to the cathode, or combinations thereof.
-Although these processes produce a very large number of electrons, far more than required by the Rose criterion, they are essentially not highly directional.
+Although these processes produce a very large number of electrons, far more than required by the Rose criterion, they are not highly directional.
 A pinhole is used to eliminate most of the electrons, with the exception of those which will travel down the column axis, thereby reducing the beam current but producing a high-quality beam.

### 2013-05-28 inc/gun/gun_design.tex
tweaks to hardware chapter

@@
 The acceleration region of our prototype UEM system has several features which are unique in comparison to standard electron guns.
-As discussed in Chapter \ref{chap:considerations}, notable requirements for a UEM electron gun include the need to handle a large photoemission area and yet to retain all of the electrons, it may not employ a pin-hole to selectively clean the electron beam.
+As discussed in Chapter \ref{chap:considerations}, notable requirements for a UEM electron gun include the need to handle a large photoemission area and yet to retain all of the electrons; it may not employ a pin-hole to selectively clean the electron beam.
 These necessities lead to several additional criteria.

### 2013-05-28 inc/hardware/column.tex
tweaks to hardware chapter

--- 2013-03-20 inc/hardware/column.tex
@@
 Apart from the laser system, the majority of an Ultrafast Electron Microscope column is either vacuum hardware or contained within the vacuum system.
-The vacuum chamber itself is almost entirely 4.5 in standard ConFlat hardware.
+The vacuum chamber itself is almost entirely standard 4.5in ConFlat (CF) hardware.
 Much has been purchased from off-the-shelf hardware suppliers.
@@
 Kimball's line of Multi-CF hardware has been especially useful in building the prototype column.
-The main chamber is a 4.5 in ``spherical cube,'' a large stainless-steel sphere which has six 4.5 in ConFlat ports oriented as a cube, with another eight 1.33 in ports on the corners of the cube.
-All ports on Multi-CF hardware also contain a set of grooves, designed as mounting points for a system of erector-set-like equipment called ``eV Parts.''
-This system is used to mount the accelerator anode and the the deflector plates which are both to be described later.
+The main chamber is a 4.5in ``spherical cube,'' a large stainless-steel sphere which has six 4.5in ConFlat ports oriented as a cube, with another eight 1.33in ports on the corners of the cube.
+All ports on Multi-CF hardware also contain a set of grooves designed as mounting points for a system of erector-set-like equipment called ``eV Parts.''
+This system is used to mount the accelerator anode and the deflector plates which are both to be described later.

### 2013-05-28 inc/hardware/laser/laser.tex
tweaks to hardware chapter

@@
 
-Although the laser system had been relatively robust, maintaining mode-locking operation was initially problematic.
+Although the laser system had been relatively robust, maintaining mode-locked operation was initially problematic.
 A cover was added above the laser cavity to prevent dust and air currents from interrupting mode locking.
 More importantly, a 1.4mm aperture (approximately four times the HW1/eM beam width at the output coupler) was added to suppress higher-order transverse modes.
-As this is an additional loss in the cavity, the output coupler was changed from 93\% to 95\% reflectivity to maintain the soliton modelocking power.
-These changes have greatly improved the mode locking stability.
+As this is an additional loss in the cavity, the output coupler was changed from 93\% to 95\% reflectivity to maintain the soliton mode-locking power.
+These changes have greatly improved the mode-locking stability.
 

### 2013-05-28 inc/hardware/magnetic_lenses/maglens.tex
tweaks to hardware chapter

@@
 Further --- as suggested by the use of the auxiliary field $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeability.
-These materials, called ``magnetic shielding'' since they capture the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.
+These materials, called ``magnetic shielding'' since they capture the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap.''
 The field lines exit the shielding in the gap and it is this field that acts as the lens.
@@
 
-The high magnetic permeability shielding is made of eleven pieces.
+The high magnetic permeability shielding is made of ten pieces.
 There are two end pieces which are revolution solids of an ``L'' shaped profile.
@@
     Side-view of the two custom large-bore magnetic lenses designed and built at UIC, seen installed in the prototype UEM.
-    The coils are visible in the gaps between the high permeability shielding ``bridges''.
+    The coils are visible in the gaps between the high permeability shielding ``bridges.''
     The bridges are attached to the tops of the end pieces by bolts to complete the shielding loop.

### 2013-05-31 inc/photocathode/setup/setup.tex
tweak photocathode setup

--- 2013-04-11 inc/photocathode/setup/setup.tex
@@
 I will then outline some future research directions that the group at UIC will continue to pursue.
-Since this is the most active current field in our group's reseach, some very new results are presented near the end of the chapter and future plans are evolving.
+Since this is the most active current field in our group's research, indeed some very new results are presented near the end of the chapter, future plans are continuously evolving.
 

### 2013-05-31 inc/photocathode/single_photon/single_photon.tex
tweak the single_photon section

@@
 
-An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.3-4.5 eV~\cite{yen_thermally_1980}, is used as a photocathode.
-When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E \approx $ 0.35 eV.
+An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.3-4.5 eV~\cite{yen_thermally_1980}, is used as a photocathode; when paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E \approx $ 0.35 eV.
 The linear dependence of the electron yield (circles) on the laser pulse energy is indicative of a single photon process, as expected since each absorbed photon may eject one electron.

### 2013-07-21 vita.tex
started vita section

\vita

\begin{center}
\begin{singlespace}

\vspace{0.4in}

\begin{tabular}{@{\hspace{0.2in}}@{\hspace{0.2in}}p{4.5in}}

{NAME:}
& Joel Adam Berger \\
\\
{EDUCATION:}
& B.S., Physics, University of Illinois at Chicago,\\
& Chicago, Illinois, 2005 \\
\\
{\Large TEACHING:}
& Something here\\

\end{tabular}

\end{singlespace}
\end{center}

### 2013-07-21 vita.tex
fix broken vita markup
added some more data to vita

@@
 
-\begin{tabular}{@{\hspace{0.2in}}@{\hspace{0.2in}}p{4.5in}}
-
-{NAME:}
-& Joel Adam Berger \\
+%\begin{tabular}{@{\hspace{0.2in}}@{\hspace{0.2in}}p{4.5in}}
+\begin{tabular}{p{1.5in}p{6in}}
+NAME: & Joel Adam Berger \\
 \\
-{EDUCATION:}
-& B.S., Physics, University of Illinois at Chicago,\\
+EDUCATION: & B.S., Physics, University of Illinois at Chicago,\\
 & Chicago, Illinois, 2005 \\
 \\
-{\Large TEACHING:}
-& Something here\\
-
+TEACHING: & Something here\\
+\\
+PROFESSIONAL MEMBERSHIPS: & Microscopy Society of America\\
+& Midwest Microscopy and Microanalysis Society\\
+\\
+SELECTED \nobreak{PUBLICATIONS:} &\\
 \end{tabular}

### 2013-07-21 vita.tex
more vita (hand crafted rather than bibtex :/)

@@
 %\begin{tabular}{@{\hspace{0.2in}}@{\hspace{0.2in}}p{4.5in}}
-\begin{tabular}{p{1.5in}p{6in}}
+\begin{tabular}{p{1.5in}p{4in}}
 NAME: & Joel Adam Berger \\
@@
 \\
-TEACHING: & Something here\\
+TEACHING: & \LaTeX{} for Technical Publishing (PHYS491 - Spring 2011)\\
+\\
+& Problem-Solving Workshop for General Physics II (Electricity and Magnetism), General Physics II (Electricity and Magnetism) Lab, General Physics I (Mechanics)\\
 \\
@@
 \\
-SELECTED \nobreak{PUBLICATIONS:} &\\
+SELECTED \nobreak{PUBLICATIONS:}
+& Berger, J. A., Greco, M. J., and Schroeder, W. A.: High-power, femtosecond, thermal-lens-shaped Yb:KGW oscillator. \uline{Opt. Express}, 16(12), 2008.\\
+\\
+& Berger, J.
+A., Hogan, J.
+T., Greco, M.
+J., Schroeder, W.
+A., Nicholls, A.
+W., and Browning, N.
+D.: DC photoelectron gun parameters for ultrafast electron microscopy.
+\uline{Microsc.
+Microanal.}, 15(04), 2009.\\
+\\
+& Berger, J. and Schroeder, W.: Semianalytic model of electron pulse propagation: Magnetic lenses and rf pulse compression cavities. \uline{Journal of Applied Physics}, 108(12):124905, 2010.\\
+\\
+& Berger, J.
+A., Rickman, B.
+L., Li, T., Nicholls, A.
+W., and Schroeder, W.
+A.: Excited-state thermionic emission in III-antimonides: Low emittance ultrafast photocathodes.
+\uline{Applied Physics Letters}, 101(19):194103, 2012.\\
 \end{tabular}

### 2013-07-23 inc/copyright.tex
include copyright notice re optics express

\noindent \ref{fig:crystal}\subref{fig:crystal_coating}, \ref{fig:laser-thermal}, \ref{fig:laser-cavity}\subref{fig:laser_stability}, \ref{fig:laser-profile} and \ref{fig:laser-performance} were previously published in Optics Express. The full article is available at \url{http://www.opticsinfobase.org/oe/abstract.cfm?uri=oe-16-12-8629}. Their use of these figures is granted to the author under the permissive non-commercial terms of the copyright transfer agreement (\url{http://www.opticsinfobase.org/oe/submit/review/copyright_permissions.cfm}).
